### 2019-11-22 FEM_CPP_settings/assumptions-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}

\title{Assumptions within the FEM}

\begin{document}
\maketitle

\begin{Rcode}{!echo,hide}
library(ggplot2)
\end{Rcode}

\section{Economic Factors}

Inflation halts in 2010, so all growth rates afterwards are real
growth.  The following are represented:
\begin{itemize}
\item Interest Rate on Wealth
\item Earning Disregard
\item National Wage Index (SA 2009 TR Intermediate Cost Assumptions,
  Table VI.F6)
\item Social Security Wage Cap
\item Medicare Cost Growth (CMS 2009 Trustees Report, p179)
\item Gross Domestic Product
\begin{description}
\item[1992-2009] Historic
\item[2010-2019] CBO analysis of President's Budget
\item[2020-2100] 2008 OASDI Trustees Report 2.1\%
\item[Labor Force] 2008 OASDI Trustees Report
\end{description}
\item Total Medicare Cost Growth (Computed)
\end{itemize}

\subsection{Interest Rate on Wealth}
This rate is used to increase household wealth year-over-year. It is a
real interest rate over and above CPI.

\begin{Rcode}{!echo,fig,savefig,label=interest}
interest <- read.table('timeseries/interest_rate.txt', skip=1)
names(interest) <- c('year','cummulative.interest')
ggplot(interest, aes(x=year, y=cummulative.interest)) + geom_line()
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{interest}
\caption{Cummulative Real Interest Rate}
\label{fig:interest}
\end{figure}

\subsection{Earning Disregard}
Below full retirement age, \$1 is deducted from SS benefits for every
\$2 earned above the limit. Above full retirement age, \$1 is deducted
for every \$3 earned above the limit.

\begin{Rcode}{!echo,fig,savefig,label=earndisreg}
disreg1 <- read.table('timeseries/eadisreg1.txt', skip=1)
disreg2 <- read.table('timeseries/eadisreg2.txt', skip=1)
names(disreg1) <- c('year','Earning.Disregard')
names(disreg2) <- names(disreg1)
disreg1$group <- 'before retirement'
disreg2$group <- 'after retirement'
disreg <- rbind(disreg1, disreg2)
disreg$Earning.Disregard <- disreg$Earning.Disregard/1e3
ggplot(disreg, aes(x=year, y=Earning.Disregard, colour=group)) +geom_line()
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{earndisreg}
\caption{Earning Disregard (Thousands) Projected 1.1\% Real Growth}
\label{fig:earndisreg}
\end{figure}

\subsection{National Wage Index}
The National Wage Index is used to calculate or update a number of quantities year
over year:
\begin{itemize}
\item AIME
\item Earnings
\item Principle Insurance Amount
\item Widow's Benefit
\end{itemize}

\begin{Rcode}{!echo,fig,savefig,label=nwi}
nwi <- read.table('timeseries/nwi.txt', skip=1)
nwi$V2 <- nwi$V2/1e3
names(nwi) <- c('year','NWI')
ggplot(nwi, aes(x=year, y=NWI)) + geom_line()
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{nwi}
\caption{National Wage Index in Thousands}
\label{fig:nwi}
\end{figure}

\subsection{Social Security Wage Cap}
This is the maximum amount of income subject to the Social Security
payroll tax. It is only used in computing taxes.

\begin{Rcode}{!echo,fig,savefig,label=sscap}
sscap <- read.table('timeseries/sscap.txt', skip=1)
names(sscap) <- c('year', 'SS.Cap')
sscap$SS.Cap <- sscap$SS.Cap / 1e3
ggplot(sscap, aes(x=year, y=SS.Cap)) + geom_line()
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{sscap}
\caption{Social Security Wage Cap (Thousands)}
\label{fig:sscap}
\end{figure}

\subsection{Total Medicare Cost Growth}
This is the projected growth in overall Medicare costs and is used to
inflate medicaid and medicare costs. This is added on top of
per-laborer GDP growth and then divided by the growth in the enrollee
pool (simulated) to figure out the per-enrollee cost growth.

\begin{Rcode}{!echo,fig,savefig,label=cummulative}
c.med <- read.table('timeseries/medgrowth.txt',skip=1)
c.gdp <- read.table('timeseries/gdp.txt', skip=4)
c.labor <- read.table('timeseries/labor.txt', skip=2)
names(c.med) <- c('year','Growth')
names(c.gdp) <- c('year','Growth')
names(c.labor) <- c('year','Growth')

c.med[,'series'] <- 'Medicare Cost above GDP'
c.gdp[,'series'] <- 'Real GDP'
c.labor[,'series'] <- 'Labor Force'

row.names(c.med) <- c.med[,'year']
row.names(c.gdp) <- c.gdp[,'year']
row.names(c.labor) <- c.labor[,'year']

c.percap <- c.gdp
c.percap[,'Growth'] <- c.gdp[,'Growth']/c.labor[,'Growth'] * 100
c.percap[,'series'] <- 'GDP Per Labor Force'

c.totmed <- c.med
c.totmed[,'series'] <- 'Total Medicare Cost'
c.totmed[,'Growth'] <- c.percap[,'Growth'] * c.med[,'Growth'] / 100

a.med <- c.med[2:nrow(c.med),]
a.gdp <- c.gdp[2:nrow(c.gdp),]
a.labor <- c.labor[2:nrow(c.labor),]
a.percap <- c.percap[2:nrow(c.percap),]
a.totmed <- c.totmed[2:nrow(c.totmed),]

c.med[,'group'] <- 'Cummulative'
c.gdp[,'group'] <- 'Cummulative'
c.labor[,'group'] <- 'Cummulative'
c.percap[,'group'] <- 'Cummulative'
c.totmed[,'group'] <- 'Cummulative'
a.med[,'group'] <- 'Annualized'
a.gdp[,'group'] <- 'Annualized'
a.labor[,'group'] <- 'Annualized'
a.percap[,'group'] <- 'Annualized'
a.totmed[,'group'] <- 'Annualized'

med.num <- as.character(c.med[2:nrow(c.med),'year'])
med.den <- as.character(c.med[1:(nrow(c.med)-1),'year'])
gdp.num <- as.character(c.gdp[2:nrow(c.gdp),'year'])
gdp.den <- as.character(c.gdp[1:(nrow(c.gdp)-1),'year'])
labor.num <- as.character(c.labor[2:nrow(c.labor),'year'])
labor.den <- as.character(c.labor[1:(nrow(c.labor)-1),'year'])
percap.num <- as.character(c.percap[2:nrow(c.percap),'year'])
percap.den <- as.character(c.percap[1:(nrow(c.percap)-1),'year'])

a.med[,'Growth'] <- c.med[med.num,'Growth'] / c.med[med.den,'Growth']
a.gdp[,'Growth'] <- c.gdp[gdp.num,'Growth'] / c.gdp[gdp.den,'Growth']
a.labor[,'Growth'] <- c.labor[labor.num,'Growth'] / c.labor[labor.den,'Growth']
a.percap[,'Growth'] <- c.percap[percap.num,'Growth'] /c.percap[percap.den,'Growth']
a.totmed[,'Growth'] <- c.totmed[med.num,'Growth'] /c.totmed[med.den,'Growth']

f <- rbind(c.gdp, c.med, c.labor, c.percap, c.totmed)
ggplot(f) + geom_line(aes(x=year, y=Growth, color=series)) + opts(legend.position = "bottom")
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{cummulative}
\caption{Cummulative Medicare Cost Growth Breakdown}
\label{fig:cummulative}
\end{figure}

\begin{Rcode}{!echo,fig,savefig,label=annualized}
f <- rbind(a.gdp, a.med, a.labor, a.percap, a.totmed)
ggplot(f) + geom_line(aes(x=year, y=Growth, color=series)) + opts(legend.position = "bottom")
\end{Rcode}

\begin{figure}[h]
\centering
\recallfig{annualized}
\caption{Annualized Medicare Cost Growth Breakdown}
\label{fig:annualized}
\end{figure}

\end{document}

### 2019-11-22 FEM_Stata/Estimation/estimate_summary-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{mathpazo,color}

\StataweaveOpts{echo=false, outfmt="frame=single, fontfamily=courier, fontsize=\small"}

\begin{document}

\begin{Statacode}{hide}
set more off
include ../../fem_env.do

*set up work directory
global estdir ../Estimates

*load estimates 
**Disease
estimates use $estdir/hearte.ster
estimates store Heart_disease
estimates use $estdir/stroke.ster
estimates store Stroke
estimates use $estdir/hibpe.ster
estimates store HBP
estimates use $estdir/diabe.ster
estimates store Diabetes
estimates use $estdir/lunge.ster
estimates store Lung
estimates use $estdir/memrye.ster
estimates store Memory_disorder

**Functional Status

estimates use $estdir/died.ster
estimates store Motality
estimates use $estdir/nhmliv.ster
estimates store Nursing_home
estimates use $estdir/iadlstat.ster
estimates store IADL
estimates use $estdir/adlstat.ster
estimates store ADL

**Predicted Medical Costs

estimates use $estdir/totmd_mcbs.ster
estimates store Total_MCBS
estimates use $estdir/totmd_meps.ster
estimates store Total_MEPS
estimates use $estdir/oopmd_meps.ster
estimates store OOP_MEPS
estimates use $estdir/mcare_pta.ster
estimates store Medicare_A
estimates use $estdir/mcare_ptb.ster
estimates store Medicare_B

**Program Participation

estimates use $estdir/ssclaim.ster 
estimates store SS_claim
estimates use $estdir/ssiclaim.ster
estimates store SSI_claim
estimates use $estdir/diclaim.ster 
estimates store DI_claim
estimates use $estdir/mcareb_takeup_newenroll.ster
estimates store MedicareB_new
estimates use $estdir/mcareb_takeup_curenroll.ster
estimates store MedicareB_current
estimates use $estdir/mcare_ptd.ster
estimates store Medicare_D

**Labor Outcomes

estimates use $estdir/work.ster
estimates store Working
estimates use $estdir/dbclaim.ster
estimates store DB_pension_claiming

\end{Statacode}

\section*{\centering Type of Estimators}

\begin{Statacode}

estimates dir

\end{Statacode}

\pagebreak
\section*{\centering Summary Characteristics}

\begin{Statacode}

estimates stat _all

\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Diseases}
\begin{Statacode}
estimates table Heart_disease Stroke HBP, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Diseases(continued)}
\begin{Statacode}
estimates table Diabetes Lung Memory_disorder, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Functional Status}
\begin{Statacode}
estimates table Motality Nursing_home IADL ADL, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Predicted Medical Costs}
\begin{Statacode} 
estimates table Total_MCBS Total_MEPS , star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Predicted Medical Costs(Continued)}
\begin{Statacode} 
estimates table OOP_MEPS Medicare_A Medicare_B, star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Program Participation}
\begin{Statacode}
estimates table SS_claim SSI_claim DI_claim, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Program Participation(Continued)}
\begin{Statacode}
estimates table MedicareB_new MedicareB_current Medicare_D, star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl1 adl2 adl3p widowed work) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}
estimates table Working DB_pension_claiming, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}
estimates table Working DB_pension_claiming, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\begin{Statacode}
qui set mem 300m
qui use $outdata/hrs17_transition.dta

*logdeltaage
qui gen logdeltaage = log(age_iwe - lage_iwe)

*llogbmi_l30 llogbmi_30p flogbmi_l30 flogbmi_30p
qui {
  local log_30 = log(30)
  mkspline llogbmi_l30 `log_30' llogbmi_30p = llogbmi
  mkspline flogbmi_l30 `log_30' flogbmi_30p = flogbmi
}

*flogq flogaime

qui gen flogq=0
qui gen flogaime=0

local age_var age_iwe

gen lage62e = floor(l`age_var') == 60 if l`age_var' < .
gen lage63e = floor(l`age_var') == 61 if l`age_var' < .

mkspline la6 58 la7 73 la7p = l`age_var'

*** GENERATE WEAVE DUMMIES
gen w3 = wave == 3
gen w4 = wave == 4
gen w5 = wave == 5
gen w6 = wave == 6
gen w7 = wave == 7

\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Diseases}
\begin{Statacode}
**marginal effects
*desease
qui estimates restore Heart_disease
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fstroke fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix heart=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\.\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Stroke
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix stroke=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\.\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore HBP
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix hbp=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\.\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Diabetes
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix diab=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\.\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Lung
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix lung=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\.\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Memory_disorder
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix mem=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\.\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

matrix disease=(heart,stroke,hbp,diab,lung,mem)
matrix colnames disease = "heart disease""stroke""hypertension""diabetes""lung disease""memory disorder"
matrix rownames disease = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "smoke" "widowed" "single" "bmi<30" "bim>30" "work"
matlist disease, lines(oneline) twidth(15)  
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Functional Health}
\begin{Statacode}
qui estimates restore Motality
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix mortality=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore Nursing_home
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix nursingh=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore IADL
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix iadl=(temp,temp1)
matrix iadl=iadl'
*matrix iadl=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore ADL
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix adl=(temp,temp1)
matrix adl=adl'
*matrix iadl=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

matrix functional=(mortality, nursingh, iadl, adl)
matrix colnames functional = "mortality""nursing home""iadl""adl"
matrix rownames functional = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "smoke" "widowed" "single" "work"
matlist functional, lines(oneline) twidth(15)  
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Predicted Medical Costs}
\begin{Statacode} 
estimates table Total_MCBS Total_MEPS , star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Program Participation}
\begin{Statacode}

qui estimates restore SS_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix ss=(temp,temp1)
matrix ss=ss'
matrix drop temp temp1

qui estimates restore SSI_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix ssi=(temp,temp1)
matrix ssi=ssi'

qui estimates restore DI_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix diclaim=(temp,temp1)
matrix diclaim=diclaim'
matrix drop temp temp1

matrix programp=(ss, ssi, diclaim)
matrix colnames programp = "ss claim""ssi claim""di claim"
matrix rownames programp = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "widowed" "single" "work"
matlist programp, lines(oneline) twidth(15)

\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}

qui estimates restore Working
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix working=(temp,temp1)
matrix working=working'

matrix labor=(working)
matrix colnames labor = "working"
matrix rownames labor = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "smoke" "widowed" "single" "work"
matlist labor, lines(oneline) twidth(15)

\end{Statacode}
\end{document}

### 2019-11-22 analysis/depratios-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{url}

\title{Dependency Ratios}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}{!echo,hide}
library(ggplot2)
library(foreign)
\end{Rcode}

Population projections are from \url{http://www.census.gov/population/www/projections/files/nation/download/NP2008_D1.csv}

\begin{Rcode}
pop <- read.csv('NP2008_D1.csv')
pop <- subset(pop, HISP==0 & RACE==0 & SEX==0)
vec1864 <- paste('POP_', 18:64, sep='')
vec1850 <- paste('POP_', 18:50, sep='')
vec65p <- paste('POP_', 65:100, sep='')
pop1864 <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec1864)
pop1850 <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec1850)
pop65p <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec65p)
pops.df <- data.frame(pop1864=unlist(as.list(pop1864)), pop1850=unlist(as.list(pop1850)), pop65p=unlist(as.list(pop65p)))
row.names(pops.df) <- names(as.list(pop1864))
print(pops.df)
\end{Rcode}

\begin{Rcode}
summ <- read.dta('../output/depratios/depratios_summary.dta')
row.names(summ) <- summ[,'year']
summ[,'pop1864'] <- pops.df[row.names(summ),1]/1e6
summ[,'pop1850'] <- pops.df[row.names(summ),2]/1e6
summ[,'pop65p'] <- pops.df[row.names(summ),3]/1e6
summ <- subset(summ, year >= 2010 & !is.na(pop1864))
\end{Rcode}

\begin{Rcode}{savefig,label=standard}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM Simulation')) + 
geom_line(aes(y=pop65p/pop1864, colour='Census')) +
opts(legend.position='bottom') + labs(y='OADR', colour='65+ Population Projection')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{standard}
\caption{OADR: Age 65+ over Ages 18-64}
\end{figure}

\begin{Rcode}{savefig,label=qalystats}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=qaly_75Q, colour='75th Percentile')) + 
geom_line(aes(y=qaly_50Q, colour='50th Percentile')) + 
geom_line(aes(y=qaly, colour='Mean')) +
geom_line(aes(y=qaly_25Q, colour='25th Percentile')) + 
geom_line(aes(y=qaly_10Q, colour='10th Percentile')) + 
opts(legend.position='bottom') + labs(colour='QALY Statistic', y='QALY')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{qalystats}
\caption{QALY Over Time}
\end{figure}

\begin{Rcode}{savefig,label=qaly65}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_65/pop1864, colour='QALY <= 0.80')) + 
geom_line(aes(y=ttl_qaly_under7_65/pop1864, colour='QALY <= 0.70')) + 
geom_line(aes(y=ttl_qaly_under6_65/pop1864, colour='QALY <= 0.60')) + 
geom_line(aes(y=ttl_qaly_under5_65/pop1864, colour='QALY <= 0.50')) +
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qaly51}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_51/pop1850, colour='QALY <= 0.80')) +
geom_line(aes(y=ttl_qaly_under7_51/pop1850, colour='QALY <= 0.70')) + 
geom_line(aes(y=ttl_qaly_under6_51/pop1850, colour='QALY <= 0.60')) + 
geom_line(aes(y=ttl_qaly_under5_51/pop1850, colour='QALY <= 0.50')) +
 opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qaly51nw}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_51_nw/pop1850, colour='QALY <= 0.80 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under7_51_nw/pop1850, colour='QALY <= 0.70 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under6_51_nw/pop1850, colour='QALY <= 0.60 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under5_51_nw/pop1850, colour='QALY <= 0.50 and Not Working')) +
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=disabled}
ggplot(summ, aes(x=year)) + geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=pop_disability/pop1850, colour='Age 51+ and Any Disability')) + 
geom_line(aes(y=ttl_adl3p/pop1850, colour='Age 51+ and ADL 3 or More')) + opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qalyweighting}
ggplot(summ, aes(x=year)) + geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=(pop_medicare_age - ttl_qaly65)/pop1864, colour='QALY-weighted FEM OADR')) + 
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{qaly65}
\caption{QALY Thresholds on Age 65+}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qaly51}
\caption{QALY Thresholds on Age 51+}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qaly51nw}
\caption{QALY Thresholds on Age 51+ and Not Working}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{disabled}
\caption{Different Definitions of Disability}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qalyweighting}
\caption{QALY-weighted Population Ratios}
\end{figure}

\begin{Rcode}{savefig,label=prevelance2010a}
s <- subset(summ, year==2010)
xes <- seq(0.85, 0.35, -0.1)
buckets <- seq(90,40,-10)

cs <- as.numeric(s[,paste('cancre_q',buckets, sep='')])
ds <- as.numeric(s[,paste('diabe_q', buckets, sep='')])
hes <- as.numeric(s[,paste('hearte_q', buckets, sep='')])
his <- as.numeric(s[,paste('hibpe_q', buckets, sep='')])
ls <- as.numeric(s[,paste('lunge_q', buckets, sep='')])
ss <- as.numeric(s[,paste('stroke_q', buckets, sep='')])
is <- as.numeric(s[,paste('iadl1_q', buckets, sep='')])
a1s <- as.numeric(s[,paste('adl1_q', buckets, sep='')])
a3s <- as.numeric(s[,paste('adl3p_q', buckets, sep='')])
ns <- as.numeric(s[,paste('nhmliv_q', buckets, sep='')])
bs <- as.numeric(s[,paste('bmiGT30_q', buckets, sep='')])

ggplot(s) +
geom_line(aes(x=xes, y=cs, colour='Cancer'), na.rm=T) +
geom_line(aes(x=xes, y=ds, colour='Diabetes'), na.rm=T) +
geom_line(aes(x=xes, y=hes, colour='Heart Disease'), na.rm=T) +
geom_line(aes(x=xes, y=his, colour='High Blood Pressure'), na.rm=T) +
geom_line(aes(x=xes, y=ls, colour='Lung Disease'), na.rm=T) +
geom_line(aes(x=xes, y=ss, colour='Stroke'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{Rcode}{savefig,label=prevelance2010b}
ggplot(s) +
geom_line(aes(x=xes, y=is, colour='IADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a1s, colour='ADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a3s, colour='ADL 3+'), na.rm=T) +
geom_line(aes(x=xes, y=ns, colour='Nursing Home'), na.rm=T) +
geom_line(aes(x=xes, y=bs, colour='BMI > 30'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelance2010a}
\caption{Prevelance of Disease by QALY in 2010}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{prevelance2010b}
\caption{Prevelance of Health Conditions by QALY in 2010}
\end{figure}

\begin{Rcode}{savefig,label=prevelance2050a}
s <- subset(summ, year==2050)

cs <- as.numeric(s[,paste('cancre_q',buckets, sep='')])
ds <- as.numeric(s[,paste('diabe_q', buckets, sep='')])
hes <- as.numeric(s[,paste('hearte_q', buckets, sep='')])
his <- as.numeric(s[,paste('hibpe_q', buckets, sep='')])
ls <- as.numeric(s[,paste('lunge_q', buckets, sep='')])
ss <- as.numeric(s[,paste('stroke_q', buckets, sep='')])
is <- as.numeric(s[,paste('iadl1_q', buckets, sep='')])
a1s <- as.numeric(s[,paste('adl1_q', buckets, sep='')])
a3s <- as.numeric(s[,paste('adl3p_q', buckets, sep='')])
ns <- as.numeric(s[,paste('nhmliv_q', buckets, sep='')])
bs <- as.numeric(s[,paste('bmiGT30_q', buckets, sep='')])

ggplot(s) +
geom_line(aes(x=xes, y=cs, colour='Cancer'), na.rm=T) +
geom_line(aes(x=xes, y=ds, colour='Diabetes'), na.rm=T) +
geom_line(aes(x=xes, y=hes, colour='Heart Disease'), na.rm=T) +
geom_line(aes(x=xes, y=his, colour='High Blood Pressure'), na.rm=T) +
geom_line(aes(x=xes, y=ls, colour='Lung Disease'), na.rm=T) +
geom_line(aes(x=xes, y=ss, colour='Stroke'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{Rcode}{savefig,label=prevelance2050b}
ggplot(s) +
geom_line(aes(x=xes, y=is, colour='IADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a1s, colour='ADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a3s, colour='ADL 3+'), na.rm=T) +
geom_line(aes(x=xes, y=ns, colour='Nursing Home'), na.rm=T) +
geom_line(aes(x=xes, y=bs, colour='BMI > 30'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelance2050a}
\caption{Prevelance of Disease by QALY in 2050}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{prevelance2050b}
\caption{Prevelance of Health Conditions by QALY in 2050}
\end{figure}

\begin{Rcode}{savefig,label=prevelanceevolutionA}
conds <- c('cancre','diabe','hearte','hibpe','lunge','stroke')
condlist <- c(paste(conds, '_qLT60', sep=''), paste(conds, '_qLT70', sep=''), paste(conds, '_qLT80', sep=''), paste(conds, '_qLT90', sep=''))

dat.l <- lapply(condlist, FUN=function(c, s) data.frame(year=as.numeric(s[,'year']),
                                                        val=s[,c],
							qaly.thresh=paste('QALY <= 0.', substr(c,nchar(c)-1, nchar(c)), sep=''),
							condition=strsplit(c,'_')[[1]][1]), s=summ)
dat <- do.call(rbind, dat.l)
ggplot(dat, aes(x=year, y=val)) + 
geom_line(aes(colour=condition)) +
scale_y_continuous(limits=c(0,1), breaks=seq(0,1,0.2)) +
facet_wrap( ~ qaly.thresh) +
labs(colour='Health Condition', x='Year', y='Prevelance') +
opts(legend.position='bottom')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelanceevolutionA}
\caption{Evolution of Prevelance by QALY Upper Threshold}
\end{figure}

\begin{Rcode}{savefig,label=prevelanceevolutionB}
conds <- c('iadl1','adl1','adl3p','nhmliv', 'bmiGT30')
condlist <- c(paste(conds, '_qLT60', sep=''), paste(conds, '_qLT70', sep=''), paste(conds, '_qLT80', sep=''), paste(conds, '_qLT90', sep=''))

dat.l <- lapply(condlist, FUN=function(c, s) data.frame(year=as.numeric(s[,'year']),
                                                        val=s[,c],
							qaly.thresh=paste('QALY <= 0.', substr(c,nchar(c)-1, nchar(c)), sep=''),
							condition=strsplit(c,'_')[[1]][1]), s=summ)
dat <- do.call(rbind, dat.l)
ggplot(dat, aes(x=year, y=val)) + 
geom_line(aes(colour=condition)) +
scale_y_continuous(limits=c(0,1), breaks=seq(0,1,0.2)) +
facet_wrap( ~ qaly.thresh) +
labs(colour='Health Condition', x='Year', y='Prevelance') +
opts(legend.position='bottom')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelanceevolutionB}
\caption{Evolution of Prevelance by QALY Upper Threshold}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/depratios/depratios_summary.dta
keep year ttl_qaly_u* ttl_qaly_a*
drop *_nw
outsheet using depratios_survivor.csv, comma replace
\end{Statacode}

\end{document}

### 2019-11-22 analysis/income-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{amsfonts}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{verbatim}
\usepackage{color}

\title{Debugging FEM Economic Data}

\begin{document}

\begin{Rcode}{!echo,hide}
library(foreign)
library(ggplot2)
stat_sum_single <- function(fun, geom='point', ...) { 
  stat_summary(fun.y=fun, colour='red', geom=geom, size = 3, ...) 
} 
agegrp <- function(ages, levels=list(b60=0:60,
f60to65=60:65, f65to70=65:70, f70to75=70:75,
f75to80=75:80, f80to85=80:85, o85=85:200)) {
 outvec <- 'none'
 for(i in 1:length(levels)) {
    outvec[min(levels[[i]]) <= ages & ages < max(levels[[i]])] <- names(levels)[i]
  }
  return(outvec)
}
\end{Rcode}

\maketitle
\listoftables
\listoffigures

\section{Variables of Interest}
Table~\ref{tab:variables} shows the variables of interest along with quick descriptions. In addition to the variables presented, there are initial value variables (prefaced with an 'f'), lag value variables ('l'), and log-transformed variables ('log').
\begin{table}[ht]
\centering
\caption{Economic Variables in the FEM}
\begin{tabular}{|c|c|l|}\hline
\textbf{Name} & \textbf{Range} & \textbf{Description} \\\hline
anydb & $\{0, 1\}$ & any defined benefit from current job\\\hline
anydc & $\{0,1\}$ & any defined contribution from current job\\\hline
db\_tenure & $[-2, 54.5]$ & defined benefit tenure\\\hline
dbage & $\mathbb{R}^+$ & age started claiming pension\\\hline
dbclaim & $\{0,1\}$ & currently claiming defined benefit pension\\\hline
dbpen & $\mathbb{R}^+$ & defined benefit pension amount\\\hline
dcwlth & $\mathbb{R}^+$ & defined contribution wealth in thousands\\\hline
dcwlthx & $\mathbb{R}^+$ & defined contribution wealth in thousands\\\hline
diben & $\mathbb{R}^+$ & amount of disability insurance benefits\\\hline
diclaim & $\{0, 1\}$& currently claiming SSDI\\\hline
gross & $\mathbb{R}^+$ & gross income\\\hline
hatota & $\mathbb{R}$ & household wealth in thousands\\\hline
hatotax & $[0, 2000]$ & household wealth in thousands, capped\\\hline
agi & $\mathbb{R}^+$ & estimated adjusted gross income\\\hline
hhearn & $\mathbb{R}^+$ & total household earnings\\\hline
hhttlinc & $\mathbb{R}^+$ & total household income, currently set equal to hhearn\\\hline
hicap & $\mathbb{R}$ & estimated household capital income\\\hline
hicap\_nonzero & $\{0, 1\}$ & hicap nonzero\\\hline
iearn & $\mathbb{R}^+$ & individual earnings in thouseands\\\hline
iearnx & $[0, 2000]$ & individual earnings in thousands, capped at 2000\\\hline
isret & $\mathbb{R}^+$ & R SoCSec Retirement Income\\\hline
net & $\mathbb{R}$ & net income\\\hline
ry\_earn & $[0, 2000000]$ & iearnx * 1000\\\hline
rssclyr & $\mathbb{R}^+$ & year person first claimed Social Security\\\hline
ssben & $\mathbb{R}^+$ & SS benefits\\\hline
ssclaim & $\{0, 1\}$ & currently claiming OASI\\\hline
ssiben & $\mathbb{R}^+$ & SSI benefits\\\hline
ssiclaim & $\mathbb{R}^+$ & currently claiming SSI\\\hline
wlth\_nonzero & $\{0, 1\}$ & hatota nonzero\\\hline
work & $\{0, 1\}$ & person currently working\\\hline
\end{tabular}
\label{tab:variables}
\end{table}

\section{Model Algorithm}

\subsection{New Cohorts}
\label{sec:new}
\begin{itemize}
\item all incoming cohorts have the same basic distribution of earnings, household wealth, and defined contribution plan wealth
\item earnings data are inflated according to the NWI factor, which is the real growth in wages from the reference year to the year when the cohort enters the simulation
\item household wealth data are inflated by CPI to the year when the cohort enters the simulation
\item DC wealth is not modified
\item household capital income is estimated via a linear regression
\end{itemize}
\subsubsection{Income Initialization}
\label{sec:init}
\begin{eqnarray}
\rm fiearn &=& {\rm iearnx} * {\rm work}\\
\rm fiearnx &=& \max\left(\min\left(200, {\rm fiearn}\right), 0\right)\\
\rm flogiearn &=& {\rm arcsinh}\left(\rm fiearn\right)/100\\
\rm flogiearnx &=& {\rm arcsinh}\left(\max\left(\min\left(200, {\rm fiearn}\right), 0\right)\right)\\
\rm iearn &=& {\rm iearnx} * \Delta_{\rm NWI} * {\rm work}\\
\rm iearnx &=& \max(\min(200, {\rm iearn}), 0)\\
\rm logiearn &=& {\rm arcsinh}\left(\rm iearn\right)/100\\
\rm logiearnx &=& {\rm arcsinh}\left(\rm iearnx\right)/100\\
\rm gross &=& {\rm iearnx} * 1000 + {\rm iearnx}_{\rm spouse} * 1000
\end{eqnarray}
In addition, hicap\_nonzero and hicap are predicted by regression
models.

\subsubsection{Household Wealth Initialization}
\begin{eqnarray}
\rm hatotax &=& \min\left(2000, {\rm hatota}\right)\\
\rm loghatota &=& {\rm arcsinh}\left(\rm hatota\right)/100\\
\rm loghatatax &=& {\rm arcsinh}\left(\rm hatotax\right)/100
\end{eqnarray}

\subsubsection{Benefits Claiming Initialization}
\begin{eqnarray}
\rm rssclyr &=& \left\{ \begin{tabular}{ll}2100&{\rm
      missing(rssclyr)}\\rssclyr&{\rm else}\end{tabular}\right.\\
\rm ssclaim &=& {\rm rssclyr} \leq {\rm year}\\
\rm rdbclyr &=& \left\{ \begin{tabular}{ll}2100&{\rm
      missing(rdbclyr)}\\rdbclyr&{\rm else}\end{tabular}\right.\\
\rm dbclaim &=& {\rm rdbclyr} \leq {\rm year}\\
\rm ldiclaim &=& \rm diclaim
\end{eqnarray}

\subsection{Single Timestep Flow}
Here are the different modules, in the order in which they are run,
and the relevant encomic calculations done in each.

\subsubsection{Health Module}
The following regressions are run conditionally:
\begin{tabular}{ll}
dbclaim &if fanydb AND NOT ldbclaim\\
ssclaim &if lage $\leq$ 68 AND NOT lssclaim AND
    lage $\geq$ 60\\
diclaim & if lage $<$ 63\\
work &if lage $<$ 73\\
\end{tabular}

\subsubsection{Interventions Module}
This module unconditionally runs the following regressions:
\begin{tabular}{l}
work\\
wlth\_nonzero\\
hicap\_nonzero\\
\end{tabular}

\subsubsection{Earnings Module}
The following regressions are run and the derivative variables
calculated as in \ref{sec:init}:
\begin{tabular}{ll}
hatota & if wlth\_nonzero\\
hicap & if hicap\_nonzero\\
iearn & if work\\
\end{tabular}

\subsubsection{Immigration Module}
This module adjusts the weights of all the records according to the
trends given in the scenario file.

\subsubsection{Initialization Module}
Here is where the calculations from Section~\ref{sec:new} actually
take place.

\subsubsection{Crossectional Module}
In the alzheimers branch, this is where the quality adjusted life
years cross-sectional model is run.

\subsubsection{Economic Outcomes Module}
The following are calculated for everyone:
\begin{eqnarray}
{\rm rssclyr} &=& \left\{\begin{tabular}{ll}year&if ssclaim\\2100&else\end{tabular}\right.\\
{\rm ry\_earn} &=& {\rm iearnx} * 1000\\
{\rm hhwealth} &=& {\rm hatotax} * 1000\\
{\rm db\_tenure} &=& {\rm db\_tenure} + \Delta_T ~{\rm if work AND
  fanydb}\\
{\rm dbpen} &=& {\rm DbBenefit}()\\
{\rm hhearn} &=& {\rm ry\_earn} + {\rm ry\_earn}_{\rm spouse}\\
{\rm hhttlinc} &=& {\rm hhearn}
\end{eqnarray}

\subsubsection{Government Expenses Module}
Government expenditures are generally benefits received by people:
\begin{eqnarray}
{\rm ssben} &=& \left\{\begin{tabular}{ll}12 * {\rm
      SSCalculator.SSBenefit}&{\rm if ssclaim}\\0&{\rm
      else}\end{tabular}\right.\\
{\rm diben} &=& \left\{\begin{tabular}{ll}{\rm
      GovExpModule.DiBenefit}&{\rm if diclaim}\\0&{\rm
      else}\end{tabular}\right.\\
{\rm ssiclaim} &=& \textbf{\color{red}POSSIBLE PROBLEM: CHECKING WITH IGOR}\\
{\rm ssiben} &=& \left\{\begin{tabular}{ll}0&{\rm NOT
      ssiclaim}\\450*12&{\rm age} < 65\\350*12&{\rm age} >=
    65\end{tabular}\right.\\
{\rm gross}&=&\\
{\rm agi}&=&\\
{\rm net}&=&
\end{eqnarray}
\textbf{\color{red} Should SSIBEN be calculated based on age or medicare
  eligibility? Should it change via an inflating factor?}

\subsubsection{Medical Costs Module}
No economic variables are calculated here.

\section{The Poverty Problem}

\begin{Rcode}{!echo,savefig,label=exemptions}
base <- read.dta('../output/ACA/summary.dta')
base <- subset(base, year >= 2010)
a <- base[,'ttl_mcare_pta_exempt'] / base[,'pop_medicare']
b <- base[,'ttl_mcare_ptb_exempt'] / base[,'pop_medicare']
dat <- data.frame(pta_exempt=a, ptb_exempt=b,
ttl_ptb=base[,'ttl_mcare_ptb_exempt'], year=base[,'year'],
pop=base[,'pop_medicare'])
ggplot(dat, aes(x=year)) + 
geom_line(aes(y=ptb_exempt, colour='Fraction')) + 
geom_line(aes(y=ttl_ptb, colour='Total Exempt (M)')) +
geom_line(aes(y=pop, colour='Medicare Eligible (M)')) +
facet_grid(colour ~ ., scales='free_y') +
opts(legend.position='bottom') + labs(colour='Series', y='')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=wlthnonzero}
agesvec <- c('5059','6064','6569','7074','7579','80')
wlthnames <- paste('fraction_wlth_nonzero_', agesvec, sep='')
wlthnonzero = base[,c('year',wlthnames)]
ggplot(wlthnonzero, aes(x=year, ymin=0, ymax=1)) +
geom_line(aes(y=fraction_wlth_nonzero_5059, colour='50-59')) + 
geom_line(aes(y=fraction_wlth_nonzero_6064, colour='60-64')) + 
geom_line(aes(y=fraction_wlth_nonzero_6569, colour='65-69')) +
geom_line(aes(y=fraction_wlth_nonzero_7074, colour='70-74')) +
geom_line(aes(y=fraction_wlth_nonzero_7579, colour='75-79')) +
geom_line(aes(y=fraction_wlth_nonzero_80, colour='80+')) +
opts(legend.position='bottom') +
labs(y='Fraction with Nonzero Wealth', colour='Age Group')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=hicapnonzero}
agesvec <- c('5059','6064','6569','7074','7576','7677','7778','7879','7980','80')
hicapnames <- paste('fraction_hicap_nonzero_', agesvec, sep='')
hicapnonzero = base[,c('year',hicapnames)]
ggplot(hicapnonzero, aes(x=year, ymin=0, ymax=1)) +
geom_line(aes(y=fraction_hicap_nonzero_5059, colour='50-59')) + 
geom_line(aes(y=fraction_hicap_nonzero_6064, colour='60-64')) + 
geom_line(aes(y=fraction_hicap_nonzero_6569, colour='65-69')) +
geom_line(aes(y=fraction_hicap_nonzero_7074, colour='70-74')) +
geom_line(aes(y=fraction_hicap_nonzero_7576, colour='75-79')) +
geom_line(aes(y=fraction_hicap_nonzero_7677, colour='75-79')) +
geom_line(aes(y=fraction_hicap_nonzero_7778, colour='75-79')) +
geom_line(aes(y=fraction_hicap_nonzero_7879, colour='75-79')) +
geom_line(aes(y=fraction_hicap_nonzero_7980, colour='75-79')) +
geom_line(aes(y=fraction_hicap_nonzero_80, colour='80+')) +
opts(legend.position='bottom') +
labs(y='Fraction with Nonzero Capital Income', colour='Age Group')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=exemptbyage}
agesvec <- c('6569','7074','7579','80pp')
exnames <- paste('ttl_mcare_pta_exempt', agesvec, sep='')
popnames <- paste('ttl_mcare_pta_enroll', agesvec, sep='')
exempt <- base[,c('year',exnames)]
exempt[,exnames] <- exempt[,exnames] / base[,popnames]
ggplot(exempt, aes(x=year, ymin=0, ymax=1)) +
geom_line(aes(y=ttl_mcare_pta_exempt6569, colour='65-69')) +
geom_line(aes(y=ttl_mcare_pta_exempt7074, colour='70-74')) +
geom_line(aes(y=ttl_mcare_pta_exempt7579, colour='75-79')) +
geom_line(aes(y=ttl_mcare_pta_exempt80pp, colour='80+')) +
opts(legend.position='bottom') +
labs(y='Fraction Exempt from Premium', colour='Age Group')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=poverty}
y2010 <- read.dta('../output/ACA/detailed_output/y2010_rep1.dta')
y2020 <- read.dta('../output/ACA/detailed_output/y2020_rep1.dta')
y2030 <- read.dta('../output/ACA/detailed_output/y2030_rep1.dta')
y2050 <- read.dta('../output/ACA/detailed_output/y2050_rep1.dta')
pop <- rbind(y2010, y2020, y2030, y2050)
pop <- subset(pop, ldied==0, select=c(year, age, poverty_level, weight))
pop[,'agec'] <- agegrp(pop[,'age'])
pop.split <- split(pop, list(pop[,'agec'], pop[,'year']))

plevel <- sapply(pop.split, FUN=function(x)
weighted.mean(x[,'poverty_level'], x[,'weight']))

n <- do.call(rbind, strsplit(names(plevel), '.', fixed=T))
f <- data.frame(poverty_level = plevel, agec=n[,1], year=n[,2])
ggplot(f) + 
geom_point(aes(x=agec, y=poverty_level, colour=year, ymin=0, ymax=1)) + 
opts(legend.position='bottom')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=povertyfrac}
pfrac <- sapply(pop.split, FUN=function(x) sum(subset(x, poverty_level
<= 1.38)[,'weight'])/sum(x[,'weight']))

g <- data.frame(poverty_fraction=pfrac, agec=n[,1], year=n[,2])
ggplot(g) + 
geom_point(aes(x=agec, y=poverty_fraction, colour=year, ymin=0, ymax=1)) + 
opts(legend.position='bottom')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=oldincome}
base <-  read.dta('../output/ACA/summary.dta')
med.gross <- base[,c('year','gross_mcare')]
hicap <- base[,c('year','hicap_mcare')]
hatota <- base[,c('year','hatota')]
lloghatotax <- base[,c('year','lloghatotax')]

names(med.gross) <- c('year','Value')
names(hicap) <- c('year','Value')
names(hatota) <- c('year','Value')
names(lloghatotax) <- c('year','Value')

h.plus.g <- hicap
h.plus.g[,'Value'] <- hicap[,'Value'] + med.gross[,'Value']

med.gross[,'series'] <- 'Average Gross'
hicap[,'series'] <- 'Average Capital Income'
hatota[,'series'] <- 'Average Household Wealth'
h.plus.g[,'series'] <- 'Average Gross plus Average Capital Income'
lloghatotax[,'series'] <- 'Average Lag of Transformed HaToTaX'

med.gross[,'scale'] <- 'small linear'
hicap[,'scale'] <- 'small linear'
hatota[,'scale'] <- 'large linear'
h.plus.g[,'scale'] <- 'small linear'
lloghatotax[,'scale'] <- 'log-ish'

f <- rbind(med.gross, hicap, h.plus.g, hatota, lloghatotax)
ggplot(f) + geom_line(aes(x=year, y=Value, colour=series)) +
facet_grid(scale ~ ., scales='free_y') +
opts(legend.position='bottom') + scale_colour_brewer(name='Income Source', 'Dark2')
\end{Rcode}

\begin{Rcode}{!echo,savefig,label=grossbreakdown}
ggplot(base) + 
geom_line(aes(x=year, y=ry_earn, colour='Yearly Earnings')) +
geom_line(aes(x=year, y=ssben, colour='SS Benefit')) + 
geom_line(aes(x=year, y=diben, colour='DI Benefit')) +
geom_line(aes(x=year, y=ssiben, colour='SSI Benefit')) +
geom_line(aes(x=year, y=dbpen, colour='DB Pension')) +
opts(legend.position='bottom')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{exemptions}
\caption{Fraction of Medicare Enrollees Below 138\% FPL}
\label{fig:exemptions}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{poverty}
\caption{Average Poverty Level by Age Group}
\label{fig:poverty}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{povertyfrac}
\caption{Fraction Below 138\% of Poverty Level by Age Group}
\label{fig:povertyfrac}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{exemptbyage}
\caption{Fraction Below 138\% of FPL by Age Group Over Time}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{oldincome}
\caption{Means for People Eligible for Medicare}
\label{fig:oldincome}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{grossbreakdown}
\caption{Mean Income Components for Medicare Eligible}
\label{fig:grossbreakdown}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{hicapnonzero}
\caption{Fraction of Population with Nonzero Capital Income}
\end{figure}

\begin{table}[ht]
\centering
\caption{Nonzero Capital Income Regression Coefficients}
\verbatiminput{../output/ACA/FEM_CPP_settings/models/hicap_nonzero.est}
\label{tab:hicapnonzero}
\end{table}

\begin{figure}[ht]
\centering
\recallfig{wlthnonzero}
\caption{Fraction of Population with Nonzero Wealth}
\end{figure}

\begin{table}[ht]
\centering
\caption{Nonzero Wealth Regression Coefficients}
\verbatiminput{../output/ACA/FEM_CPP_settings/models/wlth_nonzero.est}
\label{tab:wlthnonzero}
\end{table}

\begin{table}[ht]
\centering
\caption{Nonzero Wealth Regression Coefficients without FRBYR}
\verbatiminput{../output/noRBYR/FEM_CPP_settings/noRBYR/models/wlth_nonzero.est}
\label{tab:wlthnonzero_nofrbyr}
\end{table}

\begin{Rcode}{!echo}
covlist <- read.table('../FEM_CPP_settings/models/wlth_nonzero.est', stringsAsFactors=F, skip=2)
covlist <- covlist[1:nrow(covlist)-1,]
row.names(covlist) <- covlist[,1]
varnames <- covlist[,1]
newnames <- paste(varnames,'scaled',sep='.')
coeffs <- covlist[,2]
years <- seq(2010, 2050, 10)
dat <- lapply(years, FUN=function(y, vars, cs, newvars) {
  y <- read.dta(file.path('../output/ACA/detailed_output/', paste('y',y,'_rep1.dta', sep='')))
  y <- subset(y, ldied==0)
  c.mat <- matrix(cs, nrow=1)
  y[,newvars] <- y[,vars] * c.mat[rep.int(1,nrow(y)),]
  d <- sapply(newvars, FUN=function(v, dat) weighted.mean(dat[,v], dat[,'weight'], na.rm=T), y)
  names(d) <- newvars
  return(d)}, varnames, coeffs, newnames)
names(dat) <- as.character(years)
dat <- as.data.frame(do.call(rbind, dat))
dat[,'year'] <- as.numeric(row.names(dat))
dat.l <- as.list(dat)
print(dat)
options(warn=2)
g <- ggplot(dat)
for(i in names(dat)) {
  if(i != 'year.scaled' && i != 'year')
    g <- g + geom_line(aes_string(x='year', y=i))
}
g + opts(legend.position='right')
\end{Rcode}

\section{ROC Comparison on frbyr}

ROC curve.

\begin{Statacode}{!echo}
set mem 500M
global ghregdir "../FEM_Stata/Code"
adopath ++ "../FEM_Stata/Makedata/HRS"
adopath ++ "../FEM_Stata/Estimation"
adopath ++ "../FEM_Stata/Code"

use "/homer/c/Retire/FEM/rdata_js/hrs17r_transition.dta", clear
* FOR hacohort = 0 & 1 in wave 2 & 3 no info on SSI claiming
replace ssiclaim = -2 if inlist(hacohort, 0, 1) & inlist(wave,3,4)

replace wlth_nonzero = hatota > 0

*** CHANGE hatota, earnings VARIABLES
drop floghatota floghatotax lloghatota lloghatotax llogiearn llogiearnx flogiearn flogiearnx
set more off
foreach i in hatota hatotax iearn iearnx{
  egen flog`i' = h(f`i')
  replace flog`i' = flog`i'/100
  egen llog`i' = h(l`i')
  replace llog`i' = llog`i'/100
}

*** DEPENDENT VARIABLES
global bin_econ anyhi diclaim ssclaim dbclaim ssiclaim nhmliv work wlth_nonzero hicap_nonzero
global bin_hlth died hearte stroke cancre hibpe diabe lunge memrye
global order smkstat funcstat

*** Create obesity splines
foreach x in lobese_1 lobese_2 lobese_3 loverwt lnormwt fobese_1 fobese_2 fobese_3 foverwt fnormwt {
  cap drop `x'
}

local log_30 = log(30)
mkspline llogbmi_l30 `log_30' llogbmi_30p = llogbmi
mkspline flogbmi_l30 `log_30' flogbmi_30p = flogbmi

global bmivars llogbmi_l30 llogbmi_30p flogbmi_l30 flogbmi_30p

*** GENERATE THE AGE SPLINE VARIABLES
foreach x in lage6061 lage6263 lage64e lage6566 lage6770 lage65l lage6574 lage75l lage75p lage62e lage63e {
  cap drop `x'
}

local age_var age_iwe

gen lage6061 = floor(l`age_var') == 58 | floor(l`age_var') == 59 if l`age_var' < .
gen lage6263 = inrange(floor(l`age_var'),60,61) if l`age_var' < .
gen lage64e = floor(l`age_var') == 62 if l`age_var' < . 
gen lage6566 = floor(l`age_var') == 63 | floor(l`age_var') == 64 if l`age_var' < .
gen lage6770 = inrange(floor(l`age_var'),65,68) if l`age_var' < .

gen lage65l  = min(63,l`age_var') if l`age_var' < .
gen lage6574 = min(max(0,l`age_var'-63),73-63) if l`age_var' < .
gen lage75l = min(l`age_var', 73) if l`age_var' < . 
gen lage75p = max(0, l`age_var'-73) if l`age_var' < .

gen lage62e = floor(l`age_var') == 60 if l`age_var' < .
gen lage63e = floor(l`age_var') == 61 if l`age_var' < .

mkspline la6 58 la7 73 la7p = l`age_var'

gen logdeltaage = log(`age_var' - l`age_var')

*** GENERATE WEAVE DUMMIES
gen w3 = wave == 3
gen w4 = wave == 4
gen w5 = wave == 5
gen w6 = wave == 6
gen w7 = wave == 7

*** INDEPENDENT VARIABLES

*** Demographics
global dvars black hispan hsless college male 
*** Initial values
#d;
global fvars fhearte fstroke fcancre fhibpe fdiabe flunge fsmokev fsmoken fiadl1 fadl12 fadl3 
fwidowed fsingle fwork flogiearnx fwlth_nonzero floghatotax 
flogaime flogq fshlt fanydb frdb_na_2 frdb_na_3 frdb_na_4 fanydc flogdcwlthx ;
#d cr
*** values of health variables at t-1
global lvars_hlth lhearte lstroke lcancre lhibpe ldiabe llunge liadl1 ladl12 ladl3 lsmoken lwidowed 
*** values of econ variables at time t-1
global lvars_econ lwork llogiearnx lwlth_nonzero lloghatotax ldiclaim lssiclaim lssclaim ldbclaim lnhmliv

*** FOR MORTALITY
global allvars_died $dvars lage65l lage6574 lage75p $lvars_hlth $fvars

*** FOR CHRONIC CONDITIONS AND ORDINAL OUTCOMES
global allvars_hlth $dvars lage65l lage6574 lage75p $lvars_hlth  $fvars $bmivars logdeltaage

takestring, oldlist($allvars_hlth) newname("allvars_hearte") extlist("lhearte fhearte lstroke llunge lcancre liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_stroke") extlist("lstroke fstroke lstroke llunge liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_cancre") extlist("lhearte lstroke lcancre lhibpe ldiabe llunge fcancre liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_hibpe")  extlist("lhearte lstroke lcancre lhibpe llunge fhibpe liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_diabe")  extlist("lhearte lstroke lcancre lhibpe ldiabe llunge fdiabe liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_lunge")  extlist("lhearte lstroke lcancre lhibpe ldiabe llunge flunge liadl1 ladl12 ladl3")
takestring, oldlist($allvars_hlth) newname("allvars_memrye") extlist("lhearte lstroke lcancre lhibpe ldiabe llunge flunge")

global allvars_smkstat $allvars_hlth
global allvars_funcstat $allvars_hlth

global allvars_logbmi $allvars_hlth frbyr

*** FOR ECONOMIC OUTCOMES
global allvars_econ1 $dvars lage65l lage6574 lage75p $lvars_hlth $lvars_econ $fvars logdeltaage
global allvars_econ2 $dvars lage6061 lage62e lage63e lage64e lage6566 lage6770 $lvars_hlth $lvars_econ $fvars logdeltaage
global allvars_econ3 $dvars la6 la7 la7p $lvars_hlth $lvars_econ $fvars w3 w4 w5 w6 w7 logdeltaage
global allvars_econ4 $dvars lage65l lage6574 lage75p $lvars_hlth $lvars_econ $fvars logdeltaage

takestring, oldlist($allvars_econ1) newname("allvars_anyhi")  extlist("lage75p ldbclaim lssiclaim lnhmliv lsmoken")
takestring, oldlist($allvars_econ1) newname("allvars_diclaim")  extlist("lage75p lssiclaim lssclaim ldbclaim lnhmliv lsmoken")
takestring, oldlist($allvars_econ1) newname("allvars_dbclaim")  extlist("lage75p fwork fanydb lssiclaim ldbclaim lwork lnhmliv lsmoken")
takestring, oldlist($allvars_econ1) newname("allvars_ssiclaim")  extlist("lnhmliv lsmoken")
takestring, oldlist($allvars_econ4) newname("allvars_nhmliv")  extlist("ldiclaim lssiclaim lssclaim ldbclaim lwork llogiearnx lsmoken frdb_na_4")
takestring, oldlist($allvars_econ1) newname("allvars_iearnx")  extlist("lssiclaim lnhmliv lsmoken lage75p")

takestring, oldlist($allvars_econ2) newname("allvars_ssclaim")  extlist("lssiclaim lssclaim lnhmliv lsmoken")
takestring, oldlist($allvars_econ2) newname("allvars_work")  extlist("lssiclaim lnhmliv lsmoken")

takestring, oldlist($allvars_econ3) newname("allvars_wlth_nonzero")  extlist("lssiclaim lsmoken")
takestring, oldlist($allvars_econ3) newname("allvars_hatotax")  extlist("lssiclaim lsmoken")
takestring, oldlist($allvars_econ3) newname("allvars_hicap_nonzero") extlist("lssiclaim lsmoken")
takestring, oldlist($allvars_econ3) newname("allvars_hicap") extlist("lssiclaim, lsmoken")

/*********************************************************************/
* ESTIMATE BINARY OUTCOMES
/*********************************************************************/
probit wlth_nonzero $allvars_wlth_nonzero if wlth_nonzero!=-2&wlth_nonzero!=9
lroc, nograph
probit wlth_nonzero $allvars_wlth_nonzero frbyr if wlth_nonzero!=-2&wlth_nonzero!=9
lroc, nograph

\end{Statacode}

\end{document}

### 2019-11-22 analysis/medicare-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{datatool}

\title{Analysis of Medicare Reform Options}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}{!echo,hide}
library(foreign)
library(ggplot2)
\end{Rcode}

\begin{Statacode}{!echo,hide}
set linesize 100
global discount_rate = 0.03
\end{Statacode}

\section{Summary of Results}

Figure~\ref{fig:summary} shows a comparison of the costs of various Medicare reform scenarios. Each variant shows a particular parameter set for the examined scenarios. A complete listing of the excursions and their parameters is in Table~\ref{tab:scenarios}. The ACA schedule for Part B premium thresholds freezes them in nominal dollars until 2019 then allows them to rise with inflation, while the nominal dollars scenario never adjusts for inflation and the real dollars scenario always does.

\begin{Statacode}{!echo,hide,fig,savefig,label=aca_totmd}
use ../output/ACA/summary, clear
tsset year
tsfill
foreach v of varlist ttl_totmd_mcare {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist ttl_totmd_mcare {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected ttl_totmd_mcare year
outsheet year ttl_totmd_mcare using aca_totmd.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{aca_totmd}
\caption{NPV of Total Medical Expenditures for Medicare Eligible}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=cost_summary}
use ../output/ACA/summary, clear
gen summ_ACA = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_ACAGross = ttl_mcare
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
gen summ_nominal = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_nominalGross = ttl_mcare
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
gen summ_drop00 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop00Gross = ttl_mcare
label variable summ_drop00 "Part A Drop 0 ACA Thresholds"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
gen summ_drop03 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop03Gross = ttl_mcare
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
gen summ_drop0310 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop0310Gross = ttl_mcare
label variable summ_drop0310 "Part A Drop 3, 10 Percent"
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
gen summ_drop05 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop05Gross = ttl_mcare
label variable summ_drop05 "Part A Drop 5 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
gen summ_gdp = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
gen summ_gdpGross = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace summ_gdp = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium if year < 2012
replace summ_gdpGross = ttl_mcare if year < 2012
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
gen summ_cpi = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
gen summ_cpiGross = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace summ_cpi = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium if year < 2012
replace summ_cpiGross = ttl_mcare if year < 2012
label variable summ_cpi "Vouchers CPI"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
gen summ_age67 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age67Gross = ttl_mcare
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/elig.age70/summary, sort
gen summ_age70 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age70Gross = ttl_mcare
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
gen summ_acanolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_acanolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
gen summ_partAnolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_partAnolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
gen summ_gdpnolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_gdpnolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
gen summ_age67nolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age67nolimitGross = ttl_mcare
keep year summ_*

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Trillions of 2010 Dollars")
outsheet year summ_* using cost_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{cost_summary}
\caption{Net Present Value of Medicare Reform Options}
\label{fig:summary}
\end{figure}

\begin{table}[ht]
\centering
\caption{Net Present Value of Reform Options (Trillions 2010 Dollars)}
\recallout{cumm_cost}
\label{tab:cumm_cost}
\end{table}

\begin{Statacode}{!echo,hide,fig,savefig,label=enroll_a_summary}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll summ_ACA
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
rename ttl_mcare_pta_enroll summ_nominal
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
rename ttl_mcare_pta_enroll summ_drop03
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
rename ttl_mcare_pta_enroll summ_gdp
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
rename ttl_mcare_pta_enroll summ_age67
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_pta_enroll summ_cpi
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
rename ttl_mcare_pta_enroll summ_drop00
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
rename ttl_mcare_pta_enroll summ_drop0310
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
rename ttl_mcare_pta_enroll summ_drop05
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_mcare_pta_enroll summ_ACAnolimit
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_mcare_pta_enroll summ_partAnolimit
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_mcare_pta_enroll summ_gdpnolimit
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_mcare_pta_enroll summ_age67nolimit
keep year summ_*

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("People Enrolled in Part A (Millions)")
outsheet year summ_* using enroll_a_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{enroll_a_summary}
\caption{Annual Enrollment in Medicare Part A}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=enroll_b_summary}
use ../output/ACA/summary, clear
rename ttl_mcare_ptb_enroll summ_ACA
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
rename ttl_mcare_ptb_enroll summ_nominal
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
rename ttl_mcare_ptb_enroll summ_drop03
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
rename ttl_mcare_ptb_enroll summ_gdp
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
rename ttl_mcare_ptb_enroll summ_age67
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_ptb_enroll summ_cpi
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
rename ttl_mcare_ptb_enroll summ_drop00
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
rename ttl_mcare_ptb_enroll summ_drop0310
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
rename ttl_mcare_ptb_enroll summ_drop05
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_ACAnolimit
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_partAnolimit
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_gdpnolimit
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_age67nolimit
keep year summ_*

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("People Enrolled in Part B (Millions)")
outsheet year summ_* using enroll_b_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{enroll_b_summary}
\caption{Annual Enrollment in Medicare Part B}
\end{figure}

\begin{Statacode}{!echo,hide,savefig,label=uncovered_upper}
use ../output/ACA/summary, clear
rename ttl_uncovered_upper_mcare summ_aca
label variable summ_aca "ACA"
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03/summary, sort
rename ttl_uncovered_upper_mcare summ_parta
label variable summ_parta "Part A Drop 3 ACA Thresholds"
keep year summ_* unc_6566
merge year using ../output/voucher.gdp/summary, sort
rename ttl_uncovered_upper_mcare summ_voucher
label variable summ_voucher "Voucher GDP"
keep year summ_* unc_6566
merge year using ../output/elig.age67/summary, sort
rename ttl_uncovered_upper_mcare summ_age67
replace summ_age67 = summ_age67 + unc_6566 if year > 2010
label variable summ_age67 "Eligibility Age 67"
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_ACAnolimit
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_partAnolimit
keep year summ_* unc_6566
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_gdpnolimit
keep year summ_* unc_6566
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_age67nolimit
replace summ_age67nolimit = summ_age67nolimit + unc_6566 if year > 2010
keep year summ_* unc_6566

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_aca-summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("NPV of OOP Expenditures (Trillions)")
outsheet year summ_* using uncovered_upper.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{uncovered_upper}
\caption{Upper Bound on Total Uncovered Expenditures for Traditional Medicare Eligible Population}
\end{figure}

\begin{Statacode}{!echo,hide,savefig,label=uncovered_lower}
use ../output/ACA/summary, clear
rename ttl_uncovered_lower_mcare summ_aca
label variable summ_aca "ACA"
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03/summary, sort
rename ttl_uncovered_lower_mcare summ_parta
label variable summ_parta "Part A Drop 3 ACA Thresholds"
keep year summ_* unc_6566
merge year using ../output/voucher.gdp/summary, sort
rename ttl_uncovered_lower_mcare summ_voucher
label variable summ_voucher "Voucher GDP"
keep year summ_* unc_6566
merge year using ../output/elig.age67/summary, sort
rename ttl_uncovered_lower_mcare summ_age67
replace summ_age67 = summ_age67 + 18.0/41 * unc_6566 if year > 2010
label variable summ_age67 "Eligibility Age 67"
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_ACAnolimit
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_partAnolimit
keep year summ_* unc_6566
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_gdpnolimit
keep year summ_* unc_6566
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_age67nolimit
replace summ_age67nolimit = summ_age67nolimit + unc_6566 if year > 2010
keep year summ_* unc_6566

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_aca-summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("NPV of OOP Expenditures (Trillions)")
outsheet year summ_* using uncovered_lower.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{uncovered_lower}
\caption{Lower Bound on Total Uncovered Expenditures for Traditional Medicare Eligible Population}
\end{figure}

\begin{table}[ht]
\centering
\caption{Scenario Definitions}\vspace{0.2in}
\begin{tabular}{|l|c|c|c|c|c|c|c|}\hline
 & \textbf{Medicare} & \textbf{Part A} & \textbf{Part A} & \textbf{Part A} & \textbf{Part B} & \textbf{Part B} & \textbf{Medical Cost} \\
\textbf{Scenario} & \textbf{Eligibility} & \textbf{Premium} & \textbf{Drop} & \textbf{Premium} & \textbf{Premium} & \textbf{Premium} & \textbf{Growth} \\
& \textbf{Age} & \textbf{Schedule} & \textbf{Fraction} & \textbf{Elasticity} & \textbf{Elasticity} & \textbf{Thresholds} & \textbf{Rate}\\\hline
ACA & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
Zero Elasticity & 65 & 0\% & 0\% & 0.0 & 0.0 & ACA & SSA Intermediate\\\hline
-0.2 Elasticity & 65 & 0\% & 0\% & 0.0 & -0.2 & ACA & SSA Intermediate\\\hline
-0.5 Elasticity & 65 & 0\% & 0\% & 0.0 & -0.5 & ACA & SSA Intermediate\\\hline
Part B Nominal Dollars & 65 & 0\% & 0\% & 0.0 & -0.1 & Nominal & SSA Intermediate\\\hline
Part B Real Dollars & 65 & 0\% & 0\% & 0.0 & -0.1 & 2010 & SSA Intermediate\\\hline
Part B 25 Percent & 65 & 0\% & 0\% & 0.0 & -0.1 & N/A & SSA Intermediate\\\hline
Part A Drop 3 & 65 & ACA & 3\% & -0.1 & -0.1 & ACA & SSA Intermediate\\\hline
Vouchers GDP & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & OASDI GDP \\\hline
Vouchers GDP+1 & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & OASDI GPD + 1\%\\\hline
Eligibility Age 67 & 67 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
Eligibility Age 70 & 70 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
\end{tabular}
\label{tab:scenarios}
\end{table}

\section{Imposing a Premium on Part A}
Starting in 2010, premiums are imposed for Part A according to Table~\ref{tab:sliding}. The thresholds do not grow with inflation, so in real-dollar terms they decline over time. Any household under 138\% of the poverty line does not have any premium. Figure~\ref{fig:exemptions} shows the evolution of this population over time.

\begin{table}[ht]
\caption{Individual Sliding Scale Premiums as percentage of per-enrollee program cost}
\label{tab:sliding}
\centering
\begin{tabular}{|c|c|c|c|}\hline
\textbf{Band} & \textbf{Income Bracket} & \textbf{Part A Premium} & \textbf{Part B Premium}\\\hline
Band 1 & MAGI <= \$85,000 & 25\% & 25\%\\\hline
Band 2 & \$85,000 < MAGI <= \$107,000 & 35\% & 35\%\\\hline
Band 3 & \$107,000 < MAGI <= \$160,000  & 55\% & 55\%\\\hline
Band 4 & \$160,000 < MAGI <= \$214,000 & 65\% & 65\%\\\hline
Band 5 & \$214,000 < MAGI & 80\% & 80\%\\\hline
\end{tabular}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partB.real.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=income_pops_real}
gen b0 = income_band0/pop_medicare_age
gen b1 = income_band1/pop_medicare_age
gen b2 = income_band2/pop_medicare_age
gen b3 = income_band3/pop_medicare_age
gen b4 = income_band4/pop_medicare_age
gen b5 = income_band5/pop_medicare_age
gen pool = pop_medicare_age
global vlist b0 b1 b2 b3 b4 b5 pool
format $vlist %5.2f
outsheet year $vlist if year == 2010 | year == 2020 | year == 2036 using income_pops_real.csv, comma replace
\end{Statacode}
\begin{Statacode}{!echo,hide}
use ../output/partB.nominal.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=income_pops_nominal}
gen b0 = income_band0/pop_medicare_age
gen b1 = income_band1/pop_medicare_age
gen b2 = income_band2/pop_medicare_age
gen b3 = income_band3/pop_medicare_age
gen b4 = income_band4/pop_medicare_age
gen b5 = income_band5/pop_medicare_age
gen pool = pop_medicare_age
global vlist b0 b1 b2 b3 b4 b5 pool
format $vlist %5.2f
outsheet year $vlist if year == 2010 | year == 2020 | year == 2036 using income_pops_nominal.csv, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,Medicaid,Band 1,Band 2,Band 3,Band 4,Band 5,Total 65+}]{income_pops_real}{income_pops_real.csv}
\DTLloaddb[headers={Year,Medicaid,Band 1,Band 2,Band 3,Band 4,Band 5,Total 65+}]{income_pops_nominal}{income_pops_nominal.csv}

\begin{table}[ht]
\centering
\caption{Percentage in Real Dollar Threshold Bands by Year (Millions of People)}
\DTLdisplaydb{income_pops_real}
\label{tab:income_pops_real}
\end{table}

\begin{table}[ht]
\centering
\caption{Percentage in Nominal Dollar Threshold Bands by Year (Millions of People)}
\DTLdisplaydb{income_pops_nominal}
\label{tab:income_pops_nominal}
\end{table}

We looked at four different scenarios all based on Buchmueller, 2006. When premium jumps from nothing to the sliding-scale premium, there is a small chance that the person will drop (or not enroll in the first place). Once enrolled, all further premium adjustments are subject to a -0.1 elasticity to price and a takeup probit based on the model for Part B takeup. The average premium paid does not change significantly from scenario to scenario.

\begin{Statacode}{!echo,hide,fig,savefig,label=exemptions}
use ../output/ACA/summary, clear
gen bpop = ttl_mcare_ptb_exempt / pop_medicare
label variable bpop "Portion of Medicare Eligible Exempt from Part B Premium"
gen benroll = ttl_mcare_ptb_exempt / ttl_mcare_ptb_enroll
label variable benroll "Portion of Part B Enrolled Exempt from Premium under ACA"
keep year bpop benroll
sort year
merge year using ../output/partA.drop03/summary
gen aenroll = ttl_mcare_pta_exempt / ttl_mcare_pta_enroll
label variable aenroll "Portion of Part A Enrolled Exempt from Premium under Drop 3"
twoway connected bpop aenroll benroll year if year >= 2010, legend(cols(1))
outsheet year bpop aenroll benroll using exemptions.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{exemptions}
\caption{Fraction of Medicare Enrollees Exempt from Premium in ACA}
\label{fig:exemptions}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=drop_enrolled}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll base
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
rename ttl_mcare_pta_enroll drop03
label variable drop03 "Part A Drop 3 ACA Thresholds"
keep year base drop03
sort year
merge year using ../output/partA.drop03.real.dollars/summary
rename ttl_mcare_pta_enroll drop03Real
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
rename ttl_mcare_pta_enroll drop0325
label variable drop0325 "Part A Drop 3 25 Percent"
keep year base drop03 drop03Real drop0325
sort year
merge year using ../output/partA.drop03.10percent/summary
rename ttl_mcare_pta_enroll drop0310
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Number of Enrolled (Millions)")
outsheet year base drop03 drop03Real drop0310 drop0325 using drop_enrolled.csv if year >= 2010, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,ACA,Drop 03, Drop 03 Real,Drop 03 10 Percent,Drop 03 25 Percent}]{parta_enrolled}{drop_enrolled.csv}

\begin{table}[ht]
\centering
\caption{Enrollment in Part A}
\DTLdisplaydb{parta_enrolled}
\label{tab:parta_c_cost}
\end{table}

\begin{figure}[ht]
\centering
\recallfig{drop_enrolled}
\caption{Enrollment in Part A}
\label{fig:drop_enrolled}
\end{figure}

Part A costs (Figure~\ref{fig:drop_costs}) are relative to the enrollment, but the net costs are roughly the same, as shown in Figure~\ref{fig:parta_net_costs}.
\begin{Statacode}{!echo,hide,fig,savefig,label=drop_costs}
use ../output/ACA/summary, clear
gen base = ttl_mcare_pta
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
gen drop03 = ttl_mcare_pta
label variable drop03 "Part A Drop 3 ACA Thresholds"
keep year base drop03
sort year
merge year using ../output/partA.drop03.real.dollars/summary
rename ttl_mcare_pta drop03Real
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
rename ttl_mcare_pta drop0325
label variable drop0325 "Part A Drop 3 25 Percent"
keep year base drop03 drop03Real drop0325
sort year
merge year using ../output/partA.drop03.10percent/summary
rename ttl_mcare_pta drop0310
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base drop03 drop03Real drop0310 drop0325 using drop_costs.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{drop_costs}
\caption{Part A Total Costs with varying drop rates (excluding payroll tax revenue)}
\label{fig:drop_costs}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=parta_net_costs}
use ../output/ACA/summary, clear
gen base = ttl_mcare_pta
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
gen drop03 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop03 "Part A Drop 3 ACA Thresholds"
sort year
keep year base drop03
merge year using ../output/partA.drop03.real.dollars/summary
gen drop03Real = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
gen drop0325 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop0325 "Part A Drop 3 25 Percent"
sort year
keep year base drop03 drop03Real drop0325
merge year using ../output/partA.drop03.10percent/summary
gen drop0310 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base drop03 drop03Real drop0310 drop0325 using parta_net_costs.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{parta_net_costs}
\caption{Part A Net Costs with varying drop rates (excluding payroll tax revenue)}
\label{fig:parta_net_costs}
\end{figure}

\begin{Statacode}{!echo,hide}
tsset year
tsfill
foreach v in base drop03 {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace drop03 = drop03 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
gen c_base = sum(base) if year >= 2010
gen c_drop03 = sum(drop03) if year >= 2010
gen diffDrop03 = c_drop03 - c_base
format c_base diffDrop03 %10.2f
outsheet year c_base diffDrop03 if year >= 2010 using parta_c_cost.csv, replace comma
\end{Statacode}
\DTLloaddb[headers={Year,ACA,Drop 03}]{parta_c_cost}{parta_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Part A Options Less Net Present Value of ACA Part A (Billions 2010 Dollars)}
\DTLdisplaydb{parta_c_cost}
\label{tab:parta_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partA.drop03/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=parta_premium}
gen basePremium = 0.25 * mcare_pta
gen ptaPremium = mcare_pta_premium_paid
format basePremium ptaPremium %10.2f
outsheet year basePremium ptaPremium if year >= 2010 using parta_premium.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,Base Premium,Average Premium}]{parta_premium}{parta_premium.csv}

\begin{table}[ht]
\centering
\caption{Part A Drop 3 Nominal and Average Part A Premiums}
\DTLdisplaydb{parta_premium}
\label{tab:parta_premium}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
Here is the code for calculating the deviation:
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/partA.drop03/summary
gen drop_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen drop_a_mean = mcare_pta - mcare_pta_premium_paid
gen drop_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen drop_d_mean = mcare_ptd
gen drop_a_enrl = ttl_mcare_pta_enroll * 1e6
gen drop_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen drop_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* drop_* pop_medicare {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen baseline = base_ttl / 1e9
gen dropCost = drop_ttl / 1e9
gen enrollment_deviation = (base_ttl_a * (drop_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (drop_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (drop_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_deviation = (base_a_enrl * (drop_a_mean - base_a_mean) + base_b_enrl * (drop_b_mean - base_b_mean) + base_d_enrl * (drop_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace dropCost = dropCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_deviation = enrollment_deviation * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_deviation = outlay_deviation * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $$ syntax highlighting fix
gen cBaseline = sum(baseline) if year >= 2010
gen cDropCost = sum(dropCost) if year >= 2010
gen enrollment = sum(enrollment_deviation) if year >= 2010
gen outlay = sum(outlay_deviation) if year >= 2010
format cBaseline enrollment outlay %10.2f
outsheet year cBaseline cDropCost enrollment outlay using parta_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,Real Dollars Cost,Drop 03,Enrollment Deviation,Enrollee Outlay Deviation}]{parta_deviation}{parta_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs for Drop 03 from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{parta_deviation}
\label{tab:parta_deviation}
\end{table}

\section{Impact of Sliding Scale on Part B}

Since Medicare Part B already has a premium, we don't need to make any assumptions about people dropping when a change is introduced, we can just use the elasticity directly. The current ACA plan is to keep the premium thresholds in nominal dollars. We also ran two scenarios with thresholds in real dollars (2010) and with no thresholds where everyone not exempt pays the same premium (25\%).

Figure~\ref{fig:partB_enrollment} shows the change in enrollment fraction and Figure~\ref{fig:partB_cost} shows the change in net costs. Since the 25\% premium is a continuation of previous policy, it is not surprising that enrollment continues relatively flat. The initial large discrepancy between real and nominal dollar thresholds is due to higher average premiums under the nominal dollar scheme as shown in Figure~\ref{fig:partB_premium}. The convergence is due to the lowest threshold falling below the exemption threshold.

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_enrollment}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb_enroll
keep year base
label variable base "ACA"
merge year using ../output/partB.25percent/summary
gen p25 = ttl_mcare_ptb_enroll
label variable p25 "Part B 25 Percent"
keep year base p25
sort year
merge year using ../output/partB.real.dollars/summary
gen real = ttl_mcare_ptb_enroll
label variable real "Part B Real Dollars"
keep year base p25 real
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = ttl_mcare_ptb_enroll
label variable nominal "Part B Nominal Dollars"
twoway connected base nominal real p25 year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Number of Persons Enrolled (Millions)")
outsheet year base nominal real using partB_enrollment.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_cost}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable base "ACA"
keep year base
merge year using ../output/partB.25percent/summary
gen p25 = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable p25 "Part B 25 Percent"
keep year base p25
sort year
merge year using ../output/partB.real.dollars/summary
gen real = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable real "Part B Real Dollars"
keep year base p25 real
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable nominal "Part B Nominal Dollars"
twoway connected base nominal real p25 year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base nominal real p25 using partB_cost.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{partB_enrollment}
\caption{Medicare Part B Enrollment}
\label{fig:partB_enrollment}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{partB_cost}
\caption{Medicare Part B Net Cost (excluding payroll tax revenue)}
\label{fig:partB_cost}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=partb_c_cost}
tsset year
tsfill
foreach v in base p25 real nominal {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace p25 = p25 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace nominal = nominal * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace real = real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
gen c_base = sum(base)  if year >= 2010
gen c_p25 = sum(p25) if year >= 2010
gen c_real = sum(real) if year >= 2010
gen c_nominal = sum(nominal) if year >= 2010
gen diffP25 = c_p25 - c_base
gen diffReal = c_real - c_base
gen diffNominal = c_nominal - c_base
global vlist diffP25 diffReal diffNominal
format c_base $vlist %10.2f
outsheet year c_base $vlist if year >= 2010 using partb_c_cost.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,ACA,25\% Premium,Real Dollar Thresholds,Nominal Dollar Thresholds}]{partb_c_cost}{partb_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Part B Options Less Net Present Value of ACA Part B (Billions 2010 Dollars)}
\DTLdisplaydb{partb_c_cost}
\label{tab:partb_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partB.25percent/summary, clear
twoway connected mcare_ptb_enroll5564 mcare_ptb_enroll6569 mcare_ptb_enroll7074 mcare_ptb_enroll7579 mcare_ptb_enroll80pp year if year >= 2010, yscale(titlegap(*2)) ytitle("Fraction of Eligible Enrolled") legend(cols(1))
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=premium_summary}
use ../output/ACA/summary, clear
gen baseB = mcare_ptb_premium_paid
label variable baseB "ACA"
keep year baseB
sort year
merge year using ../output/partB.real.dollars/summary
gen real = mcare_ptb_premium_paid
label variable real "Part B Real Dollars"
keep year baseB real
sort year
merge year using ../output/partB.25percent/summary
gen partB25 = mcare_ptb_premium_paid
label variable partB25 "Part B 25 Percent"
keep year baseB real partB25
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = mcare_ptb_premium_paid
label variable nominal "Part B Nominal Dollars"
twoway connected baseB real nominal partB25 year if year >= 2010, legend(cols(1)) ytitle("Yearly Premium in 2010 Dollars")
outsheet year baseB real nominal partB25 using  premium_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{premium_summary}
\caption{Average Part B Premium Paid}
\label{fig:partB_premium}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_elasticity}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb_enroll
label variable base "ACA"
keep base year
sort year
merge year using ../output/mcare.0elas/summary
gen zero = ttl_mcare_ptb_enroll
label variable zero "Zero Elasticity"
keep base zero year
sort year
merge year using ../output/mcare.neg2elas/summary
gen neg2 = ttl_mcare_ptb_enroll
label variable neg2 "-0.2 Elasticity"
keep base zero neg2 year
sort year
merge year using ../output/mcare.neg5elas/summary
gen neg5 = ttl_mcare_ptb_enroll
label variable neg5 "-0.5 Elasticity"
twoway connected zero neg2 base neg5 year if year >= 2010, legend(cols(1)) ytitle("Number of Enrolled (Millions)")
outsheet year zero neg2 base neg5 using partB_elasticity.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{partB_elasticity}
\caption{Effect of Elasticity on Part B Enrollment}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/partB.real.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen real_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen real_a_mean = mcare_pta - mcare_pta_premium_paid
gen real_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen real_d_mean = mcare_ptd
gen real_a_enrl = ttl_mcare_pta_enroll * 1e6
gen real_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen real_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year real_*
sort year
merge year using ../output/ACA/summary
gen aca_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen aca_a_mean = mcare_pta - mcare_pta_premium_paid
gen aca_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen aca_d_mean = mcare_ptd
gen aca_a_enrl = ttl_mcare_pta_enroll * 1e6
gen aca_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen aca_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen aca_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen aca_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen aca_ttl_d = ttl_mcare_ptd * 1e9
keep year real_* aca_*
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nomin_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen nomin_a_mean = mcare_pta - mcare_pta_premium_paid
gen nomin_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen nomin_d_mean = mcare_ptd
gen nomin_a_enrl = ttl_mcare_pta_enroll * 1e6
gen nomin_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen nomin_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist real_* aca_* nomin_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen realCost = real_ttl / 1e9
gen acaCost = aca_ttl / 1e9
gen nominCost = nomin_ttl / 1e9
gen enrollment_real = (aca_ttl_a * (real_a_enrl - aca_a_enrl)/aca_a_enrl + aca_ttl_b * (real_b_enrl - aca_b_enrl)/aca_b_enrl + aca_ttl_d * (real_d_enrl - aca_d_enrl)/aca_d_enrl) / 1e9
gen outlay_real = (aca_a_enrl * (real_a_mean - aca_a_mean) + aca_b_enrl * (real_b_mean - aca_b_mean) + aca_d_enrl * (real_d_mean - aca_d_mean)) / 1e9
gen enrollment_nomin = (aca_ttl_a * (nomin_a_enrl - aca_a_enrl)/aca_a_enrl + aca_ttl_b * (nomin_b_enrl - aca_b_enrl)/aca_b_enrl + aca_ttl_d * (nomin_d_enrl - aca_d_enrl)/aca_d_enrl) / 1e9
gen outlay_nomin = (aca_a_enrl * (nomin_a_mean - aca_a_mean) + aca_b_enrl * (nomin_b_mean - aca_b_mean) + aca_d_enrl * (nomin_d_mean - aca_d_mean)) / 1e9
replace realCost = realCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace acaCost = acaCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace nominCost = nominCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_real = enrollment_real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_real = outlay_real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_nomin = enrollment_nomin * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_nomin = outlay_nomin * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cReal = sum(realCost) if year >= 2010
gen cACA = sum(acaCost) if year >= 2010
gen cNomin = sum(nominCost) if year >= 2010
gen cummEnrollmentReal = sum(enrollment_real) if year >= 2010
gen cummOutlayReal = sum(outlay_real) if year >= 2010
gen cummEnrollmentNominal = sum(enrollment_nomin) if year >= 2010
gen cummOutlayNominal = sum(outlay_nomin) if year >= 2010
global vlist cACA  cReal cummEnrollmentReal cummOutlayReal cNomin cummEnrollmentNominal cummOutlayNominal 
format $vlist %10.2f
outsheet year $vlist using partb_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{Real\\Dollars\\Cost},\shortstack{Real\\Dollars\\Enrollment\\Deviation}, \shortstack{Real\\Dollars\\Enrollee\\Outlay\\Deviation}, \shortstack{Nominal\\Dollars\\Cost},\shortstack{Nominal\\Dollars\\Enrollment\\Deviation}, \shortstack{Nominal\\Dollars\\Enrollee\\Outlay\\Deviation}}]{partb_deviation}{partb_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{partb_deviation}
\label{tab:partb_deviation}
\end{table}

\section{Medicare Vouchers can Save Money}

The analysis here assumes that the government will issue vouchers whose value will grow at a rate different from the Medicare cost growth rate. The difference between the cost of vouchers and the projected cost of Medicare would be the savings (or loss) to the government. In the base scenario, there are no vouchers and Medicare costs grow as expected. The first voucher scenario grows at the rate of GDP (normalized by labor-force growth). The second scenario grows at GDP plus 1\%. Results are shown in Figure~\ref{fig:vouchers}.

The GDP plus 1\% scenario is more expensive in the out years because the Medicare cost growth is projected to fall below GDP+1\% by 2034 and to stay below until the end of the simulation period. In the near term, either voucher scenario will save money since Medicare costs are rising at about 1.5\% above GDP.

\begin{Statacode}{!echo,hide,fig,savefig,label=voucher_a_enroll}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll base
label variable base "ACA Part A Enrollment"
keep year base
merge year using ../output/voucher.gdp/summary
rename ttl_mcare_pta_enroll gdp
label variable gdp "GDP Part A Enrollment"
sort year
keep year base gdp
merge year using ../output/voucher.gdp.plus1/summary
rename ttl_mcare_pta_enroll plus1
label variable plus1 "GDP Plus 1 Part A Enrollment"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_pta_enroll cpi
label variable cpi "CPI Part A Enrollment"
twoway connected base gdp cpi year if year >= 2010, yscale(titlegap(*2)) ytitle("Number Enrolled (Millions) in Part A") legend(cols(1))
outsheet year base gdp plus1 cpi using voucher_a_enroll.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=voucher_b_enroll}
use ../output/ACA/summary, clear
rename ttl_mcare_ptb_enroll base
label variable base "ACA Part B Enrollment"
keep year base
merge year using ../output/voucher.gdp/summary
rename ttl_mcare_ptb_enroll gdp
label variable gdp "GDP Part B Enrollment"
sort year
keep year base gdp
merge year using ../output/voucher.gdp.plus1/summary
rename ttl_mcare_ptb_enroll plus1
label variable plus1 "GDP Plus 1 Part B Enrollment"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_ptb_enroll cpi
label variable cpi "CPI Part B Enrollment"
twoway connected base gdp cpi year if year >= 2010, yscale(titlegap(*2)) ytitle("Number Enrolled (Millions) in Part B") legend(cols(1))
outsheet year base gdp plus1 cpi using voucher_b_enroll.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=vouchers}
use ../output/ACA/summary, clear
gen base_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
keep year base_medicare
label variable base_medicare "ACA"
merge year using ../output/voucher.gdp/summary
sort year
gen gdp_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace gdp_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable gdp_medicare "Vouchers GDP"
keep year base_medicare gdp_medicare
merge year using ../output/voucher.gdp.plus1/summary
gen plus1_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace plus1_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable plus1_medicare "Vouchers GDP+1"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
gen cpi_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace cpi_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable cpi_medicare "Vouchers CPI"
twoway connected base_medicare gdp_medicare cpi_medicare year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Total Medicare Cost (Billions of 2010 Dollars)")
outsheet year base_medicare gdp_medicare plus1_medicare cpi_medicare using vouchers.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{vouchers}
\caption{Comparison of Medicare Cost Growth with Vouchers - Long Term}
\label{fig:vouchers}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{voucher_a_enroll}
\caption{Number Enrolled in Part A Under Voucher Scenarios}
\label{fig:voucher_enroll}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{voucher_b_enroll}
\caption{Number Enrolled in Part B Under Voucher Scenarios}
\label{fig:voucher_enroll}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=voucher_c_cost}
tsset year
tsfill
foreach v in base_medicare gdp_medicare plus1_medicare {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base_medicare = base_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace gdp_medicare = gdp_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace plus1_medicare = plus1_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ for syntax highlights
gen c_base = sum(base_medicare) if year >= 2010
gen c_gdp = sum(gdp_medicare) if year >= 2010
gen c_plus1 = sum(plus1_medicare) if year >= 2010
gen diffGDP = c_gdp - c_base
gen diffPlus1 = c_plus1 - c_base
format c_base diffGDP diffPlus1 %10.2f
outsheet year c_base diffGDP diffPlus1 if year >= 2010 using voucher_c_cost.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year, ACA, GDP, GDP Plus 1\%}]{voucher_c_cost}{voucher_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Voucher Programs Less Net Present Value of ACA (Billions 2010 Dollars)}
\DTLdisplaydb{voucher_c_cost}
\label{tab:voucher_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/voucher.gdp/summary
gen gdp_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen gdp_a_mean = mcare_pta - mcare_pta_premium_paid
gen gdp_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen gdp_d_mean = mcare_ptd
gen gdp_a_enrl = ttl_mcare_pta_enroll * 1e6
gen gdp_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen gdp_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year base_* gdp_*
sort year
merge year using ../output/voucher.gdp.plus1/summary
gen plus_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen plus_a_mean = mcare_pta - mcare_pta_premium_paid
gen plus_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen plus_d_mean = mcare_ptd
gen plus_a_enrl = ttl_mcare_pta_enroll * 1e6
gen plus_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen plus_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* gdp_* plus_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen baseline = base_ttl / 1e9
gen gdpCost = gdp_ttl / 1e9
gen plusCost = plus_ttl / 1e9
gen enrollment_gdp = (base_ttl_a * (gdp_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (gdp_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (gdp_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_gdp = (base_a_enrl * (gdp_a_mean - base_a_mean) + base_b_enrl * (gdp_b_mean - base_b_mean) + base_d_enrl * (gdp_d_mean - base_d_mean)) / 1e9
gen enrollment_plus = (base_ttl_a * (plus_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (plus_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (plus_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_plus = (base_a_enrl * (plus_a_mean - base_a_mean) + base_b_enrl * (plus_b_mean - base_b_mean) + base_d_enrl * (plus_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace gdpCost = gdpCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace plusCost = plusCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_gdp = enrollment_gdp * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_gdp = outlay_gdp * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_plus = enrollment_plus * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_plus = outlay_plus * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cBaseline = sum(baseline) if year >= 2010
gen cGDP = sum(gdpCost) if year >= 2010
gen cPlus = sum(plusCost) if year >= 2010
gen cummEnrollmentGDP = sum(enrollment_gdp) if year >= 2010
gen cummOutlayGDP = sum(outlay_gdp) if year >= 2010
gen cummEnrollmentPlusal = sum(enrollment_plus) if year >= 2010
gen cummOutlayPlusal = sum(outlay_plus) if year >= 2010
global vlist cBaseline cGDP cummEnrollmentGDP cummOutlayGDP cPlus cummEnrollmentPlusal cummOutlayPlusal 
format $vlist %10.2f
outsheet year $vlist using voucher_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{GDP\\Cost},\shortstack{GDP\\Enrollment\\Deviation}, \shortstack{GDP\\Enrollee\\Outlay\\Deviation}, \shortstack{Plus 1\\Cost},\shortstack{Plus 1\\Enrollment\\Deviation}, \shortstack{Plus 1\\Enrollee\\Outlay\\Deviation}}]{voucher_deviation}{voucher_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Voucher Scenario Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{voucher_deviation}
\label{tab:voucher_deviation}
\end{table}

\section{Increasing Eligibility Age Reduces Total Cost and Enrollment}

Figure~\ref{fig:age_enroll} shows the number of people eligible for Medicare under three different age-based criteria.
Figure~\ref{fig:age_cost} shows that this decline in enrollment produces the expected decline in total cost, but Figure~\ref{fig:age_average} shows that the average cost per enrollee is higher, due to the elderly requiring more medical care per person.

\begin{Statacode}{!echo,hide,fig,savefig,label=age_enroll}
use ../output/ACA/summary, clear
rename pop_medicare base
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
rename pop_medicare age67
label variable age67 "Eligibility Age 67"
sort year
keep year base age67
merge year using ../output/elig.age70/summary
rename pop_medicare age70
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Millions Eligible for Medicare") legend(cols(1))
outsheet year base age67 age70 using age_enroll.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_enroll}
\caption{Medicare-eligigble population after raising eligibility age}
\label{fig:age_enroll}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=age_cost}
use ../output/ACA/summary, clear
gen base = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
gen age67 = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable age67 "Eligibility Age 67"
keep base age67 year
sort year
merge year using ../output/elig.age70/summary
gen age70 = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Total Medicare Cost (Billions of 2010 Dollars)") legend(cols(1))
outsheet year base age67 age70 using age_cost.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_cost}
\caption{Total Medicare Costs Fall when Eligibility Rises}
\label{fig:age_cost}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=age_c_cost}
tsset year
tsfill
foreach v in base age67 age70 {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace age67 = age67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace age70 = age70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen c_base = sum(base) if year >= 2010
gen c_age67 = sum(age67) if year >= 2010
gen c_age70 = sum(age70) if year >= 2010
gen diffAge67 = c_age67 - c_base
gen diffAge70 = c_age70 - c_base
global vlist c_base diffAge67 diffAge70
format $vlist %10.2f
outsheet year $vlist if year >= 2010 using age_c_cost.csv, replace comma
\end{Statacode}
\DTLloaddb[headers={Year, ACA, Eligibility Age 67, Eligibility Age 70}]{age_c_cost}{age_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Eligibility Age Options Less Net Present Value of ACA (Billions of 2010 Dollars)}
\DTLdisplaydb{age_c_cost}
\label{tab:age_c_cost}
\end{table}

\begin{Statacode}{!echo,hide,fig,savefig,label=age_average}
use ../output/ACA/summary, clear
gen base = ttl_mcare/pop_medicare
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
gen age67 = ttl_mcare/pop_medicare
label variable age67 "Eligibility Age 67"
keep base age67 year
sort year
merge year using ../output/elig.age70/summary
gen age70 = ttl_mcare/pop_medicare
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Medicare Cost per Enrollee (2010 Dollars)") legend(cols(1))
outsheet year base age67 age70 using age_average.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_average}
\caption{Average Medicare Costs Rise when Eligibility Age Rises}
\label{fig:age_average}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/elig.age67/summary
gen a67_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen a67_a_mean = mcare_pta - mcare_pta_premium_paid
gen a67_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen a67_d_mean = mcare_ptd
gen a67_a_enrl = ttl_mcare_pta_enroll * 1e6
gen a67_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen a67_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year base_* a67_*
sort year
merge year using ../output/elig.age70/summary
gen a70_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen a70_a_mean = mcare_pta - mcare_pta_premium_paid
gen a70_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen a70_d_mean = mcare_ptd
gen a70_a_enrl = ttl_mcare_pta_enroll * 1e6
gen a70_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen a70_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* a67_* a70_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
foreach v in "a_enrl" "b_enrl" "d_enrl" "a_mean" "b_mean" "d_mean" {
  replace a67_`v' = base_`v' if year < 2014
  replace a70_`v' = base_`v' if year < 2014
} 
gen baseline = base_ttl / 1e9
gen a67Cost = a67_ttl / 1e9
gen a70Cost = a70_ttl / 1e9
gen enrollment_a67 = (base_ttl_a * (a67_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (a67_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (a67_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_a67 = (base_a_enrl * (a67_a_mean - base_a_mean) + base_b_enrl * (a67_b_mean - base_b_mean) + base_d_enrl * (a67_d_mean - base_d_mean)) / 1e9
gen enrollment_a70 = (base_ttl_a * (a70_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (a70_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (a70_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_a70 = (base_a_enrl * (a70_a_mean - base_a_mean) + base_b_enrl * (a70_b_mean - base_b_mean) + base_d_enrl * (a70_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace a67Cost = a67Cost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace a70Cost = a70Cost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_a67 = enrollment_a67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_a67 = outlay_a67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_a70 = enrollment_a70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_a70 = outlay_a70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cBaseline = sum(baseline) if year >= 2010
gen cA67 = sum(a67Cost) if year >= 2010
gen cA70 = sum(a70Cost) if year >= 2010
gen cummEnrollmentA67 = sum(enrollment_a67) if year >= 2010
gen cummOutlayA67 = sum(outlay_a67) if year >= 2010
gen cummEnrollmentA70 = sum(enrollment_a70) if year >= 2010
gen cummOutlayA70 = sum(outlay_a70) if year >= 2010
global vlist cBaseline cA67 cummEnrollmentA67 cummOutlayA67 cA70 cummEnrollmentA70 cummOutlayA70
format $vlist %10.2f
outsheet year $vlist using age_deviation.csv if year >= 2010, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{Age 67\\Cost},\shortstack{Age 67\\Enrollment\\Deviation},\shortstack{Age 67\\Outlay\\Deviation},\shortstack{Age 70\\Cost},\shortstack{Age 70\\Enrollment\\Deviation},\shortstack{Age 70\\Outlay\\Deviation}}]{age_deviation}{age_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{age_deviation}
\label{tab:age_deviation}
\end{table}

\begin{Statacode}{!echo,hide}
clear all
\end{Statacode}

\end{document}

### 2019-11-22 analysis/psid_techdoc/psid_techdoc.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{lipsum}

\title{The Future America Model: Technical Documentation }
\author{Bryan Tysinger
\and
Duncan Leaf}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\listoffigures
\listoftables

\newpage

\section{Functioning of the Dynamic Model}
\subsection{Background}
History of the FEM, funding sources.

\subsection{Overview}
The defining characteristic of the model is the modeling of real rather than synthetic cohorts, all of whom 
are followed at the individual level. This allows for more heterogeneity in behavior than would be allowed 
by a cell-based approach. Also, since the PSID interviews both respondent and spouse, we can link records to 
calculate household-level outcomes, which depend on the outcomes of both spouses.

The model has three core components:

\begin{itemize}
\item The initial cohort module predicts the economic and health outcomes of new cohorts of 25/26 year-olds. This 
module takes in data from the Panel Survey of Income Dynamics (PSID) and trends calculated from other sources. It 
allows us to “generate” cohorts as the simulation proceeds, so that we can measure outcomes for the age 25+ 
population in any given year. 
\item The transition module calculates the probabilities of transiting across various health states and financial 
outcomes. The module takes as inputs risk factors such as smoking, weight, age and education, along with lagged 
health and financial states. This allows for a great deal of heterogeneity and fairly general feedback effects. 
The transition probabilities are estimated from the longitudinal data in the PSID. 
\item The policy outcomes module aggregates projections of individual-level outcomes into policy outcomes such as 
taxes, medical care costs, and disability benefits. This component takes account of public 
and private program rules to the extent allowed by the available outcomes.  
\end{itemize}

Figure 1 provides a schematic overview of the model. We start in 2009 with an initial population aged 25+ taken 
from the PSID. We then predict outcomes using our estimated transition probabilities (see section \ref{transition}). Those who 
survive make it to the end of that year, at which point we calculate policy outcomes for the year. We then move 
to the following time period (two years later), when a new cohort of 25 and 26 year-olds enters (see section \ref{newcohorts}). 
This entrance forms the new age 25+ population, which then proceeds through the transition model as before. This 
process is repeated until we reach the final year of the simulation.

\subsection{Comparison with other prominent microsimulation models}
The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
based on assumptions about health trends rather than historical time series.

FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
and childbearing are simulated.  Labor force participation is modeled in greater detail, as is education.

\section{Data Sources for Estimation}
The Panel Survey of Income Dynamics is the main data source for the model.
\subsection{Panel Survey of Income Dynamics}
The 1999-2011 waves of the Panel Survey of Income Dynamics (PSID) are used to estimate the transition models.  PSID interviews
occur every two years.  We use all individuals twenty-five or older who have formed their own households in the analysis, all heads, wives, and
"wives" in PSID terms.  When appropriately weighted, this PSID is representative of the U.S. population 25+ and older.

\subsection{Medical Expenditure Panel Survey}
The MEPS, beginning in 1996, is a set of large-scale surveys of families and individuals, their medical providers (doctors, 
hospitals, pharmacies, etc.), and employers across the United States. The Household Component (HC) of the MEPS provides 
data from individual households and their members, which is supplemented by data from their medical providers. The Household 
Component collects data from a representative sub sample of households drawn from the previous year's National Health Interview 
Survey (NHIS). Since NHIS does not include the institutionalized population, neither does MEPS: this implies that we can only 
use the MEPS to estimate medical costs for the non-elderly population. Information collected during household interviews 
include: demographic characteristics, health conditions, health status, use of medical services, sources of medical payments, 
and body weight and height. Each year the household survey includes approximately 12,000 households or 34,000 individuals. 
Sample size for those aged 25-64 is about ???.  MEPS has comparable measures of social-economic (SES) variables as those in 
PSID, including age, race/ethnicity, educational level, census region, and marital status.

\subsection{Medicare Current Beneficiary Survey}
The MCBS is a nationally representative sample of aged, disabled and institutionalized Medicare beneficiaries.  The MCBS 
attempts to interview each respondent twelve times over three years, regardless of whether he or she resides in the community, 
a facility, or transitions between community and facility settings. The disabled (under 65 years of age) and oldest-old (85 
years of age or older) are over-sampled. The first round of interviewing was conducted in 1991. Originally, the survey was a 
longitudinal sample with periodic supplements and indefinite periods of participation. In 1994, the MCBS switched to a rotating 
panel design with limited periods of participation. Each fall a new panel is introduced, with a target sample size of 12,000 
respondents and each summer a panel is retired. Institutionalized respondents are interviewed by proxy.  The MCBS contains 
comprehensive self-reported information on the health status, health care use and expenditures, health insurance coverage, 
and socioeconomic and demographic characteristics of the entire spectrum of Medicare beneficiaries.  Medicare claims data for 
beneficiaries enrolled in fee-for-service plans are also used to provide more accurate information on health care use and 
expenditures.

We estimate cost and utilization in the "part D era," using data from 2006 to 2010.

\subsection{Health and Retirement Study}
The Health and Retirement Study (HRS), waves 1998-2010 are pooled with the PSID population to estimate the mortality models. Interviews 
in the HRS occur every two years.  We use the dataset created by RAND (RAND HRS, version M) for the analysis. When appropriately weighted, 
the HRS is representative of U.S. households where at least one member is at least 51.

\section{Data Sources for Trends and Baseline Scenario}
\subsection{Data for Trends in Replenishing Cohorts}
\subsubsection{American Community Survey}
\subsubsection{National Health and Nutrition Examination Survey}

\subsection{Data for Other Projections}
\subsubsection{Wage growth, medical cost growth}
\subsubsection{Mortality reduction}

\subsection{Demographic Adjustments}
\subsubsection{Census for adjusting new cohort weights and for simulating immigration}

\section{Estimation}
\subsection{Transition Models}
\label{transition}

\section{Model for New Cohorts}
\label{newcohorts}
\subsection{Information available and empirical strategy}
\subsection{Model and Estimation}

\section{Government Revenue and Expenditures}
\subsection{Medical Costs Estimation}
\subsection{Approach to SS, DI, SSI}
\subsection{Taxes}

\section{Implementation of the Future America Model}
\subsection{Description of the Simulation}
\subsection{Assumptions}
\subsubsection{Wage Growth}
\subsubsection{Medical Cost Growth}
\subsubsection{Mortality Reduction}
\subsection{Adjustments to calibrate health costs (NHEA)}

\section{Internal Validation}
\subsection{Cross-validation}

\section{External Validation}
\subsection{Population Projections}
\subsection{Health Forecasts}
\subsection{Medical Costs and Participation}
\subsubsection{Medicare A/B/D and Medicaid enrollment and spending}
\subsubsection{Private health insurance enrollment}
\subsection{Fertility and Marriage Rates}

\end{document}

### 2019-11-22 analysis/techdoc/latex/FAM/FAM_tables.tex
First commit in git, migrated from a subversion repository

\section{Tables}

% FAM initial conditions are:
%Economic: 	Work Status, 	Earnings, 				Wealth
%Health: 		BMI Category, Smoking Category, Hypertension 
%Other: 		Education, 		Partnered, 				Partner Type, Health Insurance

\begin{table}[h]\centering
\begin{tabular}{l l l}
\textbf{Economic Outcomes} & \textbf{Health Outcomes} & \textbf{Other Outcomes}\\
Work Status & BMI Category & Education\\
Earnings & Smoking Category & Partnered\\
Wealth & Hypertension & Partner Type\\
& & Health Insurance
\end{tabular}~\\~\\
\caption{Estimated outcomes in replenishing cohorts module}
\label{tab:initial_conditions_model}
\end{table}

% FAM transition models are:
%Economic: Social Security Claiming, Disability Claiming, Supplemental Security Income Claiming, Non-Zero Capital Income, Capital Income, Non-zero Government Transfers, Government Transfers, Non-zero Wealth, Wealth, Laborforce Status, Full- or Part-time, Any Earnings if Unemployed, Any Earnings if Not in Labor Force, Earnings if Full-time, Earnings if Part-time, Earnings if Unemployed, Earnings if Not in Labor Force
%Health: Mortality, Heart Disease, Stroke, Cancer, Hypertension, Diabetes, Lung Disease, Start Smoking, Stop Smoking, ADL Status, IADL Status, Births/Paternity, Self-reported Health, BMI, Partner Death
%Marital Status (by sex): Exit Single, Exit Cohabitation, Exit Married, Single to Married, Cohabitation to Married, Married to Cohabitation
%Other: Insurance Type

\begin{sidewaystable}
\centering
\begin{tabular}{l l l l}
\textbf{Economic Outcomes} & \textbf{Health Outcomes} &  \textbf{Marital Status} & \textbf{Other Outcomes}\\
Social Security Claiming & Mortality & Exit Single & Insurance Type\\
Disability Claiming & Heart Disease & Exit Cohabitation &\\
Non-Zero Capital Income & Cancer & Exit Married & \\
Capital Income (if non-zero) & Hypertension & Single to Married &\\
Non-zero Government Transfers & Diabetes & Cohabitation to Married & \\
Government Transfers (if non-zero) & Lung Disease & Married to Cohabitation & \\
Non-zero Wealth & Start Smoking & & \\
Wealth (if non-zero) & Stop Smoking & &\\
Labor Force Status (out, unemployed, working) & ADL Status & &\\
Employed Full- or Part-time (if working) & IADL Status & &\\
Any Earnings (if Unemployed) & Births/Paternity & &\\
Any Earnings (if Not in Labor Force) & Self-reported Health & &\\
Earnings (if Full-time) & BMI & &\\
Earnings (if Part-time) & Partner Death & &\\
Earnings (if Unemployed and any) & & &\\
Earnings (if Not in Labor Force and any) & & &\\
\end{tabular}
\caption{Estimated outcomes in transitions module}
\label{tab:transition_model}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\begin{tabular}[h]{l rrrrrrrr}
\input{../tables/FAM/svy_disease_prevalence.tex}
\end{tabular}
\caption{Health condition prevalences in survey data}
\label{tab:svy_disease_prevalence}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.75}{
\begin{tabular}{l p{2.5in}p{2.5in}p{2.5in}p{2.5in}}
& \multicolumn{4}{c}{Survey}\\
Disease & \multicolumn{1}{c}{PSID/HRS} & \multicolumn{1}{c}{NHIS} & \multicolumn{1}{c}{MEPS} & \multicolumn{1}{c}{MCBS} \\
\hline
Cancer &
Has a doctor ever told you that you have cancer or a malignant tumor, excluding minor skin cancers?&
Have you ever been told be a doctor or other health professional that you had … cancer or a malignanacy of any kind? (WHEN RECODED, SKIN CANCERS WERE EXCLUDED)&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 11-21, 24-45&
Has a doctor ever told you that you had any (other) kind of cancer malignancy, or tumor other than skin cancer?\\
\hline
Heart Diseases &
Has a doctor ever told you that you had a heart attack, coronary heart disease, angina, congestive heart failure, or other heart problems?&
Four separate questions were asked about whether ever told by a docotor or oither health professional that had: CHD, Angina, MI, other heart problems.&
Have you ever been told by a doctor or health professional that you have … CHD; Angina; MI; other heart problems&
Six separate questions were asked about whether ever told by a doctor that had: Angina or MI; CHD; other heart problems (included four questions)\\
\hline
Stroke &
Has a doctor ever told you that you had a stroke?&
Have you EVER been told by a doctor or other health professional that you had a stroke?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have a stroke or TIA (transient ischemic attack)&
[Since (PREV$<$ SUPP. RD. INT. DATE),] has a doctor (ever) told (you/SP) that (you/he/she) had a stroke, a brain hemorrhage, or a cerebrovascular accident?\\
\hline
Diabetes&
Has a doctor ever told you that you have diabetes or high blood sugar?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
If Female, add: [Other than during pregnancy,] Have you hever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
Has a doctor (ever) told (you/SP) that (you/he/she) had diabetes, high blood sugar, or sugar in (your/his/her) urine? [DO NOT INCLUDE BOERDERLINE PREGNANCY, OR PRE-DIABETEIC DIABETES.]\\
\hline
Hypertension&
Has a doctor ver told you that you have high blood pressure or hypertension?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Has a doctor (ever) told (you/SP) that (you/he/she) (still) (had) (have/has) hypertension, sometimes called high blood pressure?\\
\hline
Lung Disease&
Has a doctor ever told you that you have chronic lung disease such a schronic bronchitis or emphysema? [IWER: DO NOT INCLUDE ASTHMA]&
Question 1: During the PAST 12 MONTHS, have you ever been told by a doctor or other health professional that you had chronic bronchitis? Question 2: Have you EVER been told by a docotor or other health professional that you had … emphysema?&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 127, 129-312&
Has a doctor (ever) told (you/SP) that (you/he/she) had emphysema, asthma, or COPD? [COPD=CHRONIC OBSTRUCTIVE PULMONARY DISEASE.]\\
\hline
Overweight & \multicolumn{4}{c}{\multirow{2}{*}{Self-reported body weight and height}}\\
\hhline{-~~~~}
Obese & \multicolumn{4}{c}{} \\
\hline
\end{tabular}
}
\caption{Survey questions used to determine health conditions}
\label{tab:svy_disease_questions}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Dynamically fill in numeric results using Sweave
\centering
\scalebox{0.9}{
\begin{tabular}{lllccr}
 &  &  & Type & At risk & Mean/fraction\\
\hline
\input{../tables/FAM/transitioned_outcomes_disease.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_smoking.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_logbmi.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_adlstat.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_iadlstat.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_work.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_lfpben.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_mstat.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_birth.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_financial.tex}
\hline
\end{tabular}
}
\caption[Outcomes in the transition model]{Outcomes in the transition model. Estimation sample is PSID 1999-2013 waves.}
\label{tab:transitioned_outcomes}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Write a script that dynamically generates this table from the .est files
\centering
\scalebox{0.53}{
\begin{tabular}{l*{20}{c}}
Value at time $T-1$ & \multicolumn{20}{c}{Outcome at time $T$}  \\
\hline
 & Heart  &  & & Lung  & & &  &  & Smoking  &  &  & &  &  & & Nursing & & &  Nonzero & \\
 & disease & hypertension & stroke & disease & diabetes & cancer & disability & mortality & status & BMI & Any HI & DI Claim & SS Claim & DB Claim & SSI Claim & Home & Work & Earnings & Wealth & Wealth\\
Heart disease &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Blood pressure & \checkmark &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Stroke &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Lung disease &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Diabetes & \checkmark & \checkmark & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Cancer &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Disability &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DI &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SS &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DB &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SSI &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  &  &  & \\
Work &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Earnings &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Nonzero wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Nursing home stay &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark &  &  & \checkmark & \checkmark
\end{tabular}
}
\caption[Restrictions on transition model]{Restrictions on transition model. \checkmark indicates that an outcome at time $T-1$ is allowed in the transition model for an outcome at time $T$.}
\label{tab:transition_restrictions}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{1}{
\input{../tables/FAM/controlvar_descs.tex}
}
\caption[Descriptive statistics for stock population]{Desciptive statistics for variables in 2009 PSID ages 25+ sample used as simulation stock population}
\label{tab:controlvar_descs}
\end{table}

%\begin{table}
%% \todo Dynamically fill in numeric results using Sweave and reduce the number of sig. digits after decimal
%\centering
%\scalebox{0.9}{
%\begin{tabular}{ccccrr}
% &  &  & Selection & 1992 & 2004 \\
%\hline
%\input{../tables/FAM/desc_init_conditions_binary.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_bmistat.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_smoking.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_funcstat.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_continuous.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_censcont.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_censdiscrete.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_earlyagedb.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_normalagedb.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_covars.tex}
%\hline
%\end{tabular}
%}
%\caption{Initial conditions used for estimation (1992) and simulation (2004)}
%\label{tab:desc_init_conditions}
%\end{table}

\begin{sidewaystable}
\centering
\scalebox{1}{
\input{../tables/FAM/latent_model_mean_est.tex}
}
\caption[Parameter estimates for latent model: conditional means and thresholds]{Parameter estimates for latent model: conditional means and thresholds. Sample is respondents age 25-30 in 2005-2011 PSID waves}
\label{tab:latent_model_mean_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{1}{
\input{../tables/FAM/latent_model_vcmat_est.tex}
}
\caption[Parameter estimates for latent model: parameterized covariance matrix]{Parameter estimates for latent model: parameterized covariance matrix. Sample is respondents age 25-30 in 2005-2011 PSID waves}
\label{tab:latent_model_vcmat_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_health.tex}
\caption{Health and risk factor trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_health}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_educ.tex}
\caption{Education trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_educ}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_social.tex}
\caption{Social trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_social}
\end{sidewaystable}

%\begin{table}
%\centering
%\input{../tables/FAM/NHEA_adjustment.tex}
%\caption{Per capita medical spending by payment source, age group, and year}
%\label{tab:NHEA_adjustment}
%\end{table}

%\begin{table}
%\centering
%\scalebox{0.8}{
%\input{../tables/FAM/time_series_by_year.tex}
%}
%\caption{Assumptions for each calendar year}
%\label{tab:time_series_by_year}
%\end{table}

%\begin{table}
%\centering
%\scalebox{0.8}{
%\input{../tables/FAM/time_series_by_yob.tex}
%}
%\caption[Assumptions for each birth year]{Assumptions for each birth year. In years after 1960, all values are held constant at their 1960 levels.}
%\label{tab:time_series_by_yob}
%\end{table}

\begin{table}
\centering
\input{../tables/FAM/crossval_unweighted.tex}
\caption{Crossvalidation of 1999 cohort: Mortality in 2001, 2007, and 2013}
\label{tab:crossval_unweighted}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_demog.tex}
\caption{Crossvalidation of 1999 cohort: Demographic outcomes in 2001, 2007, and 2013}
\label{tab:crossval_demog}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_binhlth.tex}
\caption{Crossvalidation of 1999 cohort: Binary health outcomes in 2001, 2007, and 2013}
\label{tab:crossval_binhlth}
\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_risk.tex}
\caption{Crossvalidation of 1999 cohort: Risk factor outcomes in 2001, 2007, and 2013}
\label{tab:crossval_risk}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_binecon.tex}
\caption{Crossvalidation of 1999 cohort: Binary economic outcomes in 2001, 2007, and 2013}
\label{tab:crossval_binecon}
\end{table}

%\begin{table}
%% \todo clean up the significant digits in this table
%% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
%\centering
%\input{../tables/FAM/crossval_cntecon.tex}
%\caption{Crossvalidation of 1999 cohort: Simulated vs reported continuous economic outcomes in 2001, 2007, and 2013}
%\label{tab:crossval_cntecon}
%\end{table}

\begin{sidewaystable}
\centering
\input{../tables/FAM/external_pop.tex}
\caption{Population forecasts: Census compared to FAM}
\label{tab:external_pop}
\end{sidewaystable}

### 2019-11-22 analysis/techdoc/latex/FAM/acknowledgments.tex
First commit in git, migrated from a subversion repository

\section{Acknowledgments}

The Future Elderly Model and Future Americans Model has been developed by a large team over the last decade.  Jay Bhattacharya, Eileen Crimmins, Christine Eibner, \'Etienne Gaudette, Geoff Joyce, Darius Lakdawalla,
Pierre-Carl Michaud, and Julie Zissimopoulos have all provided expert guidance.  Adam Gailey, Baoping Shang, and Igor Vaynman provided programming and analytic 
support during the first years of FEM development at RAND.  Jeff Sullivan then led the technical development for several years.  More recently, the University of 
Southern California research programming team has supported model development, including FAM development.  These programmers include Patricia St. Clair, Laura Gascue, Henu Zhao, and Yuhui Zheng. 
Barbara Blaylock, Malgorzata Switek, and Wendy Cheng have greatly aided model development while working as research assistants at USC.

### 2019-11-22 analysis/techdoc/latex/FAM/baseline_forecasts.tex
First commit in git, migrated from a subversion repository

\section{Baseline Forecasts}
In this section we present baseline forecasts of the Future Americans Model.  The figures show data from the PSID for the 25+ population from 1999 through 2009 and 
forecasts from the FAM for the 25+ population beginning in 2009.

\subsection{Disease Prevalence}
Figure \ref{fig:chronic_diseases_male} depicts the six chronic conditions we project for men.  And Figure \ref{fig:chronic_diseases_female} depicts the historic 
and forecasted values for women.

\begin{figure}[ht!]
\centering

\includegraphics[scale=1.0]{./img/chronic_diseases_male}
\caption{Historic and Forecasted Chronic Disease Prevalence for Men 25+}
\label{fig:chronic_diseases_male}

\includegraphics[scale=1.0]{./img/chronic_diseases_female}
\caption{Historic and Forecasted Chronic Disease Prevalence for Women 25+}
\label{fig:chronic_diseases_female} 
\end{figure}

Figure \ref{fig:adl_iadl_male} shows historic and forecasted levels for any ADL difficulties, three or more ADL difficulties, any IADL difficulties, and
two or more IADL difficulties for men 25 and older. Figure \ref{fig:adl_iadl_female} shows historic and forecasted levels for any ADL difficulties, three or 
more ADL difficulties, any IADL difficulties, and two or more IADL difficulties for women 25 and older.

\begin{figure}[ht!]
\centering

\includegraphics[scale=1.0]{./img/adl_iadl_male}
\caption{Historic and Forecasted ADL and IADL Prevalence for Men 25+}
\label{fig:adl_iadl_male}

\includegraphics[scale=1.0]{./img/adl_iadl_female}
\caption{Historic and Forecasted ADL and IADL Prevalence for Women 25+}
\label{fig:adl_iadl_female} 
\end{figure}

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation.tex
First commit in git, migrated from a subversion repository

\section{Data sources used for estimation}
The Panel Survey of Income Dynamics is the main data source for the model. We estimate models
for assigning characteristics for the replacement cohorts in Replenishing Conditions Module.  These
are summarized in Table \ref{tab:initial_conditions_model}.  We estimate transition models for 
the entire PSID population in the Transition Model Module.  Transitioned outcomes are described in
Table \ref{tab:transition_model}.

\input{data_sources_estimation_psid.tex}
\input{data_sources_estimation_hrs.tex}
%\input{data_sources_estimation_ssearnings.tex}
%\input{data_sources_estimation_nhis.tex}
\input{data_sources_estimation_meps.tex}
\input{data_sources_estimation_mcbs.tex}

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_hrs.tex
First commit in git, migrated from a subversion repository

\subsection{Health and Retirement Study}
The Health and Retirement Study (HRS), waves 1998-2012 are pooled with the PSID for estimation of mortality and 
widowhood models. The HRS has a similar structure to the PSID, with interviews occurring every two years. The HRS data
is harmonized to the PSID for all relevant variables. We use the dataset created by RAND (RAND HRS, version O) as our basis 
for the analysis. We use all cohorts in the analysis. When appropriately weighted, the HRS in 2010 is representative of U.S. households 
where at least one member is at least 51.  Compared to the PSID, the HRS includes more older Hispanics and interviews more respondents 
once they have entered nursing homes.

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_mcbs.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/psid_techdoc/psid_techdoc.tex
@@
-\documentclass{article}
-\usepackage{geometry}
-\usepackage{fancyhdr}
-\usepackage{amsmath,amsthm,amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{lipsum}
-
-\title{The Future America Model: Technical Documentation }
-\author{Bryan Tysinger
-\and
-Duncan Leaf}
-\date{\today}
-
-\begin{document}
-\maketitle
-\tableofcontents
-\listoffigures
-\listoftables
-
-\newpage
-
-\section{Functioning of the Dynamic Model}
-\subsection{Background}
-History of the FEM, funding sources.
-
-\subsection{Overview}
-The defining characteristic of the model is the modeling of real rather than synthetic cohorts, all of whom 
-are followed at the individual level. This allows for more heterogeneity in behavior than would be allowed 
-by a cell-based approach. Also, since the PSID interviews both respondent and spouse, we can link records to 
-calculate household-level outcomes, which depend on the outcomes of both spouses.
-
-The model has three core components:
-
-\begin{itemize}
-\item The initial cohort module predicts the economic and health outcomes of new cohorts of 25/26 year-olds. This 
-module takes in data from the Panel Survey of Income Dynamics (PSID) and trends calculated from other sources. It 
-allows us to “generate” cohorts as the simulation proceeds, so that we can measure outcomes for the age 25+ 
-population in any given year. 
-\item The transition module calculates the probabilities of transiting across various health states and financial 
-outcomes. The module takes as inputs risk factors such as smoking, weight, age and education, along with lagged 
-health and financial states. This allows for a great deal of heterogeneity and fairly general feedback effects. 
-The transition probabilities are estimated from the longitudinal data in the PSID. 
-\item The policy outcomes module aggregates projections of individual-level outcomes into policy outcomes such as 
-taxes, medical care costs, and disability benefits. This component takes account of public 
-and private program rules to the extent allowed by the available outcomes.  
-\end{itemize}
-
-Figure 1 provides a schematic overview of the model. We start in 2009 with an initial population aged 25+ taken 
-from the PSID. We then predict outcomes using our estimated transition probabilities (see section \ref{transition}). Those who 
-survive make it to the end of that year, at which point we calculate policy outcomes for the year. We then move 
-to the following time period (two years later), when a new cohort of 25 and 26 year-olds enters (see section \ref{newcohorts}). 
-This entrance forms the new age 25+ population, which then proceeds through the transition model as before. This 
-process is repeated until we reach the final year of the simulation.
-
-\subsection{Comparison with other prominent microsimulation models}
-The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
-model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
-based on assumptions about health trends rather than historical time series.
-
-FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
-and childbearing are simulated.  Labor force participation is modeled in greater detail, as is education.
-
-\section{Data Sources for Estimation}
-The Panel Survey of Income Dynamics is the main data source for the model.
-\subsection{Panel Survey of Income Dynamics}
-The 1999-2011 waves of the Panel Survey of Income Dynamics (PSID) are used to estimate the transition models.  PSID interviews
-occur every two years.  We use all individuals twenty-five or older who have formed their own households in the analysis, all heads, wives, and
-"wives" in PSID terms.  When appropriately weighted, this PSID is representative of the U.S. population 25+ and older.
-
-\subsection{Medical Expenditure Panel Survey}
-The MEPS, beginning in 1996, is a set of large-scale surveys of families and individuals, their medical providers (doctors, 
-hospitals, pharmacies, etc.), and employers across the United States. The Household Component (HC) of the MEPS provides 
-data from individual households and their members, which is supplemented by data from their medical providers. The Household 
-Component collects data from a representative sub sample of households drawn from the previous year's National Health Interview 
-Survey (NHIS). Since NHIS does not include the institutionalized population, neither does MEPS: this implies that we can only 
-use the MEPS to estimate medical costs for the non-elderly population. Information collected during household interviews 
-include: demographic characteristics, health conditions, health status, use of medical services, sources of medical payments, 
-and body weight and height. Each year the household survey includes approximately 12,000 households or 34,000 individuals. 
-Sample size for those aged 25-64 is about ???.  MEPS has comparable measures of social-economic (SES) variables as those in 
-PSID, including age, race/ethnicity, educational level, census region, and marital status.
-
 \subsection{Medicare Current Beneficiary Survey}
-The MCBS is a nationally representative sample of aged, disabled and institutionalized Medicare beneficiaries.  The MCBS 
-attempts to interview each respondent twelve times over three years, regardless of whether he or she resides in the community, 
-a facility, or transitions between community and facility settings. The disabled (under 65 years of age) and oldest-old (85 
-years of age or older) are over-sampled. The first round of interviewing was conducted in 1991. Originally, the survey was a 
-longitudinal sample with periodic supplements and indefinite periods of participation. In 1994, the MCBS switched to a rotating 
-panel design with limited periods of participation. Each fall a new panel is introduced, with a target sample size of 12,000 
-respondents and each summer a panel is retired. Institutionalized respondents are interviewed by proxy.  The MCBS contains 
-comprehensive self-reported information on the health status, health care use and expenditures, health insurance coverage, 
-and socioeconomic and demographic characteristics of the entire spectrum of Medicare beneficiaries.  Medicare claims data for 
-beneficiaries enrolled in fee-for-service plans are also used to provide more accurate information on health care use and 
-expenditures.
-
-We estimate cost and utilization in the "part D era," using data from 2006 to 2010.
-
-\subsection{Health and Retirement Study}
-The Health and Retirement Study (HRS), waves 1998-2010 are pooled with the PSID population to estimate the mortality models. Interviews 
-in the HRS occur every two years.  We use the dataset created by RAND (RAND HRS, version M) for the analysis. When appropriately weighted, 
-the HRS is representative of U.S. households where at least one member is at least 51.
-
-\section{Data Sources for Trends and Baseline Scenario}
-\subsection{Data for Trends in Replenishing Cohorts}
-\subsubsection{American Community Survey}
-\subsubsection{National Health and Nutrition Examination Survey}
-
-\subsection{Data for Other Projections}
-\subsubsection{Wage growth, medical cost growth}
-\subsubsection{Mortality reduction}
-
-\subsection{Demographic Adjustments}
-\subsubsection{Census for adjusting new cohort weights and for simulating immigration}
-
-\section{Estimation}
-\subsection{Transition Models}
-\label{transition}
-
-\section{Model for New Cohorts}
-\label{newcohorts}
-\subsection{Information available and empirical strategy}
-\subsection{Model and Estimation}
-
-\section{Government Revenue and Expenditures}
-\subsection{Medical Costs Estimation}
-\subsection{Approach to SS, DI, SSI}
-\subsection{Taxes}
-
-\section{Implementation of the Future America Model}
-\subsection{Description of the Simulation}
-\subsection{Assumptions}
-\subsubsection{Wage Growth}
-\subsubsection{Medical Cost Growth}
-\subsubsection{Mortality Reduction}
-\subsection{Adjustments to calibrate health costs (NHEA)}
-
-\section{Internal Validation}
-\subsection{Cross-validation}
-
-\section{External Validation}
-\subsection{Population Projections}
-\subsection{Health Forecasts}
-\subsection{Medical Costs and Participation}
-\subsubsection{Medicare A/B/D and Medicaid enrollment and spending}
-\subsubsection{Private health insurance enrollment}
-\subsection{Fertility and Marriage Rates}
-
-\end{document}
+The Medicare Current Beneficiary Survey (MCBS) is a nationally representative sample of aged, disabled 
+and institutionalized Medicare beneficiaries.  The MCBS attempts to interview each respondent twelve 
+times over three years, regardless of whether he or she resides in the community, a facility, or 
+transitions between community and facility settings. The disabled (under 65 years of age) and 
+oldest-old (85 years of age or older) are over-sampled. The first round of interviewing was conducted 
+in 1991. Originally, the survey was a longitudinal sample with periodic supplements and indefinite 
+periods of participation. In 1994, the MCBS switched to a rotating panel design with limited periods 
+of participation. Each fall a new panel is introduced, with a target sample size of 12,000 respondents 
+and each summer a panel is retired. Institutionalized respondents are interviewed by proxy.  The MCBS 
+contains comprehensive self-reported information on the health status, health care use and 
+expenditures, health insurance coverage, and socioeconomic and demographic characteristics of the 
+entire spectrum of Medicare beneficiaries.  Medicare claims data for beneficiaries enrolled in 
+fee-for-service plans are also used to provide more accurate information on health care use and 
+expenditures.  MCBS years 2007-2010 are used for estimating medical cost and enrollment models.
+See section \ref{sec:govt_revenue_and_expenditures_medcost_estimation} for discussion.

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_meps.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/psid_techdoc/psid_techdoc.tex
@@
-\documentclass{article}
-\usepackage{geometry}
-\usepackage{fancyhdr}
-\usepackage{amsmath,amsthm,amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{lipsum}
+\subsection{Medical Expenditure Panel Survey}
+\label{sec:data_sources_estimation_meps}
+The Medical Expenditure Panel Survey (MEPS), beginning in 1996, is a set of large-scale surveys of 
+families and individuals, their medical providers (doctors, hospitals, pharmacies, etc.), and 
+employers across the United States. The Household Component (HC) of the MEPS provides data from 
+individual households and their members, which is supplemented by data from their medical providers. 
+The Household Component collects data from a representative sub sample of households drawn from the 
+previous year's National Health Interview Survey (NHIS). Since NHIS does not include the 
+institutionalized population, neither does MEPS: this implies that we can only use the MEPS to 
+estimate medical costs for the non-elderly (25-64) population. Information collected during household 
+interviews include: demographic characteristics, health conditions, health status, use of medical 
+services, sources of medical payments, and body weight and height. Each year the household survey 
+includes approximately 12,000 households or 34,000 individuals. Sample size for those aged 25-64 is 
+about 15,800 in each year.  MEPS has comparable measures of social-economic (SES) variables as those in PSID, 
+including age, race/ethnicity, educational level, census region, and marital status.  We estimate expenditures 
+and utilization using 2007-2010 data.
 
-\title{The Future America Model: Technical Documentation }
-\author{Bryan Tysinger
-\and
-Duncan Leaf}
-\date{\today}
+See Section 
+\ref{sec:govt_revenue_and_expenditures_medcost_estimation} for a description.  FAM also
+uses MEPS 2001-2003 data for QALY model estimation.
 
-\begin{document}
-\maketitle
-\tableofcontents
-\listoffigures
-\listoftables
-
-\newpage
-
-\section{Functioning of the Dynamic Model}
-\subsection{Background}
-History of the FEM, funding sources.
-
-\subsection{Overview}
-The defining characteristic of the model is the modeling of real rather than synthetic cohorts, all of whom 
-are followed at the individual level. This allows for more heterogeneity in behavior than would be allowed 
-by a cell-based approach. Also, since the PSID interviews both respondent and spouse, we can link records to 
-calculate household-level outcomes, which depend on the outcomes of both spouses.
-
-The model has three core components:
-
-\begin{itemize}
-\item The initial cohort module predicts the economic and health outcomes of new cohorts of 25/26 year-olds. This 
-module takes in data from the Panel Survey of Income Dynamics (PSID) and trends calculated from other sources. It 
-allows us to “generate” cohorts as the simulation proceeds, so that we can measure outcomes for the age 25+ 
-population in any given year. 
-\item The transition module calculates the probabilities of transiting across various health states and financial 
-outcomes. The module takes as inputs risk factors such as smoking, weight, age and education, along with lagged 
-health and financial states. This allows for a great deal of heterogeneity and fairly general feedback effects. 
-The transition probabilities are estimated from the longitudinal data in the PSID. 
-\item The policy outcomes module aggregates projections of individual-level outcomes into policy outcomes such as 
-taxes, medical care costs, and disability benefits. This component takes account of public 
-and private program rules to the extent allowed by the available outcomes.  
-\end{itemize}
-
-Figure 1 provides a schematic overview of the model. We start in 2009 with an initial population aged 25+ taken 
-from the PSID. We then predict outcomes using our estimated transition probabilities (see section \ref{transition}). Those who 
-survive make it to the end of that year, at which point we calculate policy outcomes for the year. We then move 
-to the following time period (two years later), when a new cohort of 25 and 26 year-olds enters (see section \ref{newcohorts}). 
-This entrance forms the new age 25+ population, which then proceeds through the transition model as before. This 
-process is repeated until we reach the final year of the simulation.
-
-\subsection{Comparison with other prominent microsimulation models}
-The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
-model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
-based on assumptions about health trends rather than historical time series.
-
-FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
-and childbearing are simulated.  Labor force participation is modeled in greater detail, as is education.
-
-\section{Data Sources for Estimation}
-The Panel Survey of Income Dynamics is the main data source for the model.
-\subsection{Panel Survey of Income Dynamics}
-The 1999-2011 waves of the Panel Survey of Income Dynamics (PSID) are used to estimate the transition models.  PSID interviews
-occur every two years.  We use all individuals twenty-five or older who have formed their own households in the analysis, all heads, wives, and
-"wives" in PSID terms.  When appropriately weighted, this PSID is representative of the U.S. population 25+ and older.
-
-\subsection{Medical Expenditure Panel Survey}
-The MEPS, beginning in 1996, is a set of large-scale surveys of families and individuals, their medical providers (doctors, 
-hospitals, pharmacies, etc.), and employers across the United States. The Household Component (HC) of the MEPS provides 
-data from individual households and their members, which is supplemented by data from their medical providers. The Household 
-Component collects data from a representative sub sample of households drawn from the previous year's National Health Interview 
-Survey (NHIS). Since NHIS does not include the institutionalized population, neither does MEPS: this implies that we can only 
-use the MEPS to estimate medical costs for the non-elderly population. Information collected during household interviews 
-include: demographic characteristics, health conditions, health status, use of medical services, sources of medical payments, 
-and body weight and height. Each year the household survey includes approximately 12,000 households or 34,000 individuals. 
-Sample size for those aged 25-64 is about ???.  MEPS has comparable measures of social-economic (SES) variables as those in 
-PSID, including age, race/ethnicity, educational level, census region, and marital status.
-
-\subsection{Medicare Current Beneficiary Survey}
-The MCBS is a nationally representative sample of aged, disabled and institutionalized Medicare beneficiaries.  The MCBS 
-attempts to interview each respondent twelve times over three years, regardless of whether he or she resides in the community, 
-a facility, or transitions between community and facility settings. The disabled (under 65 years of age) and oldest-old (85 
-years of age or older) are over-sampled. The first round of interviewing was conducted in 1991. Originally, the survey was a 
-longitudinal sample with periodic supplements and indefinite periods of participation. In 1994, the MCBS switched to a rotating 
-panel design with limited periods of participation. Each fall a new panel is introduced, with a target sample size of 12,000 
-respondents and each summer a panel is retired. Institutionalized respondents are interviewed by proxy.  The MCBS contains 
-comprehensive self-reported information on the health status, health care use and expenditures, health insurance coverage, 
-and socioeconomic and demographic characteristics of the entire spectrum of Medicare beneficiaries.  Medicare claims data for 
-beneficiaries enrolled in fee-for-service plans are also used to provide more accurate information on health care use and 
-expenditures.
-
-We estimate cost and utilization in the "part D era," using data from 2006 to 2010.
-
-\subsection{Health and Retirement Study}
-The Health and Retirement Study (HRS), waves 1998-2010 are pooled with the PSID population to estimate the mortality models. Interviews 
-in the HRS occur every two years.  We use the dataset created by RAND (RAND HRS, version M) for the analysis. When appropriately weighted, 
-the HRS is representative of U.S. households where at least one member is at least 51.
-
-\section{Data Sources for Trends and Baseline Scenario}
-\subsection{Data for Trends in Replenishing Cohorts}
-\subsubsection{American Community Survey}
-\subsubsection{National Health and Nutrition Examination Survey}
-
-\subsection{Data for Other Projections}
-\subsubsection{Wage growth, medical cost growth}
-\subsubsection{Mortality reduction}
-
-\subsection{Demographic Adjustments}
-\subsubsection{Census for adjusting new cohort weights and for simulating immigration}
-
-\section{Estimation}
-\subsection{Transition Models}
-\label{transition}
-
-\section{Model for New Cohorts}
-\label{newcohorts}
-\subsection{Information available and empirical strategy}
-\subsection{Model and Estimation}
-
-\section{Government Revenue and Expenditures}
-\subsection{Medical Costs Estimation}
-\subsection{Approach to SS, DI, SSI}
-\subsection{Taxes}
-
-\section{Implementation of the Future America Model}
-\subsection{Description of the Simulation}
-\subsection{Assumptions}
-\subsubsection{Wage Growth}
-\subsubsection{Medical Cost Growth}
-\subsubsection{Mortality Reduction}
-\subsection{Adjustments to calibrate health costs (NHEA)}
-
-\section{Internal Validation}
-\subsection{Cross-validation}
-
-\section{External Validation}
-\subsection{Population Projections}
-\subsection{Health Forecasts}
-\subsection{Medical Costs and Participation}
-\subsubsection{Medicare A/B/D and Medicaid enrollment and spending}
-\subsubsection{Private health insurance enrollment}
-\subsection{Fertility and Marriage Rates}
-
-\end{document}
+% This is described in Section  \ref{sec:estimation_qalys_crosswalk_development}.

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_nhis.tex
First commit in git, migrated from a subversion repository

\subsection{National Health Interview Survey}
The National Health Interview Survey (NHIS) contains individual-level data on height, weight, smoking 
status, self-reported chronic conditions, income, education, and demographic variables. It is a 
repeated cross-section done every year for several decades. But the survey design has been 
significantly modified several times. Before year 1997, different subgroups of individuals were asked 
about different sets of chronic conditions, after year 1997, a selected sub-sample of the adults were 
asked a complete set of chronic conditions. The survey questions are quite similar to that in HRS. As 
a result, for projecting the trends of chronic conditions for future 51/52 year-olds, we only use data 
from 1997 to 2010. A review of survey questions is provided in Table \ref{tab:svy_disease_questions}. Information on 
weight and height were asked every year, while information on smoking was asked in selected years before 
year 1997, and has been asked annually since year 1997.

FEM uses NHIS to project prevalence of chronic conditions in future cohorts of 51-52 year olds.  The
method is discussed in Sections \ref{sec:data_sources_trends_and_baseline_entering_cohorts} and \ref{sec:new_cohorts_model_empirical_strategy}.  FEM
also relies on the Medical Expenditure Panel Survey, a subsample of NHIS respondents, for model estimation.
See section \ref{sec:data_sources_estimation_meps} for a description.

### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_psid.tex
First commit in git, migrated from a subversion repository

\subsection{Panel Survey of Income Dynamics}
The Panel Survey of Income Dynamics (PSID), waves 1999-2013 are used to estimate the transition models. 
PSID interviews occur every two years.  We create a dataset of respondents who have formed their own households, either
as single heads of households, cohabitating partners, or married partners.  These heads, wives, and "wives" (males
are automatically assigned head of household status by the PSID if they are in a couple) respond to the richest
set of PSID questions, including the health questions that are critical for our purposes.

We use all respondents age 25 and older.  When appropriately weighted, the PSID is representative of U.S. households.  
We also use the PSID as the host data for full population simulations that begin in 2009.  Respondents age 25 and 26 
are used as the basis for the synthetic cohorts that we generate, used for replenishing the sample in population 
simulations or as the basis of cohort scenarios.

The PSID continually adds new cohorts that are descendents (or new partners/spouses of descendents).  Consequently,
updating the simulation to include more recent data is straightforward.

### 2019-11-22 analysis/techdoc/latex/FAM/dynamic_model_background.tex
First commit in git, migrated from a subversion repository

\subsection{Background}
The Future Elderly Model (FEM) is a microsimulation model originally developed out of an effort to 
examine health and health care costs among the elderly Medicare population (age 65+). A description 
of the previous incarnation of the model can be found in \citet{goldman2004health}. The original work was 
founded by the Centers for Medicare and Medicaid Services and carried out by a team of researchers 
composed of Dana P. Goldman, Paul G. Shekelle, Jayanta Bhattacharya, Michael Hurd, Geoffrey F. Joyce, 
Darius N. Lakdawalla, Dawn H. Matsui, Sydne J. Newberry, Constantijn W. A. Panis and Baoping Shang.

Since then various extensions have been implemented to the original model. The most recent version of the FEM
now projects health outcomes for all Americans aged 51 and older and uses the Health and Retirement 
Study (HRS) as a host dataset rather than the Medicare Current Beneficiary Survey (MCBS).  The work 
has also been extended to include economic outcomes such as earnings, labor force participation and 
pensions. This work was funded by the National Institute on Aging through its support of the RAND 
Roybal Center for Health Policy Simulation (P30AG024968), the Department of Labor through contract 
J-9-P-2-0033, the National Institutes of Aging through the R01 grant ``Integrated Retirement 
Modeling'' (R01AG030824) and the MacArthur Foundation Research Network on an Aging Society.

This document describes the Future Americans Model (FAM), the development of the model to forecast Americans aged 25 
and older.  FAM uses the Panel Survey of Income Dynamics (PSID) as the host dataset.  In addition to modeling
health, health care costs, and economic outcomes, FAM also models life events such as changes in marital status 
and childbearing.  Development of FAM is suported by the National Institutes of Aging through the USC Roybal 
Center for Health Policy Simulation (5P30AG024968-13) and the MacArthur Foundation Research Network on an Aging Society.

% \todo add description of work extending to ages 25+ (FAM) here

### 2019-11-22 analysis/techdoc/latex/FAM/dynamic_model_comparison_with_other_microsims.tex
First commit in git, migrated from a subversion repository

\subsection{Comparison with other microsimulation models of health expenditures}
The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
based on assumptions about health trends rather than historical time series.

FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
and childbearing are simulated.  Labor force participation is modeled in greater detail, distinguishing between out-of-labor force, unemployed, 
working part-time, and working full-time.

\subsubsection{CBOLT Model}
The Congressional Budget Office (CBO) uses time-series techniques to project health expenditure 
growth in the short term and then makes an assumption on long-term growth. They use a long term 
growth of excess costs of 2.3 percentage points starting in 2020 for Medicare. They then assume a 
reduction in excess cost growth in Medicare of 1.5\% through 2083, leaving a rate of 0.9\% in 2083. 
For non-Medicare spending they assume an annual decline of 4.5\%, leading to an excess growth rate in 
2083 of 0.1\%.

\subsubsection{Centers for Medicare and Medicaid Services}
The Centers for Medicare and Medicaid Services (CMS) performs an extrapolation of medical expenditures 
over the first ten years, then computes a general equilibrium model for years 25 through 75 and 
linearly interpolates to identify medical expenditures in years 11 through 24 of their estimation. 
The core assumption they use is that excess growth of health expenditures will be one percentage point 
higher per year for years 25-75 (that is if nominal GDP growth is 4\%, health care expenditure growth 
will be 5\%).

\subsubsection{MINT Model}
Modeling Income in the Near Term (MINT) is a microsimulation model developed by the Urban Institute and others 
for the Social Security Administration to enable policy analysis of proposed changes to Social Security benefits and payroll
taxes \citet{smith2013mint}.  MINT uses the Survey of Income and Program Participation (SIPP) as the base data and simulates a range of outcomes, 
with a focus on those that will impact Social Security.  Recent extensions have included health insurance coverage and
out-of-pocket medical expenditures.  Health enters MINT via self-reported health status and self-reported work limitations.  MINT
simulates marital status and fertility.

### 2019-11-22 analysis/techdoc/latex/FAM/dynamic_model_overview.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/psid_techdoc/psid_techdoc.tex
@@
-\documentclass{article}
-\usepackage{geometry}
-\usepackage{fancyhdr}
-\usepackage{amsmath,amsthm,amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{lipsum}
-
-\title{The Future America Model: Technical Documentation }
-\author{Bryan Tysinger
-\and
-Duncan Leaf}
-\date{\today}
-
-\begin{document}
-\maketitle
-\tableofcontents
-\listoffigures
-\listoftables
-
-\newpage
-
-\section{Functioning of the Dynamic Model}
-\subsection{Background}
-History of the FEM, funding sources.
-
 \subsection{Overview}
@@
 by a cell-based approach. Also, since the PSID interviews both respondent and spouse, we can link records to 
-calculate household-level outcomes, which depend on the outcomes of both spouses.
+calculate household-level outcomes, which depend on the responses of both spouses.
 
@@
 \begin{itemize}
-\item The initial cohort module predicts the economic and health outcomes of new cohorts of 25/26 year-olds. This 
+\item The replenishing cohort module predicts the economic and health outcomes of new cohorts of 25/26 year-olds. This 
 module takes in data from the Panel Survey of Income Dynamics (PSID) and trends calculated from other sources. It 
@@
 
-Figure 1 provides a schematic overview of the model. We start in 2009 with an initial population aged 25+ taken 
-from the PSID. We then predict outcomes using our estimated transition probabilities (see section \ref{transition}). Those who 
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.65]{./img/fam_architecture.png}
+\caption{Architecture of the FAM}
+\label{fig:fam_architecture} 
+\end{figure}
+
+Figure \ref{fig:fam_architecture} provides a schematic overview of the model. In this example, we start in 2014 with an initial population aged 25+ taken 
+from the PSID. We then predict outcomes using our estimated transition probabilities (see section \ref{sec:estimation_transition_model}). Those who 
 survive make it to the end of that year, at which point we calculate policy outcomes for the year. We then move 
-to the following time period (two years later), when a new cohort of 25 and 26 year-olds enters (see section \ref{newcohorts}). 
-This entrance forms the new age 25+ population, which then proceeds through the transition model as before. This 
+to the following time period (two years later), when a replenishing cohort of 25 and 26 year-olds enters (see section 
+\ref{sec:model_for_new_cohorts}). This entrance forms the new age 25+ population, which then proceeds through the transition model as before. This 
 process is repeated until we reach the final year of the simulation.
-
-\subsection{Comparison with other prominent microsimulation models}
-The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
-model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
-based on assumptions about health trends rather than historical time series.
-
-FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
-and childbearing are simulated.  Labor force participation is modeled in greater detail, as is education.
-
-\section{Data Sources for Estimation}
-The Panel Survey of Income Dynamics is the main data source for the model.
-\subsection{Panel Survey of Income Dynamics}
-The 1999-2011 waves of the Panel Survey of Income Dynamics (PSID) are used to estimate the transition models.  PSID interviews
-occur every two years.  We use all individuals twenty-five or older who have formed their own households in the analysis, all heads, wives, and
-"wives" in PSID terms.  When appropriately weighted, this PSID is representative of the U.S. population 25+ and older.
-
-\subsection{Medical Expenditure Panel Survey}
-The MEPS, beginning in 1996, is a set of large-scale surveys of families and individuals, their medical providers (doctors, 
-hospitals, pharmacies, etc.), and employers across the United States. The Household Component (HC) of the MEPS provides 
-data from individual households and their members, which is supplemented by data from their medical providers. The Household 
-Component collects data from a representative sub sample of households drawn from the previous year's National Health Interview 
-Survey (NHIS). Since NHIS does not include the institutionalized population, neither does MEPS: this implies that we can only 
-use the MEPS to estimate medical costs for the non-elderly population. Information collected during household interviews 
-include: demographic characteristics, health conditions, health status, use of medical services, sources of medical payments, 
-and body weight and height. Each year the household survey includes approximately 12,000 households or 34,000 individuals. 
-Sample size for those aged 25-64 is about ???.  MEPS has comparable measures of social-economic (SES) variables as those in 
-PSID, including age, race/ethnicity, educational level, census region, and marital status.
-
-\subsection{Medicare Current Beneficiary Survey}
-The MCBS is a nationally representative sample of aged, disabled and institutionalized Medicare beneficiaries.  The MCBS 
-attempts to interview each respondent twelve times over three years, regardless of whether he or she resides in the community, 
-a facility, or transitions between community and facility settings. The disabled (under 65 years of age) and oldest-old (85 
-years of age or older) are over-sampled. The first round of interviewing was conducted in 1991. Originally, the survey was a 
-longitudinal sample with periodic supplements and indefinite periods of participation. In 1994, the MCBS switched to a rotating 
-panel design with limited periods of participation. Each fall a new panel is introduced, with a target sample size of 12,000 
-respondents and each summer a panel is retired. Institutionalized respondents are interviewed by proxy.  The MCBS contains 
-comprehensive self-reported information on the health status, health care use and expenditures, health insurance coverage, 
-and socioeconomic and demographic characteristics of the entire spectrum of Medicare beneficiaries.  Medicare claims data for 
-beneficiaries enrolled in fee-for-service plans are also used to provide more accurate information on health care use and 
-expenditures.
-
-We estimate cost and utilization in the "part D era," using data from 2006 to 2010.
-
-\subsection{Health and Retirement Study}
-The Health and Retirement Study (HRS), waves 1998-2010 are pooled with the PSID population to estimate the mortality models. Interviews 
-in the HRS occur every two years.  We use the dataset created by RAND (RAND HRS, version M) for the analysis. When appropriately weighted, 
-the HRS is representative of U.S. households where at least one member is at least 51.
-
-\section{Data Sources for Trends and Baseline Scenario}
-\subsection{Data for Trends in Replenishing Cohorts}
-\subsubsection{American Community Survey}
-\subsubsection{National Health and Nutrition Examination Survey}
-
-\subsection{Data for Other Projections}
-\subsubsection{Wage growth, medical cost growth}
-\subsubsection{Mortality reduction}
-
-\subsection{Demographic Adjustments}
-\subsubsection{Census for adjusting new cohort weights and for simulating immigration}
-
-\section{Estimation}
-\subsection{Transition Models}
-\label{transition}
-
-\section{Model for New Cohorts}
-\label{newcohorts}
-\subsection{Information available and empirical strategy}
-\subsection{Model and Estimation}
-
-\section{Government Revenue and Expenditures}
-\subsection{Medical Costs Estimation}
-\subsection{Approach to SS, DI, SSI}
-\subsection{Taxes}
-
-\section{Implementation of the Future America Model}
-\subsection{Description of the Simulation}
-\subsection{Assumptions}
-\subsubsection{Wage Growth}
-\subsubsection{Medical Cost Growth}
-\subsubsection{Mortality Reduction}
-\subsection{Adjustments to calibrate health costs (NHEA)}
-
-\section{Internal Validation}
-\subsection{Cross-validation}
-
-\section{External Validation}
-\subsection{Population Projections}
-\subsection{Health Forecasts}
-\subsection{Medical Costs and Participation}
-\subsubsection{Medicare A/B/D and Medicaid enrollment and spending}
-\subsubsection{Private health insurance enrollment}
-\subsection{Fertility and Marriage Rates}
-
-\end{document}

### 2019-11-22 analysis/techdoc/latex/FAM/estimation_transition_model.tex
First commit in git, migrated from a subversion repository

\subsection{Transition model}
\label{sec:estimation_transition_model} 
We consider a large set of outcomes for which we model transitions. Table 
\ref{tab:transitioned_outcomes} gives the set of outcomes considered for the 
transition model along with descriptive statistics and the population at risk 
when estimating the relationships.

Since we have a stock sample from the age 25+ population, each respondent goes 
through an individual-specific series of intervals. Hence, we have an 
unbalanced panel over the age range starting from 25 years old. Denote by 
$j_{i0}$ the first age at which respondent $i$ is observed and $j_{iT_i}$ the 
last age when he is observed. Hence we observe outcomes at ages 
$j_i = j_{i0},\ldots,j_{iT_i}$.

We first start with discrete outcomes which are absorbing states (e.g. disease 
diagnostic, mortality, benefit claiming). Record as $h_{i,j_i,m}=1$ if the 
individual outcome $m$ has occurred as of age $j_i$. We assume the 
individual-specific component of the hazard can be decomposed in a time 
invariant and variant part. The time invariant part is composed of the effect of 
observed characteristics $x_i$ that are constant over the entire life course and 
initial conditions $h_{i,j_0,-m}$ (outcomes other than the outcome 
$m$) that are determined before the first age in which each individual is observed
%\footnote{Section \ref{sec:model_development_transition_model} explains why the $h_{i,j_0,-m}$ terms are included.}. 
The time-varying part is the effect of previously diagnosed outcomes $h_{i,j_i-1,-m}$, 
on the hazard for $m$.\footnote{With some abuse of notation, $j_i-1$ denotes 
the previous age at which the respondent was observed.}  We assume an index of 
the form  $z_{m,j_i} = x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + h_{i,j_0,-m}\psi_m$. Hence, the 
latent component of the hazard is modeled as 
\begin{equation}
h^*_{i,j_i,m}= x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + h_{i,j_0,-m}\psi_m + a_{m,j_i} + \varepsilon_{i,j_i,m},
\label{eqn:transition_hzd_latent}
\end{equation}
\[
m = 1,\ldots,M_0 \mbox{, } j_i = j_{i0},\ldots,j_{i,T_i} \mbox{, } i=1,\ldots,N
\]
The term $\varepsilon_{i,j_i,m}$ is a time-varying shock specific to age $j_i$. 
We assume that this last shock is normally distributed and uncorrelated across 
diseases.  We approximate $a_{m,j_i}$ with an age spline with knots at ages 35, 45,
55, 65, and 75.
This simplification is made 
for computational reasons since the joint estimation with unrestricted age fixed 
effects for each condition would imply a large number of parameters.  The 
absorbing outcome, conditional on being at risk, is defined as 
\[
h_{i,j_i,m} = \max\{I(h^*_{i,j_i,m} > 0), h_{i,j_i-1,m}\}
%\label{eqn:transition_outcome}
\]
The occurrence of mortality censors observation of other 
outcomes in a current year. 
% Mortality is recorded from exit interviews.

A number of restrictions are placed on the way feedback is allowed in the model.  
Table \ref{tab:transition_restrictions} documents restrictions placed on the 
transition model. We also include a set of other controls. A list of such controls 
is given in Table \ref{tab:controlvar_descs} along with descriptive statistics.

We have five other types of outcomes:
\begin{enumerate}
\item First, we have binary outcomes which are not an absorbing state, such as starting smoking. We specify latent 
indices as in (\ref{eqn:transition_hzd_latent}) for these outcomes as well but where 
the lag dependent outcome also appears as a right-hand side variable. This allows for state-dependence.

\item Second, we have ordered outcomes. These outcomes are also modeled as in (\ref{eqn:transition_hzd_latent}) recognizing the observation rule is 
a function of unknown thresholds $\varsigma_m$. Similarly to binary outcomes, we allow for state-dependence by including the lagged outcome 
on the right-hand side.

\item The third type of outcomes we consider are censored outcomes, such as financial wealth. For wealth, 
there are a non-negligible number of observations with zero and negative wealth. For these, we consider 
two part models where the latent variable is specified as in (\ref{eqn:transition_hzd_latent}) but model 
probabilities only when censoring does not occur. In total, we have $M$ outcomes.

\item The fourth type of outcomes are continuous outcomes modeled with ordinary least squares.  For example,
we model transitions in log(BMI).  We allow for state-dependence by including the lagged outcome on the right-hand side.

\item The final type of models are categorical, but without an ordering.  For example, an individual can transition
to being out of the labor force, unemployed, or working (either full- or part-time).  In situations like this, we utilize
a multinomial logit model, including the lagged outcome on the right-hand side.

\end{enumerate}

The parameters 
$\mathbf{\theta}_1 = \left(\left\{\beta_m, \gamma_m, \psi_m, \varsigma_m\right\}_{m=1}^M, \right)$, 
can be estimated by maximum likelihood. Given the normality distribution assumption on the 
time-varying unobservable, the joint probability of all time-intervals until failure, right-censoring 
or death conditional on the initial conditions $h_{i,j_0,-m}$ is the product of 
normal univariate probabilities. Since these sequences, conditional on initial 
conditions, are also independent across diseases, the joint 
probability over all disease-specific sequences is simply the product of 
those probabilities.

For a given respondent observed from initial age $j_{i0}$ to a last age $j_{T_i}$, the probability of the observed health history is 
(omitting the conditioning on covariates for notational simplicity)
\[
	l^{-0}_i(\mathbf{\theta}; h_{i,j_{i0}}) = \left[\prod_{m=1}^{M-1} \prod_{j=j_{i1}}^{j_{T_i}} P_{ij,m}(\mathbf{\theta})^{(1-h_{ij-1,m})(1-h_{ij,M})} \right] \times \left[\prod_{j=j_{i1}}^{j_{T_i}} P_{ij,M}(\mathbf{\theta}) \right]
\]
We use the ${-0}$ superscript to make explicit the conditioning on $\mathbf{h}_{i,j_{i0}} = (h_{i,j_{i0},0},\ldots,h_{i,j_{i0},M})'$. We have limited information on outcomes prior to this age. 
The likelihood is a product of $M$ terms with the $m$th term containing only 
$(\beta_m, \gamma_m, \psi_m, \varsigma_m)$.  This allows the estimation
to be done seperately for each outcome.

\subsubsection{Further Details on Specific Transition Models}
This section describes the modeling strategy for particular outcomes.

\paragraph{Employment Status}
Ultimately, we wish to simulate if an individual is out of the labor force, unemployed, working part-time, or working full-time at 
time $t$.  We treat the estimation of this as a two-stage process.  In the first stage, we predict if the individual is out of 
the labor force, unemployed, or working for pay using a multinomial logit model.  Then, conditional on working for pay, we estimate if 
the individual is working part- or full-time using a probit model.

\paragraph{Earnings}
We estimate last calendar year earnings models based on the current employment status, controlling for the prior employment status.  
Of particular concern are individuals with no earnings, representing approximately twenty-five percent of the unemployed and 
seventy-eight percent of those out of the labor force.  This group is less than 0.5\% of the full- and part-time populations. We 
use a two-stage process for those out of the labor force and unemployed.  The first stage is a probit that estimates if the 
individual has any earnings.  The second stage is an OLS model of log(earnings) for those with non-zero earnings.  For 
those working full- or part-time, we estimate OLS models of log(earnings).

\paragraph{Relationship Status}
We are interested in three relationship statuses: single, cohabitating, and married.  In each case, we treat the transition
from time $t$ to time $t+1$ as a two-stage process.  In the first stage, we estimate if the individual will remain in their 
current status.  In the second stage, we estimate which of the two other states the individual will transition to, conditional
on leaving their current state.

\paragraph{Childbearing}
We estimate the number of children born in two-years separately for women and men.  We model this using an ordered probit with
three categories: no new births, one birth, and two births.  Based on the PSID data, we found the exclusion of three or more
births in a two-year period to be appropriate.

\subsubsection{Inverse Hyperbolic Sine Transformation}
One problem fitting the wealth distribution is that it has a long right tail and some negative values. We use a 
generalization of the inverse hyperbolic sine transform (IHT) presented in \citet{mackinnon1990transforming}. First denote the variable of 
interest $y$. The hyperbolic sine transform is 
\begin{equation}
y = \sinh(x) = \frac{\exp(x) - \exp(-x)}{2}
\label{eqn:sinh_y}
\end{equation}
The inverse of the hyperbolic sine transform is
\[
x = \sinh^{-1}(y) = h(y) = \log(y + (1+y^2)^{1/2})
\]
Consider the inverse transformation. We can generalize such transformation, first allowing for a 
shape parameter $\theta$,
\begin{equation}
r(y) = h(\theta y)/\theta
\label{eqn:generalized_ihs_shape}
\end{equation}
Such that we can specify the regression model as
\begin{equation}
r(y) = x\beta + \varepsilon, \varepsilon \sim \mathrm{N}(0, \sigma^2)
\label{eqn:ihs_regression_model}
\end{equation}
A further generalization is to introduce a location parameter $\omega$ such that the new 
transformation becomes
\begin{equation}
g(y) = \frac{h(\theta(y+\omega)) - h(\theta\omega)}{\theta h'(\theta \omega)}
\label{eqn:geralized_ihs_loc_scale}
\end{equation}
where $h'(a) = (1+a^2)^{-1/2}$.

We specify (\ref{eqn:ihs_regression_model}) in terms of the transformation $g$. The shape parameters 
can be estimated from the concentrated likelihood for $\theta, \omega$. We can then 
retrieve $\beta, \sigma$ by standard OLS.

Upon estimation, we can simulate 
\[
\tilde{g} = x \hat{\beta} + \sigma \tilde{\eta}
\]
where $\eta$ is a standard normal draw. Given this draw, we can retransform using 
(\ref{eqn:geralized_ihs_loc_scale}) and (\ref{eqn:sinh_y})
\begin{align*}
&h(\theta(y+\omega)) = \theta h'(\theta\omega)\tilde{g} + h(\theta\omega)\\
&\tilde{y} = \frac{\sinh\left[\theta h'(\theta\omega)\tilde{g} + h(\theta\omega)\right]-\theta\omega}{\theta}
\end{align*}

% \todo link to transition_estimates.xls on box -- make link unique to the version of the appendix in make process
The included estimates table (estimates\_FAM.xml) gives parameter estimates for the transition models.

### 2019-11-22 analysis/techdoc/latex/FAM/govt_revenue_and_expenditures_medcost_estimation.tex
First commit in git, migrated from a subversion repository

\subsection{Medical costs estimation}
\label{sec:govt_revenue_and_expenditures_medcost_estimation}
In the FAM, a cost module links a person's current state--demographics, economic status, current health, risk factors, and functional status 
to 4 types of individual medical spending. The FAM models: total medical spending (medical spending from all payment sources), Medicare 
spending\footnote{We estimate annual medical spending paid by specific parts of 
Medicare (Parts A, B, and D) and sum to get the total Medicare expenditures.}, 
Medicaid spending (medical spending paid by Medicaid), and out of pocket spending 
(medical spending by the respondent). These estimates are based on pooled weighted 
least squares regressions of each type of spending on risk factors, self-reported 
conditions, and functional status, with spending inflated to constant dollars 
using the medical component of the consumer price index.  We use the 2007-2010 
Medical Expenditure Panel Survey 
% \todo : Change the text to be something like "roughly X respondents per panel, with a new panel starting entering every Y months and lasting Z months between YYYY and YYYY." That way, it is always correct and all we do is change the years of data we're looking at. 
for these regressions for persons 
not Medicare eligible, and the 2007-2010 Medicare Current Beneficiary Survey 
% \todo : Change the text to be something like "roughly X respondents per panel, with a new panel starting entering every Y months and lasting Z months between YYYY and YYYY." That way, it is always correct and all we do is change the years of data we're looking at. 
for spending for those that are eligible for Medicare. Those eligible for 
Medicare include people eligible due to age (65+) or due to disability status. Comparisons of prevalences and question wording across these different sources are provided in Tables \ref{tab:svy_disease_prevalence} and \ref{tab:svy_disease_questions}, respectively.

In the baseline scenario, this spending estimate can be interpreted as the resources consumed by the individual given the manner in which 
medicine is practiced in the United States during the post-part D era (2006-2010). 
Models are estimated for total, Medicaid, out of pocket spending, and for the Medicare spending.

Since Medicare spending has numerous components (Parts A and B are considered here), models are needed to predict enrollment. In 2004, 
98.4\% of all Medicare enrollees, and 99\%+ of aged enrollees, were in Medicare Part A, and thus we assume that all persons eligible for 
Medicare take Part A. We use the 2007-2010 MCBS to model take up of Medicare Part B for both new enrollees into Medicare, as well as 
current enrollees without Part B. Estimates are based on weighted probit regression on various risk factors, demographic, and economic 
conditions. The PSID starting population for the FAM does not contain information on Medicare enrollment. Therefore another model of Part B 
enrollment for all persons eligible for Medicare is estimated via a probit, and used in the first year of simulation to assign initial Part 
B enrollment status. Estimation results are shown in estimates table. The MCBS data over represents the portion enrolled in Part B, having a 97\% 
enrollment rate in 2004 instead of the 93.5\% rate given by Medicare Trustee's Report. In addition to this baseline enrollment probit, we 
apply an elasticity to premiums of -0.10, based on the literature and simulation 
calibration for actual uptake through 2009 \citep{atherly2004effect,buchmueller2006price}. 
The premiums are computed using average Part B costs from the previous time step 
and the means-testing thresholds established by the ACA.

%Since both the MEPS and MCBS are known to under-predict medical spending (see, e.g., \citeauthor{selden2008aligning}, \citeyear{selden2008aligning}, and references therein), we applied adjustment factors to the predicted three types of 
%individual medical spending so that the predicted per-capita spending in FAM equal the corresponding spending in National 
%Health Expenditure Accounts (NHEA) for age group 55-64 in year 2004 and ages 65 and over in year 2010, respectively. Table 
%\ref{tab:NHEA_adjustment} shows how these adjustment factors were determined by using the ratio of expenditures in the NHEA to 
%expenditures predicted in the FAM.

Since 2006, the Medicare Current Beneficiaries Survey (MCBS) contains data on Medicare Part D. The data gives the capitated Part D payment and 
enrollment. When compared to the summary data presented in the CMS 2007 Trustee Report, the 2006 per capita cost is comparable between the MCBS 
and the CMS. However, the enrollment is underestimated in the MCBS, 53\% compared to 64.6\% according to CMS.

% \todo : add the tables or plots of Part D enrollment, payment, etc.

A cross-sectional probit model is estimated using years 2007-2010 to link demographics, economic status, current health, and 
functional status to Part D enrollment - see the estimates table. To account for both the initial under reporting of Part D 
enrollment in the MCBS, as well as the CMS prediction that Part D enrollment will rise to 75\% by 2012, the constant in the probit model is 
increased by 0.22 in 2006, to 0.56 in 2012 and beyond.  The per capita Part D cost in the MCBS matches well with the cost reported from 
CMS. An OLS regression using demographic, current health, and functional status is estimated for Part D costs based on capitated payment amounts.

The Part D enrollment and cost models are implemented in the Medical Cost module. The Part D enrollment model is executed conditional on 
the person being eligible for Medicare, and the cost model is executed conditional on the enrollment model leading a true result, after the 
Monte Carlo decision. Otherwise the person has zero Part D cost. The estimated Part D costs are added with Part A and B costs to obtain 
total Medicare cost, and any medical cost growth assumptions are then applied.

### 2019-11-22 analysis/techdoc/latex/FAM/implementation.tex
First commit in git, migrated from a subversion repository

\section{Implementation}
The FAM is implemented in multiple parts. Estimation of the transition and cross 
sectional models is performed in Stata.
The replenishing cohort model is estimated in Stata using the CMP package \citep{statacmp2011}.
The simulation is implemented in C++ for speed and flexibility.  Currently, the simulation is
run on Linux, Windows, and Mac OS X.

To match the two year structure of the PSID data used to estimate the 
transition models, the FAM simulation proceeds in two year increments. The end 
of each two year step is designed to occur on July 1st to allow for easier 
matching to population forecasts from Social Security. A simulation of the FAM 
proceeds by first loading a population representative of the age 25+ US 
population in 2009, generated from PSID. In two year increments, the FAM applies 
the transition models for mortality, health, working, wealth, earnings, and 
benefit claiming with Monte Carlo decisions to calculate the new states of the 
population. The population is also adjusted by immigration forecasts from the 
US Census Department, stratified by race and age. If incoming cohorts are being 
used, the new 25/26 year olds are added to the population. The number of new 
25/26 year olds added is consistent with estimates from the Census, stratified 
by race. Once the new states have been determined and new 25/26 year olds added, 
the cross sectional models for medical costs are performed. Summary variables 
are then computed. Computation of medical costs includes the persons that died 
to account for end of life costs. To reduce uncertainty due to the Monte 
Carlo decision rules, the simulation is performed multiple times (typically 
100), and the mean of each summary variable is calculated across repetitions.

FAM simulation takes as inputs assumptions regarding  
% [excluding for now...]
%the national wage index, 
the normal retirement age, real medical cost growth, and interest rates. The 
default assumptions are taken from the 2010 Social Security Intermediate 
scenario, adjusted for no price increases after 2010. Therefore simulation 
results are in real 2009 dollars.  
%Table \ref{tab:time_series_by_year} shows the 
%assumptions for each calendar year and Table \ref{tab:time_series_by_yob} shows 
%assumptions for each birth year.

Different simulation scenarios are implemented by changing any of the following 
components: incoming cohort model, transition models, interventions that adjust 
the probabilities of specific transition, and changes to assumptions on future 
economic conditions.

\input{./implementation_interventions.tex}

### 2019-11-22 analysis/techdoc/latex/FAM/implementation_interventions.tex
First commit in git, migrated from a subversion repository

\subsection{Intervention Module}
The intervention module can adjust characteristics of individuals when they are first read into the simulation ``init\_interventions'' or alter transitions within the 
simulation ``interventions.''  At present, init\_interventions can act on chronic diseases, ADL$/$IADL status, program participation, and some demographic characteristics.  
Interventions within the simulation can currently act on mortality, chronic diseases, and some program participation variables.

Interventions can take several forms.  The most commonly used is an adjustment to a transition probability. One can also delay the assignment of a chronic condition or cure an existing 
chronic condition.  Additional flexibility comes from selecting who is eligible for the intervention.  Some examples might help to make the interventions concrete.

\begin{itemize}
\item Example 1: Delay the enrollment into Social Security OASI by two years.  In this scenario claiming of Social Security benefits is transitioned as normal.  However, if a person is predicted to 
claim their benefits, then that status is not immediately assigned, but is instead assigned two years later.

\item Example 2: Cure hypertension for those with no other chronic diseases.  In this scenario any individual with hypertension (including those who have had hypertension for many years) is cured 
(hypertension status is set to 0), as long as they do not have other chronic diseases.  This example uses the individual’s chronic disease status as the eligibility criteria for the intervention.

\item Example 3: Reduce the incidence of hypertension for half of men aged 55 to 65 by 10\% in the first year of the simulation, gradually increasing the reduction to 20\% after 10 years.  This 
example begins to show the flexibility in the intervention module.  The eligibility criteria are more complex (half of men in a specific age range are eligible) and the intervention 
changes over time.  Mathematically, the intervention works by acting on the incidence probability, $\rho$. In the first year of the simulation, the probability is replaced by 
 $\left(1-0.5*0.1\right)\rho=0.95\rho$.  The binary outcome is then assigned based on this new probability.  Thus, at the population level, there is a 5\% reduction in incidence for men aged 55 to 65, 
as desired.  After 10 years, the probability for this eligible population becomes $\left(1-0.5*0.2\right)\rho=0.9\rho$.
\end{itemize}

More elaborate interventions can be programmed by the user.

### 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model.tex
First commit in git, migrated from a subversion repository

\section{Model for replenishing cohorts}
\label{sec:model_for_new_cohorts}
We first discuss the empirical strategy, then present the model and estimation 
results. The model for replinishing cohorts integrates information coming from trends 
among younger cohorts with the joint distribution of outcomes in the current 
population of age 25 respondents in the PSID.

%\input{new_cohorts_model_empirical_strategy.tex}
\input{new_cohorts_model_model_and_estimation.tex}
\input{new_cohorts_model_trends.tex}

### 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model_model_and_estimation.tex
First commit in git, migrated from a subversion repository

\subsection{Model and estimation}
Assume the latent model for $\mathbf{y}^*_i = (y^*_{i1},\ldots,y^*_{iM})'$,
\[
\mathbf{y}^*_i = \mathbf{\mu} + \mathbf{\varepsilon} _i, 
\]
where $\varepsilon_i$ is normally distributed with mean zero and covariance matrix $\mathbf{\Omega}$. 
It will be useful to write the model as 
\[
\mathbf{y}^*_i = \mathbf{\mu} + \mathbf{L}_\Omega \mathbf{\eta}_i,
\]
where $\mathbf{L}_\Omega$ is a lower triangular matrix such that 
$\mathbf{L}_\Omega \mathbf{L}_\Omega' = \mathbf{\Omega}$ and $\mathbf{\eta}_i = (\eta_{i1},\ldots,\eta_{iM})'$ 
are standard normal. We observe $y_i = \Gamma(y^*_i)$ which is a non-invertible mapping for a subset 
of the $M$ outcomes. For example, we have binary, ordered and censored outcomes for which integration 
is necessary.

The vector $\mathbf{\mu}$ can depend on some variables which have a stable distribution over time $\mathbf{z}_i$ 
(say race, gender and education). This way, estimation preserves the correlation with these outcomes 
without having to estimate their correlation with other outcomes. Hence, we can write 
\[
\mathbf{\mu}_i = \mathbf{z}_i \beta
\]
and the whole analysis is done conditional on $\mathbf{z}_i$.

For binary and ordered outcomes, we fix $\Omega_{m,m}=1$ which fixes the scale. Also we fix the 
location of the ordered models by fixing thresholds as $\tau_0 = -\infty$, $\tau_1 = 0$, $\tau_K = +\infty$, 
where $K$ denotes the number of categories for a particular outcome. 
We also fix to zero the correlation between selected outcomes (say earnings) and 
their selection indicator. Hence, we consider two-part models for these outcomes. Because some parameters are 
naturally bounded, we also re-parameterize the problem to guarantee an interior solution. In particular, we parameterize 
\begin{align*}
&\Omega_{m,m} = \exp(\delta_m), \mbox{   } m=m_0-1,\ldots,M \\
&\Omega_{m,n} = \tanh(\xi_{m,n})\sqrt{\Omega_{m,m}\Omega_{m,n}}, \mbox{   } m,n=1,\ldots,N \\
&\tau_{m,k} = \exp(\gamma_{m,k}) + \tau_{k-1}, \mbox{   } k=2,\ldots,K_m-1, m \mbox{ ordered}
\end{align*}
and estimate the $(\delta_{m,m}, \xi_{m,n}, \gamma_k)$ instead of the original parameters. 
The parameter values are estimated using the \emph{cmp} package in Stata \citep{statacmp2011}.
Table \ref{tab:latent_model_mean_est} gives parameter estimates for the indices while Table \ref{tab:latent_model_vcmat_est} 
gives parameter estimates of the covariance matrix in the outcomes.

%To apply trends to the future cohorts, the latent model is written as
%\[
%\mathbf{y}^*_i = \mathbf{\mu} + \mathbf{L}_\Omega \mathbf{\eta}_i.
%\]
%Each marginal has a mean change equal to $\mathrm{E}(\mathbf{y} \mid \mathbf{\mu}) = (1+\tau)g(\mathbf{\mu})$, where $\tau$ is the 
%percent change in the outcome and $g()$ is a non-linear but monotone mapping. Since 
%it is invertible, we can find the vector $\mathbf{\mu}^*$ where $\mathbf{\mu}^* = g^{-1}(\mathrm{E}(y \mid \mu)/(1+\tau))$. We use these new intercepts to simulate new outcomes.

% \todo : add an  example here for clarification

### 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model_trends.tex
First commit in git, migrated from a subversion repository

\subsection{Trends for replenishing cohorts}
Using the jointly estimated models previously described, we then assign outcomes to the replenishing cohorts, imposing trends for some health, 
risk factor, and social outcomes.  We currently impose trends on BMI, education, number of children, marital status, hypertension, and smoking 
status for these 25-26 year olds.  These trends are estimated using the National Health Interview Survey (health and risk factors) or the American Community 
Survey (social outcomes). All trends are halted after 2029. The trends are shown in Table \ref{tab:new_cohort_trends_health}, Table \ref{tab:new_cohort_trends_educ} and 
Table \ref{tab:new_cohort_trends_social}.

### 2019-11-22 analysis/techdoc/latex/FAM/validation_crossvalidation.tex
First commit in git, migrated from a subversion repository

\subsection{Cross-validation}
The cross-validation exercise randomly samples half of the PSID respondent IDs for use in estimating the transition models.  The respondents not used for estimation, 
but who were present in the PSID sample in 1999, are then simulated from 1999 through 2013.  Demographic, health, and economic outcomes are compared between the 
simulated (“FAM”) and actual (“PSID”) populations.

%These results are presented in Table \ref{tab:crossval_unweighted} - Table \ref{tab:crossval_cntecon} for 2000, 2006, and 2012, 
%with a statistical test of the difference between the average values in the two populations.

Worth noting is how the composition of the population changes in this exercise.  In 1999, the sample represents those 25 and older.  Since we follow a fixed cohort, 
the age of the population will increase to 39 and older in 2013.  This has consequences for some measures in later years where the eligible population shrinks.

\subsubsection{Demographics}
Mortality and demographic measures are presented in Tables \ref{tab:crossval_unweighted} and \ref{tab:crossval_demog}.  Mortality incidence is comparable between
the simulated and observed populations.  Demographic characteristics do not differ between the two.

\subsubsection{Health Outcomes}
Binary health outcomes are presented in Table \ref{tab:crossval_binhlth}. FAM underestimates the prevalence of ADL and IADL limitations compared to the crossvalidation sample. 
Binary outcomes, like cancer, diabetes, heart diesease, and stroke do not differ.  FAM underpredicts hypertension and lung disease compared to the crossvalidation sample.

\subsubsection{Health Risk Factors}
Risk factors are presented in Table \ref{tab:crossval_risk}.  BMI is not statistically different between the two samples.  Current smoking is not statistically different,
 but more individuals in the crossvalidation sample report being former smokers.

\subsubsection{Economic Outcomes}
Binary economic outcomes are presented in Table \ref{tab:crossval_binecon}.  FAM underpredicts claiming of federal disability and overpredicts Social Security retirement 
claiming.  Supplemental Security claiming is not statistically different between FAM and the crossvalidation sample.  Working for pay is not statistically different.

%Continuout economic outcomes are presented in Table \ref{tab:crossval_cntecon}.

On the whole, the crossvalidation exercise is reassuring.  There are differences that will be explored and improved upon in the future.

### 2019-11-22 analysis/techdoc/latex/FAM/validation_external_corroboration.tex
First commit in git, migrated from a subversion repository

\subsection{External Corroboration}
Finally, we compare FAM population forecasts to Census forecasts of the US population..  Here, we focus on the full PSID population (25 and older) and those 65 and 
older.  For this exercise, we begin the simulation in 2009 and simulate the full population through 2049.  Population projections are compared to the 2012 Census 
projections for years 2012 through 2049.  See resuts in Table \ref{tab:external_pop}.  By 2049, FAM forecasts for 25 and older remain within 2\% of Census forecasts.

### 2019-11-22 analysis/techdoc/latex/FEM/FEM_tables.tex
First commit in git, migrated from a subversion repository

\section{Tables}

\begin{sidewaystable}
\centering
\begin{tabular}[h]{l rrrrrrrr}
\input{../tables/FEM/svy_disease_prevalence.tex}
\end{tabular}
\caption{Health condition prevalences in survey data}
\label{tab:svy_disease_prevalence}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.75}{
\begin{tabular}{l p{2.5in}p{2.5in}p{2.5in}p{2.5in}}
& \multicolumn{4}{c}{Survey}\\
Disease & \multicolumn{1}{c}{HRS} & \multicolumn{1}{c}{NHIS} & \multicolumn{1}{c}{MEPS} & \multicolumn{1}{c}{MCBS} \\
\hline
Cancer &
Has a doctor ever told you that you have cancer or a malignant tumor, excluding minor skin cancers?&
Have you ever been told be a doctor or other health professional that you had … cancer or a malignanacy of any kind? (WHEN RECODED, SKIN CANCERS WERE EXCLUDED)&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 11-21, 24-45&
Has a doctor ever told you that you had any (other) kind of cancer malignancy, or tumor other than skin cancer?\\
\hline
Heart Diseases &
Has a doctor ever told you that you had a heart attack, coronary heart disease, angina, congestive heart failure, or other heart problems?&
Four separate questions were asked about whether ever told by a docotor or oither health professional that had: CHD, Angina, MI, other heart problems.&
Have you ever been told by a doctor or health professional that you have … CHD; Angina; MI; other heart problems&
Six separate questions were asked about whether ever told by a doctor that had: Angina or MI; CHD; other heart problems (included four questions)\\
\hline
Stroke &
Has a doctor ever told you that you had a stroke?&
Have you EVER been told by a doctor or other health professional that you had a stroke?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have a stroke or TIA (transient ischemic attack)&
[Since (PREV$<$ SUPP. RD. INT. DATE),] has a doctor (ever) told (you/SP) that (you/he/she) had a stroke, a brain hemorrhage, or a cerebrovascular accident?\\
\hline
Diabetes&
Has a doctor ever told you that you have diabetes or high blood sugar?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
If Female, add: [Other than during pregnancy,] Have you hever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
Has a doctor (ever) told (you/SP) that (you/he/she) had diabetes, high blood sugar, or sugar in (your/his/her) urine? [DO NOT INCLUDE BOERDERLINE PREGNANCY, OR PRE-DIABETEIC DIABETES.]\\
\hline
Hypertension&
Has a doctor ver told you that you have high blood pressure or hypertension?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Has a doctor (ever) told (you/SP) that (you/he/she) (still) (had) (have/has) hypertension, sometimes called high blood pressure?\\
\hline
Lung Disease&
Has a doctor ever told you that you have chronic lung disease such a schronic bronchitis or emphysema? [IWER: DO NOT INCLUDE ASTHMA]&
Question 1: During the PAST 12 MONTHS, have you ever been told by a doctor or other health professional that you had chronic bronchitis? Question 2: Have you EVER been told by a docotor or other health professional that you had … emphysema?&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 127, 129-312&
Has a doctor (ever) told (you/SP) that (you/he/she) had emphysema, asthma, or COPD? [COPD=CHRONIC OBSTRUCTIVE PULMONARY DISEASE.]\\
\hline
Overweight & \multicolumn{4}{c}{\multirow{2}{*}{Self-reported body weight and height}}\\
\hhline{-~~~~}
Obese & \multicolumn{4}{c}{} \\
\hline
\end{tabular}
}
\caption{Survey questions used to determine health conditions}
\label{tab:svy_disease_questions}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{0.85}{
\begin{tabular}[h]{p{0.95in}p{1.2in}p{1.6in}p{1.8in}}
\textbf{Conditions} & \textbf{Data source} & \textbf{Projection method} & \textbf{Other sources}\\
\hline
Diabetes\newline Heart disease\newline Hypertension\newline & 
National Health Interview Survey 1997-2006 & 
Use synthetic cohort approach to estimate age-specific incidence rate for each condition & 
There are other forecasts \citep{honeycutt2003dynamic,mainous2007impact} for the trends of diabetes in the U.S population; we compare their forecasts to ours and they are reasonably close \\
\hline
Overweight and obese & {
Prevalence of over-weight and obese for aged 46-56 from year 2001 to 2030, generated by Ruhm upon request} & {
Assume annual rate of change during year 2031-2050 linearly decreases from the 2030 rate to zero in 2050} & \citet{ruhm2007current} \\
\hline
Ever-smoked and smoking now & {
Forecast of prevalence of ever-smoked and smoking now for aged 45-54 from year 2005 to 2025, by \citet{levy2006smoking}} & {
For ever-smoked, assume that the prevalence at age 45-54 in year 2035 (2045) is the same as prevalence at age 35-44 (25-34) in year 2025. Assume that the annual change in prevalence at age 45-54 in year 2046-2050 the same as average in 2040-2045. For smoking-now, after year 2025, use the moving average of the past five years} & {
}\\
\hline
Any DB from current job & {
} & {
Assume annual relative declining rate for DB entitlement decrease by 2\% a year} & {
Historical trends of DB participation rates among all persons by different birth cohorts and by age, by \citet{poterba2007decline}}\\
\hline
Any DC from current job & {
} & {
Assume annual relative increasing rate for DC entitlement increase by 2\% a year until 2026 then stays the same after 2026} & {
Forecast of DC participation rates among all persons by different birth cohorts and by age, by \citet{poterba2008new}}\\
\hline
Population size 50-52\newline Male\newline Hispanic\newline Non-Hispanic black\newline & {
Census Bureau 2000-2010 Intercensal Population Estimates, 2012 National Population Estimates, and 2012 National Population Projections} & {
Projected 2060 - 2080 using linear trend based on 2040-2060} & {
} \\
\hline
\end{tabular}
}
\caption{Data sources and methods for projecting future cohort trends}
\label{tab:cohort_projection_data_methods}
\end{table}

\begin{sidewaystable}
\centering
\input{../tables/FEM/baseline_trends.tex}
\caption{Projected baseline trends for future cohorts}
\label{tab:baseline_trends}
\end{sidewaystable}

\begin{table}
\centering
\input{../tables/FEM/prevalence_1978_2004.tex}
\caption[Prevalence of obesity, hypertension, diabetes and current smokers among ages 46-56 in 1978 and 2004]{Prevalence of obesity, hypertension, diabetes and current smokers among ages 46-56 in 1978 and 2004. Prevalence in 1978 is based on NHANES II 1976-1980; Prevalence in 2004 is based on NHANES 2003-2004. 
BMI is calculated using self-reported weight and height.}
\label{tab:prevalence_1978_2004}
\end{table}

\begin{sidewaystable}
% \todo Dynamically fill in numeric results using Sweave
% For now, the numerical results can be copied from analysis/techdoc/table6.csv
\centering
\begin{tabular}{lllccr}
 &  &  & Type & At risk & Mean/fraction\\
\hline
\input{../tables/FEM/transitioned_outcomes_disease.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_smoking.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_logbmi.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_adlstat.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_iadlstat.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_lfpben.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_financial.tex}
\hline
\end{tabular}
\caption[Outcomes in the transition model]{Outcomes in the transition model. Estimation sample is HRS 1991-2008 waves.}
\label{tab:transitioned_outcomes}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Write a script that dynamically generates this table from the .est files
\centering
\scalebox{0.53}{
\begin{tabular}{l*{20}{c}}
Value at time $T-1$ & \multicolumn{20}{c}{Outcome at time $T$}  \\
\hline
 & Heart  &  & & Lung  & & &  &  & Smoking  &  &  & &  &  & & Nursing & & &  Nonzero & \\
 & disease & hypertension & stroke & disease & diabetes & cancer & disability & mortality & status & BMI & Any HI & DI Claim & SS Claim & DB Claim & SSI Claim & Home & Work & Earnings & Wealth & Wealth\\
Heart disease &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Blood pressure & \checkmark &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Stroke &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Lung disease &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Diabetes & \checkmark & \checkmark & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Cancer &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Disability &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DI &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SS &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DB &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SSI &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  &  &  & \\
Work &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Earnings &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Nonzero wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Nursing home stay &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark &  &  & \checkmark & \checkmark
\end{tabular}
}
\caption[Restrictions on transition model]{Restrictions on transition model. \checkmark indicates that an outcome at time $T-1$ is allowed in the transition model for an outcome at time $T$.}
\label{tab:transition_restrictions}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FEM/controlvar_descs.tex}
\caption[Desciptive statistics for exogeneous control variables]{Desciptive statistics for exogeneous control variables in 2004 HRS ages 51+ sample used as simulation stock population}
\label{tab:controlvar_descs}
\end{sidewaystable}

%\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
%\scalebox{0.75}{
%\input{../tables/FEM/validation_waves_4_7_9.tex}
%}
%\caption{Simulated and actual outcomes in 1998, 2004, and 2008}
%\label{tab:validation_waves_4_7_9}
%\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_unweighted.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported mortality and nursing home outcomes in 2000, 2006, and 2012}
\label{tab:crossval_unweighted}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_demog.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported demographic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_demog}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_binhlth.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported binary health outcomes in 2000, 2006, and 2012}
\label{tab:crossval_binhlth}
\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_risk.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported risk factor outcomes in 2000, 2006, and 2012}
\label{tab:crossval_risk}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_binecon.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported binary economic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_binecon}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_cntecon.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported continuous economic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_cntecon}
\end{table}

\begin{table}
\centering
\begin{tabular}{cl|rr|rccl|rr|r}
 & \multicolumn{4}{c}{MEPS 2001 (ages 51+)} & \multicolumn{2}{c}{}  & \multicolumn{4}{c}{HRS 1998 (ages 51+)}\\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{ADL limitation} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{ADL limitation} & \multicolumn{1}{c}{} \\
%\hhline{~~--~~~~--~}
 &  & No & Yes & All & & &  & No & Yes & All\\
\hhline{~----~~----}
IADL & No & 91.5 & 0.4 & 92.0 & & IADL & No & 82.0 & 10.5 & 92.5\\
limitation & Yes & 4.4 & 3.7 & 8.0 & & limitation & Yes & 3.1 & 4.5 & 7.5\\
\hhline{~----~~----}
  & All & 95.9 & 4.1 & 100.0 & &  & All & 85.1 & 14.9 & 100.0\\
\end{tabular}
\caption[Prevalence of IADL and ADL limitations among ages 51+ in MEPS 2001 and HRS 1998]
{Prevalence of IADL and ADL limitations among ages 51+ in MEPS 2001 and HRS 1998.
The IADL limitations in MEPS are defined as receiving help or supervision using the 
telephone, paying bills, taking medications, preparing light meals, doing laundry, 
or going shopping; the ADL limitations in HRS are defined as receiving help or 
supervision with personal care such as bathing, dressing, or getting around the 
house. 
The IADL limitations in HRS are defined as having any difficulty in at least one of 
the following activities: using the phone, taking medications, and managing money. 
The ADL limitations in HRS are defined as having any difficulty in at least one of the 
following activities: bathing, dressing, eating, walking across the room, and 
getting out of bed. 
}
\label{tab:meps_hrs_adl_iadl_prevalence}
\end{table}

\begin{table}
\centering
\begin{tabular}{cl|rr|rccl|rr|r}
 & \multicolumn{4}{c}{MEPS 2001 (ages 51+)} & \multicolumn{2}{c}{}  & \multicolumn{4}{c}{HRS 1998 (ages 51+)}\\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{Physical function} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{Physical function} & \multicolumn{1}{c}{} \\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{limitation} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{limitation} & \multicolumn{1}{c}{} \\
%\hhline{~~--~~~~--~}
 &  & No & Yes & All & & &  & No & Yes & All\\
\hhline{~----~~----}
IADL & No & 61.6 & 30.4 & 92.0 & & IADL & No & 60.0 & 32.5 & 92.5\\
limitation & Yes & 0.3 & 7.8 & 8.0 & & limitation & Yes & 1.0 & 6.5 & 7.5\\
\hhline{~----~~----}
  & All & 61.9 & 38.2 & 100.0 & &  & All & 61.0 & 39.0 & 100.0\\
\end{tabular}
\caption[Prevalence of IADL limitation and physical function limitation among ages 51+ in MEPS 2001 and HRS 1998]
{Prevalence of IADL limitation and physical function limitation among ages 51+ in MEPS 2001 and HRS 1998. 
The definition of IADL limitation is the same as in Table \ref{tab:meps_hrs_adl_iadl_prevalence}.
Physical function limitation in MEPS indicates that at least one of the following is true: 1) receiving help or 
supervision with bathing, dressing or walking around the house; 2) being limited in work/housework; 3) having 
difficulty walking, climbing stairs, grasping objects, reaching overhead, lifting, bending or stooping, or 
standing for long periods of time; or 4) having difficulty in hearing or vision.
Physical function limitation in HRS indicates at least one of the following is true: 1) having any difficulty in 
bathing/dressing/eating/walking across the room/getting out of bed; 2) limited in work/housework; or 3) limited in 
any other activities.
}
\label{tab:meps_hrs_physlim_iadl_prevalence}
\end{table}

\begin{table}
\centering
\scalebox{0.86}{
\begin{tabular}{l|rrrrr}
  & Model I & Model II & Model III & Model IV & Model V\\
\hline
Constant & 0.877*** & 0.898*** & 0.874*** & 0.839*** & 0.869***\\
  & (0.002) & (0.003) & (0.005) & (0.002) & (0.003)\\
Physical function limitation & -0.115*** & -0.098*** & -0.094*** &   &  \\
  & (0.004) & (0.005) & (0.004) &   &  \\
IADL limitation & -0.041 & -0.019 & -0.008 &   &  \\
  & (0.037) & (0.042) & (0.036) &   &  \\
IADL limitation * Physical function limitation & -0.150*** & -0.156*** & -0.162*** &   &  \\
  & (0.037) & (0.044) & (0.037) &   &  \\
IADL limitation, no ADL limitation &   &   &   & -0.182*** &  \\
  &   &   &   & (0.009) &  \\
Any ADL limitation &   &   &   & -0.344*** &  \\
  &   &   &   & (0.010) &  \\
Ever diagnosed with cancer &   & -0.011 & -0.015** &   & -0.030***\\
  &   & (0.009) & (0.007) &   & (0.010)\\
Ever diagnosed with diabetes &   & -0.034*** & -0.032*** &   & -0.054***\\
  &   & (0.007) & (0.005) &   & (0.007)\\
Ever diagnosed with high blood pressure &   & -0.030*** & -0.028*** &   & -0.043***\\
  &   & (0.004) & (0.004) &   & (0.005)\\
Ever diagnosed with heart disease &   & -0.024*** & -0.029*** &   & -0.055***\\
  &   & (0.006) & (0.005) &   & (0.006)\\
Ever diagnosed with lung disease &   & -0.036*** & -0.032*** &   & -0.055***\\
  &   & (0.009) & (0.007) &   & (0.010)\\
Ever diagnosed with stroke &   & -0.045*** & -0.046*** &   & -0.115***\\
  &   & (0.012) & (0.008) &   & (0.013)\\
Age 65-74 &   &   & 0.010** &   &  \\
  &   &   & (0.004) &   &  \\
Age 75 and over &   &   & 0.015*** &   &  \\
  &   &   & (0.005) &   &  \\
Male &   &   & 0.028*** &   &  \\
  &   &   & (0.004) &   &  \\
Non-Hispanic black &   &   & 0.008 &   &  \\
  &   &   & (0.007) &   &  \\
Hispanic &   &   & -0.001 &   &  \\
  &   &   & (0.007) &   &  \\
Less than HS &   &   & -0.022*** &   &  \\
  &   &   & (0.005) &   &  \\
Some college &   &   & 0.016*** &   &  \\
  &   &   & (0.005) &   &  \\
College grad &   &   & 0.037*** &   &  \\
  &   &   & (0.005) &   &  \\
Census region: Northeast &   &   & 0.003 &   &  \\
  &   &   & (0.005) &   &  \\
Census region: Midwest &   &   & 0.004 &   &  \\
  &   &   & (0.005) &   &  \\
Census region: West &   &   & -0.012** &   &  \\
  &   &   & (0.005) &   &  \\
Marital status:widowed &   &   & 0.003 &   &  \\
  &   &   & (0.005) &   &  \\
Marital status: single &   &   & -0.013*** &   &  \\
  &   &   & (0.005) &   &  \\
\hline
$N$ & 7,358 & 7,317 & 7,317 & 7,361 & 7,322\\
\hline
Adjusted $R^2$ & .24 & .27 & .29 & .18 & .11\\
\hline
\multicolumn{6}{c}{* $p<0.10$, ** $p<0.05$, *** $p<0.01$}  
\end{tabular}
}
\caption[OLS regressions of EQ-5D utility index among ages 51+ in MEPS 2001]{OLS regressions of EQ-5D utility index among ages 51+ in MEPS 2001. $p$-values in parentheses. Data source: MEPS 2001 (ages 51+). EQ-5D scoring algorithm is based on \citet{shaw2005us}.}
\label{tab:meps_eq5d_model_est}
\end{table}

\begin{table}
\centering
\begin{tabular}{l|r}
\hline
Ever diagnosed with cancer & -0.020*\\
  & (0.001)\\
Ever diagnosed with diabetes & -0.042*\\
  & (0.001)\\
Ever diagnosed with heart disease & -0.044*\\
  & (0.001)\\
Ever diagnosed with high blood pressure & -0.034*\\
  & (0.001)\\
Ever diagnosed with lung disease & -0.054*\\
  & (0.001)\\
Ever diagnosed with stroke & -0.067*\\
  & (0.002)\\
IADL limitation only & -0.160*\\
  & (0.002)\\
One or two ADL limitations & -0.099*\\
  & (0.001)\\
Three or more ADL limitations  & -0.149*\\
  & (0.002)\\
Constant & 0.881*\\
  & (0.001)\\
\hline
$N$ & 19,676  \\
\hline
Adjusted $R^2$ & 0.67\\
\hline
\multicolumn{2}{c}{* $p<0.01$}
\end{tabular}
\caption[OLS regression of the predicted EQ-5D index score against chronic conditions and FEM-type functional status specification]
{OLS regression of the predicted EQ-5D index score against chronic conditions and FEM-type functional status 
specification. $p$-values in parentheses. Data source: Health and Retirement Study, 1998. Sample included the 
age 51 and over community respondents. EQ-5D score was predicted using Model II in Table \ref{tab:meps_eq5d_model_est}.   
}
\label{tab:fem_eq5d_model_est}
\end{table}

\begin{sidewaystable}
\centering
\begin{tabular}{|l|lrr|rrrrrr|}
 \multicolumn{4}{l}{}  & \multicolumn{6}{c}{Prevalence of chronic conditions}\\
\hhline{~~~~------}
\multicolumn{1}{l}{}                  &           & Predicted   &         &        &          & Heart   &              & Lung    &       \\
\multicolumn{1}{l}{Functional status} &  Estimate & EQ-5D score & Age     & Cancer & Diabetes & disease & Hypertension & disease & Stroke\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{Healthy}} & Mean & 0.841 & 63.9 & 0.114 & 0.142 & 0.189 & 0.481 & 0.072 & 0.045\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.042 & 10.0 & 0.318 & 0.349 & 0.391 & 0.500 & 0.258 & 0.207\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.000 & 0.0 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{IADL limitation only - not in nursing home}} & Mean & 0.655 & 70.0 & 0.151 & 0.223 & 0.334 & 0.581 & 0.152 & 0.162\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.054 & 12.9 & 0.358 & 0.416 & 0.472 & 0.493 & 0.359 & 0.368\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.005 & 0.005 & 0.006 & 0.006 & 0.005 & 0.005\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{1 or 2 ADL limitations - not in nursing home}} & Mean & 0.705 & 69.1 & 0.180 & 0.291 & 0.393 & 0.665 & 0.194 & 0.162\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.057 & 12.0 & 0.384 & 0.454 & 0.488 & 0.472 & 0.395 & 0.368\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.000 & 0.1 & 0.003 & 0.003 & 0.004 & 0.003 & 0.003 & 0.003\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{3 or more ADL limitations  - not in nursing home}} & Mean & 0.636 & 70.9 & 0.181 & 0.361 & 0.467 & 0.691 & 0.216 & 0.314\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.063 & 13.2 & 0.385 & 0.480 & 0.499 & 0.462 & 0.412 & 0.464\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.005 & 0.006 & 0.006 & 0.006 & 0.005 & 0.006\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{Nursing home residency}} & Mean & 0.568 & 82.0 & 0.144 & 0.256 & 0.428 & 0.651 & 0.142 & 0.372\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.077 & 10.0 & 0.351 & 0.437 & 0.495 & 0.477 & 0.349 & 0.484\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.006 & 0.008 & 0.009 & 0.008 & 0.006 & 0.009\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{All}} & Mean & 0.808 & 65.2 & 0.125 & 0.170 & 0.229 & 0.514 & 0.094 & 0.076\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.083 & 10.9 & 0.331 & 0.375 & 0.420 & 0.500 & 0.291 & 0.266\\
\multicolumn{1}{|p{2in}|}{}  & Std. err & 0.000 & 0.0 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001\\
\hline
\end{tabular}
\caption[Average predicted EQ-5D score, age, and prevalence of chronic conditions by functional status for the stock FEM simulation sample]
{Average predicted EQ-5D score, age, and prevalence of chronic conditions by functional status for the stock FEM simulation sample 
(ages 51 and over in 2004).
EQ-5D scores were predicted according to parameter estimates in Table \ref{tab:fem_eq5d_model_est}. The predicted 
score for nursing home residents is reduced by 10\% to account for the fact that the estimation sample in Table 
\ref{tab:fem_eq5d_model_est} only includes non-nursing home residents.
}
\label{tab:fem_stock_predicted_eq5d}
\end{sidewaystable}
\clearpage

\begin{table}
% \todo Dynamically fill in numeric results using Sweave and reduce the number of sig. digits after decimal
\centering
\scalebox{0.9}{
\begin{tabular}{ccccrr}
 &  &  & Selection & 1992 & 2004 \\
\hline
\input{../tables/FEM/desc_init_conditions_binary.tex}
\hline
\input{../tables/FEM/desc_init_conditions_bmistat.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_smoking.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_funcstat.tex}
\hline
\input{../tables/FEM/desc_init_conditions_continuous.tex}
\hline
\input{../tables/FEM/desc_init_conditions_censcont.tex}
\hline
\input{../tables/FEM/desc_init_conditions_censdiscrete.tex}
\hline
\input{../tables/FEM/desc_init_conditions_earlyagedb.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_normalagedb.tex}
\hline
\input{../tables/FEM/desc_init_conditions_covars.tex}
\hline
\end{tabular}
}
\caption{Initial conditions used for estimation (1992) and simulation (2004)}
\label{tab:desc_init_conditions}
\end{table}

\begin{sidewaystable}
\centering
\scalebox{0.53}{
\input{../tables/FEM/latent_model_mean_est.tex}
}
\caption[Parameter estimates for latent model: conditional means and thresholds]{Parameter estimates for latent model: conditional means and thresholds. Sample is respondents age 50-55 in 1992 HRS wave}
\label{tab:latent_model_mean_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.53}{
\input{../tables/FEM/latent_model_vcmat_est.tex}
}
\caption[Parameter estimates for latent model: parameterized covariance matrix]{Parameter estimates for latent model: parameterized covariance matrix. Sample is respondents age 50-55 in 1992 HRS wave}
\label{tab:latent_model_vcmat_est}
\end{sidewaystable}

\begin{table}
\centering
\input{../tables/FEM/NHEA_adjustment.tex}
\caption{Per capita medical spending by payment source, age group, and year}
\label{tab:NHEA_adjustment}
\end{table}

\begin{table}
% \todo Generate this table dynamically using Sweave
% \todo add cancer, lung disease and stroke
\centering
\scalebox{0.9}{
\begin{tabular}{lrrr}
Year & 2010 & 2030 & 2050\\
\hline
\input{../tables/FEM/statusquo_results_popsize.tex}
\hline
\input{../tables/FEM/statusquo_results_conditions.tex}
\hline
\input{../tables/FEM/statusquo_results_labor.tex}
\hline
\input{../tables/FEM/statusquo_results_govrev.tex}
\hline
\input{../tables/FEM/statusquo_results_govexp.tex}
\hline
\input{../tables/FEM/statusquo_results_totmd.tex}
\hline
\end{tabular}
}
\caption{Simulation results for status quo scenario}
\label{tab:status_quo_results}
\end{table}

\begin{sidewaystable}
% \todo Generate this table dynamically using Sweave
\centering
\begin{tabular}{lrrcrrcrr}
 & \multicolumn{2}{c}{}                   &  & \multicolumn{2}{c}{Relative Change}  &  & \multicolumn{2}{c}{Absolute Change} \\
 & \multicolumn{2}{c}{``Obese 1980'' Estimates}  &  & \multicolumn{2}{c}{from Status Quo}  &  & \multicolumn{2}{c}{from Status Quo} \\
\hhline{~--~--~--}
Year & 2030 & 2050 &  & 2030 & 2050 &  & 2030 & 2050\\
\hline
\input{../tables/FEM/obesity_results_popsize.tex}
\hline
\input{../tables/FEM/obesity_results_conditions.tex}
\hline
\input{../tables/FEM/obesity_results_labor.tex}
\hline
\input{../tables/FEM/obesity_results_govrev.tex}
\hline
\input{../tables/FEM/obesity_results_govexp.tex}
\hline
\input{../tables/FEM/obesity_results_totmd.tex}
\hline
\end{tabular}
\caption{Simulation results for obesity reduction scenario compared to status quo}
\label{tab:obesity_results}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{0.8}{
\input{../tables/FEM/time_series_by_year.tex}
}
\caption{Assumptions for each calendar year}
\label{tab:time_series_by_year}
\end{table}

\begin{table}
\centering
\scalebox{0.65}{
\input{../tables/FEM/time_series_by_yob.tex}
}
\caption[Assumptions for each birth year]{Assumptions for each birth year. In years after 1960, all values are held constant at their 1960 levels.}
\label{tab:time_series_by_yob}
\end{table}

### 2019-11-22 analysis/techdoc/latex/FEM/acknowledgments.tex
First commit in git, migrated from a subversion repository

\section{Acknowledgments}

The Future Elderly Model has been developed by a large team over the last decade.  Jay Bhattacharya, Eileen Crimmins, Christine Eibner, \'Etienne Gaudette, Geoff Joyce, Darius Lakdawalla,
Pierre-Carl Michaud, and Julie Zissimopoulos have all provided expert guidance.  Adam Gailey, Baoping Shang, and Igor Vaynman provided programming and analytic 
support during the first years of development at RAND.  In more recent years, the University of Southern California research programming team has supported 
model development.  These programmers include Patricia St. Clair, Laura Gascue, Henu Zhao, and Yuhui Zheng. Barbara Blaylock and Wendy Cheng have greatly aided 
model development while working as research assistants at USC.

### 2019-11-22 analysis/techdoc/latex/FEM/baseline_forecasts.tex
First commit in git, migrated from a subversion repository

\section{Baseline Forecasts}
In this section we present baseline forecasts of the Future Elderly Model.  The figures show data from the HRS for the 55+ population from 1998 through 2012 and 
forecasts from the FEM for the 55+ population beginning in 2010.

\subsection{Disease Prevalence}
Figure \ref{fig:chronic_diseases_male} depicts the six chronic conditions we project for men.  And Figure \ref{fig:chronic_diseases_female} depicts the historic 
and forecasted values for women.

\begin{figure}[ht!]
\centering

\includegraphics[scale=1.0]{./img/chronic_diseases_male}
\caption{Historic and Forecasted Chronic Disease Prevalence for Men 55+}
\label{fig:chronic_diseases_male}

\includegraphics[scale=1.0]{./img/chronic_diseases_female}
\caption{Historic and Forecasted Chronic Disease Prevalence for Women 55+}
\label{fig:chronic_diseases_female} 
\end{figure}

Figure \ref{fig:adl_iadl_male} shows historic and forecasted levels for any ADL difficulties, three or more ADL difficulties, any IADL difficulties, and
two or more IADL difficulties for men 55 and older. Figure \ref{fig:adl_iadl_female} shows historic and forecasted levels for any ADL difficulties, three or 
more ADL difficulties, any IADL difficulties, and two or more IADL difficulties for women 55 and older.

\begin{figure}[ht!]
\centering

\includegraphics[scale=1.0]{./img/adl_iadl_male}
\caption{Historic and Forecasted ADL and IADL Prevalence for Men 55+}
\label{fig:adl_iadl_male}

\includegraphics[scale=1.0]{./img/adl_iadl_female}
\caption{Historic and Forecasted ADL and IADL Prevalence for Women 55+}
\label{fig:adl_iadl_female} 
\end{figure}

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation.tex
First commit in git, migrated from a subversion repository

\section{Data sources used for estimation}
The Health and Retirement Study is the main data source for the model. We supplemented this data with 
merged Social Security covered earnings histories and data on health trends and health care costs 
coming from 3 major health surveys in the U.S. We describe these surveys below and the samples we 
selected for the analysis. We first list the variables used in the analysis. We then give details on 
the data sources.

\begin{center}
\begin{tabular}{l l}
\multicolumn{2}{c}{\textbf{Estimated Outcomes in Initial Conditions Model}}\\
 & \\
\textbf{Economic Outcomes}	& \textbf{Health Outcomes}\\
Employment & Hypertension\\
Earnings & Heart Disease\\
Wealth & Self-Reported Health\\
Defined Contribution Pension Wealth	& BMI Status\\
Pension Plan Type	& Smoking Status\\
AIME & Functional Status\\
Social Security Quarters of Coverage & \\
Health Insurance & 
\end{tabular}\end{center}~\\~\\

\begin{center}
\begin{tabular}{l l l}
\multicolumn{3}{c}{\textbf{Estimated Outcomes in/from Transition Model}}\\
 & & \\
\textbf{Economic Outcomes} & \textbf{Health Outcomes} & \textbf{Other Outcomes}\\
Employment & Death & Income Tax Revenue\\
Earnings & Heart & Social Security Revenue\\
Wealth & Stroke & Medicare Revenue\\
Demographics & Cancer & Medical Expenses\\
Health Insurance & Hypertension & Medicare Part A Expenses\\
Disability Insurance Claim & Diabetes & Medicare Part B Expenses\\
Defined Benefit Claim & Lung Disease & Medicare Part B Enrollment\\
SSI Claim & Nursing Home &  Medicare Part D Enrollment\\
Social Security Claim & BMI  & OASI Enrollment \\
 & Smoking Status & DI enrollment\\
 & ADL Limitations & SSI enrollment\\
 & IADL Limitations	& Medicaid Enrollment\\
 & & Medicaid Expenditures
\end{tabular}
\end{center}

\input{data_sources_estimation_hrs.tex}
\input{data_sources_estimation_ssearnings.tex}
\input{data_sources_estimation_nhis.tex}
\input{data_sources_estimation_meps.tex}
\input{data_sources_estimation_mcbs.tex}

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation_hrs.tex
First commit in git, migrated from a subversion repository

\subsection{Health and Retirement Study}
The Health and Retirement Study (HRS) waves 2000-2008 are used to estimate the transition model. 
Interviews occur every two years. We use the dataset created by RAND (RAND HRS, version K) as our 
basis for the analysis. We use all cohorts in the analysis and consider sampling weights whenever 
appropriate. When appropriately weighted, the HRS in 2004 is representative of U.S. households where 
at least one member is at least 51. The HRS is also used as the host data for the simulation (pop 51+ in 2004) and for new cohorts (aged 51 and 52 in 2004).

The HRS adds new cohorts every six years. Until recently, the latest available cohort had been added 
in 2004, which is why that is the FEM's base year. The FEM is currently being updated to use the 
newly released 2010 data.

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation_meps.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_meps.tex
@@
 institutionalized population, neither does MEPS: this implies that we can only use the MEPS to 
-estimate medical costs for the non-elderly (25-64) population. Information collected during household 
+estimate medical costs for the non-elderly population. Information collected during household 
 interviews include: demographic characteristics, health conditions, health status, use of medical 
 services, sources of medical payments, and body weight and height. Each year the household survey 
-includes approximately 12,000 households or 34,000 individuals. Sample size for those aged 25-64 is 
-about 15,800 in each year.  MEPS has comparable measures of social-economic (SES) variables as those in PSID, 
-including age, race/ethnicity, educational level, census region, and marital status.  We estimate expenditures 
-and utilization using 2007-2010 data.
+includes approximately 12,000 households or 34,000 individuals. Sample size for those aged 51-64 is 
+about 4,500.  MEPS has comparable measures of social-economic (SES) variables as those in HRS, 
+including age, race/ethnicity, educational level, census region, and marital status.
 
-See Section 
-\ref{sec:govt_revenue_and_expenditures_medcost_estimation} for a description.  FAM also
-uses MEPS 2001-2003 data for QALY model estimation.
-
-% This is described in Section  \ref{sec:estimation_qalys_crosswalk_development}.
+FEM uses MEPS years 2000-2010 for cost estimation.  See Section 
+\ref{sec:govt_revenue_and_expenditures_medcost_estimation} for a description.  FEM also
+uses MEPS 2001 data for QALY model estimation. This is described in Section 
+\ref{sec:estimation_qalys}.

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation_ssearnings.tex
First commit in git, migrated from a subversion repository

\subsection{Social Security covered earnings files}
To get information on Social Security entitlements of respondents, we match the HRS data to the 
Social Security Covered Earnings files of 1992, 1993, 1998, 2004 and 2006 which provides information on 
earning histories of respondents as well as their entitlement to future Social Security benefits. We 
then construct the average indexed monthly earnings (AIME), the basis for the determination of benefit 
levels, from these earning histories. The AIME is constructed by first indexing using the National 
Wage Index (NWI) to the wage level when the respondent turns age 60. If this occurs after 2008, we 
project the evolution of the NWI using the average annual rate of change of the last 20 years (2.9\% 
nominal). We then take the 35 highest years (if less than 35 years are available, remaining years are 
considered zero earning years) and take the average. We then convert back this annual amount on a 
monthly basis and convert back to \$2004 U.S. dollars using the CPI. Quarters of coverage, which 
determine eligibility to Social Security, are defined as the sum of posted quarters to the file. A 
worker is eligible for Social Security if he has accumulated at least 40 quarters of coverage. A 
worker roughly accumulates a quarter of coverage for every \$4000 of coverage earnings up to a maximum 
of 4 per year. Not all respondents agree to have their record matched. Hence, there is the potential 
for non-representativeness. However, recent studies show that the extent of non-representativeness is 
quite small and that appropriate weighting using HRS weights mostly corrects for this problem 
\citep{kapteyn2006effects}.

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline.tex
First commit in git, migrated from a subversion repository

\section{Data sources for trends and baseline scenario}
Two types of trends need to be projected in the model. First, we need to project trends in the 
incoming cohorts (the future new age 51/52 individuals). This includes trends in health and economic 
outcomes. Second, we need to project excess aggregate growth in real income and excess growth in 
health spending.

\input{data_sources_trends_and_baseline_entering_cohorts.tex}
\input{data_sources_trends_and_baseline_other_projections.tex}
\input{data_sources_trends_and_baseline_demographic_projections.tex}

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline_demographic_projections.tex
First commit in git, migrated from a subversion repository

\subsection{Demographic adjustments}
We make two adjustments to the weighting in the HRS to match population counts. Since we deleted some 
cases from the data and only considered the set of respondents with matched Social Security records, 
this takes account of selectivity based on these characteristics. First, we post-stratify the HRS sample 
by 5 year age groups, gender and race and rebalance weights using the Census Bureau 2000-2010 Intercensal 
Population Estimates. We do this for both the host data set and the new cohorts. We scale the weights for 
future new cohorts using 2012 National Population Projections based on race and gender. Second, we post-
stratify the HRS sample of deaths between the 2002 and 2004 interview waves by 5 year age groups, gender 
and race and rebalance weights based on the Human Mortality Database.

Once the simulation begins, trends in migration and mortality are applied.  We use net migration from
the SSA Trustees report intermediate cost scenario.  Seperate mortality rate adjustment factors are 
defined for the under and over 65 age groups based on the mortality projections from the 2013 SSA
Trustees report.  The SSA projections are interpolated through 2090, then extended using GLM with log 
link through 2150.

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline_entering_cohorts.tex
First commit in git, migrated from a subversion repository

\subsection{Data for trends in entering cohorts}
\label{sec:data_sources_trends_and_baseline_entering_cohorts}
We used a multitude of data sources to compute U.S. trends. First, we used NHIS for chronic conditions 
and applied the methodology discussed in \citep{goldman2004health}. The method consists of projecting the 
experience of younger cohorts into the future until they reach age 51. The projection method is 
tailored to the synthetic cohorts observed in NHIS. For example, we observe a representative sample of 
age 35 individuals born in 1945 in 1980. We follow their disease patterns in 1980 to 1981 surveys by 
then selecting those aged 36 in 1981, accounting for mortality, etc.

We then collected information on other trends, i.e. for obesity and smoking, from 
other studies \citep{honeycutt2003dynamic,levy2006smoking,poterba2009decline,ruhm2007current,mainous2007impact}. 
Table \ref{tab:cohort_projection_data_methods} presents the sources and Table \ref{tab:baseline_trends} presents the trends we use in the baseline scenario. 
Table \ref{tab:prevalence_1978_2004} presents the prevalence of obesity, hypertension, diabetes, and current smokers in 1978 and 
2004, and the annual rates of change from 1978 to 2004.  We refer the readers to the analysis in 
\citet{goldman2004health} for information on how the trends were constructed.

### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline_other_projections.tex
First commit in git, migrated from a subversion repository

\subsection{Data for other projections}
We make two assumptions relating to real growth in wages and medical costs. Firstly, as is done in the 2009 
Social Security Trustees report intermediate cost scenario, we assume a long term real increase in wages 
(earnings) of 1.1\% per year. Next, following the Centers for Medicare and Medicaid Services, we assume 
excess real growth in medical costs (that is additional cost growth to GDP growth), as 1.5\% in 2004, 
reducing linearly to 1\% in 2033, .4\% in 2053, and -.2\% in 2083. We also include the Affordable Care 
Act cost growth targets as an optional cap on medical cost growth. Baseline medical spending figures 
presented assume those targets are met. GDP growth in the near term (through 2019) is based on CBO 
projections, with the OASDI Trustees assumption of 2\% yearly afterwards.

### 2019-11-22 analysis/techdoc/latex/FEM/dynamic_model_comparison_with_other_microsims.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/dynamic_model_comparison_with_other_microsims.tex
@@
-\subsection{Comparison with other microsimulation models of health expenditures}
-The precursor to the FAM, the FEM, was unique among models that make health expenditure projections.  It was the only 
-model that projected health trends rather than health expenditures.  It was also unique in generating mortality projections
-based on assumptions about health trends rather than historical time series.
-
-FAM extends FEM to younger ages, adding additional dimensions to the simulation.  Events over the life course, such as marital status
-and childbearing are simulated.  Labor force participation is modeled in greater detail, distinguishing between out-of-labor force, unemployed, 
-working part-time, and working full-time.
+\subsection{Comparison with other prominent microsimulation models of health expenditures}
+The FEM is unique among existing models that make health expenditure projections. It is the only model 
+that projects health trends rather than health expenditures. It is also the only model that generates 
+mortality out of assumptions on health trends rather than historical time series.
 
@@
 will be 5\%).
-
-\subsubsection{MINT Model}
-Modeling Income in the Near Term (MINT) is a microsimulation model developed by the Urban Institute and others 
-for the Social Security Administration to enable policy analysis of proposed changes to Social Security benefits and payroll
-taxes \citet{smith2013mint}.  MINT uses the Survey of Income and Program Participation (SIPP) as the base data and simulates a range of outcomes, 
-with a focus on those that will impact Social Security.  Recent extensions have included health insurance coverage and
-out-of-pocket medical expenditures.  Health enters MINT via self-reported health status and self-reported work limitations.  MINT
-simulates marital status and fertility.

### 2019-11-22 analysis/techdoc/latex/FEM/dynamic_model_overview.tex
First commit in git, migrated from a subversion repository

\subsection{Overview}

The defining characteristic of the model is the modeling of real rather than synthetic cohorts, all of whom are followed at the individual level. This allows for more heterogeneity in behavior than would be allowed by a cell-based approach. Also, since the HRS interviews both respondent and spouse, we can link records to calculate household-level outcomes such as net income and Social Security retirement benefits, which depend on the outcomes of both spouses. The omission of the population younger than age 51 sacrifices little generality, since the bulk of expenditure on the public programs we consider occurs after age 50. However, we may fail to capture behavioral responses among the young.

The model has three core components: 
\begin{itemize}
\item The initial cohort module predicts the economic and health outcomes of new cohorts of 51/52 
year-olds. This module takes in data from the Health and Retirement Study (HRS) and trends calculated 
from other sources. It allows us to ``generate'' cohorts as the simulation proceeds, so that we can 
measure outcomes for the age 51+ population in any given year. 
\item The transition module calculates the probabilities of transiting across various health states 
and financial outcomes. The module takes as inputs risk factors such as smoking, weight, age and 
education, along with lagged health and financial states. This allows for a great deal of 
heterogeneity and fairly general feedback effects. The transition probabilities are estimated from 
the longitudinal data in the Health and Retirement Study (HRS). 
\item The policy outcomes module aggregates projections of individual-level outcomes into policy 
outcomes such as taxes, medical care costs, pension benefits paid, and disability benefits. This 
component takes account of public and private program rules to the extent allowed by the available 
outcomes. Because we have access to HRS-linked restricted data from Social Security records and 
employer pension plans, we are able to realistically model retirement benefit receipt. 
\end{itemize}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{./img/fem_architecture.png}
\caption{Architecture of the FEM}
\label{fig:fem_architecture} 
\end{figure}

Figure \ref{fig:fem_architecture} provides a schematic overview of the model. We start in 2004 with 
an initial population aged 51+ taken from the HRS. We then predict outcomes using our estimated 
transition probabilities (see section \ref{sec:estimation_transition_model}). Those who survive make it to the end of that year, at 
which point we calculate policy outcomes for the year. We then move to the following time period 
(two years later), when a new cohort of 51 and 52 year-olds enters (see section \ref{sec:new_cohorts_model_empirical_strategy}). This entrance 
forms the new age 51+ population, which then proceeds through the transition model as before. This 
process is repeated until we reach the final year of the simulation.

### 2019-11-22 analysis/techdoc/latex/FEM/estimation_goodness-of-fit.tex
First commit in git, migrated from a subversion repository

\subsection{Goodness-of-fit}
To judge the goodness-of-fit of the model, we estimated parameters on the 1998-2008 estimation sample and simulated outcomes of 1998 HRS 
respondents up to 2008. We then compared simulated and actual outcomes in 1998, 2004 and 2008. Table \ref{tab:validation_waves_4_7_9} presents the results. Some 
differences exist but in general the fit is satisfactory.

%\input{../tables/FEM/validation_waves_4_7_9.tex}
%\hline
%\caption{Simulated and actual outcomes in 1998, 2004, and 2008}
%\label{tab:validation_waves_4_7_9}
%\end{longtable}

### 2019-11-22 analysis/techdoc/latex/FEM/estimation_qalys.tex
First commit in git, migrated from a subversion repository

\subsection{Quality adjusted life years}
\label{sec:estimation_qalys}

As an alternative measure of life expectancy, we compute a quality adjusted life 
year (QALY) based on the EQ-5D instrument, a widely-used health-related 
quality-of-life (HRQoL) measure\footnote{Section 
\ref{sec:model_development_qalys_hrqol} gives some background on HRQoL 
measures.}. The scoring system for EQ-5D was first developed by 
\citet{dolan1997modeling} using a UK sample. Later, a scoring system based on 
a US sample was generated \citep{shaw2005us}. The HRS does not ask the 
appropriate questions for computing EQ-5D, but the MEPS does.  
We use a crosswalk from MEPS to compute EQ-5D scores for HRS respondents not 
living in a nursing home\footnote{Section \ref{sec:model_development_qalys_eq5d} 
describes EQ-5D in MEPS. Details of the crosswalk model development are given 
in \ref{sec:model_development_qalys_crosswalk}.}.

The FEM has a more limited specification of functional status than what is available in the HRS. In order to predict 
HRQoL for the FEM simulation sample, we needed to build a bridge between the FEM-type functional status and the 
predicted EQ-5D score in HRS.  We used ordinary least squares to model the EQ-5D score predicted for non-nursing 
home in the 1998 HRS as a function of the six chronic conditions and the FEM-specification of functional status, 
The results are shown in Table \ref{tab:fem_eq5d_model_est}.

The EQ-5D scoring method is based on a community population.  Following a 
suggestion by Emmett Keeler, if a person is living in a nursing home, the QALY is 
reduced by 10\%.  We used the parameter estimates in Table 
\ref{tab:fem_eq5d_model_est} to predict EQ-5D scores for the entire FEM simulation sample 
and reduced nursing home residents' score by 10\%.  The resulting scores are representative of 
the U.S population (both in community and in nursing homes) ages 51 and over.  
Table \ref{tab:fem_stock_predicted_eq5d} summarizes the EQ-5D score using this 
model for the stock FEM simulation sample in 2004.

### 2019-11-22 analysis/techdoc/latex/FEM/estimation_transition_model.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/estimation_transition_model.tex
@@
 \label{sec:estimation_transition_model} 
-We consider a large set of outcomes for which we model transitions. Table 
-\ref{tab:transitioned_outcomes} gives the set of outcomes considered for the 
-transition model along with descriptive statistics and the population at risk 
-when estimating the relationships.
+We consider a large set of outcomes for which we model transitions. Table \ref{tab:transitioned_outcomes} gives the set of outcomes considered for the transition model 
+along with descriptive statistics and the population at risk when estimating the relationships.
 
-Since we have a stock sample from the age 25+ population, each respondent goes 
-through an individual-specific series of intervals. Hence, we have an 
-unbalanced panel over the age range starting from 25 years old. Denote by 
-$j_{i0}$ the first age at which respondent $i$ is observed and $j_{iT_i}$ the 
-last age when he is observed. Hence we observe outcomes at ages 
-$j_i = j_{i0},\ldots,j_{iT_i}$.
+Since we have a stock sample from the age 51+ population, each respondent goes through an individual-specific series of intervals. Hence, 
+we have an unbalanced panel over the age range starting from 51 years old. Denote by $j_{i0}$ the first age at which respondent $i$ is 
+observed and $j_{iT_i}$ the last age when he is observed. Hence we observe outcomes at ages $j_i = j_{i0},\ldots,j_{iT_i}$.
 
@@
 $m$) that are determined before the first age in which each individual is observed
-%\footnote{Section \ref{sec:model_development_transition_model} explains why the $h_{i,j_0,-m}$ terms are included.}. 
+\footnote{Section \ref{sec:model_development_transition_model} explains why the $h_{i,j_0,-m}$ terms are included.}. 
 The time-varying part is the effect of previously diagnosed outcomes $h_{i,j_i-1,-m}$, 
@@
 We assume that this last shock is normally distributed and uncorrelated across 
-diseases.  We approximate $a_{m,j_i}$ with an age spline with knots at ages 35, 45,
-55, 65, and 75.
-This simplification is made 
+diseases.  We approximate $a_{m,j_i}$ with an age spline. After several specification checks, 
+knots at age 65 and 75 appear to provide the best fit. This simplification is made 
 for computational reasons since the joint estimation with unrestricted age fixed 
@@
 The occurrence of mortality censors observation of other 
-outcomes in a current year. 
-% Mortality is recorded from exit interviews.
+outcomes in a current year. Mortality is recorded from exit interviews.
 
@@
 
-We have five other types of outcomes:
+We have three other types of outcomes:
 \begin{enumerate}
-\item First, we have binary outcomes which are not an absorbing state, such as starting smoking. We specify latent 
+\item First, we have binary outcomes which are not an absorbing state, such as living in a nursing home. We specify latent 
 indices as in (\ref{eqn:transition_hzd_latent}) for these outcomes as well but where 
@@
 
-\item The third type of outcomes we consider are censored outcomes, such as financial wealth. For wealth, 
-there are a non-negligible number of observations with zero and negative wealth. For these, we consider 
-two part models where the latent variable is specified as in (\ref{eqn:transition_hzd_latent}) but model 
-probabilities only when censoring does not occur. In total, we have $M$ outcomes.
-
-\item The fourth type of outcomes are continuous outcomes modeled with ordinary least squares.  For example,
-we model transitions in log(BMI).  We allow for state-dependence by including the lagged outcome on the right-hand side.
-
-\item The final type of models are categorical, but without an ordering.  For example, an individual can transition
-to being out of the labor force, unemployed, or working (either full- or part-time).  In situations like this, we utilize
-a multinomial logit model, including the lagged outcome on the right-hand side.
-
+\item The third type of outcomes we consider are censored outcomes, earnings and financial wealth. Earnings are only observed when individuals 
+work. For wealth, there are a non-negligible number of observations with zero and negative wealth. For these, we consider two part models 
+where the latent variable is specified as in (\ref{eqn:transition_hzd_latent}) but model probabilities only when censoring does not occur. In 
+total, we have $M$ outcomes.
 \end{enumerate}
@@
 
-\subsubsection{Further Details on Specific Transition Models}
-This section describes the modeling strategy for particular outcomes.
-
-\paragraph{Employment Status}
-Ultimately, we wish to simulate if an individual is out of the labor force, unemployed, working part-time, or working full-time at 
-time $t$.  We treat the estimation of this as a two-stage process.  In the first stage, we predict if the individual is out of 
-the labor force, unemployed, or working for pay using a multinomial logit model.  Then, conditional on working for pay, we estimate if 
-the individual is working part- or full-time using a probit model.
-
-\paragraph{Earnings}
-We estimate last calendar year earnings models based on the current employment status, controlling for the prior employment status.  
-Of particular concern are individuals with no earnings, representing approximately twenty-five percent of the unemployed and 
-seventy-eight percent of those out of the labor force.  This group is less than 0.5\% of the full- and part-time populations. We 
-use a two-stage process for those out of the labor force and unemployed.  The first stage is a probit that estimates if the 
-individual has any earnings.  The second stage is an OLS model of log(earnings) for those with non-zero earnings.  For 
-those working full- or part-time, we estimate OLS models of log(earnings).
-
-\paragraph{Relationship Status}
-We are interested in three relationship statuses: single, cohabitating, and married.  In each case, we treat the transition
-from time $t$ to time $t+1$ as a two-stage process.  In the first stage, we estimate if the individual will remain in their 
-current status.  In the second stage, we estimate which of the two other states the individual will transition to, conditional
-on leaving their current state.
-
-\paragraph{Childbearing}
-We estimate the number of children born in two-years separately for women and men.  We model this using an ordered probit with
-three categories: no new births, one birth, and two births.  Based on the PSID data, we found the exclusion of three or more
-births in a two-year period to be appropriate.
-
 \subsubsection{Inverse Hyperbolic Sine Transformation}
-One problem fitting the wealth distribution is that it has a long right tail and some negative values. We use a 
+One problem fitting the wealth and earnings distribution is that they have a long right tail and wealth has some negative values. We use a 
 generalization of the inverse hyperbolic sine transform (IHT) presented in \citet{mackinnon1990transforming}. First denote the variable of 
@@
 % \todo link to transition_estimates.xls on box -- make link unique to the version of the appendix in make process
-The included estimates table (estimates\_FAM.xml) gives parameter estimates for the transition models.
+%The estimates table\footnote{\input{estimates_table_link.tex}} gives parameter estimates for the transition models.
+
+%\newpage
+%\input{../tables/FEM/health_transition_est.tex}
+%\caption{Coefficient estimates and $t$-statistics for mortality and chronic condition transitions}
+%\label{tab:health_transition_est}
+%\end{longtable}
+%
+%\newpage
+%\input{../tables/FEM/econ_transition_est.tex}
+%\caption{Coefficient estimates and $t$-statistics for economic outcome transitions}
+%\label{tab:econ_transition_est}
+%\end{longtable}
+%
+%\newpage
+%\input{../tables/FEM/BMI_transition_est.tex}
+%\caption{Coefficient estimates and $t$-statistics for BMI outcome transition}
+%\label{tab:BMI_transition_est}
+%\end{longtable}
+%
+%\newpage
+%\input{../tables/FEM/ordered_transition_est.tex}
+%\caption{Coefficient estimates and $t$-statistics for ordered outcome transitions}
+%\label{tab:ordered_transition_est}
+%\end{longtable}
+%
+%\newpage
+%\input{../tables/FEM/IHS_transition_est.tex}
+%\caption{Coefficient estimates and $t$-statistics for transitioning outcomes with shape-preserving Inverse Hyperbolic Sine transformation}
+%\label{tab:IHS_transition_est}
+%\end{longtable}

### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures.tex
First commit in git, migrated from a subversion repository

\section{Government revenues and expenditures}
This gives a limited overview of how revenues and expenditures of the government are computed. These 
functions are based on 2004 rules, but we include predicted changes in program rules such changes 
based on year of birth (e.g. Normal retirement age).

We cover the following revenues and expenditures:\\

\begin{center}
\begin{tabular}{l l}
\textbf{Revenues} & \textbf{Expenditures}\\
Federal Income Tax & Social Security Retirement benefits\\
State and City Income Taxes & Social Security Disability benefits\\
Social Security Payroll Tax & Supplementary Security Income (SSI)\\
Medicare Payroll Tax & Medical Care Costs\\
Property Tax & Medicaid\\
 & Medicare (parts A, B, and D)
\end{tabular}
\end{center}

\input{govt_revenue_and_expenditures_SSben.tex}
\input{govt_revenue_and_expenditures_DIben.tex}
\input{govt_revenue_and_expenditures_SSIben.tex}
\input{govt_revenue_and_expenditures_medcost_estimation.tex}
\input{govt_revenue_and_expenditures_taxes.tex}

### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures_SSIben.tex
First commit in git, migrated from a subversion repository

\subsection{Supplemental Security Income benefits}
Self-reported receipt of supplemental security income (SSI) in the HRS provides estimates of the proportion of people receiving SSI under 
what administrative data would suggest. To correct for this bias, we link the HRS with administrative data from the social security 
administration identifying those receiving SSI. In the linked administrative data, 3.96\% of the population receives supplementary security 
income, while only 2.79\% of the sample reports social security income. We therefore estimate a probit of receiving SSI as a function of 
self-reporting social security income, as well as demographic, health, and wealth. 
%Model estimates can be found in the estimates table \footnote{\input{estimates_table_link.tex}}.

The benefit amount is taken from the average monthly benefits found in the 2004 Social Security Annual Statistical Supplement. We assign 
monthly benefit of \$450 for person aged 51 to 64, and \$350 for persons aged 65 and older.

### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures_SSben.tex
First commit in git, migrated from a subversion repository

\subsection{Social Security benefits}
Workers with 40 quarters of coverage and of age 62 are eligible to receive their retirement benefit. The benefit is calculated based on the 
Average Indexed Monthly Earnings (AIME) and the age at which benefits are first received. If an individual claims at his normal retirement 
age (NRA) (65 for those born prior to 1943, 66 for those between 1943 and 1957, and 67 thereafter), he receives his Primary Insurance 
Amount (PIA) as a monthly benefit. The PIA is a piece-wise linear function of the AIME. If a worker claims prior to his NRA, his benefit is 
lower than his PIA. If he retires after the NRA, his benefit is higher. While receiving benefits, earnings are taxed above a certain 
earning disregard level prior to the NRA. An individual is eligible to half of his spouse’s PIA, properly adjusted for the claiming age, 
if that is higher than his/her own retirement benefit. A surviving spouse is eligible to the deceased spouse’s PIA. Since we assume prices 
are constant in our simulations, we do not adjust benefits for the COLA (Cost of Living Adjustment) which usually follows inflation. We 
however adjust the PIA bend points for increases in real wages.

### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures_medcost_estimation.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/govt_revenue_and_expenditures_medcost_estimation.tex
@@
 \label{sec:govt_revenue_and_expenditures_medcost_estimation}
-In the FAM, a cost module links a person's current state--demographics, economic status, current health, risk factors, and functional status 
-to 4 types of individual medical spending. The FAM models: total medical spending (medical spending from all payment sources), Medicare 
+In the FEM, a cost module links a person's current state--demographics, economic status, current health, risk factors, and functional status 
+to 4 types of individual medical spending. The FEM models: total medical spending (medical spending from all payment sources), Medicare 
 spending\footnote{We estimate annual medical spending paid by specific parts of 
@@
 conditions, and functional status, with spending inflated to constant dollars 
-using the medical component of the consumer price index.  We use the 2007-2010 
+using the medical component of the consumer price index.  We use the 2000-2010 
 Medical Expenditure Panel Survey 
@@
 for these regressions for persons 
-not Medicare eligible, and the 2007-2010 Medicare Current Beneficiary Survey 
+not Medicare eligible, and the 2000-2010 Medicare Current Beneficiary Survey 
 % \todo : Change the text to be something like "roughly X respondents per panel, with a new panel starting entering every Y months and lasting Z months between YYYY and YYYY." That way, it is always correct and all we do is change the years of data we're looking at.
@@
 Models are estimated for total, Medicaid, out of pocket spending, and for the Medicare spending. 
+These estimates only use the MCBS dataset.
 
@@
 current enrollees without Part B. Estimates are based on weighted probit regression on various risk factors, demographic, and economic 
-conditions. The PSID starting population for the FAM does not contain information on Medicare enrollment. Therefore another model of Part B 
+conditions. The HRS starting population for the FEM does not contain information on Medicare enrollment. Therefore another model of Part B 
 enrollment for all persons eligible for Medicare is estimated via a probit, and used in the first year of simulation to assign initial Part 
@@
 
-%Since both the MEPS and MCBS are known to under-predict medical spending (see, e.g., \citeauthor{selden2008aligning}, \citeyear{selden2008aligning}, and references therein), we applied adjustment factors to the predicted three types of
-%individual medical spending so that the predicted per-capita spending in FAM equal the corresponding spending in National 
-%Health Expenditure Accounts (NHEA) for age group 55-64 in year 2004 and ages 65 and over in year 2010, respectively. Table 
-%\ref{tab:NHEA_adjustment} shows how these adjustment factors were determined by using the ratio of expenditures in the NHEA to 
-%expenditures predicted in the FAM.
+Since both the MEPS and MCBS are known to under-predict medical spending (see, e.g., \citeauthor{selden2008aligning}, \citeyear{selden2008aligning}, and references therein), we applied adjustment factors to the predicted three types of
+individual medical spending so that the predicted per-capita spending in FEM equal the corresponding spending in National 
+Health Expenditure Accounts (NHEA) for age group 55-64 in year 2004 and ages 65 and over in year 2010, respectively. Table 
+\ref{tab:NHEA_adjustment} shows how these adjustment factors were determined by using the ratio of expenditures in the NHEA to 
+expenditures predicted in the FEM.
 

### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures_taxes.tex
First commit in git, migrated from a subversion repository

\subsection{Taxes}
We consider Federal, State and City taxes paid at the household level. We also calculate Social Security taxes and Medicare taxes. HRS 
respondents are linked to their spouse in the HRS simulation. We take program rules from the OECD's Taxing Wages Publication for 2004. 
Households have basic and personal deductions based on marital status and age ($>$65). Couples are assumed to file jointly. Social Security 
benefits are partially taxed. The amount taxable increases with other income from 50\% to 85\%. Low income elderly have access to a special 
tax credit and the earned income tax credit is applied for individuals younger than 65. We calculate state and city taxes for someone 
living in Detroit, Michigan. The OECD chose this location because it is generally representative of average state and city taxes paid in 
the U.S. Since Social Security administrative data cannot be used jointly with Geocoded information in the HRS, we apply these hypothetical 
taxes to all respondents.

At the state level, there is a basic deduction for each member of the household (\$3,100) and taxable income is taxed at a flat rate of 4\%. 
At the city level, there is a small deduction of \$750 per household member and the remainder is taxed at a rate of 2.55\%. There is however 
a tax credit that decreases with income (20\% on the first 100\$ of taxes paid, 10\% on the following 50\$ and 5\% on the remaining portion).

We calculate taxes paid by the employee for Old-Age Social Insurance (SS benefits and DI) and Medicare (Medicaid and Medicare). It does not 
include the equivalent portion paid by the employer. OASI taxes of 6.2\% are levied on earnings up to \$97,500 (2004 cap) while the Medicare 
tax (1.45\%) is applied to all earnings.

% \todo : add payroll tax description here

### 2019-11-22 analysis/techdoc/latex/FEM/implementation.tex
First commit in git, migrated from a subversion repository

\section{Implementation}
The FEM is implemented in multiple parts. Estimation of the transition and cross 
sectional models is performed in Stata.
The incoming cohort model is estimated in Stata using the CMP package \citep{statacmp2011}.
The simulation is implemented in C++ to increase speed.

To match the two year structure of the Health and Retirement Study (HRS) data used to estimate the transition models, the FEM simulation 
proceeds in two year increments. The end of each two year step is designed to occur on July 1st to allow for easier matching to population 
forecasts from Social Security. A simulation of the FEM proceeds by first loading a population representative of the age 51+ US population 
in 2004, generated from HRS. In two year increments, the FEM applies the transition models for mortality, health, working, wealth, 
earnings, and benefit claiming with Monte Carlo decisions to calculate the new states of the population. The population is also adjusted by 
immigration forecasts from the US Census Department, stratified by race and age. If incoming cohorts are being used, the new 51/52 year 
olds are added to the population. The number of new 51/52 year olds added is consistent with estimates from the Census, stratified by race. 
Once the new states have been determined and new 51/52 year olds added, the cross sectional models for medical costs, and calculations for 
government expenditures and revenues are performed. Summary variables are then computed. Computation of medical costs includes the persons 
that died to account for end of life costs. Other computations, such as Social Security benefits and government tax revenues, are 
restricted to persons alive at the end of each two year interval. To eliminate uncertainty due to the Monte Carlo decision rules, the 
simulation is performed multiple times (typically 100), and the mean of each summary variable is calculated across repetitions.

FEM simulation takes as inputs assumptions regarding growth in the national wage index, normal retirement age, real medical cost growth, 
interest rates, cost of living adjustments, the consumer price index, significant gainful activity, and deferred retirement credit. The 
default assumptions are taken from the 2010 Social Security Intermediate scenario, adjusted for no price increases after 2010. Therefore 
simulation results are in real 2009 dollars.  Table \ref{tab:time_series_by_year} shows the assumptions for each calendar year and Table 
\ref{tab:time_series_by_yob} shows assumptions for each birth year.

Different simulation scenarios are implemented by changing any of the following components: incoming cohort model, transition models, 
interventions that adjust the probabilities of specific transition, and changes to assumptions on future economic conditions.

\input{./implementation_interventions.tex}

### 2019-11-22 analysis/techdoc/latex/FEM/model_development_qalys.tex
First commit in git, migrated from a subversion repository

\subsection{Quality adjusted life years}
\label{sec:model_development_qalys}

\subsubsection{Health related quality-of-life measures}
\label{sec:model_development_qalys_hrqol}
In general, HRQoL measures summarize population health by a single 
preference-based index measure. A HRQoL measure is a suitable measure of QALY. 
There are several widely-used generic HRQoL indexes, each involving a standard descriptive system: a 
multidimensional measure of health states and a corresponding scoring system to translate the descriptive 
system into a single index \citep{fryback2007us}. The scoring system is developed based on a community survey 
of preference valuation of health states in the descriptive system, using utility valuation methods like time 
trade-offs or a standard gamble.

\subsubsection{Health related quality-of-life in MEPS}
\label{sec:model_development_qalys_eq5d}
Because the health states measures in the HRS and FEM do not match the health 
states defined in any of the currently available HRQoL indexes, we used MEPS to 
create a crosswalk file for HRQoL index calculation.  MEPS collects information 
on health care cost and utilization, demographics, functional status, and medical 
conditions. Since the year 2000, it initiated a self-administered questionnaire 
for two sets of instruments: SF-12 and EQ-5D.

Seven of the twelve SF-12 questions can be used to generate another HRQoL index: 
SF-6D. However, the scoring system for SF-6D was derived from a UK sample 
\citep{brazier2004estimation} and a significant proportion of the MEPS sample 
did not give valid answer for at least one of the seven questions. Therfore, we 
decided to calculate EQ-5D index score as the HRQoL measure for FEM.

The EQ-5D instrument includes 5 questions about the extent of problems in 
mobility, self-care, daily activities, pain, and anxiety/depression. The scoring 
system for EQ-5D was first developed by \citet{dolan1997modeling} using a UK 
sample. Later, a scoring system based on a US sample was generated 
\citep{shaw2005us}. In MEPS 2001, there are 8,301 respondents aged 51 and over. 
Of those respondents, 7,439 gave valid answers for all of the five EQ-5D 
questions. We calculated EQ-5D scores for these respondents using the scoring 
algorithm based on a US sample \citep{shaw2005us}. The distribution of EQ-5D 
index scores among these respondents is shown in Figure 
\ref{fig:meps2001_age51p_eq5d_hist}.

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{./img/meps2001_age51p_eq5d_hist.png}
\caption{Distribution of EQ-5D index scores for ages 51+ in 2001 MEPS}
\label{fig:meps2001_age51p_eq5d_hist} 
\end{figure}

\subsubsection{MEPS-HRS Crosswalk development}
\label{sec:model_development_qalys_crosswalk}
The functional status measure in FEM is based on the HRS. It is a categorical 
variable including the following mutually exclusive categories: healthy, any IADL 
limitation (no ADL limitations), 1-2 ADL limitations, and 3 or more ADL limitations. 
Unfortunately the measures of IADL and ADL limitations in MEPS are quite different. 
HRS asks questions like ``Do you have any difficulty in $\ldots$'', while MEPS asks 
questions like ``Does $\ldots$help or supervision in $\ldots$.'' As Table 
\ref{tab:meps_hrs_adl_iadl_prevalence} shows, the prevalence of 
IADL limitations is relatively similar between the two surveys, while the 
prevalence of ADL limitations is much higher in HRS, relative to MEPS. This is 
reasonable since not all who have difficulty in ADLs receive help or supervision.

In order to compute EQ-5D index scores using functional status in the FEM, we needed
a set of functional status measures that is comparable across MEPS and HRS (the host 
dataset for FEM).  We explored several options for deriving such a measure.  Ultimately,
we constructed two measures. One measure indicates physical function limitation 
while the other measure indicates IADL limitation.

In MEPS, physical function 
limitation indicates that at least one of the following is true: 1) receiving help 
or supervision with bathing, dressing or walking around the house; 2) being limited 
in work/housework; 3) having difficulty walking, climbing stairs, grasping objects, 
reaching overhead, lifting, bending or stooping, or standing for long periods of time; 
or 4) having difficulty in hearing or vision. In HRS, physical function limitation indicates 
that at least one of the following is 
true: 1) having any difficulty in bathing/dressing/eating/walking across the room/getting 
out of bed; 2) limited in work/housework; or 3) limited in any other activities.

In MEPS, IADL limitation indicates receiving help or supervision using the telephone, 
paying bills, taking medications, preparing light meals, doing laundry, or going shopping.
In HRS, IADL limitation indicates having difficulty in any IADL such as using the phone, managing 
money, or taking medications.

The prevalence of our two constructed measures among ages 51 and older in MEPS (2001) and HRS (1998) is 
shown in Table \ref{tab:meps_hrs_physlim_iadl_prevalence}. The prevalences are quite similar across the two surveys.

Using MEPS 2001 data, we next use ordinary least squares to model the derived EQ-5D score as a function of six 
chronic conditions -- which are available both in HRS and MEPS, our two constructed measures of functional status, 
and an interaction term of the two measures of functional status. Three different models were considered. 
Estimation results are presented in Models I-III in Table \ref{tab:meps_eq5d_model_est}. We also show the estimation results of using only 
IADL/ADL limitation as covariates, and using only the six chronic conditions as covariates, as Models IV and V in 
Table \ref{tab:meps_eq5d_model_est}. Model II was used as the crosswalk described 
in Section \ref{sec:estimation_qalys} to calculate EQ-5D score for non-nursing 
home residents aged 51 and over in HRS 1998.

### 2019-11-22 analysis/techdoc/latex/FEM/model_development_rxexp.tex
First commit in git, migrated from a subversion repository

\subsection{Drug Expenditures}
\label{sec:model_development_rxexp}

\subsubsection{Drug Expenditures - MEPS}
\label{sec:model_development_rxexp_meps}
AHRQ produces a file of consolidated annual expenditures for each Medical Expenditure Panel Survey respondent in each calendar year.  
The total drug expenditure variable sums all amounts paid out-ot-pocket and by third party payers for each prescription
purchased in that year.  For comparison across years, we convert all amounts to 2010 dollars using the Medical CPI.

\subsubsection{Drug Expenditures - MCBS}
\label{sec:model_development_rxexp_mcbs}
The Medicare Current Beneficiary Survey produces a Prescribed Medicine Events file at the individual-event level, with cost
and utilization of prescribed medicines for the MCBS community population.  Collapsing this file to the individual provides an
estimate of prescription drug cost for each person.  For comparison across years, we convert all amounts to 2010 dollars using the Medical CPI.

There are two caveats to working with these data.  The first caveat regards how to handle the "ghost" respondents.  
"Ghosts" are individuals who enroll in Medicare, but were not
asked cost and use questions in the year of their enrollment.  For some outcomes, such as medical expenditures, the MCBS makes an 
effort to impute.  For others, such as drug utilization and expenditures, the MCBS does not.  We imputed annual drug expenditures 
for the ghosts, but for certain age ranges the drug expenditures were not reasonable.  This had the biggest effect on the 65 and 66 
year olds, for two reasons.  The first is that the 65 and 66 year olds are more likely to be ghosts, as 65 is the typical age of 
enrollment for Medicare.  The second is that the 65 and 66 year olds used for imputation (i.e., the non-"ghosts") are different.  To 
be fully present in MCBS at age 65 would require enrolling in Medicare before age 65, which happen through a different channel, such
as qualifying for Medicare due to receiving disability benefits from the federal government.

The second caveat relates to the filling in zeroes for individuals with no utilization data, but who were enrolled.  We assumed that
individuals who were not ghosts and who did not appear on the Prescribed Medicine Events file had zero prescription expenditures.

\subsubsection{Drug Expenditures - Estimation}
\label{sec:model_development_rxexp_estimation}
Due to the complexities of the age 65-66 population in the MCBS, we chose to estimate the drug expenditure models using the MEPS for 
individuals 51 to 66 and the MCBS for individuals 67 and older.  Individuals under age 65 receiving Medicare due to disability are 
estimated separately.  Since there are a number of individuals with zero expenditures, we estimate the models in two stages.  The 
first stage is a probit predicting any drug expenditures and the second is an ordinary least squares model predicting the amount, 
conditional on any.  Coefficient estimates and marginal effects are shown in the accompanying Excel workbook.

### 2019-11-22 analysis/techdoc/latex/FEM/model_development_transition_model.tex
First commit in git, migrated from a subversion repository

\subsection{Transition model}
\label{sec:model_development_transition_model}

Section \ref{sec:estimation_transition_model} describes the current FEM transition
model with a focus on discrete absorbing outcomes. In developing this model, 
it was previously assumed that the time invariant part of the hazard was composed 
of the effect of observed characteristics $x_i$ and permanent unobserved characteristics 
specific to outcome $m$, $\eta_{i,m}$.  Consequently, the index was assumed to be 
of the form 
$z_{m,j_i} = x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + \eta_{i,m}$ 
and the latent component of the hazard was modeled as
\begin{equation}
h^*_{i,j_i,m}= x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + \eta_{i,m} + a_{m,j_i} + \varepsilon_{i,j_i,m},
\label{eqn:old_transition_hzd_latent}
\end{equation}
\[
m = 1,\ldots,M_0 \mbox{, } j_i = j_{i0},\ldots,j_{i,T_i} \mbox{, } i=1,\ldots,N
\]
This is the same as (\ref{eqn:transition_hzd_latent}), except that 
(\ref{eqn:old_transition_hzd_latent}) uses unobserved characteristics $\eta_{i,m}$
instead of the effects of observed initial conditions $h_{i,j_0,-m}\psi_m$. The 
unobserved effects $\eta_{i,m}$ are persistent over time and were 
allowed to be correlated across diseases $m=1,\ldots,M$.  We assumed that these 
effects had a normal distribution with covariance matrix $\mathbf{\Omega}_\eta$.

The parameters 
$\mathbf{\theta}_1 = \left(\left\{\beta_m, \gamma_m, \varsigma_m\right\}_{m=1}^M, \mathrm{vech}(\mathbf{\Omega}_\eta)\right)$, 
could be estimated by maximum simulated likelihood. The joint probability, 
conditional on the individual frailty is the product of normal univariate 
probabilities. Similar to the joint probability in Section 
\ref{sec:estimation_transition_model}, these sequences, conditional on unobserved 
heterogeneity, are also independent across diseases.  The joint probability over 
all disease-specific sequences is simply the product of those probabilities.

For a given respondent with frailty $\eta_i$, the probability of the observed 
health history is (again, omitting the conditioning on covariates for simplicity)
\[
	l^{-0}_i(\mathbf{\theta}; \eta_i, h_{i,j_{i0}}) = \left[\prod_{m=1}^{M-1} \prod_{j=j_{i1}}^{j_{T_i}} P_{ij,m}(\mathbf{\theta}; \eta_i)^{(1-h_{ij-1,m})(1-h_{ij,M})} \right] \times \left[\prod_{j=j_{i1}}^{j_{T_i}} P_{ij,M}(\mathbf{\theta}; \eta_i) \right]
\]

To obtain the likelihood of the parameters given the observables, it is necessary 
to integrate out unobserved heterogeneity. The complication is that 
$h_{i,j_{i0},-m}$, the initial outcomes in each hazard, are not likely to be 
independent of the common unobserved heterogeneity term which needs to be integrated 
out.  A solution is to model the conditional probability distribution 
$p(\eta_i \mid \mathbf{h}_{i,j_{i0}})$ \citep{wooldridge2000framework}. Implementing 
this solution amounts to including initial outcomes at baseline (age 50) for each 
hazard. This is equivalent to writing 
\begin{align*}
&\eta_i = \Gamma h_{i0} + \alpha_i \\
&\alpha_i \sim \mathrm{N}(0, \Omega_\alpha)
\end{align*}
Therefore, this allows for permanent differences in outcomes due to differences in 
baseline outcomes. The likelihood contribution for one respondent's sequence is 
therefore given by
\begin{equation}
l_i(\mathbf{\theta}, \mathbf{h}_{i,j_{i0}}) = \int l_i(\mathbf{\theta}; \alpha_i, \mathbf{h}_{i,j_{i0}})dF(\alpha_i)
\label{eqn:likelihood_contribution}
\end{equation}

This model was estimated using maximum simulated likelihood. The likelihood
contribution (\ref{eqn:likelihood_contribution}) was replaced with a simulated 
counterpart based on $R$ draws from the distribution of $\alpha$. The BFGS
algorithm was then used to optimize over this simulated likelihood. Convergence of
the joint estimator could not be obtained, so the distribution of 
$\alpha_i$ was assumed to be degenerate. This yielded the simpler 
estimation problem describe in Section \ref{sec:estimation_transition_model}, 
where each equation is estimated separately.

### 2019-11-22 analysis/techdoc/latex/FEM/new_cohorts_model_empirical_strategy.tex
First commit in git, migrated from a subversion repository

\subsection{Information available and empirical strategy}
\label{sec:new_cohorts_model_empirical_strategy}
For the transition model, we need to first to obtain outcomes listed in Table \ref{tab:desc_init_conditions}. Ideally, we need 
information on 
\[
f_t(y_{i1},\ldots,y_{iM}) = f_t(\mathbf{y}_i)
\]
where $t$ denotes calendar time, and $\mathbf{y}_i = (y_{i1},\ldots,y_{iM})$ is a vector of outcomes 
of interest whose probability distribution at time $t$ is $f_t()$. Information on how the joint 
distribution evolves over time is not available. Trends in conditional distributions are rarely 
reported either.

Generally, we have (from published or unpublished sources) good information on trends for some moments 
of each outcome (say a mean or a fraction). That is, we have information on 
$g_{t,m}(y_{im})$,
where $g_{t,m}()$ denotes the marginal probability distribution of outcome $m$ at time $t$.

For example, we know from the NHIS repeated cross-sections that the fraction obese is increasing by 
roughly 2\% a year among 51 year olds. In statistical jargon this means we have information on how the 
mean of the marginal distribution of $y_{im}$, an indicator variable that denotes whether someone is obese, 
is evolving over time.

We also have information on the joint distribution at one point in time, say year $t_0$. For example,
 we can estimate the joint distribution on age 51 respondents in the 1992 wave of the HRS, $f_{t_0}(\mathbf{y}_i)$.

We make the assumption that only some part of $f_t(\mathbf{y}_i)$ evolves over time. In particular, 
we will model the marginal distribution of each outcome allowing for correlation across these 
marginals. The correlations will be assumed fixed while the mean of the marginals will be allowed to 
change over time.

### 2019-11-22 analysis/techdoc/latex/FEM/new_cohorts_model_model_and_estimation.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model_model_and_estimation.tex
@@
 
-%To apply trends to the future cohorts, the latent model is written as
-%\[
-%\mathbf{y}^*_i = \mathbf{\mu} + \mathbf{L}_\Omega \mathbf{\eta}_i.
-%\]
-%Each marginal has a mean change equal to $\mathrm{E}(\mathbf{y} \mid \mathbf{\mu}) = (1+\tau)g(\mathbf{\mu})$, where $\tau$ is the 
-%percent change in the outcome and $g()$ is a non-linear but monotone mapping. Since 
-%it is invertible, we can find the vector $\mathbf{\mu}^*$ where $\mathbf{\mu}^* = g^{-1}(\mathrm{E}(y \mid \mu)/(1+\tau))$. We use these new intercepts to simulate new outcomes.
-
+To apply trends to the future cohorts, the latent model is written as
+\[
+\mathbf{y}^*_i = \mathbf{\mu} + \mathbf{L}_\Omega \mathbf{\eta}_i.
+\]
+Each marginal has a mean change equal to $\mathrm{E}(\mathbf{y} \mid \mathbf{\mu}) = (1+\tau)g(\mathbf{\mu})$, where $\tau$ is the 
+percent change in the outcome and $g()$ is a non-linear but monotone mapping. Since 
+it is invertible, we can find the vector $\mathbf{\mu}^*$ where $\mathbf{\mu}^* = g^{-1}(\mathrm{E}(y \mid \mu)/(1+\tau))$. We use these new intercepts to simulate new outcomes. 
 % \todo : add an  example here for clarification

### 2019-11-22 analysis/techdoc/latex/FEM/scenarios_and_robustness_obesity_reduction.tex
First commit in git, migrated from a subversion repository

\subsection{Obesity reduction scenario}
In addition the to the status quo scenario, the Future Elderly Model can be 
used to estimate the effects of numerous possible policy changes. One such set 
of policy simulations involves changing the trends of risk factors for chronic 
conditions. This is implemented by altering the incoming cohorts. A useful 
example is an obesity reduction scenario which rolls back the prevalence of 
obesity among 50 year-olds to its 1978 level by 2030, where it remains until 
the end of the scenario, in 2050. This is accomplished by reversing the annual
rates of change for BMI category, hypertension, and diabetes shown in Table 
\ref{tab:prevalence_1978_2004}.  As seen in Table \ref{tab:obesity_results}, 
this will change the prevalence of obesity among the age 50+ in 2050. As 
compared with the status quo estimates (Table \ref{tab:status_quo_results}) the 
FEM predicts that by 2050, this will result in a change in the amount of Social 
Security benefits as well as changing combined Medicare and Medicaid expenditures.

### 2019-11-22 analysis/techdoc/latex/FEM/scenarios_and_robustness_robustness.tex
First commit in git, migrated from a subversion repository

\subsection{Robustness}
\label{sec:robustness}
A number of the restrictions on health transitions prove to be statistically significant in the health transitions model. 
%(see the table of estimates \input{estimates_table_link.tex})
Thus, we test the robustness of the FEM by analyzing the policy simulations with and without 
some of the restrictions. We test the robustness in two steps. First, we look at the effect of removing the restrictions on health outcomes 
affecting other health transitions, and compare the simulation results both for the status quo model, and for the obesity scenario discussed 
above. The model without health restrictions has less than 1\% difference in predictions relating to government expenditures in both 2030 
and 2050 (as shown in Table 28). There are slightly higher, yet still quite small increases in simulated government revenues.  Importantly, Table 29, which shows the effect 
of the obesity scenario without the health restrictions, finds very similar effects on social security and Medicare/Medicaid as under the 
FEM (2.35\% as compared with 2.28\% increase in social security benefits and a 4.31\% as compared with 4.37\% decrease in combined 
Medicare/Medicaid expenditures).

We similarly test the effect of the economic restrictions on health transitions. The economic restrictions have a larger impact on 
projected expenditures and revenues than did the health restrictions (see Table 30). However, they still have little effect on the 
conclusions drawn from the obesity scenario. The obesity scenario, in Table 31, where we remove the economic restrictions, implies the same 
2.28\% increase in social security expenditures, and a very similar 4.39\% drop in Medicaid/Medicare expenditures in 2050. Thus, while the 
imposed restrictions have some, often small, impact on the FEM baseline estimates, they have little effect on the conclusions from the 
obesity reduction scenario.

### 2019-11-22 analysis/techdoc/latex/FEM/uncertainty.tex
First commit in git, migrated from a subversion repository

\section{Uncertainty}
Validation of population-based simulation models includes uncertainty analysis (Kopec 2010). Uncertainty analysis 
is the "quantification of the differences between a model's estimates and the truth." (Citro and Hanushek 1991). Sources of 
uncertainty common in population-based microsimulation models include sampling variability from input databases, sampling 
variability from other inputs, model misspecification, and stochastic error (Citro and Hanushek 1991). Stochastic and sampling 
uncertainty can be evaluated using confidence intervals around the output point estimates.

\subsection{Confidence intervals}
The FEM produces output parameters for each repetition of the simulation. A confidence interval can be calculated from the sample 
of output parameters when the input parameters are varied with each repetition. Confidence intervals from the deterministic simulation 
only take into account the intrinsic error from the Monte Carlo stochastic element of the simulation. Repeating the deterministic estimation 
for 50 to 100 replicates is considered to reduce the Monte Carlo error to nearly zero. Stochastic uncertainty alone does not adequately 
account for uncertainty in microsimulation results.

\subsection{Bootstraping}
Sampling variability from the survey sources is considered, in the FEM, under the conditions of a resampled survey. Direct 
bootstrapping is used as a conservative technique to account for the other sources of uncertainty. The data is randomly sampled with replacement 
taking into account survey design (Yeo, H and Liu 1999). The HRS is sampled at the household level with 56 strata in recent waves. 
Observations are selected for their full panel of interviews. Each input is produced for each sample of the surveys. Bootstrapping of 
the inputs implicitly takes into account the correlation between the parameters used in the FEM, particularly from the HRS. We use dependent 
bootstrap samples for all sources of sampling uncertainty from the HRS: input database, transition parameters, and trends.

The number of bootstrap samples is started at 1000 based on what is found in the literature, but the total number used depends on the outcome 
and models being assessed. Watching the variance of the output parameter to see where it settles down offers an estimate of the number of 
repetitions needed. Each rep then has a mean and non-parametric distribution from which the 2.5 and 97.5 percentiles can be used as confidence 
intervals. The mean is calculated as the mean over all observations (Dong and Nakayama 2014). The confidence intervals are calculated as the 
mean of the upper and lower percentiles. The contributions of each source of uncertainty discussed in more detail below can be estimated by 
independently combining deterministic and sampled inputs when running the simulation.

\subsubsection{Sample uncertainty}
In the FEM, the major sources of input uncertainty come from sampling variability. The FEM uses a single wave from the HRS as the population that 
starts the simulation. In order to incorporate the sampling uncertainty in the FEm, bootstrap samples are taken from the population prior to 
population reweighting. Each bootstrap sample is reweighted to the population  before entering the simulation.

\subsubsection{Trends}
Trends are used to alter the health and economic characteristics of future cohorts before they enter the simulation are estimated from survey data or 
pulled from the literature. Because distributions of the trend parameters in each year are not available from all sources, we use 
bootstrap samples of each cohort after applying the baseline trends in order to achieve variation in the trended parameters.

\subsubsection{Transition parameter uncertainty}
Transition probabilities in the FEM are predicted for each active individual in each wave of the simulation using regression models 
estimated from the HRS. Regression models for transition probabilities are estimated using the bootstrap samples of the HRS which 
implicitly allow for correlation of covariates within models and between models. A set of transition regression models is estimated 
for each bootstrap sample.

\subsubsection{Models that predict policy outcomes}
The regression models that predict cost, utilization, and other outcomes are based on MCBS and MEPS. These data are also subject to 
sampling error and are sampled in a similar way to the HRS. The MCBS is sampled at the individual level with 100 strata. The MEPS 
is sampled at the dwelling unit level with 165 strata in recent surveys. The MEPS and MCBS 
bootsrap weights are used in the estimation of the medical spending and utilization regression models.

### 2019-11-22 analysis/techdoc/latex/FEM/validation.tex
First commit in git, migrated from a subversion repository

\section{Validation}
We perform three validation exercises:

\begin{enumerate}
\item Cross-validation
\item External validation
\item External corroboration
\end{enumerate}

Cross-validation is a test of the simulation’s internal validity that compares simulated outcomes to actual outcomes, external validation compares model forecasts 
with actual outcomes from other data sources, and external corroboration compares model forecasts to others’ forecasts.

\input{validation_crossvalidation.tex}
\input{validation_external_validation.tex}
\input{validation_external_corroboration.tex}

### 2019-11-22 analysis/techdoc/latex/FEM/validation_crossvalidation.tex
First commit in git, migrated from a subversion repository

\subsection{Cross-validation}
The cross-validation exercise randomly samples half of the HRS respondent IDs for use in estimating the transition models.  The respondents not used for estimation, 
but who were present in the HRS sample in 1998, are then simulated from 1998 through 2012.  Demographic, health, and economic outcomes are compared between the 
simulated (“FEM”) and actual (“HRS”) populations.  These results are presented in Table \ref{tab:crossval_unweighted} - Table \ref{tab:crossval_cntecon} for 2000, 2006, and 2012, 
with a statistical test of the difference between the average values in the two populations.

Worth noting is how the composition of the population changes in this exercise.  In 1998, the sample represents those 51 and older.  Since we follow a fixed cohort, 
the age of the population will increase to 65 and older in 2012.  This has consequences for some measures in later years where the eligible population shrinks.

\subsubsection{Demographics}
Demographic measures are presented in Table \ref{tab:crossval_demog}.  Demographic differences between the two populations are small.  The gender balance and fraction 
of the population that is non-Hispanic Black or Hispanic is consistent.

\subsubsection{Health Outcomes}
The FEM population has a slightly higher population with one or more ADL limitation in 2012 (19.6% vs 17.5%).  Those with any IADL limitations are not statistically 
different from one another in 2012.

The two populations are not statistically different from each other for prevalence of cancer, hypertension, or lung disease in 2012.  They do differ for diabetes 
(26.8% for the FEM, 25.0% for the HRS), heart disease (35.3% for the FEM, 32.8% for the HRS), and stroke (12.3% for FEM, 11.1% for the HRS), though the practical 
significance of these differences is not clear.

\subsubsection{Health Risk Factors}
Average BMI is slightly higher for the FEM population in 2012 (28.0 for the FEM vs. 27.7 for the HRS).  In terms of practical significance, this difference is 
equivalent to two pounds for an individual who is 5’8”.  Smoking behavior is not statistically different between the two populations in 2012.  The nursing home 
population is also not statistically different between the FEM and the HRS.

\subsubsection{Economic Outcomes}
Social Security claiming and Supplemental Security Income claiming do not differ between the FEM and the HRS by 2012.  The population working is slightly lower 
in the FEM (22.9% for the FEM vs. 24.6% for the HRS). Household wealth differs in 2012 ($362,000 for the FEM vs $339,000 for the HRS).  Earnings and capital income 
do not differ significantly between the two populations.

On the whole, the cross-validation exercise is reassuring.  Comparing simulated outcomes to actual outcomes using a set of transition models estimated on a 
separate population reveals that the majority of outcomes of interest are not statistically different.  In cases where they are, the practical difference is 
potentially low.

### 2019-11-22 analysis/techdoc/latex/FEM/validation_external_validation.tex
First commit in git, migrated from a subversion repository

\subsection{External Validation}
%NOTE: THIS SECTION NEEDS TO BE VALIDATED

The external validation exercise compares FEM full population simulations beginning in 2004 to external sources.  Here we focus on per capita benefits received from 
Social Security, Disability, and Supplemental Security Income, followed by Medicare and Medicaid.

\subsubsection{Benefits from Social Security Administration}

Conditional on a simulant receiving benefits, the FEM algorithmically assigns benefits for Old Age and Survivors, Supplemental Security Income, and Disability. Here, 
we compare simulation results to SSA figures.

For Old Age and Survivor benefits, we compare to the Social Security Administration’s December 2012 Monthly Statistical Snapshot.  Table 2 of that document indicates 
that the average OASI monthly benefit was \$1194.  FEM forecasts \$1182 for the average beneficiary for 2012.

For Supplemental Security Income we compare to Table 3 of the December 2012 Monthly Statistical Snapshot, focusing on the 65 and older population, as that is the 
only category that is directly comparable.  SSA reports that the average monthly benefit for December of 2012 was \$417.  FEM assigns \$415 to those receiving SSI.

SSA does not report a disability figure that is directly comparable to FEM forecasts.  However, SSA reports average Disability benefits by age, as well as the 
number of individuals receiving benefits at each age.  This allows us to construct the average benefit for workers 51 and older.  Based on this calculation, the 
average disabled worker 51 and older received a benefit of \$1212 in December of 2012.  Spouses of disabled workers can also receive a benefit (SSA reports a 
benefit of \$304 for spouses of disabled workers for all ages).  The 2012 FEM forecast for the average DI beneficiary, which includes both workers and 
their spouses, is \$1102.

\subsubsection{Benefits from Medicare and Medicaid}

For medical spending, we compare FEM forecasts in 2010 to National Health Expenditure Accounts measures from 2010, the most recent year for which these data 
are available.  NHEA reports total amounts by age range, which we then convert to per capita measures using the 2010 Census.  We focus on the 65-84 and 85 plus 
populations, as they are directly comparable to FEM forecasts.  We also aggregate the two groups to produce a 65 plus average.  FEM is similar to NHEA for the 
65 plus population for Medicare (\$10473 for NHEA, \$10494 for FEM) and total medical spending (\$19265 for NHEA, \$19056 for FEM).  FEM estimates are higher for 
Medicaid spending (\$2141 for NHEA, \$2818 for FEM).

### 2019-11-22 analysis/techdoc/latex/shared/dynamic_model_background.tex
First commit in git, migrated from a subversion repository

--- 2019-11-22 analysis/techdoc/latex/FAM/dynamic_model_background.tex
@@
 
-Since then various extensions have been implemented to the original model. The most recent version of the FEM
+Since then various extensions have been implemented to the original model. The most recent version 
 now projects health outcomes for all Americans aged 51 and older and uses the Health and Retirement 
@@
 J-9-P-2-0033, the National Institutes of Aging through the R01 grant ``Integrated Retirement 
-Modeling'' (R01AG030824) and the MacArthur Foundation Research Network on an Aging Society.
-
-This document describes the Future Americans Model (FAM), the development of the model to forecast Americans aged 25 
-and older.  FAM uses the Panel Survey of Income Dynamics (PSID) as the host dataset.  In addition to modeling
-health, health care costs, and economic outcomes, FAM also models life events such as changes in marital status 
-and childbearing.  Development of FAM is suported by the National Institutes of Aging through the USC Roybal 
-Center for Health Policy Simulation (5P30AG024968-13) and the MacArthur Foundation Research Network on an Aging Society.
-
+Modeling'' (R01AG030824) and the MacArthur Foundation Research Network on an Aging Society. Finally, 
+the computer code of the model was transferred from Stata to C++. This report incorporates these new 
+development efforts in the description of the model.
 % \todo add description of work extending to ages 25+ (FAM) here

### 2019-11-22 analysis/volunteer-swv.tex
First commit in git, migrated from a subversion repository

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{datatool}

\title{Analysis of Volunteering in the HRS}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}
library(foreign)
library(ggplot2)
library(survey)
hrs <- read.dta('../input_data/hrs_analytic_recoded.dta')
hrs <- subset(hrs, !is.na(wtresp) & wtresp > 0 & !died & wave %in% 3:7, select=!'r8' %in% names(hrs) | !'r9' %in% names(hrs))
hrs[,'orgnonzero'] <- hrs[,'orghours'] > 0
hrs[,'friendnonzero'] <- hrs[,'friendhours'] > 0
wave3.age <- weighted.mean(hrs[hrs[,'wave'] == 3, 'age'], hrs[hrs[,'wave'] == 3, 'wtresp'])
rhs <- 'age + black + hispan + male + hsless + college + work + hearte + stroke + cancre + hibpe + diabe + adl12 + adl3 + iadl1 + widowed + diclaim + ssclaim + dbclaim + wave'
dorgnz <- svydesign( id=~ 1, data=hrs, weights= ~ wtresp)
dfriendnz <- svydesign( id=~ 1, data=hrs, weights= ~ wtresp)
org.ever.glm <- svyglm(paste('orgnonzero', rhs, sep=' ~ ') , family=quasibinomial(link='logit'), design=dorgnz)
print(summary(org.ever.glm))
friend.ever.glm <- svyglm(paste('friendnonzero', rhs, sep=' ~ '), family=quasibinomial(link='logit'), design=dfriendnz)
print(summary(friend.ever.glm))
orgnz <- subset(hrs, orgnonzero)
frdnz <- subset(hrs, friendnonzero)
dorg <- svydesign( id=~1, data=orgnz, weights=~wtresp)
dfrd <- svydesign( id=~1, data=frdnz, weights=~wtresp)
org.lm <- svyglm(paste('orghours', rhs, sep='~'), design=dorg)
print(summary(org.lm))
frd.lm <- svyglm(paste('friendhours', rhs, sep='~'), design=dfrd)
print(summary(frd.lm))
hrs[,'orgnonzero.p'] <- predict.glm(org.ever.glm, newdata=hrs, type='response', na.action=na.pass)
hrs[,'friendnonzero.p'] <- predict.glm(friend.ever.glm, newdata=hrs, type='response', na.action=na.pass)
hrs[,'orghours.p'] <- predict.glm(org.lm, newdata=hrs, type='response', na.action=na.pass) * hrs[,'orgnonzero.p']
hrs[,'friendhours.p'] <- predict.glm(frd.lm, newdata=hrs, type='response', na.action=na.pass) * hrs[,'friendnonzero.p']
hrs2 <- hrs
hrs2[,'age'] <- wave3.age
hrs[,'orgnonzero.a'] <- predict.glm(org.ever.glm, newdata=hrs2, type='response', na.action=na.pass)
hrs[,'friendnonzero.a'] <- predict.glm(friend.ever.glm, newdata=hrs2, type='response', na.action=na.pass)
hrs[,'orghours.a'] <- predict.glm(org.lm, newdata=hrs2, type='response', na.action=na.pass) * hrs[,'orgnonzero.a']
hrs[,'friendhours.a'] <- predict.glm(frd.lm, newdata=hrs2, type='response', na.action=na.pass) * hrs[,'friendnonzero.a']
rm(hrs2)
gc()
\end{Rcode}

\begin{Rcode}{savefig,label=orgs}
hrs.s <- subset(hrs, !is.na(orghours) & orghours <= 200)
ggplot(hrs.s, aes(x=orghours, weight=wtresp/1e6)) + geom_histogram() + facet_wrap( ~ wave) + labs(y='Millions of People')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{orgs}
\caption{Weighted Histograms of Organizational Volunteer Hours}
\end{figure}

\begin{Rcode}{savefig,label=friends}
hrs.s <- subset(hrs, !is.na(friendhours) & friendhours <= 200)
ggplot(hrs.s, aes(x=friendhours, weight=wtresp/1e6)) + geom_histogram() + facet_wrap( ~ wave) + labs(y='Millions of People')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{friends}
\caption{Weighted Histograms of Friends and Family Volunteer Hours}
\end{figure}

\begin{Rcode}{}
hrs[,'orgweight'] <- ifelse(is.na(hrs[,'orghours']), NA, hrs[,'wtresp'])
hrs[,'friendweight'] <- ifelse(is.na(hrs[,'friendhours']), NA, hrs[,'wtresp'])

hrs.waves <- split(hrs, hrs[,'wave'])

orgnonzero <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero'], x[,'orgweight'], na.rm=T))
orgnonzero.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero.p'], x[,'wtresp'], na.rm=T))
orgnonzero.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero.a'], x[,'wtresp'], na.rm=T))
orgmean <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours'], x[,'orgweight'], na.rm=T))
orgmean.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours.p'], x[,'wtresp'], na.rm=T))
orgmean.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours.a'], x[,'wtresp'], na.rm=T))
orgwtfactor <- sapply(hrs.waves, FUN=function(x) sum(x[,'wtresp'], na.rm=T)/sum(x[,'orgweight'], na.rm=T))
orgtotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours'] * x[,'orgweight'], na.rm=T)/1e9) * orgwtfactor
orgtotal.p <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours.p'] * x[,'wtresp'], na.rm=T)/1e9)
orgtotal.a <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours.a'] * x[,'wtresp'], na.rm=T)/1e9)

friendnonzero <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero'], x[,'friendweight'], na.rm=T))
friendnonzero.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero.p'], x[,'wtresp'], na.rm=T))
friendnonzero.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero.a'], x[,'wtresp'], na.rm=T))
friendmean <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours'], x[,'friendweight'], na.rm=T))
friendmean.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours.p'], x[,'wtresp'], na.rm=T))
friendmean.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours.a'], x[,'wtresp'], na.rm=T))
friendwtfactor <- sapply(hrs.waves, FUN=function(x) sum(x[,'wtresp'], na.rm=T)/sum(x[,'friendweight'], na.rm=T))
friendtotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours'] * x[,'friendweight'], na.rm=T)/1e9) * friendwtfactor
friendtotal.p <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours.p'] * x[,'wtresp'], na.rm=T)/1e9)
friendtotal.a <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours.a'] * x[,'wtresp'], na.rm=T)/1e9)

iearntotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'iearn'] * x[,'wtresp'], na.rm=T)/1e9)

dat <- data.frame(year=c(1996, 1998, 2000, 2002, 2004), orgtotal, orgtotal.p, orgtotal.a, friendtotal, friendtotal.p, friendtotal.a, iearntotal, orgnonzero, friendnonzero, orgnonzero.p, friendnonzero.p, orgnonzero.a, friendnonzero.a)
\end{Rcode}

\begin{Rcode}{savefig,label=value}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal*20.85, colour='Volunteering for Organization')) + 
geom_line(aes(y=friendtotal*11.22, colour='Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{Rcode}{savefig,label=valuea}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal.a*20.85, colour='Age-Adjusted Volunteering for Organization')) + 
geom_line(aes(y=friendtotal.a*11.22, colour='Age-Adjusted Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{Rcode}{savefig,label=valuep}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal.p*20.85, colour='Predicted Volunteering for Organization')) + 
geom_line(aes(y=friendtotal.p*11.22, colour='Predicted Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{value}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{valuea}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{valuep}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{Rcode}{savefig,label=fraction}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero, colour='Fraction Volunteering for Organizations')) +
geom_line(aes(y=friendnonzero, colour='Fraction Volunteering for Friends, etc')) + 
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{Rcode}{savefig,label=fractiona}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero.a, colour='Age-Adjusted Volunteering for Organizations')) +
geom_line(aes(y=friendnonzero.a, colour='Age-Adjusted Volunteering for Friends')) +
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{Rcode}{savefig,label=fractionp}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero.p, colour='Predicted Fraction Volunteering for Organizations')) + 
geom_line(aes(y=friendnonzero.p, colour='Predicted Fraction Volunteering for Friends')) +
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{fraction}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{fractiona}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{fractionp}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{Rcode}{savefig,label=means}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean, colour='Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean, colour='Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{Rcode}{savefig,label=meansa}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean.a, colour='Age-Adjusted Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean.a, colour='Age-Adjusted Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{Rcode}{savefig,label=meansp}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean.p, colour='Predicted Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean.p, colour='Predicted Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{means}
\caption{Mean number of hours}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{meansa}
\caption{Mean number of hours}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{meansp}
\caption{Mean number of hours}
\end{figure}

\end{document}

### 2019-11-22 FEM_CPP/BariatricSurg.cpp

@@ -1,0 +1,238 @@
+#include "BariatricSurg.h"
+#include "Logger.h"
+#include "Random.h"
+#include "fem_exception.h"
+#include <sstream>
+#include <math.h>
+
+BariatricSurg::BariatricSurg(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp) : Intervention(intervention_id, tp, vp)
+{
+	params_map["lapband_start_yr"] = "2010";
+	params_map["lapband_full_yr"] = "2050";
+	params_map["lapband_full_rate"] = "0.5";
+	params_map["lapband_type"] = "1";
+	params_map["lapband_10yr_reduction"] = "0.13";
+	params_map["bs_elig_bmi1"] = "35";
+	params_map["bs_elig_bmi2"] = "40";
+
+	logbmi_model = NULL;
+
+
+	
+	bs_elig_bmi1 = new GlobalVariable("bs_elig_bmi1", 35.0, "Min BMI for Bariatric Surgery, having comorbid conditions");
+	bs_elig_bmi2 = new GlobalVariable("bs_elig_bmi2", 40.0, "Min BMI for Bariatric Surgery, without having comorbid conditions");
+	yr_step = vp->get("yr_step");
+
+	vp->addVariable(bs_elig_bmi1);
+	vp->addVariable(bs_elig_bmi2);
+		
+}
+
+BariatricSurg::~BariatricSurg(void)
+{
+}
+
+
+void BariatricSurg::setScenario(Scenario* scen) {
+	Intervention::setScenario(scen);
+
+	std::string param_name = "lapband_start_yr";
+	if(atof(params_map[param_name].c_str()) == 0.0 && params_map[param_name] != "0") {
+		// User specified a time series name to use for the factor
+		// We arent setup to do this yet, so throw exception
+		throw fem_exception("Bariatric Surgery intervention needs numbers for parameters (text given for lapband_start_yr)!");
+	} else {
+		// User specified a number to use as a constant value
+		lapband_start_yr = atoi(params_map[param_name].c_str());
+	}
+
+	param_name = "lapband_full_yr";
+	if(atof(params_map[param_name].c_str()) == 0.0 && params_map[param_name] != "0") {
+		// User specified a time series name to use for the factor
+		// We arent setup to do this yet, so throw exception
+		throw fem_exception("Bariatric Surgery intervention needs numbers for parameters (text given for lapband_full_yr)!");
+	} else {
+		// User specified a number to use as a constant value
[…]

### 2019-11-22 FEM_CPP/BariatricSurg.h

@@ -1,0 +1,47 @@
+#pragma once
+#include "Intervention.h"
+#include "GlobalVariable.h"
+#include <map>
+class BariatricSurg :
+	public Intervention
+{
+public:
+  BariatricSurg(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~BariatricSurg(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random);
+
+	/** Saves the new BMI after Bariatric Surgery and all other interventions have been applied. 
+		\param[in,out] persons The persons
+		\param[in] year The year the intervention is being applied
+		\param[in] random The Random number provider 
+	*/
+	virtual void postIntervetion(PersonVector& persons, unsigned int year, Random* random);
+
+
+	virtual inline std::string name() const { return std::string("BariatricSurg");};
+	virtual void setScenario(Scenario* scen);
+	virtual void setModelProvider(IModelProvider* mp);
+	virtual void reset();
+protected:
+	bool elig(Person* p) const;
+	
+	std::map<unsigned int, double> pre_treat_logbmi;
+	std::map<unsigned int, unsigned int> treat_year;
+	
+
+	unsigned int lapband_start_yr;
+	int lapband_full_yr;
+	double lapband_full_rate;
+	int lapband_type;
+	double lapband_10yr_reduction;
+
+	GlobalVariable* bs_elig_bmi1;
+	GlobalVariable* bs_elig_bmi2;
+	IVariable* yr_step;
+
+	double log_bs_elig_bmi1;
+	double log_bs_elig_bmi2;
+
+
+	IModel* logbmi_model;
+};

### 2019-11-22 FEM_CPP/BuiltInVariable.h

@@ -1,0 +1,23 @@
+#pragma once
+#include "Variable.h"
+#include "Vars.h"
+#include "utility.h"
+
+class BuiltInVariable :
+	public IVariable
+{
+public:
+
+	BuiltInVariable(Vars::Vars_t v) : var(v) {}
+
+	virtual ~BuiltInVariable(void) 	{}
+
+	virtual inline std::string name() const { return VarsInfo::labelOf(var);}
+	virtual inline std::string description() const {return VarsInfo::infoOf(var).desc;}
+	virtual inline double value(const Person* person) const {return person->get(var);}
+	virtual inline VarTypes::VarTypes_t type() const {return VarsInfo::typeOf(var); }
+	virtual inline bool is_missing(const Person* person) const {return person->is_missing(var);}
+
+protected:
+	Vars::Vars_t var;
+};

### 2019-11-22 FEM_CPP/ConcreteTable.cpp

@@ -1,0 +1,52 @@
+#include "ConcreteTable.h"
+#include "fem_exception.h"
+#include <sstream>
+#include "VarMap.h"
+
+ConcreteTable::ConcreteTable(std::string n, VarMap<double> table) : data(table) {
+	name = n;
+}
+
+ConcreteTable::~ConcreteTable(void)
+{
+  data.clear();
+}
+
+double ConcreteTable::Value(const TableIndex &index) const {
+	std::vector<std::string> names = data.getIndexVarNames();
+	VarMapKey k(names.size());
+	for(unsigned int i=0; i < names.size(); i++) {
+		k[i] = index.get(names[i]);
+	}
+	return(data.get(k));
+}
+
+double ConcreteTable::Value(const Person &person) const {
+	TableIndex t = getIndexTemplate();
+	t.set(person);
+	return(Value(t));
+}
+
+TableIndex ConcreteTable::getIndexTemplate(void) const {
+	TableIndex t(data.getVariableProvider());
+	std::vector<std::string> names = data.getIndexVarNames();
+	for(unsigned int i=0; i < names.size(); i++) {
+		t.addName(names[i]);
+	}
+	return(t);
+}
+
+bool ConcreteTable::isIndex(const TableIndex &index) const {
+	std::vector<std::string> names = data.getIndexVarNames();
+	VarMapKey key(names.size());
+	for(unsigned int i=0; i < names.size(); i++)
+		key[i] = index.get(names[i]);
+	return(data.isIndexKey(key));
+}
+
+bool ConcreteTable::isIndex(const Person &person) const {
+	TableIndex t = getIndexTemplate();
+	t.set(person);
+	return(isIndex(t));
+}
+

### 2019-11-22 FEM_CPP/CrossSectionalModule.cpp

@@ -1,0 +1,204 @@
+#include "CrossSectionalModule.h"
+#include "Logger.h"
+#include <sstream>
+#include <algorithm>
+#include <fstream>
+#include <math.h>
+#include "utility.h"
+#include "fem_exception.h"
+
+CrossSectionalModule::CrossSectionalModule(IVariableProvider* vp, ITimeSeriesProvider* tp) : sscalc(tp)
+{
+	mp = NULL;
+	qaly_nhm_var = vp->get("qaly_nhmliv_reduction");
+	
+	variable_provider = vp;
+}
+
+CrossSectionalModule::~CrossSectionalModule(void)
+{
+
+}
+
+
+void CrossSectionalModule::setScenario(Scenario* scen) {
+
+	Module::setScenario(scen);
+	models.clear();
+	vars_to_model.clear();
+	if(scen->contains("xsectional")) {
+		std::vector<std::string> var_names(10);
+		str_tokenize(scen->get("xsectional"), var_names);
+		for(unsigned int i = 0; i < var_names.size(); i++) 
+			if(VarsInfo::indexOf(var_names[i]) != Vars::_NONE)
+				this->vars_to_model.push_back(VarsInfo::indexOf(var_names[i]));
+	}
+	setModelProvider(mp);
+
+}
+
+void CrossSectionalModule::process(PersonVector& persons, unsigned int year, Random* random)
+{
+	Logger::log("Running Cross Sectional Module", FINE);
+	std::vector<Person*>::iterator itr;
+
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+		Person* person = *itr;
+
+		// Is the person alive?
+		if(person->test(Vars::active) && !person->test(Vars::died)) {
+
+			// Update PIA -- (Comment 7/30/2014) Is this necessary?  PIA is part of the benefit calculation elsewhere.
+			if(variable_provider->get("hrs_data")->value() == 1) {
+				const double rpia_tmp = sscalc.SsPIA(person->get(Vars::raime), person->get(Vars::rq), person->get(Vars::rbyr)+62);
+				person->set(Vars::rpia, rpia_tmp);
+			}
+
+			std::vector<Vars::Vars_t>::iterator it;
+			for(it = vars_to_model.begin(); it != vars_to_model.end(); ++it) {
+				Vars::Vars_t v = *it;
+				double qaly = 0.0;
+				switch(v) {
+				case Vars::qaly:
[…]

### 2019-11-22 FEM_CPP/EconGenModule.cpp

@@ -1,0 +1,398 @@
+#include "EconGenModule.h"
+#include <math.h>
+#include "Logger.h"
+
+EconGenModule::EconGenModule(ITimeSeriesProvider* timeSeriesProvider)
+{
+	
+//	nra = timeSeriesProvider->get("nra");
+	nwi = timeSeriesProvider->get("nwi");
+
+	// Init db_type
+	// sex, era, nra
+	db_type[0][45][55] = 9.588;
+	db_type[0][45][60] = 9.021;
+	db_type[0][45][65] = 7.568;
+	db_type[0][50][55] = 9.244;
+	db_type[0][50][60] = 8.913;
+	db_type[0][50][65] = 8.49;
+	db_type[0][55][55] = 9.094;
+	db_type[0][55][60] = 8.887;
+	db_type[0][60][60] = 8.798;
+	db_type[0][60][65] = 8.269;
+	db_type[0][55][65] = 8.341;
+
+	db_type[1][45][55] = 10.131;
+	db_type[1][45][60] = 9.712;
+	db_type[1][45][65] = 8.937;
+	db_type[1][50][55] = 9.834;
+	db_type[1][50][60] = 9.494;
+	db_type[1][50][65] = 9.184;
+	db_type[1][55][55] = 9.676;
+	db_type[1][55][60] = 9.386;
+	db_type[1][60][60] = 9.344;
+	db_type[1][60][65] = 8.886;
+	db_type[1][55][65] = 8.944;
+
+}
+
+EconGenModule::~EconGenModule(void)
+{
+}
+
+void EconGenModule::setModelProvider(IModelProvider* modp) {
+  mp = modp;
+  
+  try {
+    proptax_model = mp->get("proptax");
+  } catch (fem_exception& e) {
+    throw fem_exception("EconGenModule cannot find the required proptax model");
+  }
+  
+}
+
+void EconGenModule::process(PersonVector& persons, unsigned int year, Random* random) {
+	Logger::log("Running Economics Module", FINE);
+
+	std::vector<Person*>::iterator itr;
+	double delta_nwi = -1;
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+		Person* person = *itr;
+		if((person->test(Vars::active)) && !person->test(Vars::l2died)) {
+		  delta_nwi = nwi->Value(year-scenario->YrStep()) / nwi->Value(year - scenario->YrStep()*2.0);
+			//  For now, including grandkid care hours logic here
+			if(person->get(Vars::fkids) == 0) 
[…]

### 2019-11-22 FEM_CPP/EconGenModule.h

@@ -1,0 +1,25 @@
+#pragma once
+#include "Module.h"
+#include "TimeSeries.h"
+#include <map>
+class EconGenModule :
+	public Module
+{
+public:
+  EconGenModule(ITimeSeriesProvider* timeSeriesProvider);
+	virtual ~EconGenModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Econ Gen Module";}
+	virtual void setModelProvider(IModelProvider* modp);
+protected:
+	double updateAIME(double aime, double ry_earn, double age, bool male, unsigned int yr) const;
+	double DbBenefit(double ry_earn, double tenure, double edu, double age, int era, int nra, unsigned int yr, bool sex) const;
+	ITimeSeries* nwi;
+//	ITimeSeries* nra;
+	std::map<int, std::map<int, std::map<int, double> > > db_type;
+
+	IModelProvider* mp;
+	/// Model for predicting property taxes
+	IModel* proptax_model;
+
+};

### 2019-11-22 FEM_CPP/EquationNode.h

@@ -1,0 +1,160 @@
+#pragma once
+#include <string>
+#include <vector>
+#include <map>
+#include "Person.h"
+#include "Lookup.h"
+#include <math.h>
+#include "Variable.h"
+
+
+class INode   {
+public:
+	virtual ~INode() {}
+	virtual double value(const Person* person) const = 0;
+	virtual unsigned int getPrecedence() const = 0;
+	virtual std::string getName() const = 0;
+	virtual INode* clone() const = 0;
+};
+
+
+class LookupNode : public INode
+{
+protected:
+	const static unsigned int MAX_INPUTS = 10;
+	Lookup* lookup;
+	INode* inputs[MAX_INPUTS];
+	unsigned int preced;
+	unsigned int nReq;
+	unsigned int nInputs;
+
+public:
+	LookupNode(Lookup* lkup, unsigned int p = 99);
+	LookupNode(const LookupNode& source);
+	virtual ~LookupNode();
+	inline virtual LookupNode* clone() const { return new LookupNode(*this); }
+
+	virtual std::string getName() const;
+	inline virtual bool Complete() const {  return nInputs == nReq;         }
+	inline virtual unsigned int getPrecedence() const { return preced; }
+
+	virtual double value(const Person* person) const;
+
+	inline virtual void addNode(INode* n)    {    inputs[nInputs++] = n;     }
+
+};
+
+class VariableNode : public INode
+{
+private:
+	IVariable* var;
+
+public:
+	VariableNode(IVariable* v) { var = v; }
+	inline virtual ~VariableNode() {}
+	inline virtual double value(const Person* person) const { return var->value(person); }
+
+	inline virtual unsigned int getPrecedence() const { return 99; }
+
+	inline virtual std::string getName() const { return var->name();}
+
+	inline virtual INode* clone() const { return new VariableNode(var);}
+};
+
+class NumberNode : public INode
+{
+protected:
+	double num;
+
+public:
+	NumberNode(double n) { num = n; }
+
+	virtual inline double value(const Person* person) const { return num; }
+
+	virtual inline unsigned int getPrecedence() const { return 99; } 
+	virtual std::string getName() const;
+	inline virtual INode* clone() const { return new NumberNode(num); }
+};
+
+
+/* Operations */
+inline double add_op(double a, double b) {return a + b;}
+inline double minus_op(double a, double b) {return a - b;}
+inline double mult_op(double a, double b) {return a * b;}
[…]

### 2019-11-22 FEM_CPP/EquationParser.cpp

@@ -1,0 +1,126 @@
+
+#include "EquationParser.h"
+#include "utility.h"
+#include <stack>
+#include "fem_exception.h"
+std::vector<std::string> EquationParser::getUnknownTokens(std::string equation, NodeBuilder* builder) {
+	std::vector<std::string> delims = builder->getOps();
+	delims.push_back(" ");
+	delims.push_back(")");
+	delims.push_back(",");
+	delims.push_back("(");
+
+	std::vector<std::string> tokens;
+	str_tokenize_keep_delim(equation, tokens, delims);
+	std::vector<std::string> unknownTokens;
+	for(unsigned int i = 0; i < tokens.size(); i++) {
+		std::string s = tokens[i];
+		if (s.length() > 0 && s != "(" && s != ")" && s != "," && s != " " && !builder->containsLookup(s) && !builder->containsOp(s) && !builder->containsTerminalNode(s))
+			unknownTokens.push_back(s);
+	}
+	return unknownTokens;
+}
+
+INode* EquationParser::parseString(std::string equation, NodeBuilder* builder) {
+
+	std::vector<std::string> delims = builder->getOps();
+	delims.push_back(" ");
+	delims.push_back(")");
+	delims.push_back(",");
+	delims.push_back("(");
+
+	std::vector<std::string> tokens;
+	str_tokenize_keep_delim(equation, tokens, delims);
+	std::queue<std::string> tokens_queue;
+	for(unsigned int i = 0; i < tokens.size(); i++)
+		if(tokens[i].length() > 0 && tokens[i] != " ")
+			tokens_queue.push(tokens[i]);
+	INode* n  = parse(tokens_queue, builder);
+	return n;
+}
+
+
+INode* EquationParser::parse(std::queue<std::string>& tokens, NodeBuilder* builder) {
+	LookupNode* functionnode = NULL;
+
+	INode* active = NULL;
+	std::stack<OpNode*> stack;
+	std::string token;
+	while (!tokens.empty())
+	{
+		token = tokens.front();
+		tokens.pop();
+		if (token.length() > 0)
+		{
+			if (builder->containsOp(token))
+			{
+				OpNode* o = builder->buildOpNode(token);
+				if (stack.empty())
+					stack.push(o);
+				else if (o->getPrecedence() > stack.top()->getPrecedence())
+					stack.push(o);
+				else {
+					while (!stack.empty() && o->getPrecedence() <= stack.top()->getPrecedence())	{
[…]

### 2019-11-22 FEM_CPP/EquationParser.h

@@ -1,0 +1,21 @@
+#pragma once
+#include <string>
+#include <vector>
+#include <queue>
+#include "EquationNode.h"
+
+/** This class handles the parsing of equations in summary files and vars files.
+
+    \bug Does not check for balanced parentheses
+*/
+class EquationParser
+{
+public:
+	static std::vector<std::string> getUnknownTokens(std::string equation, NodeBuilder* builder);
+	static INode* parseString(std::string equation, NodeBuilder* builder);
+
+	
+
+private:
+	static INode* parse(std::queue<std::string>& tokens, NodeBuilder* builder);
+};

### 2019-11-22 FEM_CPP/FEM.cpp

@@ -1,0 +1,648 @@
+#include "FEM.h"
+#include "utility.h"
+#include "Logger.h"
+#include "Random.h"
+#include "fem_exception.h"
+#include "file_not_found_exception.h"
+#include <sstream>
+#include <iomanip>
+#include "ConstantTimeSeries.h"
+#include "SurvivalModel.h"
+
+FEM::FEM(Settings& s) : model_manager(&var_manager), settings(s)
+{
+	/* set up list of survival model types */
+	survival_type_names.insert("WeibullPHSurvival");
+	survival_type_names.insert("Survival");
+	
+	/* Prepare a string stream to make messages */
+	std::ostringstream ss;
+	ss.str("");
+
+	ss << "Preparing the FEM Simulator:";
+	Logger::log(ss.str(), INFO);
+
+	ss.str("");
+
+	/* Assign the equation builder to the variable and time series managers */
+	ts_manager.setBuilder(&builder);
+	table_manager.setBuilder(&builder);
+	var_manager.setBuilder(&builder);
+
+	// Setup medicare elig age time series 
+	medicare_elig_age = new ProxyTimeSeries("medicare_elig_age", "Medicare Eligibility Age");
+	ts_manager.addTimeSeries(medicare_elig_age);
+	medicare_elig_age->setSeries(ts_manager.addTimeSeries(new ConstantTimeSeries("medicare_elig_age_default", 65)));
+
+	// Setup global variables
+	scenario_variables.push_back(new GlobalVariable("mcare_ptb_coin_elas", -.2, "Medicare Part B elasticity to coinsurance"));
+	scenario_variables.push_back(new GlobalVariable("mcare_ptb_coin_chg", 0, "The % change in Medicare Part B co-insurance"));
+	scenario_variables.push_back(new GlobalVariable("tot_meps",1.05,"NHEA/MEPS adjustment for total medical expenditures"));
+	scenario_variables.push_back(new GlobalVariable("tot_mcbs",1.08,"NHEA/MCBS adjustment for total medical expenditures"));
+	scenario_variables.push_back(new GlobalVariable("mcaid_meps",1.59,"NHEA/MEPS adjustment for medicaid expenditures"));
+	scenario_variables.push_back(new GlobalVariable("mcaid_mcbs",1.48,"NHEA/MCBS adjustment for medicaid expenditures"));
+	scenario_variables.push_back(new GlobalVariable("mcare_meps",1.08,"NHEA/MEPS adjustment for medicare expenditures"));
[…]

### 2019-11-22 FEM_CPP/FEM.h

@@ -1,0 +1,189 @@
+#pragma once
+#include "Intervention.h"
+#include "Module.h"
+#include "Person.h"
+#include "InterventionModule.h"
+#include "InitInterventionModule.h"
+#include "HealthModule.h"
+#include "GlobalPreInitializationModule.h"
+#include "HRSInitializationModule.h"
+#include "PSIDInitializationModule.h"
+#include "GlobalPostInitializationModule.h"
+#include "VarCalcModule.h"
+#include "LagModule.h"
+#include "MedCostsModule.h"
+#include "EconGenModule.h"
+#include "OutputModules.h"
+#include "ImmigrationAdjModule.h"
+#include "EarningsModule.h"
+#include "GovExpModule.h"
+#include "CrossSectionalModule.h"
+#include "SummaryModule.h"
+#include "PersonVector.h"
+#include "EquationNode.h"
+#include "TimeSeriesManager.h"
+#include "TableManager.h"
+#include "VariableManager.h"
+#include "Settings.h"
+#include "ModelManager.h"
+#include "ProxyTimeSeries.h"
+#include "GlobalVariable.h"
+#include "InterventionFactory.h"
+#include "RandomProvider.h"
+#include "EducModule.h"
+#include <vector>
+#include <set>
+
+/** \mainpage The Future Elderly Model
+
+This is the main documentation for the Future Elderly Model. From here, you can access the simulation documentation as well as the documentation for the Stata code that generates the simulation inputs. This is also where information global to the model (like high-level todo items) will be located.
+
+The documentation is organized as follows:
+\li \subpage cpp The C++ Simulation Documentation
+\li \subpage stata The Stata Data Generation Code
+\li \subpage estimation The Stata Estimation Code
+
+\section longterm Long Term ToDo list
+This isn't quite a wishlist, since this stuff needs to be done, but we can keep producing results without finishing any of these items.
+
+\todo Create a regressions tests group of scenarios and settings files. 
+\li prevalence of 1+ ADLs
+\li prevalence of working (under age 65)
+\li Part A spending
+\li Part B spending
+\li total population
+\li prevalence of diabetes
+\li total population 50+ in 2050
+\li total population 65+ in 2050
+\li gov't expenditure totals in 2050 (by program)
+
+\todo Create a list of desirable unit tests.
+
+\todo Write the unit tests.
+
+\todo Create an extreme conditions group of scenarios and settings files
+\li zero mortality
+\li total mortality
+\li zero incidence of disease
[…]

### 2019-11-22 FEM_CPP/FEM_MPISlave.cpp

@@ -1,0 +1,53 @@
+#include "FEM_MPISlave.h"
+#include "utility.h"
+#include "Logger.h"
+#include "Random.h"
+#include "fem_exception.h"
+#include <sstream>
+#include <iomanip>
+#include "mpi.h"
+
+FEM_MPISlave::FEM_MPISlave(Settings& settings) : FEM(settings)
+{
+
+}
+
+FEM_MPISlave::~FEM_MPISlave(void)
+{
+}
+
+
+
+void FEM_MPISlave::runScenario(Scenario* scenario, RandomProvider* rnd_provider) {
+	std::ostringstream ss;
+	ss.str("");
+	prepScenario(scenario);
+	
+	Random* random;
+
+	MPI::Status stats;
+
+	unsigned int nyr_steps = (scenario->EndYr() - scenario->StartYr())/scenario->YrStep() + 1;
+	unsigned int nmeasures = summary_module->numMeasures();
+	double* summary_data = new double[nyr_steps*nmeasures+1];
+
+	// NOTE: MPI communicates with the actual rep number (1,2,3,...), but
+	// FEM classes uses zero-indexing for the rep number (0,1,2,...)
+	unsigned long int rep;
+	MPI::COMM_WORLD.Recv(&rep, 1, MPI::UNSIGNED_LONG, 0, MPI::ANY_TAG, stats);
+	while(rep >= 1) {
+		ss << "Repetition " << rep << "..........";
+		Logger::log(ss.str().c_str(), INFO);
+		ss.str("");
+		random = rnd_provider->getRandom(rep-1);
+		runRep(scenario, random, rep-1);
+		summary_module->getRepData((unsigned int) rep-1, summary_data+1);
+		summary_data[0] = rep-1;
+		MPI::COMM_WORLD.Send(summary_data, (int)(nyr_steps*nmeasures+1), MPI::DOUBLE, 0, 1);
+		
+		MPI::COMM_WORLD.Recv(&rep, 1, MPI::UNSIGNED_LONG, 0, MPI::ANY_TAG, stats);
+	}
+
+	delete[] summary_data;
+}
+

### 2019-11-22 FEM_CPP/FEM_MPISlave.h

@@ -1,0 +1,11 @@
+#pragma once
+#include "FEM.h"
+
+class FEM_MPISlave :
+	public FEM
+{
+public:
+	FEM_MPISlave(Settings& settings);
+	virtual ~FEM_MPISlave(void);
+	virtual void runScenario(Scenario* scenario, RandomProvider* rnd_provider);
+};

### 2019-11-22 FEM_CPP/GHREG.cpp

@@ -1,0 +1,109 @@
+#include "GHREG.h"
+#include "utility.h"
+#include "Random.h"
+#include <sstream>
+#include <limits>
+#include "fem_exception.h"
+#include <cstring>
+
+GHREG::GHREG(void)
+{
+}
+
+GHREG::~GHREG(void)
+{
+}
+
+
+
+
+GHREG::GHREG(const GHREG& source) : Regression(source)
+{
+	theta = source.theta;
+	omega = source.omega;
+	ssr = source.ssr;
+	sqrt_ssr = source.sqrt_ssr;
+	hb = source.hb;
+	dhb = source.dhb;
+
+}
+
+
+void GHREG::predict(Person* person, const Random* random) const
+{
+	double x = transform(calc_xb(person) + random->normalDist(person->getID(), predicted_var, person->getYear(), 0, 1)*sqrt_ssr);
+	predictWithProb(person, random, x);
+}
+
+
+
+std::string GHREG::describe() const
+{
+	std::string desc = Regression::describe();
+	std::stringstream strm;
+	strm << desc;
+	if(predicted_var != Vars::_NONE) { // Model is loaded?
+		strm << "\ttheta = " << theta << std::endl;
+		strm << "\tomega = " << omega << std::endl;
+		strm << "\tssr = " << ssr << std::endl;
+	}
+	return strm.str();
+} 
+
+
+double GHREG::transform(double g) const {
+	double x =  theta*dhb*g+hb;
+	double sinh = 0.5*(exp(x)-exp(-x));
+	
+	// computations
+	return (sinh-theta*omega)/theta;
+}
+
+
+void GHREG::read(std::istream& istrm, IVariableProvider* provider)
+{
+	std::string buf;
+	istrm >> buf;
+	predicted_var = VarsInfo::indexOf(buf);
+
+	// Check if the variable read in exists
+	if (predicted_var == Vars::_NONE) {
+		std::ostringstream ss;
+		ss << "Error reading GHREG model " << getName() << ": Predicted variable [" << buf << "] does not exists!";
+		throw fem_exception(ss.str().c_str());
+	}
+
+	// Eat any newline characters
+	istrm.ignore(std::numeric_limits<std::streamsize>::max(), '\n');
+
+	std::vector<double> temp_coeffs(50), temp_perturbs(50);
+	std::map<std::string, double> params;
[…]

### 2019-11-22 FEM_CPP/GHREG.h

@@ -1,0 +1,38 @@
+#pragma once
+#include "Regression.h"
+
+
+/** Implements a generalized inverse hyperbolic sine transform model.
+	This model is suited to modeling distributions with a long left tail and some negative values. For example, it used to model wealth and earnings. 
+	It is of the form \f$y = \theta^{-1}\sinh[\theta h'(\theta\omega)g + h(\theta\omega)] - \theta\omega\f$ where \f$h(x) = \sinh^{-1}(x)\f$
+	and \f$g(x) = x\beta + \sigma\eta\f$. \f$\theta\f$ and \f$\omega\f$ are shape parameters. The model is estimated used maximum likelihood.
+*/
+class GHREG :
+	public Regression
+{
+public:
+	/** Constructs a new empty model */
+	GHREG(void);
+
+	virtual ~GHREG(void);
+
+	/** Copy Constructor */
+	GHREG(const GHREG& source);
+
+	virtual void predict(Person* person, const Random* random) const;
+	virtual IModel* clone() const { return new GHREG(*this);}
+	virtual std::string describe() const;
+	virtual std::string getType() const { return "ghreg";}
+	virtual std::string getTypeDesc() const { return "ghreg";}
+
+	/** Read method for the GHREG model.
+
+	    \todo Update this method to handle the Stata interacted terms format
+	*/
+	virtual void read(std::istream& inf, IVariableProvider* provider);
+	
+protected:
+	virtual double transform(double g) const;
+	double theta, omega, ssr, sqrt_ssr, hb, dhb;
+	
+};

### 2019-11-22 FEM_CPP/GlobalPostInitializationModule.cpp

@@ -1,0 +1,92 @@
+#include "GlobalPostInitializationModule.h"
+#include "Logger.h"
+#include "SummaryModule.h"
+#include "utility.h"
+#include <math.h>
+#include <sstream>
+#include "fem_exception.h"
+
+GlobalPostInitializationModule::GlobalPostInitializationModule(IVariableProvider* vp, ITimeSeriesProvider *timeSeriesProvider)
+{
+	medicare_elig_var = vp->get("medicare_eligibility");
+	nwi = timeSeriesProvider->get("nwi");
+	cpi_yearly = timeSeriesProvider->get("cpi.yearly");
+
+}
+
+GlobalPostInitializationModule::~GlobalPostInitializationModule(void)
+{
+}
+
+
+
+void GlobalPostInitializationModule::setScenario(Scenario* scen) {
+	Module::setScenario(scen);
+	if(scen->get("psid_data")=="1")
+		ref_year = 2009;
+	else 
+		ref_year = 2010;
+}
+
+void GlobalPostInitializationModule::process(PersonVector& persons, unsigned int year, Random* random) {
+		
+	Logger::log("Running Global Post-Initialization Module", FINE);
+	/*
+	int simu_type = scenario->SimuType();
+	*/
+
+	// Scaling in CPI and NWI
+	double delta_cpi = 1.0;
+	for(unsigned int i = ref_year; i <= year; i++)
+	  delta_cpi *= (1 + cpi_yearly->Value(i));
+	/*
+		const double delta_nwi = nwi->Value(year)/nwi->Value(ref_year);
+	*/
+
+	std::vector<Person*>::iterator itr;
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+		Person* p = *itr;
+		// Only initialize just added persons
+		if(p->get(Vars::entry) == year)  {
+			/* Set the medicare part B enrollment for the initial host dataset */
+			if(medicare_elig_var->value(p) == 1.0) {
+				/* Only do this for people eligable for medicare, but also that are not dead because those that are dead will be missing some key variables */
+				if(!p->test(Vars::died))
+					init_medicare_partb_enroll->predict(p, random);
+				else
+					p->set(Vars::mcare_ptb_enroll, false);
+
+				/* Set that this person is enrolled in Medicare */
+				p->set(Vars::mcare_pta_enroll, true);
+				p->set(Vars::medicare_elig, true);
+			} else {
[…]

### 2019-11-22 FEM_CPP/GlobalPostInitializationModule.h

@@ -1,0 +1,42 @@
+/** \file Initializtion module for variables that are not specific to the host data set
+This runs AFTER the initialization modules for each host data set.
+It can use models that are specific to the host data set.
+*/
+
+#pragma once
+#include "Module.h"
+#include "TimeSeries.h"
+
+class GlobalPostInitializationModule :
+	public Module
+{
+public:
+	/** Creates the Initialization Module.
+		\param[in]  vp  The runtime variable provider used to obtain the medicare eligibility variable
+		\param[in]  timeSeriesProvider The time series provider used to obtain any time series needed by the Module
+	*/
+	GlobalPostInitializationModule(IVariableProvider* vp, ITimeSeriesProvider *timeSeriesProvider);
+	virtual ~GlobalPostInitializationModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Global Initialization Module";}
+	virtual void setModelProvider(IModelProvider* mp);
+	virtual void setScenario(Scenario* scen);
+protected:
+	/** Model for determining if a person eligable for medicare will have Medicare Part B */
+	IModel* init_medicare_partb_enroll;
+
+	/** Variable for calculating if the person is eligible for Medicare */
+	IVariable* medicare_elig_var;
+
+	/** Time series for National Wage Index */
+	ITimeSeries* nwi;
+
+	/** Time series for Consumer Price Index */ 
+	ITimeSeries* cpi_yearly;
+	
+	/** Time series for normal retirment age */
+	ITimeSeries* nra;
+
+	/** Year the scenario starts */
+	unsigned int ref_year;
+};

### 2019-11-22 FEM_CPP/GovExpModule.h

@@ -1,0 +1,145 @@
+#pragma once
+#include "Module.h"
+#include "TimeSeries.h"
+#include "Model.h"
+#include "SSCalculator.h"
+
+class GovExpModule :
+	public Module
+{
+public:
+	GovExpModule(ITimeSeriesProvider *tp);
+	virtual ~GovExpModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Govt Expenditures Module";}
+	virtual void setModelProvider(IModelProvider* mp);
+protected:
+	/* 
+	// SS Benefit calculations should be done through the SSCalculator class.  This was commented out on 7/30/2013, remove it if it is no longer needed.
+	void SSBenefit(double raime, double rq, double ry_earn, unsigned int rclyr, unsigned int rbyr, bool rl, unsigned int rdthyr,
+		double saime, double sq, double sy_earn, unsigned int sclyr, unsigned int sbyr, bool sl, unsigned int sdthyr,
+		bool rmar, unsigned int yr,
+		double &rben_g, double &rben_s, double &rben_v, double &sben_g, double &sben_s, double &sben_v);
+	*/
+	
+	/* 
+	// PIA calculations should be done through the SSCalculator class.  This was commented out on 7/30/2014, remove it if it is no longer needed.
+	double SsPIA(double raime, double rq, unsigned int rbyr, bool alive, unsigned int dthyr);
+	*/
+	double DiBenefit(double raime, double rq, double ry_earn, unsigned int rbyr, unsigned int yr);
+
+/** Struct to hold information about taxation paramters */	
+	struct tax_params {
+		/// Federal tax schedule 
+		/// brackets single (s) and couples (c)
+		double sbra[5];
+		double cbra[5];
+		double mtr[6];
+		
+		/// Basic deduction and old age deduction
+		double bded_sing;
+		double bded_coup;
+		double oded_sing;
+		double oded_coup;
+		double pded;
+		double pded_r;
+		double pded_s;
+		double pded_tcoup;
+		double pded_tsing;
+		
+		
+		/// taxation of social security benefits 
+		double base_ssa_sing;
+		double base_ssa_coup;
+		double base2_ssa_sing;
+		double base2_ssa_coup;
+
+		/// tax credit for low-income elderly
+		double tc_max1_sing;
+		double tc_max1_one65;
+		double tc_max1_both65;
[…]

### 2019-11-22 FEM_CPP/HRSInitializationModule.cpp

@@ -1,0 +1,229 @@
+#include "HRSInitializationModule.h"
+#include "Logger.h"
+#include "utility.h"
+#include <sstream>
+
+HRSInitializationModule::HRSInitializationModule(IVariableProvider* vp, ITimeSeriesProvider* tsp) {
+  variable_provider = vp;
+  timeSeriesProvider = tsp;
+  
+ 	nwi = timeSeriesProvider->get("nwi");
+	cpi_yearly = timeSeriesProvider->get("cpi.yearly");
+}
+
+HRSInitializationModule::~HRSInitializationModule() { }
+
+void HRSInitializationModule::setScenario(Scenario* scen) {
+  Module::setScenario(scen);
+	ref_year = 2010;
+}
+
+void HRSInitializationModule::process(PersonVector& persons, unsigned int year, Random* random) {
+  enabled = (variable_provider->get("hrs_data")->value() == 1);
+  if(enabled) {
+    Logger::log("Running HRS Initialization Module", FINE);
+   	int simu_type = scenario->SimuType();
+   	// Scaling in CPI and NWI
+		double delta_cpi = 1.0;
+		for(unsigned int i = ref_year; i <= year; i++)
+	  	delta_cpi *= (1 + cpi_yearly->Value(i));
+ 		const double delta_nwi = nwi->Value(year)/nwi->Value(ref_year);
+
+   	std::vector<Person*>::iterator itr;
+		for(itr = persons.begin(); itr != persons.end(); ++itr) {
+			Person* p = *itr;
+			// Only initialize just added persons
+			if(p->get(Vars::entry) == year)  {
+				
+				// Some special handling for base cohort 
+				if(year == scenario->StartYr()) {
+					// Initial AIME
+					if(simu_type == 2 || simu_type == 3) {
+						p->set(Vars::raime,  p->get(Vars::fraime));
+					}
+				}
+				
+				// If they havent claimed SS, then mark it - This should be assessed.
+				// if(p->get(Vars::rssclyr) == 2100)
+				//	p->set(Vars::ssclaim, false);
+
+				// Set missing hearta and heartae to zero for those who are deceased at simulation start
+				if(p->test(Vars::died) & p->is_missing(Vars::hearta))
[…]

### 2019-11-22 FEM_CPP/HRSInitializationModule.h

@@ -1,0 +1,44 @@
+#pragma once
+#include "Module.h"
+#include "TimeSeries.h"
+
+class HRSInitializationModule :
+	public Module
+{
+public:
+	/** Creates the Initialization Module.
+		\param[in]  vp  The runtime variable provider used to obtain the medicare eligibility variable
+		\param[in]  timeSeriesProvider The time series provider used to obtain any time series needed by the Module
+	*/
+	HRSInitializationModule(IVariableProvider* vp, ITimeSeriesProvider *timeSeriesProvider);
+	virtual ~HRSInitializationModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "HRS Initialization Module";}
+	virtual void setModelProvider(IModelProvider* mp);
+	virtual void setScenario(Scenario* scen);
+protected:
+	/** Model for predictin capital income */
+	IModel* hicap_model;
+	/** Model for predicting nonzero capital income */
+	IModel* hicap_nonzero_model;
+	
+	/* Model for predicting nonzero property taxes */
+	IModel* proptax_nonzero_model;
+	
+	/* Model for predicting nonzero igxfr */
+	IModel* igxfr_nonzero_model;
+	IModel* igxfr_model;
+	
+	/** Model for pain status */
+	IModel* painstat_model;
+	
+	/** Time series for Consumer Price Index */ 
+	ITimeSeries* cpi_yearly;
+	
+	/** Time series for National Wage Index */
+	ITimeSeries* nwi;
+	
+	/** Year the scenario starts */
+	unsigned int ref_year;
+
+};

### 2019-11-22 FEM_CPP/ImmigrationAdjModule.cpp

@@ -1,0 +1,160 @@
+#include "ImmigrationAdjModule.h"
+#include <math.h>
+#include "Logger.h"
+#include <sstream>
+#include <string>
+#include "fem_exception.h"
+#include <cstring>
+#include "utility.h"
+#include "VarMap.h"
+
+ImmigrationAdjModule::ImmigrationAdjModule(IVariableProvider* vp, ITableProvider* tabp) 
+{
+	tableProvider = tabp;
+	variable_provider=vp;
+	netMigration = tableProvider->get("immigration");
+}
+
+ImmigrationAdjModule::~ImmigrationAdjModule(void)
+{
+	clear();
+}
+
+void ImmigrationAdjModule::clear() {
+	netMigration = NULL;
+}
+
+void ImmigrationAdjModule::process(PersonVector& persons, unsigned int year, Random* random)
+{
+	Logger::log("Running Immigration Adjustment Module", FINE);
+	std::vector<Person*>::iterator itr;
+			
+	double age;
+	
+	// use dummy to shift person to lag and interim years without direct modification of person
+	// dummy helps avoid accidental changes to person that don't get undone
+	Person dummy; 
+	
+	VarMap<double> wt1(variable_provider, netMigration->getIndexTemplate().getNames());
+		
+	/* Calculate weights. Use lag age + 1 to account for population shifts in the inbetween year */
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+	  dummy = **itr; 
+	  // back up age and year to the inbetween year
+	  dummy.set(Vars::age, (int)(dummy.get(Vars::l2age)) + 1);
+	  dummy.set(Vars::year, dummy.get(Vars::year) - 1);
+	  
+	  if(dummy.test(Vars::active) && !dummy.test(Vars::l2died) && netMigration->isIndex(dummy)) {
+			if(dummy.get(Vars::weight) < 0.0) 
+				throw fem_exception("Person has negative weight before immigration adjustment (interim year)");
+			else {
+				double sumwt = wt1.isIndexKey(dummy) ? wt1.get(dummy) : 0.0;
+				wt1.set(dummy, sumwt + dummy.get(Vars::weight));
+			}
+		}					
+	}
+
+
+	/* Adjust weights */
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+		Person* person = *itr;
+		dummy = *person;
+
[…]

### 2019-11-22 FEM_CPP/ImmigrationAdjModule.h

@@ -1,0 +1,24 @@
+#pragma once
+#include "Module.h"
+#include "Table.h"
+
+/** 
+\todo Generalize \a process code to work with any size time step instead of only 2 year steps
+*/
+class ImmigrationAdjModule :
+	public Module
+{
+public:
+	ImmigrationAdjModule(IVariableProvider* vp, ITableProvider* tabp);
+	virtual ~ImmigrationAdjModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Immigration Adjustment Module";}
+	virtual void setModelProvider(IModelProvider* mp) {}
+	virtual void clear();		
+	virtual void setScenario(Scenario* scen);
+
+protected:
+	void zeroWts();
+	ITable* netMigration;
+	unsigned int yr_step; ///< the scenario time step
+};

### 2019-11-22 FEM_CPP/InitVarDelta.cpp

@@ -1,0 +1,126 @@
+#include "InitVarDelta.h"
+#include "Logger.h"
+#include "Random.h"
+#include "utility.h"
+#include <sstream>
+InitVarDelta::InitVarDelta(std::string name, Vars::Vars_t v, unsigned int intervention_id, bool r, ITimeSeriesProvider* tp, IVariableProvider* vp)  
+  : Intervention(intervention_id, tp, vp), _name(name), var(v), reduce(r)
+{
+	param_name = "p_";
+	param_name += VarsInfo::labelOf(var);
+	param_name += reduce ? "_reduce" : "_increase";
+
+	prob = 0.5;
+	params_map[param_name] = "0.5";
+
+	elig_var_param_name = "elig_p_" ;
+	elig_var_param_name += VarsInfo::labelOf(var); 
+	elig_var_param_name += reduce ? "_reduce" : "_increase";
+	params_map[elig_var_param_name] = "true";
+	elig_var = vp->get("true");
+
+	prepCategoricalVars();
+}
+
+InitVarDelta::InitVarDelta(Vars::Vars_t v, unsigned int intervention_id, bool r, ITimeSeriesProvider* tp,  IVariableProvider* vp)  
+  : Intervention(intervention_id, tp, vp), var(v), reduce(r)
+{
+	_name = "Init";
+	std::string tmp = VarsInfo::labelOf(var).substr(0,1);
+	_name += StringToUpper(tmp);
+	_name += VarsInfo::labelOf(var).substr(1);
+	_name += reduce ? "Reduction" : "Increase";
+
+	param_name = "p_";
+	param_name += VarsInfo::labelOf(var);
+	param_name += reduce ? "_reduce" : "_increase";
+
+	prob = 0.5;
+	params_map[param_name] = "0.5";
+	
+	elig_var_param_name = "elig_p_" ;
+	elig_var_param_name += VarsInfo::labelOf(var); 
+	elig_var_param_name += reduce ? "_reduce" : "_increase";
+	params_map[elig_var_param_name] = "true";
+	elig_var = vp->get("true");
+
+	prepCategoricalVars();
+}
+void InitVarDelta::prepCategoricalVars() {
+	
+	is_category_dummy = false;
+	other_category_dummies[0] = Vars::_NONE;
+	int index = 0;
+	if(VarsInfo::infoOf(var).dummy_for != Vars::_NONE) {
+		is_category_dummy = true;
+		for(unsigned int i = 0; i < Vars::NVars; i++)
[…]

### 2019-11-22 FEM_CPP/InitVarDelta.h

@@ -1,0 +1,35 @@
+#pragma once
+#include "Intervention.h"
+/** This class produces a change in the prevalence of variables for the newly entered individuals.
+
+\bug The eligibility variable cannot be a constant value, it assumes it must be a variable
+*/
+class InitVarDelta:
+	public Intervention
+{
+public:
+  InitVarDelta(std::string name, Vars::Vars_t var, unsigned int intervention_id, bool reduce, ITimeSeriesProvider* tp,  IVariableProvider* vp);
+  InitVarDelta(Vars::Vars_t var, unsigned int intervention_id, bool reduce, ITimeSeriesProvider* tp,  IVariableProvider* vp);
+	virtual ~InitVarDelta(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random) ;
+	virtual void setScenario(Scenario* scen);
+
+	virtual inline std::string name() const { return _name;}	
+
+protected:
+
+	void prepCategoricalVars();
+
+	std::string _name;
+	std::string param_name;
+	std::string elig_var_param_name;
+	Vars::Vars_t var;
+	IVariable* elig_var;
+	double prob;
+	bool reduce;
+
+	// To deal with categorical variable dummies
+	bool is_category_dummy;
+	Vars::Vars_t other_category_dummies[20];
+
+};

### 2019-11-22 FEM_CPP/Intervention.h

@@ -1,0 +1,77 @@
+#pragma once
+#include <map>
+#include "Person.h"
+#include "PersonVector.h"
+#include "Scenario.h" 
+#include "Random.h"
+#include "Model.h"
+#include "TimeSeries.h"
+#include "Variable.h"
+
+#include <set>
+
+class Intervention
+{
+public:
+  Intervention(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~Intervention(void) {}
+
+	/** Resets the intervention when a simulation is finished, clearing out any state information */
+	virtual void reset();
+	
+	/** Apply the intervention to persons for the year 
+		\param[in,out] persons The persons to apply the intervention to
+		\param[in] year The year the intervention is being applied
+		\param[in] random The Random number provider 
+	*/
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random)= 0;
+
+	/** Hook called prior to running any interventions.
+		If more than one intervention has been specified, this hook allows the intervention
+		to gather any state information before other interventions destroy that state information.
+		By default, does nothing.
+		\param[in,out] persons The persons in the simulation
+		\param[in] year The year the intervention is being applied
+		\param[in] random The Random number provider 
+	*/
+	virtual void preIntervetion(PersonVector& persons, unsigned int year, Random* random) {}
+
+	/** Hook called after running all interventions.
+		Allows the intervention to access and save any information about persons, after all 
+		interventions have been run.
+		By default, does nothing.
+		\param[in,out] persons The persons
+		\param[in] year The year the intervention is being applied
+		\param[in] random The Random number provider 
+	*/
+	virtual void postIntervetion(PersonVector& persons, unsigned int year, Random* random) {}
+
+	virtual void describe(std::ostream& strm) const;
+	virtual std::string name() const = 0;
+	virtual void setScenario(Scenario* scen);
+	virtual void setModelProvider(IModelProvider* modp) {mp = modp; }
+
+	virtual void yearEndHook(Scenario* scenario, Random* random, unsigned int year);
+
+protected: 
+	std::map<std::string, std::string> params_map;
+	unsigned int intervention_id;
+
+	ITimeSeriesProvider* tp; ///< time series provider for the interventions that require it
+	IVariableProvider* vp; ///< variable provider for the interventions that require it
+	IModelProvider* mp; ///< model provider for the interventions that require it
+
[…]

### 2019-11-22 FEM_CPP/InterventionFactory.cpp

@@ -1,0 +1,148 @@
+#include "InterventionFactory.h"
+#include "InitVarDelta.h"
+#include "VarProbIntervention.h"
+#include "OrderedVarMultIntervention.h"
+#include "DelayIntervention.h"
+#include "CureIntervention.h"
+#include "WorkTillMedicareIntervention.h"
+#include "BariatricSurg.h"
+#include "WeightLossPill.h"
+#include "ReduceDrinkDays.h"
+#include "SmokeStopIntervention.h"
+
+InterventionFactory::InterventionFactory(IVariableProvider* vp, ITimeSeriesProvider* tp)
+{
+	unsigned int index = Vars::NVars;
+
+	Vars::Vars_t init_vars[] = {
+		Vars::cancre,
+		Vars::diabe,
+		Vars::hibpe,
+		Vars::hearta,	
+		Vars::hearte,
+		Vars::lunge,
+		Vars::stroke,
+		Vars::dbclaim,
+		Vars::ssclaim,
+		Vars::nhmliv,
+		Vars::diclaim,
+		Vars::ssiclaim,
+		Vars::anyhi,
+		Vars::work,
+		Vars::wlth_nonzero,
+		Vars::smoken,
+		Vars::smokev,
+		Vars::iadl1,
+		Vars::iadl2p,
+		Vars::adl1,
+		Vars::adl2,
+		Vars::adl3p,
+		Vars::hsless,
+		Vars::college,
+		Vars::memrye,
+		Vars::alzhmr,
+		Vars::drink,
+		Vars::drinkd_e
+	};
+
+	int nInitVars = sizeof(init_vars)/sizeof(Vars::Vars_t);
+	Intervention* inter = NULL;
+	for(int i = 0; i < nInitVars; i++) {
+	  inter = new InitVarDelta(init_vars[i], index++, true, tp, vp);
+		interventions[inter->name()] = inter;
+		inter = new InitVarDelta(init_vars[i], index++, false, tp, vp);
+		interventions[inter->name()] = inter;
+	}
+
+	Vars::Vars_t vars[] = {
+		Vars::cancre,
+		Vars::diabe,
+		Vars::hibpe,
+		Vars::hearte,
+		Vars::hearta,	
+		Vars::lunge,
+		Vars::nhmliv,
+		Vars::stroke,
+		Vars::died,
+		Vars::alzhmr,
+		Vars::memrye,
+		Vars::work, 
[…]

### 2019-11-22 FEM_CPP/InterventionModule.h

@@ -1,0 +1,23 @@
+#pragma once
+#include "Module.h"
+#include "Intervention.h"
+#include "InterventionFactory.h"
+#include <vector>
+
+class InterventionModule :
+	public Module
+{
+public:
+	InterventionModule(InterventionFactory* f);
+	virtual ~InterventionModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Interventions Module";}
+	virtual void setModelProvider(IModelProvider* mp) {}
+	virtual void setScenario(Scenario* scen);
+	void reset_intervetions();
+	virtual void yearEndHook(Scenario* scenario, Random* random, unsigned int year);
+
+protected:
+	std::vector<Intervention*> interventions;
+	InterventionFactory* factory;
+};

### 2019-11-22 FEM_CPP/LogHandler.cpp

@@ -1,0 +1,92 @@
+#include "LogHandler.h"
+#include <time.h>
+
+#include "fem_exception.h"
+
+StreamLogHandler::StreamLogHandler(std::ostream& o) : ostrm(&o) { }
+StreamLogHandler::StreamLogHandler(std::ostream& o, LogLevel level) : LogHandler(level), ostrm(&o) { }
+StreamLogHandler::~StreamLogHandler(void) {}
+
+void StreamLogHandler::log(std::string msg, LogLevel level) {
+	if(level >= log_level)
+		*ostrm << msg << std::endl;
+	ostrm->flush();
+}
+
+FileLogHandler::FileLogHandler(const char* file) : outf(file) 
+{
+	if( outf.bad() || outf.fail())
+	  throw fem_exception("Could not open file "+std::string(file));
+	stream_handler = new StreamLogHandler(outf);
+}
+
+FileLogHandler::FileLogHandler(const char* file, LogLevel level) : LogHandler(level), outf(file) 
+{
+	if( outf.bad() || outf.fail())
+	  throw fem_exception("Could not open file "+std::string(file));
+	stream_handler = new StreamLogHandler(outf, level);
+}
+
+FileLogHandler::~FileLogHandler(void) {
+	delete stream_handler;
+	outf.close();
+}
+
+void FileLogHandler::log(std::string msg, LogLevel level) {
+	if(level >= log_level)
+		stream_handler->log(msg, level);
+}
+
+
+FormattingLogHandler::FormattingLogHandler(LogHandler* lh) : logHandler(lh) { }
+FormattingLogHandler::FormattingLogHandler(LogHandler* lh, LogLevel level) : LogHandler(level), logHandler(lh) { }
+
+FormattingLogHandler::~FormattingLogHandler(void) {
+	delete logHandler;
+}
+
+void FormattingLogHandler::log(std::string msg, LogLevel level) {
+	if(level >= log_level)
+		logHandler->log(format(msg, level), level);
+}
+
+
+TimeFormattingLogHandler::TimeFormattingLogHandler(LogHandler* logHandler) : FormattingLogHandler(logHandler) {}
+TimeFormattingLogHandler::TimeFormattingLogHandler(LogHandler* logHandler, LogLevel level) : FormattingLogHandler(logHandler, level) {}
+TimeFormattingLogHandler::~TimeFormattingLogHandler(void) {}
+
[…]

### 2019-11-22 FEM_CPP/LogHandler.h

@@ -1,0 +1,104 @@
+#pragma once
+#include <string>
+#include <iostream>
+#include <fstream>
+
+enum LogLevel {
+	ALL,
+	FINEST,
+	FINER,
+	FINE,
+	INFO,
+	WARNING,
+	ERROR
+};
+
+class LogHandler
+{
+public:
+	LogHandler(void) {log_level = ALL;}
+	LogHandler(LogLevel level) {log_level = level;}
+	virtual ~LogHandler(void) {}
+	virtual	void log(std::string msg, LogLevel level) = 0;
+	inline virtual void log(const char* msg, LogLevel level) {log(std::string(msg), level);}
+
+protected:
+	LogLevel log_level;
+};
+
+class StreamLogHandler : public LogHandler 
+{
+public:
+	StreamLogHandler(std::ostream& ostrm);
+	StreamLogHandler(std::ostream& ostrm, LogLevel level);
+	virtual ~StreamLogHandler(void);
+	virtual	void log(std::string msg, LogLevel level);
+
+protected:
+	std::ostream* ostrm;
+};
+
+
+class FileLogHandler : public LogHandler 
+{
+public:
+	FileLogHandler(const char* file);
+	FileLogHandler(const char* file, LogLevel level);
+	virtual ~FileLogHandler(void);
+	virtual	void log(std::string msg, LogLevel level);
+
+protected:
+	StreamLogHandler* stream_handler;
+	std::ofstream outf;
+};
+
+
+class FormattingLogHandler : public LogHandler 
+{
+public:
+	FormattingLogHandler(LogHandler* logHandler);
+	FormattingLogHandler(LogHandler* logHandler, LogLevel level);
+	virtual ~FormattingLogHandler(void);
+	virtual	void log(std::string msg, LogLevel level);
+
+protected:
+	virtual std::string format(std::string& msg, LogLevel level) = 0;
+	LogHandler* logHandler;
+};
+
+class PrefixFormattingLogHandler : public FormattingLogHandler 
+{
+public:
+	PrefixFormattingLogHandler(LogHandler* logHandler, std::string pfx) : FormattingLogHandler(logHandler), prefix(pfx) {}
+	PrefixFormattingLogHandler(LogHandler* logHandler, LogLevel level, std::string pfx) : FormattingLogHandler(logHandler, level), prefix(pfx)  {}
+	virtual ~PrefixFormattingLogHandler(void) {}
+
+protected:
[…]

### 2019-11-22 FEM_CPP/Logger.cpp

@@ -1,0 +1,19 @@
+#include "Logger.h"
+
+Logger* Logger::logger = 0;
+
+void Logger::_log(std::string& msg, LogLevel level) {
+	if(level >= log_level)
+		for(unsigned int i = 0; i < handlers.size(); i++)
+			handlers[i]->log(msg, level);
+}
+
+void Logger::_log(const char* msg, LogLevel level) {
+	log(std::string(msg), level);		
+}
+
+Logger& Logger::get() {
+	if(logger == NULL)
+		logger = new Logger();
+	return *logger;
+}

### 2019-11-22 FEM_CPP/Logger.h

@@ -1,0 +1,33 @@
+#pragma once
+#include "LogHandler.h"
+#include <string>
+#include <vector>
+
+
+
+
+class Logger
+{
+public:
+	~Logger(void) {}
+	static inline void addHandler(LogHandler* handler) {Logger::get()._addHandler(handler);}
+	static inline void log(std::string& msg, LogLevel level) {Logger::get()._log(msg, level);}
+	static inline void log(std::string msg, LogLevel level) {Logger::get()._log(msg, level);}
+	static inline void log(const char* msg, LogLevel level) {Logger::get()._log(msg, level);}
+	static inline void setLogLevel(LogLevel level) {Logger::get()._setLogLevel(level);}
+	static Logger& get();
+
+protected:
+	Logger(void) {log_level = INFO;}
+	Logger(LogLevel level) {log_level = level;}
+
+	inline void _addHandler(LogHandler* handler) {handlers.push_back(handler);}
+	void _log(std::string& msg, LogLevel level);
+	void _log(const char* msg, LogLevel level);
+	inline void _setLogLevel(LogLevel level) {log_level = level;}
+
+	std::vector<LogHandler*> handlers;
+	LogLevel log_level;
+	static Logger* logger;
+
+};

### 2019-11-22 FEM_CPP/LogitRegression.h

@@ -1,0 +1,20 @@
+#pragma once
+#include "Regression.h"
+#include <math.h>
+
+class LogitRegression :
+	public Regression
+{
+public:
+
+	LogitRegression(void);
+	LogitRegression(const LogitRegression& source);
+	virtual ~LogitRegression(void) 	{	}
+    virtual void predict(Person* person, const Random* random) const;
+    virtual void predictWithProb(Person* person, const Random* random, double prob) const;
+	virtual inline IModel* clone() const { return new LogitRegression(*this);}
+	virtual inline std::string getType() const { return "logit";}
+	virtual inline std::string getTypeDesc() const { return "logit regression";}
+protected:
+	virtual double transform(double x) const {return 1.0/(1+exp(-x));}
+};

### 2019-11-22 FEM_CPP/Lookup.h

@@ -1,0 +1,126 @@
+#pragma once
+#include <string>
+#include "TimeSeries.h"
+#include "Person.h"
+
+class INode;
+
+typedef double (*lookupFunction_1param)(double);
+typedef double (*lookupFunction_2param)(double, double);
+
+class Lookup {
+public:
+  virtual ~Lookup() {}
+	virtual unsigned int getNParams() const = 0;
+	virtual std::string getName() const = 0;
+	virtual double lookup(INode* const x[], unsigned int nInputs, const Person* person) const = 0;
+	virtual Lookup* clone() const = 0;
+};
+
+class LookupPlaceHolder : public Lookup
+{
+public:
+	LookupPlaceHolder(std::string name, unsigned int n);
+
+	virtual unsigned int getNParams() const {return nparams;}
+	virtual std::string getName() const {return name;}
+	virtual double lookup(INode* const x[], unsigned int nInputs, const Person* person) const ;
+	virtual Lookup* clone() const { return new LookupPlaceHolder(name, nparams);}
+
+
+protected:
+	unsigned int nparams;
+	std::string name;
+};
+
+class TimeSeriesLookup : public Lookup    {
+protected:
+	ITimeSeries* series;
+	std::string placeHolderName;
+public:
+
+	virtual unsigned int getNParams() const {return 1;}
+	virtual std::string getName() const;
+
+
+	virtual double lookup(INode* const x[], unsigned int nInputs, const Person* person) const;
+
+	TimeSeriesLookup(ITimeSeries* t)   {   series = t;     }
+
+	TimeSeriesLookup(std::string name)     {           placeHolderName = name;    series = NULL;    }
+	TimeSeriesLookup(const TimeSeriesLookup& source)     {           placeHolderName = source.placeHolderName;    series = source.series;    }
+
+	void setTimeSeries(ITimeSeries* its)        {            series = its;        }
+
+	void clearTimeSeries()  ;
+	virtual Lookup* clone() const {return new TimeSeriesLookup(*this);}
+};
+
+class BasicOneParamLookup : public Lookup
+{
+protected:
+	std::string name;
+	unsigned int nparams;
+	lookupFunction_1param func;
+
+public:
+	BasicOneParamLookup(std::string name, lookupFunction_1param func);
+
+	virtual inline unsigned int getNParams() const {return nparams;}
+	virtual inline std::string getName() const {return name;}
+
+	virtual inline double lookup(INode* const x[], unsigned int nInputs, const Person* person) const;
[…]

### 2019-11-22 FEM_CPP/MedCostsModule.cpp

@@ -1,0 +1,679 @@
+#include "MedCostsModule.h"
+#include "Logger.h"
+#include "SummaryModule.h"
+#include <sstream>
+#include "fem_exception.h"
+#include "ConstantTimeSeries.h"
+#include "EquationParser.h"
+#include "GlobalVariable.h"
+
+MedCostsModule::MedCostsModule(IVariableProvider* vp, ITimeSeriesProvider *tsp, NodeBuilder* bldr)
+{
+	/* Variables to hold predicted costs using the MCBS dataset for medicare elig */
+	cost_vars_mcbs.push_back(Vars::totmd_mcbs);
+	cost_vars_mcbs.push_back(Vars::caidmd_mcbs);
+	cost_vars_mcbs.push_back(Vars::oopmd_mcbs);
+	cost_vars_mcbs.push_back(Vars::mcare);
+	cost_vars_mcbs.push_back(Vars::mcare_pta);
+	cost_vars_mcbs.push_back(Vars::mcare_ptb);
+	cost_vars_mcbs.push_back(Vars::mcare_ptd);
+	
+	/* Variables to hold predicted costs using the MEPS dataset for non medicare elig*/
+	cost_vars_meps.push_back(Vars::totmd_meps);
+	cost_vars_meps.push_back(Vars::caidmd_meps);
+	cost_vars_meps.push_back(Vars::oopmd_meps);
+
+	/* Create convenience vector, cost_vars, that holds all cost variables */
+	for(unsigned int i = 0; i < cost_vars_mcbs.size(); i++)
+		cost_vars.push_back(cost_vars_mcbs[i]);
+	for(unsigned int i = 0; i < cost_vars_meps.size(); i++)
+		cost_vars.push_back(cost_vars_meps[i]);
+
+
+	event_vars.push_back(Vars::hspnit);
+	event_vars.push_back(Vars::hsptim);
+	event_vars.push_back(Vars::doctim);
+
+	medicare_elig_var = vp->get("medicare_eligibility");
+	frac_medicare_elig_var = vp->get("frac_medicare_elig");
+		
+	mcare_ptb_coin_elas = vp->get("mcare_ptb_coin_elas");
+	mcare_ptb_coin_chg = vp->get("mcare_ptb_coin_chg");
+	poverty_level = vp->get("poverty_level");
+	
+	// The summary_module will get set (and reset) in FEM.cpp for each new scenario
+	summary_module = NULL;
[…]

### 2019-11-22 FEM_CPP/MedCostsModule.h

@@ -1,0 +1,208 @@
+#pragma once
+#include "Module.h"
+#include "Model.h"
+#include "EquationNode.h"
+#include "SummaryMeasure.h"
+#include <map>
+#include <vector>
+
+
+	
+
+/** This class does regular medical costs, Medicare and Medicaid costs, and Medicare enrollment.
+
+\bug The Medicare premium calculations, and hence the enrollment calculations, depend on a full population
+of Medicare enrollees to dynamically compute the premiums. In a cohort simulation, even without any of the
+Medicar reform options enabled, Part B enrollment will likely be slightly higher than normal. Not sure
+what we can do about this.
+*/
+class MedCostsModule :
+	public Module
+{
+public:
+	/** Initializes and Creates the Medical Costs Module.
+		\param[in]  vp  The runtime variable provider used to obtain the medicare eligibility variable
+		\param[in] tsp The Time Series provider used to obtain medical cost growth and cpi
+		\param[in] bldr The NodeBuilder object used to build derived variables if necessary
+	*/
+  MedCostsModule(IVariableProvider* vp, ITimeSeriesProvider *tsp, NodeBuilder* bldr);
+	virtual ~MedCostsModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Medical Costs Module";}
+	virtual void setModelProvider(IModelProvider* mp);
+	virtual void setScenario(Scenario* scen);
+
+protected:
+	//bool IsMedicareElig(Person* p) {return p->get(Vars::age) >= medicare_elig_age || (p->test(Vars::diclaim) && p->test(Vars::ldiclaim));}
+	inline bool IsMedicareElig(Person* p) const {return medicare_elig_var->value(p) == 1.0;}
+	
+	/** Models for predicted costs based on the MCBS dataset */
+	std::vector<IModel*> cost_models_mcbs; 
+
+	/** Models for predicted costs based on the MEPS dataset */
+	std::vector<IModel*> cost_models_meps; 
+	
+	/** Models for predicting hospital and doctor events based on the MCBS dataset */
+	std::vector<IModel*> event_models_mcbs; 
+
+	/** Models for predicting hospital and doctor events based on the MEPS dataset */
+	std::vector<IModel*> event_models_meps; 
+
+	/** Model for probability of enrolling into Part D. */
[…]

### 2019-11-22 FEM_CPP/Model.h

@@ -1,0 +1,139 @@
+#pragma once
+
+#include <iostream>
+#include <istream>
+#include "Person.h"
+#include "Variable.h"
+#include "Random.h"
+
+/** Interface for all predictive models
+	A predictive model determines the value of a global or person specific variable based on the variables of that person
+	or other global variables. Subclasses implement specific model types, such as regression, probit, and ordered probit models.
+	An instance of a model must have a unique name, and this name is used by other modules to obtain a handle to the model.
+	Model instances can either be parameterized through external text files loaded at runtime, while more complex models can
+	be instantiated and setup during compile time. For models loaded at runtime through external text files, the definition of 
+	the model is contained within a plain text file named <code><i>model_name</i>.est</code>. Each model type might require 
+	a different format to this text file, however the general format is:
+	<pre>
+	model_type
+	predicted variable name
+	var<sub>1</sub> beta<sub>1</sub>
+	...  ... 
+	var<sub>N</sub> beta<sub>N</sub>
+	</pre>
+	For models that are created during runtime for external files, the model type specified in the external file is used
+	to look up the protoype model object, which is then cloned and used to read the rest of the model specification in the
+	external file.
+	
+	The primary access point for a model is the predict() method, which applies the model to a specific person and stores the result into
+	the predicted variable for that person. For some models, such as probits, a Random number generator is required to make a decision 
+	on the value of the dependent variable. The estimate() method is also useful for obtaining the predicted value, without storing the result
+	into the person object.
+*/
+class IModel
+{
+public:
+	virtual ~IModel(void) {}
+
+	/** Predicts the value of the modeled variable and stores it 
+		The results are stored back into the \a person. 
+		\param[in,out]  person  The person to use to both determine regressor values and store the predicted value into
+		\param[in] random The random number source to use if this model requires random numbers 
+	*/		
+    virtual void predict(Person* person, const Random* random) const = 0;
+
+	/** 
+	Simulate outcome using the prob argument as the probability. 
+	Store outcome in appropriate variable for \a person
+	Does NOT compute the model probability for the outcome before simulating.
+	\param[in]  person  The person to use to compute the probability
+	\param[in] random Used to simulate a random draw from the model
+	\param[in] prob The probability to use for simulating the outcome
+	*/
+	virtual void predictWithProb(Person* person, const Random* random, double prob) const = 0;
+
[…]

### 2019-11-22 FEM_CPP/MultinomialLogit.cpp

@@ -1,0 +1,175 @@
+#include "MultinomialLogit.h"
+#include <sstream>
+#include <limits>
+#include <cstring>
+#include "Logger.h"
+#include "Vars.h"
+
+MultinomialLogit::MultinomialLogit() {
+  dummy_vars = NULL;
+};
+
+MultinomialLogit::MultinomialLogit(const MultinomialLogit& source) : Regression(source) {
+  regs.clear();
+  for(std::vector<Regression*>::const_iterator i = source.regs.begin();
+      i != source.regs.end();
+      i++)
+    regs.push_back(*i);
+
+  dummy_vars = new Vars::Vars_t[regs.size()];
+  for(unsigned int i=0; i < regs.size(); i++)
+    dummy_vars[i] = source.dummy_vars[i];
+}
+
+MultinomialLogit::~MultinomialLogit(void) {
+  regs.clear();
+}
+
+void MultinomialLogit::read(std::istream& istrm, IVariableProvider* provider) {
+  double temp_coeffs[1000];
+  double temp_perts[1000];
+  IVariable* temp_vars[1000];
+  std::string buf;
+  char bufline[5000];
+  std::string firstvar("");
+
+  istrm >> buf;
+  predicted_var = VarsInfo::indexOf(buf);
+  
+  // Check if the variable read in exists
+  if (predicted_var == Vars::_NONE) {
+    std::ostringstream ss;
+    ss << "Error reading model " << getName() << ": Predicted variable [" << buf << "] does not exists!";
+    Logger::log(ss.str(), ERROR);
+  }
+  
+  // Eat any newline characters
+  istrm.ignore(std::numeric_limits<std::streamsize>::max(), '\n');
+  
+  try {
+    /* Read in coeffecients */
+    while(!istrm.eof())
+      {
+	// Read a line
+	istrm.getline(bufline, 5000);
+	// Check that a full line was read, and that it is not a comment line 
+	if(strlen(bufline) > 0 && bufline[0] != '|')  {
+	  // It is a regular variable coeffecient line.
+	  std::istringstream iss(bufline);
+	  iss >>buf >> temp_coeffs[nvars];
+	  temp_perts[nvars] = 0.0;
+	  if(iss.fail()) {
+	    std::string tstring = std::string(bufline);
+	    if(tstring.find("(dropped)")==std::string::npos) {
+	      // Something bad happened trying to read the data. 
+	      // Most likely, it tried to read the coeff but it wasnt a number
[…]

### 2019-11-22 FEM_CPP/MultinomialLogit.h

@@ -1,0 +1,33 @@
+#pragma once
+#include "Regression.h"
+#include <math.h>
+#include <vector>
+
+/** This class implements a multinomial logit regression for unordered choices.
+
+    \todo Add a unit test to confirm that this class does what it should.
+*/
+class MultinomialLogit :
+	public Regression
+{
+public:
+
+	MultinomialLogit(void);
+	MultinomialLogit(const MultinomialLogit& source);
+	virtual ~MultinomialLogit(void);
+	/** The read method for the MultinomialLogit
+
+	    \todo Update this method to handle the Stata interacted terms format
+	*/
+	virtual void read(std::istream& inf, IVariableProvider* provider);
+	virtual void predict(Person* person, const Random* random) const;
+	virtual double estimate(const Person* person) const {throw fem_exception("MLogit cannot provide an estimate");}
+	void predictWithProb(Person* person, const Random* random, double prob) const {throw fem_exception("MLogit cannot predictWithProb");}
+	virtual inline IModel* clone() const { return new MultinomialLogit(*this);}
+	virtual inline std::string getType() const { return "mlogit";}
+	virtual inline std::string getTypeDesc() const { return "multinomial logit regression";}
+	virtual std::string describe() const;
+protected:
+	std::vector<Regression*> regs;
+	Vars::Vars_t* dummy_vars;
+};

### 2019-11-22 FEM_CPP/OrderedVarMultIntervention.h

@@ -1,0 +1,45 @@
+#pragma once
+#include "Intervention.h"
+#include "OrderedProbitRegression.h"
+#include "TimeSeries.h"
+#include "Variable.h"
+
+/** This intervention adjusts the level of an ordered outcome through a multiplier.
+This assumes that the ordered outcome is a ratio scale variable (e.g., a count) 
+with values 0,1,2,...,K and these values are coded in FEM as 1,2,3,...,K+1. The 
+probability of intervening is determined by the value of the eligibility variable 
+specified in the scenario settings.  The multiplier used for intervention, also 
+specified in the scenario settings, can be a constant value or the name of a time 
+series or variable in the providers passed to the constructor.
+*/
+class OrderedVarMultIntervention:
+	public Intervention
+{
+public:
+  OrderedVarMultIntervention(std::string name, Vars::Vars_t var, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+  OrderedVarMultIntervention(Vars::Vars_t var, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~OrderedVarMultIntervention(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random) ;
+	virtual void setScenario(Scenario* scen);
+
+	virtual inline std::string name() const { return _name;}	
+
+protected:
+
+
+	std::string _name;
+	std::string param_name;
+	Vars::Vars_t var;
+	IVariable* const_mult;
+	ITimeSeries* ts_mult;
+	bool use_ts;
+	std::string elig_var_param_name;
+	IVariable* elig_var;
+	// model used to transtion var
+	OrderedProbitRegression* var_model;
+	// maximum possible value for var
+	unsigned int var_max;
+	// indicator variables for the levels of var
+  std::vector<Vars::Vars_t> dummy_vars;
+
+};

### 2019-11-22 FEM_CPP/OutputModules.h

@@ -1,0 +1,15 @@
+#pragma once
+#include "AbstractOutputModule.h"
+#include "Variable.h"
+#include <vector>
+class OutputModules : public Module
+{
+public:
+	OutputModules(IVariableProvider *vp, std::string output_types, std::string output_var_names);
+	virtual ~OutputModules(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const;
+	virtual void setScenario(Scenario* scen);
+private:
+	std::vector<AbstractOutputModule*> outputs;
+};

### 2019-11-22 FEM_CPP/OutputTextModule.h

@@ -1,0 +1,14 @@
+#pragma once
+#include "AbstractOutputModule.h"
+#include "Variable.h"
+#include <vector>
+class OutputTextModule : public AbstractOutputModule
+{
+public:
+	OutputTextModule(IVariableProvider *vp);
+	OutputTextModule(IVariableProvider *vp, std::vector<std::string> var_names);
+	virtual ~OutputTextModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Output Text Module";}
+
+};

### 2019-11-22 FEM_CPP/PersonPool.cpp

@@ -1,0 +1,58 @@
+#include "PersonPool.h"
+#include "Logger.h"
+#include <sstream>
+PersonPool* PersonPool::_pool = NULL;
+
+PersonPool::PersonPool(size_t size)
+{
+	nPersons = size;
+	nFreePersons = nPersons;
+	nUsedPersons = 0;
+	free_persons.reserve(nPersons);
+	for(size_t i = 0; i < nFreePersons; i++)
+		free_persons.push_back(new Person());
+}
+
+PersonPool::~PersonPool(void)
+{
+ 	if(nUsedPersons > 0) {
+		std::ostringstream ss;
+		ss << "Deleting Persons Pool, but there are still " << nUsedPersons << " persons in use.";
+		Logger::log(ss.str(), WARNING);
+		ss.str("");
+		ss << "This memory will be lost.";
+		Logger::log(ss.str(), WARNING);
+	}
+	for(size_t i = 0; i < nFreePersons; i++)
+		delete free_persons[i];
+}
+
+
+Person* PersonPool::_newPerson() {
+	if(nFreePersons == 0) {
+		// Double the size by creating nPerson number of free persons;
+		std::ostringstream ss;
+		ss << "Growing Persons Pool by " << nPersons << " persons";
+		Logger::log(ss.str(), WARNING);
+		nFreePersons = nPersons;
+		for(size_t i = 0; i < nFreePersons; i++)
+			free_persons.push_back(new Person());
+		nPersons += nFreePersons;
+	}
+	Person* p = free_persons.back();
+	free_persons.pop_back();
+	inuse_persons.insert(p);
+	nFreePersons--;
+	nUsedPersons++;
+	p->clear();
+	return p;
+
+}
+
+void PersonPool::_releasePerson(Person* p) {
+	p->clear();
+	inuse_persons.erase(p);
+	free_persons.push_back(p);
+	nFreePersons++;
+	nUsedPersons--;
+}

### 2019-11-22 FEM_CPP/PersonPool.h

@@ -1,0 +1,37 @@
+#pragma once
+#include "Person.h"
+#include <set>
+
+class PersonPool
+{
+public:
+	~PersonPool(void);
+
+	inline static Person* newPerson() {return get()->_newPerson();}
+	inline static void releasePerson(Person* p) {get()->_releasePerson(p);}
+	
+	inline static PersonPool* get() {
+		if(_pool == NULL)
+			initPool();
+		return _pool;
+	}
+
+	inline static void initPool(size_t size) {	_pool = new PersonPool(size);	}
+	inline static void initPool() {	initPool(DEFAULT_SIZE);	}
+	inline static void deletePool() {if(_pool != NULL) delete _pool;}
+
+protected:
+
+	static PersonPool* _pool;
+
+	PersonPool(size_t size);
+	Person* _newPerson();
+	void _releasePerson(Person* p);
+
+	size_t nPersons;
+	size_t nFreePersons;
+	size_t nUsedPersons;
+	std::set<Person*> inuse_persons;
+	std::vector<Person*> free_persons;
+	static const int DEFAULT_SIZE = 50000;
+};

### 2019-11-22 FEM_CPP/PersonVector.h

@@ -1,0 +1,38 @@
+#pragma once
+#include "Person.h"
+#include <vector>
+#include <algorithm>
+
+/** Wrapper class for a vector to hold lots of persons.
+
+*/
+class PersonVector : public std::vector<Person*>
+{
+public:
+	PersonVector(void);
+	virtual ~PersonVector(void);
+	
+	void readStata(const char* file);
+	void readDelimited(const char* file, char delim);
+	void writeDelimited(const char* file, char delim) const;
+
+	void serialize(std::ostream& ostrm) const;
+	void deserialize(std::istream& istrm);
+
+	void serialize(const char* file) const;
+	void deserialize(const char* file);
+
+	virtual void clear();
+	virtual void remove(std::vector<Person*>::iterator pos);
+
+	inline void sortByID() { std::sort(begin(), end(), personComparer); }
+
+	/** A function to check for duplicate IDs. not something to be run frequently, as
+	    it cycles through all persons.
+
+	    @return returns false if all IDs are unique, throws an exception otherwise
+	*/
+	bool checkIDs() const;
+protected:
+	void buildSpouses();
+};

### 2019-11-22 FEM_CPP/ProxyTimeSeries.cpp

@@ -1,0 +1,40 @@
+#include "ProxyTimeSeries.h"
+#include "fem_exception.h"
+
+
+ProxyTimeSeries::ProxyTimeSeries(void)
+{
+	series = NULL;
+}
+
+
+
+double ProxyTimeSeries::Value(unsigned int year) const {
+	checkSeries();
+	return series->Value(year);
+}
+
+unsigned int ProxyTimeSeries::StartYear() const {
+	if(series != NULL)
+	  return series->StartYear();
+	else throw fem_exception(name + " ProxyTimeSeries::StartYear has no underlying series");
+}
+
+unsigned int ProxyTimeSeries::EndYear() const {
+	if(series != NULL)
+	  return series->EndYear();
+	else throw fem_exception(name + " ProxyTimeSeries::EndYear has no underlying series");
+}
+
+ITimeSeries* ProxyTimeSeries::CrossSection(unsigned int year_start, unsigned int year_end) const {
+	checkSeries();
+	return series->CrossSection(year_start, year_end);
+}
+
+void ProxyTimeSeries::checkSeries() const {
+	if(series == NULL) {
+		std::ostringstream ss;
+		ss << "Proxy time series [" << name << "] has no underlying implementation.";
+		throw fem_exception(ss.str());
+	}
+}

### 2019-11-22 FEM_CPP/Random.cpp

@@ -1,0 +1,30 @@
+#include "Random.h"
+#include "utility.h"
+
+std::vector<double> Random::mvnormDist( unsigned int id, const std::vector<unsigned int>& process, unsigned int year, const std::vector<double>& m, const std::vector< std::vector<double> >& C) const {
+	unsigned int size = m.size();
+	
+	// check sizes
+	if(process.size() != size || C.size() != size)
+		throw fem_exception("Dimension mismatch in multivariate normal random generator");
+	
+	std::vector<double> draw(size);
+	std::vector<double> z(size);	
+		
+	for(unsigned int i=0; i < size; i++) {
+		z[i] = normalDist(id, process[i], year, 0, 1);
+	}
+		
+	for(unsigned int i=0; i < size; i++) {
+		draw[i] = m[i];
+		// check that row i of C has length i+1
+		if(C[i].size() != i+1)
+			throw fem_exception("Error in multivariate normal random generator: covariance parameters are not lower triangular");
+
+		for(unsigned int j=0; j <= i; j++) {
+			draw[i] += C[i][j]*z[j];
+		}
+	}
+		
+	return(draw);
+}

### 2019-11-22 FEM_CPP/RandomBasic.h

@@ -1,0 +1,16 @@
+#pragma once
+
+#include "Random.h"
+
+class RandomBasic : public Random
+{
+public:
+	RandomBasic(unsigned int r) : Random(r)  {}
+	virtual ~RandomBasic() {} 
+	
+	virtual double normalDist( unsigned int id,  unsigned int process, unsigned int year, double m, double s ) const;
+	virtual double uniformDist( unsigned int id,  unsigned int random_process, unsigned int year ) const;
+	virtual unsigned int randomIndex(unsigned int id, Random_t rand) const {return 0;}
+	
+	
+};

### 2019-11-22 FEM_CPP/RandomBasicProvider.cpp

@@ -1,0 +1,8 @@
+#include "RandomBasicProvider.h"
+
+Random* RandomBasicProvider::getRandom(unsigned int rep) {
+	if(random != NULL) 
+		delete random;
+	random = new RandomBasic(rep);
+	return random;
+}

### 2019-11-22 FEM_CPP/RandomBlowfish.cpp

@@ -1,0 +1,436 @@
+#include "RandomBlowfish.h"
+#include <climits>
+#include <cmath>
+#include "utility.h"
+
+
+
+// The specification of the blowfish algorithm initializes the S and P
+// boxes mentioned above to the hexadecimal digits of PI. 
+void RandomBlowfish::set_pi()
+{
+  int j;
+  // The hexadecimal digits of pi, arranged as four s_boxes and one p_array, as per the Blowfish default.
+  // From : http://www.schneier.com/code/constants.txt
+
+/*
+   =============================================================
+   The hex digits of pi, arranged as four s_boxes & one p_array,
+   as per the Blowfish default. These have passed muster w/ Eric
+   Young's set of test vectors. Enjoy.  -Mike Schaudies
+   =============================================================
+
+*/
+
+unsigned long sbox0[] = {
+0xd1310ba6L, 0x98dfb5acL, 0x2ffd72dbL, 0xd01adfb7L, 0xb8e1afedL, 0x6a267e96L,
+0xba7c9045L, 0xf12c7f99L, 0x24a19947L, 0xb3916cf7L, 0x0801f2e2L, 0x858efc16L,
+0x636920d8L, 0x71574e69L, 0xa458fea3L, 0xf4933d7eL, 0x0d95748fL, 0x728eb658L,
+0x718bcd58L, 0x82154aeeL, 0x7b54a41dL, 0xc25a59b5L, 0x9c30d539L, 0x2af26013L,
+0xc5d1b023L, 0x286085f0L, 0xca417918L, 0xb8db38efL, 0x8e79dcb0L, 0x603a180eL,
[…]

### 2019-11-22 FEM_CPP/RandomBlowfishProvider.cpp

@@ -1,0 +1,36 @@
+#include "RandomBlowfishProvider.h"
+#include <sstream>
+#include <fstream>
+#include "fem_exception.h"
+
+
+
+RandomBlowfishProvider::~RandomBlowfishProvider(void) {
+	for(unsigned int i = 0; i < random_blowfishes.size(); i++)
+		delete random_blowfishes[i];
+}
+
+void RandomBlowfishProvider::readKeyFile(const char* file) {
+	
+	std::ifstream inf(file);
+	if( inf.bad() || inf.fail())
+	  throw fem_exception("Could not open file "+std::string(file));
+	unsigned int r = 0;
+	while(!inf.eof()) {
+		unsigned int key;
+		inf >> key;
+		random_blowfishes.push_back(new RandomBlowfish(key, r));
+		keys.push_back(key);
+		r++;
+	}
+}
+
+Random* RandomBlowfishProvider::getRandom(unsigned int rep) {
+	if(rep >= this->maxReps()) {
+		std::ostringstream ss;
+		ss << "Requested Random numbers for rep " << rep+1 << ", but only " << maxReps() << " available!";
+		throw fem_exception(ss.str());
+	}
+	return random_blowfishes[rep];
+}
+

### 2019-11-22 FEM_CPP/Regression.h

@@ -1,0 +1,153 @@
+#pragma once
+#include "Model.h"
+
+/** The <i>Regression</i> class serves as a base class for models of the form <code>y = f(<b>X'B</b>)</code>.
+	For the basic regression model, the transformation <code>f(x)=1</code>. Subclasses implement other 
+	transformation types. The values of the regressors <b>X</b> can either be taken from core built-in variables
+	such as <code>age</code> or derived variables such as <code>lage65p</code> (an age spline variable). The model specification
+	is read from a flat text file with the following format:
+	<pre>
+	regress
+	predicted var name
+	var<sub>1</sub> beta<sub>1</sub>
+	...  ... 
+	var<sub>N</sub> beta<sub>N</sub>
+	</pre>
+	The first line indicates that this is a simple regression model and therefore this class should be used. The next line indicates
+	the name of the predicted variable. The next N lines give regressor coeffecient for the N regressors. 
+*/
+class Regression :
+	public IModel
+{
+public:
+	/** Constructs an empty regression model. */
+	Regression(void);
+	
+	/** Constructs an new regression model based on source model. 
+		\param[in]  source  The regression model to copy
+	*/
+	Regression(const Regression& source);
+
+	/** Contructor from data, sets esigma to zero if no argument given */
+	Regression(int nv, Vars::Vars_t pvar, std::string n, const double cfs[],  IVariable* vs[], const double cps[], const double esd = 0.0);
+
+	virtual ~Regression(void);
+
+	/** Predicts the new value of the modeled variable.
+		The results are stored back into the \a person. 
+		\param[in,out]  person  The person to use to both determine regressor values and store the predicted value into
+		\param[in] random Used to draw a normal error with variance given by the error term in the model
+	*/
+    virtual void predict(Person* person, const Random* random) const;
+	
+	/** 
+	Simulate outcome using the prob argument instead of an internally computed probability. 
+	Store outcome in appropriate variable for \a person
+	Does NOT compute the model probability for the outcome before simulating.
+	\param[in]  person  The person to use to compute the probability
+	\param[in] random Used to simulate a random draw from the model
+	\param[in] prob The probability to use for simulating the outcome
+	*/
+	virtual void predictWithProb(Person* person, const Random* random, double prob) const;
+	
+	/** Estimates the new value of the modeled variable.
+		The results are not stored back into the \a person. 
+		\param[in]  person  The person to use to both determine regressor values
[…]

### 2019-11-22 FEM_CPP/SSCalculator.cpp

@@ -1,0 +1,444 @@
+#include "SSCalculator.h"
+#include "fem_exception.h"
+#include <math.h>
+SSCalculator::SSCalculator(ITimeSeriesProvider *timeSeriesProvider)
+{
+	nra = timeSeriesProvider->get("nra");
+	cola = timeSeriesProvider->get("cola");
+	nwi = timeSeriesProvider->get("nwi");
+	drc = timeSeriesProvider->get("drc");
+	eadisreg1 = timeSeriesProvider->get("eadisreg1");
+	eadisreg2 = timeSeriesProvider->get("eadisreg2");
+}
+
+
+SSCalculator::~SSCalculator() {
+}
+
+double SSCalculator::SSBenefit(Person* person, const int cyr, const int rclyr) {
+
+	// Social Security Benefit Function for United States
+	// Benefits are computed as if claimed on 6/1/Claim Year and received through 6/1/(Claim Year+1)
+
+	// Computes all three benefit types: retirement, spousal, and widow
+	// Prefixes indicate what what benefit a variable applies to: (r)etirement, (s)pouses, (w)idow
+
+	// Year and month born
+  const int rbyr = (int) person->get(Vars::rbyr);
+  const int rbmonth = (int) person->get(Vars::rbmonth);
+
+	// Age in months and yrs at claiming
+	const int rage_months = (rclyr - rbyr)*12 + (6-rbmonth);
+	const int rage_yrs = rage_months/12;
+	
+	// Normal retirement age in months and years, based on birth year
+	const int nra_months = (int) nra->Value(rbyr);
+	const int nra_yrs = nra_months/12;
+	
+	// Compute benefit amounts based on the age at the claiming year
+
+	/*** Compute Retirement Benefits ***/
+	double rben = RetBenefit(person, cyr, rclyr);
+	// Apply the COLA adjustment, which is given by the change in the COLA series from age 62 to the current year
+	rben	*= cola->Value(cyr-1) / cola->Value(rbyr+62-1);
+	
+
+	/*** Compute Benefits as Spouse ***/
+
+	double sben = 0.0;
+	
+	// Determine eligibility for spousal retirement benefits
+	// First, must be married
+	if(person->test(Vars::married)) {
+
+		// Get the Spouse
+		const Person* spouse = person->getSpouse();
+
+		// Spouse's year and month of birth
+		const int sbyr = (int) spouse->get(Vars::rbyr);
[…]

### 2019-11-22 FEM_CPP/ScenarioVector.cpp

@@ -1,0 +1,126 @@
+#include "ScenarioVector.h"
+#include "utility.h"
+#include <map>
+#include <fstream>
+#include <sstream>
+#include "fem_exception.h"
+#include <iostream>
+#include <cstdlib>
+#include <cstring>
+
+ScenarioVector::ScenarioVector(void)
+{
+}
+
+ScenarioVector::~ScenarioVector(void)
+{
+	for(unsigned int i = 0; i < size(); i++)
+		delete (*this)[i];
+}
+
+
+
+void ScenarioVector::clear(void)
+{
+	for(unsigned int i = 0; i < size(); i++)
+		delete (*this)[i];
+	std::vector<Scenario*>::clear();
+}
+
+
+void ScenarioVector::readDelimited(const char* file, char delim)
+{
+	std::ifstream istrm(file);
+	if( istrm.bad() || istrm.fail())
+	  throw fem_exception("Could not open file "+std::string(file));
+
+	std::string delim_str(1, delim);
+	const int BUFSIZE = 5000;
+	char buf[BUFSIZE];
+
+	// Read the first line, with column headers
+	istrm.getline(buf, BUFSIZE);
+
+	// Trim off any potential extra whitespaces or carriage returns that may have been thrown in by Excel
+	std::string str(buf);
+	trim(str);
+
+	// Tokenize the string based on the delim, and place the resulting strings in the col_names vector
+	std::vector<std::string> col_names;
+	str_tokenize(str, col_names, delim_str);
+
+	// Change all column names to lower case
+	for(unsigned int i = 0; i < col_names.size(); i++)	{
+		StringToLower(col_names[i]);
+	}
+
+
+	std::vector<std::string> tokens;        
+    while (!istrm.eof()) {
+			// Read a scenario line
+			istrm.getline(buf, BUFSIZE);
+			
+			if(strlen(buf) > 0) {
+				Scenario* s = new Scenario();
+				tokens.clear();
+
+				// Trim off any potential extra whitespaces or carriage returns that may have been thrown in by Excel
+				str = std::string(buf);
+				trim(str);
+
+				// Tokenize to extract the values
+				str_tokenize(str, tokens, delim_str);
+
+				// The number of values should be the same as number of column headers. If not, something is wrong
+				if(tokens.size() != col_names.size()) {
+					std::ostringstream ss;
[…]

### 2019-11-22 FEM_CPP/Settings.h

@@ -1,0 +1,20 @@
+#pragma once
+#include <string>
+#include <map>
+#include <ostream>
+class Settings
+{
+public:
+	Settings(void);
+	~Settings(void);
+
+	void readSettings(const char* file);
+
+	std::string get(std::string name);
+		
+	inline void set(std::string name, std::string val) {params[name] = val;}
+	void dumpSettings(std::ostream& strm);
+
+protected:
+	std::map<std::string, std::string> params;
+};

### 2019-11-22 FEM_CPP/ShiftedProbitRegression.cpp

@@ -1,0 +1,102 @@
+#include "ShiftedProbitRegression.h"
+#include <sstream>
+#include <iomanip>
+#include "fem_exception.h"
+#include "Random.h"
+#include <limits>
+#include <cstring>
+
+ShiftedProbitRegression::ShiftedProbitRegression(void)
+{
+	name = "shifted_probit";
+}
+
+ShiftedProbitRegression::ShiftedProbitRegression(const ShiftedProbitRegression& source) : ProbitRegression(source)
+{
+	this->xb_shifts = source.xb_shifts;
+}
+
+double ShiftedProbitRegression::getShift(unsigned int year) const {
+  if(xb_shifts.count(year) > 0)
+    return xb_shifts.at(year);
+  else {
+    std::ostringstream ss;
+    ss << "ShiftedProbitRegression " << getName() << " does not have shift data for " << year;
+    throw fem_exception(ss.str());
+  }
+}
+
+
+void ShiftedProbitRegression::read(std::istream& istrm, IVariableProvider* provider)
+{
+	double temp_coeffs[1000];
+	IVariable* temp_vars[1000];
+	std::string buf;
+	char bufline[5000];
+	istrm >> buf;
+	predicted_var = VarsInfo::indexOf(buf);
+
+	// Check if the variable read in exists
+	if (predicted_var == Vars::_NONE) {
+		std::ostringstream ss;
+		ss << "Error reading shifted probit model " << getName() << ": Predicted variable [" << buf << "] does not exists!";
+		throw fem_exception(ss.str().c_str());
+	}
+
+	// Eat any newline characters
+	istrm.ignore(std::numeric_limits<std::streamsize>::max(), '\n');
+	
+	bool reading_shifts = false;
+	try {
+		/* Read in coeffecients */
+		while(!istrm.eof())
+		{
+			istrm.getline(bufline, 5000);
+			if(strlen(bufline) > 0 && !(bufline[0]=='o' && bufline[1]=='.'))  {
+				std::istringstream iss(bufline);
+				char firstchar;
+				iss.read(&firstchar, 1);
+				iss.putback(firstchar);
+				if(firstchar == '|') {
+					reading_shifts = true;
+					// eat a line;
+				}
+				if(!reading_shifts) {
+					iss >>buf >> temp_coeffs[nvars];
+					temp_vars[nvars] = provider->get(buf);
+					nvars++;
+
+				} else if(firstchar != '|') {
+					unsigned int year;
+					double shift;
[…]

### 2019-11-22 FEM_CPP/ShiftedProbitRegression.h

@@ -1,0 +1,21 @@
+#pragma once
+#include "ProbitRegression.h"
+#include "utility.h"
+
+class ShiftedProbitRegression :
+	public ProbitRegression
+{
+public:
+
+	ShiftedProbitRegression(void);
+	ShiftedProbitRegression(const ShiftedProbitRegression& source);
+	virtual ~ShiftedProbitRegression(void) 	{	}
+	virtual IModel* clone() const { return new ShiftedProbitRegression(*this);}
+	virtual std::string getType() const { return "shifted_probit";}
+	virtual std::string getTypeDesc() const { return "shifted probit regression";}
+	virtual void read(std::istream& inf, IVariableProvider* provider);
+protected:
+	virtual double calc_xb(const Person* person) const { return ProbitRegression::calc_xb(person) + getShift(person->getYear()); }
+	std::map<unsigned int, double> xb_shifts;
+	double getShift(unsigned int year) const;
+};

### 2019-11-22 FEM_CPP/SummaryModule.cpp

@@ -1,0 +1,846 @@
+#include "SummaryModule.h"
+#include "utility.h"
+#include "EquationParser.h"
+#include "Variable.h"
+#include "Logger.h"
+#include <fstream>
+#include <sstream>
+#include <algorithm>
+#include "fem_exception.h"
+#include <cstring>
+
+const double SummaryModule::MISSING_VAL = 8.9885e+307;
+
+SummaryModule::SummaryModule(const char* file, NodeBuilder* builder, IVariableProvider* vp) {
+	nreps = 0;
+
+	std::ifstream istrm(file);
+	if( istrm.bad() || istrm.fail())
+	  throw fem_exception("Could not open file "+std::string(file));
+	std::string line;
+	char buffer[5000];
+	while(!istrm.eof()) {
+		// Read a line
+		istrm.getline(buffer, 5000);
+
+		// Check if its a valid, non comment, line
+		if(strlen(buffer) > 0 && buffer[0] != '|') {
+			std::vector<std::string> parts;
+			// Trim off any potential extra whitespaces or carriage returns
+			std::string str(buffer);
+			trim(str);
+
+			// Split by commas
+			str_tokenize(str, parts, ",");
+
+			// Trim off any whitespaces
+			for(unsigned int i = 0; i < parts.size(); i++)
+				trim(parts[i]);
+
+			if(parts[0].size() > 32) {
+				Logger::log("Stata might not accept summary measure name with more than 32 characters: " + parts[0],WARNING);
+			}
+
+			// Create a new summary measure based on the definition
+			/** \bug If the parts vector doesn't have enough parts, this causes a segfault */
+			addMeasure( new SummaryMeasure(
+				vp->get(parts[2]), 
+				EquationParser::parseString(parts[6], builder), 
+				parts[0], 
+				parts[1], 
+				atof(parts[4].c_str()), 
+				EquationParser::parseString(parts[5], builder), 
+				parts[3]));
+		}
+	}
+
+	istrm.close();
+	if(numMeasures() > MAX_VARS)
+	  throw fem_exception("Summary Module tried to create more than MAX_VARS measures (remember, each measure also has a sigma)\n\t\tYou can adjust SummaryModule::MAX_VARS upwards and recompile without fear of breaking anything (just consuming more memory)");
+}
+
+void SummaryModule::addMeasure(SummaryMeasure* sm) {
+  if(measures.count(sm->getName()) == 0)
+    measures[sm->getName()] = sm;
+  else
[…]

### 2019-11-22 FEM_CPP/SummaryModule.h

@@ -1,0 +1,59 @@
+#pragma once
+#include "Module.h"
+#include <map>
+#include "SummaryMeasure.h"
+class SummaryModule :
+	public Module
+{
+ public:
+  /** This is the constructor for the SummaryModule class. It creates all the summary measure objects
+      required for this run of the simulation.
+
+      \bug If you swap the weight and the factor values in the summary output text file, 
+      no errors are thrown, but no outputs are generated, either (all values zero due to
+      the atof function processing the string weight for scale and coming up with zero).
+  */
+	SummaryModule(const char* file, NodeBuilder* builder, IVariableProvider* vp) ;
+	virtual ~SummaryModule(void);
+	virtual void process(PersonVector& persons, unsigned int year, Random* random);
+	virtual std::string description() const {return "Summary Module";}
+	virtual void setModelProvider(IModelProvider* mp) {};
+	void scenarioFinished();
+	virtual void setScenario(Scenario* scen);
+	virtual inline unsigned int numMeasures() const {return measures.size();}
+	virtual void output();
+	virtual void outputByRep();
+	void addRepData(unsigned int rep, double* data);	
+	void getRepData(unsigned int rep, double* data) const;
+	double getValue(unsigned int year, std::string name, Random* random) const;
+	bool hasMeasure(std::string name) const {return measures.count(name) > 0;}
+	SummaryMeasure* getMeasure(std::string name);
+	void addMeasure(SummaryMeasure* sm);
+
+	/** Denotes a missing value. Used when there is no persons to compute a measure over */
+	static const double MISSING_VAL;
+protected:
+	std::map<std::string, SummaryMeasure*> measures;
+	std::vector<std::map<unsigned int, std::map<std::string, double> >* > year_measure_val_vec;
+	std::map<unsigned int, std::map<std::string, double> > year_measure_mean;
+	unsigned short nreps;
+
+	void prepStataDescriptors();
+	void prepStataVarLabels();
+
+	void prepStataDescriptorsByRep();
+	void prepStataVarLabelsByRep();
+
+	static const unsigned int BUF_SIZE = 5000;
+	char buf[BUF_SIZE];
+
+	static const unsigned int MAX_VARS = 4000;
+
+	static const unsigned int MAX_DESCRIPTOR = MAX_VARS*118+2;
+	char descriptor[MAX_DESCRIPTOR];
+
+	static const unsigned int MAX_VARLABELS = MAX_VARS*81;
+	char var_labels[MAX_VARLABELS];
+
+
+};

### 2019-11-22 FEM_CPP/Table.h

@@ -1,0 +1,95 @@
+#pragma once
+#include <string>
+#include <vector>
+#include <map>
+#include "Variable.h"
+#include "VarMap.h"
+
+/**
+The TableIndex class defines objects used for indexing entries in derivatives of the ITable class
+*/
+class TableIndex
+{
+	public:
+		TableIndex();
+		
+		TableIndex(IVariableProvider* vp);
+
+		/**
+		Adds element name to TableIndex but does not set a value for the element
+		@param[in] name The name of the index element
+		*/
+		void addName(std::string name) {name2val[name] = 0;}
+
+		/**
+		Sets values of the TableIndex to match values of person
+		@param[in] person The person object from which to get index values
+		*/
+		void set(const Person &person);
+		
+		/**
+		Set value for a single element of TableIndex identified by name
+		@param[in] name The name of the index element to set
+		@param[in] value The value to assign to the index element
+		*/
+		void set(std::string name, double value);
+
+		/**
+		Get value for a single element of TableIndex identified by name
+		@param[in] name The name of the index element to get
+		*/
+		double get(std::string name) const;
+			
+		/**
+		Returns true if TableIndex element values have been modified since creation
+		*/
+		bool isSet() const {return modified;};
+		
+		void setVariableProvider(IVariableProvider* vp) { variableProvider=vp; };
+		
+		/**
+		Returns a vector of strings containing the variable name for each index element
+		*/
+		std::vector<std::string> getNames(void);
+		
+	private:
+		bool modified;
+		std::map<std::string, double> name2val;
+		IVariableProvider* variableProvider;
+};
+
+class ITable
+{
+public:
+	virtual ~ITable(void) {}
+	virtual double Value(const TableIndex &index) const = 0;
+	virtual double Value(const Person &person) const = 0;
+  virtual std::string getName() const = 0;
+  virtual void setName(std::string name)= 0;
+  virtual std::string getDescription() const = 0;
+  virtual void setDescription(std::string desc)= 0;
+  virtual bool isIndex(const TableIndex &index) const = 0;
+ 	virtual bool isIndex(const Person &person) const = 0;
+
+  	
+  /** 
+  Returns a TableIndex object with key names from the table, but no values filled in. 
+  Once the values are filled in, the TableIndex object can be used to index elements in the table.
+  */
+  virtual TableIndex getIndexTemplate() const = 0;
+
[…]

### 2019-11-22 FEM_CPP/TimeScaledProbitRegression.h

@@ -1,0 +1,27 @@
+#pragma once
+#include "ProbitRegression.h"
+#include "utility.h"
+
+class TimeScaledProbitRegression :
+	public ProbitRegression
+{
+public:
+
+	TimeScaledProbitRegression(void);
+	TimeScaledProbitRegression(const TimeScaledProbitRegression& source);
+	virtual ~TimeScaledProbitRegression(void) 	{	}
+	virtual inline IModel* clone() const { return new TimeScaledProbitRegression(*this);}
+	virtual inline std::string getType() const { return "time_scaled_probit";}
+	virtual inline std::string getTypeDesc() const { return "time rescaled probit regression";}
+
+	/** Reads the model. Just like a normal probit specification, except the line after the dependent variable contains the modeled number of years in the transition */
+	virtual void read(std::istream& inf, IVariableProvider* provider);
+
+	/** Calculates the probability of the event by calculating P = 1 - (1 - CDF(xb))^(yr_step/modeled years), then applies the random draw */
+	virtual void predict(Person* person, const Random* random) const;
+protected:
+	
+	/** The modeled number of years for the transition model. */
+	int modeled_delta_year;
+	IVariable* yr_step;
+};

### 2019-11-22 FEM_CPP/TimeSeries.h

@@ -1,0 +1,31 @@
+#pragma once
+#include <string>
+#include <vector>
+class ITimeSeries
+{
+public:
+	virtual ~ITimeSeries(void) {}
+    virtual double Value(unsigned int year) const = 0;
+    virtual unsigned int StartYear() const = 0;
+    virtual unsigned int EndYear() const = 0;
+    virtual ITimeSeries* CrossSection(unsigned int year_start, unsigned int year_end) const = 0;
+    virtual std::string getName() const = 0;
+    virtual void setName(std::string name)= 0;
+    virtual std::string getDescription() const = 0;
+    virtual void setDescription(std::string desc)= 0;
+
+	static ITimeSeries* Read(std::string timeseries_name, const char* filename);
+};
+
+
+
+class ITimeSeriesProvider
+{
+public:
+	virtual ~ITimeSeriesProvider(void) {}
+	virtual ITimeSeries* get(std::string name) = 0;
+	virtual void getAll(std::vector<ITimeSeries*> &vec) = 0;
+	virtual bool hasSeries(std::string name) = 0;
+};
+
+

### 2019-11-22 FEM_CPP/TimeSeriesManager.cpp

@@ -1,0 +1,93 @@
+#include "TimeSeriesManager.h"
+#include "utility.h"
+#include <fstream>
+#include "fem_exception.h"
+TimeSeriesManager::TimeSeriesManager(void)
+{
+	builder = NULL;
+}
+
+TimeSeriesManager::~TimeSeriesManager(void)
+{
+	for(std::map<std::string, ITimeSeries*>::iterator itr = series_map.begin(); itr != series_map.end(); ++itr)
+		delete (*itr).second;
+
+	for(std::map<std::string, ProxyTimeSeries*>::iterator itr = proxy_series_map.begin(); itr != proxy_series_map.end(); ++itr)
+		delete (*itr).second;
+}
+
+
+ITimeSeries* TimeSeriesManager::get(std::string name) {
+  if(proxy_series_map.count(name) > 0)
+    return proxy_series_map[name];
+  else
+    throw missing_var_exception("TimeSeriesManager does not have entry named " + name, name);
+}
+
+ITimeSeries* TimeSeriesManager::addTimeSeries(ITimeSeries* series){
+	// If a time series with this name already exists, then delete the existing version
+	if(series_map.count(series->getName()))
+		delete series_map[series->getName()];
+	series_map[series->getName()] = series;
+	
+	// If there is not yet a proxy time series with this name, then create it
+	// and add a lookup node for this time series to the node builder
+	if(proxy_series_map.count(series->getName()) == 0) {
+		ProxyTimeSeries* p = new ProxyTimeSeries(series->getName(), series->getDescription());
+		proxy_series_map[series->getName()] = p;
+		if(builder != NULL)
+			builder->addLookup(new LookupNode(new TimeSeriesLookup(p)));
+	}
+
+	// Have the proxy for this time series point to this new time series
+	proxy_series_map[series->getName()]->setSeries(series);
+	
+	return proxy_series_map[series->getName()];
+}
+
+void TimeSeriesManager::setBuilder(NodeBuilder* b) {
+	builder = b;
+	if(builder != NULL) {
+		std::map<std::string, ProxyTimeSeries*>::iterator itr;
+		for(itr = proxy_series_map.begin(); itr != proxy_series_map.end(); ++itr)
+			if(!builder->containsLookup((*itr).first))
[…]

### 2019-11-22 FEM_CPP/VarCalcModule.cpp

@@ -1,0 +1,26 @@
+#include "VarCalcModule.h"
+#include "Logger.h"
+
+VarCalcModule::VarCalcModule(void)
+{
+}
+
+VarCalcModule::~VarCalcModule(void)
+{
+}
+
+void VarCalcModule::process(PersonVector& persons, unsigned int year, Random* random) {
+	
+	Logger::log("Running Variable Calculations Module", FINE);
+	std::vector<Person*>::iterator itr;
+	for(itr = persons.begin(); itr != persons.end(); ++itr) {
+		Person* person = *itr;
+		if(person->test(Vars::active) && !person->test(Vars::l2died)) {
+			/* Increment the year by the year step */
+			person->set(Vars::year, person->get(Vars::year)+scenario->YrStep());
+
+			/* Calculate the new age */
+			person->set(Vars::age, person->get(Vars::year) - person->get(Vars::rbyr) + (7.0 - person->get(Vars::rbmonth))/12);
+		}
+	}
+}

### 2019-11-22 FEM_CPP/VarMap.cpp

@@ -1,0 +1,284 @@
+#include <cstring>
+#include "VarMap.h"
+
+/*
+// elementwise comparison
+bool VarMapKey::operator<(const VarMapKey &x) {
+	for(unsigned int i=0; i < size(); i++) {
+		if(this->at(i) < x[i])
+			return true;
+		else if(this->at(i) > x[i])
+			return false;
+	}
+	// if not returned yet, they are equal
+	return false;
+}
+*/
+
+VarMapKey::VarMapKey(size_t length) : std::vector<double>() { 
+	resize(length); 
+}
+
+template <class val_type> void VarMap<val_type>::read(const char* mapdef_file) {
+	std::ifstream inf(mapdef_file);
+	if( inf.bad() || inf.fail())
+		throw fem_exception("Could not open file: " + std::string(mapdef_file));
+	char buf[5000];
+
+	clear();
+	
+	// First line has comments -- read and ignore;
+	inf.getline(buf, 5000);
+
+	read(inf);
+
+	inf.close();
+}
+
+
+template <class val_type> void VarMap<val_type>::read(std::istream& inf) {
+	char buf[5000];
+	const char* delims = " \t"; // delimiters of variable names
+
+	int line = 1;
+	// Read definition variables from next line
+	size_t nvars = 0;
+	inf.getline(buf, 5000);
+	
+	if(!inf.eof()) {
+		// parse out variable names
+		char* namep = std::strtok(buf, delims);
+		while(namep != NULL) {
+			std::string vname = std::string(namep);
+			// it would be good to check if vname is a valid variable name, but the variable provider might not 
+			// be set until after this read. so, we'll allow invalid variable names here and throw an exception 
+			// later when a table lookup is attempted.
+			varNames.push_back(vname);
+			nvars++;
+			namep = std::strtok(NULL, delims);
+		}
+	} else {
+		throw fem_exception("Could not read first line of table");
+	}
+	
+	indexLength = nvars;
+	/** \todo Should it be an exception if indexLength==0? */
+	/* not sure because it's possible that someone might have a 
+	   special table that gives the same value for all persons
+	*/
+	
+	// read the table data
+	while(!inf.eof()) {
+		line++;
+		// Read a line
+		inf.getline(buf, 5000);	
+		
+		// Check that the line is valid and not a comment
+		if(strlen(buf) > 0 && buf[0] != '|')  {
[…]

### 2019-11-22 FEM_CPP/VarProbIntervention.cpp

@@ -1,0 +1,112 @@
+#include "VarProbIntervention.h"
+#include "Logger.h"
+#include "Random.h"
+#include "utility.h"
+#include "ConcreteTimeSeries.h"
+#include "GlobalVariable.h"
+#include "ProbitRegression.h"
+
+#include <sstream>
+VarProbIntervention::VarProbIntervention(std::string name, Vars::Vars_t v, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp)  
+  :  Intervention(intervention_id, tp, vp), _name(name), var(v)
+{
+	param_name = "mult_p";
+	param_name += VarsInfo::labelOf(var);
+
+	elig_var_param_name = "elig_mult_p";
+	elig_var_param_name += VarsInfo::labelOf(var);
+
+	this->const_mult = vp->addVariable(new GlobalVariable(param_name + "_const_mult", 0.5));
+	this->use_ts = false;
+	elig_var = vp->get("true");
+	std::ostringstream ss;
+	ss << const_mult->name();
+	params_map[param_name] = ss.str();
+	params_map[elig_var_param_name] = "true";
+
+}
+
+VarProbIntervention::VarProbIntervention(Vars::Vars_t v, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp)
+  : Intervention(intervention_id, tp, vp), var(v)
+{
+	_name = "P";
+	std::string tmp = VarsInfo::labelOf(var).substr(0,1);
+	_name += StringToUpper(tmp);
+	_name += VarsInfo::labelOf(var).substr(1);
+	_name += "Mult";
+
+	param_name = "mult_p";
+	param_name += VarsInfo::labelOf(var);
+
+	elig_var_param_name = "elig_mult_p";
+	elig_var_param_name += VarsInfo::labelOf(var);
+
+	this->const_mult = vp->addVariable(new GlobalVariable(param_name + "_const_mult", 0.5));
+	this->use_ts = false;
+	elig_var = vp->get("true");
+	std::ostringstream ss;
+	ss << const_mult->name();
+	params_map[param_name] = ss.str();
+	params_map[elig_var_param_name] = "true";
+}
+
+VarProbIntervention::~VarProbIntervention(void)
+{
+}
+
+
+void VarProbIntervention::setScenario(Scenario* scen) {
+	Intervention::setScenario(scen);
+
[…]

### 2019-11-22 FEM_CPP/VarProbIntervention.h

@@ -1,0 +1,35 @@
+#pragma once
+#include "Intervention.h"
+#include "TimeSeries.h"
+#include "Variable.h"
+
+/** This intervention adjusts the probability of an outcome through a multiplier.
+The probability of intervening is determined by the value of the eligibility 
+variable specified in the scenario settings.  The multiplier used for intervention,
+specified in the scenario settings, can be a constant value or the name of a time 
+series or variable in the providers passed to the constructor.
+*/
+class VarProbIntervention:
+	public Intervention
+{
+public:
+  VarProbIntervention(std::string name, Vars::Vars_t var, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+  VarProbIntervention(Vars::Vars_t var, unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~VarProbIntervention(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random) ;
+	virtual void setScenario(Scenario* scen);
+
+	virtual inline std::string name() const { return _name;}	
+
+protected:
+
+
+	std::string _name;
+	std::string param_name;
+	Vars::Vars_t var;
+	IVariable* const_mult;
+	ITimeSeries* ts_mult;
+	bool use_ts;
+	std::string elig_var_param_name;
+	IVariable* elig_var;
+};

### 2019-11-22 FEM_CPP/VariableManager.cpp

@@ -1,0 +1,178 @@
+#include "VariableManager.h"
+#include "utility.h"
+#include "Logger.h"
+#include "EquationParser.h"
+#include "EquationNode.h"
+#include "DerivedVariable.h"
+#include "BuiltInVariable.h"
+#include "fem_exception.h"
+#include "file_not_found_exception.h"
+#include <fstream>
+#include <cstring>
+
+VariableManager::VariableManager(void)
+{
+	builder = NULL;
+	for(unsigned int i = 0; i < Vars::NVars; i++)
+		addVariable(new BuiltInVariable((Vars::Vars_t)i));
+}
+
+VariableManager::~VariableManager(void)
+{
+	for(std::map<std::string, IVariable*>::iterator itr = var_map.begin(); itr != var_map.end(); ++itr)
+		delete (*itr).second;
+
+
+	for(std::map<std::string, ProxyVariable*>::iterator itr = proxy_var_map.begin(); itr != proxy_var_map.end(); ++itr)
+		delete (*itr).second;
+}
+
+
+IVariable* VariableManager::get(std::string name) {
+  if(exists(name))
+    return proxy_var_map[name];
+  throw fem_exception(std::string(name + " is not a valid variable").c_str());
+}
+
+bool inline VariableManager::exists(std::string name) const {
+  return proxy_var_map.count(name) != 0;
+}
+
+IVariable* VariableManager::addVariable(IVariable* var){
+
+	// If a variable with this name already exists, then delete the existing version
+	if(var_map.count(var->name()))
+		delete var_map[var->name()];
+	var_map[var->name()] = var;
+	
+	// If there is not yet a proxy variable with this name, then create it
+	// and add a variale node for this variable to the node builder
+	if(proxy_var_map.count(var->name()) == 0) {
+		ProxyVariable* p = new ProxyVariable(var->name(), var->description());
+		proxy_var_map[var->name()] = p;
+		if(builder != NULL)
+			builder->addTerminalNode(new VariableNode(p));
+	}
+
+	// Have the proxy for this variable point to the actual variable
+	proxy_var_map[var->name()]->setVar(var);
+	return proxy_var_map[var->name()];
+}
+
+void VariableManager::setBuilder(NodeBuilder* b) {
+	builder = b;
+	if(builder != NULL) {
+		std::map<std::string, ProxyVariable*>::iterator itr;
[…]

### 2019-11-22 FEM_CPP/WeightLossPill.h

@@ -1,0 +1,18 @@
+#pragma once
+#include "Intervention.h"
+#include <set>
+class WeightLossPill :
+	public Intervention
+{
+public:
+  WeightLossPill(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~WeightLossPill(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random);
+	virtual inline std::string name() const { return std::string("WeightLossPill");};
+	virtual void setScenario(Scenario* scen);
+protected:
+	bool elig(Person* p) const;
+
+	unsigned int start_yr;
+	double reduction;
+};

### 2019-11-22 FEM_CPP/WorkTillMedicareIntervention.h

@@ -1,0 +1,19 @@
+#pragma once
+#include "Intervention.h"
+#include "Variable.h"
+
+class WorkTillMedicareIntervention :
+	public Intervention
+{
+public:
+  WorkTillMedicareIntervention(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp);
+	virtual ~WorkTillMedicareIntervention(void);
+	virtual void intervene(PersonVector& persons, unsigned int year, Random* random);
+	virtual void setScenario(Scenario* scen);
+
+	virtual inline std::string name() const {return "WorkTillMedicare";}
+	
+
+protected:
+	IVariable* medicare_elig_var;
+};

### 2019-11-22 FEM_CPP/file_not_found_exception.h

@@ -1,0 +1,45 @@
+#pragma once
+#include <exception>
+
+class file_not_found_exception : public fem_exception
+{
+protected:
+  // Description: Filename tried to open
+  std::string filename;
+public:
+  // Description: Initializes the message.
+  //
+  // Arguments:
+  //   message: Message for this exception.
+	//   filename: the file that cannot be opened
+	file_not_found_exception( const std::string& message, const std::string& fname ) : fem_exception(message),
+    filename(fname) { }
+   // Description: Initializes the message.
+  //
+  // Arguments:
+  //   message: Message for this exception.
+  //   filename: the file that cannot be opened
+  file_not_found_exception( const char* message, const char* fname ) : fem_exception(message),
+    filename(fname) { }
+
+
+   // Description: Initializes the message.
+  //
+  // Arguments:
+  //   filename: the file that cannot be opened
+	file_not_found_exception( const std::string& fname ) : fem_exception("Could not open file" + fname),
+    filename(fname) { }
+   // Description: Initializes the message.
+  //
+  // Arguments:
+  //   filename: the file that cannot be opened
+	  file_not_found_exception( const char* fname ) : fem_exception("Could not open file" + std::string(fname)),
+    filename(fname) { }
+
+  // Description: Destructor.
+  ~file_not_found_exception()  throw()  { }
+  // Description: Returns description of this exception.
+  //
+  // Returns: Description of this exception.
+  virtual const char *what() const throw() { std::string s = theMessage + ", " + filename; return s.c_str(); }
+};

### 2019-11-22 FEM_CPP/missing_var_exception.cpp

@@ -1,0 +1,16 @@
+#include "missing_var_exception.h"
+
+
+
+std::string missing_var_exception::default_msg(const std::string& vname)  {
+	std::string msg("Attempted to read missing variable [");
+	msg += vname;
+	msg += "]";
+	return msg;
+}
+
+
+std::string missing_var_exception::default_msg(const char* vname)  {
+	std::string varname_str(vname);
+	return default_msg(varname_str);
+}

### 2019-11-22 FEM_CPP/missing_var_exception.h

@@ -1,0 +1,41 @@
+#pragma once
+#include "fem_exception.h"
+
+/** Exception thrown when the value of a variable flagged as missing is requested
+*/
+class missing_var_exception :
+	public fem_exception
+{
+public:
+	/** Creates a missing_var_exception with the given message and variable name */
+	missing_var_exception(const std::string& message, const std::string& vname) : fem_exception(message), varname(vname) { }
+	
+	/** Creates a missing_var_exception with the default message for the given variable name 
+		The default message is "Attempted to read missing variable [varname]"
+	*/
+	missing_var_exception(const std::string& vname) : fem_exception(default_msg(vname)), varname(vname) { }
+
+	/** Creates a missing_var_exception with the given message and variable name */
+	missing_var_exception(const char* message, const char* vname) : fem_exception(message), varname(vname) { }
+
+	/** Creates a missing_var_exception with the default message for the given variable name 
+		The default message is "Attempted to read missing variable [varname]"
+	*/
+	missing_var_exception(const char* vname) : fem_exception(default_msg(vname)), varname(vname) { }
+
+	/** Destructor */
+	virtual ~missing_var_exception(void) throw()  { }
+
+	/** Returns the name of the missing variable */
+	virtual std::string getVarname(void) {return varname;}
+
+protected:
+  /** The name of the variable with a missing value */
+  std::string varname;
+
+  /** Creates the default message from the given variable name */
+  std::string default_msg(const std::string& varname);
+
+  /** Creates the default message from the given variable name */
+  std::string default_msg(const char* vname);
+};

### 2019-11-22 FEM_CPP/testing/mvnorm_main.cpp

@@ -1,0 +1,173 @@
+#include <cmath>
+#include <iostream>
+#include <iomanip>
+#include <vector>
+#include "../RandomBlowfish.h"
+#include "../RandomBasic.h"
+#include "../RandomTable.h"
+
+/** length of random vectors used for testing */
+#define TEST_SIZE 5
+/** number of Monte Carlo draws used for testing */
+#define NDRAWS 5000000
+
+/** Testing code for multivarite normal random generation. 
+Creates NDRAWS Monte Carlo draws, estimate sample moments and compare to true values.
+\todo This only tests the implementation through RandomBlowfish.  RandomBasic and RandomTable implementations should also be tested.
+*/
+int main(int argc, char** argv) {
+    RandomBlowfish blowfish_test(1091638739, 1);
+ 
+ 		// set up mean and covariance parameters
+ 		std::vector<double> mean_true(TEST_SIZE);
+ 		std::vector< std::vector<double> > C(TEST_SIZE);
+ 		for(unsigned int i=0; i < TEST_SIZE; i++) {
+ 			mean_true[i] = 100*(i+1); 
+ 			C[i].resize(i+1);
+ 			for(unsigned int j=0; j <= i; j++) {
+ 				C[i][j] = (i+1) + (j+1);
+ 				if(i != j) {
+ 					int s = 0;
+ 					int t = ((j+1) % (i+1)) - ((TEST_SIZE / 2)+1);
+ 					if(t < 0)
+ 						s = -1;
+ 					else if(t > 0)
+ 						s = 1;
+ 					C[i][j] *= s;
+ 				}
+ 			}
+   	}
+   	double cov_true[TEST_SIZE][TEST_SIZE];
+   	for(unsigned int i=0; i < TEST_SIZE; i++)
+   		for(unsigned int j=0; j <= i; j++) {
+   			cov_true[i][j] = cov_true[j][i] = 0.0;
+   			for(unsigned int k=0; k <= std::min(i,j); k++) {
+   				cov_true[i][j] += C[i][k] * C[j][k];
+   			}
+   			cov_true[j][i] = cov_true[i][j];
+   		}
+   			
+   	// generate random draws
+   	unsigned int id = 10000;
+   	std::vector<unsigned int> process(TEST_SIZE);
+   	for(unsigned int i=0; i < TEST_SIZE; i++)
[…]

### 2019-11-22 analysis/disease_survival/ggkm.R

@@ -1,0 +1,72 @@
+# This code has some modifications for better formatting.  The original code can be found
+# at https://gist.github.com/araastat/9927677
+
+#’ Create a Kaplan-Meier plot using ggplot2
+#’
+#’ @param sfit a \code{\link[survival]{survfit}} object
+#’ @param returns logical: if \code{TRUE}, return an ggplot object
+#’ @param xlabs x-axis label
+#’ @param ylabs y-axis label
+#’ @param ystratalabs The strata labels. \code{Default = levels(summary(sfit)$strata)}
+#’ @param ystrataname The legend name. Default = “Strata”
+#’ @param timeby numeric: control the granularity along the time-axis
+#’ @param main plot title
+#’ @param pval logical: add the pvalue to the plot?
+#’ @return a ggplot is made. if returns=TRUE, then an ggplot object
+#’ is returned
+#’ @author Abhijit Dasgupta with contributions by Gil Tomas
+#’ \url{http://statbandit.wordpress.com/2011/03/08/an-enhanced-kaplan-meier-plot/}
+#’ @export
+#’ @examples
+#’ \dontrun{
+#’ data(colon)
+#’  fit <- survfit(Surv(time,status)~rx, data=colon)
+#'  ggkm(fit, timeby=500)
+#' }
+ggkm <- function(sfit, returns = FALSE,
+xlabs = "Time", ylabs = "survival probability",
+ystratalabs = NULL, ystrataname = NULL,
+timeby = 100, main = "Kaplan-Meier Plot",
+pval = TRUE, ...) {
+require(plyr)
+require(ggplot2)
+require(survival)
+require(gridExtra)
+if(is.null(ystratalabs)) {
+   ystratalabs <- as.character(levels(summary(sfit)$strata))
+}
+m <- max(nchar(ystratalabs))
+if(is.null(ystrataname)) ystrataname <- "Strata"
+times <- seq(0, max(sfit$time), by = timeby)
+.df <- data.frame(time = sfit$time, n.risk = sfit$n.risk,
+    n.event = sfit$n.event, surv = sfit$surv, strata = summary(sfit, censored = T)$strata,
+    upper = sfit$upper, lower = sfit$lower)
+levels(.df$strata) <- ystratalabs
+zeros <- data.frame(time = 0, surv = 1, strata = factor(ystratalabs, levels=levels(.df$strata)),
[…]

### 2019-11-22 analysis/psid_marriage/marriage_plots.R

@@ -1,0 +1,516 @@
+# This file creates plots and corresponding data in .csv files for PSID marriage outcomes
+
+library(ggplot2)
+library(gridExtra)
+library(gtable)
+
+# set the starting year for plots here:
+year.start = 2009
+
+# image parameters (units=inches)
+img.height = 8
+img.width = 8
+img.res = 200
+
+sname = c("psid")
+sdesc = c("PSID Default")
+
+print("Loading sources:")
+for(s in sname) { 
+	# load each scenario's summary output
+	fname = paste("../../output/",s,"/",s,"_summary.Rdata",sep="")
+	print(fname)
+	load(file=fname) 
+	ls()
+	
+	# drop records before year.start
+	eval(parse(text=paste(s,"_summary=",s,"_summary[",s,"_summary$year > 2009,]",sep="")))
+}
+ls()
+
+# single line plot
+plotline1 <- function(varsrc, varname, xlabel, ylabel, ymult, ymin, ymax, title) {
+	df.src = eval(parse(text=varsrc))
+	pl = qplot(df.src[,"year"], ymult*df.src[,varname], geom="Line", xlab=xlabel, ylab=ylabel, ylim=c(ymin,ymax), main=title)
+	ggsave(filename=paste(varsrc,"_",varname,".pdf",sep=""), plot=pl, width=img.width, height=img.height, units="in")
+	write.csv(file=paste(varsrc,"_",varname,".csv",sep=""), df.src[c("year",varname)], row.names=FALSE)
+	return(pl)
+}
+
+# line plot of different series from same source
+plotline2 <- function(varsrc, xvar, yvars, ydescs, xlabel, ylabel, ymult, ymin, ymax, title) {
+	# create individual data frames and append them together
+	fprefix = paste(varsrc,paste(yvars, collapse="_"),sep="_")
+	df.master = NULL
+	for(i in 1:length(yvars)) {
+		df.temp = eval(parse(text=varsrc))
+		df.temp["series"] = rep(ydescs[i], nrow(df.temp))
+		df.temp = df.temp[c(xvar, yvars[i], "series")]
+		names(df.temp)[2] = "y"
+		df.master = rbind(df.master, df.temp)
+	}
+	
+	# create plot
[…]

### 2019-11-22 analysis/techdoc/latex/FEM/get_estimates_table_url.py

@@ -1,0 +1,40 @@
+##### Open the link file.
+with open("estimates_table_link.tex") as ifile:
+	cururl = ifile.read()
+
+##### Extract the URL.
+cururl = cururl[5:-1]
+
+##### Interact w/ user to update URL
+disptxt = '''\n
+The technical documentation is currently using this URL
+to link to the estimates table:\n\n''' + cururl + '''\n
+If you want to use a different URL, please type it now
+and then press Enter/Return. Otherwise, don't type
+anything and just press Enter/Return to continue using 
+the current URL.\n
+New URL: '''
+
+newurl = raw_input(disptxt)
+
+if newurl == "":
+	print "\nYou did not type a new URL."
+	newurl = cururl
+else:
+	print "\nThe URL has been updated."
+
+print '''
+The technical documentation will use this URL to link 
+to the table of estimates:\n\n''' + newurl
+
+#### Overwrite the link file with the URL (new or old).
+
+newurl = "\\url{" + newurl + "}"
+with open("estimates_table_link.tex", 'w') as ofile:
+	ofile.write(newurl)
+
+print '''
+Finished updating link to estimates table. Now you
+need to remake (or finish making) the technical
+documentation to include the updated URL.
+'''

### 2019-11-22 bootstrap_directories.py

@@ -1,0 +1,22 @@
+import os
+
+x_max = int(os.environ["MAXBREP"])
+
+x = 1
+while (x <= x_max):
+	if not os.path.isdir(str(os.environ["INPUT"]) + "/input_rep" + str(x)):
+		os.makedirs(str(os.environ["INPUT"]) + "/input_rep" + str(x))
+
+	if not os.path.isdir(str(os.environ["STER"]) + "/models_rep" + str(x)):
+		os.makedirs(str(os.environ["STER"]) + "/models_rep" + str(x))
+
+	if not os.path.isdir(str(os.environ["STER"]) + "/models_rep" + str(x) + "/crossvalidation"):
+		os.makedirs(str(os.environ["STER"]) + "/models_rep" + str(x) + "/crossvalidation")
+
+	if not os.path.isdir(str(os.environ["EST"]) + "/models_rep" + str(x)):
+		os.makedirs(str(os.environ["EST"]) + "/models_rep" + str(x))
+
+	x += 1
+
+with open("bootstrap_directories.txt","w") as file:
+	file.write("Bootstrap directories created")

### 2019-11-22 new51_dependency_bstrend.py

@@ -1,0 +1,16 @@
+import os
+
+x_max = int(os.environ["MAXBREP"])
+
+x = 1
+with open("new51_dependency_bstrend.txt","w") as file:
+	while (x <= x_max):
+		y = 2004
+		file.write("\n")
+		file.write("BCOHORTSB" + str(x) + " := $(addsuffix _bstrend.dta, $(DATADIR)/input_rep" + str(x) + "/new51s)\n")
+		file.write("FEM_Stata/Makedata/HRS/new51s_bstrend_brep_" + str(x) + ".txt: ${BCOHORTSB" + str(x) + "}\n")
+		file.write("\n")
+		file.write("$(DATADIR)/input_rep" + str(x) + "/new51s_bstrend.dta: $(COMMON) FEM_Stata/Makedata/HRS/new51_simulate_bootstrap.do $(DATADIR)/new51s_status_quo.dta $(DATADIR)/input_rep" + str(x) + "/bootstrap_sample.dta $(DATADIR)/pop5152_projection_2150.dta \n")
+		file.write("	cd FEM_Stata/Makedata/HRS && SCENARIO=bstrend TREND=status_quo BREP=" + str(x) + " FYEAR=$(FYEAR) LYEAR=$(LYEAR) $(STATA) new51_simulate_bootstrap.do\n")
+		
+		x += 1

### 2019-11-22 new51_dependency_input_bstrend.py

@@ -1,0 +1,16 @@
+import os
+
+x_max = int(os.environ["MAXBREP"])
+
+x = 1
+with open("new51_dependency_input_bstrend.txt","w") as file:
+	while (x <= x_max):
+		y = 2004
+		file.write("\n")
+		file.write("BCOHORTSIB" + str(x) + " := $(addsuffix _input_bstrend.dta, $(DATADIR)/input_rep" + str(x) + "/new51s)\n")
+		file.write("FEM_Stata/Makedata/HRS/new51s_input_bstrend_brep_" + str(x) + ".txt: ${BCOHORTSIB" + str(x) + "}\n")
+		file.write("\n")
+		file.write("$(DATADIR)/input_rep" + str(x) + "/new51s_input_bstrend.dta: $(COMMON) FEM_Stata/Makedata/HRS/new51_simulate_bootstrap.do $(DATADIR)/input_rep" + str(x) +"/new51s_input.dta $(DATADIR)/input_rep" + str(x) + "/bootstrap_sample.dta $(BASEDIR)/pop5152_projection_2080.dta \n")
+		file.write("	cd FEM_Stata/Makedata/HRS && SCENARIO=input_bstrend TREND=status_quo BREP=" + str(x) + " FYEAR=$(FYEAR) LYEAR=$(LYEAR) $(STATA) new51_simulate_bootstrap.do\n")
+		
+		x += 1

### 2019-11-22 nomake.setup.sh

@@ -1,0 +1,23 @@
+#
+# ROOT=/sch-projects/public-data-projects/FEM/pstclair/psid2
+# export ROOT
+CURDIR=$(pwd)
+ROOT=$CURDIR
+export ROOT
+
+export MCBSDIR=/sch-projects/dua-data-projects/FEM/MCBS/$LOGNAME/rdata_psid
+export HRSDIR=/sch-projects/dua-data-projects/FEM/HRS/$LOGNAME/rdata_psid 
+export HRSRESTRICT=/sch-data-library/dua-data/HRS/Restricted/FEM 
+export HRSSENSITIVE=/sch-data-library/dua-data/HRS/Sensitive 
+export MCBSRESTRICT=/sch-data-library/dua-data/MCBS/rawdata 
+export HRSPUB=/sch-data-library/public-data/HRS/Unrestricted 
+export RANDVER=k 
+export HRSPROJECTS=/sch-projects/public-data-projects/HRS 
+export MEPS=/sch-data-library/public-data/MEPS 
+export NHIS=/sch-data-library/public-data/NHIS 
+export NHANES=/sch-data-library/public-data/NHANES 
+export HMD=/sch-data-library/public-data/HumanMortalityDatabase 
+export PSIDSAS=$CURDIR/FEM_Sas/PSID/Programs/ 
+export PSIDSTATA=$CURDIR/FEM_Stata/Makedata/PSID 
+export PSIDPUB=/nfs/sch-data1/data-library/public-data/PSID/
+

### 2019-11-22 run.stata11.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+set -e
+nice stata11-mp -b $1
+set +e
+grep -C5 -E "^r\([0-9]+\)" ${1%do}log
+let $?
+exit
+

### 2019-11-22 run.stata14.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+nice stata14-mp -b $1
+grep -C5 -E "^r\([0-9]+\)" ${1%do}log
+let $?
+exit

### 2020-07-21 FEM_R/PH_Sensitivity_vis.Rmd
Added a number of WIP visualisation scripts written in R. Also a couple of R notebooks showing sensitivity analysis and PH interventions

---
title: "Public Health Intervention Visualisation"
output: html_notebook
---

This notebook will keep all visualisations for the English Future Elderly Model public health intervention scenarios.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(ggplot2)
```

# Smoke Stop Cohort

```{r}
# Read in all the data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
t10 <- read_dta('output/TestC10/TestC10_summary.dta')
t20 <- read_dta('output/TestC20/TestC20_summary.dta')
t50 <- read_dta('output/TestC50/TestC50_summary.dta')
t80 <- read_dta('output/TestC80/TestC80_summary.dta')
t100 <- read_dta('output/TestC100/TestC100_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
t10$age <- (cohort$year - 2012) + 50
t20$age <- (cohort$year - 2012) + 50
t50$age <- (cohort$year - 2012) + 50
t80$age <- (cohort$year - 2012) + 50
t100$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
t10$survival <- t10$m_endpop_all / t10$m_endpop_all[1]
t20$survival <- t20$m_endpop_all / t20$m_endpop_all[1]
t50$survival <- t50$m_endpop_all / t50$m_endpop_all[1]
t80$survival <- t80$m_endpop_all / t80$m_endpop_all[1]
t100$survival <- t100$m_endpop_all / t100$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=t10, size=0.3, col='green') +
        geom_line(data=t50, size=0.3, col='blue') +
        geom_line(data=t100, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=t10, size=0.3, col='green') +
        geom_line(data=t50, size=0.3, col='blue') +
        geom_line(data=t100, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=t10, size=0.3, col='green') +
        geom_line(data=t50, size=0.3, col='blue') +
        geom_line(data=t100, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=t10, size=0.3, col='green') +
        geom_line(data=t50, size=0.3, col='blue') +
        geom_line(data=t100, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=t10, size=0.3, col='green') +
        geom_line(data=t50, size=0.3, col='blue') +
        geom_line(data=t100, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
```

# Smoke Stop Population

```{r}
# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
tp10 <- read_dta('output/TestP10/TestP10_summary.dta')
tp20 <- read_dta('output/TestP20/TestP20_summary.dta')
tp50 <- read_dta('output/TestP50/TestP50_summary.dta')
tp80 <- read_dta('output/TestP80/TestP80_summary.dta')
tp100 <- read_dta('output/TestP100/TestP100_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=tp10, size=0.3, col='green') +
        geom_line(data=tp50, size=0.3, col='blue') +
        geom_line(data=tp100, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=tp10, size=0.3, col='green') +
        geom_line(data=tp50, size=0.3, col='blue') +
        geom_line(data=tp100, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=tp10, size=0.3, col='green') +
        geom_line(data=tp50, size=0.3, col='blue') +
        geom_line(data=tp100, size=0.3, col='red') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=tp10, size=0.3, col='green') +
        geom_line(data=tp50, size=0.3, col='blue') +
        geom_line(data=tp100, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=tp10, size=0.3, col='green') +
        geom_line(data=tp50, size=0.3, col='blue') +
        geom_line(data=tp100, size=0.3, col='red') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')
```

# Drinking Cohort

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
dc10 <- read_dta('output/DC10/DC10_summary.dta')
dc20 <- read_dta('output/DC20/DC20_summary.dta')
dc50 <- read_dta('output/DC50/DC50_summary.dta')
dc80 <- read_dta('output/DC80/DC80_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
dc10$age <- (cohort$year - 2012) + 50
dc20$age <- (cohort$year - 2012) + 50
dc50$age <- (cohort$year - 2012) + 50
dc80$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
dc10$survival <- dc10$m_endpop_all / dc10$m_endpop_all[1]
dc20$survival <- dc20$m_endpop_all / dc20$m_endpop_all[1]
dc50$survival <- dc50$m_endpop_all / dc50$m_endpop_all[1]
dc80$survival <- dc80$m_endpop_all / dc80$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')

ggplot(cohort, aes(x=age, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Drinking by Age', y='Prevalence of Drinking', x='Age')

ggplot(cohort, aes(x=age, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes by Age', y='Prevalence of Diabetes', x='Age')

ggplot(cohort, aes(x=age, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension by Age', y='Prevalence of Hypertension', x='Age')

ggplot(cohort, aes(x=age, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease by Age', y='Prevalence of Heart Disease', x='Age')

ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

ggplot(cohort, aes(x=age, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Stroke by Age', y='Prevalence of Stroke', x='Age')

```

# Drinking Population

```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
dp10 <- read_dta('output/DP10/DP10_summary.dta')
dp20 <- read_dta('output/DP20/DP20_summary.dta')
dp50 <- read_dta('output/DP50/DP50_summary.dta')
dp80 <- read_dta('output/DP80/DP80_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

ggplot(pop, aes(x=year, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Drinking by Age', y='Prevalence of Drinking', x='Age')

ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes by Age', y='Prevalence of Diabetes', x='Age')

ggplot(pop, aes(x=year, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension by Age', y='Prevalence of Hypertension', x='Age')

ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease by Age', y='Prevalence of Heart Disease', x='Age')

ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

ggplot(pop, aes(x=year, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dp10, size=0.3, col='green') +
        geom_line(data=dp50, size=0.3, col='blue') +
        geom_line(data=dp80, size=0.3, col='red') +
        labs(title='Prevalence of Stroke by Age', y='Prevalence of Stroke', x='Age')

```

# Exercise Status Cohort

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
ec10 <- read_dta('output/ExstatC10/ExstatC10_summary.dta')
ec50 <- read_dta('output/ExstatC50/ExstatC50_summary.dta')
ec80 <- read_dta('output/ExstatC80/ExstatC80_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
ec10$age <- (cohort$year - 2012) + 50
ec50$age <- (cohort$year - 2012) + 50
ec80$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
ec10$survival <- ec10$m_endpop_all / ec10$m_endpop_all[1]
ec50$survival <- ec50$m_endpop_all / ec50$m_endpop_all[1]
ec80$survival <- ec80$m_endpop_all / ec80$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=ec10, size=0.3, col='green') +
        geom_line(data=ec50, size=0.3, col='blue') +
        geom_line(data=ec80, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')

ggplot(cohort, aes(x=age, y=a_bmi_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Average BMI by Age', y='Average BMI', x='Age')

ggplot(cohort, aes(x=age, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes by Age', y='Prevalence of Diabetes', x='Age')

ggplot(cohort, aes(x=age, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension by Age', y='Prevalence of Hypertension', x='Age')

ggplot(cohort, aes(x=age, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease by Age', y='Prevalence of Heart Disease', x='Age')

ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

ggplot(cohort, aes(x=age, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Stroke by Age', y='Prevalence of Stroke', x='Age')

```

# Exercise Status Population

```{r}

pop <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
ep10 <- read_dta('output/ExstatP10/ExstatC10_summary.dta')
ep50 <- read_dta('output/ExstatP50/ExstatC50_summary.dta')
ep80 <- read_dta('output/ExstatP80/ExstatC80_summary.dta')

# Now plot all survival curves together
ggplot(pop, aes(x=year, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=ep10, size=0.3, col='green') +
        geom_line(data=ep50, size=0.3, col='blue') +
        geom_line(data=ep80, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')

ggplot(pop, aes(x=year, y=a_bmi_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Average BMI Over Time', y='Average BMI', x='Age')

ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes Over Time', y='Prevalence of Diabetes', x='Age')

ggplot(pop, aes(x=year, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension Over Time', y='Prevalence of Hypertension', x='Age')

ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease Over Time', y='Prevalence of Heart Disease', x='Age')

ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Cancer Over Time', y='Prevalence of Cancer', x='Age')

ggplot(pop, aes(x=year, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=dc10, size=0.3, col='green') +
        geom_line(data=dc50, size=0.3, col='blue') +
        geom_line(data=dc80, size=0.3, col='red') +
        labs(title='Prevalence of Stroke Over Time', y='Prevalence of Stroke', x='Age')

```

### 2020-07-21 FEM_R/pubHealthInterventionVis_Notebook.Rmd
Added a number of WIP visualisation scripts written in R. Also a couple of R notebooks showing sensitivity analysis and PH interventions

---
title: "Public Health Intervention Visualisation"
output: html_notebook
---

This notebook will keep all visualisations for the English Future Elderly Model public health intervention scenarios.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(ggplot2)
```

# COMMIT

The commit smoking interventions are inspired by this paper. <!-- TODO: Add link to COMMIT paper here -->

The intervention resulted in a 3% increase in the smoking quit rate for light/moderate smokers (anyone smoking less than 25 cigs/day).

```{r}
 # First read in the baseline cohort data file and the commit scenario outputs
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
commit3 <- read_dta('output/COMMIT_cSmoken3/COMMIT_cSmoken3_summary.dta')

# As these are all cohort datasets we can calculate their age by (year - 2012 + 50) 
cohort$age <- (cohort$year - 2012) + 50
commit3$age <- (cohort$year - 2012) + 50

# Now calculate survival rates for all 3 (normalise pop_count to inital value)
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
commit3$survival <- commit3$m_endpop_all / commit3$m_endpop_all[1]

# Plot survival curve
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.5) +
        geom_line(data=commit3, col='red', size=0.5)

# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
```

# Smoke Stop Cohort Interventions

```{r}
# First read in the baseline cohort data file and the commit scenario outputs
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
x169 <- read_dta('output/Smoke_Stop_Cohort169/Smoke_Stop_Cohort169_summary.dta')

# As these are all cohort datasets we can calculate their age by (year - 2012 + 50)
cohort$age <- (cohort$year - 2012) + 50
x169$age <- (cohort$year - 2012) + 50

# Now calculate survival rates for all 3 (normalise pop_count to inital value)
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
x169$survival <- x169$m_endpop_all / x169$m_endpop_all[1]

# Plot survival curve
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.5) +
        geom_line(data=x169, col='red', size=0.5)

# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
```

# Smoke Stop Population Interventions

```{r}
# First read in the baseline cohort data file and the commit scenario outputs
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
p169 <- read_dta('output/Smoke_Stop_Pop169/Smoke_Stop_Pop169_summary.dta')

# No need to calculate year as this is population scenario (replenishing)
# Also means can't calculate survival rates

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')
```

# Custom Smoke Stop Interventions (Cohort)

This is the custom intervention written in C++ - SmokeStopIntervention.cpp

```{r}

# Read in data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
SSPC <- read_dta('output/SmokeStopInterventionC/SmokeStopInterventionC_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
SSPC$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
SSPC$survival <- SSPC$m_endpop_all / SSPC$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# Custom Smoke Stop Interventions (Population)

```{r}

# Load data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
SSPP <- read_dta('output/SmokeStopInterventionP/SmokeStopInterventionP_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```

### 2020-08-12 FEM_R/ELSA_US_comparison.Rmd
Lots of changes to R scripts analysing outputs. Of note, started an R notebook for imputing missing data in ELSA before running through the model

---
title: "Comparing Understanding Society with ELSA"
output: html_notebook
---

This document exists to compare the distributions of key variables (mainly risk behaviours: smoking, drinking, exercise (and obesity)) between Understanding Society (US) and the English Longitudinal Survey of Ageing (ELSA).

# Setup

```{r "setup", include=FALSE}

# Set up the environment
US_dir <- "/home/luke/Documents/E_FEM_clean/Understanding_Society/UKDA-6614-stata/stata/stata11_se/ukhls_w2/"
ELSA_dir <- "/home/luke/Documents/E_FEM/UKDA-5050-stata/stata/stata11_se/"

require(haven)
require(dplyr)
require(ggplot2)

```

# Smoking Distributions by Age by Sex

```{r}

```

### 2020-08-12 FEM_R/ELSA_imputation.Rmd
Lots of changes to R scripts analysing outputs. Of note, started an R notebook for imputing missing data in ELSA before running through the model

---
title: "Multiple Imputation of Missing Data in ELSA"
output: html_notebook
---

This document will run through the steps taken to impute large amounts of missing data in ELSA.

Missing data in ELSA is a significant problem to running the FEM, as a number of key predictive variables have a high degree of missingness and/or are not missing at random. For example, BMI data is missing for every odd wave, and the even waves are missing a considerable proportion of that data.

First, the extent of missing information will be assessed, then a few methods for imputing these missing values will be trialled.

```{r "setup", include=FALSE}

```

### 2020-08-12 FEM_R/risk_behaviour_burden.Rmd
Lots of changes to R scripts analysing outputs. Of note, started an R notebook for imputing missing data in ELSA before running through the model

---
title: "Burden of Risk Behaviours"
output: html_notebook
---

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(ggplot2)
```

# Baseline Plots (Cohort)

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# No Smoking Scenario (Cohort)

```{r}

# Read in all the data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
no_smoke_C <- read_dta('output/CureSmoken_C/CureSmoken_C_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
no_smoke_C$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
no_smoke_C$survival <- no_smoke_C$m_endpop_all / no_smoke_C$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# SmokeStopFull (cohort)

```{r}

# Read in all the data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
SSF_C <- read_dta('output/SmokeStopFull_C/SmokeStopFull_C_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
SSF_C$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
SSF_C$survival <- SSF_C$m_endpop_all / SSF_C$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# No Smoking Scenario (Population)

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
no_smoke_P <- read_dta('output/CureSmoken_P/CureSmoken_P_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```

# SmokeStopFull (Population)

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
SSF_P <- read_dta('output/SmokeStopFull_P/SmokeStopFull_P_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```

### 2020-08-19 FEM_R/ELSA_imputation.Rmd
Multiple scripts in R based on imputing missing information in ELSA

@@
 
+require(haven)
+require(mice)
+require(lattice)
+require(tidyverse)
+require(Hmisc)
+
+FEM_dir <- "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/"
+knitr::opts_knit$set(root.dir = FEM_dir)
+
+set.seed(500) # Set seed for reproducibility
+
 ```
+
+First we set up the environment, changing the working directory and including the libraries we will be using for this task. We have also set the seed at '5000' to ensure reproducibility.
+
+Next step is to read in the harmonized ELSA data file.
+
+```{r}
+
+# Read in the data.
+H_ELSA_base <- read_dta('H_ELSA_f_2002-2016.dta')
+
+```
+
+Now let's describe the data using the describe function from the Hmisc package (takes some time and the output is LARGE):
+
+```{r}
+
+#describe(H_ELSA_base)
+
+```
+
+Now lets try to summarise some of the important variables, to see what extent they are missing values.
+
+```{r}
+
+bmi <- H_ELSA_base %>% select(matches('r[0-9]bmi$'))
+#summary(bmi)
+
+smokenVars <- H_ELSA_base %>% select(matches('r[0-9]smoken$'))
+#summary(smokenVars)
+
+smokevVars <- H_ELSA_base %>% select(matches('r[0-9]smokev$'))
+#summary(smokevVars)
+
+drinkVars <- H_ELSA_base %>% select(matches('r[0-9]drink$'))
+drinkdVars <- H_ELSA_base %>% select(matches('r[0-9]drinkd_e$'))
+```
+
+The output for this is too obtuse, and we are only really interested in the extent of missingness in a variable.
+Therefore, we are going to use a function specifically looking at how much data is missing.
+
+```{r}
+
+# Calculate percentage of missing cases
+pMiss <- function(x) {sum(is.na(x)) / length(x) * 100}
+
+apply(bmi, 2, pMiss) # Asses missing in BMI
+#apply(bmi, 1, pMiss)
+print(' ')
+apply(smokenVars, 2, pMiss)
+print(' ')
+apply(smokevVars, 2, pMiss)
+print(' ')
+apply(drinkVars, 2, pMiss)
+print(' ')
+apply(drinkdVars, 2, pMiss)
+
+cancreVars <- H_ELSA_base %>% select(matches('r[0-9]cancre$'))
+apply(cancreVars, 2, pMiss)
+```
+
+Let's just look at extracting important variables and looking for those with complete cases.
+
+```{r}
+
+# Select only the vars we want to impute, plus those to use in the imputation
+keyVars <- H_ELSA_base %>% select(contains(c('idauniq', 'iwindy', 'rabyear', 'ragender', 'raracem', 'raeducl', 'hlthlm', 'work', 'itearn', 'retemp',
+                                            'vgactx_e', 'mdactx_e', 'ltactx_e', 'ipubpen', 'drink', 'drinkd', 'bmi', 'smoken', 'smokev', 'smokef')))
+
+completeData <- complete.cases(keyVars)
+sum(completeData)
+
+someVars <- H_ELSA_base %>% select(contains(c('idauniq', 'rabyear', 'raeducl', 'drink', 'smoken', 'smokev')))
+completeSome <- complete.cases(someVars)
+sum(completeSome)
+
+```
+
+So that's **zero** complete cases when a relatively small group of important variables are chosen, and just over 1000 when we drop to 6 key variables. Not good.
+
+Now we've produced some numbers to explain just how bad the extent of missing data is in ELSA, we have to attempt to deal with it.
+First up is going to be a big and dirty multiple imputation model, but not how it is traditionally supposed to be used.
+Instead of using each new dataset to estimate a model and then pooling the results of that model to get an aggregate, we are going to pool the results of the datasets themselves and take an average, to impute the missing information in place.
+
+The next step is to add a few variables back in that we might want, namely iwstat and radyear.
+We need to make sure the mortality model is improved, as this was the whole reason for the imputation model in the first place.
+
+```{r}
+
+impVars2 <- H_ELSA_base %>% select(matches(c('idauniq', 'r[0-9]iwstat', 'rabyear$', 'radyear', 'ragender$', 'raeducl$', 'r[0-9]walkra$', 'r[0-9]dressa$', 'r[0-9]batha$', 
+                                            'r[0-9]eata$', 'r[0-9]beda$', 'r[0-9]toilta$', 'r[0-9]mapa$', 'r[0-9]phonea$', 'r[0-9]moneya$', 'r[0-9]medsa$', 'r[0-9]shopa$', 'r[0-9]mealsa$',
+                                            'r[0-9]housewka$', 'r[0-9]hibpe$', 'r[0-9]diabe$', 'r[0-9]cancre$', 'r[0-9]lunge$', 'r[0-9]hearte$', 'r[0-9]stroke$',
+                                            'r[0-9]bmi$', 'r[0-9]smokev$', 'r[0-9]smoken$', 'r[0-9]smokef$', 'r[0-9]drink$', 'r[0-9]vgactx_e$', 'r[0-9]mdactx_e$', 'r[0-9]ltactx_e$',
+                                            'r[0-9]drinkd_e$')))
+
+impVars2 <- impVars2 %>% select(-matches('s[0-9]idauniq'))
+
+colnames(impVars2)
+ncol(impVars2)
+
+# Assign impossible bmi values as missing (values with BMI < 10. Only happens in wave 8, 4 cases)
+impVars2$r8bmi[impVars2$r8bmi < 10] <- NA
+
+# Run the imputation
+# Tolerance had to be set to low as I kept getting the error:
+# Error in solve.default(xtx + diag(pen)) : system is computationally singular: reciprocal condition number = 4.00027e-17 (number changes when different vars are included)
+# Removed a number of variables also to try and get around this error, as it is saying that 1 or more of the variables are linearly dependant (or so small they are treated as zero - unlikely)
+impData2 <-mice(impVars2, m = 5, maxit=5, seed=500, tol=1e-25)
+
+saveRDS(impData2, file = "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/ELSA_imputed_R.mids")
+
+```
+
+```{r}
+
+impData2 <- readRDS(file = "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/ELSA_imputed_R.mids")
+
+# Now collect all imputed datasets in long format
+completedData <- complete(impData2, 'long')
+
+# Aggregate the data
+a<-aggregate(completedData , by = list(completedData$.id), FUN= mean)
+
+# Remove vars generated in imputation
+#colnames(a)
+final <- a %>% select(-c('Group.1', '.id', '.imp'))
+
+final_rename <- final
+colnames(final_rename) <- paste0(colnames(final_rename), '_imp')
+
+#colnames(final_rename)
+
+final_rename %>% rename(idauniq = idauniq_imp)
+
+#colnames(final_rename)
+
+write_path <- '/home/luke/Documents/E_FEM_clean/E_FEM/input_data/ELSA_imputed_R.dta'
+
+write_dta(final_rename, write_path)
+
+```
+
+Lots of processing still to be done on the imputed data file:
+- Round all binary vars to the nearest whole number. This will make sure everyone is either a 0 or 1
+- smokef var need to be rounded to the nearest whole number also, but we need to make sure that for anyone with smoken value of 0, smokef is also 0
+
+```{r}
+
+# Select all binary vars (and idauniq var)
+binVars <- final_rename %>% select(contains(c('walkra', 'dressa', 'batha', 'eata', 'beda', 'toilta', 'mapa', 'phonea', 'moneya', 'medsa', 'shopa', 'mealsa', 'housewka', 'hibpe', 'diabe', 'cancre',
+                                              )))
+
+# Now round all binary vars to nearest whole number
+
+```
+
+OLD STUFF / DEVELOPMENT
+
+```{r}
+
+impVars <- H_ELSA_base %>% select(matches(c('idauniq', 'rabyear$', 'ragender$', 'raeducl$', 'r[0-9]walkra$', 'r[0-9]dressa$', 'r[0-9]batha$', 
+                                            'r[0-9]eata$', 'r[0-9]beda$', 'r[0-9]toilta$', 'r[0-9]mapa$', 'r[0-9]phonea$', 'r[0-9]moneya$', 'r[0-9]medsa$', 'r[0-9]shopa$', 'r[0-9]mealsa$',
+                                            'r[0-9]housewka$', 'r[0-9]hibpe$', 'r[0-9]diabe$', 'r[0-9]cancre$', 'r[0-9]lunge$', 'r[0-9]hearte$', 'r[0-9]stroke$',
+                                            'r[0-9]bmi$', 'r[0-9]smokev$', 'r[0-9]smoken$', 'r[0-9]smokef$', 'r[0-9]drink$', 'r[0-9]vgactx_e$', 'r[0-9]mdactx_e$', 'r[0-9]ltactx_e$',
+                                            'r[0-9]drinkd_e$')))
+
+# 'r[0-9]retemp$', 'r[0-9]psyche$', 'r[0-9]arthre$',
+# 'r[0-9]ipubpen$', 'r[0-9]itearn$', 'h[0-9]atotf$',
+
+#colnames(impVars)
+ncol(impVars)
+
+# Assign impossible bmi values as missing (values with BMI < 10. Only happens in wave 8, 4 cases)
+impVars$r8bmi[impVars$r8bmi < 10] <- NA
+
+# Run the imputation
+# Tolerance had to be set to low as I kept getting the error:
+# Error in solve.default(xtx + diag(pen)) : system is computationally singular: reciprocal condition number = 4.00027e-17 (number changes when different vars are included)
+# Removed a number of variables also to try and get around this error, as it is saying that 1 or more of the variables are linearly dependant (or so small they are treated as zero - unlikely)
+impData <-mice(impVars, m = 5, maxit=5, seed=500, tol=1e-25)
+
+```
+Complete! Had to remove a couple of variables that we would rather have kept but at least it has worked. We can try to include these variables now iteratively and see where we fall down.
+
+Next steps are to:
+- Check that all the information has actually been imputed in the output dataset 'impData'
+- Compare the distributions of the imputed data to the original data
+- Save a dataset with the newly imputed data
+  - Run this dataset through the FEM to see if it runs, and what difference it has produced
+- Iteratively include more variables in the imputation step, and see where it fails. Try to include as many variables as possible and identify those that break the models (linearly dependant)
+
+```{r}
+
+# Now collect all imputed datasets in long format
+completedData <- complete(impData, 'long')
+
+contVars <- impData %>% select(contains(c('id', 'bmi', 'smokef', 'drinkd_e')))
+
+# Aggregate the data
+a<-aggregate(contVars , by = list(contVars$.id), FUN= mean)
+
+# Remove vars generated in imputation
+colnames(a)
+#final <- a %>% select(-c('Group.1', '.id'))
+
+```
+
+Now try again but combining all variables, not just the continuous vars.
+
+```{r}
+
+# Now collect all imputed datasets in long format
+completedData2 <- complete(impData, 'long')
+
+# Aggregate the data
+a<-aggregate(completedData2 , by = list(completedData2$.id), FUN= mean)
+
+# Remove vars generated in imputation
+colnames(a)
+final <- a %>% select(-c('Group.1', '.id'))
+
+```

### 2020-08-19 FEM_R/mul_imp_BMI.r

@@ -10,8 +10,8 @@ require(tidyverse)
 set.seed(500) # Set seed for reproducibility
 
 # Read in the reshaped_data and get a summary
-#H_ELSA_base <- read_dta('H_ELSA_pre_reshape.dta')
-H_ELSA_base <- read_dta('H_ELSA_f_2002-2016.dta')
+H_ELSA_base <- read_dta('H_ELSA_pre_reshape.dta')
+#H_ELSA_base <- read_dta('H_ELSA_f_2002-2016.dta')
 summary(H_ELSA_base)
 
 # Select only the vars we want to impute, plus those to use in the imputation
@@ -30,7 +30,7 @@ completedData <- complete(impData, 'long')
 # Extract only bmi vars
 bmiVars <- completedData %>% select(contains(c('id', 'bmi')))
 
-# Aggregate the data by 
+# Aggregate the data
 a<-aggregate(bmiVars , by = list(bmiVars$.id),FUN= mean)
 
 # Remove vars generated in imputation

### 2020-08-19 FEM_R/ELSA_data_check.Rmd
Couple more R scripts for assessing missing data in ELSA and visualising FEM outputs

--- 2020-08-12 FEM_R/ELSA_imputation.Rmd
@@
 ---
-title: "Multiple Imputation of Missing Data in ELSA"
+title: "R Notebook"
 output: html_notebook
@@
 
-This document will run through the steps taken to impute large amounts of missing data in ELSA.
-
-Missing data in ELSA is a significant problem to running the FEM, as a number of key predictive variables have a high degree of missingness and/or are not missing at random.
-For example, BMI data is missing for every odd wave, and the even waves are missing a considerable proportion of that data.
-
-First, the extent of missing information will be assessed, then a few methods for imputing these missing values will be trialled.
+This document will assess the level of missing data in some key variables to inform decisions around what data to use, and what needs to be imputed.
 
@@
 
+require(haven)
+require(mice)
+require(lattice)
+require(tidyverse)
+require(Hmisc)
+
+FEM_dir <- "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/"
+knitr::opts_knit$set(root.dir = FEM_dir)
+
+set.seed(500) # Set seed for reproducibility
+
 ```
+
+Read in the data.
+
+```{r}
+
+# Read in the data.
+H_ELSA_base <- read_dta('H_ELSA_f_2002-2016.dta')
+
+```
+
+Now declare the function to assess the level of missing data.
+
+```{r}
+
+# Calculate percentage of missing cases
+pMiss <- function(x) {sum(is.na(x)) / length(x) * 100}
+
+```
+
+Now select key variables from ELSA and apply the function to assess missingness.
+
+```{r}
+
+# Cancer
+cancre <- H_ELSA_base %>% select(matches('r[0-9]cancre$'))
+apply(cancre, 2, pMiss)
+print(' ')
+
+# Diabetes
+diabe <- H_ELSA_base %>% select(matches('r[0-9]diabe$'))
+apply(diabe, 2, pMiss)
+
+```

### 2020-08-19 FEM_R/risk_behaviour_burden.Rmd
Couple more R scripts for assessing missing data in ELSA and visualising FEM outputs

@@
 ggplot(cohort, aes(x=age, y=p_cancre_all)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
+
+```
+
+# Baseline plots (Population)
+
+```{r}
+
+pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Plot Average age at death
+ggplot(pop, aes(x=year, y=d_age_all)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')
+# Smoking variables and chronic diseases
+ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
+ggplot(pop, aes(x=year, y=p_smoken_all)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
+ggplot(pop, aes(x=year, y=p_lunge_all)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
+ggplot(pop, aes(x=year, y=p_cancre_all)) +
         geom_line(col='black', size=0.3) +

### 2020-09-09 FEM_R/ELSA_data_check.Rmd
Added lots of R scripts for visualising and checking data

@@
 apply(diabe, 2, pMiss)
+print(' ')
+
+# Lung Disease
+lunge <- H_ELSA_base %>% select(matches('r[0-9]lunge$'))
+apply(lunge, 2, pMiss)
+print(' ')
+
+# smoken and smokev and smokef
+smoken <- H_ELSA_base %>% select(matches('r[0-9]smoken$'))
+apply(smoken, 2, pMiss)
+print(' ')
+smokev <- H_ELSA_base %>% select(matches('r[0-9]smokev$'))
+apply(smokev, 2, pMiss)
+print(' ')
+smokef <- H_ELSA_base %>% select(matches('r[0-9]smokef$'))
+apply(smokef, 2, pMiss)
+print(' ')
+
+# drink
+drink <- H_ELSA_base %>% select(matches('r[0-9]drink$'))
+apply(drink, 2, pMiss)
+print(' ')
 

### 2020-09-09 FEM_R/ELSA_risk_behaviours.Rmd
Added lots of R scripts for visualising and checking data

---
title: "Burden of Risk Behaviours"
output: html_notebook
---

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(tidyr)
require(ggplot2)
```

# Baseline Plots (Cohort)

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        labs(title='Cohort Survival', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=n_smoke_start_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Smoke Start', y='# Started Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Smoking (Cohort)', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Lung Disease (Cohort)', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Cancer (Cohort)', y='Prevalence of Cancer', x='Age')
ggplot(cohort, aes(x=age, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Drink (Cohort)', y='Prevalence of Drink', x='Age')
ggplot(cohort, aes(x=age, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Stroke (Cohort)', y='Prevalence of Drink', x='Age')

```

# Baseline plots (Population)

```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_smooth(method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        labs(title='Average Age at Death', y='Age', x='Year')
# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Smoking', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Lung Disease', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of Cancer', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        geom_smooth(method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        labs(title='Prevalence of Drink', y='Proportion of Total Pop.', x='Year')

ggplot(pop, aes(x=year, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of HBP', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_drinkd1)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of HBP in teetotalers', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_drinkd2)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of HBP in low drinkers', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_drinkd3)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of HBP in moderate drinkers', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_drinkd4)) +
        geom_line(col='black', size=0.3) +
        labs(title='Prevalence of HBP in heavy drinkers', y='Proportion of Total Pop.', x='Year')

ggplot(pop, aes(x=year, y=a_bmi_all)) +
        geom_line(col='black', size=0.3) +
        labs(title='Average BMI', y='BMI', x='Year')

ggplot(pop, aes(x=year, y=a_adlstat_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(col='red', size=0.3, aes(x=year, y=a_adlstat_drinkd1, color="No Drinking")) +
        geom_line(col='blue', size=0.3, aes(x=year, y=a_adlstat_drinkd2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=a_adlstat_drinkd3)) +
        geom_line(col='orange', size=0.3, aes(x=year, y=a_adlstat_drinkd4)) +
        labs(title='Average level of Disability', y='Disability Level', x='Year')

```

# Show differences within drinkers and exercisers

```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')

colours <- c("Total Population" = "black", "No Drinking" = "red", "Light Drinker" = "blue", "Moderate Drinker" = "Dark Green", "Heavy Drinker" = "orange")

ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.5) +
        labs(title='Prevalence of Heart Disease', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.5, aes(x=year, y=p_hearte_all, color="Total Population")) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_hearte_drinkd1, color="No Drinking")) +
        labs(title='Prevalence of Heart Disease', y='Proportion of Total Pop.', x='Year', color="Legend") +
        scale_color_manual(values = colours, labels = c("Total Population", "No Drinking"))
ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_hearte_drinkd1, color="No Drinking")) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_hearte_drinkd2, color=group)) +
        labs(title='Prevalence of Heart Disease', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_hearte_drinkd1)) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_hearte_drinkd2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=p_hearte_drinkd3)) +
        labs(title='Prevalence of Heart Disease', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_hearte_drinkd1)) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_hearte_drinkd2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=p_hearte_drinkd3)) +
        geom_line(col='orange', size=0.3, aes(x=year, y=p_hearte_drinkd4)) +
        labs(title='Prevalence of Heart Disease', y='Proportion of Total Pop.', x='Year')

ggplot(pop, aes(x=year, y=p_asthmae_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_asthmae_drinkd1)) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_asthmae_drinkd2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=p_asthmae_drinkd3)) +
        geom_line(col='orange', size=0.3, aes(x=year, y=p_asthmae_drinkd4)) +
        labs(title='Prevalence of Asthma', y='Proportion of Total Pop.', x='Year')

ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_diabe_drinkd1)) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_diabe_drinkd2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=p_diabe_drinkd3)) +
        geom_line(col='orange', size=0.3, aes(x=year, y=p_diabe_drinkd4)) +
        labs(title='Prevalence of Diabetes', y='Proportion of Total Pop.', x='Year')

ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.5) +
        geom_line(col='red', size=0.3, aes(x=year, y=p_diabe_exstat1)) +
        geom_line(col='blue', size=0.3, aes(x=year, y=p_diabe_exstat2)) +
        geom_line(col='dark green', size=0.3, aes(x=year, y=p_diabe_exstat3)) +
        labs(title='Prevalence of Diabetes', y='Proportion of Total Pop.', x='Year')

```

# Reshaping data before plotting for legend
```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')

# Specify colour palette
colours <- c("Total Population" = "black", "No Alcohol" = "red", "Light Drinker" = "blue", "Moderate Drinker" = "Dark Green", "Heavy Drinker" = "orange")

pop_copy <- rename(pop, "Total Population" = p_hearte_all, "No Alcohol" = p_hearte_drinkd1, "Light Drinker" = p_hearte_drinkd2, "Moderate Drinker" = p_hearte_drinkd3, "Heavy Drinker" = p_hearte_drinkd4)

pop_gathered <- gather(pop_copy, key = Subpopulation, value = Rate, c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker", "Heavy Drinker"))
#pop_gathered <- gather(pop_copy, key = Subpopulation, value = Rate, c("Total Population", "No Alcohol"))
#pop_gathered <- gather(pop_copy, key = Subpopulation, value = Rate, c("Total Population"))
#pop_gathered <- gather(pop_copy, key = Subpopulation, value = Rate, c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker"))

pop_gathered$Subpopulation <- factor(pop_gathered$Subpopulation, levels = c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker", "Heavy Drinker"))
#pop_gathered$Subpopulation <- factor(pop_gathered$Subpopulation, levels = c("Total Population", "No Alcohol"))
#pop_gathered$Subpopulation <- factor(pop_gathered$Subpopulation, levels = c("Total Population"))
#pop_gathered$Subpopulation <- factor(pop_gathered$Subpopulation, levels = c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker"))

ggplot(pop_gathered, aes(x=year, y= Rate, group = Subpopulation, colour = Subpopulation)) + 
        geom_line() +
        labs(title='Heart Disease', 
             y='Proportion of Total Pop.', 
             x='Year') +
        scale_color_manual(name = "Subpopulation", labels=c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker", "Heavy Drinker"), values = colours)

```

# Init no smoking interventions (cohort and pop)

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
initNoSmokeC <- read_dta('output/initNoSmokeC/initNoSmokeC_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
initNoSmokeC$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
initNoSmokeC$survival <- initNoSmokeC$m_endpop_all / initNoSmokeC$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeC, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeC, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeC, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeC, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeC, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
initNoSmokeP <- read_dta('output/initNoSmokeP/initNoSmokeP_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeP, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeP, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeP, size=0.3, col='red') +
        labs(title='Prevalence of Smoking', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeP, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=initNoSmokeP, size=0.3, col='red') +
        labs(title='Prevalence of Cancer', y='Prevalence of Cancer', x='Year')

```

# Do some silly shit again to get a legend

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
initNoSmokeP <- read_dta('output/initNoSmokeP/initNoSmokeP_summary.dta')

colours_smoke <- c("Baseline" = "black", "No Smoking" = "red")

pop_lunge <- select(pop, c(year, p_lunge_all))
noSmoke <- select(initNoSmokeP, c(year, p_lunge_all))

#combined <- merge(pop_lunge, noSmoke)
combined <- merge.data.frame(pop_lunge, noSmoke, by="year")
combined <- rename(combined, "Baseline" = p_lunge_all.x, "No Smoking" = p_lunge_all.y)

comb_gather <- gather(combined, key = Scenario, value = Rate, c("Baseline", "No Smoking"))

ggplot(comb_gather, aes(x=year, y= Rate, group = Scenario, colour = Scenario)) + 
        geom_line() +
        labs(title='Lung Disease', 
             y='Proportion of Total Pop.', 
             x='Year') +
        scale_color_manual(name = "Scenario", labels=c("Baseline", "No Smoking"), values = colours_smoke)

```

# Init no drinking interventions (cohort and pop)

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
noDrinkC <- read_dta('output/initNoDrinkC/initNoDrinkC_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
noDrinkC$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
noDrinkC$survival <- noDrinkC$m_endpop_all / noDrinkC$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
ggplot(cohort, aes(x=age, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Drink (Cohort)', y='Prevalence of Drink', x='Age')
ggplot(cohort, aes(x=age, y=a_bmi_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Average BMI by Age', y='Average BMI', x='Age')
ggplot(cohort, aes(x=age, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes by Age', y='Prevalence of Diabetes', x='Age')
ggplot(cohort, aes(x=age, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension by Age', y='Prevalence of Hypertension', x='Age')
ggplot(cohort, aes(x=age, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease by Age', y='Prevalence of Heart Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
ggplot(cohort, aes(x=age, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Stroke by Age', y='Prevalence of Stroke', x='Age')
ggplot(cohort, aes(x=age, y=p_asthmae_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkC, size=0.3, col='red') +
        labs(title='Prevalence of Asthma by Age', y='Prevalence of Asthma', x='Age')

```

```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
noDrinkP <- read_dta('output/initNoDrinkP/initNoDrinkP_summary.dta')

# Now plot all survival curves together
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_smooth(method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        geom_smooth(data=noDrinkP, method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        labs(title='Average Age at Death Over Time', y='Average Age at Death', x='Year')
ggplot(pop, aes(x=year, y=p_drink_all)) +
        geom_line(col='black', size=0.3) +
        geom_smooth(method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        geom_smooth(data=noDrinkP, method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        labs(title='Prevalence of Drink', y='Proportion of Total Pop.', x='Year')
ggplot(pop, aes(x=year, y=a_bmi_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Average BMI Over Time', y='Average BMI', x='Year')
ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Diabetes Over Time', y='Prevalence of Diabetes', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Hypertension Over Time', y='Prevalence of Hypertension', x='Year')
ggplot(pop, aes(x=year, y=p_hearte_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Heart Disease Over Time', y='Prevalence of Heart Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Cancer Over Time', y='Prevalence of Cancer', x='Year')
ggplot(pop, aes(x=year, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Stroke Over Time', y='Prevalence of Stroke', x='Year')
ggplot(pop, aes(x=year, y=p_asthmae_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=noDrinkP, size=0.3, col='red') +
        labs(title='Prevalence of Asthma Over Time', y='Prevalence of Asthma', x='Year')

```

# No Smoking Scenario (Cohort)

```{r}

# Read in all the data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
no_smoke_C <- read_dta('output/CureSmoken_C/CureSmoken_C_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
no_smoke_C$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
no_smoke_C$survival <- no_smoke_C$m_endpop_all / no_smoke_C$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
ggplot(cohort, aes(x=age, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_C, size=0.3, col='green') +
        labs(title='Prevalence of High Blood Pressure  by Age', y='Prevalence of Cancer', x='Age')

```

# SmokeStopFull (cohort)

```{r}

# Read in all the data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
SSF_C <- read_dta('output/SmokeStopFull_C/SmokeStopFull_C_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
SSF_C$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
SSF_C$survival <- SSF_C$m_endpop_all / SSF_C$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_C, size=0.3, col='green') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# No Smoking Scenario (Population)

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
no_smoke_P <- read_dta('output/CureSmoken_P/CureSmoken_P_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=no_smoke_P, size=0.3, col='green') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```

# SmokeStopFull (Population)

```{r}

# Read in all the data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
SSF_P <- read_dta('output/SmokeStopFull_P/SmokeStopFull_P_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSF_P, size=0.3, col='green') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```

### 2020-09-09 FEM_R/ERSA_plots.Rmd
Added lots of R scripts for visualising and checking data

---
title: "ERSA PLOTS"
output: html_notebook
---

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(tidyr)
require(ggplot2)
```

# COHORT

```{r}

cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')

colours.mf <- c("Male" = "red", "Female" = "blue")

# Generate age var
cohort$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
cohort$survivalm <- cohort$m_endpop_m / cohort$m_endpop_m[1]
cohort$survivalf <- cohort$m_endpop_f / cohort$m_endpop_f[1]

cohort_cp <- rename(cohort, Male = survivalm, Female = survivalf)

cohort_gath <- gather(cohort_cp, key = Subpopulation, value = Rate, c("Male", "Female"))
#pop_gathered$Subpopulation <- factor(pop_gathered$Subpopulation, levels = c("Total Population", "No Alcohol", "Light Drinker", "Moderate Drinker", "Heavy Drinker"))
cohort_gath$Subpopulation <- factor(cohort_gath$Subpopulation, levels = c("Female", "Male"))

ggplot(cohort_gath, aes(x=year, y= Rate, group = Subpopulation, colour = Subpopulation)) + 
        geom_line() +
        labs(title='Cohort Survival', 
             y='Survival', 
             x='Year') +
        scale_color_manual(name = "Subpopulation", labels=c("Female", "Male"), values = colours.mf)

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        labs(title='Cohort Survival', y='Survival', x='Age')

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='red', size=0.3, aes(y=survivalm)) +
        geom_line(col='blue', size=0.3, aes(y=survivalf)) +
        labs(title='Cohort Survival', y='Survival', x='Age')

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2020-09-09 FEM_R/pubHealthInterventionVis_Notebook.Rmd
Added lots of R scripts for visualising and checking data

@@
         labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
+ggplot(cohort, aes(x=age, y=n_smoke_start_all)) +
+        geom_line(col='black', size=0.3) +
+        geom_line(data=x169, size=0.3, col='red') +
+        labs(title='Smoke Start', y='# Started Smoking', x='Age')
 ggplot(cohort, aes(x=age, y=p_smoken_all)) +

### 2020-10-30 FEM_R/visualise_chron_diseases.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(tidyr)
require(ggplot2)
```

```{r}

pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
diabe10 <- read_dta('output/ELSA_pDiabe10/ELSA_pDiabe10_summary.dta')
diabe50 <- read_dta('output/ELSA_pDiabe50/ELSA_pDiabe50_summary.dta')
diabe100 <- read_dta('output/ELSA_pDiabe100/ELSA_pDiabe100_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        #geom_smooth(method="auto", se=FALSE, fullrange=FALSE, level=0.95) +
        geom_line(data=diabe10, size=0.3, col='blue') +
        geom_line(data=diabe50, size=0.3, col='red') +
        geom_line(data=diabe100, size=0.3, col='dark green') +
        labs(title='Average Age at Death', y='Age', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=p_diabe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=diabe10, size=0.3, col='blue') +
        geom_line(data=diabe50, size=0.3, col='red') +
        geom_line(data=diabe100, size=0.3, col='dark green') +
        labs(title='Diabetes', y='Prevalence of Diabetes', x='Year')
ggplot(pop, aes(x=year, y=p_hibpe_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=diabe10, size=0.3, col='blue') +
        geom_line(data=diabe50, size=0.3, col='red') +
        geom_line(data=diabe100, size=0.3, col='dark green') +
        labs(title='Prevalence of Hypertension', y='Prevalence of Hypertension', x='Year')
ggplot(pop, aes(x=year, y=p_stroke_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=diabe10, size=0.3, col='blue') +
        geom_line(data=diabe50, size=0.3, col='red') +
        geom_line(data=diabe100, size=0.3, col='dark green') +
        labs(title='Prevalence of Stroke', y='Prevalence of Stroke', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=diabe10, size=0.3, col='blue') +
        geom_line(data=diabe50, size=0.3, col='red') +
        geom_line(data=diabe100, size=0.3, col='dark green') +
        labs(title='Prevalence of Cancer', y='Prevalence of Cancer', x='Year')

```

### 2020-10-31 FEM_R/IJM_paper1_vis_WIP.Rmd

---
title: "Visualisations for FEM IJM paper 1"
output: html_notebook
---

This notebook will contain all the visualisations for the first paper on the FEM, to be submitted to the International Journal of Microsimulation.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(gridExtra)

```

Setup is complete, working directory is set as the output dir, and the necessary packages have been loaded.

# Disability Prevalence

```{r}

# This code block will handle calculating the proportion of respondents that are disabled at baseline and at the year 2040 (28 years, 14 waves into simulation)
# We will calculate this information for each repetition so we can produce confidence intervals or some other extra information

# First read in the output datasets (by rep)
baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_by_rep.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_by_rep.dta')

```

Below we define a function to get the prevalence of anyadl both at 2012 and 2040. The prevalence of anyadl will always be the same in 2012, as this information is produced straight from the stock population which is the same for every scenario. The 2040 prevalence will change however, so we should report the relative change via a graph or table (or both). As we have the XXXXX_by_rep.dta files also, we can calculate the confidence intervals for the 2040 prevalence (the 2012 confint's will always be 0).

```{r}

get_disability_prev <- function(data) {

  # Select only the columns we want, and filter to have a separate dataset for 2012 and 2040
  adl <- select(data, year, rep, a_anyadl_all)
  adl_2012 <- filter(adl, year == 2012)
  adl_2040 <- filter(adl, year == 2040)

  # Calculate mean, standard deviation and size of the samples for later
  m_12 <- mean(adl_2012$a_anyadl_all)
  sd_12 <- sd(adl_2012$a_anyadl_all)
  n_12 <- length(adl_2012$a_anyadl_all)

  m_40 <- mean(adl_2040$a_anyadl_all)
  sd_40 <- sd(adl_2040$a_anyadl_all)
  n_40 <- length(adl_2040$a_anyadl_all)

  # Now calculate the error and confidence intervals for both 2012 and then 2040
  error_12 <- qnorm(0.975) * sd_12 / sqrt(n_12)
  left_12 <- m_12 - error_12
  right_12 <- m_12 + error_12

  error_40 <- qnorm(0.975) * sd_40 / sqrt(n_40)
  left_40 <- m_40 - error_40
  right_40 <- m_40 + error_40

  # Collect information into vectors to add to a dataframe
  avg <- c(m_12, m_40)
  error <- c(error_12, error_40)
  left <- c(left_12, left_40)
  right <- c(right_12, right_40)

  df <- data.frame(avg, error, left, right)

  rownames(df) <- c('2012', '2040')

  return(df)
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_by_rep.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_by_rep.dta')
diabe100 <- read_dta('ELSA_pdiabe100/ELSA_pdiabe100_by_rep.dta')
hearte100 <- read_dta('ELSA_phearte100/ELSA_phearte100_by_rep.dta')

df_base <- get_disability_prev(baseline)
df_cancre100 <- get_disability_prev(cancre100)
df_diabe100 <- get_disability_prev(diabe100)
df_hearte100 <- get_disability_prev(hearte100)

```

```{r}

# Now collect all the information into a single dataset for visualisation
# Columns are going to be: prev_2012, error_2012, prev_2040, error_2040
prev_2012 <- c(df_base$avg[1], df_cancre100$avg[1], df_diabe100$avg[1], df_hearte100$avg[1])
prev_2040 <- c(df_base$avg[2], df_cancre100$avg[2], df_diabe100$avg[2], df_hearte100$avg[2])
error_2012 <- c(df_base$error[1], df_cancre100$error[1], df_diabe100$error[1], df_hearte100$error[1])
error_2040 <- c(df_base$error[2], df_cancre100$error[2], df_diabe100$error[2], df_hearte100$error[2])

full_df <- data.frame(prev_2012, error_2012, prev_2040, error_2040)
full_df$scenario <- c('Baseline', 'Cancer-100', 'Diabetes-100', 'Heart Disease-100')

```

Going to have another go at the get_disability_prev() function to get the output in a better format. We don't need to report the 2012 prevalence or error as it will always be the same. Instead, have 1 function to calculate the 2040 prevalence, then another function that calculates the 2012 and finds the difference between the 2.

```{r}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_by_rep.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_by_rep.dta')
diabe100 <- read_dta('ELSA_pdiabe100/ELSA_pdiabe100_by_rep.dta')

b_adl <- select(baseline, year, rep, a_anyadl_all)
#c_adl <- select(baseline, year, rep, a_anyadl_all)

b_2012 <- filter(b_adl, year == 2012)
b_2040 <- filter(b_adl, year == 2040)

m <- mean(b_2040$a_anyadl_all)
sd <- sd(b_2040$a_anyadl_all)
n <- length(b_2040$a_anyadl_all)

error <- qnorm(0.975) * sd / sqrt(n)

res <- c(mean(b_2012$a_anyadl_all), mean(b_2040$a_anyadl_all), error)

```

Now collect the above into a function and run it on a list with lapply.

```{r}

get_dis_prev3 <- function(data, target_year=2040) {

  adl <- select(data, year, rep, a_anyadl_all)

  a_2012 <- filter(adl, year == 2012)
  a_ty <- filter(adl, year == target_year)

  sd <- sd(a_ty$a_anyadl_all)
  n <- length(a_ty$a_anyadl_all)

  error <- qnorm(0.975) * sd / sqrt(n)

  res <- c(mean(a_2012$a_anyadl_all), mean(a_ty$a_anyadl_all), error)

  return(res)
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_by_rep.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_by_rep.dta')
diabe100 <- read_dta('ELSA_pdiabe100/ELSA_pdiabe100_by_rep.dta')
hearte100 <- read_dta('ELSA_phearte100/ELSA_phearte100_by_rep.dta')

base <- get_dis_prev3(baseline, 2050)
cancre_100 <- get_dis_prev3(cancre100, 2050)
diabe_100 <- get_dis_prev3(diabe100, 2050)
hearte_100 <- get_dis_prev3(hearte100, 2050)

df <- data.frame(rbind(base, cancre_100, diabe_100, hearte_100))
colnames(df) <- c('p2012', 'p2040', 'error_2040')

df$scenario <- c('Baseline', 'Cancer-100', 'Diabetes-100', 'Heart Disease-100')
df$diff_to_baseline <- df$p2040 - df$p2040[1]

```

```{r}

p1 <- ggplot(df) +
        geom_bar(aes(x=scenario, y=p2040), stat="identity", fill='blue', alpha=0.7) +
        geom_errorbar(aes(x=scenario, ymin=p2040 - error_2040, ymax=p2040 + error_2040), width=0.4, colour='red', alpha=0.9, size=0.5)

p2 <- ggplot(df) +
        geom_bar(aes(x=scenario, y=diff_to_baseline), stat="identity", fill='blue', alpha=0.7) +
        geom_errorbar(aes(x=scenario, ymin = diff_to_baseline - error_2040, ymax = diff_to_baseline + error_2040), width=0.4, colour='red', alpha=0.9, size=0.5)

#multiplot(p1, p2, cols=2)

grid.arrange(p1, p2, nrow=2)
```

# Chronic Disease Prevalence (Base)
## Projections

```{r}

visPop_chronDis2 <- function(base, ten, fifty, hundred, disease) {

  # Construct some varnames using argument to keep function generic
  scen_base <- paste0(disease, '_base')
  scen10 <- paste0(disease, '_10')
  scen50 <- paste0(disease, '_50')
  scen100 <- paste0(disease, '_100')

  # Define colours for graphing
  if (disease == 'cancre') {
    colours <- c('cancre_base' = 'black', 'cancre_10' = 'blue', 'cancre_50' = 'dark green', 'cancre_100' = 'red')
    chronDis <- 'Cancer'
  } else if (disease == 'diabe') {
    colours <- c('diabe_base' = 'black', 'diabe_10' = 'blue', 'diabe_50' = 'dark green', 'diabe_100' = 'red')
    chronDis <- 'Diabetes'
  } else if (disease == 'hearte') {
    colours <- c('hearte_base' = 'black', 'hearte_10' = 'blue', 'hearte_50' = 'dark green', 'hearte_100' = 'red')
    chronDis <- 'Heart Disease'
  } else {
    print('The disease entered is wrong, please pick one of cancer, diabetes, heart disease.')
  }

  # These are the variables we want to visualise here
  var_select <- c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')

  # Select only the vars we want to see
  dis_base <- select(base, all_of(var_select))
  dis_10 <- select(ten, all_of(var_select))
  dis_50 <- select(fifty, all_of(var_select))
  dis_100 <- select(hundred, all_of(var_select))

  # Rename for unique colnames when merging
  dis_base <- rename(dis_base, 'cancre_base' = p_cancre_all, 'diabe_base' = p_diabe_all, 'hearte_base' = p_hearte_all)
  dis_10 <- rename(dis_10, 'cancre_10' = p_cancre_all, 'diabe_10' = p_diabe_all, 'hearte_10' = p_hearte_all)
  dis_50 <- rename(dis_50, 'cancre_50' = p_cancre_all, 'diabe_50' = p_diabe_all, 'hearte_50' = p_hearte_all)
  dis_100 <- rename(dis_100, 'cancre_100' = p_cancre_all, 'diabe_100' = p_diabe_all, 'hearte_100' = p_hearte_all)
  # Collect in a list to facilitate merge with multiple df's
  df_list <- list(dis_base, dis_10, dis_50, dis_100)

  # Merge on year var
  combined <- df_list %>% reduce(inner_join, by = 'year')

  # Reshape the data to long, can then extract the scenarios we want and visualise
  melted <- melt(combined, id.var = 'year', variable.name = 'scenario')

  scen_subset <- filter(melted, scenario == scen_base | scenario == scen10 | scenario == scen50 | scenario == scen100)
  scen_subset$scenario <- factor(as.character(scen_subset$scenario),
                                 levels = c(scen_base, scen10, scen50, scen100))

  #plot1 <- 
  ggplot(scen_subset, aes(x=year, y=value, group = scenario, colour = scenario)) +
              geom_line() +
              labs(title = chronDis, 
                  y = 'Prevalence',
                  x = 'Year') +
              scale_colour_manual(name = "Scenario", labels = c('Baseline', '10%', '50%', '100%'), values = colours)

  #outfile <- paste0('/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/ChronDisease/', disease, '/cancre_prev_reduc.png')
  #ggsave(plot=plot1, filename=outfile, height=5, width=10)
}

```

## Baseline Projections

```{r}

visProjectedBaselineCD <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/') {

  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                melt(id.var='year', value.name = 'Prevalence')

  # Create the tiled plot of disease prevalence
  tiled <- ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
            geom_line() +
            facet_grid(. ~ variable) +
            theme(panel.spacing.x = unit(1.5, 'lines')) +
            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'baseline_prevalence.png', 
         plot = tiled,
         path = path,
         width = 10,
         height = 5)

  # Show plot
  tiled
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBaselineCD(baseline)

```

Now we have the baseline chronDis projections, we want to show this split by different groups. I.e. split into obese/not obese, or by age splines.

```{r}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

hearteMelt <- baseline %>%
              select(c('year','p_hearte_all', 'p_hearte_obese', 'p_hearte_notObese')) %>%
              rename('Total' = p_hearte_all, 'Obese' = p_hearte_obese, 'Healthy Weight' = p_hearte_notObese) %>%
              melt(id.var='year', value.name = 'Prevalence')

#ggplot(data=hearteMelt, aes(x=year, y=Prevalence, color=variable)) +
#  geom_line()

visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/') {

  # Generate varnames
  all <- paste0('p_', cd, '_all')
  obese <- paste0('p_', cd, '_obese')
  notObese <- paste0('p_', cd, '_notObese')
  adl <- paste0('p_', cd, '_anyadl')
  noadl <- paste0('p_', cd, '_noadl')

  obeseMelt <- base %>%
              select(c('year', all, obese, notObese)) %>%
              rename('All' = all, 'Obese' = obese, 'Healthy Weight' = notObese) %>%
              melt(id.var='year', value.name = 'Prevalence')

  adlMelt <- base %>%
              select(c('year', all, adl, noadl)) %>%
              rename('All' = all, 'Disabled' = adl, 'Able' = noadl) %>%
              melt(id.var='year', value.name = 'Prevalence')

  p1 <- ggplot(data=obeseMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line()

  f1 <- paste0(cd, '_prev_by_obesity.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f1, 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  p2 <- ggplot(data=adlMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line()

  f2 <- paste0(cd, '_prev_by_disability.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f2, 
         plot = p2,
         path = path,
         width = 6,
         height = 5)

}

visProjectedBySubpop(baseline, cd = 'cancre')
visProjectedBySubpop(baseline, cd = 'diabe')
visProjectedBySubpop(baseline, cd = 'hearte')

```

## Initial prevalence

```{r}

visInitBasePrev <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/') {

  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                filter(year == 2012) %>%
                melt(id.var = 'year', value.name = 'Prevalence') %>%
                subset(select = -year)

  bar <- ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
          geom_bar(stat = 'identity') +
          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'init_baseline_prevalence.png', 
         plot = bar,
         path = path,
         width = 7,
         height = 5)

  bar
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visInitBasePrev(baseline)

```

Now the initial prevalence for the 'all' group is visualised, let's look into some of the subgroups. i.e. obese, maybe an age comparison

# OLD WORK (Depracated)

The code below is depracated. It is the first try at a function to visualise cohort population outputs.

```{r}

visPop_chronDis <- function(base, ten, fifty, hundred, disease) {

  # Construct some varnames using argument to keep function generic
  scen_base <- paste0(disease, '_base')
  scen_10 <- paste0(disease, '_10')
  scen_50 <- paste0(disease, '_50')
  scen_100 <- paste0(disease, '_100')

  # Define colours for graphing
  if (disease == 'cancre') {
    colours <- c('cancre_base' = 'black', 'cancre_10' = 'blue', 'cancre_50' = 'dark green', 'cancre_100' = 'red')
    chronDis <- 'Cancer'
  } else if (disease == 'diabe') {
    colours <- c('diabe_base' = 'black', 'diabe_10' = 'blue', 'diabe_50' = 'dark green', 'diabe_100' = 'red')
    chronDis <- 'Diabetes'
  } else if (disease == 'hearte') {
    colours <- c('hearte_base' = 'black', 'hearte_10' = 'blue', 'hearte_50' = 'dark green', 'hearte_100' = 'red')
    chronDis <- 'Heart Disease'
  } else {
    print('The disease entered is wrong, please pick one of cancer, diabetes, heart disease.')
  }

  # These are the variables we want to visualise here
  var_select <- c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')

  # Select only the vars we want to see
  dis_base <- select(base, all_of(var_select))
  dis_10 <- select(ten, all_of(var_select))
  dis_50 <- select(fifty, all_of(var_select))
  dis_100 <- select(hundred, all_of(var_select))

  # Rename for unique colnames when merging
  dis_base <- rename(dis_base, 'cancre_base' = p_cancre_all, 'diabe_base' = p_diabe_all, 'hearte_base' = p_hearte_all)
  dis_10 <- rename(dis_10, 'cancre_10' = p_cancre_all, 'diabe_10' = p_diabe_all, 'hearte_10' = p_hearte_all)
  dis_50 <- rename(dis_50, 'cancre_50' = p_cancre_all, 'diabe_50' = p_diabe_all, 'hearte_50' = p_hearte_all)
  dis_100 <- rename(dis_100, 'cancre_100' = p_cancre_all, 'diabe_100' = p_diabe_all, 'hearte_100' = p_hearte_all)
  # Collect in a list to facilitate merge with multiple df's
  df_list <- list(dis_base, dis_10, dis_50, dis_100)

  # Merge on year var
  combined <- df_list %>% reduce(inner_join, by = 'year')

  scen_gather <- gather(combined, key = Scenario, value = Rate, c(all_of(scen_base), all_of(scen_10), all_of(scen_50), all_of(scen_100)))

  ggplot(scen_gather, aes(x=year, y=Rate, group = Scenario, colour = Scenario)) +
    geom_line() +
    labs(title = chronDis, 
         y = 'Prevalence',
         x = 'Year') +
    scale_colour_manual(name = "Scenario", labels = c('Baseline', '10%', '50%', '100%'), values = colours)

}

pop <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
cancre10 <- read_dta('ELSA_pCancer10/ELSA_pCancer10_summary.dta')
cancre50 <- read_dta('ELSA_pCancer50/ELSA_pCancer50_summary.dta')
cancre100 <- read_dta('ELSA_pCancer100/ELSA_pCancer100_summary.dta')

visPop_chronDis(pop, cancre10, cancre50, cancre100, 'cancre')
visPop_chronDis(pop, cancre10, cancre50, cancre100, 'diabe')
visPop_chronDis(pop, cancre10, cancre50, cancre100, 'hearte')

pop <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
diabe10 <- read_dta('ELSA_pDiabe10/ELSA_pDiabe10_summary.dta')
diabe50 <- read_dta('ELSA_pDiabe50/ELSA_pDiabe50_summary.dta')
diabe100 <- read_dta('ELSA_pDiabe100/ELSA_pDiabe100_summary.dta')

visPop_chronDis(pop, diabe10, diabe50, diabe100, 'cancre')
visPop_chronDis(pop, diabe10, diabe50, diabe100, 'diabe')
visPop_chronDis(pop, diabe10, diabe50, diabe100, 'hearte')

pop <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
hearte10 <- read_dta('ELSA_pHearte10/ELSA_pHearte10_summary.dta')
hearte50 <- read_dta('ELSA_pHearte50/ELSA_pHearte50_summary.dta')
hearte100 <- read_dta('ELSA_pHearte100/ELSA_pHearte100_summary.dta')

visPop_chronDis(pop, hearte10, hearte50, hearte100, 'cancre')
visPop_chronDis(pop, hearte10, hearte50, hearte100, 'diabe')
visPop_chronDis(pop, hearte10, hearte50, hearte100, 'hearte')

```

### 2020-11-08 FEM_R/IJM_paper1_vis_WIP.Rmd
Added chunk for visualising the ReduceBMI intervention for testing

@@
 
+# REDUCE BMI Intervention
+
+```{r}
+
+base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
+pop_rbmi <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')
+coh_rbmi <- read_dta('ELSA_coh_BMI_reduc_obese/ELSA_coh_BMI_reduc_obese_summary.dta')
+
+ggplot(data = base, aes(x=year, y=a_bmi_all)) +
+  geom_line()
+
+ggplot(data = pop_rbmi, aes(x=year, y=a_bmi_all)) +
+  geom_line()
+
+ggplot(data=base, aes(x=year, y=p_cancre_all)) +
+  geom_line()
+ggplot(data=pop_rbmi, aes(x=year, y=p_cancre_all)) +
+  geom_line()
+
+ggplot(data=base, aes(x=year, y=p_diabe_all)) +
+  geom_line()
+ggplot(data=pop_rbmi, aes(x=year, y=p_diabe_all)) +
+  geom_line()
+
+ggplot(data=base, aes(x=year, y=p_hearte_all)) +
+  geom_line()
+ggplot(data=pop_rbmi, aes(x=year, y=p_hearte_all)) +
+  geom_line()
+
+```
+
 # OLD WORK (Depracated)

### 2020-11-16 FEM_R/IJM_paper_vis_NEAT.Rmd
Copied visualisation script into FEM_R from the IJM paper directory. Added chunk to visualise cross-validation 2 output

---
title: "IJM Paper Visualisations NEAT"
output: html_notebook
---

This notebook will hold the finished visualisation functions written in IJM_paper1_vis_WIP.Rmd.

This is just so we can come to a single notebook and rerun ALL visualisations that will go into the IJM FEM paper.

Start with setup.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(plyr)
require(data.table)

```

# Validation Stuff

## Cross-validation 2

In this scenario, transition models were estimated from waves 1-4, and the stock population was based on wave 5, with a view to simulating from wave 5-8 (2010-2016). We want to check whether there is a big difference as the model takes over from ELSA data from 2008-2010, in terms of chronic disease prevalence etc.

```{r}

vis_CV2_prev <- function(long, CV2, path) {

  # Keep only waves 1 - 4
  longC <- subset(long, wave < 5)
  longC <- as.data.table(longC)

  # Replace all NA with 0
  longC[is.na(longC)] <- 0

  # Have to drop r*clust vars as they contain strings
  longC <- select(longC, -contains('clust'))

  # Now get the weighted mean by wave of every column based on cwtresp
  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]

  # ELSA_long is now processed to the point where we can start extracting and combining variables
  # Better start by adding year to l2 and reordering the columns
  l2$year <- seq(from=2002, to = 2008, by = 2)
  setcolorder(l2, c('year', 'wave'))

  # Now lets extract chronic disease variables from both dataframes and do a row combine
  # Start with cancre, diabe, hearte
  cd_l2 <- select(l2, c(year, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl, anyiadl)) #, adl1, adl2, adl3p))
  cd_CV2 <- select(CV2, c(year, p_cancre_all, p_diabe_all, 
                            p_hearte_all, p_lunge_all, p_smoken_all, 
                            a_bmi_all, p_work_all,  p_anyadl_all, p_anyiadl_all))
                            #p_adl1_all, p_adl2_all, p_adl3p_all))

  CV2_rename <- cd_CV2 %>% dplyr::rename('cancre' = p_cancre_all, 
                                        'diabe' = p_diabe_all, 
                                        'hearte' = p_hearte_all,
                                        'lunge' = p_lunge_all,
                                        'smoken' = p_smoken_all,
                                        'bmi' = a_bmi_all,
                                        'work' = p_work_all,
                                        'anyadl' = p_anyadl_all,
                                        'anyiadl' = p_anyiadl_all)
                                        #'adl1' = p_adl1_all,
                                        #'adl2' = p_adl2_all,
                                        #'adl3p' = p_adl3p_all)

  combine <- rbind(cd_l2, CV2_rename)

  melted <- melt(combine, id.var = 'year', value.name = 'Prevalence')

  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
          geom_line() +
          facet_wrap(. ~ variable, scales = "free")
          #theme(panel.spacing.x = unit(1.5, 'lines')) +
          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
}

# Load in whole stock pop, plus the CV2 summary output file
long <- read_dta('../input_data/ELSA_long.dta')
CV2 <- read_dta('ELSA_CrossValidation2/ELSA_CrossValidation2_summary.dta')

vis_CV2_prev(long, CV2, path = 'path')

```

# ReduceBMI (NEW SCENARIO I - FULL POPULATION)

## Baseline Initial Prevalence

```{r}
### visInitBasePrev
# Inputs:   base
#           path
#
# Output:   bar - 
#
# This function... 
visInitBasePrev <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                filter(year == 2012) %>%
                melt(id.var = 'year', value.name = 'Prevalence') %>%
                subset(select = -year)

  bar <- ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
          geom_bar(stat = 'identity') +
          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'init_baseline_prevalence.png', 
         plot = bar,
         path = path,
         width = 7,
         height = 5)

  bar
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visInitBasePrev(baseline)

```

## Baseline Projections

```{r}
### visProjectedBaselineCD
# Inputs:   base - summary output file from the ELSA_Baseline scenario
#           path - path to the directory where the resulting plot will be saved
#
# Output:   tiled - A 3 column tiled plot with the projected prevalence of Cancer, Diabetes, and Heart Disease on the same x axis, with year on the y
#
# This function visualises the prevalence of Cancer, Diabetes, and Heart Disease in the Baseline projection of the FEM
visProjectedBaselineCD <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                melt(id.var='year', value.name = 'Prevalence')

  # Create the tiled plot of disease prevalence
  tiled <- ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
            geom_line() +
            facet_grid(. ~ variable) +
            theme(panel.spacing.x = unit(1.5, 'lines')) +
            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'baseline_prevalence.png', 
         plot = tiled,
         path = path,
         width = 10,
         height = 5)

  # Show plot
  tiled
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBaselineCD(baseline)

```

## ChronDisease Prevalence Comparison

```{r}

visCDComparison <- function(base, int, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # For file naming and stuff
  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  }

  var <- paste0('p_', cd, '_all')

  selBase <- baseline %>%
              select(c('year', all_of(var))) %>%
              rename('Baseline' = var)

  selInt <- int %>%
              select(c('year', all_of(var))) %>%
              rename('Intervention' = var)

  melted <- merge(selBase, selInt, by = 'year') %>%
              melt(id.var = 'year', value.name = 'Scenario')

  p <- ggplot(data = melted, aes(x = year, y = Scenario, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f <- paste0(dis, '_prev_intervention.png')

  ggsave(filename = f, 
           plot = p,
           path = path,
           width = 6,
           height = 5)

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visCDComparison(baseline, int, 'cancre')
visCDComparison(baseline, int, 'diabe')
visCDComparison(baseline, int, 'hearte')

```

## Disability Prevalence Comparison

```{r}

# First do % change from 2012 - 2050

visDisabilityChange <- function(base, int, targetYear, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selInt <- int %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Intervention' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  # Merge base and intervention df's
  selected <- inner_join(selBase, selInt, by = 'year')

  # Calculate the difference between targetYear prev and initial prev
  newline <- sapply(selected, function(x) x[2] - x[1])

  # Add the difference as a new row, then transpose and sort out the column names
  selected <- rbind(selected, newline)
  new <- data.frame(t(selected))
  colnames(new) <- c('y2012', 'y2050', 'Difference')

  # Calculate % change, add onto the df and move rownames into their own row
  new$change <- (new$Difference / new$y2012) * 100
  new <- cbind(Scenario = rownames(new), new)
  rownames(new) <- 1:nrow(new)

  # Delete first row, artifacts from the previous preparation
  new <- slice(new, -1)

  p <- ggplot(data = new, aes(x = Scenario, y = change)) +
      geom_bar(stat = 'identity') +
      labs(title = '% Change in Disability Prevalence 2012 - 2050', x = 'Scenario',y = '% Change')

  ggsave(filename = 'disability_prev_change.png', 
           plot = p,
           path = path,
           width = 6,
           height = 5)
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visDisabilityChange(baseline, int, 2050)

```

```{r}

visDisabilityChangeOverTime <- function(base, int, targetYear, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all)

  selInt <- int %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Intervention' = p_anyadl_all)

  # Merge base and intervention df's
  selected <- inner_join(selBase, selInt, by = 'year') %>%
                melt(id.var = 'year', value.name = 'Prevalence')

  p <- ggplot(data = selected, aes(x = year, y = Prevalence, color = variable)) +
      geom_line() +
      labs(title = 'Change in Prevalence of Disability 2012 - 2060', x = 'Year', y = 'Prevalence')

  ggsave(filename = 'disability_prev_change_overTime.png', 
           plot = p,
           path = path,
           width = 6,
           height = 5)

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visDisabilityChangeOverTime(baseline, int, 2050)

```

# Smoking Cessation (SCENARIO II - COHORT)

## Baseline CD prevalence and survival curve

```{r}

visCohortBase <- function(cohort, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_2/') {
  # Generate age var
  cohort$Age <- (cohort$year - 2012) + 50
  # Survival curve
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]

  cdMelt <- cohort %>%
              select(c('Age', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
              rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
              melt(id.var='Age', value.name='Prevalence')

  p1 <- ggplot(data=cohort, aes(x=Age, y=Survival)) +
          geom_line() +
          labs(title = 'Survival Curve')

  ggsave(filename = 'cohort_survival_curve.png', 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  p2 <- ggplot(data=cdMelt, aes(x=Age, y=Prevalence)) +
          geom_line() +
          facet_grid(. ~ variable) +
          theme(panel.spacing.x = unit(1.5, 'lines')) +
          labs(title = 'Chronic Disease Prevalence in Cohort')

  ggsave(filename = 'cohort_base_prev.png', 
         plot = p2,
         path = path,
         width = 10,
         height = 5)

  p1
  p2

}

cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')

visCohortBase(cohort)

```

## Intervention CD prev and survival curve comparison

```{r}

cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
noSmoke <- read_dta('ELSA_CnoSmoke/ELSA_CnoSmoke_summary.dta')

visSmokIntervention <- function(cohort, intervention, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_2/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)

  survivals <- survivals %>%
                 rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, 'No Smoking' = intervention.Survival) %>%
                 melt(id.var = 'Age', value.name = 'Survival')

  p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
        geom_line() +
        labs(title = 'Baseline vs Non-smoking', x = 'Age', y = 'Survival')

  ggsave(filename = 'cohort_and_intervention_survival_curve.png', 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  # Now do the chronic disease comparison graphs
  # Collect data for each disease separately
  cancre <- data.frame(cohort$Age, cohort$p_cancre_all, noSmoke$p_cancre_all)
  diabe <- data.frame(cohort$Age, cohort$p_diabe_all, noSmoke$p_diabe_all)
  hearte <- data.frame(cohort$Age, cohort$p_hearte_all, noSmoke$p_hearte_all)
  lunge <- data.frame(cohort$Age, cohort$p_lunge_all, noSmoke$p_lunge_all)

  cancre <- cancre %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_cancre_all, 'No Smoking' = noSmoke.p_cancre_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  diabe <- diabe %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_diabe_all, 'No Smoking' = noSmoke.p_diabe_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  hearte <- hearte %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_hearte_all, 'No Smoking' = noSmoke.p_hearte_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  lunge <- lunge %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_lunge_all, 'No Smoking' = noSmoke.p_lunge_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  p2 <- ggplot(data = cancre, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Cancer', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'cancre_cohort_intervention.png', 
         plot = p2,
         path = path,
         width = 7,
         height = 5)

  p3 <- ggplot(data = diabe, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Diabetes', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'diabe_cohort_intervention.png', 
         plot = p3,
         path = path,
         width = 7,
         height = 5)

  p4 <- ggplot(data = hearte, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Heart Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'hearte_cohort_intervention.png', 
         plot = p4,
         path = path,
         width = 7,
         height = 5)

  p5 <- ggplot(data = lunge, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Lung Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'lunge_cohort_intervention.png', 
         plot = p5,
         path = path,
         width = 7,
         height = 5)

  p1
  p2
  p3
  p4
  p5

}

visSmokIntervention(cohort, noSmoke)

```

# OLD - DEFUNKT CHUNKS

# Chronic Disease Prevalence (OLD SCENARIO I - FULL POPULATION)

## Baseline Initial Prevalence

```{r}
### visInitBasePrev
# Inputs:   base
#           path
#
# Output:   bar - 
#
# This function... 
visInitBasePrev <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                filter(year == 2012) %>%
                melt(id.var = 'year', value.name = 'Prevalence') %>%
                subset(select = -year)

  bar <- ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
          geom_bar(stat = 'identity') +
          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'init_baseline_prevalence.png', 
         plot = bar,
         path = path,
         width = 7,
         height = 5)

  bar
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visInitBasePrev(baseline)

```

## Baseline Projections

```{r}
### visProjectedBaselineCD
# Inputs:   base - summary output file from the ELSA_Baseline scenario
#           path - path to the directory where the resulting plot will be saved
#
# Output:   tiled - A 3 column tiled plot with the projected prevalence of Cancer, Diabetes, and Heart Disease on the same x axis, with year on the y
#
# This function visualises the prevalence of Cancer, Diabetes, and Heart Disease in the Baseline projection of the FEM
visProjectedBaselineCD <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                melt(id.var='year', value.name = 'Prevalence')

  # Create the tiled plot of disease prevalence
  tiled <- ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
            geom_line() +
            facet_grid(. ~ variable) +
            theme(panel.spacing.x = unit(1.5, 'lines')) +
            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'baseline_prevalence.png', 
         plot = tiled,
         path = path,
         width = 10,
         height = 5)

  # Show plot
  tiled
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBaselineCD(baseline)

```

```{r}
### visProjectedBySubpop
# Inputs:   base - 
#           path - 
#           cd - chronic disease to visualise
#
# Output:   None
#
# This function visualises the baseline prevalence of a chronic disease broken down by subpopulations. Currently, the subpopulations included are: 
# - obese/not obese 
# - disabled (anyadl==1)/not disabled 
# - age splines (55-64,65-74,75-84,85p)
visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  }

  # Generate varnames
  all <- paste0('p_', cd, '_all')
  # obesity
  obese <- paste0('p_', cd, '_obese')
  notObese <- paste0('p_', cd, '_notObese')
  # disability
  adl <- paste0('p_', cd, '_anyadl')
  noadl <- paste0('p_', cd, '_noadl')
  # age
  a5564 <- paste0('p_', cd, '_5564')
  a6574 <- paste0('p_', cd, '_6574')
  a7584 <- paste0('p_', cd, '_7584')
  a85p <- paste0('p_', cd, '_85p')

  obeseMelt <- base %>%
              select(c('year', all_of(all), all_of(obese), all_of(notObese))) %>%
              rename('All' = all, 'Obese' = obese, 'Healthy Weight' = notObese) %>%
              melt(id.var='year', value.name = 'Prevalence')

  adlMelt <- base %>%
              select(c('year', all_of(all), adl, noadl)) %>%
              rename('All' = all, 'Disabled' = adl, 'Able' = noadl) %>%
              melt(id.var='year', value.name = 'Prevalence')

  ageMelt <- base %>%
              select(c('year', a5564, a6574, a7584, a85p)) %>%
              rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
              melt(id.var = 'year', value.name = 'Prevalence')

  p1 <- ggplot(data=obeseMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f1 <- paste0(cd, '_prev_by_obesity.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f1, 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  p2 <- ggplot(data=adlMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f2 <- paste0(cd, '_prev_by_disability.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f2, 
         plot = p2,
         path = path,
         width = 6,
         height = 5)

  p3 <- ggplot(data=ageMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f3 <- paste0(cd, '_prev_by_age.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f3, 
         plot = p3,
         path = path,
         width = 6,
         height = 5)
  p1
  p2
  p3

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBySubpop(baseline, cd = 'cancre')
visProjectedBySubpop(baseline, cd = 'diabe')
visProjectedBySubpop(baseline, cd = 'hearte')

```

## Disability Comparison 2012 - 2040

This one is slightly more tricky to think about so need a plan.

I want to compare the prevalence of anyadl in 2012 and 2050 (or 2040, or 2060). So first I need to extract the prevalence of anyadl from all scenarios I want to compare.

```{r}

visDisabilityComparison <- function(base, ten, fifty, hundred, targetYear, cd,  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC10 <- ten %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('10%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC50 <- fifty %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('50%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC100 <- hundred %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('100%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selected <- inner_join(selBase, selC10)
  selected <- inner_join(selected, selC50)
  selected <- inner_join(selected, selC100)
  selected <- select(selected, -year)

  newline <- sapply(selected, function(x) x[2] - x[1])

  selected <- rbind(selected, newline)
  new <- data.frame(t(selected))
  colnames(new) <- c('y2012', 'y2050', 'Difference')

  new$change <- (new$Difference / new$y2012) * 100
  #new <- rename(new, '%Increase' = change)

  new <- cbind(Scenario = rownames(new), new)
  rownames(new) <- 1:nrow(new)
  new$Scenario <- factor(new$Scenario, levels = c('Baseline', '10%', '50%', '100%'))

  x_name <- paste0(cd, ' Reduction Scenario')

  p <-ggplot(data = new, aes(x = Scenario, y = change)) +
        geom_bar(stat = 'identity') +
        labs(title = 'Increase in Prevalence of Disability (2012 - 2050)', x = x_name, y = '% Increase')

  fname <- paste0(cd, '_reduc_disability_prev.png')

  ggsave(filename = fname, 
         plot = p,
         path = path,
         width = 6,
         height = 5)

  p

}

base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

cancre10 <- read_dta('ELSA_pcancre10/ELSA_pcancre10_summary.dta')
cancre50 <- read_dta('ELSA_pcancre50/ELSA_pcancre50_summary.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_summary.dta')

diabe10 <- read_dta('ELSA_pdiabe10/ELSA_pdiabe10_summary.dta')
diabe50 <- read_dta('ELSA_pdiabe50/ELSA_pdiabe50_summary.dta')
diabe100 <- read_dta('ELSA_pdiabe100/ELSA_pdiabe100_summary.dta')

hearte10 <- read_dta('ELSA_phearte10/ELSA_phearte10_summary.dta')
hearte50 <- read_dta('ELSA_phearte50/ELSA_phearte50_summary.dta')
hearte100 <- read_dta('ELSA_phearte100/ELSA_phearte100_summary.dta')

visDisabilityComparison(base, cancre10, cancre50, cancre100, 2050, 'Cancer')
visDisabilityComparison(base, diabe10, diabe50, diabe100, 2050, 'Diabetes')
visDisabilityComparison(base, hearte10, hearte50, hearte100, 2050, 'Heart Disease')

```

### 2020-11-16 FEM_R/IJM_paper_vis_NEAT.Rmd
Quick

@@
   # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl, anyiadl)) #, adl1, adl2, adl3p))
-  cd_CV2 <- select(CV2, c(year, p_cancre_all, p_diabe_all, 
+  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl, anyiadl)) #, adl1, adl2, adl3p))
+  cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
                             p_hearte_all, p_lunge_all, p_smoken_all, 
@@
 
-  CV2_rename <- cd_CV2 %>% dplyr::rename('cancre' = p_cancre_all, 
+  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_all,
+                                        'cancre' = p_cancre_all, 
                                         'diabe' = p_diabe_all, 

### 2020-11-23 FEM_R/model_analysis.Rmd
Created the model analysis R Notebook. So far read in all T-test files and in process of formatting tables

---
title: "R Notebook"
output: html_notebook
---

This notebook is for debugging purposes. The current plan is to write a Make rule for running this notebook automatically at the end of a model run.

The purpose of this notebook is:
- Present the T-tests altogether in one document.
- Present all transition models in same document.
- Add visualisation for cross-validation 2 output

Do all this for both minimal models and cross-validation.

```{r setup, include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(formattable)
#require(haven)
#require(reshape2)
#require(ggplot2)
#require(gridExtra)

```

```{r include=FALSE}

# Specify a function to rename the dataframes, so we can run in an sapply()
rename_first_var <- function(testCSV, prefix) {
  testCSV[,1] <- paste0(testCSV[,1], '_', prefix)
  return(testCSV)
}

```

First, we will collect and present all the T-tests together. We will add the prefix 'min_' or 'cv_' onto the variable names so we don't get mixed up.

```{r}
# names for T-tests directory and filename stem
min <- "ELSA_minimal/T-tests/"
stem <- "minimal_all_waves_"

# Read in all the T-test csv files
min_uw <- read.csv(paste0(min, stem, "unweighted.csv"))
min_demog <- read.csv(paste0(min, stem, "demog.csv"))
min_binhlth <- read.csv(paste0(min, stem, "binhlth.csv"))
min_binecon <- read.csv(paste0(min, stem, "binecon.csv"))
min_risk <- read.csv(paste0(min, stem, "risk.csv"))

# Collect in a list to run a list apply
min_list <- list(min_uw, min_demog, min_binhlth, min_binecon, min_risk)
# Add prefix 'min_' onto every variable name for clarity
min_list <- lapply(min_list, rename_first_var, prefix='min')

```

```{r}
# names for T-tests directory and filename stem
CV1 <- "ELSA_CrossValidation1/T-tests/"
stem <- "CV1_all_waves_"
# Read in files
cv_uw <- read.csv(paste0(CV1, stem, "unweighted.csv"))
cv_demog <- read.csv(paste0(CV1, stem, "demog.csv"))
cv_binhlth <- read.csv(paste0(CV1, stem, "binhlth.csv"))
cv_binecon <- read.csv(paste0(CV1, stem, "binecon.csv"))
cv_risk <- read.csv(paste0(CV1, stem, "risk.csv"))
# Now collect into list and rename vars for clarity
cv_list <- list(cv_uw, cv_demog, cv_binhlth, cv_binecon, cv_risk)
cv_list <- lapply(cv_list, rename_first_var, prefix='CV')
```

Both Minimal and Cross-validation T-tests are now renamed and collected into a separate list object each. We will now combine like tests together (i.e. demog with demog, binhlth with binhlth).

```{r include=FALSE}
# Set up vector with all p-value column names. Require these for table formatting
pValNames <- c('p_value_wave3', 'p_value_wave4', 'p_value_wave5', 'p_value_wave6', 'p_value_wave7', 'p_value_wave8')
```

### Unweighted (died)
```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
uw <- rbind(min_list[[1]], cv_list[[1]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
uw <- arrange(uw, variable)
#uw
formattable(uw, 
            align = c('l', rep('r', ncol(uw) - 1)),
            list(area(col=pValNames) 
                  ~ proportion_bar(color='green', 0.01)))
```
### Demographics
```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
demog <- rbind(min_list[[2]], cv_list[[2]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
demog <- arrange(demog, variable)
formattable(demog, 
            align = c('l', rep('r', ncol(demog) - 1)),
            list(area(col=pValNames) 
                  ~ proportion_bar(color='green', 0.01)))
```
### Binary Health
```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
binhlth <- rbind(min_list[[3]], cv_list[[3]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
binhlth <- arrange(binhlth, variable)
formattable(binhlth, 
            align = c('l', rep('r', ncol(binhlth) - 1)),
            list(area(col=pValNames) 
                  ~ proportion_bar(color='green', 0.01)))
```
### Binary Economic
```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
binecon <- rbind(min_list[[4]], cv_list[[4]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
binecon <- arrange(binecon, variable)
formattable(binecon, 
            align = c('l', rep('r', ncol(binecon) - 1)),
            list(area(col=pValNames) 
                  ~ proportion_bar(color='green', 0.01)))
```
### Risk Behaviours
```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
risk <- rbind(min_list[[5]], cv_list[[5]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
risk <- arrange(risk, variable)
formattable(risk, 
            align = c('l', rep('r', ncol(risk) - 1)),
            list(area(col=pValNames) 
                  ~ proportion_bar(color='green', 0.01)))
```

```{r echo = FALSE, results = 'asis'}
# Row bind to collect both dfs together
risk <- rbind(min_list[[5]], cv_list[[5]])
# Reorder rows in alphabetical order. This ensures that like variables are clumped together
risk <- arrange(risk, variable)
formattable(risk, 
            align = c('l', rep('r', ncol(risk) - 1)),
            list(area(col=pValNames) ~ color_tile('transparent', 'red')))
```

### 2020-11-24 FEM_R/model_analysis.Rmd
Generated a table with Reactable package including sticky columns and some row styling. This will be the template for all T-test tables soon, just need to add colour gradients and maybe some box plots

@@
 require(formattable)
+require(reactable)
 #require(haven)
@@
 
-# Specify a function to rename the dataframes, so we can run in an sapply()
+# Need this code block for reactable tables to be rendered properly in HTML
+# Fix from here: https://www.gitmemory.com/issue/glin/reactable/43/618802061
+htmltools::tagList(
+  reactR::html_dependency_react(),
+  reactR::html_dependency_reacttools(),
+  reactable::reactable(iris)
+)
+
+# function to rename the dataframes, so we can run in an sapply()
 rename_first_var <- function(testCSV, prefix) {
@@
   return(testCSV)
+}
+
+# function for shortening column names
+shorten_columns <- function(testCSV) {
+  testCSV <- testCSV %>%
+                rename_at(vars(starts_with('fem_mean_wave')), funs(sub('fem_mean_wave', 'FEM_', .))) %>%
+                rename_at(vars(starts_with('elsa_mean_wave')), funs(sub('elsa_mean_wave', 'ELSA_', .))) %>%
+                rename_at(vars(starts_with('p_value_wave')), funs(sub('p_value_wave', 'P_Value_', .))) %>%
+                rename('Variable' = variable)
 }
@@
 
-# Collect in a list to run a list apply
+# Collect in a list to do some formatting
 min_list <- list(min_uw, min_demog, min_binhlth, min_binecon, min_risk)
@@
 min_list <- lapply(min_list, rename_first_var, prefix='min')
+# Rename columns to be more succinct
+min_list <- lapply(min_list, shorten_columns)
 
@@
 cv_risk <- read.csv(paste0(CV1, stem, "risk.csv"))
-# Now collect into list and rename vars for clarity
+# Now collect into list and rename vars and cols for clarity
 cv_list <- list(cv_uw, cv_demog, cv_binhlth, cv_binecon, cv_risk)
 cv_list <- lapply(cv_list, rename_first_var, prefix='CV')
+cv_list <- lapply(cv_list, shorten_columns)
 ```
@@
 # Set up vector with all p-value column names. Require these for table formatting
-pValNames <- c('p_value_wave3', 'p_value_wave4', 'p_value_wave5', 'p_value_wave6', 'p_value_wave7', 'p_value_wave8')
+pValNames <- c('P_Value_3', 'P_Value_4', 'P_Value_5', 'P_Value_6', 'P_Value_7', 'P_Value_8')
 ```
@@
 # Reorder rows in alphabetical order. This ensures that like variables are clumped together
-uw <- arrange(uw, variable)
+uw <- arrange(uw, Variable)
 #uw
@@
 # Reorder rows in alphabetical order. This ensures that like variables are clumped together
-demog <- arrange(demog, variable)
+demog <- arrange(demog, Variable)
 formattable(demog, 
@@
 # Reorder rows in alphabetical order. This ensures that like variables are clumped together
-binhlth <- arrange(binhlth, variable)
+binhlth <- arrange(binhlth, Variable)
 formattable(binhlth, 
@@
 # Reorder rows in alphabetical order. This ensures that like variables are clumped together
-binecon <- arrange(binecon, variable)
+binecon <- arrange(binecon, Variable)
 formattable(binecon, 
@@
 # Reorder rows in alphabetical order. This ensures that like variables are clumped together
-risk <- arrange(risk, variable)
+risk <- arrange(risk, Variable)
 formattable(risk, 
@@
 ```
+
+# Reactable Package
+
+```{r}
+
+stickyStyle <- list(position = "sticky", left = 0, background = "#fff", zIndex = 1,
+                    borderRight = "1px solid #eee")
+
+reactable(risk,
+          defaultColDef = colDef(
+            header = function(value) gsub('_', ' ', value, fixed=TRUE),
+            cell = function(value) format(value, nsmall = 2),
+            align='center',
+            headerStyle = list(background = "#f7f7f8")
+          ),
+          columns = list(
+            Variable = colDef(minWidth = 160,
+                              style = stickyStyle,
+                              headerStyle = stickyStyle)
+          ),
+          bordered = TRUE,
+          highlight = TRUE,
+          striped = TRUE)
+
+```

### 2020-11-24 FEM_R/model_analysis.Rmd
Implemented the reactable tables for T-tests and happy with the output

@@
 ```{r setup, include=FALSE}
-# Need to set up the environment
+# Need to set up the working directory
 workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
@@
 
-First, we will collect and present all the T-tests together. We will add the prefix 'min_' or 'cv_' onto the variable names so we don't get mixed up.
+<!-- First, we will collect and present all the T-tests together. We will add the prefix 'min_' or 'cv_' onto the variable names so we don't get mixed up. -->
 
-```{r}
+```{r include=FALSE}
 # names for T-tests directory and filename stem
@@
 min_list <- lapply(min_list, shorten_columns)
+# Replace na with 0
+min_list <- lapply(min_list, function(x) { x[is.na(x)] <- 0; x} )
 
@@
 
-```{r}
+```{r include=FALSE}
 # names for T-tests directory and filename stem
@@
 cv_list <- lapply(cv_list, shorten_columns)
+cv_list <- lapply(cv_list, function(x) { x[is.na(x)] <- 0; x} )
 ```
 
-Both Minimal and Cross-validation T-tests are now renamed and collected into a separate list object each. We will now combine like tests together (i.e. demog with demog, binhlth with binhlth).
+<!-- Both Minimal and Cross-validation T-tests are now renamed and collected into a separate list object each.
+We will now combine like tests together (i.e.
+demog with demog, binhlth with binhlth).
+-->
 
@@
 
+<!-- Lets build the combined tables now. -->
+
+```{r include=FALSE}
+# Row bind to collect both dfs together
+uw <- rbind(min_list[[1]], cv_list[[1]])
+# Reorder rows in alphabetical order. This ensures that like variables are clumped together
+uw <- arrange(uw, Variable)
+
+# demog
+demog <- rbind(min_list[[2]], cv_list[[2]])
+demog <- arrange(demog, Variable)
+
+# binhlth
+binhlth <- rbind(min_list[[3]], cv_list[[3]])
+binhlth <- arrange(binhlth, Variable)
+
+# binecon
+binecon <- rbind(min_list[[4]], cv_list[[4]])
+binecon <- arrange(binecon, Variable)
+
+# risk
+risk <- rbind(min_list[[5]], cv_list[[5]])
+risk <- arrange(risk, Variable)
+```
+
+# Reactable Tables
+
+The [reactable](https://glin.github.io/reactable/index.html) package has the ability to include lots of complex styling, such as additional colour, formatting and even inline graphs built dynamically from the input data.
+
+For the purpose of this document, we are going to include a sticky first column (so we can always see the variable names whilst scrolling), a colour gradient from white to green based on the P Value (low = white, high = green), and a red cell background for any P values of 0.
+
+```{r echo = FALSE}
+## This block will define the logic for creating the tables with Reactable
+# The tables are formatted exaclty the same for each group of vars so makes sense to define this only once
+
+# createTTT - create T-Test Table
+# Have a default column definition, to handle the basic formatting, then specific formatting for the first column (Variable) and P Val cols
+createTTT <- function(t) {
+  # To enable a 'sticky' first column, that always shows as you scroll across the table
+  stickyStyle <- list(position = "sticky", left = 0, background = "#fff", zIndex = 1,
+                      borderRight = "1px solid #eee")
+
+  # Define a colour ramp to use for P Value cell gradients
+  greenPal <- function(x) rgb(colorRamp(c("#ffffff", "#008000"))(x), maxColorValue = 255)
+
+  # Add in the green colour gradient for P Value cells. Also add a red block for value of 0
+  pValColDef <- colDef(
+    style = function(value) {
+      color <- greenPal(value)
+      if (value == 0) {
+        color <- "#ff4c4c"
+      }
+      list(background = color)
+    },
+    minWidth = 110
+  )
+
+  res <- reactable(t,
+            defaultColDef = colDef(
+              header = function(value) gsub('_', ' ', value, fixed=TRUE),
+              cell = function(value) format(value, nsmall = 2),
+              align='center',
+              style = function(value) {
+
+              }
+              #headerStyle = list(background = "#f7f7f8")
+            ),
+            columns = list(
+              Variable = colDef(
+                  minWidth = 160,
+                  style = stickyStyle,
+                  headerStyle = stickyStyle),
+              P_Value_3 = pValColDef,
+              P_Value_4 = pValColDef,
+              P_Value_5 = pValColDef,
+              P_Value_6 = pValColDef,
+              P_Value_7 = pValColDef,
+              P_Value_8 = pValColDef
+            ),
+            bordered = TRUE,
+            highlight = TRUE,
+            striped = TRUE)
+
+  return(res)
+}
+
+```
+
+### Unweighted (Died)
+
+```{r echo = FALSE}
+# Create the actual table!
+uwTab <- createTTT(uw)
+uwTab
+```
+
+### Demographics
+
+```{r echo = FALSE}
+# Create the actual table!
+demogTab <- createTTT(demog)
+demogTab
+```
+
+### Binary Health
+```{r echo = FALSE}
+# Create the actual table!
+binhlthTab <- createTTT(binhlth)
+binhlthTab
+```
+### Binary Economic
+```{r echo = FALSE}
+# Create the actual table!
+bineconTab <- createTTT(binecon)
+bineconTab
+```
+### Risk Behaviours
+```{r echo = FALSE}
+# Create the actual table!
+riskTab <- createTTT(risk)
+riskTab
+```
+
+# DEFUNKT
 ### Unweighted (died)
@@
 ```
-
-# Reactable Package
-
-```{r}
-
-stickyStyle <- list(position = "sticky", left = 0, background = "#fff", zIndex = 1,
-                    borderRight = "1px solid #eee")
-
-reactable(risk,
-          defaultColDef = colDef(
-            header = function(value) gsub('_', ' ', value, fixed=TRUE),
-            cell = function(value) format(value, nsmall = 2),
-            align='center',
-            headerStyle = list(background = "#f7f7f8")
-          ),
-          columns = list(
-            Variable = colDef(minWidth = 160,
-                              style = stickyStyle,
-                              headerStyle = stickyStyle)
-          ),
-          bordered = TRUE,
-          highlight = TRUE,
-          striped = TRUE)
-
-```

### 2020-11-24 FEM_R/model_analysis.Rmd
Removed old formattable tables from model analysis notebook

@@
 ```
-
-# DEFUNKT
-### Unweighted (died)
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-uw <- rbind(min_list[[1]], cv_list[[1]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-uw <- arrange(uw, Variable)
-#uw
-formattable(uw, 
-            align = c('l', rep('r', ncol(uw) - 1)),
-            list(area(col=pValNames) 
-                  ~ proportion_bar(color='green', 0.01)))
-```
-### Demographics
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-demog <- rbind(min_list[[2]], cv_list[[2]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-demog <- arrange(demog, Variable)
-formattable(demog, 
-            align = c('l', rep('r', ncol(demog) - 1)),
-            list(area(col=pValNames) 
-                  ~ proportion_bar(color='green', 0.01)))
-```
-### Binary Health
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-binhlth <- rbind(min_list[[3]], cv_list[[3]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-binhlth <- arrange(binhlth, Variable)
-formattable(binhlth, 
-            align = c('l', rep('r', ncol(binhlth) - 1)),
-            list(area(col=pValNames) 
-                  ~ proportion_bar(color='green', 0.01)))
-```
-### Binary Economic
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-binecon <- rbind(min_list[[4]], cv_list[[4]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-binecon <- arrange(binecon, Variable)
-formattable(binecon, 
-            align = c('l', rep('r', ncol(binecon) - 1)),
-            list(area(col=pValNames) 
-                  ~ proportion_bar(color='green', 0.01)))
-```
-### Risk Behaviours
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-risk <- rbind(min_list[[5]], cv_list[[5]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-risk <- arrange(risk, Variable)
-formattable(risk, 
-            align = c('l', rep('r', ncol(risk) - 1)),
-            list(area(col=pValNames) 
-                  ~ proportion_bar(color='green', 0.01)))
-```
-
-```{r echo = FALSE, results = 'asis'}
-# Row bind to collect both dfs together
-risk <- rbind(min_list[[5]], cv_list[[5]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-risk <- arrange(risk, Variable)
-formattable(risk, 
-            align = c('l', rep('r', ncol(risk) - 1)),
-            list(area(col=pValNames) ~ color_tile('transparent', 'red')))
-```

### 2020-11-24 FEM_R/model_analysis.Rmd
Added visualisation of handover plots for cross-validation 2

@@
 require(reactable)
-#require(haven)
+require(haven)
 #require(reshape2)
@@
 ```
+
+# Cross Validation 2 - 'Handover' plots
+
+```{r echo = FALSE}
+vis_CV2_prev <- function(long, CV2) {
+
+  # Keep only waves 1 - 4
+  longC <- subset(long, wave < 5)
+  longC <- as.data.table(longC)
+
+  # Replace all NA with 0
+  longC[is.na(longC)] <- 0
+
+  # Have to drop r*clust vars as they contain strings
+  longC <- select(longC, -contains('clust'))
+
+  # Now get the weighted mean by wave of every column based on cwtresp
+  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]
+
+  # ELSA_long is now processed to the point where we can start extracting and combining variables
+  # Better start by adding year to l2 and reordering the columns
+  l2$year <- seq(from=2002, to = 2008, by = 2)
+  setcolorder(l2, c('year', 'wave'))
+
+  # Now lets extract chronic disease variables from both dataframes and do a row combine
+  # Start with cancre, diabe, hearte
+  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl)) #, anyiadl, adl1, adl2, adl3p))
+  cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
+                            p_hearte_all, p_lunge_all, p_smoken_all, 
+                            a_bmi_all, p_work_all,  p_anyadl_all))
+                            #, p_anyiadl_all p_adl1_all, p_adl2_all, p_adl3p_all))
+
+  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_all,
+                                        'cancre' = p_cancre_all, 
+                                        'diabe' = p_diabe_all, 
+                                        'hearte' = p_hearte_all,
+                                        'lunge' = p_lunge_all,
+                                        'smoken' = p_smoken_all,
+                                        'bmi' = a_bmi_all,
+                                        'work' = p_work_all,
+                                        'anyadl' = p_anyadl_all)
+                                        #'anyiadl' = p_anyiadl_all)
+                                        #'adl1' = p_adl1_all,
+                                        #'adl2' = p_adl2_all,
+                                        #'adl3p' = p_adl3p_all)
+
+  combine <- rbind(cd_l2, CV2_rename)
+
+  melted <- melt(combine, id.var = 'year', value.name = 'Prevalence')
+
+  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
+          geom_line() +
+          facet_wrap(. ~ variable, scales = "free")
+          #theme(panel.spacing.x = unit(1.5, 'lines')) +
+          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
+}
+
+# Load in whole stock pop, plus the CV2 summary output file
+long <- read_dta('../input_data/ELSA_long.dta')
+CV2 <- read_dta('ELSA_CrossValidation2/ELSA_CrossValidation2_summary.dta')
+
+vis_CV2_prev(long, CV2)
+```

### 2020-11-25 FEM_R/model_analysis.Rmd
Added call to require(data.table) so the CV2 visualisations would work

@@
 require(haven)
+require(data.table)
 #require(reshape2)
@@
 
+  pValColDef2 <- colDef(
+    style = function(value) {
+      if (value >= 0.05) {
+        color <- "#4ca64c"
+      }
+      if (value < 0.05) {
+        color <- "#ff6666"
+      }
+      list(background = color)
+    },
+    minWidth = 110
+  )
+
   res <- reactable(t,
@@
               cell = function(value) format(value, nsmall = 2),
-              align='center',
-              style = function(value) {
-
-              }
-              #headerStyle = list(background = "#f7f7f8")
+              align = 'center'
             ),
@@
                   headerStyle = stickyStyle),
-              P_Value_3 = pValColDef,
-              P_Value_4 = pValColDef,
-              P_Value_5 = pValColDef,
-              P_Value_6 = pValColDef,
-              P_Value_7 = pValColDef,
-              P_Value_8 = pValColDef
+              P_Value_3 = pValColDef2,
+              P_Value_4 = pValColDef2,
+              P_Value_5 = pValColDef2,
+              P_Value_6 = pValColDef2,
+              P_Value_7 = pValColDef2,
+              P_Value_8 = pValColDef2
             ),

### 2020-11-26 FEM_R/model_analysis.Rmd
Changed name of analysis document
Big refactor to rename ugly and inconsistent "cross-validation/crossvalidation/cross_validation" file names for crossval. Now replaced with "CV" everywhere

@@
 ---
-title: "R Notebook"
+title: "FEM Model Analysis"
 output: html_notebook
@@
 # names for T-tests directory and filename stem
-CV1 <- "ELSA_CrossValidation1/T-tests/"
+CV1 <- "ELSA_CV1/T-tests/"
 stem <- "CV1_all_waves_"
@@
 long <- read_dta('../input_data/ELSA_long.dta')
-CV2 <- read_dta('ELSA_CrossValidation2/ELSA_CrossValidation2_summary.dta')
+CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
 

### 2020-11-27 FEM_R/model_analysis.Rmd
More changes to transition models trying to improve on the T-tests

@@
 require(tidyverse)
-require(formattable)
+#require(formattable)
 require(reactable)
@@
 ```
+
+# Github Reporting
+
+It would be nice if we could copy paste tables directly from this document into a github issue.
+For this reason, we will look into a package that can print tables directly into markdown or something similar, which we can then place straight into a Github issue.
+
+```{r}
+
+#kable
+
+```

### 2020-12-09 FEM_R/model_analysis.Rmd
Tweaks to smkint model, still WIP

@@
                   headerStyle = stickyStyle),
+              P_Value_1 = pValColDef2,
+              P_Value_2 = pValColDef2,
               P_Value_3 = pValColDef2,
@@
 
-# Cross Validation 2 - 'Handover' plots
+# Cross Validation 2 - 'Handover' plots (Still WIP need fixing)
 

### 2020-12-11 FEM_R/model_analysis.Rmd

@@
   # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl)) #, anyiadl, adl1, adl2, adl3p))
+  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl)) #, anyiadl, adl1, adl2, adl3p))
   cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
                             p_hearte_all, p_lunge_all, p_smoken_all, 
-                            a_bmi_all, p_work_all,  p_anyadl_all))
+                            a_bmi_all, p_employed_all,  p_anyadl_all))
                             #, p_anyiadl_all p_adl1_all, p_adl2_all, p_adl3p_all))
@@
                                         'bmi' = a_bmi_all,
-                                        'work' = p_work_all,
+                                        'employed' = p_employed_all,
                                         'anyadl' = p_anyadl_all)

### 2020-12-15 FEM_R/model_analysis.Rmd

@@
 min_binecon <- read.csv(paste0(min, stem, "binecon.csv"))
+min_cntecon <- read.csv(paste0(min, stem, "cntecon.csv"))
 min_risk <- read.csv(paste0(min, stem, "risk.csv"))
@@
 # Collect in a list to do some formatting
-min_list <- list(min_uw, min_demog, min_binhlth, min_binecon, min_risk)
+min_list <- list(min_uw, min_demog, min_binhlth, min_binecon, min_cntecon, min_risk)
 # Add prefix 'min_' onto every variable name for clarity
@@
 cv_binecon <- read.csv(paste0(CV1, stem, "binecon.csv"))
+cv_cntecon <- read.csv(paste0(CV1, stem, "cntecon.csv"))
 cv_risk <- read.csv(paste0(CV1, stem, "risk.csv"))
 # Now collect into list and rename vars and cols for clarity
-cv_list <- list(cv_uw, cv_demog, cv_binhlth, cv_binecon, cv_risk)
+cv_list <- list(cv_uw, cv_demog, cv_binhlth, cv_binecon, cv_cntecon, cv_risk)
 cv_list <- lapply(cv_list, rename_first_var, prefix='CV')
@@
 
+# cntecon
+cntecon <- rbind(min_list[[5]], cv_list[[5]])
+cntecon <- arrange(cntecon, Variable)
+
 # risk
-risk <- rbind(min_list[[5]], cv_list[[5]])
+risk <- rbind(min_list[[6]], cv_list[[6]])
 risk <- arrange(risk, Variable)
@@
 ### Unweighted (Died)
-
 ```{r echo = FALSE}
@@
 ```
-
 ### Demographics
-
 ```{r echo = FALSE}
@@
 ```
-
 ### Binary Health
@@
 bineconTab
+```
+### Count Economic
+```{r echo = FALSE}
+# Create the actual table!
+cnteconTab <- createTTT(cntecon)
+cnteconTab
 ```

### 2021-01-04 FEM_R/IJM_paper_vis_NEAT.Rmd
Fixed some visualisations written for previous intervention scenarios, and added more visualisations into debug document

--- 2020-11-16 FEM_R/IJM_paper_vis_NEAT.Rmd
@@
   # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl, anyiadl)) #, adl1, adl2, adl3p))
-  cd_CV2 <- select(CV2, c(year, p_cancre_all, p_diabe_all, 
+  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl, anyiadl)) #, adl1, adl2, adl3p))
+  cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
                             p_hearte_all, p_lunge_all, p_smoken_all, 
-                            a_bmi_all, p_work_all,  p_anyadl_all, p_anyiadl_all))
+                            a_bmi_all, p_employed_all,  p_anyadl_all, p_anyiadl_all))
                             #p_adl1_all, p_adl2_all, p_adl3p_all))
 
-  CV2_rename <- cd_CV2 %>% dplyr::rename('cancre' = p_cancre_all, 
+  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_all,
+                                        'cancre' = p_cancre_all, 
                                         'diabe' = p_diabe_all, 
@@
                                         'bmi' = a_bmi_all,
-                                        'work' = p_work_all,
+                                        'employed' = p_employed_all,
                                         'anyadl' = p_anyadl_all,
@@
                 select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
-                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
+                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                 filter(year == 2012) %>%
@@
                 select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
-                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
+                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                 melt(id.var='year', value.name = 'Prevalence')

### 2021-01-04 FEM_R/model_analysis.Rmd
Fixed some visualisations written for previous intervention scenarios, and added more visualisations into debug document

--- 2020-12-11 FEM_R/model_analysis.Rmd
@@
 title: "FEM Model Analysis"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 require(tidyverse)
-#require(formattable)
 require(reactable)
@@
 require(data.table)
-#require(reshape2)
-#require(ggplot2)
-#require(gridExtra)
+require(reshape2)
+require(ggplot2)
+require(plyr)
 
@@
 
-# function to rename the dataframes, so we can run in an sapply()
+# function to rename the dataframes, so we can run in an lapply()
 rename_first_var <- function(testCSV, prefix) {
@@
                 rename_at(vars(starts_with('p_value_wave')), funs(sub('p_value_wave', 'P_Value_', .))) %>%
-                rename('Variable' = variable)
+                dplyr::rename('Variable' = variable)
 }
@@
 
+# Intervention Scenarios
+
+### Base Initial Prevalence of Chronic Disease
+
+```{r echo = FALSE}
+visInitBasePrev <- function(base) {
+
+  cdMelted <- base %>%
+                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
+                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
+                filter(year == 2012) %>%
+                melt(id.var = 'year', value.name = 'Prevalence') %>%
+                subset(select = -year)
+
+  ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
+          geom_bar(stat = 'identity') +
+          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')
+}
+
+baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+visInitBasePrev(baseline)
+```
+
+### Baseline Projections
+
+```{r echo = FALSE}
+visProjectedBaselineCD <- function(base) {
+
+  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
+  cdMelted <- base %>%
+                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
+                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
+                melt(id.var='year', value.name = 'Prevalence')
+
+  # Create the tiled plot of disease prevalence
+  ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
+            geom_line() +
+            facet_grid(. ~ variable) +
+            theme(panel.spacing.x = unit(1.5, 'lines')) +
+            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
+}
+
+baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+visProjectedBaselineCD(baseline)
+```
+
+### Baseline Projections by Subpopulation
+
+```{r echo = FALSE}
+visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {
+
+  if(cd == 'cancre') {
+    dis <- 'Cancer'
+  } else if(cd == 'diabe') {
+    dis <- 'Diabetes'
+  } else if(cd == 'hearte') {
+    dis <- 'Heart Disease'
+  }
+
+  # Generate varnames
+  all <- paste0('p_', cd, '_all')
+  # age
+  a5564 <- paste0('p_', cd, '_5564')
+  a6574 <- paste0('p_', cd, '_6574')
+  a7584 <- paste0('p_', cd, '_7584')
+  a85p <- paste0('p_', cd, '_85p')
+
+  ageMelt <- base %>%
+              select(c('year', a5564, a6574, a7584, a85p)) %>%
+              dplyr::rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
+              melt(id.var = 'year', value.name = 'Prevalence')
+
+  ggplot(data=ageMelt, aes(x=year, y=Prevalence, color = variable)) +
+          geom_line() +
+          labs(title = dis, x = 'Year', y = 'Prevalence')
+}
+
+baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+visProjectedBySubpop(baseline, cd = 'cancre')
+visProjectedBySubpop(baseline, cd = 'diabe')
+visProjectedBySubpop(baseline, cd = 'hearte')
+```
+
 # Github Reporting

### 2021-01-04 FEM_R/model_analysis.Rmd
Added more visualisations to model_analysis.Rmd and fixed problems created by rmarkdown update

@@
 title: "FEM Model Analysis"
-output:
-  html_document:
-    df_print: paged
+output: html_notebook
 ---
@@
 
-  melted <- melt(combine, id.var = 'year', value.name = 'Prevalence')
+  melted <- reshape2::melt(combine, id.var = 'year', value.name = 'Prevalence')
 
@@
                 filter(year == 2012) %>%
-                melt(id.var = 'year', value.name = 'Prevalence') %>%
+                reshape2::melt(id.var = 'year', value.name = 'Prevalence') %>%
                 subset(select = -year)
@@
                 dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
-                melt(id.var='year', value.name = 'Prevalence')
+                reshape2::melt(id.var='year', value.name = 'Prevalence')
 
@@
               dplyr::rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
-              melt(id.var = 'year', value.name = 'Prevalence')
+              reshape2::melt(id.var = 'year', value.name = 'Prevalence')
 

### 2021-01-08 FEM_R/mortality_adjustment.Rmd
Started writing an R Notebook detailing how the mortality adjustment tables are calculated

---
title: "Generate Mortality Adjustment for FEM"
output: html_notebook
author: Luke Archer
email: l.archer@leeds.ac.uk
---

This notebook will go through the steps taken to generate the mortality adjustment table for the FEM.

The mortality adjustment is a table that specifies how the mortality rate should be adjusted over time in the FEM. As a rule, the mortality rate is decreasing over time. The original US version divided the mortality adjustment into a rate for those aged below 65, and those aged above. As the data we are using contains information by gender and single year of age, we will instead attempt to make a version of this with a separate rate by gender and single year of age.

```{r setup, include = FALSE}

# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(readODS)

```

First lets read in the .xls document containing the life tables. The life tables are divided into period and cohort life tables, which are further subdivided by gender, resulting in 4 separate life tables: a cohort and period life table for men, and the same for women.

```{r}

#ppp <- read_excel('../Mortality_Projections/engppp18ex.ods')
m_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 3,
                     skip = 9,
                     row_names = TRUE)
f_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 4,
                     skip = 9,
                     row_names = TRUE)
m_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 5,
                     skip = 9,
                     row_names = FALSE)
f_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 6,
                     skip = 9,
                     row_names = FALSE)

m_period <- m_period[-1,]
f_period <- f_period[-1,]
m_cohort <- m_cohort[-1,]
f_cohort <- f_cohort[-1,]

```

Before using this data to calculate mortality adjustment, we will produce one version using the assumptions laid out in [this](http://www.ssisi.ie/SSISI172Naqvi&WhelanRevised20190614.pdf) paper assessing different methodologies of estimating future life expectancies (focussing on Table VII on page 15). The paper has a detailed comparison between UK (ONS) and Irish (CSO) outputs, providing information on the assumptions made to produce the UK outputs. Here, we will use these assumptions to generate a mortality adjustment table, and we can use this to compare with a more granular version we will create soon using the life tables we loaded in in the previous chunk.

SIDELINE THE ABOVE, not doing this first might have a go later.

```{r}

#PASS

```

Forget the previous paragraph and chunk for a bit, instead we're going to try to use the cohort life tables to work out an age and gender specific life expectancy for each year.

```{r}

# First drop all ages below 50 for both male and female dfs
#male <- tibble::rownames_to_column(m_cohort) # create copy of df so we can run this chunk independantly if we make mistakes
male <- m_cohort %>% rename(Age = 'Attained age (years)')
female <- f_cohort %>% rename(Age = 'Attained age (years)')

# Filter out ages below 50
male <- male %>% filter(Age >= 50)
female <- female %>% filter(Age >= 50)

# Replace rownames with values from Age row
rownames(male) <- male$Age
rownames(female) <- female$Age
male <- subset(male, select = -Age)
female <- subset(female, select = -Age)

# Testing where the value for 50 yr old in 2012 is
#foo <- male[1,33]
#print(foo)
# Now normalise to the value for a 50 year old in 2012
male_norm <- male / male[1,32]
female_norm <- female / female[1,32]

male_norm

# Divide all by 1 to flip the values, so values below 1 mean reduction in mortality, and vice versa
m_inv <- 1 / male_norm
f_inv <- 1 / female_norm

m_inv

```

HAD A BETTER IDEA! Not going to remove the above chunk but I think I have figured out a mistake, going to try again. I think instead of scaling to single value for a 50 year old in 2012, I think we need to scale to the values for all ages in 2012. Then, we would have a value of 1 for every age in 2012, and the values would change over time, at a different rate for each age.

```{r}
# Rename age column and filter out ages below 50
male <- m_cohort %>% 
            rename(Age = 'Attained age (years)') %>%
            filter(Age >= 50)

female <- f_cohort %>% 
            rename(Age = 'Attained age (years)') %>%
            filter(Age >= 50)

# Replace rownames with values from Age row
rownames(male) <- male$Age
rownames(female) <- female$Age
male <- subset(male, select = -Age)
female <- subset(female, select = -Age)

# Add prefix to colnames so we can reference them with $ operator
#colnames(male) <- paste("y", colnames(male), sep = '')
#colnames(female) <- paste("y", colnames(female), sep = '')

# Scale all values to the 2012 column
male_norm <- male / male[,32]
female_norm <- female / female[,32]

# Take the inverse (we want lower values to represent reduced mortality)
m_inv <- 1 / male_norm
f_inv <- 2 / female_norm

m_inv

```

This is much more like it I think. We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and older years show an increase (where mortlity used to be higher).

The next step now is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model). We can do this by reshaping the data, which should then create multiple rows, one for each gender, age, and year combination.

```{r}
# Copy dfs so we don't have to rerun other chunks
m_copy <- m_inv
f_copy <- f_inv

# Add a column for gender and reinsert the rownames back as a column for reshaping
m_copy <- cbind(Male = 1, rownames(m_copy), m_copy)
rownames(m_copy) <- 1:nrow(m_copy)
f_copy <- cbind(Male = 0, rownames(f_copy), f_copy)
rownames(f_copy) <- 1:nrow(f_copy)
# Rename age column
m_copy <- rename(m_copy, Age = "rownames(m_copy)")
f_copy <- rename(f_copy, Age = "rownames(f_copy)")

# First attempt at reshaping with gather
m_long <- gather(m_copy, year, adjustment, 3:ncol(m_long), factor_key = TRUE)
# Looks good!

```

### 2021-01-10 FEM_R/mortality_adjustment.Rmd
Added an R notebook file that handles the production of the mortality adjustment table, based on life tables supplied by the ONS. Had to add a column definition into ELSA_vars.txt, and included the settings file in ELSA_baseline.csv for testing. All works now!

@@
 
-First lets read in the .xls document containing the life tables.
-The life tables are divided into period and cohort life tables, which are further subdivided by gender, resulting in 4 separate life tables: a cohort and period life table for men, and the same for women.
+First read in the .xls document containing the life tables. The life tables are divided into period and cohort tables, which are further subdivided by gender, resulting in 4 separate life tables.
 
 ```{r}
-
-#ppp <- read_excel('../Mortality_Projections/engppp18ex.ods')
+# Read in male and female, cohort and period data separately 
 m_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
@@
 
+# First row is empty so remove
 m_period <- m_period[-1,]
@@
 f_cohort <- f_cohort[-1,]
-
 ```
 
-Before using this data to calculate mortality adjustment, we will produce one version using the assumptions laid out in [this](http://www.ssisi.ie/SSISI172Naqvi&WhelanRevised20190614.pdf) paper assessing different methodologies of estimating future life expectancies (focussing on Table VII on page 15).
-The paper has a detailed comparison between UK (ONS) and Irish (CSO) outputs, providing information on the assumptions made to produce the UK outputs.
-Here, we will use these assumptions to generate a mortality adjustment table, and we can use this to compare with a more granular version we will create soon using the life tables we loaded in in the previous chunk.
+Steps:
 
-SIDELINE THE ABOVE, not doing this first might have a go later.
-
-```{r}
-
-#PASS
-
-```
-
-Forget the previous paragraph and chunk for a bit, instead we're going to try to use the cohort life tables to work out an age and gender specific life expectancy for each year.
-
-```{r}
-
-# First drop all ages below 50 for both male and female dfs
-#male <- tibble::rownames_to_column(m_cohort) # create copy of df so we can run this chunk independantly if we make mistakes
-male <- m_cohort %>% rename(Age = 'Attained age (years)')
-female <- f_cohort %>% rename(Age = 'Attained age (years)')
-
-# Filter out ages below 50
-male <- male %>% filter(Age >= 50)
-female <- female %>% filter(Age >= 50)
-
-# Replace rownames with values from Age row
-rownames(male) <- male$Age
-rownames(female) <- female$Age
-male <- subset(male, select = -Age)
-female <- subset(female, select = -Age)
-
-# Testing where the value for 50 yr old in 2012 is
-#foo <- male[1,33]
-#print(foo)
-# Now normalise to the value for a 50 year old in 2012
-male_norm <- male / male[1,32]
-female_norm <- female / female[1,32]
-
-male_norm
-
-# Divide all by 1 to flip the values, so values below 1 mean reduction in mortality, and vice versa
-m_inv <- 1 / male_norm
-f_inv <- 1 / female_norm
-
-m_inv
-
-```
-
-HAD A BETTER IDEA!
-Not going to remove the above chunk but I think I have figured out a mistake, going to try again.
-I think instead of scaling to single value for a 50 year old in 2012, I think we need to scale to the values for all ages in 2012.
-Then, we would have a value of 1 for every age in 2012, and the values would change over time, at a different rate for each age.
+- Drop ages below 50
+- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
+- Normalise matrix to 2012 values (column 32)
+- Calculate inverse (value < 1 == decreased chance of mortality)
 
@@
 male <- m_cohort %>% 
-            rename(Age = 'Attained age (years)') %>%
-            filter(Age >= 50)
+            rename(age = 'Attained age (years)') %>%
+            filter(age >= 50)
 
 female <- f_cohort %>% 
-            rename(Age = 'Attained age (years)') %>%
-            filter(Age >= 50)
+            rename(age = 'Attained age (years)') %>%
+            filter(age >= 50)
 
 # Replace rownames with values from Age row
-rownames(male) <- male$Age
-rownames(female) <- female$Age
-male <- subset(male, select = -Age)
-female <- subset(female, select = -Age)
-
-# Add prefix to colnames so we can reference them with $ operator
-#colnames(male) <- paste("y", colnames(male), sep = '')
-#colnames(female) <- paste("y", colnames(female), sep = '')
+rownames(male) <- male$age
+rownames(female) <- female$age
+male <- subset(male, select = -age)
+female <- subset(female, select = -age)
 
@@
 
+head(female_norm)
+
 # Take the inverse (we want lower values to represent reduced mortality)
 m_inv <- 1 / male_norm
-f_inv <- 2 / female_norm
+f_inv <- 1 / female_norm
 
-m_inv
-
+head(f_inv)
 ```
 
-This is much more like it I think.
-We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and older years show an increase (where mortlity used to be higher).
+We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and previous years show an increase (where mortality used to be higher).
 
-The next step now is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model).
-We can do this by reshaping the data, which should then create multiple rows, one for each gender, age, and year combination.
+The next step is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model).
+We can do this with tidyr::gather.
 
@@
 # Add a column for gender and reinsert the rownames back as a column for reshaping
-m_copy <- cbind(Male = 1, rownames(m_copy), m_copy)
+m_copy <- cbind(male = 1, rownames(m_copy), m_copy)
 rownames(m_copy) <- 1:nrow(m_copy)
-f_copy <- cbind(Male = 0, rownames(f_copy), f_copy)
+f_copy <- cbind(male = 0, rownames(f_copy), f_copy)
 rownames(f_copy) <- 1:nrow(f_copy)
 # Rename age column
-m_copy <- rename(m_copy, Age = "rownames(m_copy)")
-f_copy <- rename(f_copy, Age = "rownames(f_copy)")
+m_copy <- rename(m_copy, age = "rownames(m_copy)")
+f_copy <- rename(f_copy, age = "rownames(f_copy)")
 
 # First attempt at reshaping with gather
-m_long <- gather(m_copy, year, adjustment, 3:ncol(m_long), factor_key = TRUE)
-# Looks good!
+m_long <- gather(m_copy, year, adjustment, 3:ncol(m_copy), factor_key = TRUE)
+f_long <- gather(f_copy, year, adjustment, 3:ncol(f_copy), factor_key = TRUE)
 
+head(f_long) # Looks good!
 ```
+
+Last step is just to row bind and stick everything together, then can manually delete the 'adjustment' column heading before adding to a folder called 'mortality_adj_ELSA' in FEM_CPP_settings.
+
+```{r}
+# Combine male and female data
+adjustment <- rbind(m_long, f_long)
+
+head(adjustment)
+
+# Write to folder outside trunk so we can't accidentally replace it later
+write_tsv(adjustment, path = '/home/luke/Documents/E_FEM_clean/Mortality_Projections/mortality_adj.txt', 
+                      append = FALSE, 
+                      col_names = TRUE)
+```

### 2021-01-10 FEM_R/model_analysis.Rmd
Added survival curve to debug document

@@
 
+### Survival Curve
+
+```{r echo = FALSE}
+cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
+
+cohort$age <- (cohort$year - 2012) + 50
+
+cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
+cohort$survival_m <- cohort$m_endpop_m / cohort$m_endpop_m[1]
+cohort$survival_f <- cohort$m_endpop_f / cohort$m_endpop_f[1]
+
+# plot up survival curve
+ggplot(cohort, aes(x=age, y=survival)) +
+        geom_line(col='black', size=0.3) +
+        labs(title='Cohort Survival: All', y='Survival', x='Age')
+```
+
 # Github Reporting

### 2021-01-11 FEM_R/model_analysis.Rmd
Copied work on the AgeUK validation from the barebones FEM version over to development.

@@
 
+# AgeUK Almanac External Validation
+
+'The Age UK almanac of disease profiles in later life', published in October 2015, contains information on the prevalence of major diseases, conditions, and syndromes affecting older people in England, broken down by age and sex.
+
+```{r}
+# Need a function here to handle the repetition
+get_prevalence <- function(FEM) {
+
+}
+
+# Now read in male and female AgeUK data
+male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
+female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+# Rename columns
+colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
+colnames(female_ageUK) <- colnames(male_ageUK)
+
+# Collect conditions into a list for extracting from FEM
+# Will need to be written in FEM shorthand
+keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+
+# Now get FEM output for 2014
+FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+FEM_2014 <- FEM_2014[keep_vars]
+# extract males and females
+mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+
+m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
+f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
+
+m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
+f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
+colnames(m6064_prev) <- 'FEM_male_6064'
+colnames(f6064_prev) <- 'FEM_female_6064'
+
+prev_df <- data.frame(m6064_prev, f6064_prev)
+
+# correct colnames in ageUK stuff
+male_ageUK <- male_ageUK %>% filter(conditions == c('Hypertension', 'Stroke', 'Diabetes', 'Dementia'))
+
+# Now calculate prevalence by age group for each chronic disease
+# prevalence = total / affected
+# Diabetes
+male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
+female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
+colnames(male_prev) <- 'FEM_male'
+colnames(female_prev) <- 'FEM_female'
+
+```
+
 # Github Reporting

### 2021-01-12 FEM_R/model_analysis.Rmd
WIP few chunks in model_analysis.Rmd for comparing simulation to AgeUK almanac

@@
 # Need a function here to handle the repetition
-get_prevalence <- function(FEM) {
+get_FEM_prev <- function(FEM, age_range, keep_vars) {
+  # Keep only important vars
+  FEM <- FEM[keep_vars]
+  # Separate by gender
+  mFEM <- FEM %>% filter(male == 1)
+  fFEM <- FEM %>% filter(male == 0)
+  # Filter by age from age_range
+  mFEM <- mFEM %>% filter(age %in% (age_range))
+  fFEM <- fFEM %>% filter(age %in% (age_range))
+  # Get proportion of total pop. (prevalence)
+  mFEM <- data.frame((colSums(mFEM) / nrow(mFEM)) * 100)
+  fFEM <- data.frame((colSums(fFEM) / nrow(fFEM)) * 100)
 
+  # produce a string for column naming
+  age <- paste0(age_range[1], '-', age_range[5])
+
+  colnames(mFEM) <- paste0('FEM_male_', age)
+  colnames(fFEM) <- paste0('FEM_female_', age)
+
+  #male <- cbind(male, mFEM)
+
+  # Collect into 1 df
+  prev_df <- data.frame(mFEM, fFEM)
+  return(prev_df)
+
+  #return(out)
+}
+
+package <- function() {
+  # Set up objects for holding and returning the prevalences, separate dfs for male and female
+  out <- list()
+  male <- data.frame()
+  female <- data.frame()
+
+  age_list <- c('Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+')
+  age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)
 }
@@
 FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-FEM_2014 <- FEM_2014[keep_vars]
-# extract males and females
+
+test <- get_FEM_prev(FEM_2014, 60:64, keep_vars)
+
+#FEM_2014 <- FEM_2014[keep_vars]
+## extract males and females
 mFEM_2014 <- FEM_2014 %>% filter(male == 1)
@@
 
+# AgeUK conditions to extract
+#condlist <- c('Dementia', 'Hypertension', 'Diabetes', 'Stroke')
+
 # correct colnames in ageUK stuff
-male_ageUK <- male_ageUK %>% filter(conditions == c('Hypertension', 'Stroke', 'Diabetes', 'Dementia'))
+#male_ageUK <- male_ageUK %>% filter(Condition == 'Dementia' | Condition == 'Hypertension' | Condition == 'Diabetes' | Condition == 'Stroke')
+#male_ageUK <- male_ageUK %>% filter(Condition %in% condlist)
 
@@
 # Diabetes
-male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
-female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
-colnames(male_prev) <- 'FEM_male'
-colnames(female_prev) <- 'FEM_female'
+#male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
+#female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
+#colnames(male_prev) <- 'FEM_male'
+#colnames(female_prev) <- 'FEM_female'
+
+```
+
+```{r}
+
+tmp <- function(m, f, age) {
+
+  m_age <- m %>% filter(age %in% (age))
+  f_age <- f %>% filter(age %in% (age))
+
+  m_prev <- data.frame((colSums(m_age) / nrow(m_age)) * 100)
+  f_prev <- data.frame((colSums(f_age) / nrow(f_age)) * 100)
+
+  list <- list(m_prev, f_prev)
+  return(list)
+}
+
+# Now read in male and female AgeUK data
+male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
+female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+# Rename columns
+colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
+colnames(female_ageUK) <- colnames(male_ageUK)
+
+# Collect conditions into a list for extracting from FEM
+# Will need to be written in FEM shorthand
+keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+
+# Now get FEM output for 2014
+FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+FEM <- FEM_2014[keep_vars]
+
+# extract males and females
+mFEM <- FEM %>% filter(male == 1)
+fFEM <- FEM %>% filter(male == 0)
+
+age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)
+
+m6064 <- mFEM %>% filter(age %in% (60:64))
+f6064 <- fFEM %>% filter(age %in% (60:64))
+
+m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
+f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
+colnames(m6064_prev) <- 'FEM_male_6064'
+colnames(f6064_prev) <- 'FEM_female_6064'
+
+male <- list()
+female <- list()
+
+mf_list <- tmp(mFEM, fFEM, 60:64)
+male <- list(male, mf_list[1])
+female <- list(female, mf_list[2])
 

### 2021-01-20 FEM_R/mortality_adjustment.Rmd
Updated the mortality adjustment to use a different data source. Now using mortality rates (qx) to calculate the adjustment, which is much more similar to the US values now

@@
 require(readODS)
+require(ggplot2)
 
@@
 ```
+
+# DISREGARD ALL THE ABOVE
+
+It's all wrong. Start again but can reuse some of the code for another dataset.
+
+# Second Attempt Using Mortality Rates (qx)
+
+```{r}
+m_period <- read_ods('../Mortality_Projections/engppp18qx.ods',
+                     sheet = 3,
+                     skip = 6,
+                     row_names = TRUE)
+f_period <- read_ods('../Mortality_Projections/engppp18qx.ods',
+                     sheet = 4,
+                     skip = 6,
+                     row_names = TRUE)
+m_cohort <- read_ods('../Mortality_Projections/engppp18qx.ods',
+                     sheet = 5,
+                     skip = 6,
+                     row_names = FALSE)
+f_cohort <- read_ods('../Mortality_Projections/engppp18qx.ods',
+                     sheet = 6,
+                     skip = 6,
+                     row_names = FALSE)
+
+```
+
+Steps:
+
+- Drop ages below 50
+- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
+- Calculate ratio between values for consecutive years
+
+```{r}
+# Rename age column and filter out ages below 50
+male <- m_cohort %>% 
+            rename(age = 'age (years)') %>%
+            filter(age >= 50)
+
+female <- f_cohort %>% 
+            rename(age = 'age (years)') %>%
+            filter(age >= 50)
+
+# Replace rownames with values from Age row
+rownames(male) <- male$age
+rownames(female) <- female$age
+male <- subset(male, select = -age)
+female <- subset(female, select = -age)
+
+ratio_df <- data.frame(matrix(vector(mode = 'numeric', length = nrow(male) * ncol(male)), nrow = nrow(male), ncol = ncol(male)))
+
+for(i in nrow(male)) {
+  for(j in ncol(male)) {
+    ratio_df[i,j] <- male[i,j] / male[i-1, j-1]
+  }
+}
+
+```
+
+```{r}
+# Rename age column and filter out ages below 50
+male <- m_cohort %>% 
+            rename(age = 'age (years)') %>%
+            filter(age >= 50)
+
+female <- f_cohort %>% 
+            rename(age = 'age (years)') %>%
+            filter(age >= 50)
+
+# Replace rownames with values from Age row
+rownames(male) <- male$age
+rownames(female) <- female$age
+male <- subset(male, select = -age)
+female <- subset(female, select = -age)
+
+# Scale all values to the 2012 column
+# SET TO 2004 FOR TESTING AGAINST US, NEED TO CHANGE THIS BACK 
+male_norm <- male / male[,24]
+female_norm <- female / female[,24]
+
+head(female_norm)
+```
+
+```{r}
+# Copy dfs so we don't have to rerun other chunks
+m_copy <- male_norm
+f_copy <- female_norm
+
+# Add a column for gender and reinsert the rownames back as a column for reshaping
+m_copy <- cbind(male = 1, rownames(m_copy), m_copy)
+rownames(m_copy) <- 1:nrow(m_copy)
+f_copy <- cbind(male = 0, rownames(f_copy), f_copy)
+rownames(f_copy) <- 1:nrow(f_copy)
+# Rename age column
+m_copy <- rename(m_copy, age = "rownames(m_copy)")
+f_copy <- rename(f_copy, age = "rownames(f_copy)")
+
+# First attempt at reshaping with gather
+m_long <- gather(m_copy, year, adjustment, 3:ncol(m_copy), factor_key = TRUE)
+f_long <- gather(f_copy, year, adjustment, 3:ncol(f_copy), factor_key = TRUE)
+
+head(f_long) # Looks good!
+```
+
+Write out the file to E_FEM_clean/Mortality_Projections/.
+
+```{r}
+# Combine male and female data
+adjustment <- rbind(m_long, f_long)
+
+head(adjustment)
+
+# Write to folder outside trunk so we can't accidentally replace it later
+write_tsv(adjustment, path = '/home/luke/Documents/E_FEM_clean/Mortality_Projections/mortality_adj.txt', 
+                      append = FALSE, 
+                      col_names = TRUE)
+```
+
+```{r}
+# Test against US data
+us_adj <- read_tsv('../Mortality_Projections/mortality_adj_US.txt', col_names = TRUE, skip = 1)
+
+adj <- adjustment
+
+adj$year <- as.numeric(as.character(adj$year))
+us_adj$year <- as.numeric(as.character(us_adj$year))
+
+# Extract only 60 year olds from my dataset
+# Drop all years before 1990 as this is when the US data starts
+adj_60 <- adj %>%
+                filter(age == 60) %>%
+                select(-age) %>%
+                filter(year > 1989) %>%
+                rename(group = male)
+
+# Drop all years after 2068 for US as this is when Eng data stops
+# Also drop over 65 values
+adj_us <- us_adj %>%
+                filter(year < 2069) %>%
+                filter(l2age65pd == 0) %>%
+                rename(group = l2age65pd)
+adj_us$group <- 2
+
+```
+
+NOW PLOT!!
+
+```{r}
+
+# Now combine the data and start plotting
+combined <- rbind(adj_60, adj_us)
+
+combined$group[combined$group == 1] <- 'male'
+combined$group[combined$group == 0] <- 'female'
+combined$group[combined$group == 2] <- 'US'
+
+ggplot(data = combined, aes(x = year, y = adjustment, color = group)) +
+      geom_line(size = 0.5) +
+      labs(title = 'Age 60')
+
+cut <- combined %>% filter(group != 'US')
+
+ggplot(data = cut, aes(x = year, y = adjustment, color = group)) +
+      geom_line()
+
+cut2 <- cut %>% filter(group == 'female')
+
+ggplot(data = cut2, aes(x = year, y = adjustment, color = group)) +
+      geom_line()
+
+```
+
+```{r}
+# Test against US data
+us_adj2 <- read_tsv('../Mortality_Projections/mortality_adj_US.txt', col_names = TRUE, skip = 1)
+
+adj2 <- adjustment
+
+adj2$year <- as.numeric(as.character(adj2$year))
+us_adj$year <- as.numeric(as.character(us_adj$year))
+# Extract only 60 year olds from my dataset
+# Drop all years before 1990 as this is when the US data starts
+adj_60 <- adj2 %>%
+                filter(age == 75) %>%
+                select(-age) %>%
+                filter(year > 1989) %>%
+                rename(group = male)
+
+# Drop all years after 2068 for US as this is when Eng data stops
+# Also drop over 65 values
+adj_us <- us_adj %>%
+                filter(year < 2069) %>%
+                filter(l2age65pd == 1) %>%
+                rename(group = l2age65pd)
+adj_us$group <- 2
+
+```
+```{r}
+
+# Now combine the data and start plotting
+combined <- rbind(adj_60, adj_us)
+
+combined$group[combined$group == 1] <- 'male'
+combined$group[combined$group == 0] <- 'female'
+combined$group[combined$group == 2] <- 'US'
+
+ggplot(data = combined, aes(x = year, y = adjustment, color = group)) +
+      geom_line(size = 0.5) +
+      labs(title = 'Age 75')
+
+cut <- combined %>% filter(group != 'US')
+
+ggplot(data = cut, aes(x = year, y = adjustment, color = group)) +
+      geom_line()
+
+cut2 <- cut %>% filter(group == 'female')
+
+ggplot(data = cut2, aes(x = year, y = adjustment, color = group)) +
+      geom_line()
+
+```

### 2021-01-25 FEM_R/mortality_adjustment.Rmd
Finished replacing the mortality adjustment using ONS data on mortality rate (qx)

@@
 ```
-
-First read in the .xls document containing the life tables. The life tables are divided into period and cohort tables, which are further subdivided by gender, resulting in 4 separate life tables.
-
-```{r}
-# Read in male and female, cohort and period data separately 
-m_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
-                     sheet = 3,
-                     skip = 9,
-                     row_names = TRUE)
-f_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
-                     sheet = 4,
-                     skip = 9,
-                     row_names = TRUE)
-m_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
-                     sheet = 5,
-                     skip = 9,
-                     row_names = FALSE)
-f_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
-                     sheet = 6,
-                     skip = 9,
-                     row_names = FALSE)
-
-# First row is empty so remove
-m_period <- m_period[-1,]
-f_period <- f_period[-1,]
-m_cohort <- m_cohort[-1,]
-f_cohort <- f_cohort[-1,]
-```
-
-Steps:
-
-- Drop ages below 50
-- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
-- Normalise matrix to 2012 values (column 32)
-- Calculate inverse (value < 1 == decreased chance of mortality)
-
-```{r}
-# Rename age column and filter out ages below 50
-male <- m_cohort %>% 
-            rename(age = 'Attained age (years)') %>%
-            filter(age >= 50)
-
-female <- f_cohort %>% 
-            rename(age = 'Attained age (years)') %>%
-            filter(age >= 50)
-
-# Replace rownames with values from Age row
-rownames(male) <- male$age
-rownames(female) <- female$age
-male <- subset(male, select = -age)
-female <- subset(female, select = -age)
-
-# Scale all values to the 2012 column
-male_norm <- male / male[,32]
-female_norm <- female / female[,32]
-
-head(female_norm)
-
-# Take the inverse (we want lower values to represent reduced mortality)
-m_inv <- 1 / male_norm
-f_inv <- 1 / female_norm
-
-head(f_inv)
-```
-
-We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and previous years show an increase (where mortality used to be higher).
-
-The next step is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model).
-We can do this with tidyr::gather.
-
-```{r}
-# Copy dfs so we don't have to rerun other chunks
-m_copy <- m_inv
-f_copy <- f_inv
-
-# Add a column for gender and reinsert the rownames back as a column for reshaping
-m_copy <- cbind(male = 1, rownames(m_copy), m_copy)
-rownames(m_copy) <- 1:nrow(m_copy)
-f_copy <- cbind(male = 0, rownames(f_copy), f_copy)
-rownames(f_copy) <- 1:nrow(f_copy)
-# Rename age column
-m_copy <- rename(m_copy, age = "rownames(m_copy)")
-f_copy <- rename(f_copy, age = "rownames(f_copy)")
-
-# First attempt at reshaping with gather
-m_long <- gather(m_copy, year, adjustment, 3:ncol(m_copy), factor_key = TRUE)
-f_long <- gather(f_copy, year, adjustment, 3:ncol(f_copy), factor_key = TRUE)
-
-head(f_long) # Looks good!
-```
-
-Last step is just to row bind and stick everything together, then can manually delete the 'adjustment' column heading before adding to a folder called 'mortality_adj_ELSA' in FEM_CPP_settings.
-
-```{r}
-# Combine male and female data
-adjustment <- rbind(m_long, f_long)
-
-head(adjustment)
-
-# Write to folder outside trunk so we can't accidentally replace it later
-write_tsv(adjustment, path = '/home/luke/Documents/E_FEM_clean/Mortality_Projections/mortality_adj.txt', 
-                      append = FALSE, 
-                      col_names = TRUE)
-```
-
-# DISREGARD ALL THE ABOVE
-
-It's all wrong. Start again but can reuse some of the code for another dataset.
 
@@
 male <- m_cohort %>% 
-            rename(age = 'age (years)') %>%
+            dplyr::rename(age = 'age (years)') %>%
             filter(age >= 50)
@@
 female <- f_cohort %>% 
-            rename(age = 'age (years)') %>%
+            dplyr::rename(age = 'age (years)') %>%
             filter(age >= 50)
@@
 # Rename age column
-m_copy <- rename(m_copy, age = "rownames(m_copy)")
-f_copy <- rename(f_copy, age = "rownames(f_copy)")
+m_copy <- dplyr::rename(m_copy, age_elsa = "rownames(m_copy)")
+f_copy <- dplyr::rename(f_copy, age_elsa = "rownames(f_copy)")
 
@@
 ```
+
+## Now check against US mortality adjustment
 
@@
 ```
+
+# Old Defunkt
+
+First read in the .xls document containing the life tables. The life tables are divided into period and cohort tables, which are further subdivided by gender, resulting in 4 separate life tables.
+
+```{r}
+# Read in male and female, cohort and period data separately 
+m_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
+                     sheet = 3,
+                     skip = 9,
+                     row_names = TRUE)
+f_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
+                     sheet = 4,
+                     skip = 9,
+                     row_names = TRUE)
+m_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
+                     sheet = 5,
+                     skip = 9,
+                     row_names = FALSE)
+f_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
+                     sheet = 6,
+                     skip = 9,
+                     row_names = FALSE)
+
+# First row is empty so remove
+m_period <- m_period[-1,]
+f_period <- f_period[-1,]
+m_cohort <- m_cohort[-1,]
+f_cohort <- f_cohort[-1,]
+```
+
+Steps:
+
+- Drop ages below 50
+- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
+- Normalise matrix to 2012 values (column 32)
+- Calculate inverse (value < 1 == decreased chance of mortality)
+
+```{r}
+# Rename age column and filter out ages below 50
+
+male <- m_cohort %>% 
+            dplyr::rename(age = "Attained age (years)") %>%
+            filter(age >= 50)
+
+female <- f_cohort %>% 
+            dplyr::rename(age = "Attained age (years)") %>%
+            filter(age >= 50)
+
+# Replace rownames with values from Age row
+rownames(male) <- male$age
+rownames(female) <- female$age
+male <- subset(male, select = -age)
+female <- subset(female, select = -age)
+
+# Scale all values to the 2012 column
+male_norm <- male / male[,32]
+female_norm <- female / female[,32]
+
+head(female_norm)
+
+# Take the inverse (we want lower values to represent reduced mortality)
+m_inv <- 1 / male_norm
+f_inv <- 1 / female_norm
+
+head(f_inv)
+```
+
+We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and previous years show an increase (where mortality used to be higher).
+
+The next step is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model).
+We can do this with tidyr::gather.
+
+```{r}
+# Copy dfs so we don't have to rerun other chunks
+m_copy <- m_inv
+f_copy <- f_inv
+
+# Add a column for gender and reinsert the rownames back as a column for reshaping
+m_copy <- cbind(male = 1, rownames(m_copy), m_copy)
+rownames(m_copy) <- 1:nrow(m_copy)
+f_copy <- cbind(male = 0, rownames(f_copy), f_copy)
+rownames(f_copy) <- 1:nrow(f_copy)
+# Rename age column
+m_copy <- dplyr::rename(m_copy, age = "rownames(m_copy)")
+f_copy <- dplyr::rename(f_copy, age = "rownames(f_copy)")
+
+# First attempt at reshaping with gather
+m_long <- gather(m_copy, year, adjustment, 3:ncol(m_copy), factor_key = TRUE)
+f_long <- gather(f_copy, year, adjustment, 3:ncol(f_copy), factor_key = TRUE)
+
+head(f_long) # Looks good!
+```
+
+Last step is just to row bind and stick everything together, then can manually delete the 'adjustment' column heading before adding to a folder called 'mortality_adj_ELSA' in FEM_CPP_settings.
+
+```{r}
+# Combine male and female data
+adjustment <- rbind(m_long, f_long)
+
+head(adjustment)
+
+# Write to folder outside trunk so we can't accidentally replace it later
+write_tsv(adjustment, path = '/home/luke/Documents/E_FEM_clean/Mortality_Projections/mortality_adj.txt', 
+                      append = FALSE, 
+                      col_names = TRUE)
+```
+
+# DISREGARD ALL THE ABOVE
+
+It's all wrong. Start again but can reuse some of the code for another dataset.

### 2021-01-27 FEM_R/model_analysis.Rmd
More WIP stuff for the AgeUK alamanac validation
Defined multiple new output subpopulations in summary_output_gen. Added these to measures_subpop_ELSA so we can use the summary out for the ageUK validation
Removed a chunk that visualised things based on subpopulations as the subpopulations have now changed

@@
 
-### Baseline Projections by Subpopulation
-
-```{r echo = FALSE}
-visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {
-
-  if(cd == 'cancre') {
-    dis <- 'Cancer'
-  } else if(cd == 'diabe') {
-    dis <- 'Diabetes'
-  } else if(cd == 'hearte') {
-    dis <- 'Heart Disease'
-  }
-
-  # Generate varnames
-  all <- paste0('p_', cd, '_all')
-  # age
-  a5564 <- paste0('p_', cd, '_5564')
-  a6574 <- paste0('p_', cd, '_6574')
-  a7584 <- paste0('p_', cd, '_7584')
-  a85p <- paste0('p_', cd, '_85p')
-
-  ageMelt <- base %>%
-              select(c('year', a5564, a6574, a7584, a85p)) %>%
-              dplyr::rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
-              reshape2::melt(id.var = 'year', value.name = 'Prevalence')
-
-  ggplot(data=ageMelt, aes(x=year, y=Prevalence, color = variable)) +
-          geom_line() +
-          labs(title = dis, x = 'Year', y = 'Prevalence')
-}
-
-baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-
-visProjectedBySubpop(baseline, cd = 'cancre')
-visProjectedBySubpop(baseline, cd = 'diabe')
-visProjectedBySubpop(baseline, cd = 'hearte')
-```
-
 ### Survival Curve
@@
 
+mf_list <- tmp(mFEM, fFEM, 65:69)
+male <- list(male, mf_list[1])
+female <- list(female, mf_list[2])
+
 ```

### 2021-01-27 FEM_R/validation.Rmd
Started the validation work in a separate notebook to model analysis

--- 2021-01-27 FEM_R/model_analysis.Rmd
@@
 ---
-title: "FEM Model Analysis"
+title: "External Validation"
 output: html_notebook
 ---
-
-This notebook is for debugging purposes. The current plan is to write a Make rule for running this notebook automatically at the end of a model run.
-
-The purpose of this notebook is:
-- Present the T-tests altogether in one document.
-- Present all transition models in same document.
-- Add visualisation for cross-validation 2 output
-
-Do all this for both minimal models and cross-validation.
 
@@
 require(tidyverse)
-require(reactable)
+#require(reactable)
 require(haven)
-require(data.table)
+#require(data.table)
 require(reshape2)
 require(ggplot2)
-require(plyr)
+#require(plyr)
 
-```
-
-```{r include=FALSE}
-
-# Need this code block for reactable tables to be rendered properly in HTML
-# Fix from here: https://www.gitmemory.com/issue/glin/reactable/43/618802061
-htmltools::tagList(
-  reactR::html_dependency_react(),
-  reactR::html_dependency_reacttools(),
-  reactable::reactable(iris)
-)
-
-# function to rename the dataframes, so we can run in an lapply()
-rename_first_var <- function(testCSV, prefix) {
-  testCSV[,1] <- paste0(testCSV[,1], '_', prefix)
-  return(testCSV)
-}
-
-# function for shortening column names
-shorten_columns <- function(testCSV) {
-  testCSV <- testCSV %>%
-                rename_at(vars(starts_with('fem_mean_wave')), funs(sub('fem_mean_wave', 'FEM_', .))) %>%
-                rename_at(vars(starts_with('elsa_mean_wave')), funs(sub('elsa_mean_wave', 'ELSA_', .))) %>%
-                rename_at(vars(starts_with('p_value_wave')), funs(sub('p_value_wave', 'P_Value_', .))) %>%
-                dplyr::rename('Variable' = variable)
-}
-
-```
-
-<!-- First, we will collect and present all the T-tests together. We will add the prefix 'min_' or 'cv_' onto the variable names so we don't get mixed up. -->
-
-```{r include=FALSE}
-# names for T-tests directory and filename stem
-min <- "ELSA_minimal/T-tests/"
-stem <- "minimal_all_waves_"
-
-# Read in all the T-test csv files
-min_uw <- read.csv(paste0(min, stem, "unweighted.csv"))
-min_demog <- read.csv(paste0(min, stem, "demog.csv"))
-min_binhlth <- read.csv(paste0(min, stem, "binhlth.csv"))
-min_binecon <- read.csv(paste0(min, stem, "binecon.csv"))
-min_risk <- read.csv(paste0(min, stem, "risk.csv"))
-
-# Collect in a list to do some formatting
-min_list <- list(min_uw, min_demog, min_binhlth, min_binecon, min_risk)
-# Add prefix 'min_' onto every variable name for clarity
-min_list <- lapply(min_list, rename_first_var, prefix='min')
-# Rename columns to be more succinct
-min_list <- lapply(min_list, shorten_columns)
-# Replace na with 0
-min_list <- lapply(min_list, function(x) { x[is.na(x)] <- 0; x} )
-
-```
-
-```{r include=FALSE}
-# names for T-tests directory and filename stem
-CV1 <- "ELSA_CV1/T-tests/"
-stem <- "CV1_all_waves_"
-# Read in files
-cv_uw <- read.csv(paste0(CV1, stem, "unweighted.csv"))
-cv_demog <- read.csv(paste0(CV1, stem, "demog.csv"))
-cv_binhlth <- read.csv(paste0(CV1, stem, "binhlth.csv"))
-cv_binecon <- read.csv(paste0(CV1, stem, "binecon.csv"))
-cv_risk <- read.csv(paste0(CV1, stem, "risk.csv"))
-# Now collect into list and rename vars and cols for clarity
-cv_list <- list(cv_uw, cv_demog, cv_binhlth, cv_binecon, cv_risk)
-cv_list <- lapply(cv_list, rename_first_var, prefix='CV')
-cv_list <- lapply(cv_list, shorten_columns)
-cv_list <- lapply(cv_list, function(x) { x[is.na(x)] <- 0; x} )
-```
-
-<!-- Both Minimal and Cross-validation T-tests are now renamed and collected into a separate list object each.
-We will now combine like tests together (i.e.
-demog with demog, binhlth with binhlth).
--->
-
-```{r include=FALSE}
-# Set up vector with all p-value column names. Require these for table formatting
-pValNames <- c('P_Value_3', 'P_Value_4', 'P_Value_5', 'P_Value_6', 'P_Value_7', 'P_Value_8')
-```
-
-<!-- Lets build the combined tables now. -->
-
-```{r include=FALSE}
-# Row bind to collect both dfs together
-uw <- rbind(min_list[[1]], cv_list[[1]])
-# Reorder rows in alphabetical order. This ensures that like variables are clumped together
-uw <- arrange(uw, Variable)
-
-# demog
-demog <- rbind(min_list[[2]], cv_list[[2]])
-demog <- arrange(demog, Variable)
-
-# binhlth
-binhlth <- rbind(min_list[[3]], cv_list[[3]])
-binhlth <- arrange(binhlth, Variable)
-
-# binecon
-binecon <- rbind(min_list[[4]], cv_list[[4]])
-binecon <- arrange(binecon, Variable)
-
-# risk
-risk <- rbind(min_list[[5]], cv_list[[5]])
-risk <- arrange(risk, Variable)
-```
-
-# Reactable Tables
-
-The [reactable](https://glin.github.io/reactable/index.html) package has the ability to include lots of complex styling, such as additional colour, formatting and even inline graphs built dynamically from the input data.
-
-For the purpose of this document, we are going to include a sticky first column (so we can always see the variable names whilst scrolling), a colour gradient from white to green based on the P Value (low = white, high = green), and a red cell background for any P values of 0.
-
-```{r echo = FALSE}
-## This block will define the logic for creating the tables with Reactable
-# The tables are formatted exaclty the same for each group of vars so makes sense to define this only once
-
-# createTTT - create T-Test Table
-# Have a default column definition, to handle the basic formatting, then specific formatting for the first column (Variable) and P Val cols
-createTTT <- function(t) {
-  # To enable a 'sticky' first column, that always shows as you scroll across the table
-  stickyStyle <- list(position = "sticky", left = 0, background = "#fff", zIndex = 1,
-                      borderRight = "1px solid #eee")
-
-  # Define a colour ramp to use for P Value cell gradients
-  greenPal <- function(x) rgb(colorRamp(c("#ffffff", "#008000"))(x), maxColorValue = 255)
-
-  # Add in the green colour gradient for P Value cells. Also add a red block for value of 0
-  pValColDef <- colDef(
-    style = function(value) {
-      color <- greenPal(value)
-      if (value == 0) {
-        color <- "#ff4c4c"
-      }
-      list(background = color)
-    },
-    minWidth = 110
-  )
-
-  pValColDef2 <- colDef(
-    style = function(value) {
-      if (value >= 0.05) {
-        color <- "#4ca64c"
-      }
-      if (value < 0.05) {
-        color <- "#ff6666"
-      }
-      list(background = color)
-    },
-    minWidth = 110
-  )
-
-  res <- reactable(t,
-            defaultColDef = colDef(
-              header = function(value) gsub('_', ' ', value, fixed=TRUE),
-              cell = function(value) format(value, nsmall = 2),
-              align = 'center'
-            ),
-            columns = list(
-              Variable = colDef(
-                  minWidth = 160,
-                  style = stickyStyle,
-                  headerStyle = stickyStyle),
-              P_Value_1 = pValColDef2,
-              P_Value_2 = pValColDef2,
-              P_Value_3 = pValColDef2,
-              P_Value_4 = pValColDef2,
-              P_Value_5 = pValColDef2,
-              P_Value_6 = pValColDef2,
-              P_Value_7 = pValColDef2,
-              P_Value_8 = pValColDef2
-            ),
-            bordered = TRUE,
-            highlight = TRUE,
-            striped = TRUE)
-
-  return(res)
-}
-
-```
-
-### Unweighted (Died)
-
-```{r echo = FALSE}
-# Create the actual table!
-uwTab <- createTTT(uw)
-uwTab
-```
-
-### Demographics
-
-```{r echo = FALSE}
-# Create the actual table!
-demogTab <- createTTT(demog)
-demogTab
-```
-
-### Binary Health
-```{r echo = FALSE}
-# Create the actual table!
-binhlthTab <- createTTT(binhlth)
-binhlthTab
-```
-### Binary Economic
-```{r echo = FALSE}
-# Create the actual table!
-bineconTab <- createTTT(binecon)
-bineconTab
-```
-### Risk Behaviours
-```{r echo = FALSE}
-# Create the actual table!
-riskTab <- createTTT(risk)
-riskTab
-```
-
-# Cross Validation 2 - 'Handover' plots (Still WIP need fixing)
-
-```{r echo = FALSE}
-vis_CV2_prev <- function(long, CV2) {
-
-  # Keep only waves 1 - 4
-  longC <- subset(long, wave < 5)
-  longC <- as.data.table(longC)
-
-  # Replace all NA with 0
-  longC[is.na(longC)] <- 0
-
-  # Have to drop r*clust vars as they contain strings
-  longC <- select(longC, -contains('clust'))
-
-  # Now get the weighted mean by wave of every column based on cwtresp
-  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]
-
-  # ELSA_long is now processed to the point where we can start extracting and combining variables
-  # Better start by adding year to l2 and reordering the columns
-  l2$year <- seq(from=2002, to = 2008, by = 2)
-  setcolorder(l2, c('year', 'wave'))
-
-  # Now lets extract chronic disease variables from both dataframes and do a row combine
-  # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl)) #, anyiadl, adl1, adl2, adl3p))
-  cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
-                            p_hearte_all, p_lunge_all, p_smoken_all, 
-                            a_bmi_all, p_employed_all,  p_anyadl_all))
-                            #, p_anyiadl_all p_adl1_all, p_adl2_all, p_adl3p_all))
-
-  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_all,
-                                        'cancre' = p_cancre_all, 
-                                        'diabe' = p_diabe_all, 
-                                        'hearte' = p_hearte_all,
-                                        'lunge' = p_lunge_all,
-                                        'smoken' = p_smoken_all,
-                                        'bmi' = a_bmi_all,
-                                        'employed' = p_employed_all,
-                                        'anyadl' = p_anyadl_all)
-                                        #'anyiadl' = p_anyiadl_all)
-                                        #'adl1' = p_adl1_all,
-                                        #'adl2' = p_adl2_all,
-                                        #'adl3p' = p_adl3p_all)
-
-  combine <- rbind(cd_l2, CV2_rename)
-
-  melted <- reshape2::melt(combine, id.var = 'year', value.name = 'Prevalence')
-
-  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
-          geom_line() +
-          facet_wrap(. ~ variable, scales = "free")
-          #theme(panel.spacing.x = unit(1.5, 'lines')) +
-          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
-}
-
-# Load in whole stock pop, plus the CV2 summary output file
-long <- read_dta('../input_data/ELSA_long.dta')
-CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
-
-vis_CV2_prev(long, CV2)
-```
-
-# Intervention Scenarios
-
-### Base Initial Prevalence of Chronic Disease
-
-```{r echo = FALSE}
-visInitBasePrev <- function(base) {
-
-  cdMelted <- base %>%
-                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
-                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
-                filter(year == 2012) %>%
-                reshape2::melt(id.var = 'year', value.name = 'Prevalence') %>%
-                subset(select = -year)
-
-  ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
-          geom_bar(stat = 'identity') +
-          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')
-}
-
-baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-
-visInitBasePrev(baseline)
-```
-
-### Baseline Projections
-
-```{r echo = FALSE}
-visProjectedBaselineCD <- function(base) {
-
-  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
-  cdMelted <- base %>%
-                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
-                dplyr::rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
-                reshape2::melt(id.var='year', value.name = 'Prevalence')
-
-  # Create the tiled plot of disease prevalence
-  ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
-            geom_line() +
-            facet_grid(. ~ variable) +
-            theme(panel.spacing.x = unit(1.5, 'lines')) +
-            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
-}
-
-baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-
-visProjectedBaselineCD(baseline)
-```
-
-### Survival Curve
-
-```{r echo = FALSE}
-cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
-
-cohort$age <- (cohort$year - 2012) + 50
-
-cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
-cohort$survival_m <- cohort$m_endpop_m / cohort$m_endpop_m[1]
-cohort$survival_f <- cohort$m_endpop_f / cohort$m_endpop_f[1]
-
-# plot up survival curve
-ggplot(cohort, aes(x=age, y=survival)) +
-        geom_line(col='black', size=0.3) +
-        labs(title='Cohort Survival: All', y='Survival', x='Age')
 ```
@@
 'The Age UK almanac of disease profiles in later life', published in October 2015, contains information on the prevalence of major diseases, conditions, and syndromes affecting older people in England, broken down by age and sex.
+
+```{r}
+# Just read in the data in this chunk, then don't have to have everything in the environment all the time
+#AgeUK
+mAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
+fAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+
+# FEM 2014 detailed output
+FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+## extract males and females
+mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+
+# FEM 2014 summary
+FEM_sum <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+```
+
+```{r}
+# Rename columns
+colnames(mAGE_UK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
+colnames(fAGE_UK) <- colnames(mAGE_UK)
+
+# Collect conditions into a list for extracting from FEM
+# Will need to be written in FEM shorthand
+keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+
+m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
+f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
+
+m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
+f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
+colnames(m6064_prev) <- 'FEM_male_6064'
+colnames(f6064_prev) <- 'FEM_female_6064'
+
+prev_df <- data.frame(m6064_prev, f6064_prev)
+
+```
+
+```{r}
+male_au <- mAGE_UK
+female_au <- fAGE_UK
+
+#mFEM <- mFEM_2014
+#fFEM <- fFEM_2014
+
+FEM <- FEM_sum
+
+ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+
+m_ages <- paste0('m', ages)
+
+m_ages
+```
+
+## First Try
 
@@
 ```
-
-# Github Reporting
-
-It would be nice if we could copy paste tables directly from this document into a github issue.
-For this reason, we will look into a package that can print tables directly into markdown or something similar, which we can then place straight into a Github issue.
-
-```{r}
-
-#kable
-
-```

### 2021-02-02 FEM_R/validation.Rmd
WIP validation notebook

@@
 # FEM 2014 detailed output
-FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
 ## extract males and females
-mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
 
@@
 ```{r}
-# Rename columns
-colnames(mAGE_UK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
-colnames(fAGE_UK) <- colnames(mAGE_UK)
+# Take copy of FEM summary
+FEM <- FEM_sum
 
-# Collect conditions into a list for extracting from FEM
-# Will need to be written in FEM shorthand
-keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+# Keep only year 2014 from FEM summary
+FEM <- filter(FEM, year == 2014)
 
-m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
-f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
+# Specify the age groups to extract from FEM
+ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# Add m and f for male and female variables
+m_ages <- paste0('m_', ages)
+f_ages <- paste0('f_', ages)
 
-m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
-f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
-colnames(m6064_prev) <- 'FEM_male_6064'
-colnames(f6064_prev) <- 'FEM_female_6064'
+# Collect the age suffixes into a vector
+keep_vars <- c(m_ages, f_ages)
+# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
 
-prev_df <- data.frame(m6064_prev, f6064_prev)
+# Don't like using loops but this one was just too simple to bother with figuring out the apply version
+final <- c()
+ticker <- 1
+# Generating the column (variable) names we want to keep in a loop
+for(item in diseases) {
+  for(anotherItem in keep_vars) {
+    final[ticker] <- paste0('p_', item, '_', anotherItem)
+    ticker <- ticker + 1
+  }
+}
+
+# Keep the right vars using varnames generated in loop
+FEM <- FEM %>%
+          select(c('year', all_of(final)))
+
+# Rename some vars to remove 'p_' before merging
+FEM <- FEM %>%
+          rename_at(vars(starts_with('p_')),
+                    funs(str_replace(., 'p_', ''))) %>%
+          select(!year)
+
+pivoted.FEM <- pivot_longer(data = FEM,
+                            cols = hearte_m_6064:demene_f_100p, 
+                            names_to = c('Condition', 'Gender', 'Age_Group'), 
+                            names_sep = '_')
+
+# Convert value from fraction to percentage
+pivoted.FEM$value <- pivoted.FEM$value * 100
+
+```
+
+Now process the AgeUK data and combine male and female
+
+```{r}
+# Copy over age uk data
+male_au <- mAGE_UK
+female_au <- fAGE_UK
+
+# Rename cols to enable reshaping
+male_au <- male_au %>%
+                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
+                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
+                       m_9094 = X90.94, m_9599 = X95.99, m_100 = Over.100) %>%
+                select(!Total)
+
+female_au <- female_au %>%
+                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
+                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
+                       f_9094 = X90.94, f_9599 = X95.99, f_100 = Over.100) %>%
+                select(!Total)
+
+# Now need to reshape these datasets to long format
+pivoted.male <- pivot_longer(data = male_au,
+                             cols = m_6064:m_100,
+                             names_to = c('Gender', 'Age_Group'),
+                             names_sep = '_')
+
+pivoted.female <- pivot_longer(data = female_au,
+                             cols = f_6064:f_100,
+                             names_to = c('Gender', 'Age_Group'),
+                             names_sep = '_')
+
+# Now row bind these frames together, combining male and female data
+ageUK <-rbind(pivoted.male, pivoted.female)
+
+# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# extract all heart related conditions for combining
+ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# Aggregate the values by gender and age group
+ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# Rename columns and add condition column back
+colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+ageUK_heart$Condition <- 'Heart Problems'
+
+# Now drop the heart problems from original ageUK and replace with aggregate
+ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+ageUK <- rbind(ageUK, ageUK_heart)
+
+# Rename a few just for completeness
+ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+
+# Finally, add column to specify that this is ageUK data
+ageUK$source <- "ageUK"
 
@@
 ```{r}
-male_au <- mAGE_UK
-female_au <- fAGE_UK
 
-#mFEM <- mFEM_2014
-#fFEM <- fFEM_2014
+# COME ON!!!!!!
+# Steps:
+#   - Rename FEM rows to match AgeUK
+#   - Add column to specify between FEM and AgeUK
+#   - Merge datasets, maintain long format
+#   - ggplot2 some nice grouped bar charts
+#   - Also think about some nice reactable tables maybe? With some custom styling
 
-FEM <- FEM_sum
+# Make copy
+newFEM <- pivoted.FEM
 
-ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# Rename all the conditions to fit ageUK
+# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+#       Therefore only validation from this will be that FEM is higher than ageUK
+newFEM$Condition <- gsub("hearte", "Hearte Problems", newFEM$Condition)
+newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
 
-m_ages <- paste0('m', ages)
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
 
-m_ages
+```
+
+```{r}
+
+# Now row bind the datasets
+combined <- rbind(newFEM, ageUK)
+# Sort
+combined <- arrange(combined, source, Condition, Gender, Age_Group)
+
+```
+
+```{r}
+
+# Lets do some plotting!
+
 ```
@@
 ```
+
+```{r}
+# Rename columns
+colnames(mAGE_UK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
+colnames(fAGE_UK) <- colnames(mAGE_UK)
+
+# Collect conditions into a list for extracting from FEM
+# Will need to be written in FEM shorthand
+keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+
+m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
+f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
+
+m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
+f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
+colnames(m6064_prev) <- 'FEM_male_6064'
+colnames(f6064_prev) <- 'FEM_female_6064'
+
+prev_df <- data.frame(m6064_prev, f6064_prev)
+
+```

### 2021-02-02 FEM_R/validation.Rmd
Removed code from first try

@@
 ```
-
-## First Try
-
-```{r}
-# Need a function here to handle the repetition
-get_FEM_prev <- function(FEM, age_range, keep_vars) {
-  # Keep only important vars
-  FEM <- FEM[keep_vars]
-  # Separate by gender
-  mFEM <- FEM %>% filter(male == 1)
-  fFEM <- FEM %>% filter(male == 0)
-  # Filter by age from age_range
-  mFEM <- mFEM %>% filter(age %in% (age_range))
-  fFEM <- fFEM %>% filter(age %in% (age_range))
-  # Get proportion of total pop. (prevalence)
-  mFEM <- data.frame((colSums(mFEM) / nrow(mFEM)) * 100)
-  fFEM <- data.frame((colSums(fFEM) / nrow(fFEM)) * 100)
-
-  # produce a string for column naming
-  age <- paste0(age_range[1], '-', age_range[5])
-
-  colnames(mFEM) <- paste0('FEM_male_', age)
-  colnames(fFEM) <- paste0('FEM_female_', age)
-
-  #male <- cbind(male, mFEM)
-
-  # Collect into 1 df
-  prev_df <- data.frame(mFEM, fFEM)
-  return(prev_df)
-
-  #return(out)
-}
-
-package <- function() {
-  # Set up objects for holding and returning the prevalences, separate dfs for male and female
-  out <- list()
-  male <- data.frame()
-  female <- data.frame()
-
-  age_list <- c('Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+')
-  age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)
-}
-
-# Now read in male and female AgeUK data
-male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
-female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
-# Rename columns
-colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
-colnames(female_ageUK) <- colnames(male_ageUK)
-
-# Collect conditions into a list for extracting from FEM
-# Will need to be written in FEM shorthand
-keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
-
-# Now get FEM output for 2014
-FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-
-test <- get_FEM_prev(FEM_2014, 60:64, keep_vars)
-
-#FEM_2014 <- FEM_2014[keep_vars]
-## extract males and females
-mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-fFEM_2014 <- FEM_2014 %>% filter(male == 0)
-
-m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
-f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
-
-m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
-f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
-colnames(m6064_prev) <- 'FEM_male_6064'
-colnames(f6064_prev) <- 'FEM_female_6064'
-
-prev_df <- data.frame(m6064_prev, f6064_prev)
-
-# AgeUK conditions to extract
-#condlist <- c('Dementia', 'Hypertension', 'Diabetes', 'Stroke')
-
-# correct colnames in ageUK stuff
-#male_ageUK <- male_ageUK %>% filter(Condition == 'Dementia' | Condition == 'Hypertension' | Condition == 'Diabetes' | Condition == 'Stroke')
-#male_ageUK <- male_ageUK %>% filter(Condition %in% condlist)
-
-# Now calculate prevalence by age group for each chronic disease
-# prevalence = total / affected
-# Diabetes
-#male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
-#female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
-#colnames(male_prev) <- 'FEM_male'
-#colnames(female_prev) <- 'FEM_female'
-
-```
-
-```{r}
-
-tmp <- function(m, f, age) {
-
-  m_age <- m %>% filter(age %in% (age))
-  f_age <- f %>% filter(age %in% (age))
-
-  m_prev <- data.frame((colSums(m_age) / nrow(m_age)) * 100)
-  f_prev <- data.frame((colSums(f_age) / nrow(f_age)) * 100)
-
-  list <- list(m_prev, f_prev)
-  return(list)
-}
-
-# Now read in male and female AgeUK data
-male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
-female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
-# Rename columns
-colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
-colnames(female_ageUK) <- colnames(male_ageUK)
-
-# Collect conditions into a list for extracting from FEM
-# Will need to be written in FEM shorthand
-keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
-
-# Now get FEM output for 2014
-FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-FEM <- FEM_2014[keep_vars]
-
-# extract males and females
-mFEM <- FEM %>% filter(male == 1)
-fFEM <- FEM %>% filter(male == 0)
-
-age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)
-
-m6064 <- mFEM %>% filter(age %in% (60:64))
-f6064 <- fFEM %>% filter(age %in% (60:64))
-
-m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
-f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
-colnames(m6064_prev) <- 'FEM_male_6064'
-colnames(f6064_prev) <- 'FEM_female_6064'
-
-male <- list()
-female <- list()
-
-mf_list <- tmp(mFEM, fFEM, 60:64)
-male <- list(male, mf_list[1])
-female <- list(female, mf_list[2])
-
-mf_list <- tmp(mFEM, fFEM, 65:69)
-male <- list(male, mf_list[1])
-female <- list(female, mf_list[2])
-
-```
-
-```{r}
-# Rename columns
-colnames(mAGE_UK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
-colnames(fAGE_UK) <- colnames(mAGE_UK)
-
-# Collect conditions into a list for extracting from FEM
-# Will need to be written in FEM shorthand
-keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
-
-m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
-f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
-
-m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
-f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
-colnames(m6064_prev) <- 'FEM_male_6064'
-colnames(f6064_prev) <- 'FEM_female_6064'
-
-prev_df <- data.frame(m6064_prev, f6064_prev)
-
-```

### 2021-02-02 FEM_R/validation.Rmd
WIP - mostly done but trying to figure out the best way to show all the plots together

@@
                        m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
-                       m_9094 = X90.94, m_9599 = X95.99, m_100 = Over.100) %>%
+                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
                 select(!Total)
@@
                        f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
-                       f_9094 = X90.94, f_9599 = X95.99, f_100 = Over.100) %>%
+                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
                 select(!Total)
@@
 pivoted.male <- pivot_longer(data = male_au,
-                             cols = m_6064:m_100,
+                             cols = m_6064:m_100p,
                              names_to = c('Gender', 'Age_Group'),
@@
 pivoted.female <- pivot_longer(data = female_au,
-                             cols = f_6064:f_100,
+                             cols = f_6064:f_100p,
                              names_to = c('Gender', 'Age_Group'),
@@
 
+# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+ageUK <- filter(ageUK, Condition != 'Asthma', Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+
 ```
@@
 #       Therefore only validation from this will be that FEM is higher than ageUK
-newFEM$Condition <- gsub("hearte", "Hearte Problems", newFEM$Condition)
+newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
 newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
@@
 combined <- rbind(newFEM, ageUK)
+# Specify factor levels for age group
+combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
 # Sort
@@
 
+# First try to plot the whole dataset in a facet grid
+# This isn't perfect, and it's pretty complicated to plot altogether
+ggplot(data = combined, aes(x = Age_Group, y = value, width = 0.7)) +
+  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+  facet_wrap(~Condition)
+
 ```
+
+```{r}
+
+# Now instead, lets plot an individual condition - Hypertension
+
+# First extract all hypertension info
+hyper <- filter(combined, Condition == 'Hypertension')
+
+# Big dirty plot
+ggplot(data = hyper, aes(x = Age_Group, y = value, width = 0.7)) +
+  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+  labs(title = 'Hypertension', x = 'Age Group', y = 'Proportion')
+
+```
+
+```{r}
+
+# Next, lets try an all female and all male example
+
+```

### 2021-02-03 FEM_R/model_analysis.Rmd
Added validation visualisations to model_analysis.Rmd

--- 2021-01-11 FEM_R/model_analysis.Rmd
@@
 
-```{r}
-# Need a function here to handle the repetition
-get_prevalence <- function(FEM) {
+```{r echo = FALSE}
 
+#AgeUK
+male_au <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
+female_au <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+
+# FEM 2014 detailed output
+#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+## extract males and females
+#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+
+# FEM 2014 summary
+FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Keep only year 2014 from FEM summary
+FEM <- filter(FEM, year == 2014)
+
+# Specify the age groups to extract from FEM
+ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# Add m and f for male and female variables
+m_ages <- paste0('m_', ages)
+f_ages <- paste0('f_', ages)
+
+# Collect the age suffixes into a vector
+keep_vars <- c(m_ages, f_ages)
+# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+
+# Don't like using loops but this one was just too simple to bother with figuring out the apply version
+final <- c()
+ticker <- 1
+# Generating the column (variable) names we want to keep in a loop
+for(item in diseases) {
+  for(anotherItem in keep_vars) {
+    final[ticker] <- paste0('p_', item, '_', anotherItem)
+    ticker <- ticker + 1
+  }
 }
 
-# Now read in male and female AgeUK data
-male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
-female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
-# Rename columns
-colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
-colnames(female_ageUK) <- colnames(male_ageUK)
+# Keep the right vars using varnames generated in loop
+FEM <- FEM %>%
+          select(c('year', all_of(final)))
 
-# Collect conditions into a list for extracting from FEM
-# Will need to be written in FEM shorthand
-keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+# Rename some vars to remove 'p_' before merging
+FEM <- FEM %>%
+          rename_at(vars(starts_with('p_')),
+                    funs(str_replace(., 'p_', ''))) %>%
+          select(!year)
 
-# Now get FEM output for 2014
-FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-FEM_2014 <- FEM_2014[keep_vars]
-# extract males and females
-mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+pivoted.FEM <- pivot_longer(data = FEM,
+                            cols = hearte_m_6064:demene_f_100p, 
+                            names_to = c('Condition', 'Gender', 'Age_Group'), 
+                            names_sep = '_')
 
-m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
-f6064 <- fFEM_2014 %>% filter(age %in% (60:64))
+# Convert value from fraction to percentage
+pivoted.FEM$value <- pivoted.FEM$value * 100
 
-m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
-f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
-colnames(m6064_prev) <- 'FEM_male_6064'
-colnames(f6064_prev) <- 'FEM_female_6064'
+# Rename cols to enable reshaping
+male_au <- male_au %>%
+                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
+                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
+                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
+                select(!Total)
 
-prev_df <- data.frame(m6064_prev, f6064_prev)
+female_au <- female_au %>%
+                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
+                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
+                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
+                select(!Total)
 
-# correct colnames in ageUK stuff
-male_ageUK <- male_ageUK %>% filter(conditions == c('Hypertension', 'Stroke', 'Diabetes', 'Dementia'))
+# Now need to reshape these datasets to long format
+pivoted.male <- pivot_longer(data = male_au,
+                             cols = m_6064:m_100p,
+                             names_to = c('Gender', 'Age_Group'),
+                             names_sep = '_')
 
-# Now calculate prevalence by age group for each chronic disease
-# prevalence = total / affected
-# Diabetes
-male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
-female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
-colnames(male_prev) <- 'FEM_male'
-colnames(female_prev) <- 'FEM_female'
+pivoted.female <- pivot_longer(data = female_au,
+                             cols = f_6064:f_100p,
+                             names_to = c('Gender', 'Age_Group'),
+                             names_sep = '_')
+
+# Now row bind these frames together, combining male and female data
+ageUK <-rbind(pivoted.male, pivoted.female)
+
+# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# extract all heart related conditions for combining
+ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# Aggregate the values by gender and age group
+ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# Rename columns and add condition column back
+colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+ageUK_heart$Condition <- 'Heart Problems'
+
+# Now drop the heart problems from original ageUK and replace with aggregate
+ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+ageUK <- rbind(ageUK, ageUK_heart)
+
+# Rename a few just for completeness
+ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+
+# Finally, add column to specify that this is ageUK data
+ageUK$source <- "ageUK"
+
+# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+ageUK <- filter(ageUK, Condition != 'Asthma', Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+
+# Make copy
+newFEM <- pivoted.FEM
+
+# Rename all the conditions to fit ageUK
+# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+#       Therefore only validation from this will be that FEM is higher than ageUK
+newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+
+# Now row bind the datasets
+combined <- rbind(newFEM, ageUK)
+# Specify factor levels for age group
+combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
+# Sort
+combined <- arrange(combined, source, Condition, Gender, Age_Group)
+
+# Function to filter the combined dataset and visualise
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {
+
+  if(gender == 'm') {
+    full_gender <- 'Male'
+  }
+  else if(gender == 'f') {
+    full_gender <- 'Female'
+  }
+
+  # filter by gender and condition
+  data <- data %>%
+            filter(Gender == gender) %>%
+            filter(Condition == cd)
+
+  # now plot!
+  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
+          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+
+  # Save plot into in E-FEM_Paper/figures/
+  ggsave(filename = paste0(cd, '_', gender, '.png'),
+         plot = p,
+         path = path,
+         width = 7,
+         height = 5)
+
+  p
+}
+
+# Collect all conditions (and both genders) into a list, and apply the function above to that list
+cd_list <- unique(combined$Condition)
+gender_list <- unique(combined$Gender)
+
+for(cd in cd_list) {
+  for(gen in gender_list) {
+    visAgeUKValid(combined, cd, gen)
+  }
+}
 

### 2021-02-03 FEM_R/validation.Rmd
Added validation visualisations to model_analysis.Rmd

@@
 # Next, lets try an all female and all male example
+comb <- combined
+
+# extract genders
+male <- filter(combined, Gender == 'm')
+female <- filter(combined, Gender == 'f')
+
+# try plotting each separately
+
+ggplot(data = male, aes(x = Age_Group, y = value, width = 0.7)) +
+  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+  facet_wrap(~Condition) +
+  labs(title = 'Male')
+
+ggplot(data = female, aes(x = Age_Group, y = value, width = 0.7)) +
+  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+  facet_wrap(~Condition) +
+  labs(title = 'Female')
 
 ```
+
+```{r}
+
+# Function to filter the combined dataset and visualise
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {
+
+  if(gender == 'm') {
+    full_gender <- 'Male'
+  }
+  else if(gender == 'f') {
+    full_gender <- 'Female'
+  }
+
+  # filter by gender and condition
+  data <- data %>%
+            filter(Gender == gender) %>%
+            filter(Condition == cd)
+
+  # now plot!
+  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
+          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+
+  # Save plot into in E-FEM_Paper/figures/
+  ggsave(filename = paste0(cd, '_', gender, '.png'),
+         plot = p,
+         path = path,
+         width = 7,
+         height = 5)
+
+  p
+}
+
+visAgeUKValid(combined, 'Hypertension', 'm')
+
+```
+
+```{r}
+
+# Collect all conditions (and both genders) into a list, and apply the function above to that list
+cd_list <- unique(combined$Condition)
+gender_list <- unique(combined$Gender)
+
+for(cd in cd_list) {
+  for(gen in gender_list) {
+    visAgeUKValid(combined, cd, gen)
+  }
+}
+
+```

### 2021-02-03 FEM_R/validation.Rmd
Updated the validation to include Asthma

@@
 # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')
+diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene', 'asthmae')
 
@@
 pivoted.FEM <- pivot_longer(data = FEM,
-                            cols = hearte_m_6064:demene_f_100p, 
+                            cols = hearte_m_6064:asthmae_f_100p, 
                             names_to = c('Condition', 'Gender', 'Age_Group'), 
@@
 #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
-ageUK <- filter(ageUK, Condition != 'Asthma', Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
 
@@
 newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
 

### 2021-02-08 FEM_R/model_analysis.Rmd
Added ageUK validation to model_analysis.Rmd and changed the size of the images being saved

@@
 # FEM 2014 summary
-FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+FEM <- read_dta('ELSA_AgeUK_valid/ELSA_AgeUK_valid_summary.dta')
 
@@
          path = path,
-         width = 7,
-         height = 5)
+         width = 6,
+         height = 4)
 
-  p
+  return(p)
 }
 
+```
+
+```{r}
 # Collect all conditions (and both genders) into a list, and apply the function above to that list
@@
   for(gen in gender_list) {
-    visAgeUKValid(combined, cd, gen)
+    plot <- visAgeUKValid(combined, cd, gen)
+    print(plot)
   }
 }
-
 ```

### 2021-02-17 FEM_R/model_analysis.Rmd
Added lots more validation to RNotebook file, and added some of these to model_analysis.Rmd

@@
 
+Second go at this cross-validation, attempting to solve some issues by using the 60 plus population.
+
+```{r echo = FALSE}
+vis_CV2_prev <- function(long, CV2) {
+
+  # Keep only waves 1 - 4
+  longC <- subset(long, wave < 5)
+  longC <- as.data.table(longC)
+
+  # Replace all NA with 0
+  longC[is.na(longC)] <- 0
+
+  # Have to drop r*clust vars as they contain strings
+  longC <- select(longC, -contains('clust'))
+
+  # Now get the weighted mean by wave of every column based on cwtresp
+  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]
+
+  # ELSA_long is now processed to the point where we can start extracting and combining variables
+  # Better start by adding year to l2 and reordering the columns
+  l2$year <- seq(from=2002, to = 2008, by = 2)
+  setcolorder(l2, c('year', 'wave'))
+
+  # Now lets extract chronic disease variables from both dataframes and do a row combine
+  # Start with cancre, diabe, hearte
+  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl)) #, anyiadl, adl1, adl2, adl3p))
+  cd_CV2 <- select(CV2, c(year, a_age_60p, p_cancre_60p, p_diabe_60p, 
+                            p_hearte_60p, p_lunge_60p, p_smoken_60p, 
+                            a_bmi_60p, p_employed_60p,  p_anyadl_60p))
+                            #, p_anyiadl_60p p_adl1_60p, p_adl2_60p, p_adl3p_60p))
+
+  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_60p,
+                                        'cancre' = p_cancre_60p, 
+                                        'diabe' = p_diabe_60p, 
+                                        'hearte' = p_hearte_60p,
+                                        'lunge' = p_lunge_60p,
+                                        'smoken' = p_smoken_60p,
+                                        'bmi' = a_bmi_60p,
+                                        'employed' = p_employed_60p,
+                                        'anyadl' = p_anyadl_60p)
+                                        #'anyiadl' = p_anyiadl_60p)
+                                        #'adl1' = p_adl1_60p,
+                                        #'adl2' = p_adl2_60p,
+                                        #'adl3p' = p_adl3p_60p)
+
+  combine <- rbind(cd_l2, CV2_rename)
+
+  melted <- reshape2::melt(combine, id.var = 'year', value.name = 'Prevalence')
+
+  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
+          geom_line() +
+          facet_wrap(. ~ variable, scales = "free")
+          #theme(panel.spacing.x = unit(1.5, 'lines')) +
+          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
+}
+
+# Load in whole stock pop, plus the CV2 summary output file
+long <- read_dta('../input_data/ELSA_long.dta')
+CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
+
+vis_CV2_prev(long, CV2)
+```
+
 # Intervention Scenarios
@@
 
+# Smoking
+
+```{r}
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Smoking prevalence stats
+smok.5059 <- 15.2
+smok.60p <- 10.2
+
+# FEM summary, filter for 2018 and 60p
+#FEM.smok <- FEM_sum
+FEM.smok <- FEM_base
+FEM.smok <- FEM.smok %>%
+                filter(year == 2018) %>%
+                select(contains('smoken')) %>%
+                select(contains('60p'), contains('5059')) %>%
+                select(contains('p_'))
+
+FEM.smok <- FEM.smok * 100
+
+smok.combined <- data.frame(FEM = FEM.smok[1,1], AgeUK = 10.2)
+smok.combined <- data.frame(source = c('FEM', 'FEM', 'Observed', 'Observed'), age = c('50 - 59', '60+', '50 - 59', '60+'), value = c(FEM.smok[1,2], FEM.smok[1,1], 15.2, 10.2))
+
+smok.plot <-ggplot(data = smok.combined, aes(x = age, y = value)) +
+                geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+                labs(title = 'Comparison of smoking prevalence in the elderly in 2018', x = ' ', y = 'Prevalence (%)')
+
+ggsave(filename = 'smoking_prev_2018.png',
+       plot = smok.plot,
+       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/',
+       width = 7,
+       height = 5)
+
+smok.plot
+
+```
+
 # Github Reporting

### 2021-02-17 FEM_R/validation.Rmd
Added lots more validation to RNotebook file, and added some of these to model_analysis.Rmd

@@
 #require(plyr)
+require(readODS)
 
@@
 # FEM 2014 summary
-FEM_sum <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+FEM_sum <- read_dta('ELSA_AgeUK_valid/ELSA_AgeUK_valid_summary.dta')
 
@@
   # Save plot into in E-FEM_Paper/figures/
-  ggsave(filename = paste0(cd, '_', gender, '.png'),
-         plot = p,
-         path = path,
-         width = 7,
-         height = 5)
+  #ggsave(filename = paste0(cd, '_', gender, '.png'),
+  #       plot = p,
+  #       path = path,
+  #       width = 7,
+  #       height = 5)
 
@@
 ```
+
+# Smoking
+
+## Action on Smoking and Health
+
+```{r}
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Smoking prevalence stats
+# Stats found here: https://ash.org.uk/information-and-resources/fact-sheets/smoking-statistics-who-smokes-and-how-much/
+# Page 3, table 4
+smok.5059 <- 15.2
+smok.60p <- 10.2
+
+# FEM summary, filter for 2018 and 60p
+#FEM.smok <- FEM_sum
+FEM.smok <- FEM_base
+FEM.smok <- FEM.smok %>%
+                filter(year == 2018) %>%
+                select(contains('smoken')) %>%
+                select(contains('60p'), contains('5059')) %>%
+                select(contains('p_'))
+
+FEM.smok <- FEM.smok * 100
+
+smok.combined <- data.frame(FEM = FEM.smok[1,1], AgeUK = 10.2)
+smok.combined <- data.frame(source = c('FEM', 'FEM', 'Observed', 'Observed'), age = c('50 - 59', '60+', '50 - 59', '60+'), value = c(FEM.smok[1,2], FEM.smok[1,1], 15.2, 10.2))
+
+smok.plot <-ggplot(data = smok.combined, aes(x = age, y = value)) +
+                geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+                labs(title = 'Comparison of smoking prevalence in the elderly in 2018', x = ' ', y = 'Prevalence (%)')
+
+ggsave(filename = 'smoking_prev_2018.png',
+       plot = smok.plot,
+       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/',
+       width = 7,
+       height = 5)
+
+smok.plot
+
+```
+
+## ONS
+
+Found some more data on smoking in the UK from the ONS, perhaps more reliable than the other? Might be a good idea to show both if they are all within a couple of percent.
+
+```{r}
+# FEM
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# ONS - Male
+m.ons.smok <- read_ods(path = '/home/luke/Documents/E_FEM_clean/Validation/adultsmokinghabitsinengland2019final.ods',
+                sheet = 4,
+                skip = 8,
+                col_names = TRUE,
+                row_names = FALSE,
+                range = "E1:F22")
+# ONS - Female
+f.ons.smok <- read_ods(path = '/home/luke/Documents/E_FEM_clean/Validation/adultsmokinghabitsinengland2019final.ods',
+                sheet = 4,
+                skip = 8,
+                col_names = TRUE,
+                row_names = FALSE,
+                range = "L1:M22")
+# ONS - All
+all.ons.smok <- read_ods(path = '/home/luke/Documents/E_FEM_clean/Validation/adultsmokinghabitsinengland2019final.ods',
+                sheet = 4,
+                skip = 8,
+                col_names = TRUE,
+                row_names = FALSE,
+                range = "S1:T22")
+
+# Remove empty row
+m.ons.smok <- m.ons.smok %>% slice(-1)
+f.ons.smok <- f.ons.smok %>% slice(-1)
+all.ons.smok <- all.ons.smok %>% slice(-1)
+
+# Add year column
+m.ons.smok$Year <- c(2000:2019)
+f.ons.smok$Year <- c(2000:2019)
+all.ons.smok$Year <- c(2000:2019)
+
+# Add Gender col
+m.ons.smok$Gender <- 'Male'
+f.ons.smok$Gender <- 'Female'
+all.ons.smok$Gender <- 'All'
+
+# Rearrange columns before rbind
+m.ons.smok <- subset(m.ons.smok, select = c(3,4,1,2))
+f.ons.smok <- subset(f.ons.smok, select = c(3,4,1,2))
+all.ons.smok <- subset(all.ons.smok, select = c(3,4,1,2))
+
+# Filter out years before 2012
+m.ons.smok <- dplyr::filter(m.ons.smok, Year >= 2012)
+f.ons.smok <- dplyr::filter(f.ons.smok, Year >= 2012)
+all.ons.smok <- dplyr::filter(all.ons.smok, Year >= 2012)
+
+# Combine male, female, all and rename 60 plus column
+ons.smok <- rbind(m.ons.smok, f.ons.smok, all.ons.smok)
+ons.smok <- dplyr::rename(ons.smok, '60p' = '60 and over')
+
+# Reshape data to long
+pivot.ons.smok <- pivot_longer(data = ons.smok,
+                               cols = '50-59':'60p',
+                               names_to = 'Age',
+                               values_to = 'Value')
+
+pivot.ons.smok$Source <- 'ONS'
+
+# ONS data wrangled, now get FEM into shape
+FEM.smok <- FEM_base
+# Get year out to put back later
+Year <- FEM.smok$year
+FEM.smok <- FEM.smok %>%
+                select(contains('smoken')) %>%
+                select(contains('p_'), -starts_with('n_')) %>%
+                select(contains('60p'), contains('5059')) %>%
+                dplyr::rename('60p' = p_smoken_60p, '50-59' = p_smoken_5059)
+
+FEM.smok <- subset(FEM.smok, select = c(2, 1))
+FEM.smok <- cbind(Year, FEM.smok)
+
+FEM.smok <- filter(FEM.smok, year >= 2012, year < 2022)
+
+pivot.FEM.smok <- pivot_longer(data = FEM.smok,
+                               cols = '50-59':'60p',
+                               names_to = 'Age',
+                               values_to = 'Value')
+pivot.FEM.smok$Source <- 'FEM'
+pivot.FEM.smok$Value <- pivot.FEM.smok$Value * 100
+
+pivot.ons.smok.all <- pivot.ons.smok %>%
+                          filter(Gender == 'All') %>%
+                          select(-Gender)
+
+combined <- rbind(pivot.FEM.smok, pivot.ons.smok.all)
+
+ggplot(data = combined, aes(x = Year, y = Value)) +
+  geom_bar(aes(fill = Source), stat = 'identity', position = 'dodge')
+
+```
+
+# ADLs
+
+Just having a play around with this data, wanted to visualise how levels of anyADL change over time.
+
+```{r}
+
+FEM.new <- FEM_base
+
+FEM.adl <- FEM.new %>%
+              select(contains('year'), contains('anyadl')) %>%
+              select(-starts_with('i_'), -starts_with('n_'), -starts_with('a_')) %>%
+              select(-contains('_m'), -contains('_f')) %>%
+              select(-contains('5059'), -contains('60p'))
+
+p.FEM.adl <- pivot_longer(data = FEM.adl,
+                          cols = p_anyadl_100p:p_anyadl_all,
+                          names_to = c('Prevalence', 'Condition', 'Age'),
+                          names_sep = '_')
+
+p.FEM.adl <- select(p.FEM.adl, -Prevalence, -Condition)
+
+p.FEM.adl$Age <- factor(p.FEM.adl$Age, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p', 'all'))
+
+ggplot(data = p.FEM.adl, aes(x = year, y = value, colour = Age)) +
+  geom_line(aes(fill = Age))
+
+```
+
+# ADL Comparison (Wittenberg)
+
+Data for this comparison came from [this paper](https://people.uea.ac.uk/en/publications/projections-of-demand-and-expenditure-on-adult-social-care-2015-to-2040(d1f44389-4e52-4cb1-9b66-ecc7594d345e).html) - Projections of Demand and Expenditure on Adult Social Care 2015 to 2040.
+On page 6, base case projections are outlined.
+Wittenberg says that from 2015 to 2040, the population having at least 1 difficulty in ADL or IADL would rise by 67%, and from 2015 to 2070 would rise by 116%.
+When considering only ADLs, the increase was 69% from 2015 to 2040, and 124% from 2015 to 2070.
+Written more clearly:
+
+ADL & IADL
+2015 - 2040: 67% increase
+2015 - 2070: 116% increase
+
+ADL only
+2015 - 2040: 69% increase
+2015 - 2070: 124% increase
+
+```{r}
+
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Get FEM output and extract only p_anyadl_all
+FEM.adl <- FEM_base %>%
+              select(contains('year'), contains('anyadl')) %>%
+              select(-starts_with('i_'), -starts_with('n_'), -starts_with('a_')) %>%
+              select(-contains('_m'), -contains('_f')) %>%
+              select(contains('year'), contains('_all'))
+
+# Now calculate % increase between 2015 & 2040. Need to take average between 2014 & 2016 values
+FEM.2015 <- data.frame(year = 2015, ((FEM.adl[2,2] + FEM.adl[3,2]) / 2))
+# extract 2040
+FEM.2040 <- filter(FEM.adl, year == 2040)
+# Row bind
+FEM.2015_2040 <- rbind(FEM.2015, FEM.2040)
+
+# Now calculate percentage change
+FEM.adl.perchange <- ((FEM.2015_2040[2,2] - FEM.2015_2040[1,2]) / FEM.2015_2040[2,2]) * 100
+
+# Not good! Showing an 8% reduction in anyadl when we should be seeing a ~70% increase
+# Have changed some transition models to see if we can change this significantly
+# Now 14% reduction, not good. Need to have a good look at these later.
+
+```
+
+# Stroke Comparison (Wittenberg)
+
+Data from [this paper](https://www.semanticscholar.org/paper/The-future-incidence%2C-prevalence-and-costs-of-in-UK-King-Wittenberg/fdaffc8cd7790556e57c77c0b195c8f7a64c4189) by King et al.
+(2020).
+Their model predicts an increase in prevalence of 120% from 2015 to 2035.
+Also, paper projects that incidence of stroke will increase by 60% each year in the same period.
+Going to start by comparing the prevalence, but when I come back to this I think it would be interesting to compare incidence also.
+
+```{r}
+
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Get FEM output and extract only p_anyadl_all
+FEM.stroke <- FEM_base %>%
+              select(contains('year'), contains('stroke')) %>%
+              select(-starts_with('i_'), -starts_with('n_'), -starts_with('a_')) %>%
+              select(contains('year'), contains('_all'))
+# select(-contains('_m'), -contains('_f')) %>%
+
+FEM.stroke.2015 <- data.frame(year = 2015, ((FEM.stroke[2,2] + FEM.stroke[3,2]) / 2))
+FEM.stroke.2040 <- filter(FEM.stroke, year == 2040)
+FEM.stroke.2015_2040 <- rbind(FEM.stroke.2015, FEM.stroke.2040)
+
+# Now calculate percentage change
+FEM.stroke.perchange <- ((FEM.stroke.2015_2040[2,2] - FEM.stroke.2015_2040[1,2]) / FEM.stroke.2015_2040[2,2]) * 100
+
+# Output showing a ~29% increase in stroke, nowhere near the 120% increase Wittenberg predicts
+
+```
+
+```{r}
+
+FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+# Get FEM output and extract only p_anyadl_all
+FEM.dem <- FEM_base %>%
+              select(contains('year'), contains('demene')) %>%
+              select(-starts_with('i_'), -starts_with('n_'), -starts_with('a_')) %>%
+              select(contains('year'), contains('_all'))
+# select(-contains('_m'), -contains('_f')) %>%
+
+FEM.dem.2015 <- data.frame(year = 2015, ((FEM.dem[2,2] + FEM.dem[3,2]) / 2))
+FEM.dem.2040 <- filter(FEM.dem, year == 2040)
+FEM.dem.2015_2040 <- rbind(FEM.dem.2015, FEM.dem.2040)
+
+# Now calculate percentage change
+FEM.dem.perchange <- ((FEM.dem.2015_2040[2,2] - FEM.dem.2015_2040[1,2]) / FEM.dem.2015_2040[2,2]) * 100
+
+```
+
+# Bland Altman method of Validation
+
+```{r}
+# Contains method for bland-altman validation 
+require(MethComp)
+
+```

### 2021-02-17 FEM_R/model_analysis_core.Rmd
Creating copies of some files for the core model, including a new covar_definitions file with stripped back right hand variables

--- 2021-02-17 FEM_R/model_analysis.Rmd
@@
 
-baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+baseline <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
 
@@
 ```{r echo = FALSE}
-cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
+cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')
 
@@
 ```{r}
-FEM_base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+FEM_base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
 

### 2021-02-24 FEM_R/results.Rmd
Many changes to get core scenarios working. Added ELSA_core_scen.csv file for running the core scenarios

---
title: "R Notebook"
output: html_notebook
---

This notebook will hold some initial testing and visualisation for the results section of the IJM paper.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
#require(plyr)
#require(data.table)

```

```{r}

# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

```

Data has been read in, lets create some functions to visualise.

```{r}

compareSurvivalCurve <- function(cohort, intervention, int_name,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
                 melt(id.var = 'Age', value.name = 'Survival')

  p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
        geom_line() +
        labs(title = paste('Baseline vs', int_name), x = 'Age', y = 'Survival')

  ggsave(filename = paste0(int_name, '_survival_curve.png'),
         plot = p1,
         path = path)

  return(p1)

}

```

```{r}

compareDeathAge <- function(base, int, int_name, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)

  print(colnames(aDeath))

  aDeath <- aDeath %>%
                dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
                reshape2::melt(id.var = 'Year', value.name = 'Age')

  p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
        geom_line() +
        labs(title = paste('Baseline vs', int_name), x = 'Year', y = 'Age at Death')

  ggsave(filename = paste0(int_name, '_death_age.png'),
         plot = p1,
         path = path)

  return(p1)
}
```

# Hearte

## Cohort

```{r}

coh <- cohort
h <- heart_c

compareSurvivalCurve(coh, h, 'No Heart Disease')

```

```{r}

coh <- cohort
h <- heart_c

cohortCDVis <- function(cohort, intervention, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  cohMelt <- cohort %>%
              select(c('Age', all_of(cds))) %>%
              melt(id.var='Age', value.name = 'Prevalence_cohort')

  intMelt <- intervention %>%
              select(c('Age', all_of(cds))) %>%
              melt(id.var='Age', value.name = 'Prevalence_intervention')

  merged <- merge(cohMelt, intMelt)

  pivot.merged <- pivot_longer(data = merged,
                             cols = Prevalence_cohort:Prevalence_intervention,
                             names_to = c('Prevalence', 'source'),
                             names_sep = '_')
  pivot.merged <- select(pivot.merged, -Prevalence)

  p1 <- ggplot(data = pivot.merged, aes(x = Age, y = value, color = source)) +
          geom_line() +
          facet_wrap(. ~ variable, scales = "free")

  ggsave(filename = 'noHearte_cd_lattice.png',
         plot = p1,
         path = path)

  return(p1)

}

cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')

cohortCDVis(coh, h, cds)

```

## Population

```{r}

b <- base
h_p <- heart_p

compareDeathAge(b, h_p, 'No Heart Disease')

```

# Smoking

## Average Age at Death

```{r}

b2 <- base
s <- smok

compareDeathAge(b2, smok, 'No Smoking')

```

## Chronic Diseases

```{r}

baseCDVis <- function(base, int, cds, cds.rename, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  selBase <- base %>%
              select(c('year', all_of(cds)))
  colnames(selBase) <- c('Year', cds.rename)

  selInt <- int %>%
              select(c('year', all_of(cds)))
  colnames(selInt) <- c('Year', cds.rename)

  pivot.selBase <- pivot_longer(data = selBase,
                              cols = 'Lung Disease':Diabetes,
                              names_to = 'Condition',
                              values_to = 'Prevalence')
  pivot.selBase$scen <- 'Baseline'

  pivot.selInt <- pivot_longer(data = selInt,
                              cols = 'Lung Disease':Diabetes,
                              names_to = 'Condition',
                              values_to = 'Prevalence')
  pivot.selInt$scen <- 'Intervention'

  merged <- rbind(pivot.selBase, pivot.selInt)

  p <- ggplot(data = merged, aes(x = Year, y = Prevalence, color = scen)) +
          geom_line() +
          facet_wrap(. ~ Condition, scales = "free")
          labs(x = 'Year', y = 'Prevalence')

  ggsave(filename = 'noSmoke_cd_lattice.png',
         plot = p,
         path = path)

  return(p)
}

cds <- c('p_lunge_all', 'p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
cds.rename <- c('Lung Disease', 'Heart', 'Cancer', 'Stroke', 'Diabetes')
baseCDVis(b2, s, cds, cds.rename)

```

```{r}

visCDComparison <- function(base, int, int_name, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # For file naming and stuff
  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  } else if(cd == 'lunge') {
    dis <- 'Lung Disease'
  } else if(cd == 'stroke') {
    dis <- 'Stroke'
  }

  var <- paste0('p_', cd, '_all')

  selBase <- base %>%
              select(c('year', all_of(var))) %>%
              dplyr::rename('Baseline' = var)

  selInt <- int %>%
              select(c('year', all_of(var))) %>%
              dplyr::rename('Intervention' = var)

  melted <- merge(selBase, selInt, by = 'year') %>%
              reshape2::melt(id.var = 'year', value.name = 'Scenario')

  p <- ggplot(data = melted, aes(x = year, y = Scenario, color = variable)) +
          geom_line() +
          labs(title = paste(int_name, dis, sep = ' - '), x = 'Year', y = 'Prevalence')

  ggsave(filename = paste0(int_name, '_', dis, '_prev.png'),
         plot = p,
         path = path)

  return(p)

}

visCDComparison(b2, s, 'No Smoke', 'cancre')
visCDComparison(b2, s, 'No Smoke', 'diabe')
visCDComparison(b2, s, 'No Smoke', 'hearte')
visCDComparison(b2, s, 'No Smoke', 'lunge')
visCDComparison(b2, s, 'No Smoke', 'stroke')

```

# BMI

```{r}

c <- cohort
bmi1 <- obese1
bmi2 <- obese2
bmi3 <- nomorbid

compareSurvivalCurveBMI <- function(cohort, int1, int2, int3,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  int1$Age <- (int1$year - 2012) + 50
  int2$Age <- (int2$year - 2012) + 50
  int3$Age <- (int3$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  int1$Survival <- int1$m_endpop_all / int1$m_endpop_all[1]
  int2$Survival <- int2$m_endpop_all / int2$m_endpop_all[1]
  int3$Survival <- int3$m_endpop_all / int3$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, int1$Survival, int2$Survival, int3$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Cohort' = cohort.Survival, 'No Obese 1' = int1.Survival, 'No Obese 2' = int2.Survival, 'No Morbid' = int3.Survival)

  pivot.survival <- pivot_longer(data = survivals,
                               cols = Cohort:'No Morbid',
                               names_to = 'Scenario',
                               values_to = 'Survival')

  p1 <-ggplot(data = pivot.survival, aes(x=Age, y=Survival, color = Scenario)) +
        geom_line() +
        labs(title = 'BMI Intervention Survival', x = 'Age', y = 'Survival')

  ggsave(filename = 'BMI_survival_curves.png',
         plot = p1,
         path = path)

  return(p1)

}

compareSurvivalCurveBMI(cohort, bmi1, bmi2, bmi3)

```

```{r}

c <- cohort
bmi1 <- obese1
bmi2 <- obese2
bmi3 <- nomorbid

# # Calculate age and survival curves
# c$Age <- (c$year - 2012) + 50
# bmi1$Age <- (bmi1$year - 2012) + 50
# bmi2$Age <- (bmi2$year - 2012) + 50
# bmi3$Age <- (bmi3$year - 2012) + 50
# c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
# bmi1$Survival <- bmi1$m_endpop_all / bmi1$m_endpop_all[1]
# bmi2$Survival <- bmi2$m_endpop_all / bmi2$m_endpop_all[1]
# bmi3$Survival <- bmi3$m_endpop_all / bmi3$m_endpop_all[1]
# 
# cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')
# 
# coh <- c %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_cohort = p_hearte_all, cancer_cohort = p_cancre_all, 
#                           stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
#                           bmi_cohort = a_bmi_all)
# 
# b1 <- bmi1 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi1 = p_hearte_all, cancer_bmi1 = p_cancre_all, 
#                           stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
#                           bmi_bmi1 = a_bmi_all)
# 
# b2 <- bmi2 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi2 = p_hearte_all, cancer_bmi2 = p_cancre_all, 
#                           stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
#                           bmi_bmi2 = a_bmi_all)
# 
# b3 <- bmi3 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi3 = p_hearte_all, cancer_bmi3 = p_cancre_all, 
#                           stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
#                           bmi_bmi3 = a_bmi_all)
# 
# merged <- Reduce(merge, list(coh, b1, b2, b3))
# 
# pivot.merged <- pivot_longer(data = merged,
#                              cols = heart_cohort:bmi_bmi3,
#                              names_to = c('Feature', 'Scenario'),
#                              names_sep = '_',
#                              values_to = 'Value')
# 
# ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()

visCDComparisonBMI <- function(cohort, int1, int2, int3, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  int1$Age <- (int1$year - 2012) + 50
  int2$Age <- (int2$year - 2012) + 50
  int3$Age <- (int3$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  int1$Survival <- int1$m_endpop_all / int1$m_endpop_all[1]
  int2$Survival <- int2$m_endpop_all / int2$m_endpop_all[1]
  int3$Survival <- int3$m_endpop_all / int3$m_endpop_all[1]

  coh <- cohort %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_cohort = p_hearte_all, cancer_cohort = p_cancre_all, 
                            stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
                            bmi_cohort = a_bmi_all)

  b1 <- int1 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi1 = p_hearte_all, cancer_bmi1 = p_cancre_all, 
                            stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
                            bmi_bmi1 = a_bmi_all)

  b2 <- int2 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi2 = p_hearte_all, cancer_bmi2 = p_cancre_all, 
                            stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
                            bmi_bmi2 = a_bmi_all)

  b3 <- int3 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi3 = p_hearte_all, cancer_bmi3 = p_cancre_all, 
                            stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
                            bmi_bmi3 = a_bmi_all)

  merged <- Reduce(merge, list(coh, b1, b2, b3))

  pivot.merged <- pivot_longer(data = merged,
                               cols = heart_cohort:bmi_bmi3,
                               names_to = c('Feature', 'Scenario'),
                               names_sep = '_',
                               values_to = 'Value')

  p1 <-ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Average BMI', x = 'Age', y = 'BMI')

  ggsave(filename = 'BMI_abmi.png',
         plot = p1,
         path = path)

  p2 <- ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Heart Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_phearte.png',
         plot = p2,
         path = path)

  p3 <- ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Stroke', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_pstroke.png',
         plot = p3,
         path = path)

  p4 <- ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Diabetes', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_pdiabe.png',
         plot = p4,
         path = path)

  return(list(p1, p2, p3, p4))
}

cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')

visList <- visCDComparisonBMI(c, bmi1, bmi2, bmi3, cds)

for(graph in visList) {
  print(graph)
}

```

```{r}

visCDComparison(c, bmi3, 'No Morbid', 'cancre')
visCDComparison(c, bmi3, 'No Morbid', 'diabe')
visCDComparison(c, bmi3, 'No Morbid', 'stroke')
visCDComparison(c, bmi3, 'No Morbid', 'hearte')
visCDComparison(c, bmi3, 'No Morbid', 'lunge')

```

### 2021-04-19 FEM_R/model_analysis_core.Rmd
Many changes to get core scenarios working. Added ELSA_core_scen.csv file for running the core scenarios
Lots of small changes, committing altogether as would take forever to do piece by piece

@@
 #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-FEM <- read_dta('ELSA_AgeUK_valid/ELSA_AgeUK_valid_summary.dta')
+FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
 
@@
 # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene', 'asthmae')
+diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'demene', 'asthmae')
 
@@
 # Rename a few just for completeness
-ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
 ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
@@
 ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+ageUK <- filter(ageUK, Condition != 'Cancer, recent')
 
@@
 newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
-newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
 newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)

### 2021-04-19 FEM_R/results.Rmd
Lots of small changes, committing altogether as would take forever to do piece by piece

@@
 
-This notebook will hold some initial testing and visualisation for the results section of the IJM paper.
+# FEM Results
 
@@
 # Need to set up the environment
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
+workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
 knitr::opts_knit$set(root.dir = workingDir)
@@
 require(ggplot2)
+require(ggpubr)
 #require(plyr)
@@
 ```{r}
-
+## Read in all the data
 # Baselines
@@
 # Hearte (cohort and pop)
-heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
+#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
 heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
@@
 
+# Drinking
+nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
+noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
 ```
 
-Data has been read in, lets create some functions to visualise.
+## Functions
 
-```{r}
+Define some functions here to run common visualisations.
 
-compareSurvivalCurve <- function(cohort, intervention, int_name,
+```{r "functions", include=FALSE}
+
+compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
                                  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
@@
 
+  if(save) {
     ggsave(filename = paste0(int_name, '_survival_curve.png'),
@@
          path = path)
+  }
 
   return(p1)
-
 }
 
-```
-
-```{r}
-
-compareDeathAge <- function(base, int, int_name, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
   aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)
-
-  print(colnames(aDeath))
 
@@
   p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = paste('Baseline vs', int_name), x = 'Year', y = 'Age at Death')
 
+  if(save) {
     ggsave(filename = paste0(int_name, '_death_age.png'),
@@
          path = path)
+  }
 
@@
 }
-```
 
-# Hearte
-
-## Cohort
-
-```{r}
-
-coh <- cohort
-h <- heart_c
-
-compareSurvivalCurve(coh, h, 'No Heart Disease')
-
-```
-
-```{r}
-
-coh <- cohort
-h <- heart_c
-
-cohortCDVis <- function(cohort, intervention, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+cohortCDVis <- function(cohort, intervention, cds, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
@@
           geom_line() +
-          facet_wrap(. ~ variable, scales = "free")
+          facet_wrap(. ~ variable, scales = "free", ncol=2)
 
+  if(save) {
     ggsave(filename = 'noHearte_cd_lattice.png',
@@
          path = path)
+  }
 
   return(p1)
-
 }
 
-cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
+baseCDVis <- function(base, intervention, int.name, cds, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
-cohortCDVis(coh, h, cds)
+  baseMelt <- base %>%
+              select(c('year', all_of(cds))) %>%
+              melt(id.var='year', value.name = 'Prevalence_base')
 
-```
+  intMelt <- intervention %>%
+              select(c('year', all_of(cds))) %>%
+              melt(id.var='year', value.name = 'Prevalence_intervention')
 
-## Population
+  merged <- merge(baseMelt, intMelt)
 
-```{r}
+  pivot.merged <- pivot_longer(data = merged,
+                             cols = Prevalence_base:Prevalence_intervention,
+                             names_to = c('Prevalence', 'source'),
+                             names_sep = '_')
+  pivot.merged <- select(pivot.merged, -Prevalence)
 
-b <- base
-h_p <- heart_p
+  p1 <- ggplot(data = pivot.merged, aes(x = year, y = value, color = source)) +
+          geom_line() +
+          facet_wrap(. ~ variable, scales = "free", ncol=2) +
+          labs(title = int.name)
 
-compareDeathAge(b, h_p, 'No Heart Disease')
+  return(p1)
+}
 
-```
-
-# Smoking
-
-## Average Age at Death
-
-```{r}
-
-b2 <- base
-s <- smok
-
-compareDeathAge(b2, smok, 'No Smoking')
-
-```
-
-## Chronic Diseases
-
-```{r}
-
-baseCDVis <- function(base, int, cds, cds.rename, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+baseCDVis2 <- function(base, int, cds, cds.rename, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
   selBase <- base %>%
-              select(c('year', all_of(cds)))
+              dplyr::select(c('year', all_of(cds)))
   colnames(selBase) <- c('Year', cds.rename)
@@
   selInt <- int %>%
-              select(c('year', all_of(cds)))
+              dplyr::select(c('year', all_of(cds)))
   colnames(selInt) <- c('Year', cds.rename)
@@
 
+  if(save) {
     ggsave(filename = 'noSmoke_cd_lattice.png',
@@
 
-cds <- c('p_lunge_all', 'p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
-cds.rename <- c('Lung Disease', 'Heart', 'Cancer', 'Stroke', 'Diabetes')
-baseCDVis(b2, s, cds, cds.rename)
-
-```
-
-```{r}
-
-visCDComparison <- function(base, int, int_name, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+visCDComparison <- function(base, int, int_name, cd, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
@@
     dis <- 'Stroke'
+  } else if(cd == 'smoken') {
+    dis <- 'Smoking'
+  } else if(cd == 'employed') {
+    dis <- 'Employed'
+  } else if(cd == 'retired') {
+    dis <- 'Retired'
+  } else if(cd == 'anyadl') {
+    dis <- 'ADLs'
+  } else if(cd == 'anyiadl') {
+    dis <- 'IADLs'
   }
@@
   selBase <- base %>%
-              select(c('year', all_of(var))) %>%
+              dplyr::select(c('year', all_of(var))) %>%
               dplyr::rename('Baseline' = var)
@@
   selInt <- int %>%
-              select(c('year', all_of(var))) %>%
+              dplyr::select(c('year', all_of(var))) %>%
               dplyr::rename('Intervention' = var)
@@
 
+  if(save) {
     ggsave(filename = paste0(int_name, '_', dis, '_prev.png'),
@@
          path = path)
+  }
 
   return(p)
-
 }
 
-visCDComparison(b2, s, 'No Smoke', 'cancre')
-visCDComparison(b2, s, 'No Smoke', 'diabe')
-visCDComparison(b2, s, 'No Smoke', 'hearte')
-visCDComparison(b2, s, 'No Smoke', 'lunge')
-visCDComparison(b2, s, 'No Smoke', 'stroke')
+visCDComparisonCustom <- function(base, int, int_name, cd, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+  # For file naming and stuff
+  if(cd == 'lunge') {
+    tit <- 'Lung Disease'
+    y.lab <- ' '
+    prev <- TRUE
+  } else if(cd == 'smoken') {
+    tit <- 'Smoking'
+    y.lab <- 'Prevalence'
+    prev <- TRUE
+  } else if(cd == 'age') {
+    tit <- 'Average Age'
+    y.lab <- 'Age'
+    prev <- FALSE
+  }
 
-```
+  if(prev) {
+    var <- paste0('p_', cd, '_all')
+  } else {
+    var <- paste0('a_', cd, '_all')
+  }
 
-# BMI
+  selBase <- base %>%
+              dplyr::select(c('year', all_of(var))) %>%
+              dplyr::rename('Baseline' = var)
+  selInt <- int %>%
+              dplyr::select(c('year', all_of(var))) %>%
+              dplyr::rename('Intervention' = var)
+  melted <- merge(selBase, selInt, by = 'year') %>%
+              reshape2::melt(id.var = 'year', value.name = 'Prevalence') %>%
+              dplyr::rename('Scenario' = variable)
 
-```{r}
+  p <- ggplot(data = melted, aes(x = year, y = Prevalence, color = Scenario)) +
+          geom_line() +
+          labs(title = tit, x = 'Year', y = y.lab)
 
-c <- cohort
-bmi1 <- obese1
-bmi2 <- obese2
-bmi3 <- nomorbid
+  return(p)
+}
 
-compareSurvivalCurveBMI <- function(cohort, int1, int2, int3,
+compareSurvivalCurveBMI <- function(cohort, int1, int2, int3, save=FALSE,
                                  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
@@
 
+  if(save) {
     ggsave(filename = 'BMI_survival_curves.png',
@@
          path = path)
+  }
 
   return(p1)
-
 }
 
-compareSurvivalCurveBMI(cohort, bmi1, bmi2, bmi3)
-
-```
-
-```{r}
-
-c <- cohort
-bmi1 <- obese1
-bmi2 <- obese2
-bmi3 <- nomorbid
-
-# # Calculate age and survival curves
-# c$Age <- (c$year - 2012) + 50
-# bmi1$Age <- (bmi1$year - 2012) + 50
-# bmi2$Age <- (bmi2$year - 2012) + 50
-# bmi3$Age <- (bmi3$year - 2012) + 50
-# c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
-# bmi1$Survival <- bmi1$m_endpop_all / bmi1$m_endpop_all[1]
-# bmi2$Survival <- bmi2$m_endpop_all / bmi2$m_endpop_all[1]
-# bmi3$Survival <- bmi3$m_endpop_all / bmi3$m_endpop_all[1]
-# 
-# cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')
-# 
-# coh <- c %>%
-#             select(Age, all_of(cds)) %>%
-#             dplyr::rename(heart_cohort = p_hearte_all, cancer_cohort = p_cancre_all, 
-#                           stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
-#                           bmi_cohort = a_bmi_all)
-# 
-# b1 <- bmi1 %>%
-#             select(Age, all_of(cds)) %>%
-#             dplyr::rename(heart_bmi1 = p_hearte_all, cancer_bmi1 = p_cancre_all, 
-#                           stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
-#                           bmi_bmi1 = a_bmi_all)
-# 
-# b2 <- bmi2 %>%
-#             select(Age, all_of(cds)) %>%
-#             dplyr::rename(heart_bmi2 = p_hearte_all, cancer_bmi2 = p_cancre_all, 
-#                           stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
-#                           bmi_bmi2 = a_bmi_all)
-# 
-# b3 <- bmi3 %>%
-#             select(Age, all_of(cds)) %>%
-#             dplyr::rename(heart_bmi3 = p_hearte_all, cancer_bmi3 = p_cancre_all, 
-#                           stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
-#                           bmi_bmi3 = a_bmi_all)
-# 
-# merged <- Reduce(merge, list(coh, b1, b2, b3))
-# 
-# pivot.merged <- pivot_longer(data = merged,
-#                              cols = heart_cohort:bmi_bmi3,
-#                              names_to = c('Feature', 'Scenario'),
-#                              names_sep = '_',
-#                              values_to = 'Value')
-# 
-# ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
-#   geom_line()
-# 
-# ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
-#   geom_line()
-# 
-# ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
-#   geom_line()
-# 
-# ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
-#   geom_line()
-
-visCDComparisonBMI <- function(cohort, int1, int2, int3, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+visCDComparisonBMI <- function(cohort, int1, int2, int3, cds, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
 
@@
                             stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
-                            bmi_cohort = a_bmi_all)
+                            bmi_cohort = a_bmi_all, employed_cohort = p_employed_all,
+                            retired_cohort = p_retired_all, adl_cohort = p_anyadl_all,
+                            iadl_cohort = p_anyiadl_all)
 
@@
                             stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
-                            bmi_bmi1 = a_bmi_all)
+                            bmi_bmi1 = a_bmi_all, employed_bmi1 = p_employed_all,
+                            retired_bmi1 = p_retired_all, adl_bmi1 = p_anyadl_all,
+                            iadl_bmi1 = p_anyiadl_all)
 
@@
                             stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
-                            bmi_bmi2 = a_bmi_all)
+                            bmi_bmi2 = a_bmi_all, employed_bmi2 = p_employed_all,
+                            retired_bmi2 = p_retired_all, adl_bmi2 = p_anyadl_all,
+                            iadl_bmi2 = p_anyiadl_all)
 
@@
                             stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
-                            bmi_bmi3 = a_bmi_all)
+                            bmi_bmi3 = a_bmi_all, employed_bmi3 = p_employed_all,
+                            retired_bmi3 = p_retired_all, adl_bmi3 = p_anyadl_all,
+                            iadl_bmi3 = p_anyiadl_all)
 
@@
   pivot.merged <- pivot_longer(data = merged,
-                               cols = heart_cohort:bmi_bmi3,
+                               cols = heart_cohort:iadl_bmi3,
                                names_to = c('Feature', 'Scenario'),
@@
 
+  if(save) {
     ggsave(filename = 'BMI_abmi.png',
@@
 
+  if(save) {
     ggsave(filename = 'BMI_phearte.png',
@@
 
+  if(save) {
     ggsave(filename = 'BMI_pstroke.png',
@@
 
+  if(save) {
     ggsave(filename = 'BMI_pdiabe.png',
@@
          path = path)
-
-  return(list(p1, p2, p3, p4))
   }
 
-cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')
+  p5 <- ggplot(data = filter(pivot.merged, Feature == 'employed'), aes(x = Age, y = Value, color = Scenario)) +
+        geom_line() +
+        labs(title = 'Employed', x = 'Age', y = 'Prevalence')
+
+  p6 <- ggplot(data = filter(pivot.merged, Feature == 'retired'), aes(x = Age, y = Value, color = Scenario)) +
+        geom_line() +
+        labs(title = 'Retired', x = 'Age', y = 'Prevalence')
+
+  p7 <- ggplot(data = filter(pivot.merged, Feature == 'adl'), aes(x = Age, y = Value, color = Scenario)) +
+        geom_line() +
+        labs(title = 'ADLs', x = 'Age', y = 'Prevalence')
+
+  p8 <- ggplot(data = filter(pivot.merged, Feature == 'iadl'), aes(x = Age, y = Value, color = Scenario)) +
+        geom_line() +
+        labs(title = 'IADLs', x = 'Age', y = 'Prevalence')
+
+  return(list(p1, p2, p3, p4, p5, p6, p7, p8))
+}
+
+```
+
+## Prevalence Change (2012-2050)
+
+Here we will look at the % change in prevalence from 2012 - 2050, and compare these between variables.
+
+Todo:
+- Visualise results
+  - Bar chart
+  - Fancy table
+
+```{r "prev_change"}
+
+b <- base
+
+b <- b %>%
+        select(contains(c('year', 'cancre', 'diabe', 'hearte', 'lunge', 'stroke', 'demene', 'employed', 'anyadl', 'anyiadl', 'adl2', 'adl3p')))
+
+b.2012 <- b %>% filter(year == 2012)
+b.2050 <- b %>% filter(year == 2050)
+
+cancre.change <- b.2050$p_cancre_all / b.2012$p_cancre_all
+diabe.change <- b.2050$p_diabe_all / b.2012$p_diabe_all
+hearte.change <- b.2050$p_hearte_all / b.2012$p_hearte_all
+lunge.change <- b.2050$p_lunge_all / b.2012$p_lunge_all
+stroke.change <- b.2050$p_stroke_all / b.2012$p_stroke_all
+demene.change <- b.2050$p_demene_all / b.2012$p_demene_all
+employed.change <- b.2050$p_employed_all / b.2012$p_employed_all
+adl.change <- b.2050$p_anyadl_all / b.2012$p_anyadl_all
+iadl.change <- b.2050$p_anyiadl_all / b.2012$p_anyiadl_all
+adl2.change <- b.2050$p_adl2_all / b.2012$p_adl2_all
+adl3p.change <- b.2050$p_adl3p_all / b.2012$p_adl3p_all
+
+print(paste0('Cancer: ', cancre.change))
+print(paste0('Diabetes: ', diabe.change))
+print(paste0('Heart Disease: ', hearte.change))
+print(paste0('Lung Disease: ', lunge.change))
+print(paste0('Stroke: ', stroke.change))
+print(paste0('Dementia: ', demene.change))
+print(paste0('Employment: ', employed.change))
+print(paste0('AnyADL: ', adl.change))
+print(paste0('AnyIADL: ', iadl.change))
+print(paste0('ADL2: ', adl2.change))
+print(paste0('ADL3p: ', adl3p.change))
+
+```
+
+## Heart Disease Intervention (Population)
+
+```{r}
+
+b <- base
+h_p <- heart_p
+
+compareDeathAge(b, h_p, 'No Heart Disease', save=TRUE)
+
+cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'p_hibpe_all')
+baseCDVis(b, h_p, 'No Heart Disease', cds)
+
+econs <- c('p_employed_all', 'p_retired_all', 'p_anyadl_all', 'p_anyiadl_all', 'p_adl2_all', 'p_adl3p_all')
+baseCDVis(b, h_p, 'No Heart Disease', econs)
+
+```
+
+## Smoking Intervention (Cohort)
+
+```{r}
+
+c <- cohort
+s <- smok
+
+compareSurvivalCurve(c, smok, 'No Smoking')
+
+cds <- c('p_lunge_all', 'p_hearte_all', 'p_cancre_all', 'p_smoken_all', 'p_stroke_all', 'p_diabe_all')
+cds.rename <- c('Lung Disease', 'Heart', 'Cancer', 'Smoking', 'Stroke', 'Diabetes')
+baseCDVis2(c, s, cds, cds.rename)
+
+visCDComparison(c, s, 'No Smoke', 'cancre')
+visCDComparison(c, s, 'No Smoke', 'diabe')
+visCDComparison(c, s, 'No Smoke', 'hearte')
+visCDComparison(c, s, 'No Smoke', 'lunge')
+visCDComparison(c, s, 'No Smoke', 'stroke')
+visCDComparison(c, s, 'No Smoke', 'smoken')
+visCDComparison(c, s, 'No Smoke', 'employed')
+visCDComparison(c, s, 'No Smoke', 'retired')
+visCDComparison(c, s, 'No Smoke', 'anyadl')
+visCDComparison(c, s, 'No Smoke', 'anyiadl')
+
+smok.prev <- visCDComparisonCustom(c, s, 'No Smoke', 'smoken')
+smok.lung <- visCDComparisonCustom(c, s, 'No Smoke', 'lunge')
+#smok.age <- visCDComparisonCustom(b2, s, 'No Smoke', 'age')
+
+ggarrange(smok.prev, smok.lung,
+          labels = c('A', 'B', 'C'),
+          common.legend = TRUE)
+```
+
+```{r}
+
+# Lets get the results by age group, sex and education level
+# For this we need detailed output
+#c_det <- read_dta('ELSA_core_cohort')
+#s <- smok
+
+cohortVisByGroup <- function(cohort, intervention, cds, groups, save=FALSE) {
+
+  # Calculate age and survival curves
+  cohort$Age <- (cohort$year - 2012) + 50
+  intervention$Age <- (intervention$year - 2012) + 50
+  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
+  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
+
+  cohMelt <- cohort %>%
+              select(c('Age', all_of(cds), all_of(groups))) %>%
+              melt(id.var=c('Age', 'male', 'educ'), value.name = 'Prevalence_cohort')
+
+  intMelt <- intervention %>%
+              select(c('Age', all_of(cds), all_of(groups))) %>%
+              melt(id.var=c('Age', 'male', 'educ'), value.name = 'Prevalence_intervention')
+
+  merged <- merge(cohMelt, intMelt)
+
+  pivot.merged <- pivot_longer(data = merged,
+                             cols = Prevalence_cohort:Prevalence_intervention,
+                             names_to = c('Prevalence', 'source'),
+                             names_sep = '_')
+  pivot.merged <- select(pivot.merged, -Prevalence)
+
+  return(pivot.merged)
+}
+
+#cds <- c('p_lunge_all', 'p_hearte_all', 'p_cancre_all', 'p_smoken_all', 'p_stroke_all', 'p_diabe_all')
+#groups <- c('male', 'age', 'educ')
+
+#output <- cohortVisByGroup(c, s, cds, groups)
+
+```
+
+## BMI Reduction Interventions
+
+```{r}
+
+c <- cohort
+bmi1 <- obese1
+bmi2 <- obese2
+bmi3 <- nomorbid
+
+compareSurvivalCurveBMI(cohort, bmi1, bmi2, bmi3)
+
+cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all', 
+         'p_employed_all', 'p_retired_all', 'p_anyadl_all', 'p_anyiadl_all')
 
@@
 
-```
-
-```{r}
-
 visCDComparison(c, bmi3, 'No Morbid', 'cancre')
@@
 visCDComparison(c, bmi3, 'No Morbid', 'lunge')
+visCDComparison(c, bmi3, 'No Morbid', 'employed')
+visCDComparison(c, bmi3, 'No Morbid', 'retired')
+visCDComparison(c, bmi3, 'No Morbid', 'anyadl')
+visCDComparison(c, bmi3, 'No Morbid', 'anyiadl')
 
 ```
+
+## Quit Alcohol Intervention (Just out of interest)
+
+```{r}
+
+b3 <- base
+nod <- nodrink
+nopd <- noBadDrink
+
+compareDeathAge(b3, nod, "No Drink")
+compareDeathAge(b3, nopd, "No Problem Drink")
+
+cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
+baseCDVis(b3, nod, 'No Drink', cds)
+
+econs <- c('p_employed_all', 'p_retired_all', 'p_anyadl_all', 'p_anyiadl_all')
+baseCDVis(b3, nod, 'No Drink', econs)
+
+baseCDVis(b3, nopd, 'No Problem Drink', cds)
+baseCDVis(b3, nopd, 'No Problem Drink', econs)
+
+```
+
+# OLD DEFUNKT CODE
+
+### Hearte Cohort Intervention
+```{r}
+
+coh <- cohort
+h <- heart_c
+
+# Survival Curves
+compareSurvivalCurve(coh, h, 'No Heart Disease')
+
+# Chronic Diseases
+cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
+cohortCDVis(coh, h, cds)
+
+# Economic Factors
+econs <- c('p_employed_all', 'p_retired_all', 'p_anyadl_all', 'p_anyiadl_all')
+cohortCDVis(coh, h, econs)
+
+```

### 2021-04-19 FEM_CPP/HealthModule.cpp

@@ -306,8 +306,8 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
 						person->set(Vars::smokev,person->get(Vars::l2smokev));
 						// Clean up other vars - didn't start, too
 						person->set(Vars::smoke_start,0.0);
-						// Set intensity to non-smoker (smkint == 0)
-						person->set(Vars::smkint, 0.0);
+						// Set intensity to non-smoker (heavy_smoker == 0)
+						person->set(Vars::heavy_smoker, 0.0);
 					}	
 					else {
 						// Didn't stop, so maintain smkstat, smoken, smokev, smkint
@@ -341,7 +341,7 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
 						person->set(Vars::smoken,person->get(Vars::l2smoken));
 						person->set(Vars::smokev,person->get(Vars::l2smokev));
 						//person->set(Vars::smkint,person->get(Vars::l2smkint));
-                        person->set(Vars::smkint, 0.0);
+                        person->set(Vars::heavy_smoker, 0.0);
 						// Clean up other vars - didn't stop, too
 						person->set(Vars::smoke_stop,0.0);
 					}
@@ -351,12 +351,13 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
 			// If we change drink to 0 then we need to make sure heavy_drinker and freq_drinker are also 0
 			if (elsa_data) {
 			    // previously drinker
-			    if(person->get(Vars::l2drink)) {
+			    if(person->test(Vars::l2drink)) {
 			        // now quit
-			        if(person->get(Vars::drink)) {
+			        if(person->get(Vars::drink) == 0) {
 			            // set heavy and freq to 0
-			            person->set(Vars::heavy_drinker, 0.0);
-			            person->set(Vars::freq_drinker, 0.0);
+			            //person->set(Vars::heavy_drinker, 0.0);
+			            //person->set(Vars::freq_drinker, 0.0);
+			            person->set(Vars::problem_drinker, 0.0);
 			        }
 			    }
 			}

### 2021-04-19 FEM_CPP/Vars.cpp

@@ -620,6 +620,9 @@ VarsInfo::VarsInfo() {
     info[Vars::freq_drinker]        = VarInfo("freq_drinker", "Frequent Drinker (>5 days/week)", VarTypes::Boolean);
     info[Vars::l2freq_drinker]      = VarInfo("l2freq_drinker", "Lag of Frequent Drinker (>5 days/week)", VarTypes::Boolean);
     info[Vars::pfreq_drinker]       = VarInfo("pfreq_drinker", "Probability of Frequent Drinker", VarTypes::Double);
+    info[Vars::problem_drinker]     = VarInfo("problem_drinker", "Problem Drinker (binge/frequent)", VarTypes::Boolean);
+    info[Vars::l2problem_drinker]   = VarInfo("l2problem_drinker", "Lag of Problem Drinker (binge/frequent)", VarTypes::Boolean);
+    info[Vars::pproblem_drinker]    = VarInfo("pproblem_drinker", "Probability of Problem Drinker (binge/frequent)", VarTypes::Double);
 	//info[Vars::l2drinkd]			= VarInfo("l2drinkd", "Lag of Number of days per week drinks alcohol, Range [0,7]", VarTypes::Short);
 	//info[Vars::l2drinkwn]			= VarInfo("l2drinkwn", "Lag of number of drinks per week", VarTypes::Double);
 	//info[Vars::drinkd1]				= VarInfo("drinkd1", "drinkd 1 [0, 1]", VarTypes::Boolean, Vars::drinkd_stat, 1);
@@ -667,6 +670,9 @@ VarsInfo::VarsInfo() {
     info[Vars::psmkint1]            = VarInfo("psmkint1", "Probability of smoking intensity level 1", VarTypes::Double);
     info[Vars::psmkint2]            = VarInfo("psmkint2", "Probability of smoking intensity level 2", VarTypes::Double);
     info[Vars::psmkint3]            = VarInfo("psmkint3", "Probability of smoking intensity level 3", VarTypes::Double);
+    info[Vars::heavy_smoker]        = VarInfo("heavy_smoker", "Heavy Smoker (10+ cigs/day)", VarTypes::Boolean);
[…]

### 2021-04-24 FEM_R/model_analysis_core.Rmd
Removed cancer and heart disease graphs from validation

@@
 # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'demene', 'asthmae')
+diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
 
@@
 pivoted.FEM <- pivot_longer(data = FEM,
-                            cols = hearte_m_6064:asthmae_f_100p, 
+                            cols = hibpe_m_6064:asthmae_f_100p, 
                             names_to = c('Condition', 'Gender', 'Age_Group'), 
@@
 # extract all heart related conditions for combining
-ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
 # Aggregate the values by gender and age group
-ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
 # Rename columns and add condition column back
-colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-ageUK_heart$Condition <- 'Heart Problems'
+#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+#ageUK_heart$Condition <- 'Heart Problems'
 
@@
 ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-ageUK <- rbind(ageUK, ageUK_heart)
+#ageUK <- rbind(ageUK, ageUK_heart)
 
@@
 #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
 
@@
 ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
-ageUK <- filter(ageUK, Condition != 'Cancer, recent')
+ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
 
@@
 #       Therefore only validation from this will be that FEM is higher than ageUK
-newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
 newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
@@
 newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
-newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
 #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
@@
 newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
+
+#force Make to run this again
 

### 2021-04-30 FEM_R/results.Rmd
Couple of small changes here and there, also started script to visualise by subgroup but its very much WIP

@@
 
+```{r}
+visCDComparison(c, bmi1, 'No Obesity', 'cancre')
+visCDComparison(c, bmi1, 'No Obesity', 'diabe')
+visCDComparison(c, bmi1, 'No Obesity', 'stroke')
+visCDComparison(c, bmi1, 'No Obesity', 'hearte')
+visCDComparison(c, bmi1, 'No Obesity', 'lunge')
+visCDComparison(c, bmi1, 'No Obesity', 'employed')
+visCDComparison(c, bmi1, 'No Obesity', 'retired')
+visCDComparison(c, bmi1, 'No Obesity', 'anyadl')
+visCDComparison(c, bmi1, 'No Obesity', 'anyiadl')
+```
+
 ## Quit Alcohol Intervention (Just out of interest)

### 2021-04-30 FEM_R/visualise_by_subgroup.Rmd
Couple of small changes here and there, also started script to visualise by subgroup but its very much WIP

---
title: "R Notebook"
output: html_notebook
---

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(ggpubr)
#require(plyr)
#require(data.table)

rm(workingDir)
```

```{r}
## Read in all the data
# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking
nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

## Detailed Output Appending

```{r}
# Here we will start to use the detailed output (appended?) to get the rates of chronic diseases by specific subgroups
# Might be an idea to write a general function to get proportion by specific sub group
# Will need arguments: dataset, target variable, subgroups

# save path for ease
base.detail.path <- 'ELSA_core_base/detailed_output/'

# First step is to append all the detailed output files together
base.detail.filelist <- list.files(base.detail.path)
# read first file
base.detail <- read_dta(paste0(base.detail.path, base.detail.filelist[1]))

# read in all the remaining files in a loop and add to base.detail
for (f in base.detail.filelist[-1]) {
  new.df <- read_dta(paste0(base.detail.path, f))
  base.detail <- rbind(base.detail, new.df)
}

rm(new.df, base.detail.path, base.detail.filelist, f)
```

The above chunk has created a MASSIVE stata data file containing the detailed output for every repetition for every year in the simulation. We can use this to do some more detailed analysis by subgroup.

Lets start with the Heart disease intervention. Will need to create a detailed output file for this intervention scenario as well.

```{r}
heart.detail.path <- 'ELSA_core_remove_hearte_p/detailed_output/'
heart.detail.filelist <- list.files(heart.detail.path)
heart.detail <- read_dta(paste0(heart.detail.path, heart.detail.filelist[1]))

for (f in heart.detail.filelist[-1]) {
  new.df <- read_dta(paste0(heart.detail.path, f))
  heart.detail <- rbind(heart.detail, new.df)
}

rm(new.df, heart.detail.path, heart.detail.filelist, f)
```

## Visualising

Now we have both base and hearte_p detailed output files. What do we do first?

What are our target visualisations?

We want to get the heart disease intervention comparisons by:
- gender
- education level
- age groups
- exstat? Might be interesting

Good start. How do we achieve this? We need the function now to manipulate these files until only the target variables and sub groups remain. I think it is fine to run the function once for each target/sub group pair. Will mean running multiple times on the large datasets but I think this will be fine for now. We can always change this later, and create another function to make reduced datasets.

```{r}
# After a little bit of google fu I think this is actually quite a simple task:
# https://stackoverflow.com/questions/37057784/summary-of-proportions-by-group

base.sumOut <- base.detail %>%
                    count(age, hearte) %>%
                    mutate(prop = prop.table(n))

base.sumOut

heart.sumOut <- heart.detail %>%
                    count(age, hearte) %>%
                    mutate(prop = prop.table(n))
heart.sumOut

# lets plot just to have a look

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

#filter(combined, hearte ==1)

ggplot(data = combined, aes(x = age, y = prop, color = source)) +
  geom_line() +
  facet_wrap('hearte')

rm(base.sumOut, heart.sumOut, combined)
```

```{r}
# Lets try some more groups before we stick it into a function
base.sumOut <- base.detail %>%
                    count(age, male, hearte) %>%
                    mutate(prop = prop.table(n))

base.sumOut

heart.sumOut <- heart.detail %>%
                    count(age, male, hearte) %>%
                    mutate(prop = prop.table(n))
heart.sumOut

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

ggplot(data = filter(combined, hearte == 1), aes(x = age, y = prop, color = source)) +
  geom_line() +
  facet_wrap(vars(male), labeller = "label_both")

rm(base.sumOut, heart.sumOut, combined)
```

Is this what we would expect? I think this shows that women are more likely to have heart disease than men in the FEM. This can't be correct...

Let's look at overall numbers by gender (not broken down by age).

```{r}
base.sumOut <- base.detail %>%
                    count(male, hearte) %>%
                    mutate(prop = prop.table(n))
base.sumOut

heart.sumOut <- heart.detail %>%
                    count(male, hearte) %>%
                    mutate(prop = prop.table(n))
heart.sumOut

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

combined$male <- factor(combined$male)

ggplot(data = filter(combined, hearte == 1), aes(x = male, y = prop, color = source)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge")
```

Now lets try and do it so that the proportions are relative.

```{r}
base.sumOut <- base.detail %>%
                    group_by(age, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
heart.sumOut <- heart.detail %>%
                    group_by(age, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
head(base.sumOut)
head(heart.sumOut)

# lets plot just to have a look

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

#filter(combined, hearte ==1)

ggplot(data = combined, aes(x = age, y = freq, color = source)) +
  geom_line() +
  facet_wrap('hearte')

rm(base.sumOut, heart.sumOut, combined)
```
This is much better! The frequency's are relative in this attempt, which makes it much easier to understand and compare. Now to recreate the next 2 graphs with relative frequencies instead of proportions.

```{r}
# Lets try some more groups before we stick it into a function
base.sumOut <- base.detail %>%
                    group_by(age, male, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
heart.sumOut <- heart.detail %>%
                    group_by(age, male, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
head(base.sumOut)
head(heart.sumOut)

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

ggplot(data = filter(combined, hearte == 1), aes(x = age, y = freq, color = source)) +
  geom_line() +
  facet_wrap(vars(male), labeller = "label_both")

rm(base.sumOut, heart.sumOut, combined)
```

```{r}
base.sumOut <- base.detail %>%
                    group_by(male, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
heart.sumOut <- heart.detail %>%
                    group_by(male, hearte) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
head(base.sumOut)
head(heart.sumOut)

# give both a source then rbind
base.sumOut$source <- 'Baseline'
heart.sumOut$source <- 'No Heart Disease'

combined <- rbind(base.sumOut, heart.sumOut)

combined$male <- factor(combined$male)

ggplot(data = filter(combined, hearte == 1), aes(x = male, y = freq, color = source)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge")
```

```{r}
# FUnction time

relFreq.subgroup <- function(base, int, target, subgroup) {

  base.sumOut <- base %>%
                    group_by(!!subgroup, !!target) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
  heart.sumOut <- int %>%
                    group_by(!!subgroup, !!target) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))

  # give both a source then rbind
  base.sumOut$source <- 'Baseline'
  heart.sumOut$source <- 'No Heart Disease'

  combined <- rbind(base.sumOut, heart.sumOut)

  ggplot(data = combined, aes(x = !!subgroup, y = freq, color = source)) +
  geom_line() +
  facet_wrap(target, labeller = "label_both")
}

relFreq.subgroup(base.detail, heart.detail, quo(hearte), quo(age))
```

```{r}
relFreq.subgroup.factor <- function(base, int, target, subgroup) {
  base.sumOut <- base %>%
                    group_by(!!subgroup, !!target) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))
  heart.sumOut <- int %>%
                    group_by(!!subgroup, !!target) %>%
                    summarise(n = n()) %>%
                    mutate(freq = n / sum(n))

  # give both a source then rbind
  base.sumOut$source <- 'Baseline'
  heart.sumOut$source <- 'No Heart Disease'

  combined <- rbind(base.sumOut, heart.sumOut)

  #combined$target <- factor(combined$target)

  ggplot(data = filter(combined, !!target == 1), aes(x = !!subgroup, y = freq, color = source)) +
    geom_bar(aes(fill = source), stat = "identity", position = "dodge")
}

relFreq.subgroup.factor(base.detail, heart.detail, quo(hearte), quo(male))
```

### 2021-04-30 FEM_R/calculate_dflys_from_summary.Rmd
Started working on dflys, added some derived vars to ELSA_vars.txt and included them in measures_subpop_ELSA to work with in summary output files

---
title: "R Notebook"
output: html_notebook
---

Going to have a go at producing estimates of disease free life years and disability free life years.

To do this we need to use the 'disabled' and 'anydisease' variables that are defined in ELSA_vars.txt.

First setup and reading in data files.

# Setup
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
#require(reshape2)
require(ggplot2)
#require(ggpubr)
#require(plyr)
#require(data.table)

```

# Read in Data
```{r}
## Read in all the data
# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking
nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

# Disability Free Life Years

Going to attempt disease free first.

What we are looking for is a change in the total number of people with no diseases. i.e. change in n_anydisease_all

Bryan said to 'sum it up for the whole duration of the simulation'. I don't know what this means, but if I have a good go at it now I can always ask him next time we meet.

```{r}
c <- cohort
s <- smok

cSel <- c %>%
          select('year', n_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_disabled_all, 'total' = m_endpop_all)

sSel <- s %>%
          select('year', n_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_disabled_all, 'total' = m_endpop_all)

#combined <- merge(cSel, sSel, by='year')

# Add source and rbind
cSel$source <- 'cohort'
sSel$source <- 'intervention'

combined <- rbind(cSel, sSel)

ggplot(combined, aes(x = year, y = disabled, color = source)) +
  geom_line()
```

The graph above shows why we cannot rely on graphical interpretations of the 'n_' var alone.

Let's try to get the number of disabled years per person.

```{r}
cSel2 <- cSel
sSel2 <- sSel

# Population is in millions initially
cSel2$total <- cSel2$total * 1000000
sSel2$total <- sSel2$total * 1000000

cSel2$dis_by_person <- cSel2$disabled / cSel2$total
sSel2$dis_by_person <- sSel2$disabled / sSel2$total

combined2 <- rbind(cSel2, sSel2)

ggplot(data = combined2, aes(x = year, y = dis_by_person, color = source)) +
  geom_line()

```

## Try again, this time with the correct variable (not disabled)

```{r}
c <- cohort
s <- smok

cSel <- c %>%
          select('year', n_not_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_not_disabled_all, 'total' = m_endpop_all)

sSel <- s %>%
          select('year', n_not_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_not_disabled_all, 'total' = m_endpop_all)

#combined <- merge(cSel, sSel, by='year')

# Add source and rbind
cSel$source <- 'cohort'
sSel$source <- 'intervention'

combined <- rbind(cSel, sSel)

ggplot(combined, aes(x = year, y = disabled, color = source)) +
  geom_line()
```

```{r}
cSel2 <- cSel
sSel2 <- sSel

# Population is in millions initially
cSel2$total <- cSel2$total * 1000000
sSel2$total <- sSel2$total * 1000000

cSel2$dis_by_person <- cSel2$disabled / cSel2$total
sSel2$dis_by_person <- sSel2$disabled / sSel2$total

combined2 <- rbind(cSel2, sSel2)

ggplot(data = combined2, aes(x = year, y = dis_by_person, color = source)) +
  geom_line()

```

### 2021-05-05 FEM_R/model_analysis_core.Rmd
Added lots of new vars to roc_validation.do, fixed some errors, and added educ levels to output subgroups for analysis

@@
                 geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
-                labs(title = 'Comparison of smoking prevalence in the elderly in 2018', x = ' ', y = 'Prevalence (%)')
+                labs(title = 'Prevalence of Smoking (2018)', x = ' ', y = 'Prevalence (%)')
 

### 2021-05-06 FEM_R/results.Rmd
Changed a lot of the visualisations to use geom_smooth() instead of geom_line()

--- 2021-04-19 FEM_R/results.Rmd
@@
   return(p1)
+
+}
+
+visCDComparisonCohort <- function(cohort, int, int_name, cd, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+
+  # For file naming and stuff
+  if(cd == 'cancre') {
+    dis <- 'Cancer'
+  } else if(cd == 'diabe') {
+    dis <- 'Diabetes'
+  } else if(cd == 'hearte') {
+    dis <- 'Heart Disease'
+  } else if(cd == 'lunge') {
+    dis <- 'Lung Disease'
+  } else if(cd == 'stroke') {
+    dis <- 'Stroke'
+  } else if(cd == 'smoken') {
+    dis <- 'Smoking'
+  } else if(cd == 'employed') {
+    dis <- 'Employed'
+  } else if(cd == 'retired') {
+    dis <- 'Retired'
+  } else if(cd == 'anyadl') {
+    dis <- 'ADLs'
+  } else if(cd == 'anyiadl') {
+    dis <- 'IADLs'
+  }
+
+  var <- paste0('p_', cd, '_all')
+
+  # Calculate age and survival curves
+  cohort$Age <- (cohort$year - 2012) + 50
+  intervention$Age <- (intervention$year - 2012) + 50
+  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
+  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
+
+  selcohort <- cohort %>%
+              dplyr::select(c('Age', all_of(var))) %>%
+              dplyr::rename('Cohort' = var)
+
+  selInt <- int %>%
+              dplyr::select(c('Age', all_of(var))) %>%
+              dplyr::rename('Intervention' = var)
+
+  melted <- merge(selcohort, selInt, by = 'year') %>%
+              reshape2::melt(id.var = 'Age', value.name = 'Scenario')
+
+  p <- ggplot(data = melted, aes(x = Age, y = Scenario, color = variable)) +
+          geom_smooth(lwd = 0.5) +
+          labs(title = paste(int_name, dis, sep = ' - '), x = 'Year', y = 'Prevalence')
+
+  if(save) {
+    ggsave(filename = paste0(int_name, '_', dis, '_prev.png'),
+         plot = p,
+         path = path)
+  }
+
+  return(p)
 }
@@
   p <- ggplot(data = melted, aes(x = year, y = Scenario, color = variable)) +
-          geom_line() +
+          geom_smooth(lwd = 0.5) +
           labs(title = paste(int_name, dis, sep = ' - '), x = 'Year', y = 'Prevalence')
@@
   p1 <-ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Average BMI', x = 'Age', y = 'BMI')
@@
   p2 <- ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Heart Disease', x = 'Age', y = 'Prevalence')
@@
   p3 <- ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Stroke', x = 'Age', y = 'Prevalence')
@@
   p4 <- ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Diabetes', x = 'Age', y = 'Prevalence')
@@
   p5 <- ggplot(data = filter(pivot.merged, Feature == 'employed'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Employed', x = 'Age', y = 'Prevalence')
@@
   p6 <- ggplot(data = filter(pivot.merged, Feature == 'retired'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'Retired', x = 'Age', y = 'Prevalence')
@@
   p7 <- ggplot(data = filter(pivot.merged, Feature == 'adl'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'ADLs', x = 'Age', y = 'Prevalence')
@@
   p8 <- ggplot(data = filter(pivot.merged, Feature == 'iadl'), aes(x = Age, y = Value, color = Scenario)) +
-        geom_line() +
+        geom_smooth(lwd = 0.5) +
         labs(title = 'IADLs', x = 'Age', y = 'Prevalence')
@@
 
-visCDComparison(c, s, 'No Smoke', 'cancre')
-visCDComparison(c, s, 'No Smoke', 'diabe')
-visCDComparison(c, s, 'No Smoke', 'hearte')
-visCDComparison(c, s, 'No Smoke', 'lunge')
-visCDComparison(c, s, 'No Smoke', 'stroke')
-visCDComparison(c, s, 'No Smoke', 'smoken')
-visCDComparison(c, s, 'No Smoke', 'employed')
-visCDComparison(c, s, 'No Smoke', 'retired')
-visCDComparison(c, s, 'No Smoke', 'anyadl')
-visCDComparison(c, s, 'No Smoke', 'anyiadl')
+visCDComparisonCohort(c, s, 'No Smoke', 'cancre')
+visCDComparisonCohort(c, s, 'No Smoke', 'diabe')
+visCDComparisonCohort(c, s, 'No Smoke', 'hearte')
+visCDComparisonCohort(c, s, 'No Smoke', 'lunge')
+visCDComparisonCohort(c, s, 'No Smoke', 'stroke')
+visCDComparisonCohort(c, s, 'No Smoke', 'smoken')
+visCDComparisonCohort(c, s, 'No Smoke', 'employed')
+visCDComparisonCohort(c, s, 'No Smoke', 'retired')
+visCDComparisonCohort(c, s, 'No Smoke', 'anyadl')
+visCDComparisonCohort(c, s, 'No Smoke', 'anyiadl')
 
@@
 
-visList <- visCDComparisonBMI(c, bmi1, bmi2, bmi3, cds)
+visList <- visCDComparisonBMI(c, bmi1, bmi2, bmi3, cds, save=TRUE)
 
@@
 
+```
+
+```{r}
+visCDComparison(c, bmi1, 'No Obesity', 'cancre')
+visCDComparison(c, bmi1, 'No Obesity', 'diabe', save=TRUE)
+visCDComparison(c, bmi1, 'No Obesity', 'stroke')
+visCDComparison(c, bmi1, 'No Obesity', 'hearte')
+visCDComparison(c, bmi1, 'No Obesity', 'lunge')
+visCDComparison(c, bmi1, 'No Obesity', 'employed')
+visCDComparison(c, bmi1, 'No Obesity', 'retired')
+visCDComparison(c, bmi1, 'No Obesity', 'anyadl')
+visCDComparison(c, bmi1, 'No Obesity', 'anyiadl')
 ```

### 2021-05-10 FEM_R/calculate_dflys_from_summary.Rmd
Found a better way to calculate disability free and disease free life years from summary

@@
 # Baselines
-base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
+#base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
 cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')
@@
 # Hearte (cohort and pop)
-#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
-heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
+heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
+#heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
 
@@
 
-# BMI
-nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
-obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
+# BMI (cohort)
+#nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
+#obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
 obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')
 
-# Drinking
-nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
-noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
+# Drinking (pop)
+nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
+nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
+#noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
 ```
 
-# Disability Free Life Years
+# IDEA!!!
+
+Think we need to get the average of disabled var and multiply by the total endpop of corresponding year.
+This way, we'll calculate the average number of diseased (or disabled) years, and then we can do the 
+opposite calculation to get non-diseased or non-disabled.
+
+Got this idea from looking at the `summary_file_statistics.do` file in input_data/detailed_output/, lines 72-77. Check out the whole idea, this script goes:
+
+- Calculate total_lifeyear for each person m_endpop_xxx * 2
+- Calculate summed_total_lifeyear for whole population ( sum(total_lifeyear) )
+- Calculate average_lifeyear (summed_total_lifeyear / m_endpop_xxx)
+
+Same can be done for disability free life years and disease free life years, see script for details.
+
+## LifeYears
+
+### Functions
+```{r}
+cohortLifeyear <- function(c) {
+  c.tot.lfyrs <- c$m_endpop_all * 2
+  c.sum.tot.lfyrs <- sum(c.tot.lfyrs)
+  c.avg.lfyr <- c.sum.tot.lfyrs / c$m_endpop_all[1]
+  return(c.avg.lfyr)
+}
+
+lifeyearComparison <- function(c, i, int.name) {
+
+  c.avg.lfyr <- cohortLifeyear(c)
+
+  i.avg.lfyr <- cohortLifeyear(i)
+
+  print(int.name)
+  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.lfyr, digits=4)))
+  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.lfyr, digits=4)))
+  print(paste('Increase:                        ', signif(x = (i.avg.lfyr - c.avg.lfyr), digits=4)))
+}
+
+diseaseVarLifeyear <- function(c, var) {
+
+  prevVar <- paste0('p_', var, '_all')
+
+  c.tot.var.lfyrs <- c[prevVar] * c$m_endpop_all * 2
+  c.sum.var.lfyrs <- sum(c.tot.var.lfyrs)
+  c.avg.var.lfyrs <- c.sum.var.lfyrs / c$m_endpop_all[1]
+
+  print(paste0(var, ':    ', signif(c.avg.var.lfyrs, digits=4)))
+}
+
+```
+
+### Cohort
+```{r}
+# Cohort life years
+c <- cohort
+c.ly <- cohortLifeyear(c)
+print(paste0('Cohort Life Years: ', signif(c.ly, digits=4)))
+print('-------------------')
+
+# Cohort disease life years
+diseaseVarLifeyear(c, 'cancre')
+diseaseVarLifeyear(c, 'diabe')
+diseaseVarLifeyear(c, 'hearte')
+diseaseVarLifeyear(c, 'lunge')
+diseaseVarLifeyear(c, 'hibpe')
+diseaseVarLifeyear(c, 'stroke')
+diseaseVarLifeyear(c, 'alzhe')
+diseaseVarLifeyear(c, 'demene')
+diseaseVarLifeyear(c, 'smoken')
+diseaseVarLifeyear(c, 'drink')
+diseaseVarLifeyear(c, 'heavy_smoker')
+diseaseVarLifeyear(c, 'problem_drinker')
+```
+
+### Smoking
+```{r}
+# Cohort
+c <- cohort
+# No smoking
+s <- smok
+
+lifeyearComparison(c, s, 'No Smoking')
+```
+
+### Heart Disease
+```{r}
+# Cohort
+c <- cohort
+# No Heart Disease
+h <- heart_c
+
+lifeyearComparison(c, h, 'No Heart Disease')
+```
+
+### Obesity
+```{r}
+c <- cohort
+o <- obese1
+lifeyearComparison(c, o, 'No Obesity')
+```
+
+### Drink
+```{r}
+c <- cohort
+d <- nodrink_c
+lifeyearComparison(c, d, 'No Drink')
+```
+
+## Disability
+
+### Functions
+```{r}
+disabilityLifeYearsCohort <- function(c) {
+  total_dis <- c$a_disabled_all * c$m_endpop_all
+  sum_tot_dis <- sum(total_dis)
+  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
+  return(avg_dis)
+}
+
+disabilityLifeYearComp <- function(c, i, int.name) {
+
+  c.avg.dis <- disabilityLifeYearsCohort(c)
+
+  i.avg.dis <- disabilityLifeYearsCohort(i)
+
+  print(int.name)
+  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+
+disabilityFreeLifeYearsCohort <- function(c) {
+  total_disfree <- c$a_not_disabled_all * c$m_endpop_all
+  sum_tot_disfree <- sum(total_disfree)
+  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
+  return(avg_disfree)
+}
+```
+
+### Cohort
+```{r}
+# Cohort life years
+c <- cohort
+c.dly <- disabilityLifeYearsCohort(c)
+print(paste0('Cohort Disabled Life Years: ', signif(c.dly, digits=4)))
+print('-------------------')
+c.dfly <- disabilityFreeLifeYearsCohort(c)
+print(paste0('Cohort Disability Free Life Years: ', signif(c.dfly, digits=4)))
+
+# Cohort disease life years
+#diseaseVarLifeyear(c, 'cancre')
+#diseaseVarLifeyear(c, 'diabe')
+#diseaseVarLifeyear(c, 'hearte')
+#diseaseVarLifeyear(c, 'lunge')
+#diseaseVarLifeyear(c, 'hibpe')
+#diseaseVarLifeyear(c, 'stroke')
+#diseaseVarLifeyear(c, 'alzhe')
+#diseaseVarLifeyear(c, 'demene')
+#diseaseVarLifeyear(c, 'smoken')
+#diseaseVarLifeyear(c, 'drink')
+#diseaseVarLifeyear(c, 'heavy_smoker')
+#diseaseVarLifeyear(c, 'problem_drinker')
+```
+
+### Smoking
+```{r}
+c <- cohort
+# No Heart Disease
+s <- smok
+
+total_dflys_s <- s$a_disabled_all * s$m_endpop_all
+sum_tot_dflys_s <- sum(total_dflys_s)
+avg_dflys_s <- sum_tot_dflys_s / s$m_endpop_all[1]
+
+avg_dflys
+avg_dflys_s
+
+```
+
+### Heart Disease
+
+```{r}
+
+c <- cohort
+
+total_dflys <- c$a_disabled_all * c$m_endpop_all
+sum_tot_dflys <- sum(total_dflys)
+avg_dflys <- sum_tot_dflys / c$m_endpop_all[1]
+
+# No Heart Disease
+h <- heart_c
+
+total_dflys_h <- h$a_disabled_all * h$m_endpop_all
+sum_tot_dflys_h <- sum(total_dflys_h)
+avg_dflys_h <- sum_tot_dflys_h / h$m_endpop_all[1]
+
+avg_dflys
+avg_dflys_h
+
+```
+
+### Not Disabled
+```{r}
+
+disabilityFreeComparison <- function(c, i, int.name) {
+  tot.df <- c$a_not_disabled_all * c$m_endpop_all
+  sum.tot.df <- sum(tot.df)
+  avg.df <- sum.tot.df / c$m_endpop_all[1]
+
+  tot.df.i <- i$a_not_disabled_all * i$m_endpop_all
+  sum.tot.df.i <- sum(tot.df.i)
+  avg.df.i <- sum.tot.df.i / i$m_endpop_all[1]
+
+  print(int.name)
+  print(paste('Cohort Average Disability Free:', signif(avg.df, digits=4)))
+  print(paste('Intervention Disability Free:', signif(avg.df.i, digits=4)))
+  print(paste('Difference:', signif(x = (avg.df.i - avg.df), digits=4)))
+}
+
+c <- cohort
+# No Heart Disease
+h <- heart_c
+
+disabilityFreeComparison(c, h, 'No Heart Disease')
+
+```
+
+```{r}
+c <- cohort
+s <- smok
+
+disabilityFreeComparison(c, s, 'No Smoking')
+```
+
+# Defunkt
+## Disability Free Life Years
 
@@
 
-## Try again, this time with the correct variable (not disabled)
+### Try again, this time with the correct variable (not disabled)
 

### 2021-05-28 FEM_R/ELSA_data_check.Rmd
Lots of little changes for ROC validation, couple of R scripts focused on calculating DFLYs

@@
 
+rm(cancre, diabe, lunge, smoken, smokev, smokef, drink)
+
 ```
+
+```{r}
+
+# Now get the proportion of missing data in variables in raw stock file
+stock <- read_dta('ELSA_stock_preImpute.dta')
+
+```
+
+```{r}
+
+pMiss(stock$age)
+pMiss(stock$male)
+pMiss(stock$bmi)
+pMiss(stock$logbmi)
+pMiss(stock$smoken)
+pMiss(stock$smokev)
+pMiss(stock$educ)
+
+print('-----------')
+
+pMiss(stock$cancre)
+pMiss(stock$diabe)
+pMiss(stock$hearte)
+pMiss(stock$stroke)
+pMiss(stock$lunge)
+pMiss(stock$hibpe)
+
+```

### 2021-05-28 FEM_R/calculate_dflys_from_summary.Rmd
Lots of little changes for ROC validation, couple of R scripts focused on calculating DFLYs

@@
 diseaseVarLifeyear(c, 'demene')
-diseaseVarLifeyear(c, 'smoken')
-diseaseVarLifeyear(c, 'drink')
-diseaseVarLifeyear(c, 'heavy_smoker')
-diseaseVarLifeyear(c, 'problem_drinker')
+#diseaseVarLifeyear(c, 'smoken')
+#diseaseVarLifeyear(c, 'drink')
+#diseaseVarLifeyear(c, 'heavy_smoker')
+#diseaseVarLifeyear(c, 'problem_drinker')
 ```
@@
 disabilityLifeYearComp <- function(c, i, int.name) {
-
   c.avg.dis <- disabilityLifeYearsCohort(c)
-
   i.avg.dis <- disabilityLifeYearsCohort(i)
-
   print(int.name)
@@
 }
+
+disabilityFreeLifeYearsComp <- function(c, i, int.name) {
+  c.avg.dis <- disabilityFreeLifeYearsCohort(c)
+  i.avg.dis <- disabilityFreeLifeYearsCohort(i)
+  print(int.name)
+  print(paste('Cohort Average Disability Free Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Disability Free Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+
+disVarDFLY <- function(c, var) {
+  prevVar <- paste0('p_', var, '_all')
+
+  c.tot.var.dfly <- c[prevVar] * c$a_not_disabled_all * c$m_endpop_all
+  c.sum.var.dfly <- sum(c.tot.var.dfly)
+  c.avg.var.dfly <- c.sum.var.dfly / c$m_endpop_all[1]
+
+  print(paste0(var, ':    ', signif(c.avg.var.dfly, digits=4)))
+}
 ```
@@
 print(paste0('Cohort Disability Free Life Years: ', signif(c.dfly, digits=4)))
+print('-------------------')
 
-# Cohort disease life years
-#diseaseVarLifeyear(c, 'cancre')
-#diseaseVarLifeyear(c, 'diabe')
-#diseaseVarLifeyear(c, 'hearte')
-#diseaseVarLifeyear(c, 'lunge')
-#diseaseVarLifeyear(c, 'hibpe')
-#diseaseVarLifeyear(c, 'stroke')
-#diseaseVarLifeyear(c, 'alzhe')
-#diseaseVarLifeyear(c, 'demene')
-#diseaseVarLifeyear(c, 'smoken')
-#diseaseVarLifeyear(c, 'drink')
-#diseaseVarLifeyear(c, 'heavy_smoker')
-#diseaseVarLifeyear(c, 'problem_drinker')
+# Cohort disability life years
+disVarDFLY(c, 'cancre')
+disVarDFLY(c, 'diabe')
+disVarDFLY(c, 'hearte')
+disVarDFLY(c, 'lunge')
+disVarDFLY(c, 'hibpe')
+disVarDFLY(c, 'stroke')
+disVarDFLY(c, 'alzhe')
+disVarDFLY(c, 'demene')
+#disVarDFLY(c, 'smoken')
+#disVarDFLY(c, 'drink')
+#disVarDFLY(c, 'heavy_smoker')
+#disVarDFLY(c, 'problem_drinker')
 ```
@@
 
-total_dflys_s <- s$a_disabled_all * s$m_endpop_all
-sum_tot_dflys_s <- sum(total_dflys_s)
-avg_dflys_s <- sum_tot_dflys_s / s$m_endpop_all[1]
-
-avg_dflys
-avg_dflys_s
-
+disabilityFreeLifeYearsComp(c, s, 'No Smoking')
 ```
@@
 ```{r}
-
 c <- cohort
-
-total_dflys <- c$a_disabled_all * c$m_endpop_all
-sum_tot_dflys <- sum(total_dflys)
-avg_dflys <- sum_tot_dflys / c$m_endpop_all[1]
-
 # No Heart Disease
 h <- heart_c
-
-total_dflys_h <- h$a_disabled_all * h$m_endpop_all
-sum_tot_dflys_h <- sum(total_dflys_h)
-avg_dflys_h <- sum_tot_dflys_h / h$m_endpop_all[1]
-
-avg_dflys
-avg_dflys_h
-
+disabilityFreeLifeYearsComp(c, h, 'No Heart Disease')
+cohortLifeyear(c)
 ```
@@
 ```{r}
-
-disabilityFreeComparison <- function(c, i, int.name) {
-  tot.df <- c$a_not_disabled_all * c$m_endpop_all
-  sum.tot.df <- sum(tot.df)
-  avg.df <- sum.tot.df / c$m_endpop_all[1]
-
-  tot.df.i <- i$a_not_disabled_all * i$m_endpop_all
-  sum.tot.df.i <- sum(tot.df.i)
-  avg.df.i <- sum.tot.df.i / i$m_endpop_all[1]
-
-  print(int.name)
-  print(paste('Cohort Average Disability Free:', signif(avg.df, digits=4)))
-  print(paste('Intervention Disability Free:', signif(avg.df.i, digits=4)))
-  print(paste('Difference:', signif(x = (avg.df.i - avg.df), digits=4)))
-}
-
 c <- cohort
@@
 
-disabilityFreeComparison(c, h, 'No Heart Disease')
-
+disabilityFreeLifeYearsComp(c, h, 'No Heart Disease')
 ```

### 2021-05-28 FEM_R/dfly_detailed_output.Rmd
Lots of little changes for ROC validation, couple of R scripts focused on calculating DFLYs

---
title: "R Notebook"
output: html_notebook
---

This script will attempt a couple of things using detailed output data files. We will attempt:
- Disability free life years (and disease free)
- Treatment on the treated analysis

First we're going to have a go at replicating the file tot_test.do, saved in E_FEM_clean/Bryan_scripts

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(ggpubr)
#require(plyr)
#require(data.table)

theme('minimal')
```

Setup is complete (including the working directory), so let's read in some of the data files.

We are using detailed output, so need the appended data files that should be in the output folders.

```{r}
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_append.dta')
heart <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_append.dta')
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_append.dta')
```

Now start with the cohort file, and try to calculate a host of things like life years, dflys, and other things like that.

Generate stats for lung disease and heart disease.

```{r}

# generate flag for initial wave with disease and year died (if known)
new <- cohort %>%
          group_by(hhidpn, mcrep, year) %>%

```

```{r}

aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot.dta')

```

```{r}

educ <- aData %>%
          group_by(educ) %>%
          summarise(improv.ly = mean(improved_lifeyears), improv.dfly = mean(improved_dflys))

educ

```

```{r}
educ2 <- aData %>%
          group_by(educ) %>%
          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))

educ2
```

### 2021-05-28 FEM_R/summary_statistics_forPaper.Rmd
Lots of little changes for ROC validation, couple of R scripts focused on calculating DFLYs

---
title: "Starting Populations Summary Stats"
output: html_notebook
---

# Setup

## WorkDir & Libraries
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(ggplot2)
require(Hmisc)

rm(workingDir)
```

## Data
```{r}

stock <- read_dta('ELSA_stock.dta')
repl <- read_dta('ELSA_repl.dta')
trans <- read_dta('ELSA_transition.dta')

```

# Summary Statistics
## Stock
```{r}
# What do we need to know?
# Population average of:
# - Year
# - Age
# - Gender
# - BMI
#   - % Distribution in BMI categories (<25, 25-30, >30)
# - Smoking Status
#   - Ever Smoked
#   - Current Smoker
# - Education Level
#   - Level 1
#   - Level 2
#   - Level 3
# - Initial Disease Prevalence
#   - Cancer
#   - Diabetes
#   - Heart Disease
#   - MORE!

# Try this:
# https://thatdatatho.com/easily-create-descriptive-summary-statistic-tables-r-studio/

stock.sum <- stock %>%
            summarise(count = n(),
                      year = median(entry, na.rm = TRUE),
                      wave = median(wave, na.rm = TRUE),
                      age.mean = mean(age, na.rm = TRUE),
                      age.sd = sd(age, na.rm = TRUE),
                      male = mean(male, na.rm = TRUE),
                      bmi = mean(bmi, na.rm = TRUE))

stock.sum

#bmi.cut <- cut(stock$bmi, breaks = c(20, 25, 30))
```

# Weighted Means
```{r}
weighted.mean(stock$male, stock$weight)

weighted.mean(stock$cancre, stock$weight)
weighted.mean(stock$diabe, stock$weight)
weighted.mean(stock$hearte, stock$weight)
weighted.mean(stock$stroke, stock$weight)
weighted.mean(stock$lunge, stock$weight)
weighted.mean(stock$hibpe, stock$weight)

```

```{r}
s2 <- stock

s2$bmi_cat <- cut(s2$bmi, breaks = c(25, 30))

#weighted.mean()
```

```{r}
require(survey)

stock.design <- svydesign(id = ~idauniq,
                          weights = ~weight,
                          data = stock)

svymean(~male, stock.design, na.rm=TRUE)
svymean(~cancre, stock.design, na.rm=TRUE)
svymean(~age, stock.design, na.rm=TRUE)

print('Standard deviation')
dev <- 0.1444 * sqrt(length(stock.design$variables))
dev

```

```{r}
require(survey)
weighted.survey.means <- function(init.pop, transition=FALSE) {

  if(!transition) {
    design <- svydesign(id = ~idauniq,
                      weights = ~weight,
                      data = init.pop)
  } else if(transition) {
    # Error from missing cwtresp var - drop all that are missing to solve
    init.pop <- init.pop[complete.cases(init.pop$cwtresp),]
    design <- svydesign(id = ~idauniq,
                      weights = ~cwtresp,
                      data = init.pop)
  }
  print(svymean(~age, design, na.rm=TRUE))
  print(svyvar(~age, design, na.rm=TRUE))
  print(svymean(~male, design, na.rm=TRUE))
  print(svymean(~bmi, design, na.rm=TRUE))
  print(svyvar(~bmi, design, na.rm=TRUE))
  print(svymean(~smoken, design, na.rm=TRUE))
  print(svymean(~smokev, design, na.rm=TRUE))
  print(svymean(~hsless, design, na.rm=TRUE))
  print(svymean(~college, design, na.rm=TRUE))
  print('---------------------')
  print(svymean(~cancre, design, na.rm=TRUE))
  print(svymean(~diabe, design, na.rm=TRUE))
  print(svymean(~hearte, design, na.rm=TRUE))
  print(svymean(~stroke, design, na.rm=TRUE))
  print(svymean(~lunge, design, na.rm=TRUE))
  print(svymean(~hibpe, design, na.rm=TRUE))
}

weighted.survey.means(stock)
```

```{r}
# Replenishing
weighted.survey.means(repl)
```

```{r}
# Transition
weighted.survey.means(trans, transition=TRUE)
```

```{r}
# Now calculate incidence rates of chronic diseases (& some other stuff)
# incidence = total no. new cases / number of people at risk

# I think we'll need to group by year (or wave), and get counts of chronic disease ==1

getIncidenceRate <- function(t, var) {

  l2var <- paste0('l2', var)

  select.vars <- c(var, l2var)

  t <- select(t, year, all_of(select.vars))
  t$incidence.counts <- (t[,var] == 1) & (t[,l2var] == 0)

  t <- t %>%
          summarise(n = n(), newIn = sum(incidence.counts, na.rm=TRUE)) %>%
          mutate(incidence.rate = (newIn / n) * 100)

  return(t)
}

getIncidenceRate(trans, 'cancre')
getIncidenceRate(trans, 'diabe')
getIncidenceRate(trans, 'hearte')
getIncidenceRate(trans, 'stroke')
getIncidenceRate(trans, 'lunge')
getIncidenceRate(trans, 'hibpe')
```

### 2021-06-07 FEM_R/dfly_detailed_output.Rmd
Trying some big changes for ROC curves and calculated dflys in an R script

@@
 theme('minimal')
+rm(workingDir)
 ```
 
-Setup is complete (including the working directory), so let's read in some of the data files.
+Setup is complete, so let's read in some of the data files.
 
@@
 new <- cohort %>%
-          group_by(hhidpn, mcrep, year) %>%
+          group_by(hhidpn, mcrep) %>%
+          arrange(year, .by_group = TRUE) %>%
+          mutate(hearte_init = min(year[hearte == 1]), hearte_died = max(year[hearte == 1])) %>%
+          mutate(lunge_init = min(year[lunge == 1]), lunge_died = max(year[lunge == 1]))
+
+new$hearte_init[is.infinite(new$hearte_init)] <- 0
+new$hearte_died[is.infinite(new$hearte_died)] <- 0
+new$lunge_init[is.infinite(new$lunge_init)] <- 0
+new$lunge_died[is.infinite(new$lunge_died)] <- 0
 
 ```
+
+# Using output from stata script treatment_on_treated.do or some version of it
 

### 2021-06-24 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Lots of little changes getting ready to submit the paper to IJM

---
title: "R Notebook"
output: html_notebook
---

# Setup

## Libraries
```{r "setup", include=FALSE}
require(tidyverse)
require(haven)
require(ggplot2)
require(Hmisc)

theme_set(theme_bw())

workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

rm(workingDir)
```

## Starting Populations
```{r}
stock <- read_dta('input_data/ELSA_stock.dta')
repl <- read_dta('input_data/ELSA_repl.dta')
trans <- read_dta('input_data/ELSA_transition.dta')
```

# Methods

## Summary Stats for Starting Populations

### Function Declaration
```{r}
require(survey)
weighted.survey.means <- function(init.pop, transition=FALSE) {

  if(!transition) {
    design <- svydesign(id = ~idauniq,
                      weights = ~weight,
                      data = init.pop)
  } else if(transition) {
    # Error from missing cwtresp var - drop all that are missing to solve
    init.pop <- init.pop[complete.cases(init.pop$cwtresp),]
    design <- svydesign(id = ~idauniq,
                      weights = ~cwtresp,
                      data = init.pop)
  }
  print(svymean(~age, design, na.rm=TRUE))
  print(svyvar(~age, design, na.rm=TRUE))
  print(svymean(~male, design, na.rm=TRUE))
  print(svymean(~bmi, design, na.rm=TRUE))
  print(svyvar(~bmi, design, na.rm=TRUE))
  print(svymean(~smoken, design, na.rm=TRUE))
  print(svymean(~smokev, design, na.rm=TRUE))
  print(svymean(~hsless, design, na.rm=TRUE))
  print(svymean(~college, design, na.rm=TRUE))
  print('---------------------')
  print(svymean(~cancre, design, na.rm=TRUE))
  print(svymean(~diabe, design, na.rm=TRUE))
  print(svymean(~hearte, design, na.rm=TRUE))
  print(svymean(~stroke, design, na.rm=TRUE))
  print(svymean(~lunge, design, na.rm=TRUE))
  print(svymean(~hibpe, design, na.rm=TRUE))
  print(svymean(~alzhe, design, na.rm=TRUE))
  print(svymean(~demene, design, na.rm=TRUE))
  print(svymean(~hchole, design, na.rm=TRUE))
}
```

### Stock
```{r}
weighted.survey.means(stock)
```

### Replenishing
```{r}
weighted.survey.means(repl)
```

### Transition
```{r}
weighted.survey.means(trans, transition=TRUE)
```

#### Incidence
```{r}
getIncidenceRate <- function(t, var) {

  l2var <- paste0('l2', var)

  select.vars <- c(var, l2var)

  t <- select(t, year, all_of(select.vars))
  t$incidence.counts <- (t[,var] == 1) & (t[,l2var] == 0)

  t <- t %>%
          summarise(n = n(), newIn = sum(incidence.counts, na.rm=TRUE)) %>%
          mutate(incidence.rate = (newIn / n) * 100)

  t$var <- var

  return(t)
}

getIncidenceRate(trans, 'cancre')
getIncidenceRate(trans, 'diabe')
getIncidenceRate(trans, 'hearte')
getIncidenceRate(trans, 'stroke')
getIncidenceRate(trans, 'lunge')
getIncidenceRate(trans, 'hibpe')
getIncidenceRate(trans, 'alzhe')
getIncidenceRate(trans, 'demene')
getIncidenceRate(trans, 'hchole')
```

# Validation

## Handovers

Handover plots are generated from a Stata script that can be called via a Make rule (`make handovers`). The core_complete set of scenarios must be run prior to this with the `55p_f_l` & `55p_m_l` subpopulations.

## ROC plots

ROC plots are also generated via a Stata script. The entire workflow for generating ROC plots (including running two scenarios for 500 repetitions) is handled by another Make rule (`make roc`).

## AgeUK validation

```{r echo = FALSE}
#AgeUK
male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE.csv')
female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE.csv')

# FEM 2014 detailed output
#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
## extract males and females
#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
#fFEM_2014 <- FEM_2014 %>% filter(male == 0)

# FEM 2014 summary
#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')

# Keep only year 2014 from FEM summary
FEM <- filter(FEM, year == 2014)

# Specify the age groups to extract from FEM
ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
# Add m and f for male and female variables
m_ages <- paste0('m_', ages)
f_ages <- paste0('f_', ages)

# Collect the age suffixes into a vector
keep_vars <- c(m_ages, f_ages)
# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')

# Don't like using loops but this one was just too simple to bother with figuring out the apply version
final <- c()
ticker <- 1
# Generating the column (variable) names we want to keep in a loop
for(item in diseases) {
  for(anotherItem in keep_vars) {
    final[ticker] <- paste0('p_', item, '_', anotherItem)
    ticker <- ticker + 1
  }
}

# Keep the right vars using varnames generated in loop
FEM <- FEM %>%
          select(c('year', all_of(final)))

# Rename some vars to remove 'p_' before merging
FEM <- FEM %>%
          rename_at(vars(starts_with('p_')),
                    funs(str_replace(., 'p_', ''))) %>%
          select(!year)

pivoted.FEM <- pivot_longer(data = FEM,
                            cols = hibpe_m_6064:asthmae_f_100p, 
                            names_to = c('Condition', 'Gender', 'Age_Group'), 
                            names_sep = '_')

# Convert value from fraction to percentage
pivoted.FEM$value <- pivoted.FEM$value * 100

# Rename cols to enable reshaping
male_au <- male_au %>%
                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
                select(!Total)

female_au <- female_au %>%
                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
                select(!Total)

# Now need to reshape these datasets to long format
pivoted.male <- pivot_longer(data = male_au,
                             cols = m_6064:m_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

pivoted.female <- pivot_longer(data = female_au,
                             cols = f_6064:f_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

# Now row bind these frames together, combining male and female data
ageUK <-rbind(pivoted.male, pivoted.female)

# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
# extract all heart related conditions for combining
#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
# Aggregate the values by gender and age group
#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
# Rename columns and add condition column back
#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
#ageUK_heart$Condition <- 'Heart Problems'

# Now drop the heart problems from original ageUK and replace with aggregate
ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
#ageUK <- rbind(ageUK, ageUK_heart)

# Rename a few just for completeness
#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)

# Finally, add column to specify that this is ageUK data
ageUK$source <- "ageUK"

# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')

# Make copy
newFEM <- pivoted.FEM

# Rename all the conditions to fit ageUK
# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
#       Therefore only validation from this will be that FEM is higher than ageUK
#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)

#force Make to run this again

# Add column to specify this is FEM data
newFEM$source <- "FEM"

# Now row bind the datasets
combined <- rbind(newFEM, ageUK)
# Specify factor levels for age group
combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
# Sort
combined <- arrange(combined, source, Condition, Gender, Age_Group)

# Function to filter the combined dataset and visualise
visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {

  if(gender == 'm') {
    full_gender <- 'Male'
  }
  else if(gender == 'f') {
    full_gender <- 'Female'
  }

  # filter by gender and condition
  data <- data %>%
            filter(Gender == gender) %>%
            filter(Condition == cd)

  # now plot!
  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = paste0(cd, '_', gender, '.png'),
         plot = p,
         path = path,
         width = 6,
         height = 4)

  return(p)
}

```

```{r}
# Collect all conditions (and both genders) into a list, and apply the function above to that list
cd_list <- unique(combined$Condition)
gender_list <- unique(combined$Gender)

for(cd in cd_list) {
  for(gen in gender_list) {
    plot <- visAgeUKValid(combined, cd, gen)
    print(plot)
  }
}
```

## Smoking
```{r}
FEM_base <- read_dta('output/COMPLETE/ELSA_core_base/ELSA_core_base_summary.dta')

# Smoking prevalence stats
smok.5059 <- 15.2
smok.60p <- 10.2

# FEM summary, filter for 2018 and 60p
#FEM.smok <- FEM_sum
FEM.smok <- FEM_base
FEM.smok <- FEM.smok %>%
                filter(year == 2018) %>%
                select(contains('smoken')) %>%
                select(contains('60p'), contains('5059')) %>%
                select(contains('p_'))

FEM.smok <- FEM.smok * 100

smok.combined <- data.frame(FEM = FEM.smok[1,1], ASH = 10.2)
smok.combined <- data.frame(source = c('FEM', 'FEM', 'ASH', 'ASH'), age = c('50 - 59', '60+', '50 - 59', '60+'), value = c(FEM.smok[1,2], FEM.smok[1,1], 15.2, 10.2))

smok.plot <-ggplot(data = smok.combined, aes(x = age, y = value)) +
                geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
                labs(x = ' ', y = 'Prevalence (%)')

ggsave(filename = 'smoking_prev_2018.png',
       plot = smok.plot,
       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/smok_prev/',
       width = 7,
       height = 5)

smok.plot

```

# Results

## Heart Disease Intervention
```{r}
compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)

  aDeath <- aDeath %>%
                dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
                reshape2::melt(id.var = 'Year', value.name = 'Age')

  p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
        geom_smooth(lwd = 0.5, se=FALSE) +
        labs(x = 'Year', y = 'Age at Death') +

  if(save) {
    ggsave(filename = 'hearte_dAge_all.png',
         plot = p1,
         path = path)
  }

  return(p1)
}
```

### Life-Expectancy
```{r}
b <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
h_p <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

compareDeathAge(b, h_p, 'No Heart Disease', save=FALSE)
```

#### With Confidence Intervals
```{r}
b_rep <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_by_rep.dta')
h_rep <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_by_rep.dta')

compareDeathAgeConfints <- function(b, i, int_name=FALSE, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  aDet <- data.frame(b$year, b$rep, b$d_age_all, i$d_age_all)

  aDet2 <- aDet %>% 
            dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_all, !!quo_name(int_name) := i.d_age_all) %>%
            pivot_longer(
                cols = Baseline:'No Heart Disease',
                names_to = 'Source',
                values_to = 'Life_Exp'
            )

  p <- ggplot(data = aDet2, aes(x = Year, y = Life_Exp, color = Source)) +
          geom_smooth(lwd = 0.5) +
          labs(y = 'Life Expectancy')

  if(save) {
    ggsave(filename = 'hearte_dAge_all.png',
         plot = p,
         path = path)
  }

  p
}

compareDeathAgeConfints(b_rep, h_rep, int_name = 'No Heart Disease', save=TRUE)

```

### Split by Activity Level
```{r}
b <- b_rep
h <- h_rep

d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
d.exstat1 <- d.exstat1 %>%
                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
d.exstat1.pivot <- pivot_longer(d.exstat1,
                              cols = Baseline:Intervention,
                              names_to = 'Source',
                              values_to = 'Life_Expectancy')
p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
      geom_smooth(lwd = 0.5) +
      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
p1

d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
d.exstat2 <- d.exstat2 %>%
                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
d.exstat2.pivot <- pivot_longer(d.exstat2,
                              cols = Baseline:Intervention,
                              names_to = 'Source',
                              values_to = 'Life_Expectancy')
p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
      geom_smooth(lwd = 0.5) +
      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
      scale_fill_grey()
p2

d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
d.exstat3 <- d.exstat3 %>%
                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
d.exstat3.pivot <- pivot_longer(d.exstat3,
                              cols = Baseline:Intervention,
                              names_to = 'Source',
                              values_to = 'Life_Expectancy')
p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
      geom_smooth(lwd = 0.5) +
      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
p3

path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'

ggsave(filename = 'hearte_dAge_exstat1.png',
         plot = p1,
         path = path)

ggsave(filename = 'hearte_dAge_exstat2.png',
         plot = p2,
         path = path)

ggsave(filename = 'hearte_dAge_exstat3.png',
         plot = p3,
         path = path)

```

## Smoking Intervention
```{r}
compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$rep, cohort$Survival, intervention$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  #survivals <- survivals %>%
  #               dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
  #               melt(id.var = 'Age', value.name = 'Survival')

  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Rep' = cohort.rep, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
                 pivot_longer(
                   cols = Baseline:'No Smoking',
                   names_to = 'Source',
                   values_to = 'Life_Exp'
                 )

  p1 <-ggplot(data = survivals, aes(x=Age, y=Life_Exp, color = Source)) +
        geom_smooth(lwd = 0.5) +
        labs(x = 'Age', y = 'Survival')

  if(save) {
    ggsave(filename = paste0(int_name, '_survival_curve.png'),
         plot = p1,
         path = path)
  }

  return(p1)
}
```

### Survival Curve
```{r}
c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')

compareSurvivalCurve(c, s, 'No Smoking', save=FALSE)
```

#### Confidence Intervals
```{r}
compareSurvivalCurveConfints <- function(cohort, intervention, int_name, save=FALSE,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$rep, cohort$Survival, intervention$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Rep' = cohort.rep, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
                 melt(id.var = 'Age', value.name = 'Survival')

  p <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
        geom_line() +
        labs(x = 'Age', y = 'Survival')

  if(save) {
    ggsave(filename = paste0(int_name, '_survival_curve.png'),
         plot = p,
         path = path)
  }

  return(p)
}

c_rep <- read_dta('output/COMPLETE/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
s_rep <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')

compareSurvivalCurveConfints(c_rep, s_rep)
```

### Lung Disease Prevalence All
```{r}
c <- c
s <- s

# Calculate age and survival curves
c$Age <- (c$year - 2012) + 50
s$Age <- (s$year - 2012) + 50
c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
s$Survival <- s$m_endpop_all / s$m_endpop_all[1]

d.all <- data.frame(c$Age, c$p_lunge_all, s$p_lunge_all)
d.all <- d.all %>%
                dplyr::rename('Age' = c.Age, 'Baseline' = c.p_lunge_all, 'Intervention' = s.p_lunge_all)
d.all.pivot <- pivot_longer(d.all,
                              cols = Baseline:Intervention,
                              names_to = 'Source',
                              values_to = 'Prevalence')
p1 <- ggplot(data = d.all.pivot, aes(x = Age, y = Prevalence, color = Source)) +
      geom_smooth(lwd = 0.5, se=FALSE) +
      labs(x = 'Age', y = 'Prevalence')
p1

path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'

ggsave(filename = 'smok_lunge_all.png',
         plot = p1,
         path = path)
```

### Life Years & DFLYS
```{r}
aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot_test.dta')

all2 <- aData %>%
          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))

all2
```

#### Split by Education
```{r}
educ2 <- aData %>%
          group_by(educ) %>%
          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))

educ2
```

### 2021-06-24 FEM_R/dfly_detailed_output.Rmd
Lots of little changes getting ready to submit the paper to IJM

--- 2021-05-28 FEM_R/dfly_detailed_output.Rmd
@@
 
-theme('minimal')
+rm(workingDir)
 ```
 
-Setup is complete (including the working directory), so let's read in some of the data files.
+Setup is complete, so let's read in some of the data files.
 
@@
 cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_append.dta')
-heart <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_append.dta')
+#heart <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_append.dta')
 smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_append.dta')
@@
 # generate flag for initial wave with disease and year died (if known)
-new <- cohort %>%
-          group_by(hhidpn, mcrep, year) %>%
+new.c <- cohort %>%
+          group_by(hhidpn, mcrep) %>%
+          arrange(year, .by_group = TRUE) %>%
+          mutate(hearte_init = min(year[hearte == 1]), hearte_died = max(year[hearte == 1])) %>%
+          mutate(lunge_init = min(year[lunge == 1]), lunge_died = max(year[lunge == 1]))
+
+# Replace all infinite values with 0 (infinites generated in mutate calls above)
+new$hearte_init[is.infinite(new$hearte_init)] <- 0
+new$hearte_died[is.infinite(new$hearte_died)] <- 0
+new$lunge_init[is.infinite(new$lunge_init)] <- 0
+new$lunge_died[is.infinite(new$lunge_died)] <- 0
+
+# now generate flag variable if
+
+#rm(new)
+```
+
+## Using output from stata script treatment_on_treated.do or some version of it
+
+```{r}
+
+aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot_test.dta')
+
+```
+
+### Whole Population
+```{r}
+
+all <- aData %>%
+        summarise(improv.ly = mean(improved_lifeyears), improv.dfly = mean(improved_dflys))
+
+all
 
@@
 ```{r}
+all2 <- aData %>%
+          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
+                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))
 
-aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot.dta')
-
+all2
 ```
 
+### By Education
 ```{r}

### 2021-06-24 FEM_R/model_analysis_core.Rmd
Lots of little changes getting ready to submit the paper to IJM

@@
 # Need to set up the working directory
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
+workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/COMPLETE/"
 knitr::opts_knit$set(root.dir = workingDir)
@@
 # Load in whole stock pop, plus the CV2 summary output file
-long <- read_dta('../input_data/ELSA_long.dta')
+long <- read_dta('../../input_data/ELSA_long.dta')
 CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
@@
 #AgeUK
-male_au <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
-female_au <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
+female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
 
@@
 # Function to filter the combined dataset and visualise
-visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
 
@@
        plot = smok.plot,
-       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/',
+       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/',
        width = 7,

### 2021-06-28 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Now automatically generate script with all figures after running core_scen

@@
 require(Hmisc)
+require(reshape2)
 
@@
 ```{r}
-compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
-
-  aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)
-
-  aDeath <- aDeath %>%
-                dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
-                reshape2::melt(id.var = 'Year', value.name = 'Age')
-
-  p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
-        geom_smooth(lwd = 0.5, se=FALSE) +
-        labs(x = 'Year', y = 'Age at Death') +
-
-  if(save) {
-    ggsave(filename = 'hearte_dAge_all.png',
-         plot = p1,
-         path = path)
-  }
-
-  return(p1)
-}
+# compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+#   
+#   aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)
+#   
+#   aDeath <- aDeath %>%
+#                 dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
+#                 reshape2::melt(id.var = 'Year', value.name = 'Age')
+#   
+#   p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
+#         geom_smooth(lwd = 0.5, se=FALSE) +
+#         labs(x = 'Year', y = 'Age at Death') +
+#   
+#   if(save) {
+#     ggsave(filename = 'hearte_dAge_all.png',
+#          plot = p1,
+#          path = path)
+#   }
+#   
+#   return(p1)
+# }
 ```
@@
 ```{r}
-b <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
-h_p <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
-
-compareDeathAge(b, h_p, 'No Heart Disease', save=FALSE)
+# b <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
+# h_p <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
+# 
+# compareDeathAge(b, h_p, 'No Heart Disease', save=FALSE)
 ```
@@
 
+### SplitbyEduc
+```{r}
+b <- b_rep
+h <- h_rep
+
+d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
+d.educ1 <- d.educ1 %>%
+                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
+d.educ1.pivot <- pivot_longer(d.educ1,
+                              cols = Baseline:Intervention,
+                              names_to = 'Source',
+                              values_to = 'Life_Expectancy')
+p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+      geom_smooth(lwd = 0.5) +
+      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
+p1
+
+d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
+d.educ2 <- d.educ2 %>%
+                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
+d.educ2.pivot <- pivot_longer(d.educ2,
+                              cols = Baseline:Intervention,
+                              names_to = 'Source',
+                              values_to = 'Life_Expectancy')
+p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+      geom_smooth(lwd = 0.5) +
+      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
+      scale_fill_grey()
+p2
+
+d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
+d.educ3 <- d.educ3 %>%
+                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
+d.educ3.pivot <- pivot_longer(d.educ3,
+                              cols = Baseline:Intervention,
+                              names_to = 'Source',
+                              values_to = 'Life_Expectancy')
+p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+      geom_smooth(lwd = 0.5) +
+      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
+p3
+
+path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
+
+ggsave(filename = 'hearte_dAge_educ1.png',
+         plot = p1,
+         path = path)
+
+ggsave(filename = 'hearte_dAge_educ2.png',
+         plot = p2,
+         path = path)
+
+ggsave(filename = 'hearte_dAge_educ3.png',
+         plot = p3,
+         path = path)
+
+```
+
 ## Smoking Intervention
 ```{r}
-compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
-                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
-
-  # Calculate age and survival curves
-  cohort$Age <- (cohort$year - 2012) + 50
-  intervention$Age <- (intervention$year - 2012) + 50
-  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
-  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
-
-  survivals <- data.frame(cohort$Age, cohort$rep, cohort$Survival, intervention$Survival)
-
-  # Had a small problem when trying to rename using the variable name given as function argument
-  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
-  #survivals <- survivals %>%
+# compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
+#                                  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+#   
+#   # Calculate age and survival curves
+#   cohort$Age <- (cohort$year - 2012) + 50
+#   intervention$Age <- (intervention$year - 2012) + 50
+#   cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
+#   intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
+#   
+#   survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)
+#   
+#   # Had a small problem when trying to rename using the variable name given as function argument
+#   # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
+#   survivals <- survivals %>%
 #                  dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
 #                  melt(id.var = 'Age', value.name = 'Survival')
-
-  survivals <- survivals %>%
-                 dplyr::rename('Age' = cohort.Age, 'Rep' = cohort.rep, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
-                 pivot_longer(
-                   cols = Baseline:'No Smoking',
-                   names_to = 'Source',
-                   values_to = 'Life_Exp'
-                 )
-
-  p1 <-ggplot(data = survivals, aes(x=Age, y=Life_Exp, color = Source)) +
-        geom_smooth(lwd = 0.5) +
-        labs(x = 'Age', y = 'Survival')
-
-  if(save) {
-    ggsave(filename = paste0(int_name, '_survival_curve.png'),
-         plot = p1,
-         path = path)
-  }
-
-  return(p1)
-}
+#   
+#   p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
+#         geom_line() +
+#         labs(x = 'Age', y = 'Survival')
+#   
+#   if(save) {
+#     ggsave(filename = paste0(int_name, '_survival_curve.png'),
+#          plot = p1,
+#          path = path)
+#   }
+#   
+#   return(p1)
+# }
 ```
@@
 ```{r}
-c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
-s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')
-
-compareSurvivalCurve(c, s, 'No Smoking', save=FALSE)
+# c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')
+# s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
+# 
+# compareSurvivalCurve(c, s, 'No Smoking', save=FALSE)
 ```
@@
                  dplyr::rename('Age' = cohort.Age, 'Rep' = cohort.rep, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
-                 melt(id.var = 'Age', value.name = 'Survival')
+                 pivot_longer(
+                   cols = Baseline:'No Smoking',
+                   names_to = 'Source',
+                   values_to = 'Life_Exp'
+                 )
 
-  p <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
-        geom_line() +
+  p <-ggplot(data = survivals, aes(x=Age, y=Life_Exp, color = Source)) +
+        geom_smooth(lwd = 0.5) +
         labs(x = 'Age', y = 'Survival')
@@
 
-compareSurvivalCurveConfints(c_rep, s_rep)
+compareSurvivalCurveConfints(c_rep, s_rep, 'No Smoking', save=TRUE)
 ```
@@
 ```{r}
-c <- c
-s <- s
+c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')
+s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
 
@@
       geom_smooth(lwd = 0.5, se=FALSE) +
+      labs(x = 'Age', y = 'Prevalence')
+p1
+
+path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'
+
+#ggsave(filename = 'smok_lunge_all.png',
+#         plot = p1,
+#         path = path)
+```
+
+```{r}
+c <- c_rep
+s <- s_rep
+
+# Calculate age and survival curves
+c$Age <- (c$year - 2012) + 50
+s$Age <- (s$year - 2012) + 50
+c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
+s$Survival <- s$m_endpop_all / s$m_endpop_all[1]
+
+d.all <- data.frame(c$Age, c$rep, c$p_lunge_all, s$p_lunge_all)
+d.all <- d.all %>%
+                dplyr::rename('Age' = c.Age, 'Rep' = c.rep, 'Baseline' = c.p_lunge_all, 'Intervention' = s.p_lunge_all)
+d.all.pivot <- pivot_longer(d.all,
+                              cols = Baseline:Intervention,
+                              names_to = 'Source',
+                              values_to = 'Prevalence')
+p1 <- ggplot(data = d.all.pivot, aes(x = Age, y = Prevalence, color = Source)) +
+      geom_smooth(lwd = 0.5) +
       labs(x = 'Age', y = 'Prevalence')

### 2021-06-28 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Fixed crossvalidation script error, also reduced the analysis to bare minimum for speedy runs

@@
 ```{r echo = FALSE}
-#AgeUK
-male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE.csv')
-female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE.csv')
-
-# FEM 2014 detailed output
-#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-## extract males and females
-#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
-
-# FEM 2014 summary
-#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
-
-# Keep only year 2014 from FEM summary
-FEM <- filter(FEM, year == 2014)
-
-# Specify the age groups to extract from FEM
-ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
-# Add m and f for male and female variables
-m_ages <- paste0('m_', ages)
-f_ages <- paste0('f_', ages)
-
-# Collect the age suffixes into a vector
-keep_vars <- c(m_ages, f_ages)
-# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
-
-# Don't like using loops but this one was just too simple to bother with figuring out the apply version
-final <- c()
-ticker <- 1
-# Generating the column (variable) names we want to keep in a loop
-for(item in diseases) {
-  for(anotherItem in keep_vars) {
-    final[ticker] <- paste0('p_', item, '_', anotherItem)
-    ticker <- ticker + 1
-  }
-}
-
-# Keep the right vars using varnames generated in loop
-FEM <- FEM %>%
-          select(c('year', all_of(final)))
-
-# Rename some vars to remove 'p_' before merging
-FEM <- FEM %>%
-          rename_at(vars(starts_with('p_')),
-                    funs(str_replace(., 'p_', ''))) %>%
-          select(!year)
-
-pivoted.FEM <- pivot_longer(data = FEM,
-                            cols = hibpe_m_6064:asthmae_f_100p, 
-                            names_to = c('Condition', 'Gender', 'Age_Group'), 
-                            names_sep = '_')
-
-# Convert value from fraction to percentage
-pivoted.FEM$value <- pivoted.FEM$value * 100
-
-# Rename cols to enable reshaping
-male_au <- male_au %>%
-                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
-                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
-                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
-                select(!Total)
-
-female_au <- female_au %>%
-                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
-                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
-                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
-                select(!Total)
-
-# Now need to reshape these datasets to long format
-pivoted.male <- pivot_longer(data = male_au,
-                             cols = m_6064:m_100p,
-                             names_to = c('Gender', 'Age_Group'),
-                             names_sep = '_')
-
-pivoted.female <- pivot_longer(data = female_au,
-                             cols = f_6064:f_100p,
-                             names_to = c('Gender', 'Age_Group'),
-                             names_sep = '_')
-
-# Now row bind these frames together, combining male and female data
-ageUK <-rbind(pivoted.male, pivoted.female)
-
-# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
-# extract all heart related conditions for combining
-#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
-# Aggregate the values by gender and age group
-#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
-# Rename columns and add condition column back
-#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-#ageUK_heart$Condition <- 'Heart Problems'
-
-# Now drop the heart problems from original ageUK and replace with aggregate
-ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-#ageUK <- rbind(ageUK, ageUK_heart)
-
-# Rename a few just for completeness
-#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
-
-# Finally, add column to specify that this is ageUK data
-ageUK$source <- "ageUK"
-
-# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
-#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
-ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
-ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
-
-# Make copy
-newFEM <- pivoted.FEM
-
-# Rename all the conditions to fit ageUK
-# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
-#       Therefore only validation from this will be that FEM is higher than ageUK
-#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
-newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
-newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
-newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
-#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
-#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
-newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
-newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
-
-#force Make to run this again
-
-# Add column to specify this is FEM data
-newFEM$source <- "FEM"
-
-# Now row bind the datasets
-combined <- rbind(newFEM, ageUK)
-# Specify factor levels for age group
-combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
-# Sort
-combined <- arrange(combined, source, Condition, Gender, Age_Group)
-
-# Function to filter the combined dataset and visualise
-visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
-
-  if(gender == 'm') {
-    full_gender <- 'Male'
-  }
-  else if(gender == 'f') {
-    full_gender <- 'Female'
-  }
-
-  # filter by gender and condition
-  data <- data %>%
-            filter(Gender == gender) %>%
-            filter(Condition == cd)
-
-  # now plot!
-  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
-          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
-          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
-
-  # Save plot into in E-FEM_Paper/figures/
-  ggsave(filename = paste0(cd, '_', gender, '.png'),
-         plot = p,
-         path = path,
-         width = 6,
-         height = 4)
-
-  return(p)
-}
+# #AgeUK
+# male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE.csv')
+# female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE.csv')
+# 
+# # FEM 2014 detailed output
+# #FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+# ## extract males and females
+# #mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+# #fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+# 
+# # FEM 2014 summary
+# #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+# FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+# 
+# # Keep only year 2014 from FEM summary
+# FEM <- filter(FEM, year == 2014)
+# 
+# # Specify the age groups to extract from FEM
+# ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# # Add m and f for male and female variables
+# m_ages <- paste0('m_', ages)
+# f_ages <- paste0('f_', ages)
+# 
+# # Collect the age suffixes into a vector
+# keep_vars <- c(m_ages, f_ages)
+# # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+# diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
+# 
+# # Don't like using loops but this one was just too simple to bother with figuring out the apply version
+# final <- c()
+# ticker <- 1
+# # Generating the column (variable) names we want to keep in a loop
+# for(item in diseases) {
+#   for(anotherItem in keep_vars) {
+#     final[ticker] <- paste0('p_', item, '_', anotherItem)
+#     ticker <- ticker + 1
+#   }
+# }
+# 
+# # Keep the right vars using varnames generated in loop
+# FEM <- FEM %>%
+#           select(c('year', all_of(final)))
+# 
+# # Rename some vars to remove 'p_' before merging
+# FEM <- FEM %>%
+#           rename_at(vars(starts_with('p_')),
+#                     funs(str_replace(., 'p_', ''))) %>%
+#           select(!year)
+# 
+# pivoted.FEM <- pivot_longer(data = FEM,
+#                             cols = hibpe_m_6064:asthmae_f_100p, 
+#                             names_to = c('Condition', 'Gender', 'Age_Group'), 
+#                             names_sep = '_')
+# 
+# # Convert value from fraction to percentage
+# pivoted.FEM$value <- pivoted.FEM$value * 100
+# 
+# # Rename cols to enable reshaping
+# male_au <- male_au %>%
+#                 dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
+#                        m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
+#                        m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
+#                 select(!Total)
+# 
+# female_au <- female_au %>%
+#                 dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
+#                        f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
+#                        f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
+#                 select(!Total)
+# 
+# # Now need to reshape these datasets to long format
+# pivoted.male <- pivot_longer(data = male_au,
+#                              cols = m_6064:m_100p,
+#                              names_to = c('Gender', 'Age_Group'),
+#                              names_sep = '_')
+# 
+# pivoted.female <- pivot_longer(data = female_au,
+#                              cols = f_6064:f_100p,
+#                              names_to = c('Gender', 'Age_Group'),
+#                              names_sep = '_')
+# 
+# # Now row bind these frames together, combining male and female data
+# ageUK <-rbind(pivoted.male, pivoted.female)
+# 
+# # Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# # extract all heart related conditions for combining
+# #ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# # Aggregate the values by gender and age group
+# #ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# # Rename columns and add condition column back
+# #colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+# #ageUK_heart$Condition <- 'Heart Problems'
+# 
+# # Now drop the heart problems from original ageUK and replace with aggregate
+# ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+# #ageUK <- rbind(ageUK, ageUK_heart)
+# 
+# # Rename a few just for completeness
+# #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+# #ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+# 
+# # Finally, add column to specify that this is ageUK data
+# ageUK$source <- "ageUK"
+# 
+# # REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+# #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+# ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+# ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
+# 
+# # Make copy
+# newFEM <- pivoted.FEM
+# 
+# # Rename all the conditions to fit ageUK
+# # NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+# #       Therefore only validation from this will be that FEM is higher than ageUK
+# #newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+# newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+# newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+# newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+# #newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+# #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+# newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+# newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
+# 
+# 
+# #force Make to run this again
+# 
+# # Add column to specify this is FEM data
+# newFEM$source <- "FEM"
+# 
+# # Now row bind the datasets
+# combined <- rbind(newFEM, ageUK)
+# # Specify factor levels for age group
+# combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
+# # Sort
+# combined <- arrange(combined, source, Condition, Gender, Age_Group)
+# 
+# # Function to filter the combined dataset and visualise
+# visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
+#   
+#   if(gender == 'm') {
+#     full_gender <- 'Male'
+#   }
+#   else if(gender == 'f') {
+#     full_gender <- 'Female'
+#   }
+#   
+#   # filter by gender and condition
+#   data <- data %>%
+#             filter(Gender == gender) %>%
+#             filter(Condition == cd)
+#   
+#   # now plot!
+#   p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
+#           geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+#           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+#   
+#   # Save plot into in E-FEM_Paper/figures/
+#   ggsave(filename = paste0(cd, '_', gender, '.png'),
+#          plot = p,
+#          path = path,
+#          width = 6,
+#          height = 4)
+#   
+#   return(p)
+# }
 
@@
 
-```{r}
+```{r echo=FALSE}
 # Collect all conditions (and both genders) into a list, and apply the function above to that list
-cd_list <- unique(combined$Condition)
-gender_list <- unique(combined$Gender)
-
-for(cd in cd_list) {
-  for(gen in gender_list) {
-    plot <- visAgeUKValid(combined, cd, gen)
-    print(plot)
-  }
-}
+# cd_list <- unique(combined$Condition)
+# gender_list <- unique(combined$Gender)
+# 
+# for(cd in cd_list) {
+#   for(gen in gender_list) {
+#     plot <- visAgeUKValid(combined, cd, gen)
+#     print(plot)
+#   }
+# }
 ```
@@
 ## Heart Disease Intervention
-```{r}
+```{r echo=FALSE}
 # compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
@@
 ### Life-Expectancy
-```{r}
+```{r echo=FALSE}
 # b <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
@@
 ## Smoking Intervention
-```{r}
+```{r echo=FALSE}
 # compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
@@
 ### Survival Curve
-```{r}
+```{r echo=FALSE}
 # c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')

### 2021-06-28 FEM_R/model_analysis_core.Rmd
Fixed crossvalidation script error, also reduced the analysis to bare minimum for speedy runs

@@
 ```{r echo = FALSE}
-
-#AgeUK
-male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
-female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
-
-# FEM 2014 detailed output
-#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-## extract males and females
-#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
-
-# FEM 2014 summary
-#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
-
-# Keep only year 2014 from FEM summary
-FEM <- filter(FEM, year == 2014)
-
-# Specify the age groups to extract from FEM
-ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
-# Add m and f for male and female variables
-m_ages <- paste0('m_', ages)
-f_ages <- paste0('f_', ages)
-
-# Collect the age suffixes into a vector
-keep_vars <- c(m_ages, f_ages)
-# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
-
-# Don't like using loops but this one was just too simple to bother with figuring out the apply version
-final <- c()
-ticker <- 1
-# Generating the column (variable) names we want to keep in a loop
-for(item in diseases) {
-  for(anotherItem in keep_vars) {
-    final[ticker] <- paste0('p_', item, '_', anotherItem)
-    ticker <- ticker + 1
-  }
-}
-
-# Keep the right vars using varnames generated in loop
-FEM <- FEM %>%
-          select(c('year', all_of(final)))
-
-# Rename some vars to remove 'p_' before merging
-FEM <- FEM %>%
-          rename_at(vars(starts_with('p_')),
-                    funs(str_replace(., 'p_', ''))) %>%
-          select(!year)
-
-pivoted.FEM <- pivot_longer(data = FEM,
-                            cols = hibpe_m_6064:asthmae_f_100p, 
-                            names_to = c('Condition', 'Gender', 'Age_Group'), 
-                            names_sep = '_')
-
-# Convert value from fraction to percentage
-pivoted.FEM$value <- pivoted.FEM$value * 100
-
-# Rename cols to enable reshaping
-male_au <- male_au %>%
-                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
-                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
-                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
-                select(!Total)
-
-female_au <- female_au %>%
-                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
-                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
-                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
-                select(!Total)
-
-# Now need to reshape these datasets to long format
-pivoted.male <- pivot_longer(data = male_au,
-                             cols = m_6064:m_100p,
-                             names_to = c('Gender', 'Age_Group'),
-                             names_sep = '_')
-
-pivoted.female <- pivot_longer(data = female_au,
-                             cols = f_6064:f_100p,
-                             names_to = c('Gender', 'Age_Group'),
-                             names_sep = '_')
-
-# Now row bind these frames together, combining male and female data
-ageUK <-rbind(pivoted.male, pivoted.female)
-
-# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
-# extract all heart related conditions for combining
-#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
-# Aggregate the values by gender and age group
-#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
-# Rename columns and add condition column back
-#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-#ageUK_heart$Condition <- 'Heart Problems'
-
-# Now drop the heart problems from original ageUK and replace with aggregate
-ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-#ageUK <- rbind(ageUK, ageUK_heart)
-
-# Rename a few just for completeness
-#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
-
-# Finally, add column to specify that this is ageUK data
-ageUK$source <- "ageUK"
-
-# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
-#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
-ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
-ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
-
-# Make copy
-newFEM <- pivoted.FEM
-
-# Rename all the conditions to fit ageUK
-# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
-#       Therefore only validation from this will be that FEM is higher than ageUK
-#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
-newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
-newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
-newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
-#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
-#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
-newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
-newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
-
-#force Make to run this again
-
-# Add column to specify this is FEM data
-newFEM$source <- "FEM"
-
-# Now row bind the datasets
-combined <- rbind(newFEM, ageUK)
-# Specify factor levels for age group
-combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
-# Sort
-combined <- arrange(combined, source, Condition, Gender, Age_Group)
-
-# Function to filter the combined dataset and visualise
-visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
-
-  if(gender == 'm') {
-    full_gender <- 'Male'
-  }
-  else if(gender == 'f') {
-    full_gender <- 'Female'
-  }
-
-  # filter by gender and condition
-  data <- data %>%
-            filter(Gender == gender) %>%
-            filter(Condition == cd)
-
-  # now plot!
-  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
-          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
-          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
-
-  # Save plot into in E-FEM_Paper/figures/
-  ggsave(filename = paste0(cd, '_', gender, '.png'),
-         plot = p,
-         path = path,
-         width = 6,
-         height = 4)
-
-  return(p)
-}
+# #AgeUK
+# male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
+# female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
+# 
+# # FEM 2014 detailed output
+# #FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+# ## extract males and females
+# #mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+# #fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+# 
+# # FEM 2014 summary
+# #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+# FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+# 
+# # Keep only year 2014 from FEM summary
+# FEM <- filter(FEM, year == 2014)
+# 
+# # Specify the age groups to extract from FEM
+# ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# # Add m and f for male and female variables
+# m_ages <- paste0('m_', ages)
+# f_ages <- paste0('f_', ages)
+# 
+# # Collect the age suffixes into a vector
+# keep_vars <- c(m_ages, f_ages)
+# # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+# diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
+# 
+# # Don't like using loops but this one was just too simple to bother with figuring out the apply version
+# final <- c()
+# ticker <- 1
+# # Generating the column (variable) names we want to keep in a loop
+# for(item in diseases) {
+#   for(anotherItem in keep_vars) {
+#     final[ticker] <- paste0('p_', item, '_', anotherItem)
+#     ticker <- ticker + 1
+#   }
+# }
+# 
+# # Keep the right vars using varnames generated in loop
+# FEM <- FEM %>%
+#           select(c('year', all_of(final)))
+# 
+# # Rename some vars to remove 'p_' before merging
+# FEM <- FEM %>%
+#           rename_at(vars(starts_with('p_')),
+#                     funs(str_replace(., 'p_', ''))) %>%
+#           select(!year)
+# 
+# pivoted.FEM <- pivot_longer(data = FEM,
+#                             cols = hibpe_m_6064:asthmae_f_100p, 
+#                             names_to = c('Condition', 'Gender', 'Age_Group'), 
+#                             names_sep = '_')
+# 
+# # Convert value from fraction to percentage
+# pivoted.FEM$value <- pivoted.FEM$value * 100
+# 
+# # Rename cols to enable reshaping
+# male_au <- male_au %>%
+#                 dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
+#                        m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
+#                        m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
+#                 select(!Total)
+# 
+# female_au <- female_au %>%
+#                 dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
+#                        f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
+#                        f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
+#                 select(!Total)
+# 
+# # Now need to reshape these datasets to long format
+# pivoted.male <- pivot_longer(data = male_au,
+#                              cols = m_6064:m_100p,
+#                              names_to = c('Gender', 'Age_Group'),
+#                              names_sep = '_')
+# 
+# pivoted.female <- pivot_longer(data = female_au,
+#                              cols = f_6064:f_100p,
+#                              names_to = c('Gender', 'Age_Group'),
+#                              names_sep = '_')
+# 
+# # Now row bind these frames together, combining male and female data
+# ageUK <-rbind(pivoted.male, pivoted.female)
+# 
+# # Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# # extract all heart related conditions for combining
+# #ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# # Aggregate the values by gender and age group
+# #ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# # Rename columns and add condition column back
+# #colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+# #ageUK_heart$Condition <- 'Heart Problems'
+# 
+# # Now drop the heart problems from original ageUK and replace with aggregate
+# ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+# #ageUK <- rbind(ageUK, ageUK_heart)
+# 
+# # Rename a few just for completeness
+# #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+# #ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+# 
+# # Finally, add column to specify that this is ageUK data
+# ageUK$source <- "ageUK"
+# 
+# # REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+# #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+# ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+# ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
+# 
+# # Make copy
+# newFEM <- pivoted.FEM
+# 
+# # Rename all the conditions to fit ageUK
+# # NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+# #       Therefore only validation from this will be that FEM is higher than ageUK
+# #newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+# newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+# newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+# newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+# #newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+# #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+# newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+# newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
+# 
+# 
+# #force Make to run this again
+# 
+# # Add column to specify this is FEM data
+# newFEM$source <- "FEM"
+# 
+# # Now row bind the datasets
+# combined <- rbind(newFEM, ageUK)
+# # Specify factor levels for age group
+# combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
+# # Sort
+# combined <- arrange(combined, source, Condition, Gender, Age_Group)
+# 
+# # Function to filter the combined dataset and visualise
+# visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
+#   
+#   if(gender == 'm') {
+#     full_gender <- 'Male'
+#   }
+#   else if(gender == 'f') {
+#     full_gender <- 'Female'
+#   }
+#   
+#   # filter by gender and condition
+#   data <- data %>%
+#             filter(Gender == gender) %>%
+#             filter(Condition == cd)
+#   
+#   # now plot!
+#   p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
+#           geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+#           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+#   
+#   # Save plot into in E-FEM_Paper/figures/
+#   ggsave(filename = paste0(cd, '_', gender, '.png'),
+#          plot = p,
+#          path = path,
+#          width = 6,
+#          height = 4)
+#   
+#   return(p)
+# }
 
@@
 
-```{r}
-# Collect all conditions (and both genders) into a list, and apply the function above to that list
-cd_list <- unique(combined$Condition)
-gender_list <- unique(combined$Gender)
-
-for(cd in cd_list) {
-  for(gen in gender_list) {
-    plot <- visAgeUKValid(combined, cd, gen)
-    print(plot)
-  }
-}
+```{r echo=FALSE}
+# # Collect all conditions (and both genders) into a list, and apply the function above to that list
+# cd_list <- unique(combined$Condition)
+# gender_list <- unique(combined$Gender)
+# 
+# for(cd in cd_list) {
+#   for(gen in gender_list) {
+#     plot <- visAgeUKValid(combined, cd, gen)
+#     print(plot)
+#   }
+# }
 ```

### 2021-08-18 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Committing all previous changes before cleaning up the repo and beginning work on a new feature

@@
 ```{r echo=FALSE}
-# Collect all conditions (and both genders) into a list, and apply the function above to that list
+# # Collect all conditions (and both genders) into a list, and apply the function above to that list
 # cd_list <- unique(combined$Condition)
@@
 ### Split by Activity Level
-```{r}
-b <- b_rep
-h <- h_rep
-
-d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
-d.exstat1 <- d.exstat1 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
-d.exstat1.pivot <- pivot_longer(d.exstat1,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-p1
-
-d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
-d.exstat2 <- d.exstat2 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
-d.exstat2.pivot <- pivot_longer(d.exstat2,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-      scale_fill_grey()
-p2
-
-d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
-d.exstat3 <- d.exstat3 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
-d.exstat3.pivot <- pivot_longer(d.exstat3,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-p3
-
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'
-
-ggsave(filename = 'hearte_dAge_exstat1.png',
-         plot = p1,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_exstat2.png',
-         plot = p2,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_exstat3.png',
-         plot = p3,
-         path = path)
+```{r echo=FALSE}
+# b_rep <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_by_rep.dta')
+# h_rep <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_by_rep.dta')
+# 
+# b <- b_rep
+# h <- h_rep
+# 
+# d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
+# d.exstat1 <- d.exstat1 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
+# d.exstat1.pivot <- pivot_longer(d.exstat1,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
+# p1
+# 
+# d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
+# d.exstat2 <- d.exstat2 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
+# d.exstat2.pivot <- pivot_longer(d.exstat2,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
+#       scale_fill_grey()
+# p2
+# 
+# d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
+# d.exstat3 <- d.exstat3 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
+# d.exstat3.pivot <- pivot_longer(d.exstat3,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
+# p3
+# 
+# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'
+# 
+# ggsave(filename = 'hearte_dAge_exstat1.png',
+#          plot = p1,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_exstat2.png',
+#          plot = p2,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_exstat3.png',
+#          plot = p3,
+#          path = path)
 
@@
 ### SplitbyEduc
-```{r}
-b <- b_rep
-h <- h_rep
-
-d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
-d.educ1 <- d.educ1 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
-d.educ1.pivot <- pivot_longer(d.educ1,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-p1
-
-d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
-d.educ2 <- d.educ2 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
-d.educ2.pivot <- pivot_longer(d.educ2,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-      scale_fill_grey()
-p2
-
-d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
-d.educ3 <- d.educ3 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
-d.educ3.pivot <- pivot_longer(d.educ3,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-p3
-
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
-
-ggsave(filename = 'hearte_dAge_educ1.png',
-         plot = p1,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_educ2.png',
-         plot = p2,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_educ3.png',
-         plot = p3,
-         path = path)
+```{r echo=FALSE}
+# b <- b_rep
+# h <- h_rep
+# 
+# d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
+# d.educ1 <- d.educ1 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
+# d.educ1.pivot <- pivot_longer(d.educ1,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
+# p1
+# 
+# d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
+# d.educ2 <- d.educ2 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
+# d.educ2.pivot <- pivot_longer(d.educ2,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
+#       scale_fill_grey()
+# p2
+# 
+# d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
+# d.educ3 <- d.educ3 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
+# d.educ3.pivot <- pivot_longer(d.educ3,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
+# p3
+# 
+# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
+# 
+# ggsave(filename = 'hearte_dAge_educ1.png',
+#          plot = p1,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_educ2.png',
+#          plot = p2,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_educ3.png',
+#          plot = p3,
+#          path = path)
 
@@
 
-c_rep <- read_dta('output/COMPLETE/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
+c_rep <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
 s_rep <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')
@@
 ```
+
+# Comparing Output Values
+```{r}
+b <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
+h <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
+
+new.df <- data.frame(b$year, b$d_age_all, h$d_age_all)
+
+new.df$diff <- new.df$h.d_age_all - new.df$b.d_age_all
+
+new.df
+```

### 2021-08-18 FEM_R/model_analysis_core.Rmd
Committing all previous changes before cleaning up the repo and beginning work on a new feature

--- 2021-06-24 FEM_R/model_analysis_core.Rmd
@@
 ```{r echo = FALSE}
-
 #AgeUK
@@
 
-```{r}
+```{r echo=FALSE}
 # Collect all conditions (and both genders) into a list, and apply the function above to that list

### 2021-09-03 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Added confidence intervals to AgeUK data in the external validation

@@
 ```{r echo = FALSE}
-# #AgeUK
-# male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE.csv')
-# female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE.csv')
-# 
-# # FEM 2014 detailed output
-# #FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-# ## extract males and females
-# #mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-# #fFEM_2014 <- FEM_2014 %>% filter(male == 0)
-# 
-# # FEM 2014 summary
-# #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-# FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
-# 
-# # Keep only year 2014 from FEM summary
-# FEM <- filter(FEM, year == 2014)
-# 
-# # Specify the age groups to extract from FEM
-# ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
-# # Add m and f for male and female variables
-# m_ages <- paste0('m_', ages)
-# f_ages <- paste0('f_', ages)
-# 
-# # Collect the age suffixes into a vector
-# keep_vars <- c(m_ages, f_ages)
-# # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-# diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
-# 
-# # Don't like using loops but this one was just too simple to bother with figuring out the apply version
-# final <- c()
-# ticker <- 1
-# # Generating the column (variable) names we want to keep in a loop
-# for(item in diseases) {
-#   for(anotherItem in keep_vars) {
-#     final[ticker] <- paste0('p_', item, '_', anotherItem)
-#     ticker <- ticker + 1
-#   }
-# }
-# 
-# # Keep the right vars using varnames generated in loop
-# FEM <- FEM %>%
-#           select(c('year', all_of(final)))
-# 
-# # Rename some vars to remove 'p_' before merging
-# FEM <- FEM %>%
-#           rename_at(vars(starts_with('p_')),
-#                     funs(str_replace(., 'p_', ''))) %>%
-#           select(!year)
-# 
-# pivoted.FEM <- pivot_longer(data = FEM,
-#                             cols = hibpe_m_6064:asthmae_f_100p,
-#                             names_to = c('Condition', 'Gender', 'Age_Group'),
-#                             names_sep = '_')
-# 
-# # Convert value from fraction to percentage
-# pivoted.FEM$value <- pivoted.FEM$value * 100
-# 
+#AgeUK
+male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE_CI.csv')
+female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE_CI.csv')
+
+# FEM 2014 detailed output
+#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+## extract males and females
+#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+
+# FEM 2014 summary
+#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+
+# Keep only year 2014 from FEM summary
+FEM <- filter(FEM, year == 2014)
+
+# Specify the age groups to extract from FEM
+ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# Add m and f for male and female variables
+m_ages <- paste0('m_', ages)
+f_ages <- paste0('f_', ages)
+
+# Collect the age suffixes into a vector
+keep_vars <- c(m_ages, f_ages)
+# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
+
+# Don't like using loops but this one was just too simple to bother with figuring out the apply version
+final <- c()
+ticker <- 1
+# Generating the column (variable) names we want to keep in a loop
+for(item in diseases) {
+  for(anotherItem in keep_vars) {
+    final[ticker] <- paste0('p_', item, '_', anotherItem)
+    ticker <- ticker + 1
+  }
+}
+
+# Keep the right vars using varnames generated in loop
+FEM <- FEM %>%
+          select(c('year', all_of(final)))
+
+# Rename some vars to remove 'p_' before merging
+FEM <- FEM %>%
+          rename_at(vars(starts_with('p_')),
+                    funs(str_replace(., 'p_', ''))) %>%
+          select(!year)
+
+pivoted.FEM <- pivot_longer(data = FEM,
+                            cols = hibpe_m_6064:asthmae_f_100p,
+                            names_to = c('Condition', 'Gender', 'Age_Group'),
+                            names_sep = '_')
+
+# Convert value from fraction to percentage
+pivoted.FEM$value <- pivoted.FEM$value * 100
+# Set gender codes to upper case
+pivoted.FEM$Gender <- toupper(pivoted.FEM$Gender)
+
 # # Rename cols to enable reshaping
@@
 #                 select(!Total)
-# 
-# # Now need to reshape these datasets to long format
-# pivoted.male <- pivot_longer(data = male_au,
-#                              cols = m_6064:m_100p,
-#                              names_to = c('Gender', 'Age_Group'),
-#                              names_sep = '_')
-# 
-# pivoted.female <- pivot_longer(data = female_au,
-#                              cols = f_6064:f_100p,
-#                              names_to = c('Gender', 'Age_Group'),
-#                              names_sep = '_')
-# 
-# # Now row bind these frames together, combining male and female data
-# ageUK <-rbind(pivoted.male, pivoted.female)
-# 
-# # Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
-# # extract all heart related conditions for combining
-# #ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
-# # Aggregate the values by gender and age group
-# #ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
-# # Rename columns and add condition column back
-# #colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-# #ageUK_heart$Condition <- 'Heart Problems'
-# 
-# # Now drop the heart problems from original ageUK and replace with aggregate
-# ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-# #ageUK <- rbind(ageUK, ageUK_heart)
-# 
-# # Rename a few just for completeness
-# #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-# #ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
-# 
-# # Finally, add column to specify that this is ageUK data
-# ageUK$source <- "ageUK"
-# 
-# # REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
-# #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
-# ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
-# ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
-# 
-# # Make copy
-# newFEM <- pivoted.FEM
-# 
-# # Rename all the conditions to fit ageUK
-# # NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
-# #       Therefore only validation from this will be that FEM is higher than ageUK
-# #newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
-# newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
-# newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
-# newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
-# #newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
-# #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
-# newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
-# newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
-# 
-# 
-# #force Make to run this again
-# 
-# # Add column to specify this is FEM data
-# newFEM$source <- "FEM"
-# 
-# # Now row bind the datasets
-# combined <- rbind(newFEM, ageUK)
-# # Specify factor levels for age group
-# combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
-# # Sort
-# combined <- arrange(combined, source, Condition, Gender, Age_Group)
-# 
-# # Function to filter the combined dataset and visualise
-# visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
-# 
-#   if(gender == 'm') {
-#     full_gender <- 'Male'
-#   }
-#   else if(gender == 'f') {
-#     full_gender <- 'Female'
-#   }
-# 
-#   # filter by gender and condition
-#   data <- data %>%
-#             filter(Gender == gender) %>%
-#             filter(Condition == cd)
-# 
-#   # now plot!
-#   p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
-#           geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
-#           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
-# 
-#   # Save plot into in E-FEM_Paper/figures/
-#   ggsave(filename = paste0(cd, '_', gender, '.png'),
-#          plot = p,
-#          path = path,
-#          width = 6,
-#          height = 4)
-# 
-#   return(p)
-# }
+
+male_au <- male_au %>%
+                dplyr::rename(M_6064_total = M_60.64_total, M_6064_lower = M_60.64_l, M_6064_upper = M_60.64_u, 
+                              M_6569_total = M_65.69_total, M_6569_lower = M_65.69_l, M_6569_upper = M_65.69_u,
+                              M_7074_total = M_70.74_total, M_7074_lower = M_70.74_l, M_7074_upper = M_70.74_u,
+                              M_7579_total = M_75.79_total, M_7579_lower = M_75.79_l, M_7579_upper = M_75.79_u,
+                              M_8084_total = M_80.84_total, M_8084_lower = M_80.84_l, M_8084_upper = M_80.84_u,
+                              M_8589_total = M_85.89_total, M_8589_lower = M_85.89_l, M_8589_upper = M_85.89_u,
+                              M_9094_total = M_90.94_total, M_9094_lower = M_90.94_l, M_9094_upper = M_90.94_u,
+                              M_9599_total = M_95.99_total, M_9599_lower = M_95.99_l, M_9599_upper = M_95.99_u,
+                              M_100p_total = M_Over.100_total, M_100p_lower = M_Over.100_l, M_100p_upper = M_Over.100_u) %>%
+                select(-M_Total_total, -M_Total_l, -M_Total_u)
+
+female_au <- female_au %>%
+                dplyr::rename(F_6064_total = F_60.64_total, F_6064_lower = F_60.64_l, F_6064_upper = F_60.64_u, 
+                              F_6569_total = F_65.69_total, F_6569_lower = F_65.69_l, F_6569_upper = F_65.69_u,
+                              F_7074_total = F_70.74_total, F_7074_lower = F_70.74_l, F_7074_upper = F_70.74_u,
+                              F_7579_total = F_75.79_total, F_7579_lower = F_75.79_l, F_7579_upper = F_75.79_u,
+                              F_8084_total = F_80.84_total, F_8084_lower = F_80.84_l, F_8084_upper = F_80.84_u,
+                              F_8589_total = F_85.89_total, F_8589_lower = F_85.89_l, F_8589_upper = F_85.89_u,
+                              F_9094_total = F_90.94_total, F_9094_lower = F_90.94_l, F_9094_upper = F_90.94_u,
+                              F_9599_total = F_95.99_total, F_9599_lower = F_95.99_l, F_9599_upper = F_95.99_u,
+                              F_100p_total = F_Over.100_total, F_100p_lower = F_Over.100_l, F_100p_upper = F_Over.100_u) %>%
+                select(-F_Total_total, -F_Total_l, -F_Total_u)
+
+# Now need to reshape these datasets to long format
+pivoted.male <- pivot_longer(data = male_au,
+                             cols = M_6064_total:M_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
+                             names_sep = '_')
+
+pivoted.female <- pivot_longer(data = female_au,
+                             cols = F_6064_total:F_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
+                             names_sep = '_')
+
+# Now bring the confidence intervals back into columns because ggplot needs that format
+# to plot them. Seems pretty daft to be honest but thats what we've got
+pivoted.male <- pivot_wider(pivoted.male,
+                            names_from = Type,
+                            values_from = value)
+pivoted.female <- pivot_wider(pivoted.female,
+                            names_from = Type,
+                            values_from = value)
+
+# Now row bind these frames together, combining male and female data
+ageUK <-rbind(pivoted.male, pivoted.female)
+
+# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# extract all heart related conditions for combining
+#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# Aggregate the values by gender and age group
+#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# Rename columns and add condition column back
+#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+#ageUK_heart$Condition <- 'Heart Problems'
+
+# Now drop the heart problems from original ageUK and replace with aggregate
+ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+#ageUK <- rbind(ageUK, ageUK_heart)
+
+# Rename a few just for completeness
+#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+
+# Finally, add column to specify that this is ageUK data
+ageUK$source <- "ageUK"
+
+# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
+
+# Make copy
+newFEM <- pivoted.FEM
+
+# Rename all the conditions to fit ageUK
+# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+#       Therefore only validation from this will be that FEM is higher than ageUK
+#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
+
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+newFEM$total <- newFEM$value
+newFEM <- select(newFEM, -value)
+
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+# Add column to specify these are all totals (no confidence intervals)
+newFEM$lower <- NA
+newFEM$upper <- NA
+
+# Now row bind the datasets
+combined <- rbind(newFEM, ageUK)
+# Specify factor levels for age group
+combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
+# Sort
+combined <- arrange(combined, source, Condition, Gender, Age_Group)
+
+# Function to filter the combined dataset and visualise
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK') {
+
+  if(gender == 'M') {
+    full_gender <- 'Male'
+  }
+  else if(gender == 'F') {
+    full_gender <- 'Female'
+  }
+
+  # filter by gender and condition
+  data <- data %>%
+            filter(Gender == gender) %>%
+            filter(Condition == cd)
+
+  # now plot!
+  p <- ggplot(data = data, aes(x = Age_Group, y = total)) +
+          geom_bar(aes(fill = source), stat = "identity", position = position_dodge2()) +
+          geom_errorbar(aes(ymin = lower, ymax = upper), position = position_dodge2(width = 1, padding = 0.5)) +
+          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+
+  # Save plot into in E-FEM_Paper/figures/
+  ggsave(filename = paste0(cd, '_', gender, '.png'),
+        plot = p,
+        path = path,
+        width = 7,
+        height = 5)
+
+  return(p)
+}
 
@@
 ```{r echo=FALSE}
-# # Collect all conditions (and both genders) into a list, and apply the function above to that list
-# cd_list <- unique(combined$Condition)
-# gender_list <- unique(combined$Gender)
-# 
-# for(cd in cd_list) {
-#   for(gen in gender_list) {
-#     plot <- visAgeUKValid(combined, cd, gen)
-#     print(plot)
-#   }
-# }
+# Collect all conditions (and both genders) into a list, and apply the function above to that list
+cd_list <- unique(combined$Condition)
+gender_list <- unique(combined$Gender)
+
+for(cd in cd_list) {
+  for(gen in gender_list) {
+    plot <- visAgeUKValid(combined, cd, gen)
+    print(plot)
+  } 
+}
 ```
@@
 
+group.colours <- c(FEM = '#f8766d', ASH = '#00FF  00')
+
 smok.plot <-ggplot(data = smok.combined, aes(x = age, y = value)) +
                 geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+                scale_fill_manual(values=group.colours) +
                 labs(x = ' ', y = 'Prevalence (%)')

### 2021-09-03 FEM_R/validation.Rmd
Added confidence intervals to AgeUK data in the external validation

@@
 #AgeUK
-mAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
-fAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
+mAGE_UK <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
+fAGE_UK <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
 
@@
 # FEM 2014 summary
-FEM_sum <- read_dta('ELSA_AgeUK_valid/ELSA_AgeUK_valid_summary.dta')
+FEM_sum <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
 
@@
 
-  p
+  return(p)
 }
@@
   for(gen in gender_list) {
-    visAgeUKValid(combined, cd, gen)
+    plot <- visAgeUKValid(combined, cd, gen)
+    print(plot)
   }
@@
 
+```
+
+## With Error Bars
+```{r}
+# Function to filter the combined dataset and visualise
+visAgeUKValidError <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {
+
+  if(gender == 'm') {
+    full_gender <- 'Male'
+  }
+  else if(gender == 'f') {
+    full_gender <- 'Female'
+  }
+
+  # filter by gender and condition
+  data <- data %>%
+            filter(Gender == gender) %>%
+            filter(Condition == cd)
+
+  # now plot!
+  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
+          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+
+  # Save plot into in E-FEM_Paper/figures/
+  #ggsave(filename = paste0(cd, '_', gender, '.png'),
+  #       plot = p,
+  #       path = path,
+  #       width = 7,
+  #       height = 5)
+
+  return(p)
+}
+
+visAgeUKValid(combined, 'Hypertension', 'm')
 ```

### 2021-09-09 FEM_R/model_analysis_core.Rmd
Lots of little changes made for updating the model to use ELSA g.2 release. Set core models to run until 2018, imputed lots of newly missing data, updated the crossvalidation and model analysis script in R, and handled reweighting issue with missing cwtresp information

@@
               P_Value_7 = pValColDef2,
-              P_Value_8 = pValColDef2
+              P_Value_8 = pValColDef2,
+              P_Value_9 = pValColDef2
             ),

### 2021-09-10 FEM_R/model_analysis_core.Rmd
Added logic for the confidence intervals on the AgeUK figures into the model_analysis_core.Rmd script

@@
 #AgeUK
-male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
-female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
-
-# FEM 2014 detailed output
-#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-## extract males and females
-#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE_CI.csv')
+female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE_CI.csv')
 
 # FEM 2014 summary
-#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
 
@@
 pivoted.FEM$value <- pivoted.FEM$value * 100
+# Set gender codes to upper case
+pivoted.FEM$Gender <- toupper(pivoted.FEM$Gender)
 
-# Rename cols to enable reshaping
+# Rename variables for the pivot
 male_au <- male_au %>%
-                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74,
-                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89,
-                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
-                select(!Total)
+                dplyr::rename(M_6064_total = M_60.64_total, M_6064_lower = M_60.64_l, M_6064_upper = M_60.64_u, 
+                              M_6569_total = M_65.69_total, M_6569_lower = M_65.69_l, M_6569_upper = M_65.69_u,
+                              M_7074_total = M_70.74_total, M_7074_lower = M_70.74_l, M_7074_upper = M_70.74_u,
+                              M_7579_total = M_75.79_total, M_7579_lower = M_75.79_l, M_7579_upper = M_75.79_u,
+                              M_8084_total = M_80.84_total, M_8084_lower = M_80.84_l, M_8084_upper = M_80.84_u,
+                              M_8589_total = M_85.89_total, M_8589_lower = M_85.89_l, M_8589_upper = M_85.89_u,
+                              M_9094_total = M_90.94_total, M_9094_lower = M_90.94_l, M_9094_upper = M_90.94_u,
+                              M_9599_total = M_95.99_total, M_9599_lower = M_95.99_l, M_9599_upper = M_95.99_u,
+                              M_100p_total = M_Over.100_total, M_100p_lower = M_Over.100_l, M_100p_upper = M_Over.100_u) %>%
+                select(-M_Total_total, -M_Total_l, -M_Total_u)
 
 female_au <- female_au %>%
-                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74,
-                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89,
-                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
-                select(!Total)
+                dplyr::rename(F_6064_total = F_60.64_total, F_6064_lower = F_60.64_l, F_6064_upper = F_60.64_u, 
+                              F_6569_total = F_65.69_total, F_6569_lower = F_65.69_l, F_6569_upper = F_65.69_u,
+                              F_7074_total = F_70.74_total, F_7074_lower = F_70.74_l, F_7074_upper = F_70.74_u,
+                              F_7579_total = F_75.79_total, F_7579_lower = F_75.79_l, F_7579_upper = F_75.79_u,
+                              F_8084_total = F_80.84_total, F_8084_lower = F_80.84_l, F_8084_upper = F_80.84_u,
+                              F_8589_total = F_85.89_total, F_8589_lower = F_85.89_l, F_8589_upper = F_85.89_u,
+                              F_9094_total = F_90.94_total, F_9094_lower = F_90.94_l, F_9094_upper = F_90.94_u,
+                              F_9599_total = F_95.99_total, F_9599_lower = F_95.99_l, F_9599_upper = F_95.99_u,
+                              F_100p_total = F_Over.100_total, F_100p_lower = F_Over.100_l, F_100p_upper = F_Over.100_u) %>%
+                select(-F_Total_total, -F_Total_l, -F_Total_u)
 
@@
 pivoted.male <- pivot_longer(data = male_au,
-                             cols = m_6064:m_100p,
-                             names_to = c('Gender', 'Age_Group'),
+                             cols = M_6064_total:M_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
                              names_sep = '_')
@@
 pivoted.female <- pivot_longer(data = female_au,
-                             cols = f_6064:f_100p,
-                             names_to = c('Gender', 'Age_Group'),
+                             cols = F_6064_total:F_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
                              names_sep = '_')
+
+# Now bring the confidence intervals back into columns because ggplot needs that format
+# to plot them. Seems pretty daft to be honest but thats what we've got
+pivoted.male <- pivot_wider(pivoted.male,
+                            names_from = Type,
+                            values_from = value)
+pivoted.female <- pivot_wider(pivoted.female,
+                            names_from = Type,
+                            values_from = value)
 
@@
 
-# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
-# extract all heart related conditions for combining
-#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
-# Aggregate the values by gender and age group
-#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
-# Rename columns and add condition column back
-#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-#ageUK_heart$Condition <- 'Heart Problems'
-
 # Now drop the heart problems from original ageUK and replace with aggregate
 ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-#ageUK <- rbind(ageUK, ageUK_heart)
-
-# Rename a few just for completeness
-#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
 
@@
 
-#force Make to run this again
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+newFEM$total <- newFEM$value
+newFEM <- select(newFEM, -value)
 
@@
 newFEM$source <- "FEM"
+# Add column to specify these are all totals (no confidence intervals)
+newFEM$lower <- NA
+newFEM$upper <- NA
 
@@
 # Function to filter the combined dataset and visualise
-visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK') {
 
-  if(gender == 'm') {
+  if(gender == 'M') {
     full_gender <- 'Male'
   }
-  else if(gender == 'f') {
+  else if(gender == 'F') {
     full_gender <- 'Female'
@@
   # now plot!
-  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
-          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+  p <- ggplot(data = data, aes(x = Age_Group, y = total)) +
+          geom_bar(aes(fill = source), stat = "identity", position = position_dodge2()) +
+          geom_errorbar(aes(ymin = lower, ymax = upper), position = position_dodge2(width = 1, padding = 0.5)) +
           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
@@
   # Save plot into in E-FEM_Paper/figures/
-  ggsave(filename = paste0(cd, '_', gender, '.png'),
-         plot = p,
-         path = path,
-         width = 6,
-         height = 4)
+  #ggsave(filename = paste0(cd, '_', gender, '.png'),
+  #      plot = p,
+  #      path = path,
+  #      width = 7,
+  #      height = 5)
 
@@
 
-```{r}
+```{r echo=False}
 FEM_base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')

### 2021-09-10 FEM_R/model_analysis_core.Rmd
Moved AgeUK validation data into the E_FEM filesystem (previously sat 1 level up). Changed paths in model_analysis_core.Rmd to point to new directory

@@
 #AgeUK
-male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE_CI.csv')
-female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE_CI.csv')
+male_au <- read.csv('../../FEM_R/AgeUK_Validation/AgeUK-chronDisease-MALE_CI.csv')
+female_au <- read.csv('../../FEM_R/AgeUK_Validation/AgeUK-chronDisease-FEMALE_CI.csv')
 
 # FEM 2014 summary
-FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
 
@@
 
-```{r echo=False}
+```{r echo=FALSE}
 FEM_base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
@@
 
-ggsave(filename = 'smoking_prev_2018.png',
-       plot = smok.plot,
-       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/',
-       width = 7,
-       height = 5)
+#ggsave(filename = 'smoking_prev_2018.png',
+#       plot = smok.plot,
+#       path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/',
+#       width = 7,
+#       height = 5)
 

### 2021-09-13 FEM_R/model_analysis_core.Rmd
Modified the model_analysis_core.Rmd script to include the T-test table for continuous economic variables (wealth and income)

@@
 # Set up vector with all p-value column names. Require these for table formatting
-pValNames <- c('P_Value_3', 'P_Value_4', 'P_Value_5', 'P_Value_6', 'P_Value_7', 'P_Value_8')
+pValNames <- c('P_Value_3', 'P_Value_4', 'P_Value_5', 'P_Value_6', 'P_Value_7', 'P_Value_8', 'P_Value_9')
 ```
@@
 ## This block will define the logic for creating the tables with Reactable
-# The tables are formatted exaclty the same for each group of vars so makes sense to define this only once
+# The tables are formatted exactly the same for each group of vars so makes sense to define this only once
 
@@
 ```
+### Count Economic
+```{r echo = FALSE}
+# Create the actual table!
+cnteconTab <- createTTT(cntecon)
+cnteconTab
+```
 ### Risk Behaviours
@@
 riskTab
-```
-
-# Cross Validation 2 - 'Handover' plots (Still WIP need fixing)
-
-```{r echo = FALSE}
-vis_CV2_prev <- function(long, CV2) {
-
-  # Keep only waves 1 - 4
-  longC <- subset(long, wave < 5)
-  longC <- as.data.table(longC)
-
-  # Replace all NA with 0
-  longC[is.na(longC)] <- 0
-
-  # Have to drop r*clust vars as they contain strings
-  longC <- select(longC, -contains('clust'))
-
-  # Now get the weighted mean by wave of every column based on cwtresp
-  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]
-
-  # ELSA_long is now processed to the point where we can start extracting and combining variables
-  # Better start by adding year to l2 and reordering the columns
-  l2$year <- seq(from=2002, to = 2008, by = 2)
-  setcolorder(l2, c('year', 'wave'))
-
-  # Now lets extract chronic disease variables from both dataframes and do a row combine
-  # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl)) #, anyiadl, adl1, adl2, adl3p))
-  cd_CV2 <- select(CV2, c(year, a_age_all, p_cancre_all, p_diabe_all, 
-                            p_hearte_all, p_lunge_all, p_smoken_all, 
-                            a_bmi_all, p_employed_all,  p_anyadl_all))
-                            #, p_anyiadl_all p_adl1_all, p_adl2_all, p_adl3p_all))
-
-  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_all,
-                                        'cancre' = p_cancre_all, 
-                                        'diabe' = p_diabe_all, 
-                                        'hearte' = p_hearte_all,
-                                        'lunge' = p_lunge_all,
-                                        'smoken' = p_smoken_all,
-                                        'bmi' = a_bmi_all,
-                                        'employed' = p_employed_all,
-                                        'anyadl' = p_anyadl_all)
-                                        #'anyiadl' = p_anyiadl_all)
-                                        #'adl1' = p_adl1_all,
-                                        #'adl2' = p_adl2_all,
-                                        #'adl3p' = p_adl3p_all)
-
-  combine <- rbind(cd_l2, CV2_rename)
-
-  melted <- reshape2::melt(combine, id.var = 'year', value.name = 'Prevalence')
-
-  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
-          geom_line() +
-          facet_wrap(. ~ variable, scales = "free")
-          #theme(panel.spacing.x = unit(1.5, 'lines')) +
-          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
-}
-
-# Load in whole stock pop, plus the CV2 summary output file
-long <- read_dta('../../input_data/ELSA_long.dta')
-CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
-
-vis_CV2_prev(long, CV2)
-```
-
-Second go at this cross-validation, attempting to solve some issues by using the 60 plus population.
-
-```{r echo = FALSE}
-vis_CV2_prev <- function(long, CV2) {
-
-  # Keep only waves 1 - 4
-  longC <- subset(long, wave < 5)
-  longC <- as.data.table(longC)
-
-  # Replace all NA with 0
-  longC[is.na(longC)] <- 0
-
-  # Have to drop r*clust vars as they contain strings
-  longC <- select(longC, -contains('clust'))
-
-  # Now get the weighted mean by wave of every column based on cwtresp
-  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]
-
-  # ELSA_long is now processed to the point where we can start extracting and combining variables
-  # Better start by adding year to l2 and reordering the columns
-  l2$year <- seq(from=2002, to = 2008, by = 2)
-  setcolorder(l2, c('year', 'wave'))
-
-  # Now lets extract chronic disease variables from both dataframes and do a row combine
-  # Start with cancre, diabe, hearte
-  cd_l2 <- select(l2, c(year, age, cancre, diabe, hearte, lunge, smoken, bmi, employed, anyadl)) #, anyiadl, adl1, adl2, adl3p))
-  cd_CV2 <- select(CV2, c(year, a_age_60p, p_cancre_60p, p_diabe_60p, 
-                            p_hearte_60p, p_lunge_60p, p_smoken_60p, 
-                            a_bmi_60p, p_employed_60p,  p_anyadl_60p))
-                            #, p_anyiadl_60p p_adl1_60p, p_adl2_60p, p_adl3p_60p))
-
-  CV2_rename <- cd_CV2 %>% dplyr::rename('age' = a_age_60p,
-                                        'cancre' = p_cancre_60p, 
-                                        'diabe' = p_diabe_60p, 
-                                        'hearte' = p_hearte_60p,
-                                        'lunge' = p_lunge_60p,
-                                        'smoken' = p_smoken_60p,
-                                        'bmi' = a_bmi_60p,
-                                        'employed' = p_employed_60p,
-                                        'anyadl' = p_anyadl_60p)
-                                        #'anyiadl' = p_anyiadl_60p)
-                                        #'adl1' = p_adl1_60p,
-                                        #'adl2' = p_adl2_60p,
-                                        #'adl3p' = p_adl3p_60p)
-
-  combine <- rbind(cd_l2, CV2_rename)
-
-  melted <- reshape2::melt(combine, id.var = 'year', value.name = 'Prevalence')
-
-  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
-          geom_line() +
-          facet_wrap(. ~ variable, scales = "free")
-          #theme(panel.spacing.x = unit(1.5, 'lines')) +
-          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
-}
-
-# Load in whole stock pop, plus the CV2 summary output file
-long <- read_dta('../../input_data/ELSA_long.dta')
-CV2 <- read_dta('ELSA_CV2/ELSA_CV2_summary.dta')
-
-vis_CV2_prev(long, CV2)
 ```

### 2021-10-27 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added a script to visualise all the wealth-health-socioeconomic outputs, as well as do the replacement for the replacement intervention. Replacement is done as a first try with multiple imputation but very crude

---
title: "R Notebook"
output: html_notebook
---

```{r "setup", include=FALSE}

require(tidyverse)
require(haven)
require(ggplot2)

# for multiple imputation replacement
require(mice)
require(data.table)

workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

rm(workingDir)
```

## Data
```{r}

cohort <- read_dta("output/WEALTH/ELSA_cohort/ELSA_cohort_summary.dta")

```

## Survival Curve
```{r}

# Function to compare survival curves between quintiles

## THIS IS NOT USEFUL!!!
# Because we calculate survival using the m_endpop_* variable (which is a 
# population level variable) it will always give the same value for every
# subgroup.
# Need to talk to Bryan about a way to analyse this within subgroup.
survivalCurveSubgroup <- function(cohort) {

  # Age and survival for each quintile
  cohort$Age <- (cohort$year - 2012) + 50
  cohort$q1 <- cohort$m_endpop_wq1 / cohort$m_endpop_wq1[1]
  cohort$q2 <- cohort$m_endpop_wq2 / cohort$m_endpop_wq2[1]
  cohort$q3 <- cohort$m_endpop_wq3 / cohort$m_endpop_wq3[1]
  cohort$q4 <- cohort$m_endpop_wq4 / cohort$m_endpop_wq4[1]
  cohort$q5 <- cohort$m_endpop_wq5 / cohort$m_endpop_wq5[1]

  # Cut dataframe down to just the age and survivals
  survivals <- cohort %>% select(Age, q1, q2, q3, q4, q5)

  # pivot the data to long format for ggplot
  survivals <- pivot_longer(data = survivals,
                            cols = q1:q5,
                            names_to = "Quintile",
                            values_to = "Survival")

  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Survival, color = Quintile)) +
    geom_smooth(se = FALSE, linetype = 'solid', size=0.4)

  return(p)

}

survivalCurveSubgroup(cohort)

```

### Testing
```{r echo=FALSE}

check <- cohort %>% select(year, m_endpop_wq1, m_endpop_wq2, m_endpop_wq3, m_endpop_wq4, m_endpop_wq5)

check2 <- cohort %>% select(year, p_anydisease_wq1, p_anydisease_wq2, p_anydisease_wq3, p_anydisease_wq4, p_anydisease_wq5)

check3 <- cohort %>% select(year, p_disabled_wq1, p_disabled_wq2, p_disabled_wq3, p_disabled_wq4, p_disabled_wq5)

```

```{r}
rm(check, check2, check3)
```

## Prevalences
```{r}

allSubgroupPrevalence <- function(cohort, var) {

  #Age
  cohort$Age <- (cohort$year - 2012) + 50

  # Build varnames
  varnames <- c()
  for (i in 1:5) {
    varnames[i] <- paste0('p_', var, '_wq', i)
  }

  # Select just the things we want
  prev.all <- cohort %>% select(Age, all_of(varnames))

  # # pivot the data to long format for ggplot
  survivals <- pivot_longer(data = prev.all,
                            cols = varnames[1]:varnames[5],
                            names_to = "Quintile",
                            values_to = "Disease_Prev")

  # plot with geom_smooth
  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Disease_Prev, color = Quintile)) +
    geom_smooth(se=FALSE, size=0.4)

  return(p)

}

allSubgroupPrevalence(cohort, 'anydisease')
allSubgroupPrevalence(cohort, 'disabled')
allSubgroupPrevalence(cohort, 'severeCondition')
allSubgroupPrevalence(cohort, 'mildCondition')

```

```{r}

someSubgroupPrevalence <- function(cohort, var, subgroups) {

  #Age
  cohort$Age <- (cohort$year - 2012) + 50

  # Build varnames
  varnames <- c()
  ticker <- 1
  for (i in subgroups) {
    varnames[ticker] <- paste0('p_', var, '_wq', i)
    ticker <- ticker + 1
  }

  # Select just the things we want
  #prev.all <- cohort %>% select(Age, p_anydisease_wq1, p_anydisease_wq2, p_anydisease_wq3, p_anydisease_wq4, p_anydisease_wq5)
  prev.all <- cohort %>% select(Age, all_of(varnames))

  # # pivot the data to long format for ggplot
  # survivals <- pivot_longer(data = prev.all,
  #                           cols = p_anydisease_wq1:p_anydisease_wq5,
  #                           names_to = "Quintile",
  #                           values_to = "Disease_Prev")

  survivals <- pivot_longer(data = prev.all,
                            cols = varnames[1]:varnames[length(varnames)],
                            names_to = "Quintile",
                            values_to = "Disease_Prev")

  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Disease_Prev, color = Quintile)) +
    geom_smooth(se=FALSE, size=0.4)

  return(p)

}

someSubgroupPrevalence(cohort, 'anydisease', subgroups = c(1,3))
someSubgroupPrevalence(cohort, 'disabled', subgroups = c(1,3))
someSubgroupPrevalence(cohort, 'disabled', subgroups = c(1,4,5))

```

# Replacement Counterfactual

```{r}
## Read in ELSA_repl.dta file for replacement
repl <- read_dta('input_data/ELSA_repl.dta')
```

In this counterfactual, we will be using wealth as an input to health without directly using wealth. We want to replace the health and risk behaviour status distribution of the poorest quintile with that of the median or higher quintile to assess how differently the health outcomes would be.

Practically, I think we need to have a few different versions of this to see the impact we can create.

1. Just risk behaviours. 
  - Drinking, smoking, BMI, exercise status, more?
2. Risk behaviours and disability
3. Risk behaviours, disability, and health status

NEED TO REMEMBER ACCOUNTING!! If changing smoking status especially, need to make sure that the smoke_start and smoke_stop variables both make sense after changes.

## Ideas

### IPF

One way to go about this is IPF, so lets figure out what we need for IPF.

- Seed
  - Just the variables of interest organised into a matrix of rows = length(people) & cols = length(variables)
- Constraints
  - Totals of each variable by age and sex

IPF actually could be a bit tricky, as I would need to collapse my seed down to counts by group, and would then need to integerise the population after this happens. I could do this using the algorithm from humanleague that is built for this purpose. Could even do the whole thing using humanleague functions? I know they are efficient and it would be nice to cite them. However this is an added layer of complexity which I could probably do to stay away from.

### Multiple imputation

This is not a bad idea, because I can run a multiple imputation by age and sex just like I can do IPF. The `mice` package in R looks the best, and we can do different imputation methods for different variables (i.e. different methods for BMI - continuous, drinking & smoking - ordered).

```{r}
## REMEMBER ACCOUNTING! SMOKING VARS AND EXSTAT

# which variables to select and impute
# Need to do this in parts because some of the lags wouldn't impute when 
# the current vars were included. Specifically smoken, smokev, and problem_drinker
demog.vars <- c('idauniq', 'wealth_quint', 'age', 'male')
risk.vars.fir <- c('smokev', 'smoken', 'heavy_smoker', 'drink', 
                   'problem_drinker', 'logbmi',  'exstat', 'l2drink', 
                   'l2logbmi', 'l2exstat')
risk.vars.sec <- c('heavy_smoker', 'drink', 'logbmi',  'exstat', 'l2smoken',
                   'l2smokev', 'l2heavy_smoker', 'l2drink', 'l2problem_drinker',
                   'l2logbmi', 'l2exstat')

## Have a go at multiple imputation using distribution from median quintile first
# First generate a test dataset with only the variables we need to impute and age,sex
test.fir <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
                  select(all_of(demog.vars), all_of(risk.vars.fir))

test.sec <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
                  select(all_of(demog.vars), all_of(risk.vars.sec))

# Remove data for the risk behaviours for wealth_quint == 1
# Doing this in chunks as not sure of a better way
#test.1 <- test[test$wealth_quint == 1]
#test.3 <- test[test$wealth_quint == 3]

test.1.fir <- filter(test.fir, wealth_quint == 1)
test.1.sec <- filter(test.sec, wealth_quint == 1)
test.3.fir <- filter(test.fir, wealth_quint == 3)
test.3.sec <- filter(test.sec, wealth_quint == 3)

# replace all risk var information with NA for lowest quintile
test.1.fir[risk.vars.fir] <- NA
test.1.sec[risk.vars.sec] <- NA

# Now put them all back together with a rbind
test.fir <- rbind(test.1.fir, test.3.fir)
test.sec <- rbind(test.1.sec, test.3.sec)

# Now have a go at imputing all missings
imp <- mice(data=test.fir, m=5, maxit=10, printFlag = FALSE)

# get a complete version of the dataset
test2.fir <- complete(imp)

imp2 <- mice(data=test.sec, m=5, maxit=10, printFlag = FALSE)

test2.sec <- complete(imp2)

test2 <- merge(test2.fir, test2.sec, by = 'idauniq')

rm(test.1.fir, test.1.sec, test.3.fir, test.3.sec, risk.vars.fir, risk.vars.sec,
   test.fir, test.sec)
```

```{r}

id.var <- 'idauniq'
demog.vars <- c('wealth_quint', 'age', 'male')
risk.vars <- c('smokev', 'l2smokev', 'smoken', 'l2smoken', 'heavy_smoker',
               'l2heavy_smoker', 'drink', 'l2drink', 'problem_drinker',
               'l2problem_drinker', 'logbmi', 'l2logbmi', 'exstat',
               'l2exstat')

test <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
                  select(all_of(id.var), all_of(demog.vars), all_of(risk.vars))

#test.1 <- test[test$wealth_quint == 1]
#test.3 <- test[test$wealth_quint == 3]

test.1 <- test %>% filter(wealth_quint == 1)
test.3 <- test %>% filter(wealth_quint == 3)

test.1[risk.vars] <- NA

test <- rbind(test.1, test.3)

imp <- mice(data=test, m=5, maxit=10, printFlag = FALSE)

# get a complete version of the dataset
test2 <- complete(imp)

### ACCOUNTING
# if smoken == 1 then most likely l2smoken == 1 also
if (is.na(test2$l2smoken)) {
  test2$l2smoken <- test2$smoken
}

# if l2smokev == 1 then smokev must also == 1
if (is.na(test2$smokev)) {
  test2$smokev <- test2$l2smokev
  # also, if l2smoken == 1 then l2smokev must also equal 1
  if (test2$l2smoken == 1) {
    test2$l2smokev <- 1
  }
  # same for smoken
  if (test2$smoken == 1)
    test2$l2smoken <- 1
}

# no variability in problem_drinker -> l2problem_drinker in median wealth group,
# so assume same here
if (is.na(test2$l2problem_drinker)) {
  test2$l2problem_drinker <- test2$problem_drinker
}

## OK!
## The test2 dataset now contains fully imputed data for quintile 1 based on
## values from quintile 3 so far. Can easily turn this into a function and 
## use different quintiles (or just do them without the function)
```

```{r}
## Step 2 is now to replace the values from the original repl dataset and save this as an alternative for running
# through the model

# copy this so we don't mess with the original repl
repl.replaced <- repl %>% filter(wealth_quint == 1) %>%
                          select(-all_of(risk.vars), -all_of(demog.vars))

# get just the quint 1 people
quint1.replaced <- test2 %>% filter(wealth_quint == 1)

# merge the modified repl with the newly replaced quint1 people
repl.replaced <- merge(repl.replaced, quint1.replaced, by='idauniq', all.x = TRUE)

# Now filter all the quint1 people from original repl and row bind the new people
repl2 <- repl %>% filter(wealth_quint != 1)

repl.final <- rbind(repl.replaced, repl2)

```

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final, 'input_data/ELSA_repl_modified_quint1.dta')
```

### 2021-10-29 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added code for generating replacement repl populations, along with some of the accounting needed and some imputation to get the population to run through the model (more needed before its working correctly)

@@
 
+# Data manipulation and visualisation
 require(tidyverse)
+require(ggplot2)
+
+# Handling Stata data files
 require(haven)
-require(ggplot2)
+require(foreign)
 
@@
 require(mice)
-require(data.table)
 
@@
 # if smoken == 1 then most likely l2smoken == 1 also
-if (is.na(test2$l2smoken)) {
-  test2$l2smoken <- test2$smoken
-}
+# if (is.na(test2$l2smoken)) {
+#   test2$l2smoken <- test2$smoken
+# }
+test2$l2smoken[is.na(test2$l2smoken)] <- test2$smoken
 
 # if l2smokev == 1 then smokev must also == 1
-if (is.na(test2$smokev)) {
-  test2$smokev <- test2$l2smokev
-  # also, if l2smoken == 1 then l2smokev must also equal 1
-  if (test2$l2smoken == 1) {
-    test2$l2smokev <- 1
-  }
-  # same for smoken
-  if (test2$smoken == 1)
-    test2$l2smoken <- 1
-}
+# if (is.na(test2$smokev)) {
+#   test2$smokev <- test2$l2smokev
+#   # also, if l2smoken == 1 then l2smokev must also equal 1
+#   if (test2$l2smoken == 1) {
+#     test2$l2smokev <- 1
+#   }
+#   # same for smoken
+#   if (test2$smoken == 1)
+#     test2$l2smoken <- 1
+# }
+test2$smokev[is.na(test2$smokev)] <- test2$l2smokev
+test2$l2smokev[test2$l2smoken == 1] <- 1
+test2$l2smoken[test2$smoken == 1] <- 1
 
@@
 # so assume same here
-if (is.na(test2$l2problem_drinker)) {
-  test2$l2problem_drinker <- test2$problem_drinker
-}
+# if (is.na(test2$l2problem_drinker)) {
+#   test2$l2problem_drinker <- test2$problem_drinker
+# }
+test2$l2problem_drinker[is.na(test2$l2problem_drinker)] <- test2$problem_drinker
 
@@
 ## use different quintiles (or just do them without the function)
+
+rm(imp, test, test.1, test.3)
 ```
@@
 ```{r}
-## Step 2 is now to replace the values from the original repl dataset and save this as an alternative for running
-# through the model
+## Step 2 is now to replace the values from the original repl dataset and save 
+# this as an alternative for running through the model
 
@@
 
+rm(quint1.replaced, repl.replaced, repl2)
+```
+
+Need to deal with some missing data and problems that are created by loading the data in R and saving to dta format.
+
+Specifically, loading data in R loses the special missing characters that stata keeps and can deal with, just replaces with a simple (.) missing.
+One of the problem vars is educl (spouse education level), which is needed for the education imputation model.
+We can replace all missings with 0 as educl is used in the equation educl * married, which will be 0 if not married anyway.
+
+```{r}
+
+## Whatever we do in here have to do to the baseline repl as well just to make 
+# sure things are standardised
+
+repl.final2 <- repl.final
+repl2 <- repl
+
+## handle missing educl data as described above
+repl.final2$educl[is.na(repl.final2$educl)] <- 0
+repl2$educl[is.na(repl$educl)] <- 0
+
+table(repl.final2$educl)
+summary(repl.final2$educl)
+print('---------')
+
+repl.final2$radadeduage[is.na(repl.final2$radadeduage)] <- 0
+repl2$radadeduage[is.na(repl$radadeduage)] <- 0
+
+repl.final2$ramomeduage[is.na(repl.final2$ramomeduage)] <- 0
+repl2$ramomeduage[is.na(repl$ramomeduage)] <- 0
+
+## USE THIS SPACE TO HOTDECK LOTS OF MISSING VARIABLES (start with lnly)
 ```
@@
 ## FINALLY SAVE THE DATA
-write_dta(repl.final, 'input_data/ELSA_repl_modified_quint1.dta')
+write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
+write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
 ```

### 2021-10-29 FEM_R/Wealth_Health_Socioeconomics.Rmd
Hotdecked some variables to make both the repl_standard and repl_modified_quint1 populations run through simulation. Can now start assessing the outcome of the intervention

@@
 
+# for hotdecking
+require(VIM)
+
 workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
@@
 ```
+
+# Replacement Counterfactual
+
+```{r}
+## Read in ELSA_repl.dta file for replacement
+repl <- read_dta('input_data/ELSA_repl.dta')
+```
+
+In this counterfactual, we will be using wealth as an input to health without directly using wealth.
+We want to replace the health and risk behaviour status distribution of the poorest quintile with that of the median or higher quintile to assess how differently the health outcomes would be.
+
+Practically, I think we need to have a few different versions of this to see the impact we can create.
+
+1. Just risk behaviours. 
+  - Drinking, smoking, BMI, exercise status, more?
+2. Risk behaviours and disability
+3. Risk behaviours, disability, and health status
+
+NEED TO REMEMBER ACCOUNTING!! If changing smoking status especially, need to make sure that the smoke_start and smoke_stop variables both make sense after changes.
+
+## Ideas
+
+### IPF
+
+One way to go about this is IPF, so lets figure out what we need for IPF.
+
+- Seed
+  - Just the variables of interest organised into a matrix of rows = length(people) & cols = length(variables)
+- Constraints
+  - Totals of each variable by age and sex
+
+IPF actually could be a bit tricky, as I would need to collapse my seed down to counts by group, and would then need to integerise the population after this happens.
+I could do this using the algorithm from humanleague that is built for this purpose.
+Could even do the whole thing using humanleague functions?
+I know they are efficient and it would be nice to cite them.
+However this is an added layer of complexity which I could probably do to stay away from.
+
+### Multiple imputation
+
+This is not a bad idea, because I can run a multiple imputation by age and sex just like I can do IPF.
+The `mice` package in R looks the best, and we can do different imputation methods for different variables (i.e.
+different methods for BMI - continuous, drinking & smoking - ordered).
+
+```{r}
+# Set up lists of id vars, demographic vars (for imputation) and risk vars to be imputed
+id.var <- 'idauniq'
+demog.vars <- c('wealth_quint', 'age', 'male')
+risk.vars <- c('smokev', 'l2smokev', 'smoken', 'l2smoken', 'heavy_smoker',
+               'l2heavy_smoker', 'drink', 'l2drink', 'problem_drinker',
+               'l2problem_drinker', 'logbmi', 'l2logbmi', 'exstat',
+               'l2exstat')
+
+test <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
+                  select(all_of(id.var), all_of(demog.vars), all_of(risk.vars))
+
+# separate test df by quintile so we can remove data for quint1
+test.1 <- test %>% filter(wealth_quint == 1)
+test.3 <- test %>% filter(wealth_quint == 3)
+test.1[risk.vars] <- NA
+test <- rbind(test.1, test.3)
+
+# Run mi & get a complete version of the imputed dataset
+imp <- mice(data=test, m=5, maxit=10, printFlag = FALSE)
+test2 <- complete(imp)
+
+### ACCOUNTING
+# if smoken == 1 then most likely l2smoken == 1 also
+test2$l2smoken[is.na(test2$l2smoken)] <- test2$smoken
+
+# if l2smokev == 1 then smokev must also == 1
+test2$smokev[is.na(test2$smokev)] <- test2$l2smokev
+# also, if l2smoken == 1 then l2smokev must also equal 1
+test2$l2smokev[test2$l2smoken == 1] <- 1
+# same for smoken
+test2$l2smoken[test2$smoken == 1] <- 1
+
+# no variability in problem_drinker -> l2problem_drinker in median wealth group,
+# so assume same here
+test2$l2problem_drinker[is.na(test2$l2problem_drinker)] <- test2$problem_drinker
+
+## OK!
+## The test2 dataset now contains fully imputed data for quintile 1 based on
+## values from quintile 3 so far. Can easily turn this into a function and 
+## use different quintiles (or just do them without the function)
+
+rm(imp, test, test.1, test.3)
+```
+
+```{r}
+## Next step is to replace the values from the original repl dataset and save 
+# this as an alternative for running through the model
+
+# copy this so we don't mess with the original repl
+repl.replaced <- repl %>% filter(wealth_quint == 1) %>%
+                          select(-all_of(risk.vars), -all_of(demog.vars))
+
+# get just the quint 1 people
+quint1.replaced <- test2 %>% filter(wealth_quint == 1)
+
+# merge the modified repl with the newly replaced quint1 people
+repl.replaced <- merge(repl.replaced, quint1.replaced, by='idauniq', all.x = TRUE)
+
+# Now filter all the quint1 people from original repl and row bind the new people
+repl2 <- repl %>% filter(wealth_quint != 1)
+
+repl.final <- rbind(repl.replaced, repl2)
+
+rm(quint1.replaced, repl.replaced, repl2)
+```
+
+Need to deal with some missing data and problems that are created by loading the data in R and saving to dta format.
+
+Specifically, loading data in R loses the special missing characters that stata keeps and can deal with, just replaces with a simple (.) missing.
+One of the problem vars is educl (spouse education level), which is needed for the education imputation model.
+We can replace all missings with 0 as educl is used in the equation educl * married, which will be 0 if not married anyway.
+
+```{r}
+
+## Whatever we do in here have to do to the baseline repl as well just to make 
+# sure things are standardised
+
+repl.final2 <- repl.final
+repl2 <- repl
+
+## handle missing educl data as described above
+repl.final2$educl[is.na(repl.final2$educl)] <- 0
+repl2$educl[is.na(repl$educl)] <- 0
+
+#table(repl.final2$educl)
+#summary(repl.final2$educl)
+#print('---------')
+
+repl.final2$radadeduage[is.na(repl.final2$radadeduage)] <- 0
+repl2$radadeduage[is.na(repl$radadeduage)] <- 0
+
+repl.final2$ramomeduage[is.na(repl.final2$ramomeduage)] <- 0
+repl2$ramomeduage[is.na(repl$ramomeduage)] <- 0
+
+## USE THIS SPACE TO HOTDECK LOTS OF MISSING VARIABLES (start with lnly)
+hotdeck.vars <- c('lnly', 'workstat', 'smokev')
+
+repl.final2 <- hotdeck(repl.final2, variable = hotdeck.vars)
+repl2 <- hotdeck(repl2, variable = hotdeck.vars)
+```
+
+```{r}
+## FINALLY SAVE THE DATA
+write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
+write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
+```
+
+# Output Visualisation
 
@@
 ```
-
-# Replacement Counterfactual
-
-```{r}
-## Read in ELSA_repl.dta file for replacement
-repl <- read_dta('input_data/ELSA_repl.dta')
-```
-
-In this counterfactual, we will be using wealth as an input to health without directly using wealth.
-We want to replace the health and risk behaviour status distribution of the poorest quintile with that of the median or higher quintile to assess how differently the health outcomes would be.
-
-Practically, I think we need to have a few different versions of this to see the impact we can create.
-
-1. Just risk behaviours. 
-  - Drinking, smoking, BMI, exercise status, more?
-2. Risk behaviours and disability
-3. Risk behaviours, disability, and health status
-
-NEED TO REMEMBER ACCOUNTING!! If changing smoking status especially, need to make sure that the smoke_start and smoke_stop variables both make sense after changes.
-
-## Ideas
-
-### IPF
-
-One way to go about this is IPF, so lets figure out what we need for IPF.
-
-- Seed
-  - Just the variables of interest organised into a matrix of rows = length(people) & cols = length(variables)
-- Constraints
-  - Totals of each variable by age and sex
-
-IPF actually could be a bit tricky, as I would need to collapse my seed down to counts by group, and would then need to integerise the population after this happens.
-I could do this using the algorithm from humanleague that is built for this purpose.
-Could even do the whole thing using humanleague functions?
-I know they are efficient and it would be nice to cite them.
-However this is an added layer of complexity which I could probably do to stay away from.
-
-### Multiple imputation
-
-This is not a bad idea, because I can run a multiple imputation by age and sex just like I can do IPF.
-The `mice` package in R looks the best, and we can do different imputation methods for different variables (i.e.
-different methods for BMI - continuous, drinking & smoking - ordered).
-
-```{r}
-## REMEMBER ACCOUNTING! SMOKING VARS AND EXSTAT
-
-# which variables to select and impute
-# Need to do this in parts because some of the lags wouldn't impute when 
-# the current vars were included. Specifically smoken, smokev, and problem_drinker
-demog.vars <- c('idauniq', 'wealth_quint', 'age', 'male')
-risk.vars.fir <- c('smokev', 'smoken', 'heavy_smoker', 'drink', 
-                   'problem_drinker', 'logbmi',  'exstat', 'l2drink', 
-                   'l2logbmi', 'l2exstat')
-risk.vars.sec <- c('heavy_smoker', 'drink', 'logbmi',  'exstat', 'l2smoken',
-                   'l2smokev', 'l2heavy_smoker', 'l2drink', 'l2problem_drinker',
-                   'l2logbmi', 'l2exstat')
-
-## Have a go at multiple imputation using distribution from median quintile first
-# First generate a test dataset with only the variables we need to impute and age,sex
-test.fir <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
-                  select(all_of(demog.vars), all_of(risk.vars.fir))
-
-test.sec <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
-                  select(all_of(demog.vars), all_of(risk.vars.sec))
-
-# Remove data for the risk behaviours for wealth_quint == 1
-# Doing this in chunks as not sure of a better way
-#test.1 <- test[test$wealth_quint == 1]
-#test.3 <- test[test$wealth_quint == 3]
-
-test.1.fir <- filter(test.fir, wealth_quint == 1)
-test.1.sec <- filter(test.sec, wealth_quint == 1)
-test.3.fir <- filter(test.fir, wealth_quint == 3)
-test.3.sec <- filter(test.sec, wealth_quint == 3)
-
-# replace all risk var information with NA for lowest quintile
-test.1.fir[risk.vars.fir] <- NA
-test.1.sec[risk.vars.sec] <- NA
-
-# Now put them all back together with a rbind
-test.fir <- rbind(test.1.fir, test.3.fir)
-test.sec <- rbind(test.1.sec, test.3.sec)
-
-# Now have a go at imputing all missings
-imp <- mice(data=test.fir, m=5, maxit=10, printFlag = FALSE)
-
-# get a complete version of the dataset
-test2.fir <- complete(imp)
-
-imp2 <- mice(data=test.sec, m=5, maxit=10, printFlag = FALSE)
-
-test2.sec <- complete(imp2)
-
-test2 <- merge(test2.fir, test2.sec, by = 'idauniq')
-
-rm(test.1.fir, test.1.sec, test.3.fir, test.3.sec, risk.vars.fir, risk.vars.sec,
-   test.fir, test.sec)
-```
-
-```{r}
-
-id.var <- 'idauniq'
-demog.vars <- c('wealth_quint', 'age', 'male')
-risk.vars <- c('smokev', 'l2smokev', 'smoken', 'l2smoken', 'heavy_smoker',
-               'l2heavy_smoker', 'drink', 'l2drink', 'problem_drinker',
-               'l2problem_drinker', 'logbmi', 'l2logbmi', 'exstat',
-               'l2exstat')
-
-test <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 3)) %>%
-                  select(all_of(id.var), all_of(demog.vars), all_of(risk.vars))
-
-#test.1 <- test[test$wealth_quint == 1]
-#test.3 <- test[test$wealth_quint == 3]
-
-test.1 <- test %>% filter(wealth_quint == 1)
-test.3 <- test %>% filter(wealth_quint == 3)
-
-test.1[risk.vars] <- NA
-
-test <- rbind(test.1, test.3)
-
-imp <- mice(data=test, m=5, maxit=10, printFlag = FALSE)
-
-# get a complete version of the dataset
-test2 <- complete(imp)
-
-### ACCOUNTING
-# if smoken == 1 then most likely l2smoken == 1 also
-# if (is.na(test2$l2smoken)) {
-#   test2$l2smoken <- test2$smoken
-# }
-test2$l2smoken[is.na(test2$l2smoken)] <- test2$smoken
-
-# if l2smokev == 1 then smokev must also == 1
-# if (is.na(test2$smokev)) {
-#   test2$smokev <- test2$l2smokev
-#   # also, if l2smoken == 1 then l2smokev must also equal 1
-#   if (test2$l2smoken == 1) {
-#     test2$l2smokev <- 1
-#   }
-#   # same for smoken
-#   if (test2$smoken == 1)
-#     test2$l2smoken <- 1
-# }
-test2$smokev[is.na(test2$smokev)] <- test2$l2smokev
-test2$l2smokev[test2$l2smoken == 1] <- 1
-test2$l2smoken[test2$smoken == 1] <- 1
-
-# no variability in problem_drinker -> l2problem_drinker in median wealth group,
-# so assume same here
-# if (is.na(test2$l2problem_drinker)) {
-#   test2$l2problem_drinker <- test2$problem_drinker
-# }
-test2$l2problem_drinker[is.na(test2$l2problem_drinker)] <- test2$problem_drinker
-
-## OK!
-## The test2 dataset now contains fully imputed data for quintile 1 based on
-## values from quintile 3 so far. Can easily turn this into a function and 
-## use different quintiles (or just do them without the function)
-
-rm(imp, test, test.1, test.3)
-```
-
-```{r}
-## Step 2 is now to replace the values from the original repl dataset and save 
-# this as an alternative for running through the model
-
-# copy this so we don't mess with the original repl
-repl.replaced <- repl %>% filter(wealth_quint == 1) %>%
-                          select(-all_of(risk.vars), -all_of(demog.vars))
-
-# get just the quint 1 people
-quint1.replaced <- test2 %>% filter(wealth_quint == 1)
-
-# merge the modified repl with the newly replaced quint1 people
-repl.replaced <- merge(repl.replaced, quint1.replaced, by='idauniq', all.x = TRUE)
-
-# Now filter all the quint1 people from original repl and row bind the new people
-repl2 <- repl %>% filter(wealth_quint != 1)
-
-repl.final <- rbind(repl.replaced, repl2)
-
-rm(quint1.replaced, repl.replaced, repl2)
-```
-
-Need to deal with some missing data and problems that are created by loading the data in R and saving to dta format.
-
-Specifically, loading data in R loses the special missing characters that stata keeps and can deal with, just replaces with a simple (.) missing.
-One of the problem vars is educl (spouse education level), which is needed for the education imputation model.
-We can replace all missings with 0 as educl is used in the equation educl * married, which will be 0 if not married anyway.
-
-```{r}
-
-## Whatever we do in here have to do to the baseline repl as well just to make 
-# sure things are standardised
-
-repl.final2 <- repl.final
-repl2 <- repl
-
-## handle missing educl data as described above
-repl.final2$educl[is.na(repl.final2$educl)] <- 0
-repl2$educl[is.na(repl$educl)] <- 0
-
-table(repl.final2$educl)
-summary(repl.final2$educl)
-print('---------')
-
-repl.final2$radadeduage[is.na(repl.final2$radadeduage)] <- 0
-repl2$radadeduage[is.na(repl$radadeduage)] <- 0
-
-repl.final2$ramomeduage[is.na(repl.final2$ramomeduage)] <- 0
-repl2$ramomeduage[is.na(repl$ramomeduage)] <- 0
-
-## USE THIS SPACE TO HOTDECK LOTS OF MISSING VARIABLES (start with lnly)
-```
-
-```{r}
-## FINALLY SAVE THE DATA
-write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
-write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
-```

### 2021-11-02 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added visualisations for the first replacement scenario (risk behaviours switched with quint3) (some visualisations are broken). Also generated a new replacement population, where risk behaviours, disability status, and health status are switched from quintile 4 onto quintile 1

@@
 
+#### Risk Behaviours only
+
 ```{r}
@@
 rm(imp, test, test.1, test.3)
+
 ```
@@
 ## FINALLY SAVE THE DATA
-write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
+write_dta(repl.final2, 'input_data/ELSA_repl_risk_quint3.dta', version = 12)
 write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
+
+rm(repl.final2, repl2, hotdeck.vars, demog.vars, risk.vars, id.var,
+   repl.final, test2)
+```
+
+#### Risk Behaviours, ADLs, and Health Status
+
+```{r}
+# Set up lists of id vars, demographic vars (for imputation) and risk vars to be imputed
+id.var <- 'idauniq'
+demog.vars <- c('wealth_quint', 'age', 'male')
+change.vars <- c('smokev', 'l2smokev', 'smoken', 'l2smoken', 'heavy_smoker',
+               'l2heavy_smoker', 
+               'drink', 'l2drink', 'problem_drinker', 'l2problem_drinker', 
+               'logbmi', 'l2logbmi', 
+               'exstat', 'l2exstat', 
+               'adlstat', 'l2adlstat', 'iadlstat', 'l2iadlstat',
+               'cancre', 'l2cancre', 'diabe', 'l2diabe', 'hearte', 'l2hearte',
+               'lunge', 'l2lunge', 'hibpe', 'l2hibpe', 'stroke', 'l2stroke',
+               'asthmae', 'l2asthmae')
+
+test <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 4)) %>%
+                  select(all_of(id.var), all_of(demog.vars), all_of(change.vars))
+
+# separate test df by quintile so we can remove data for quint1
+test.1 <- test %>% filter(wealth_quint == 1)
+test.4 <- test %>% filter(wealth_quint == 4)
+test.1[change.vars] <- NA
+test <- rbind(test.1, test.4)
+
+# Run mi & get a complete version of the imputed dataset
+imp <- mice(data=test, m=5, maxit=10, printFlag = FALSE)
+imputed <- complete(imp)
+
+## OK!
+## The test2 dataset now contains fully imputed data for quintile 1 based on
+## values from quintile 3 so far. Can easily turn this into a function and
+## use different quintiles (or just do them without the function)
+
+rm(imp, test, test.1, test.4)
+
+```
+
+```{r}
+### ACCOUNTING
+## The imputation doesn't work perfectly, and a lot of the lagged variables are not imputed in that step as they are correlated with their current
+
+# copy so we don't mess with original
+accounted <- imputed
+
+## CHRONIC CONDITIONS
+# Handle chronic conditions first as they are the easiest to impute
+absorbing_vars <- c('cancre', 'diabe', 'hearte',
+                     'lunge', 'hibpe', 'stroke',
+                     'asthmae')
+
+## If var is absorbing, then 0 in var.current must also be 0 in var.lagged
+for (var in absorbing_vars) {
+  accounted[[paste0('l2', var)]][accounted[[var]] == 0] <- 0
+  accounted[[paste0('l2', var)]][accounted[[var]] == 1] <- 1
+}
+
+## specific issue with stroke
+# Stroke didn't impute, but looking at data for the quantile we swapped with
+# not a single case of stroke, so going to set all to 0
+accounted$stroke <- 0
+accounted$l2stroke <- 0
+
+## ADLs/IADLs
+# if adlstat or iadlstat == 1, then lag variable should also be 1
+# i.e. if no disabilitys in current, then previous also no disabilities
+accounted$l2adlstat[accounted$adlstat == 1] <- 1
+accounted$l2iadlstat[accounted$iadlstat == 1] <- 1
+
+## RISK BEHAVIOURS
+# if smoken == 1 then most likely l2smoken == 1 also
+accounted$l2smoken[is.na(accounted$l2smoken)] <- accounted$smoken[is.na(accounted$l2smoken)]
+
+# if l2smokev == 1 then smokev must also == 1
+accounted$smokev[is.na(accounted$smokev)] <- accounted$l2smokev[is.na(accounted$smokev)]
+# also, if l2smoken == 1 then l2smokev must also equal 1
+accounted$l2smokev[accounted$l2smoken == 1] <- 1
+# and if l2smokev == 1 then smokev == 1
+accounted$smokev[accounted$l2smokev == 1] <- 1
+# same for smoken
+accounted$l2smoken[accounted$smoken == 1] <- 1
+# going to assume that someone who is a heavy smoker now was a heavy smoker in lag as well
+accounted$l2heavy_smoker[is.na(accounted$l2heavy_smoker)] <- accounted$heavy_smoker[is.na(accounted$l2heavy_smoker)]
+
+# if person is a problem drinker then they must also be drink == 1
+accounted$drink[accounted$l2problem_drinker == 1] <- 1
+# probably unlikely that someone would change drink status, so set lag == current
+accounted$l2drink[is.na(accounted$l2drink)] <- accounted$drink[is.na(accounted$l2drink)]
+# no variability in problem_drinker -> l2problem_drinker in median wealth group,
+# so assume same here
+accounted$l2problem_drinker[is.na(accounted$l2problem_drinker)] <- accounted$problem_drinker
+```
+
+```{r}
+## Next step is to replace the values from the original repl dataset and save 
+# this as an alternative for running through the model
+
+# copy this so we don't mess with the original repl
+repl.quint1 <- repl %>% filter(wealth_quint == 1) %>%
+                          select(-all_of(change.vars), -all_of(demog.vars))
+
+# get just the quint 1 people
+quint1.replaced <- accounted %>% filter(wealth_quint == 1)
+
+# merge the modified repl with the newly replaced quint1 people
+repl.replaced <- merge(repl.quint1, quint1.replaced, by='idauniq', all.x = TRUE)
+
+# Now filter all the quint1 people from original repl and row bind the new people
+repl2 <- repl %>% filter(wealth_quint != 1)
+
+repl.final <- rbind(repl.replaced, repl2)
+
+rm(repl.quint1, quint1.replaced, repl.replaced, repl2, accounted, imputed)
+```
+
+```{r}
+
+## IMPUTING DATA TO RUN THROUGH MODEL
+
+## Whatever we do in here have to do to the baseline repl as well just to make 
+# sure things are standardised
+
+repl.final2 <- repl.final
+repl2 <- repl
+
+## handle missing educl data as described above
+repl.final2$educl[is.na(repl.final2$educl)] <- 0
+repl2$educl[is.na(repl$educl)] <- 0
+
+#table(repl.final2$educl)
+#summary(repl.final2$educl)
+#print('---------')
+
+repl.final2$radadeduage[is.na(repl.final2$radadeduage)] <- 0
+repl2$radadeduage[is.na(repl$radadeduage)] <- 0
+
+repl.final2$ramomeduage[is.na(repl.final2$ramomeduage)] <- 0
+repl2$ramomeduage[is.na(repl$ramomeduage)] <- 0
+
+## USE THIS SPACE TO HOTDECK LOTS OF MISSING VARIABLES (start with lnly)
+hotdeck.vars <- c('lnly', 'workstat', 'smokev')
+
+repl.final2 <- hotdeck(repl.final2, variable = hotdeck.vars)
+repl2 <- hotdeck(repl2, variable = hotdeck.vars)
+
+rm(hotdeck.vars, var)
+```
+
+```{r}
+## FINALLY SAVE THE DATA
+write_dta(repl.final2, 'input_data/ELSA_repl_rdh_quint4.dta', version = 12)
+write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
+
+rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var)
 ```
@@
 
-## Data
+## Baseline
+
+### Data
 ```{r}
@@
 
-## Survival Curve
+### Survival Curve
 ```{r}
@@
 # Function to compare survival curves between quintiles
-
-## THIS IS NOT USEFUL!!!
-# Because we calculate survival using the m_endpop_* variable (which is a 
-# population level variable) it will always give the same value for every
-# subgroup.
-# Need to talk to Bryan about a way to analyse this within subgroup.
 survivalCurveSubgroup <- function(cohort) {
@@
 
-## Prevalences
+### Prevalences
 ```{r}
@@
   # Select just the things we want
-  #prev.all <- cohort %>% select(Age, p_anydisease_wq1, p_anydisease_wq2, p_anydisease_wq3, p_anydisease_wq4, p_anydisease_wq5)
   prev.all <- cohort %>% select(Age, all_of(varnames))
-
-  # # pivot the data to long format for ggplot
-  # survivals <- pivot_longer(data = prev.all,
-  #                           cols = p_anydisease_wq1:p_anydisease_wq5,
-  #                           names_to = "Quintile",
-  #                           values_to = "Disease_Prev")
 
@@
 ```
+
+## Replacement Scenarios
+
+```{r}
+cohort <- read_dta("output/WEALTH/ELSA_cohort/ELSA_cohort_summary.dta")
+replaced1 <- read_dta("output/WEALTH/ELSA_risk_quint3/ELSA_risk_quint3_summary.dta")
+replaced2 <- read_dta("output/WEALTH/ELSA_rdh_quint4/ELSA_rdh_quint4_summary.dta")
+```
+
+### Survival Curve
+```{r "quint3"}
+
+survivalCurveReplaced <- function(cohort, replaced) {
+
+  # Age and survival for each quintile
+  cohort$Age <- (cohort$year - 2012) + 50
+  cohort$q1 <- cohort$m_endpop_wq1 / cohort$m_endpop_wq1[1]
+  cohort$q3 <- cohort$m_endpop_wq3 / cohort$m_endpop_wq3[1]
+  replaced$Age <- (replaced$year - 2012) + 50
+  replaced$q1 <- replaced$m_endpop_wq1 / replaced$m_endpop_wq1[1]
+
+  # Cut dataframe down to just the age and survivals
+  #survivals <- cohort %>% select(Age, q1, q2, q3, q4, q5)
+  survivals <- data.frame(cohort$Age, cohort$q1, cohort$q3, replaced$q1)
+
+  survivals <- survivals %>%
+                  dplyr::rename('Age' = cohort.Age, 
+                                'CohortQ1' = cohort.q1, 
+                                'CohortQ3' = cohort.q3, 
+                                'ReplacedQ1' = replaced.q1) %>%
+                  pivot_longer(cols = CohortQ1:ReplacedQ1,
+                               names_to = 'Source',
+                               values_to = 'Survival')
+
+  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Survival, color = Source)) +
+    geom_smooth(se = FALSE, linetype = 'solid', size=0.4)
+
+  # return.object <- list(survivals, plot)
+  # 
+  # return(return.object)
+  return(p)
+}
+
+survivalCurveReplaced(cohort, replaced1)
+
+```
+```{r "quint4"}
+
+survivalCurveReplaced <- function(cohort, replaced) {
+
+  # Age and survival for each quintile
+  cohort$Age <- (cohort$year - 2012) + 50
+  cohort$q1 <- cohort$m_endpop_wq1 / cohort$m_endpop_wq1[1]
+  cohort$q4 <- cohort$m_endpop_wq4 / cohort$m_endpop_wq4[1]
+  replaced$Age <- (replaced$year - 2012) + 50
+  replaced$q1 <- replaced$m_endpop_wq1 / replaced$m_endpop_wq1[1]
+
+  # Cut dataframe down to just the age and survivals
+  #survivals <- cohort %>% select(Age, q1, q2, q3, q4, q5)
+  survivals <- data.frame(cohort$Age, cohort$q1, cohort$q4, replaced$q1)
+
+  survivals <- survivals %>%
+                  dplyr::rename('Age' = cohort.Age, 
+                                'CohortQ1' = cohort.q1, 
+                                'CohortQ3' = cohort.q4, 
+                                'ReplacedQ1' = replaced.q1) %>%
+                  pivot_longer(cols = CohortQ1:ReplacedQ1,
+                               names_to = 'Source',
+                               values_to = 'Survival')
+
+  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Survival, color = Source)) +
+    geom_smooth(se = FALSE, linetype = 'solid', size=0.4)
+
+  # return.object <- list(survivals, plot)
+  # 
+  # return(return.object)
+  return(p)
+}
+
+survivalCurveReplaced(cohort, replaced2)
+
+```
+
+### Prevalences
+```{r}
+
+subgroupPrevReplaced <- function(cohort, replaced, var, subgroups) {
+
+  # Age
+  cohort$Age <- (cohort$year - 2012) + 50
+  replaced$Age <- (replaced$year - 2012) + 50
+
+  # Build varnames
+  varnames <- c()
+  ticker <- 1
+  for (i in subgroups) {
+    varnames[ticker] <- paste0('p_', var, '_wq', i)
+    print(varnames[ticker])
+    ticker <- ticker + 1
+  }
+
+  # Select vars we've just built
+  cohort.prev <- cohort %>% select(Age, all_of(varnames))
+  replaced.prev <- cohort %>% select(Age, all_of(varnames))
+
+  prevs <- merge(cohort.prev, replaced.prev, by = 'Age', suffixes = c('.Cohort', '.Replaced'))
+
+  print(colnames(prevs))
+
+  prevs <- pivot_longer(data = prevs,
+                        cols = paste0(varnames[1], '.Cohort'):paste0(varnames[length(varnames)], '.Replaced'),
+                        names_to = 'Source',
+                        values_to = 'Prevalence')
+  print(colnames(prevs))
+
+  p <- ggplot(data = prevs, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+      geom_smooth(se=FALSE, size=0.4)
+
+  return(p)
+
+}
+
+subgroupPrevReplaced(cohort, replaced1, var = 'anydisease', subgroups = c(1,3))
+# subgroupPrevReplaced(cohort, replaced1, var = 'disabled', subgroups = c(1,3))
+# subgroupPrevReplaced(cohort, replaced1, var = 'severeCondition', subgroups = c(1,3))
+# subgroupPrevReplaced(cohort, replaced1, var = 'mildCondition', subgroups = c(1,3))
+# 
+# subgroupPrevReplaced(cohort, replaced2, var = 'anydisease', subgroups = c(1,4))
+# subgroupPrevReplaced(cohort, replaced2, var = 'disabled', subgroups = c(1,4))
+# subgroupPrevReplaced(cohort, replaced2, var = 'severeCondition', subgroups = c(1,4))
+# subgroupPrevReplaced(cohort, replaced2, var = 'mildCondition', subgroups = c(1,4))
+
+#someSubgroupPrevalence(cohort, var = 'anydisease', subgroups = c(1,3))
+
+```
+
+##### Any Disease
+```{r}
+
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_anydisease_wq1, p_anydisease_wq4) %>%
+                dplyr::rename('OrigPoor' = p_anydisease_wq1, 'OrigWealth' = p_anydisease_wq4)
+r2.prev <- r2 %>% select(Age, p_anydisease_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_anydisease_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  geom_smooth(se=FALSE, size=0.4) +
+  labs(title = 'Any Disease')
+
+```
+
+##### Disabled
+```{r}
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_disabled_wq1, p_disabled_wq4) %>%
+                dplyr::rename('OrigPoor' = p_disabled_wq1, 'OrigWealth' = p_disabled_wq4)
+r2.prev <- r2 %>% select(Age, p_disabled_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_disabled_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  #geom_smooth(se=FALSE, size=0.4) +
+  geom_line() +
+  labs(title = 'Disabled')
+```
+
+##### Severe Condition
+```{r}
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_severeCondition_wq1, p_severeCondition_wq4) %>%
+                dplyr::rename('OrigPoor' = p_severeCondition_wq1, 'OrigWealth' = p_severeCondition_wq4)
+r2.prev <- r2 %>% select(Age, p_severeCondition_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_severeCondition_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  geom_smooth(se=FALSE, size=0.4) +
+  labs(title = 'Severe Condition')
+```
+
+##### Mild Condition
+```{r}
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_mildCondition_wq1, p_mildCondition_wq4) %>%
+                dplyr::rename('OrigPoor' = p_mildCondition_wq1, 'OrigWealth' = p_mildCondition_wq4)
+r2.prev <- r2 %>% select(Age, p_mildCondition_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_mildCondition_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  geom_smooth(se=FALSE, size=0.4) +
+  labs(title = 'Mild Condition')
+```
+
+##### cancre
+```{r}
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_cancre_wq1, p_cancre_wq4) %>%
+                dplyr::rename('OrigPoor' = p_cancre_wq1, 'OrigWealth' = p_cancre_wq4)
+r2.prev <- r2 %>% select(Age, p_cancre_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_cancre_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  geom_smooth(se=FALSE, size=0.4)
+
+c <- cohort
+r2 <- replaced2
+
+c$Age <- (c$year - 2012) + 50
+r2$Age <- (r2$year - 2012) + 50
+
+c.prev <- c %>% select(Age, p_hearte_wq1, p_hearte_wq4) %>%
+                dplyr::rename('OrigPoor' = p_hearte_wq1, 'OrigWealth' = p_hearte_wq4)
+r2.prev <- r2 %>% select(Age, p_hearte_wq1) %>%
+                dplyr::rename('ReplacedPoor' = p_hearte_wq1)
+
+combined <- merge(c.prev, r2.prev, by = 'Age')
+
+combined.pivot <- pivot_longer(data = combined,
+                               cols = OrigPoor:ReplacedPoor,
+                               names_to = 'Source',
+                               values_to = 'Prevalence')
+
+ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
+  geom_smooth(se=FALSE, size=0.4)
+```

### 2021-11-02 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added some code to check the initial conditions of the repl_standard population (eventually will collect into function to run for each initial population). Will also eventually look at these specifics by wealth_quintile. Also used ggplot to generate BMI histograms and overlay

@@
 require(haven)
-require(foreign)
 
@@
 
-rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var)
+rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var, repl)
 ```
@@
 ## Baseline
+
+### Initial Conditions
+
+We want to check the initial conditions of each population (or at least for each quintile). Focus on these variables first:
+- Gender
+- Risk behaviours (Smoking, Drinking, BMI, Exercise status)
+- ADLs/IADLs (Disabled?)
+- Severe/Mild Condition?
+
+```{r "summary output FAIL"}
+cohort <- read_dta('output/WEALTH/ELSA_cohort/ELSA_cohort_summary.dta')
+
+### For first run through of this lets look at the summary output. Half of the work is already done for us here
+
+select.vars <- c('male', 'smoke', 'drink', 'exstat', 'bmi', 'disabled', 'disease', 'Condition')
+
+## To check the initial conditions we only need 2012 info (as this is before simulation)
+# Then remove the _all _m & _f columns as we only want outputs by wealth quintile
+init.c <- cohort %>% filter(year == 2012) %>%
+                      select(-ends_with(c('_all', '_m', '_f'))) %>%
+                      select(year, contains(all_of(select.vars))) %>%
+                      select(-starts_with(c('i', 'n')))
+# remove the p_ prefix
+colnames(init.c) <- gsub('p_', '', colnames(init.c))
+# now pivot to get data by quintile
+init.c.pivot <-  pivot_longer(data = init.c,
+                              cols = male_wq1:severeCondition_wq5,
+                              names_to = c('Variable', 'Quintile'),
+                              names_sep = '_w',
+                              values_to = 'Prevalence')
+
+## Summary output is actually making this a bit more difficult than it has to be
+# Instead lets look at the repl populations and make a function to do all this
+
+rm(init.c, init.c.pivot, cohort)
+```
+
+```{r "repl_populations"}
+
+repl <- read_dta('input_data/ELSA_repl_standard.dta')
+repl <- zap_labels(repl)
+repl.2012 <- repl %>% filter(year == 2012)
+
+## First step is to group by wealth_quint and then start producing summary stats
+repl.sum <- repl.2012 %>%
+                group_by(wealth_quint) %>%
+                summarise(
+                  count = n(),
+                  mean_age = mean(age),
+                  mean_male = mean(male),
+                  smoken = mean(smoken),
+                  smokev = mean(smokev),
+                  heavy_smoker = mean(heavy_smoker),
+                  mean_BMI = mean(bmi, na.rm = TRUE),
+                  sd_BMI = sd(bmi, na.rm = TRUE),
+                  missing_BMI_percent = ((sum(is.na(bmi)) / n()) * 100)
+                )
+
+## histogram of BMI by quintile
+repl.bmi <- repl.2012 %>% select(wealth_quint, bmi) %>%
+                          filter(wealth_quint %in% c(1,3))
+repl.bmi <- drop_na(repl.bmi)
+repl.bmi$wealth_quint <- as.factor(repl.bmi$wealth_quint)
+
+ggplot(repl.bmi, aes(x = bmi, color = wealth_quint)) +
+  geom_histogram()
+```
 

### 2021-11-03 FEM_R/Wealth_Health_Socioeconomics.Rmd
Fixed accounting of adlstat/iadlstat dummys in generation of the ELSA_repl_rdh_quint4.dta file. Also fixed the visualisation function for prevalences by subgroup between baseline and intervention output files

@@
 accounted$l2iadlstat[accounted$iadlstat == 1] <- 1
+# Now do all the dummys also
+# ADLs & l2ADLs
+accounted$adl1[accounted$adlstat == 2] <- 1
+accounted$adl1[accounted$adlstat != 2] <- 0
+accounted$adl2[accounted$adlstat == 3] <- 1
+accounted$adl2[accounted$adlstat != 3] <- 0
+accounted$adl3p[accounted$adlstat == 4] <- 1
+accounted$adl3p[accounted$adlstat != 4] <- 0
+# IADLs & l2IADLs
+accounted$iadl1[accounted$iadlstat == 2] <- 1
+accounted$iadl1[accounted$iadlstat != 2] <- 0
+accounted$iadl2p[accounted$iadlstat == 3] <- 1
+accounted$iadl2p[accounted$iadlstat != 3] <- 0
 
@@
 
+# Have to remove these too as they were handled in accounting
+adl.vars <- c('adl1', 'adl2', 'adl3p', 'iadl1', 'iadl2p')
+
 # copy this so we don't mess with the original repl
 repl.quint1 <- repl %>% filter(wealth_quint == 1) %>%
-                          select(-all_of(change.vars), -all_of(demog.vars))
+                          select(-all_of(change.vars), -all_of(demog.vars),
+                                 -all_of(adl.vars))
 
@@
     varnames[ticker] <- paste0('p_', var, '_wq', i)
-    print(varnames[ticker])
     ticker <- ticker + 1
@@
   # Select vars we've just built
-  cohort.prev <- cohort %>% select(Age, all_of(varnames))
-  replaced.prev <- cohort %>% select(Age, all_of(varnames))
+  cohort.prev <- cohort %>% select(Age, all_of(varnames)) %>%
+                            dplyr::rename('OrigPoor' = varnames[1],
+                                          'OrigWealthy' = varnames[2])
+  replaced.prev <- replaced %>% select(Age, varnames[1]) %>%
+                              dplyr::rename('ReplacedPoor' = varnames[1])
 
-  prevs <- merge(cohort.prev, replaced.prev, by = 'Age', suffixes = c('.Cohort', '.Replaced'))
-
-  print(colnames(prevs))
+  prevs <- merge(cohort.prev, replaced.prev, by = 'Age')
 
   prevs <- pivot_longer(data = prevs,
-                        cols = paste0(varnames[1], '.Cohort'):paste0(varnames[length(varnames)], '.Replaced'),
+                        cols = OrigPoor:ReplacedPoor,
                         names_to = 'Source',
                         values_to = 'Prevalence')
-  print(colnames(prevs))
 
   p <- ggplot(data = prevs, mapping = aes(x = Age, y = Prevalence, color = Source)) +
-      geom_smooth(se=FALSE, size=0.4)
+      geom_smooth(se=FALSE, size=0.4) +
+      labs(title = paste0('Var: ', var, '; Quintiles: [', subgroups[1], ', ', subgroups[2], ']'))
 
@@
 subgroupPrevReplaced(cohort, replaced1, var = 'anydisease', subgroups = c(1,3))
-# subgroupPrevReplaced(cohort, replaced1, var = 'disabled', subgroups = c(1,3))
-# subgroupPrevReplaced(cohort, replaced1, var = 'severeCondition', subgroups = c(1,3))
-# subgroupPrevReplaced(cohort, replaced1, var = 'mildCondition', subgroups = c(1,3))
-# 
-# subgroupPrevReplaced(cohort, replaced2, var = 'anydisease', subgroups = c(1,4))
-# subgroupPrevReplaced(cohort, replaced2, var = 'disabled', subgroups = c(1,4))
-# subgroupPrevReplaced(cohort, replaced2, var = 'severeCondition', subgroups = c(1,4))
-# subgroupPrevReplaced(cohort, replaced2, var = 'mildCondition', subgroups = c(1,4))
+subgroupPrevReplaced(cohort, replaced1, var = 'disabled', subgroups = c(1,3))
+subgroupPrevReplaced(cohort, replaced1, var = 'severeCondition', subgroups = c(1,3))
+subgroupPrevReplaced(cohort, replaced1, var = 'mildCondition', subgroups = c(1,3))
 
-#someSubgroupPrevalence(cohort, var = 'anydisease', subgroups = c(1,3))
+subgroupPrevReplaced(cohort, replaced2, var = 'anydisease', subgroups = c(1,4))
+subgroupPrevReplaced(cohort, replaced2, var = 'disabled', subgroups = c(1,4))
+subgroupPrevReplaced(cohort, replaced2, var = 'severeCondition', subgroups = c(1,4))
+subgroupPrevReplaced(cohort, replaced2, var = 'mildCondition', subgroups = c(1,4))
 

### 2021-11-22 FEM_R/impute_alcbase.Rmd
Imputed alcbase information for drinkers who report not having drank anything in the week before the survey

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(mice)

```

## Multiple imputation

We will do multiple imputation of

```{r}
wv <- read_dta('input_data/H_ELSA_g2_wv_specific.dta')
```

Now set up lists of the variables we're going to keep and impute. This will have to be the wave specific variables 
```{r}

id.var <- 'idauniq'
demog.vars <- c('ragender', 'raeducl')
age.vars <- c('r4agey', 'r5agey', 'r6agey', 'r7agey', 'r8agey', 'r9agey')
scako.vars <- c('r4scako', 'r5scako', 'r6scako', 'r7scako', 'r8scako', 'r9scako')
alc.vars <- c('r4alcbase', 'r5alcbase', 'r6alcbase', 'r7alcbase', 'r8alcbase', 
              'r9alcbase')

# first extract just these vars from original data
test <- wv %>% select(all_of(id.var), all_of(demog.vars), all_of(age.vars),
                      all_of(scako.vars), all_of(alc.vars))

# Calculate the annual and monthly consumption
for (v in alc.vars) {
  newVarName.a <- paste0(v, '_a')
  test[[newVarName.a]] <- test[[v]] * 52

  newVarName.m <- paste0(v, '_m')
  test[[newVarName.m]] <- test[[v]] * 4
}

#imp <- mice(data=test, m=5, maxit=10, printFlag = TRUE)

```

Right, data is prepared for mice, but we need to make it work in a very specific way.
Information must be imputed within scako groups. So I think a multi-layered for loop is the only way to do this. I am envisaging:
- Loop over alc.vars
- Loop over `wv'scako values
- Generate a df of

```{r}

miss.list <- list()
# generate a separate df for each wave
imp.list <- list()
for (x in 4:9) {
  scak <- paste0('r', x, 'scako')
  alcb <- paste0('r', x, 'alcbase')

  newDF <- test %>% select(all_of(id.var), all_of(demog.vars),
                                  paste0('r', x, 'agey'),
                                  all_of(scak), 
                                  all_of(alcb))

  # Extract records where scako != 1:7 (i.e. missing or not drank in last 12 months)
  scako.missings <- newDF[!newDF[[scak]] %in% 1:7,]

  # Add to a list so we can rejoin later
  miss.list[[x]] <- scako.missings

  # Now select only the non-missing drinkers
  impDF <- newDF[newDF[[scak]] %in% 1:7,]

  #impDF[[alcb]][impDF[[alcb]] == 0.0,] <- NA
  impDF[[alcb]][impDF[[alcb]] == 0] <- NA

  imp <- mice(data = impDF, m=5, maxit=10, printFlag = FALSE)
  imp.complete <- rbind(complete(imp), miss.list[[x]])

  # Make sure anyone who said they haven't drank in last 12 months has alcbase == 0
  imp.complete[[alcb]][imp.complete[[scak]] == 8] <- 0
  # and finally drop anything that isn't wave specific for the merge in the next chunk
  imp.complete <- imp.complete %>% select(-ragender, -raeducl)

  # save to the list
  imp.list[[x]] <- imp.complete
}

```

```{r}

# Last step is to merge all the dataframes together on the id var and save 
# so it can be merged with the original data in stata
# Also remove all other vars but the alcbase vars as we only want the imputed versions of them
final.imputed <- imp.list %>% discard(is.null) %>%
                reduce(inner_join, by = "idauniq")

# drop age and scako too? Yes for first attempt
final.imputed <- final.imputed %>% select(-contains('scako'), -contains('agey'))

write_dta(final.imputed, path = 'input_data/alcbase_imputed.dta')
```

### 2022-03-01 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
Changed some figures to greyscale in IJM analyses

--- 2021-06-28 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
@@
 require(reshape2)
-
 theme_set(theme_bw())
-
 workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
 knitr::opts_knit$set(root.dir = workingDir)
-
 rm(workingDir)
@@
   t <- t %>%
-          summarise(n = n(), newIn = sum(incidence.counts, na.rm=TRUE)) %>%
+          dplyr::summarise(n = n(), newIn = sum(incidence.counts, na.rm=TRUE)) %>%
           mutate(incidence.rate = (newIn / n) * 100)
@@
 }
-
 getIncidenceRate(trans, 'cancre')
@@
 ```{r echo = FALSE}
-# #AgeUK
-# male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE.csv')
-# female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE.csv')
-# 
-# # FEM 2014 detailed output
-# #FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
-# ## extract males and females
-# #mFEM_2014 <- FEM_2014 %>% filter(male == 1)
-# #fFEM_2014 <- FEM_2014 %>% filter(male == 0)
-# 
-# # FEM 2014 summary
-# #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
-# FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
-# 
-# # Keep only year 2014 from FEM summary
-# FEM <- filter(FEM, year == 2014)
-# 
-# # Specify the age groups to extract from FEM
-# ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
-# # Add m and f for male and female variables
-# m_ages <- paste0('m_', ages)
-# f_ages <- paste0('f_', ages)
-# 
-# # Collect the age suffixes into a vector
-# keep_vars <- c(m_ages, f_ages)
-# # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
-# diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
-# 
-# # Don't like using loops but this one was just too simple to bother with figuring out the apply version
-# final <- c()
-# ticker <- 1
-# # Generating the column (variable) names we want to keep in a loop
-# for(item in diseases) {
-#   for(anotherItem in keep_vars) {
-#     final[ticker] <- paste0('p_', item, '_', anotherItem)
-#     ticker <- ticker + 1
-#   }
-# }
-# 
-# # Keep the right vars using varnames generated in loop
-# FEM <- FEM %>%
-#           select(c('year', all_of(final)))
-# 
-# # Rename some vars to remove 'p_' before merging
-# FEM <- FEM %>%
-#           rename_at(vars(starts_with('p_')),
-#                     funs(str_replace(., 'p_', ''))) %>%
-#           select(!year)
-# 
-# pivoted.FEM <- pivot_longer(data = FEM,
-#                             cols = hibpe_m_6064:asthmae_f_100p, 
-#                             names_to = c('Condition', 'Gender', 'Age_Group'), 
-#                             names_sep = '_')
-# 
-# # Convert value from fraction to percentage
-# pivoted.FEM$value <- pivoted.FEM$value * 100
-# 
+#AgeUK
+male_au <- read.csv('../Validation/AgeUK-chronDisease-MALE_CI.csv')
+female_au <- read.csv('../Validation/AgeUK-chronDisease-FEMALE_CI.csv')
+# FEM 2014 detailed output
+#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
+## extract males and females
+#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
+#fFEM_2014 <- FEM_2014 %>% filter(male == 0)
+# FEM 2014 summary
+#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+FEM <- read_dta('output/COMPLETE/ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
+# Keep only year 2014 from FEM summary
+FEM <- filter(FEM, year == 2014)
+# Specify the age groups to extract from FEM
+ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
+# Add m and f for male and female variables
+m_ages <- paste0('m_', ages)
+f_ages <- paste0('f_', ages)
+# Collect the age suffixes into a vector
+keep_vars <- c(m_ages, f_ages)
+# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
+diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
+# Don't like using loops but this one was just too simple to bother with figuring out the apply version
+final <- c()
+ticker <- 1
+# Generating the column (variable) names we want to keep in a loop
+for(item in diseases) {
+  for(anotherItem in keep_vars) {
+    final[ticker] <- paste0('p_', item, '_', anotherItem)
+    ticker <- ticker + 1
+  }
+}
+# Keep the right vars using varnames generated in loop
+FEM <- FEM %>%
+          select(c('year', all_of(final)))
+# Rename some vars to remove 'p_' before merging
+FEM <- FEM %>%
+          rename_at(vars(starts_with('p_')),
+                    funs(str_replace(., 'p_', ''))) %>%
+          select(!year)
+pivoted.FEM <- pivot_longer(data = FEM,
+                            cols = hibpe_m_6064:asthmae_f_100p,
+                            names_to = c('Condition', 'Gender', 'Age_Group'),
+                            names_sep = '_')
+# Convert value from fraction to percentage
+pivoted.FEM$value <- pivoted.FEM$value * 100
+# Set gender codes to upper case
+pivoted.FEM$Gender <- toupper(pivoted.FEM$Gender)
 # # Rename cols to enable reshaping
@@
 #                 select(!Total)
-# 
-# # Now need to reshape these datasets to long format
-# pivoted.male <- pivot_longer(data = male_au,
-#                              cols = m_6064:m_100p,
-#                              names_to = c('Gender', 'Age_Group'),
-#                              names_sep = '_')
-# 
-# pivoted.female <- pivot_longer(data = female_au,
-#                              cols = f_6064:f_100p,
-#                              names_to = c('Gender', 'Age_Group'),
-#                              names_sep = '_')
-# 
-# # Now row bind these frames together, combining male and female data
-# ageUK <-rbind(pivoted.male, pivoted.female)
-# 
-# # Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
-# # extract all heart related conditions for combining
-# #ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
-# # Aggregate the values by gender and age group
-# #ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
-# # Rename columns and add condition column back
-# #colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
-# #ageUK_heart$Condition <- 'Heart Problems'
-# 
-# # Now drop the heart problems from original ageUK and replace with aggregate
-# ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
-# #ageUK <- rbind(ageUK, ageUK_heart)
-# 
-# # Rename a few just for completeness
-# #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
-# #ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
-# 
-# # Finally, add column to specify that this is ageUK data
-# ageUK$source <- "ageUK"
-# 
-# # REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
-# #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
-# ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
-# ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
-# 
-# # Make copy
-# newFEM <- pivoted.FEM
-# 
-# # Rename all the conditions to fit ageUK
-# # NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
-# #       Therefore only validation from this will be that FEM is higher than ageUK
-# #newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
-# newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
-# newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
-# newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
-# #newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
-# #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
-# newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
-# newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
-# 
-# 
-# #force Make to run this again
-# 
-# # Add column to specify this is FEM data
-# newFEM$source <- "FEM"
-# 
-# # Now row bind the datasets
-# combined <- rbind(newFEM, ageUK)
-# # Specify factor levels for age group
-# combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
-# # Sort
-# combined <- arrange(combined, source, Condition, Gender, Age_Group)
-# 
-# # Function to filter the combined dataset and visualise
-# visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
-#   
-#   if(gender == 'm') {
-#     full_gender <- 'Male'
-#   }
-#   else if(gender == 'f') {
-#     full_gender <- 'Female'
-#   }
-#   
-#   # filter by gender and condition
-#   data <- data %>%
-#             filter(Gender == gender) %>%
-#             filter(Condition == cd)
-#   
-#   # now plot!
-#   p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
-#           geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
-#           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
-#   
-#   # Save plot into in E-FEM_Paper/figures/
-#   ggsave(filename = paste0(cd, '_', gender, '.png'),
-#          plot = p,
-#          path = path,
-#          width = 6,
-#          height = 4)
-#   
-#   return(p)
-# }
+male_au <- male_au %>%
+                dplyr::rename(M_6064_total = M_60.64_total, M_6064_lower = M_60.64_l, M_6064_upper = M_60.64_u, 
+                              M_6569_total = M_65.69_total, M_6569_lower = M_65.69_l, M_6569_upper = M_65.69_u,
+                              M_7074_total = M_70.74_total, M_7074_lower = M_70.74_l, M_7074_upper = M_70.74_u,
+                              M_7579_total = M_75.79_total, M_7579_lower = M_75.79_l, M_7579_upper = M_75.79_u,
+                              M_8084_total = M_80.84_total, M_8084_lower = M_80.84_l, M_8084_upper = M_80.84_u,
+                              M_8589_total = M_85.89_total, M_8589_lower = M_85.89_l, M_8589_upper = M_85.89_u,
+                              M_9094_total = M_90.94_total, M_9094_lower = M_90.94_l, M_9094_upper = M_90.94_u,
+                              M_9599_total = M_95.99_total, M_9599_lower = M_95.99_l, M_9599_upper = M_95.99_u,
+                              M_100p_total = M_Over.100_total, M_100p_lower = M_Over.100_l, M_100p_upper = M_Over.100_u) %>%
+                select(-M_Total_total, -M_Total_l, -M_Total_u)
+female_au <- female_au %>%
+                dplyr::rename(F_6064_total = F_60.64_total, F_6064_lower = F_60.64_l, F_6064_upper = F_60.64_u, 
+                              F_6569_total = F_65.69_total, F_6569_lower = F_65.69_l, F_6569_upper = F_65.69_u,
+                              F_7074_total = F_70.74_total, F_7074_lower = F_70.74_l, F_7074_upper = F_70.74_u,
+                              F_7579_total = F_75.79_total, F_7579_lower = F_75.79_l, F_7579_upper = F_75.79_u,
+                              F_8084_total = F_80.84_total, F_8084_lower = F_80.84_l, F_8084_upper = F_80.84_u,
+                              F_8589_total = F_85.89_total, F_8589_lower = F_85.89_l, F_8589_upper = F_85.89_u,
+                              F_9094_total = F_90.94_total, F_9094_lower = F_90.94_l, F_9094_upper = F_90.94_u,
+                              F_9599_total = F_95.99_total, F_9599_lower = F_95.99_l, F_9599_upper = F_95.99_u,
+                              F_100p_total = F_Over.100_total, F_100p_lower = F_Over.100_l, F_100p_upper = F_Over.100_u) %>%
+                select(-F_Total_total, -F_Total_l, -F_Total_u)
+# Now need to reshape these datasets to long format
+pivoted.male <- pivot_longer(data = male_au,
+                             cols = M_6064_total:M_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
+                             names_sep = '_')
+pivoted.female <- pivot_longer(data = female_au,
+                             cols = F_6064_total:F_100p_upper,
+                             names_to = c('Gender', 'Age_Group', 'Type'),
+                             names_sep = '_')
+# Now bring the confidence intervals back into columns because ggplot needs that format
+# to plot them. Seems pretty daft to be honest but thats what we've got
+pivoted.male <- pivot_wider(pivoted.male,
+                            names_from = Type,
+                            values_from = value)
+pivoted.female <- pivot_wider(pivoted.female,
+                            names_from = Type,
+                            values_from = value)
+# Now row bind these frames together, combining male and female data
+ageUK <-rbind(pivoted.male, pivoted.female)
+# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
+# extract all heart related conditions for combining
+#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
+# Aggregate the values by gender and age group
+#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
+# Rename columns and add condition column back
+#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
+#ageUK_heart$Condition <- 'Heart Problems'
+# Now drop the heart problems from original ageUK and replace with aggregate
+ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
+#ageUK <- rbind(ageUK, ageUK_heart)
+# Rename a few just for completeness
+#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
+#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
+# Finally, add column to specify that this is ageUK data
+ageUK$source <- "ageUK"
+# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
+#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
+ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
+ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
+# Make copy
+newFEM <- pivoted.FEM
+# Rename all the conditions to fit ageUK
+# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
+#       Therefore only validation from this will be that FEM is higher than ageUK
+#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
+newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
+newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
+newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
+#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
+#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
+newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
+newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+newFEM$total <- newFEM$value
+newFEM <- select(newFEM, -value)
+# Add column to specify this is FEM data
+newFEM$source <- "FEM"
+# Add column to specify these are all totals (no confidence intervals)
+newFEM$lower <- NA
+newFEM$upper <- NA
+# Now row bind the datasets
+combined <- rbind(newFEM, ageUK)
+# Specify factor levels for age group
+combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
+# Sort
+combined <- arrange(combined, source, Condition, Gender, Age_Group)
+# Function to filter the combined dataset and visualise
+visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK') {
 
+  if(gender == 'M') {
+    full_gender <- 'Male'
+  }
+  else if(gender == 'F') {
+    full_gender <- 'Female'
+  }
+
+  # filter by gender and condition
+  data <- data %>%
+            filter(Gender == gender) %>%
+            filter(Condition == cd)
+
+  # now plot!
+  p <- ggplot(data = data, aes(x = Age_Group, y = total)) +
+          geom_bar(aes(fill = source), stat = "identity", position = position_dodge2()) +
+          geom_errorbar(aes(ymin = lower, ymax = upper), position = position_dodge2(width = 1, padding = 0.5)) +
+          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
+
+  p <- p + scale_fill_grey(start = .2, end = .8)
+
+  # Save plot into in E-FEM_Paper/figures/
+  ggsave(filename = paste0(cd, '_', gender, '.png'),
+        plot = p,
+        path = path,
+        width = 7,
+        height = 5)
+
+  return(p)
+}
 ```
@@
 # Collect all conditions (and both genders) into a list, and apply the function above to that list
-# cd_list <- unique(combined$Condition)
-# gender_list <- unique(combined$Gender)
-# 
-# for(cd in cd_list) {
-#   for(gen in gender_list) {
-#     plot <- visAgeUKValid(combined, cd, gen)
-#     print(plot)
-#   }
-# }
+cd_list <- unique(combined$Condition)
+gender_list <- unique(combined$Gender)
+for(cd in cd_list) {
+  for(gen in gender_list) {
+    plot <- visAgeUKValid(combined, cd, gen)
+    print(plot)
+  } 
+}
 ```
@@
 FEM_base <- read_dta('output/COMPLETE/ELSA_core_base/ELSA_core_base_summary.dta')
-
 # Smoking prevalence stats
@@
 smok.60p <- 10.2
-
 # FEM summary, filter for 2018 and 60p
@@
                 select(contains('p_'))
-
 FEM.smok <- FEM.smok * 100
-
 smok.combined <- data.frame(FEM = FEM.smok[1,1], ASH = 10.2)
 smok.combined <- data.frame(source = c('FEM', 'FEM', 'ASH', 'ASH'), age = c('50 - 59', '60+', '50 - 59', '60+'), value = c(FEM.smok[1,2], FEM.smok[1,1], 15.2, 10.2))
-
+group.colours <- c(FEM = '#00bfc4', ASH = '#f8766d')
 smok.plot <-ggplot(data = smok.combined, aes(x = age, y = value)) +
                 geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
+                scale_fill_manual(values=group.colours) +
                 labs(x = ' ', y = 'Prevalence (%)')
+
+smok.plot <- smok.plot + scale_fill_grey(start = .2, end = .8)
 
@@
        height = 5)
-
 smok.plot
-
 ```
@@
 h_rep <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_by_rep.dta')
-
 compareDeathAgeConfints <- function(b, i, int_name=FALSE, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
@@
 }
-
 compareDeathAgeConfints(b_rep, h_rep, int_name = 'No Heart Disease', save=TRUE)
-
 ```
@@
 ### Split by Activity Level
-```{r}
-b <- b_rep
-h <- h_rep
-
-d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
-d.exstat1 <- d.exstat1 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
-d.exstat1.pivot <- pivot_longer(d.exstat1,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-p1
-
-d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
-d.exstat2 <- d.exstat2 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
-d.exstat2.pivot <- pivot_longer(d.exstat2,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-      scale_fill_grey()
-p2
-
-d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
-d.exstat3 <- d.exstat3 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
-d.exstat3.pivot <- pivot_longer(d.exstat3,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-p3
-
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'
-
-ggsave(filename = 'hearte_dAge_exstat1.png',
-         plot = p1,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_exstat2.png',
-         plot = p2,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_exstat3.png',
-         plot = p3,
-         path = path)
-
+```{r echo=FALSE}
+# b_rep <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_by_rep.dta')
+# h_rep <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_by_rep.dta')
+# 
+# b <- b_rep
+# h <- h_rep
+# 
+# d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
+# d.exstat1 <- d.exstat1 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
+# d.exstat1.pivot <- pivot_longer(d.exstat1,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
+# p1
+# 
+# d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
+# d.exstat2 <- d.exstat2 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
+# d.exstat2.pivot <- pivot_longer(d.exstat2,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
+#       scale_fill_grey()
+# p2
+# 
+# d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
+# d.exstat3 <- d.exstat3 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
+# d.exstat3.pivot <- pivot_longer(d.exstat3,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
+# p3
+# 
+# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'
+# 
+# ggsave(filename = 'hearte_dAge_exstat1.png',
+#          plot = p1,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_exstat2.png',
+#          plot = p2,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_exstat3.png',
+#          plot = p3,
+#          path = path)
 ```
@@
 ### SplitbyEduc
-```{r}
-b <- b_rep
-h <- h_rep
-
-d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
-d.educ1 <- d.educ1 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
-d.educ1.pivot <- pivot_longer(d.educ1,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-p1
-
-d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
-d.educ2 <- d.educ2 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
-d.educ2.pivot <- pivot_longer(d.educ2,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-      scale_fill_grey()
-p2
-
-d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
-d.educ3 <- d.educ3 %>%
-                dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
-d.educ3.pivot <- pivot_longer(d.educ3,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Life_Expectancy')
-p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-p3
-
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
-
-ggsave(filename = 'hearte_dAge_educ1.png',
-         plot = p1,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_educ2.png',
-         plot = p2,
-         path = path)
-
-ggsave(filename = 'hearte_dAge_educ3.png',
-         plot = p3,
-         path = path)
-
+```{r echo=FALSE}
+# b <- b_rep
+# h <- h_rep
+# 
+# d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
+# d.educ1 <- d.educ1 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
+# d.educ1.pivot <- pivot_longer(d.educ1,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
+# p1
+# 
+# d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
+# d.educ2 <- d.educ2 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
+# d.educ2.pivot <- pivot_longer(d.educ2,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
+#       scale_fill_grey()
+# p2
+# 
+# d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
+# d.educ3 <- d.educ3 %>%
+#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
+# d.educ3.pivot <- pivot_longer(d.educ3,
+#                               cols = Baseline:Intervention,
+#                               names_to = 'Source',
+#                               values_to = 'Life_Expectancy')
+# p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
+#       geom_smooth(lwd = 0.5) +
+#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
+# p3
+# 
+# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
+# 
+# ggsave(filename = 'hearte_dAge_educ1.png',
+#          plot = p1,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_educ2.png',
+#          plot = p2,
+#          path = path)
+# 
+# ggsave(filename = 'hearte_dAge_educ3.png',
+#          plot = p3,
+#          path = path)
 ```
@@
 ```{r echo=FALSE}
-# compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
-#                                  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
-#   
-#   # Calculate age and survival curves
-#   cohort$Age <- (cohort$year - 2012) + 50
-#   intervention$Age <- (intervention$year - 2012) + 50
-#   cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
-#   intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
-#   
-#   survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)
-#   
-#   # Had a small problem when trying to rename using the variable name given as function argument
-#   # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
-#   survivals <- survivals %>%
-#                  dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
-#                  melt(id.var = 'Age', value.name = 'Survival')
-#   
-#   p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
-#         geom_line() +
-#         labs(x = 'Age', y = 'Survival')
-#   
-#   if(save) {
-#     ggsave(filename = paste0(int_name, '_survival_curve.png'),
-#          plot = p1,
-#          path = path)
-#   }
-#   
-#   return(p1)
-# }
+compareSurvivalCurve <- function(cohort, intervention, int_name, save=FALSE,
+                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
+
+  # Calculate age and survival curves
+  cohort$Age <- (cohort$year - 2012) + 50
+  intervention$Age <- (intervention$year - 2012) + 50
+  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
+  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
+
+  survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)
+
+  # Had a small problem when trying to rename using the variable name given as function argument
+  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
+  survivals <- survivals %>%
+                 dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
+                 melt(id.var = 'Age', value.name = 'Survival')
+
+  p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
+        geom_line() +
+        labs(x = 'Age', y = 'Survival')
+
+  if(save) {
+    ggsave(filename = paste0(int_name, '_survival_curve.png'),
+         plot = p1,
+         path = path)
+  }
+
+  return(p1)
+}
 ```
@@
 ```{r echo=FALSE}
-# c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')
-# s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
-# 
-# compareSurvivalCurve(c, s, 'No Smoking', save=FALSE)
+c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')
+s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
+
+compareSurvivalCurve(c, s, 'No Smoking', save=FALSE)
 ```
@@
 }
-
-c_rep <- read_dta('output/COMPLETE/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
+c_rep <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
 s_rep <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')
-
 compareSurvivalCurveConfints(c_rep, s_rep, 'No Smoking', save=TRUE)
@@
 s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
-
 # Calculate age and survival curves
@@
 s$Survival <- s$m_endpop_all / s$m_endpop_all[1]
-
 d.all <- data.frame(c$Age, c$p_lunge_all, s$p_lunge_all)
@@
 p1
-
 path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'
-
 #ggsave(filename = 'smok_lunge_all.png',
@@
 s <- s_rep
-
 # Calculate age and survival curves
@@
 s$Survival <- s$m_endpop_all / s$m_endpop_all[1]
-
 d.all <- data.frame(c$Age, c$rep, c$p_lunge_all, s$p_lunge_all)
@@
 p1
-
 path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'
-
 ggsave(filename = 'smok_lunge_all.png',
@@
 aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot_test.dta')
-
 all2 <- aData %>%
@@
                     base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))
-
 all2
@@
                     base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))
-
 educ2
 ```
+
+# Comparing Output Values
+```{r}
+b <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
+h <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
+new.df <- data.frame(b$year, b$d_age_all, h$d_age_all)
+new.df$diff <- new.df$h.d_age_all - new.df$b.d_age_all
+new.df
+```

### 2022-03-01 FEM_R/impute_alcbase.Rmd

@@
 write_dta(final.imputed, path = 'input_data/alcbase_imputed.dta')
+
+rm(final.imputed, imp.list, imp.complete)
 ```
+
+## Annual based
+
+```{r}
+
+miss.list <- list()
+# generate a separate df for each wave
+imp.list <- list()
+for (x in 4:9) {
+  scak <- paste0('r', x, 'scako')
+  alcb <- paste0('r', x, 'alcbase_a')
+
+  newDF <- test %>% select(all_of(id.var), all_of(demog.vars),
+                                  paste0('r', x, 'agey'),
+                                  all_of(scak), 
+                                  all_of(alcb))
+
+  # Extract records where scako != 1:7 (i.e. missing or not drank in last 12 months)
+  scako.missings <- newDF[!newDF[[scak]] %in% 1:7,]
+
+  # Add to a list so we can rejoin later
+  miss.list[[x]] <- scako.missings
+
+  # Now select only the non-missing drinkers
+  impDF <- newDF[newDF[[scak]] %in% 1:7,]
+
+  #impDF[[alcb]][impDF[[alcb]] == 0.0,] <- NA
+  impDF[[alcb]][impDF[[alcb]] == 0] <- NA
+
+  imp <- mice(data = impDF, m=5, maxit=10, printFlag = FALSE)
+  imp.complete <- rbind(complete(imp), miss.list[[x]])
+
+  # Make sure anyone who said they haven't drank in last 12 months has alcbase == 0
+  imp.complete[[alcb]][imp.complete[[scak]] == 8] <- 0
+  # and finally drop anything that isn't wave specific for the merge in the next chunk
+  imp.complete <- imp.complete %>% select(-ragender, -raeducl)
+
+  # save to the list
+  imp.list[[x]] <- imp.complete
+}
+
+```
+
+```{r}
+
+final.imputed <- imp.list %>% discard(is.null) %>%
+                reduce(inner_join, by = "idauniq")
+
+# drop age and scako too? Yes for first attempt
+final.imputed <- final.imputed %>% select(-contains('scako'), -contains('agey'))
+
+write_dta(final.imputed, path = 'input_data/alcbase_imputed2.dta')
+
+```

### 2022-03-01 FEM_R/model_analysis_core.Rmd

@@
 
+  p <- p + scale_fill_grey(start = .2, end = .7)
   # Save plot into in E-FEM_Paper/figures/

### 2022-05-30 FEM_R/BMI_kdensity_plots.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)

```

```{r}
ELSA <- read_dta('input_data/ELSA_long.dta')
raw.ELSA <- read_dta('input_data/H_ELSA_g2_wv_specific.dta')
```

```{r}
bmi <- ELSA %>% select(wave, male, bmi)

bmi.m <- bmi %>% filter(male == 1)
bmi.f <- bmi %>% filter(male == 0)

bmi.m.2 <- bmi.m %>% filter(wave == 2) %>%
                        select(bmi) %>%
                        drop_na()
bmi.f.2 <- bmi.f %>% filter(wave == 2) %>%
                        select(bmi) %>%
                        drop_na()
```

```{r}
m.dens <- density(bmi.m.2$bmi)
f.dens <- density(bmi.f.2$bmi)
```

```{r}
plot(m.dens, col='blue', lwd=2)
lines(f.dens, col='red', lwd=2)
```

### 2022-05-30 FEM_R/FEM_ALL_PLOTS_AND_ANALYSES.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

--- 2022-03-01 FEM_R/IJM_PAPER1_ALL_ANALYSES.Rmd
@@
 ---
-title: "R Notebook"
+title: "FEM_All_plots_and_analyses"
 output: html_notebook
@@
 
-# Setup
+This notebook holds all the plots and analyses that have been undertaken using the FEM, including many different avenues of validation and results.
 
+# SETUP
 ## Libraries
@@
 require(reshape2)
-theme_set(theme_bw())
+
 workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
@@
 
-# Methods
+# Summary Stats for Starting Pops
 
-## Summary Stats for Starting Populations
-
-### Function Declaration
+## Function Declaration
 ```{r}
@@
 
-#### Incidence
+## Incidence
 ```{r}
@@
 Handover plots are generated from a Stata script that can be called via a Make rule (`make handovers`).
-The core_complete set of scenarios must be run prior to this with the `55p_f_l` & `55p_m_l` subpopulations.
 
@@
   # Save plot into in E-FEM_Paper/figures/
-  ggsave(filename = paste0(cd, '_', gender, '.png'),
-        plot = p,
-        path = path,
-        width = 7,
-        height = 5)
+  #ggsave(filename = paste0(cd, '_', gender, '.png'),
+  #      plot = p,
+  #      path = path,
+  #      width = 7,
+  #      height = 5)
 
@@
 
-## Smoking
+## Smoking (ASH)
 ```{r}
@@
 ## Heart Disease Intervention
-```{r echo=FALSE}
-# compareDeathAge <- function(base, int, int_name, save=FALSE, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
-#   
-#   aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)
-#   
-#   aDeath <- aDeath %>%
-#                 dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
-#                 reshape2::melt(id.var = 'Year', value.name = 'Age')
-#   
-#   p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
-#         geom_smooth(lwd = 0.5, se=FALSE) +
-#         labs(x = 'Year', y = 'Age at Death') +
-#   
-#   if(save) {
-#     ggsave(filename = 'hearte_dAge_all.png',
-#          plot = p1,
-#          path = path)
-#   }
-#   
-#   return(p1)
-# }
-```
 
-### Life-Expectancy
-```{r echo=FALSE}
-# b <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
-# h_p <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
-# 
-# compareDeathAge(b, h_p, 'No Heart Disease', save=FALSE)
-```
-
-#### With Confidence Intervals
+### Death Age
+With confidence intervals:
 ```{r}
@@
 compareDeathAgeConfints(b_rep, h_rep, int_name = 'No Heart Disease', save=TRUE)
-```
-
-### Split by Activity Level
-```{r echo=FALSE}
-# b_rep <- read_dta('output/SCENARIO/ELSA_core_base/ELSA_core_base_by_rep.dta')
-# h_rep <- read_dta('output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_by_rep.dta')
-# 
-# b <- b_rep
-# h <- h_rep
-# 
-# d.exstat1 <- data.frame(b$year, b$rep, b$d_age_exstat1, h$d_age_exstat1)
-# d.exstat1 <- d.exstat1 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat1, 'Intervention' = h.d_age_exstat1)
-# d.exstat1.pivot <- pivot_longer(d.exstat1,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p1 <- ggplot(data = d.exstat1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-# p1
-# 
-# d.exstat2 <- data.frame(b$year, b$rep, b$d_age_exstat2, h$d_age_exstat2)
-# d.exstat2 <- d.exstat2 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat2, 'Intervention' = h.d_age_exstat2)
-# d.exstat2.pivot <- pivot_longer(d.exstat2,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p2 <- ggplot(data = d.exstat2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-#       scale_fill_grey()
-# p2
-# 
-# d.exstat3 <- data.frame(b$year, b$rep, b$d_age_exstat3, h$d_age_exstat3)
-# d.exstat3 <- d.exstat3 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_exstat3, 'Intervention' = h.d_age_exstat3)
-# d.exstat3.pivot <- pivot_longer(d.exstat3,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p3 <- ggplot(data = d.exstat3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-# p3
-# 
-# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/'
-# 
-# ggsave(filename = 'hearte_dAge_exstat1.png',
-#          plot = p1,
-#          path = path)
-# 
-# ggsave(filename = 'hearte_dAge_exstat2.png',
-#          plot = p2,
-#          path = path)
-# 
-# ggsave(filename = 'hearte_dAge_exstat3.png',
-#          plot = p3,
-#          path = path)
-```
-
-### SplitbyEduc
-```{r echo=FALSE}
-# b <- b_rep
-# h <- h_rep
-# 
-# d.educ1 <- data.frame(b$year, b$rep, b$d_age_educ1, h$d_age_educ1)
-# d.educ1 <- d.educ1 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ1, 'Intervention' = h.d_age_educ1)
-# d.educ1.pivot <- pivot_longer(d.educ1,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p1 <- ggplot(data = d.educ1.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'Low', x = 'Year', y = 'Life Expectancy')
-# p1
-# 
-# d.educ2 <- data.frame(b$year, b$rep, b$d_age_educ2, h$d_age_educ2)
-# d.educ2 <- d.educ2 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ2, 'Intervention' = h.d_age_educ2)
-# d.educ2.pivot <- pivot_longer(d.educ2,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p2 <- ggplot(data = d.educ2.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'Moderate', x = 'Year', y = 'Life Expectancy') +
-#       scale_fill_grey()
-# p2
-# 
-# d.educ3 <- data.frame(b$year, b$rep, b$d_age_educ3, h$d_age_educ3)
-# d.educ3 <- d.educ3 %>%
-#                 dplyr::rename('Year' = b.year, 'Rep' = b.rep, 'Baseline' = b.d_age_educ3, 'Intervention' = h.d_age_educ3)
-# d.educ3.pivot <- pivot_longer(d.educ3,
-#                               cols = Baseline:Intervention,
-#                               names_to = 'Source',
-#                               values_to = 'Life_Expectancy')
-# p3 <- ggplot(data = d.educ3.pivot, aes(x = Year, y = Life_Expectancy, color = Source)) +
-#       geom_smooth(lwd = 0.5) +
-#       labs(title = 'High', x = 'Year', y = 'Life Expectancy')
-# p3
-# 
-# path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/test/'
-# 
-# ggsave(filename = 'hearte_dAge_educ1.png',
-#          plot = p1,
-#          path = path)
-# 
-# ggsave(filename = 'hearte_dAge_educ2.png',
-#          plot = p2,
-#          path = path)
-# 
-# ggsave(filename = 'hearte_dAge_educ3.png',
-#          plot = p3,
-#          path = path)
 ```
@@
 ```
-
-#### Confidence Intervals
-```{r}
-compareSurvivalCurveConfints <- function(cohort, intervention, int_name, save=FALSE,
-                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {
-
-  # Calculate age and survival curves
-  cohort$Age <- (cohort$year - 2012) + 50
-  intervention$Age <- (intervention$year - 2012) + 50
-  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
-  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]
-
-  survivals <- data.frame(cohort$Age, cohort$rep, cohort$Survival, intervention$Survival)
-
-  # Had a small problem when trying to rename using the variable name given as function argument
-  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
-  survivals <- survivals %>%
-                 dplyr::rename('Age' = cohort.Age, 'Rep' = cohort.rep, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
-                 pivot_longer(
-                   cols = Baseline:'No Smoking',
-                   names_to = 'Source',
-                   values_to = 'Life_Exp'
-                 )
-
-  p <-ggplot(data = survivals, aes(x=Age, y=Life_Exp, color = Source)) +
-        geom_smooth(lwd = 0.5) +
-        labs(x = 'Age', y = 'Survival')
-
-  if(save) {
-    ggsave(filename = paste0(int_name, '_survival_curve.png'),
-         plot = p,
-         path = path)
-  }
-
-  return(p)
-}
-c_rep <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_by_rep.dta')
-s_rep <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_by_rep.dta')
-compareSurvivalCurveConfints(c_rep, s_rep, 'No Smoking', save=TRUE)
-```
-
-### Lung Disease Prevalence All
-```{r}
-c <- read_dta('output/SCENARIO/ELSA_core_cohort/ELSA_core_cohort_summary.dta')
-s <- read_dta('output/SCENARIO/ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
-# Calculate age and survival curves
-c$Age <- (c$year - 2012) + 50
-s$Age <- (s$year - 2012) + 50
-c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
-s$Survival <- s$m_endpop_all / s$m_endpop_all[1]
-d.all <- data.frame(c$Age, c$p_lunge_all, s$p_lunge_all)
-d.all <- d.all %>%
-                dplyr::rename('Age' = c.Age, 'Baseline' = c.p_lunge_all, 'Intervention' = s.p_lunge_all)
-d.all.pivot <- pivot_longer(d.all,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Prevalence')
-p1 <- ggplot(data = d.all.pivot, aes(x = Age, y = Prevalence, color = Source)) +
-      geom_smooth(lwd = 0.5, se=FALSE) +
-      labs(x = 'Age', y = 'Prevalence')
-p1
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'
-#ggsave(filename = 'smok_lunge_all.png',
-#         plot = p1,
-#         path = path)
-```
-
-```{r}
-c <- c_rep
-s <- s_rep
-# Calculate age and survival curves
-c$Age <- (c$year - 2012) + 50
-s$Age <- (s$year - 2012) + 50
-c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
-s$Survival <- s$m_endpop_all / s$m_endpop_all[1]
-d.all <- data.frame(c$Age, c$rep, c$p_lunge_all, s$p_lunge_all)
-d.all <- d.all %>%
-                dplyr::rename('Age' = c.Age, 'Rep' = c.rep, 'Baseline' = c.p_lunge_all, 'Intervention' = s.p_lunge_all)
-d.all.pivot <- pivot_longer(d.all,
-                              cols = Baseline:Intervention,
-                              names_to = 'Source',
-                              values_to = 'Prevalence')
-p1 <- ggplot(data = d.all.pivot, aes(x = Age, y = Prevalence, color = Source)) +
-      geom_smooth(lwd = 0.5) +
-      labs(x = 'Age', y = 'Prevalence')
-p1
-path <- '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results'
-ggsave(filename = 'smok_lunge_all.png',
-         plot = p1,
-         path = path)
-```
-
-### Life Years & DFLYS
-```{r}
-aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot_test.dta')
-all2 <- aData %>%
-          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
-                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))
-all2
-```
-
-#### Split by Education
-```{r}
-educ2 <- aData %>%
-          group_by(educ) %>%
-          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
-                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))
-educ2
-```
-
-# Comparing Output Values
-```{r}
-b <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_base/ELSA_core_base_summary.dta')
-h <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/output/SCENARIO/ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
-new.df <- data.frame(b$year, b$d_age_all, h$d_age_all)
-new.df$diff <- new.df$h.d_age_all - new.df$b.d_age_all
-new.df
-```

### 2022-05-30 FEM_R/OLD_STUFF/IJM_paper1_vis_WIP.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

--- 2020-11-08 FEM_R/IJM_paper1_vis_WIP.Rmd
@@
 
+ggplot(data=base, aes(x=year, y=p_anyadl_all)) +
+  geom_line()
+ggplot(data=pop_rbmi, aes(x=year, y=p_anyadl_all)) +
+  geom_line()
+
 ```
+
+<!--### Baseline Projections by Subpopulation
+
+```{r echo = FALSE}
+visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {
+
+  if(cd == 'cancre') {
+    dis <- 'Cancer'
+  } else if(cd == 'diabe') {
+    dis <- 'Diabetes'
+  } else if(cd == 'hearte') {
+    dis <- 'Heart Disease'
+  }
+
+  # Generate varnames
+  all <- paste0('p_', cd, '_all')
+  # age
+  a5564 <- paste0('p_', cd, '_5564')
+  a6574 <- paste0('p_', cd, '_6574')
+  a7584 <- paste0('p_', cd, '_7584')
+  a85p <- paste0('p_', cd, '_85p')
+
+  ageMelt <- base %>%
+              select(c('year', a5564, a6574, a7584, a85p)) %>%
+              dplyr::rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
+              reshape2::melt(id.var = 'year', value.name = 'Prevalence')
+
+  ggplot(data=ageMelt, aes(x=year, y=Prevalence, color = variable)) +
+          geom_line() +
+          labs(title = dis, x = 'Year', y = 'Prevalence')
+}
+
+baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
+
+visProjectedBySubpop(baseline, cd = 'cancre')
+visProjectedBySubpop(baseline, cd = 'diabe')
+visProjectedBySubpop(baseline, cd = 'hearte')
+```
+
+-->
 

### 2022-05-30 FEM_R/OLD_STUFF/PH_Sensitivity_vis.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

--- 2020-07-21 FEM_R/PH_Sensitivity_vis.Rmd
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Average BMI by Age', y='Average BMI', x='Age')
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Prevalence of Diabetes by Age', y='Prevalence of Diabetes', x='Age')
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Prevalence of Hypertension by Age', y='Prevalence of Hypertension', x='Age')
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Prevalence of Heart Disease by Age', y='Prevalence of Heart Disease', x='Age')
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
+        geom_line(data=ec10, size=0.3, col='green') +
+        geom_line(data=ec50, size=0.3, col='blue') +
+        geom_line(data=ec80, size=0.3, col='red') +
         labs(title='Prevalence of Stroke by Age', y='Prevalence of Stroke', x='Age')
@@
 
-pop <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
-ep10 <- read_dta('output/ExstatP10/ExstatC10_summary.dta')
-ep50 <- read_dta('output/ExstatP50/ExstatC50_summary.dta')
-ep80 <- read_dta('output/ExstatP80/ExstatC80_summary.dta')
+pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
+ep10 <- read_dta('output/ExstatP10/ExstatP10_summary.dta')
+ep50 <- read_dta('output/ExstatP50/ExstatP50_summary.dta')
+ep80 <- read_dta('output/ExstatP80/ExstatP80_summary.dta')
 
 # Now plot all survival curves together
-ggplot(pop, aes(x=year, y=survival)) +
+ggplot(pop, aes(x=year, y=d_age_all)) +
         geom_line(col='black', size=0.3) +
@@
         geom_line(data=ep80, size=0.3, col='red') +
-        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
+        labs(title='Average Age at Death Over Time', y='Average Age at Death', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Average BMI Over Time', y='Average BMI', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Average BMI Over Time', y='Average BMI', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Prevalence of Diabetes Over Time', y='Prevalence of Diabetes', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Prevalence of Diabetes Over Time', y='Prevalence of Diabetes', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Prevalence of Hypertension Over Time', y='Prevalence of Hypertension', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Prevalence of Hypertension Over Time', y='Prevalence of Hypertension', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Prevalence of Heart Disease Over Time', y='Prevalence of Heart Disease', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Prevalence of Heart Disease Over Time', y='Prevalence of Heart Disease', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Prevalence of Cancer Over Time', y='Prevalence of Cancer', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Prevalence of Cancer Over Time', y='Prevalence of Cancer', x='Year')
 
@@
         geom_line(col='black', size=0.3) +
-        geom_line(data=dc10, size=0.3, col='green') +
-        geom_line(data=dc50, size=0.3, col='blue') +
-        geom_line(data=dc80, size=0.3, col='red') +
-        labs(title='Prevalence of Stroke Over Time', y='Prevalence of Stroke', x='Age')
+        geom_line(data=ep10, size=0.3, col='green') +
+        geom_line(data=ep50, size=0.3, col='blue') +
+        geom_line(data=ep80, size=0.3, col='red') +
+        labs(title='Prevalence of Stroke Over Time', y='Prevalence of Stroke', x='Year')
 

### 2022-05-30 FEM_R/OLD_STUFF/summary_statistics_forPaper.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

--- 2021-05-28 FEM_R/summary_statistics_forPaper.Rmd
@@
 #bmi.cut <- cut(stock$bmi, breaks = c(20, 25, 30))
+
+rm(stock.sum)
 ```
@@
 weighted.mean(stock$male, stock$weight)
-
+print('-------')
 weighted.mean(stock$cancre, stock$weight)
@@
 weighted.mean(stock$hibpe, stock$weight)
+weighted.mean(stock$alzhe, stock$weight)
+weighted.mean(stock$demene, stock$weight)
+weighted.mean(stock$hchole, stock$weight)
 
@@
   print(svymean(~hibpe, design, na.rm=TRUE))
+  print(svymean(~alzhe, design, na.rm=TRUE))
+  print(svymean(~demene, design, na.rm=TRUE))
+  print(svymean(~hchole, design, na.rm=TRUE))
 }
@@
 getIncidenceRate(trans, 'hibpe')
+getIncidenceRate(trans, 'alzhe')
+getIncidenceRate(trans, 'demene')
+getIncidenceRate(trans, 'hchole')
 ```

### 2022-05-30 FEM_CPP/AlcoholLockdown.cpp

@@ -1,0 +1,216 @@
+#include "AlcoholLockdown.h"
+#include "Logger.h"
+#include "Random.h"
+#include "fem_exception.h"
+#include <sstream>
+#include <math.h>
+#include <random>
+
+
+
+
+
+
+AlcoholLockdown::AlcoholLockdown(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp) :
+        Intervention(intervention_id, tp, vp)
+{
+    params_map["al_start_yr"] = "2020";
+}
+
+AlcoholLockdown::~AlcoholLockdown(void)
+{
+}
+
+
+
+void AlcoholLockdown::setScenario(Scenario* scen) {
+    Intervention::setScenario(scen);
+
+    std::string param_name = "al_start_yr";
+    if(atof(params_map[param_name].c_str()) == 0.0 && params_map[param_name] != "0") {
+        // User specified a time series name to use for the factor
+        // We arent setup to do this yet, so throw exception
+        throw fem_exception("Alcohol lockdown intervention needs numbers for parameters (text given for al_start_yr)!");
+    } else {
+        // User specified a number to use as a constant value
+        start_yr = atoi(params_map[param_name].c_str());
+    }
+}
+
+
+
+void AlcoholLockdown::intervene(PersonVector& persons, unsigned int year, Random* random)
+{
+    std::ostringstream ss;
+    ss << "Running Intervention: " << name() << std::endl;
+    ss << "\t" << "start year = " << start_yr << std::endl;
+    Logger::log(ss.str(), FINE);
+    ss.str("");
+    if(year <= start_yr)
+        return;
+    std::vector<Person*>::iterator itr;
+    for(itr = persons.begin(); itr != persons.end(); ++itr) {
+        Person* person = *itr;
+        if(person->test(Vars::active) && !person->test(Vars::l2died)) {
+            //Are they eligible? (yes everyone is at this stage)
+            if(elig(person)) {
+
+                // Take a random subsection of 50% of people to change their consumption habits (20% down, 30% up)
+                bool draw = random->uniformDist(person->getID(), this->intervention_id, year) > 0.5;
+
+                if(draw) {
+                    // First set treated
+                    person->set(Vars::alc_ldown_treated, true);
+
+                    // TODO: Find the paper/gov link where the 20% decrease and 30% increase are quoted
+                    // First attempt at this intervention will run as follows (based on values from somewhere, find these):
[…]

### 2022-05-30 FEM_CPP/Vars.h

@@ -672,6 +672,7 @@ namespace Vars
         osteoe,                     ///< Osteoporosis ever
         l2osteoe,                   ///< Lag of Osteoporosis ever
         posteoe,                    ///< Probability of Osteoporosis ever
+        alc_ldown_treated,          ///< Recieved the alcohol lockdown consumption change intervention
 		NVars,                    ///< A counter for the number of valid variables
 		_NONE                     ///< A special value for no variable, used in mappings where there is no valid second
 		};

### 2022-05-30 FEM_R/chron_disease_vis.r

@@ -1,145 +1,0 @@
-getwd()
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
-setwd(workingDir)
-
-require(haven)
-require(dplyr)
-
-baseline <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
-
-cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
-
-
-
-###
-commit3 <- read_dta('output/COMMIT_cSmoken3/COMMIT_cSmoken3_summary.dta')
-
-commit30 <- read_dta('output/COMMIT_cSmoken30/COMMIT_cSmoken30_summary.dta')
-
-
-## Visualise the 3% intervention
-plot(cohort$year, cohort$n_smoken_all, type='l', col='red')
-lines(commit3$year, commit3$n_smoken_all, col='green')
-
-plot(commit3$year, commit3$p_smoken_all, type='l', col='green')
-lines(cohort$year, cohort$p_smoken_all, col='red')
-
-plot(commit3$year, commit3$n_lunge_all, type='l', col='green')
-lines(cohort$year, cohort$n_lunge_all, col='red')
-
-plot(commit3$year, commit3$p_lunge_all, type='l', col='green')
-lines(cohort$year, cohort$p_lunge_all, col='red')
-
-plot(commit3$year, commit3$p_cancre_all, type='l', col='green')
-lines(cohort$year, cohort$p_cancre_all, col='red')
-
-plot(commit3$year, commit3$p_diabe_all, type='l', col='green')
-lines(cohort$year, cohort$p_diabe_all, col='red')
-
-
-## Visualise the 30% intervention
-plot(cohort$year, cohort$n_smoken_all, type='l', col='red')
-lines(commit30$year, commit30$n_smoken_all, col='green')
-
-plot(commit30$year, commit30$p_smoken_all, type='l', col='green')
-lines(cohort$year, cohort$p_smoken_all, col='red')
-
-plot(commit30$year, commit30$n_lunge_all, type='l', col='green')
-lines(cohort$year, cohort$n_lunge_all, col='red')
-
-plot(commit30$year, commit30$p_cancre_all, type='l', col='green')
-lines(cohort$year, cohort$p_cancre_all, col='red')
-
-plot(commit30$year, commit30$p_diabe_all, type='l', col='green')
[…]

### 2022-05-30 FEM_R/utilities.R

@@ -1,0 +1,113 @@
+
+# Function to calculate AUDIT-C scores and the accompanying AUDIT-C group in
+# Understanding Society data. This function is specific to the COVID-19 dataset.
+calc_audit_C19 <- function(indresp_data, wave_letter) {
+
+  # generate the variable namesf (all follow naming convention 'c{wave_letter}_audit#')
+  abst <- paste0('c', wave_letter, '_auditc1_cv')
+  q1 <- paste0('c', wave_letter, '_auditc3_cv')
+  q2 <- paste0('c', wave_letter, '_auditc4')
+  q3 <- paste0('c', wave_letter, '_auditc5_cv')
+
+  dat <- indresp_data %>% select(pidp, all_of(abst), all_of(q1), all_of(q2), all_of(q3))
+
+  # 2. Generate new audit_score var
+  dat$audit_score <- 0
+
+  # 2.5. Check for nullness in audit vars and set score to 0 if all vars are present
+  dat$audit_score[dat[[q1]] < 0] <- NA
+  dat$audit_score[dat[[q2]] < 0] <- NA
+  dat$audit_score[dat[[q3]] < 0] <- NA
+
+  # 3. Populate new variable
+  # Abstainers
+  dat$audit_score[dat[[abst]] == 2] <- 0
+  # Q1
+  dat$audit_score[which(dat[[q1]] == 1)] <- 0
+  dat$audit_score[which(dat[[q1]] == 2)] <- 1
+  dat$audit_score[which(dat[[q1]] == 3)] <- 2
+  dat$audit_score[which(dat[[q1]] == 4)] <- 3
+  dat$audit_score[which(dat[[q1]] == 5)] <- 4
+  dat$audit_score[which(dat[[q1]] == 6)] <- 4
+  # Q2
+  dat$audit_score[which(dat[[q2]] == 1)] <- dat$audit_score[which(dat[[q2]] == 1)] + 0
+  dat$audit_score[which(dat[[q2]] == 2)] <- dat$audit_score[which(dat[[q2]] == 2)] + 1
+  dat$audit_score[which(dat[[q2]] == 3)] <- dat$audit_score[which(dat[[q2]] == 3)] + 2
+  dat$audit_score[which(dat[[q2]] == 4)] <- dat$audit_score[which(dat[[q2]] == 4)] + 3
[…]

### 2022-05-30 FEM_R/vis_funcs_WIP.r

@@ -1,82 +1,0 @@
-# This script is for development of functions for visualising FEM outputs together
-
-# Set working directory
-getwd()
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
-setwd(workingDir)
-
-# Include dependency packages
-require(haven)
-require(dplyr)
-
-# Show 4 graphs in one image
-par(mfrow=c(2,2))
-
-# Function 1
-# This is for cohort outputs
-cohortVis <- function(cohortOutputList) {
-  
-  # First read in the baseline cohort output summary
-  cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
-  
-  # Convert output scenario names in cohortOutputList to full paths to then read in the data
-  pathList <- list()
-  i <- 1
-  for (element in cohortOutputList) {
-    pathList[[i]] <- paste0('output/', element, '/', element, '_summary.dta')
-    i <- i + 1
-  }
-  # Now read in datafiles to a list
-  dataList <- lapply(pathList, read_dta)
-  
-  # Check everything is working with the input
-  #lapply(dataList, names)
-  
-  # Now need to calculate survival rates 
-  
-}
-
-
-commitList <- list('COMMIT_cSmoken3', 'COMMIT_cSmoken30')
-
-cohortVis(commitList)
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-pathList <- list()
-
-i <- 1
-for (element in commitList) {
-  pathList[[i]] <- paste0('output/', element, '/', element, '_summary.dta')
-  i <- i + 1
-}
-
-ticker <- 1
-for (item in commitList) {
-  path <- paste0('output/', item, '/', item, '_summary.dta')
-  print(path)
-  #varname <- paste0(item, ticker)
-  paste0(item, ticker) <- read_dta(path)
-  ticker <- ticker + 1
-}
-
-

### 2022-07-12 FEM_R/LYs_and_DFLYs.Rmd
Committing lots of changes over a long period of time. Lots of work towards the lockdown alcohol and minimum unit price interventions. Should have been better at committing this stuff but theres been a lot of switching between different ideas

---
title: "LifeYears and Disease/Disability Free LifeYears"
output: html_notebook
---

# SETUP
## Libraries
```{r "setup", include=FALSE}
require(tidyverse)
require(haven)
require(ggplot2)

workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

# LYs and DFLYs

We are going to calculate Life Years and Disability Free Life Years for the full population outputs, then we can compare the baseline case to the Minimum Unit Price intervention. We need detailed output for this, and there is a script in stata that has done this for a cohort output that I think we can learn from. Some of the key points I know about this so far:

- We need to create a dummy variable for each person to determine whether they are alive at each wave. Summing this variable per person will be the LIFEYEARS
- Following this, we can create a dummy var for each person to determine whether they have a chronic disease and another for disability. Summing these and calculating Life Years minus x will give us disease and disability free Life Years

First get hold of the appended detailed output data.
```{r}
path <- 'input_data/detailed_output/'

cohort <- read_dta(paste0(path, 'ELSA_cohort_append.dta'))
mup <- read_dta(paste0(path, 'ELSA_alcInt_cohort_append.dta'))
```

```{r}

## Think it through - Steps:
# 1. Add a variable for lifeyears
  # Var needs to == 1 if not died and 0 if died
  # With this var we can calculate the total number of lifeyears for each person
# 2. Calculate lifeyears per person per repetition
  # Sum lifeyears per person
  # Divide sum lifeyears by number of reps

coh <- cohort

#base2 <- base %>% group_by(hhidpn, mcrep) %>%
#                  mutate(smoken_init = ())

# generate lifeyear for people who have not died
coh$lifeyear <- 1 - coh$died
# generate the number of repetitions for each person
coh$nreps <- max(coh$mcrep) + 1
# now generate the lifeyear intermediate (lifeyears per person)
coh <- coh %>% 
  group_by(hhidpn, year) %>%
  mutate(nreps = n()) %>%
  ungroup() %>%
  group_by(hhidpn) %>%
  mutate(n_LY = sum(lifeyear),
         mean_LY = n_LY / nreps,
         n_DFLY = sum(nodisease),
         mean_DFLY = n_DFLY / nreps,
         two_yr_LY = mean_LY * 2,
         two_yr_DFLY = mean_DFLY * 2,
         weight_intermed = sum(weight),
         weight2 = weight_intermed / nreps)

mean <- coh %>% summarise(LY = mean(mean_LY),
                          DFLY = mean(mean_DFLY))

```

```{r}

```

### 2022-07-12 FEM_R/dev_output_vis_and_test_alcohol.Rmd
Committing lots of changes over a long period of time. Lots of work towards the lockdown alcohol and minimum unit price interventions. Should have been better at committing this stuff but theres been a lot of switching between different ideas

---
title: "R Notebook"
output: html_notebook
---

# SETUP
## Libraries
```{r "setup", include=FALSE}
require(tidyverse)
require(haven)
require(ggplot2)
require(Hmisc)
require(reshape2)

workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

## Load in data
```{r}
base_path <- 'output/ALCOHOL/ELSA_full/'
lockdown_path <- 'output/ALCOHOL/ELSA_lockdown/'

ldown <- read_dta(paste0(lockdown_path, 'ELSA_lockdown_summary.dta'))

base <- read_dta(paste0(base_path, 'ELSA_full_summary.dta'))

mup <- read_dta('output/ALCOHOL/ELSA_alcInt_full/ELSA_alcInt_full_summary.dta')
#mup2 <- read_dta('output/ALCOHOL/ELSA_alcInt2_full/ELSA_alcInt2_full_summary.dta')
```

Now write a generic function to visualise a single continuous variable over time. Lets try with average unit consumption.

# Vis single continous

```{r}
vis_continuous_output <- function(data, var, name, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  # select column we want and rename for the visualisation
  selected <- data %>% select(year, all_of(var)) %>%
                        filter(year < end.yr) %>%
                        rename(!!quo_name(name) := var)
  # plot
  ggplot(selected, aes_string(x = 'year', y = name)) +
    geom_line()
}

vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption')
vis_continuous_output(ldown, 'a_alcbase_alc_ldown_treated', 'alcohol_consumption')
```

```{r}
vis_continuous_output(base, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption', horizon = 20)
vis_continuous_output(mup, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
```

```{r}
vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
vis_continuous_output(ldown, 'a_alcbase_alc_ldown_treated', 'alcohol_consumption', horizon = 50)
```

Now lets do a visualisation of an ordinal variable, in this case alcstat or alcstat4.

```{r}

vis_ordinal_output <- function(data, key.var.list, suffix, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  # add prefix and suffix
  key.var.list <- lapply(1:length(key.var.list), function(k) paste0('p_', key.var.list[[k]], suffix))
  # unlist and get the vars for pivoting
  first.var <- unlist(key.var.list[1])
  last.var <- unlist(key.var.list[length(key.var.list)])

  # select vars and pivot
  selected <- data %>% select(year, contains(all_of(unlist(key.var.list)))) %>%
                        filter(year < end.yr) %>%
                        pivot_longer(cols = first.var:last.var,
                                     names_to = 'Variable',
                                     values_to = 'Prevalence')
  # visualise
  ggplot(selected, aes(x = year, y = Prevalence, color = Variable)) +
    geom_line() +
    labs(title = 'Single Scenario Ordinal', x = 'Year')
}

```

```{r}
vis_ordinal_output(base, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all', horizon = 50)
#vis_ordinal_output(ldown, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all', horizon = 50)
vis_ordinal_output(mup, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all')
#vis_ordinal_output(mup2, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all')
```

```{r}

vis_ordinal_out_compare <- function(base, scen2, prefix, var, suffix, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  key.var <- paste0(prefix, var, suffix)

  base.select <- base %>% select(year, all_of(key.var)) %>%
                            rename(Baseline = key.var) %>%
                            filter(year < end.yr)
  scen2.select <- scen2 %>% select(year, all_of(key.var)) %>%
                            rename(Intervention = key.var) %>%
                            filter(year < end.yr)

  merged <- merge(base.select, scen2.select, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Baseline:Intervention,
                          names_to = 'Scenario',
                          values_to = 'Proportion')

  ggplot(pivoted, aes(x = year, y = Proportion, color = Scenario)) +
    geom_line() +
    labs(title = var, x = 'Year')

}

vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'abstainer', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'moderate', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'increasingRisk', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'highRisk', '_all', horizon = 20)

```

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'abstainer', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'moderate', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'increasingRisk', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'highRisk', '_all', horizon = 50)
```

```{r}
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'abstainer', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'moderate', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'increasingRisk', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'highRisk', '_all', horizon = 50)
```

NOW LOOK AT SOME HEALTH OUTCOMES IN THE MUP CASE.

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'hearte', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'diabe', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'alzhe', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'demene', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'd_', 'age', '_highRisk', horizon = 50)
```

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'hearte', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'diabe', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'alzhe', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'demene', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'd_', 'age', '_increasingRisk', horizon = 50)
```

# Internal Validation of alcbase and alcstat

```{r}

long <- read_dta('input_data/ELSA_long.dta')
cv1 <- read_dta('output/ALCOHOL/ELSA_CV1/ELSA_CV1_summary.dta')

```

```{r}

# 
# 
# var <- 'abstainer'
# #var <- enquo(var)
# 
# # , abstainer, moderate, increasingRisk, highRisk
# 
# long.alcstat.sum <- long %>% select(year, .data[[var]]) %>%
#                       drop_na() %>%
#                       group_by(year, .data[[var]]) %>%
#                       summarise(n = n())
# 
# long.alcstat.sum2 <- long.alcstat.sum %>%
#                       mutate(prop = n/sum(n) * 100) %>%
#                       filter(.data[[var]] == 1) %>%
#                       select(-.data[[var]], -n) %>%
#                       rename(Raw = prop)
# 
# cv1.alcstat.sum <- cv1 %>% select(year, p_abstainer_all) %>%
#                             rename(Simulated = p_abstainer_all) %>%
#                             mutate(Simulated = Simulated * 100)
# 
# merged <- merge(long.alcstat.sum2, cv1.alcstat.sum, by = 'year')
# 
# pivoted <- pivot_longer(data = merged,
#                         cols = Raw:Simulated,
#                         names_to = 'Data',
#                         values_to = 'Percentage')
# 
# ggplot(pivoted, aes(x = year, y = Percentage, color = Data)) +
#   geom_line()

compare_cv_binary <- function(long, cv1, var) {

  #var <- enquo(var)

  long.sum <- long %>% select(year, .data[[var]]) %>%
                      drop_na() %>%
                      group_by(year, .data[[var]]) %>%
                      summarise(n = n()) %>%
                      mutate(prop = n/sum(n) * 100) %>%
                      filter(.data[[var]] == 1) %>%
                      select(-.data[[var]], -n) %>%
                      rename(Raw = prop)

  p.var <- paste0('p_', var, '_all')

  cv1.sum <- cv1 %>% select(year, p.var) %>%
                              rename(Simulated = p.var) %>%
                              mutate(Simulated = Simulated * 100)
  merged <- merge(long.sum, cv1.sum, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Raw:Simulated,
                          names_to = 'Data',
                          values_to = 'Percentage')
  ggplot(pivoted, aes(x = year, y = Percentage, color = Data)) +
    geom_line() +
    labs(title = var, x = 'Year')

  #ggsave(filename = paste(var, '_crossval.png'),
  #       path = 'Deliverables/Conferences/NordicDemographic/cv')

}

compare_cv_binary(long, cv1, var = 'abstainer')
compare_cv_binary(long, cv1, var = 'moderate')
compare_cv_binary(long, cv1, var = 'increasingRisk')
compare_cv_binary(long, cv1, var = 'highRisk')

```

```{r}
long.alcbase.sum <- long %>% select(year, alcbase) %>%
                      drop_na() %>%
                      group_by(year) %>%
                      summarise(avg_alcbase = mean(alcbase))

cv1.alcbase.sum <- cv1 %>% select(year, a_alcbase_all)

compare_cv_continuous <- function(long, cv1, var) {

  long.sum <- long %>% select(year, .data[[var]]) %>%
                      drop_na() %>%
                      group_by(year) %>%
                      summarise(avg_alcbase = mean(alcbase)) %>%
                      rename(Raw = avg_alcbase)

  a.var <- paste0('a_', var, '_all')

  cv1.sum <- cv1 %>% select(year, a.var) %>%
                              rename(Simulated = a.var)

  merged <- merge(long.sum, cv1.sum, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Raw:Simulated,
                          names_to = 'Data',
                          values_to = 'Consumption')
  ggplot(pivoted, aes(x = year, y = Consumption, color = Data)) +
    geom_line() +
    labs(title = var, x = 'Year')

  ggsave(filename = paste(var, '_crossval.png'),
         path = 'Deliverables/Conferences/NordicDemographic/cv')

}

#compare_cv_continuous(long, cv1, 'alcbase')
```

## Remove files
```{r}
rm(cv1, cv1.alcbase.sum, ldown, long, long.alcbase.sum)
```

# LYs and DFLYs

We are going to calculate Life Years and Disability Free Life Years for the full population outputs, then we can compare the baseline case to the Minimum Unit Price intervention. We need detailed output for this, and there is a script in stata that has done this for a cohort output that I think we can learn from. Some of the key points I know about this so far:

- We need to create a dummy variable for each person to determine whether they are alive at each wave. Summing this variable per person will be the LIFEYEARS
- Following this, we can create a dummy var for each person to determine whether they have a chronic disease and another for disability. Summing these and calculating Life Years minus x will give us disease and disability free Life Years

First get hold of the appended detailed output data.
```{r}
# path <- 'input_data/detailed_output/'
# 
# base <- read_dta(paste0(path, 'ELSA_cohort_append.dta'))
# mup <- read_dta(paste0(path, 'ELSA_alcInt_cohort_append.dta'))
```

```{r}

## Think it through - Steps:
# 1. Add a variable for lifeyears
  # Var needs to == 1 if not died and 0 if died
  # With this var we can calculate the total number of lifeyears for each person
# 2. Calculate lifeyears per person per repetition
  # Sum lifeyears per person
  # Divide sum lifeyears by number of reps

# generate lifeyear for people who have not died
# base$lifeyear <- 1 - base$died
# # generate the number of repetitions for each person
# base$nreps <- max(base$mcrep) + 1
# # now generate the lifeyear intermediate (lifeyears per person)
# base2 <- base %>% group_by(hhidpn, year) %>%
#                     mutate(nreps = n()) %>%
#                     ungroup() %>%
#                     group_by(hhidpn) %>%
#                     mutate(n_LY = sum(lifeyear),
#                            mean_LY = n_LY / nreps) %>%
#                     mutate(n_DFLY = sum(not_disabled),
#                            mean_DFLY = n_DFLY / nreps) %>%
#                     mutate(two_yr_LY = mean_LY * 2,
#                            two_yr_DFLY = mean_DFLY * 2)
#   
# 
# mean <- base2 %>% summarise(LY = mean(mean_LY),
#                             DFLY = mean(mean_DFLY))

```

```{r}
source('../visualisation_utilities.r')
```

# MultipleContinousOut

```{r}
vis_multi_continuous_out_compare <- function(data1, data2, prefix, var, suffix.list, horizon = 50, save = FALSE) {

  # calculate horizon
  start.yr <- 2012
  end.yr <- start.yr + horizon

  # add prefix and suffix
  key.var.list <- lapply(1:length(suffix.list), function(k) paste0(prefix, '_', var, '_', suffix.list[[k]]))

  data1.select <- data1 %>% select(year, all_of(unlist(key.var.list))) %>%
                            rename(Baseline_Moderate = key.var.list[[1]],
                                   Baseline_Increasing = key.var.list[[2]],
                                   Baseline_High = key.var.list[[3]]) %>%
                          filter(year < end.yr) %>%
                          filter(year > start.yr + 2)
  data2.select <- data2 %>% select(year, all_of(unlist(key.var.list))) %>%
                          rename(Intervention_Moderate = key.var.list[[1]],
                                 Intervention_Increasing = key.var.list[[2]],
                                 Intervention_High = key.var.list[[3]]) %>%
                          filter(year < end.yr) %>%
                          filter(year > start.yr + 2)

  merged <- merge(data1.select, data2.select, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Baseline_Moderate:Intervention_High,
                          names_sep = '_',
                          names_to = c('Scenario', 'Subgroup'),
                          values_to = 'Proportion')

  #ggplot(pivoted, aes(x = year, colour = Scenario)) +
  #  geom_line(aes(y = Proportion))

  p <- ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
        geom_line() +
        scale_linetype_manual(values = c("TRUE" = "solid", "FALSE" = "dashed")) +
        guides(linetype = FALSE)

  if(save) {
    ggsave(filename = 'alcbase_comparison_multisuffix.png',
           path = 'Deliverables/Conferences/NordicDemographic/ordinal_comparisons')
  }

  p
}
```

```{r}
data1 <- base
data2 <- mup
prefix <- 'a'
var <- 'alcbase'
suffix.list <- c('moderate', 'increasingRisk', 'highRisk')
horizon <- 50

vis_multi_continuous_out_compare(data1, data2, prefix, var, suffix.list, horizon = 30)
```

```{r}
data1 <- base
data2 <- mup
prefix <- 'a'
var <- 'alcbase'
suffix.list <- c('moderate', 'increasingRisk', 'highRisk')
horizon <- 50

# calculate horizon
start.yr <- 2012
end.yr <- start.yr + horizon

# add prefix and suffix
key.var.list <- lapply(1:length(suffix.list), function(k) paste0(prefix, '_', var, '_', suffix.list[[k]]))

data1.select <- data1 %>% select(year, all_of(unlist(key.var.list))) %>%
                          rename(Baseline_Moderate = key.var.list[[1]],
                                 Baseline_Increasing = key.var.list[[2]],
                                 Baseline_High = key.var.list[[3]]) %>%
                        filter(year < end.yr) %>%
                        filter(year > start.yr + 2)
data2.select <- data2 %>% select(year, all_of(unlist(key.var.list))) %>%
                        rename(Intervention_Moderate = key.var.list[[1]],
                               Intervention_Increasing = key.var.list[[2]],
                               Intervention_High = key.var.list[[3]]) %>%
                        filter(year < end.yr) %>%
                        filter(year > start.yr + 2)

merged <- merge(data1.select, data2.select, by = 'year')

pivoted <- pivot_longer(data = merged,
                        cols = Baseline_Moderate:Intervention_High,
                        names_sep = '_',
                        names_to = c('Scenario', 'Subgroup'),
                        values_to = 'Proportion')

#ggplot(pivoted, aes(x = year, colour = Scenario)) +
#  geom_line(aes(y = Proportion))

ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
  geom_line() +
  scale_linetype_manual(name = 'Scenario', values = c("TRUE" = "solid", "FALSE" = "dashed"), guide = 'legend', labels = c('Baseline', 'Intervention')) +
  guides(linetype = FALSE)

ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
  geom_line() +
  scale_linetype_manual(name = 'Scenario', values = c("TRUE" = "solid", "FALSE" = "dashed"), guide = 'legend', labels = c('Intervention', 'Baseline'))

#ggplot(pivoted, aes(x = year, y = Proportion, group =))
```

# Reduction

Now producing graphs to show reduction in disease prevalence over time because the prevalence graphs are a bit weird.

```{r}

base1 <- base
mup1 <- mup

base2 <- base1 %>% select(year, p_hearte_all, p_hearte_moderate, p_hearte_increasingRisk, p_hearte_highRisk) %>%
                    rename(Baseline_All = p_hearte_all, 
                           Baseline_Moderate = p_hearte_moderate, 
                           Baseline_IncreasingRisk = p_hearte_increasingRisk,
                           Baseline_HighRisk = p_hearte_highRisk)

mup2 <- mup1 %>% select(year, p_hearte_all, p_hearte_moderate, p_hearte_increasingRisk, p_hearte_highRisk) %>%
                    rename(Intervention_All = p_hearte_all, 
                           Intervention_Moderate = p_hearte_moderate, 
                           Intervention_IncreasingRisk = p_hearte_increasingRisk,
                           Intervention_HighRisk = p_hearte_highRisk)

merged <- merge(base2, mup2, by = 'year')

differenced <- merged %>% mutate(Moderate = Baseline_Moderate - Intervention_Moderate,
                                 IncreasingRisk = Baseline_IncreasingRisk - Intervention_IncreasingRisk,
                                 HighRisk = Baseline_HighRisk - Intervention_HighRisk)

transformed <- differenced %>% group_by(year) %>%
                                summarise(mean_Mod = mean(Moderate),
                                          min_Mod = min(Moderate),
                                          max_Mod = max(Moderate),
                                          mean_Inc = mean(IncreasingRisk),
                                          min_Inc = min(IncreasingRisk),
                                          max_Inc = max(IncreasingRisk),
                                          mean_High = mean(HighRisk),
                                          min_High = min(HighRisk),
                                          max_High = max(HighRisk),)

pivoted <- pivot_longer(data = transformed,
                        cols = mean_Mod:max_High,
                        names_sep = '_',
                        names_to = c('Stat', 'Subgroup'),
                        values_to = 'Prevalence')

ggplot(pivoted, aes(x = year, y = Prevalence, group = interaction(Stat, Subgroup), colour = Subgroup)) +
  geom_line()

```

```{r}

```

### 2022-07-12 FEM_R/dfly_detailed_output.Rmd
Committing lots of changes over a long period of time. Lots of work towards the lockdown alcohol and minimum unit price interventions. Should have been better at committing this stuff but theres been a lot of switching between different ideas

@@
 # Need to set up the environment
-workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
+workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
 knitr::opts_knit$set(root.dir = workingDir)
@@
 require(haven)
-require(reshape2)
+#require(reshape2)
 require(ggplot2)
-require(ggpubr)
+#require(ggpubr)
 #require(plyr)
@@
 ```{r}
-cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_append.dta')
+#cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_append.dta')
 #heart <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_append.dta')
-smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_append.dta')
+#smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_append.dta')
+
+path <- 'input_data/detailed_output/'
+
+cohort <- read_dta(paste0(path, 'ELSA_cohort_append.dta'))
+mup <- read_dta(paste0(path, 'ELSA_alcInt_cohort_append.dta'))
 ```
@@
           mutate(hearte_init = min(year[hearte == 1]), hearte_died = max(year[hearte == 1])) %>%
-          mutate(lunge_init = min(year[lunge == 1]), lunge_died = max(year[lunge == 1]))
+          mutate(diabe_init = min(year[diabe == 1]), diabe_died = max(year[diabe == 1]))
 
 # Replace all infinite values with 0 (infinites generated in mutate calls above)
-new$hearte_init[is.infinite(new$hearte_init)] <- 0
-new$hearte_died[is.infinite(new$hearte_died)] <- 0
-new$lunge_init[is.infinite(new$lunge_init)] <- 0
-new$lunge_died[is.infinite(new$lunge_died)] <- 0
+new.c$hearte_init[is.infinite(new.c$hearte_init)] <- 0
+new.c$hearte_died[is.infinite(new.c$hearte_died)] <- 0
+new.c$diabe_init[is.infinite(new.c$diabe_init)] <- 0
+new.c$diabe_died[is.infinite(new.c$diabe_died)] <- 0
 

### 2022-07-12 FEM_CPP/AlcoholIntervention50p2.cpp

@@ -1,4 +1,4 @@
-#include "AlcoholIntervention50p.h"
+#include "AlcoholIntervention50p2.h"
 #include "Logger.h"
 #include "fem_exception.h"
 #include <sstream>
@@ -10,19 +10,19 @@
 
 
 
-AlcoholIntervention50p::AlcoholIntervention50p(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp) :
+AlcoholIntervention50p2::AlcoholIntervention50p2(unsigned int intervention_id, ITimeSeriesProvider* tp, IVariableProvider* vp) :
         Intervention(intervention_id, tp, vp)
 {
     params_map["ai50p_start_yr"] = "2012";
 }
 
-AlcoholIntervention50p::~AlcoholIntervention50p(void)
+AlcoholIntervention50p2::~AlcoholIntervention50p2(void)
 {
 }
 
 
 
-void AlcoholIntervention50p::setScenario(Scenario* scen) {
+void AlcoholIntervention50p2::setScenario(Scenario* scen) {
     Intervention::setScenario(scen);
 
     std::string param_name = "ai50p_start_yr";
@@ -38,7 +38,7 @@ void AlcoholIntervention50p::setScenario(Scenario* scen) {
 
 
 
-void AlcoholIntervention50p::intervene(PersonVector& persons, unsigned int year, Random* random)
+void AlcoholIntervention50p2::intervene(PersonVector& persons, unsigned int year, Random* random)
 {
     std::ostringstream ss;
     ss << "Running Intervention: " << name() << std::endl;
@@ -82,23 +82,21 @@ void AlcoholIntervention50p::intervene(PersonVector& persons, unsigned int year,
                     person->set(Vars::alcbase, person->get(Vars::alcbase) - 4.2);
                 }*/
 
-                person->set(Vars::mup_treated, 1);
-
                 // Changing to a percentage reduction in consumption
                 // Moderate (1.5% reduction)
                 if(person->test(Vars::moderate)) {
-                    person->set(Vars::alcbase, person->get(Vars::alcbase) * 0.985);
-                    //person->set(Vars::alcbase, person->get(Vars::alcbase) * 0.5);
+                    //person->set(Vars::alcbase, person->get(Vars::alcbase) * 0.985);
+                    person->set(Vars::alcbase, person->get(Vars::alcbase) - 0.1);
                 }
                 // Increasing Risk (3.9% reduction)
                 if(person->test(Vars::increasingRisk)) {
[…]

### 2022-07-12 FEM_CPP/HealthModule.cpp

@@ -356,8 +356,8 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
 			            // set alcstat4 to 1 as this is for abstainers
 			            person->set(Vars::alcstat4, 1);
 
-                        // Now also set all the dummys to zero (previously was abstainer but thats been removed)
-			            //person->set(Vars::abstainer, 1);
+                        // Now also set all the dummys to zero
+			            person->set(Vars::abstainer, 1);
                         person->set(Vars::moderate, 0);
                         person->set(Vars::increasingRisk, 0);
                         person->set(Vars::highRisk, 0);
@@ -378,7 +378,7 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
 			            //person->set(Vars::alcstat4, 2);
 			            // Also set dummy
 			            //person->set(Vars::moderate, 1);
-                        //person->set(Vars::abstainer, 0);
+                        person->set(Vars::abstainer, 0);
                         //person->set(Vars::increasingRisk, 0);
                         //person->set(Vars::highRisk, 0);
 			        }
@@ -431,8 +431,17 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
                     person->set(Vars::highRisk, 0);
                     // set alcbase to equal the correct var
                     person->set(Vars::alcbase, person->get(Vars::alcbase_inc));
-                    person->set(Vars::alcbase_mod, 0);
+                    //person->set(Vars::alcbase_mod, 0);
                     person->set(Vars::alcbase_high, 0);
+                    // fill the alcbase_mod for male and female differently
+                    if(person->test(Vars::male)) {
+                        // if male
+                        person->set(Vars::alcbase_mod, 21);
+                    }
+                    if(!person->test(Vars::male)) {
+                        // if female
+                        person->set(Vars::alcbase_mod, 15);
+                    }
                 }
                 // high risk
                 if(person->get(Vars::alcstat) == 3) {
@@ -445,7 +454,120 @@ void HealthModule::process(PersonVector& persons, unsigned int year, Random* ran
                     person->set(Vars::alcbase, person->get(Vars::alcbase_high));
                     person->set(Vars::alcbase_inc, 0);
                     person->set(Vars::alcbase_mod, 0);
+                    // fill the alcbase_mod and alcbase_inc for male and female differently
+                    if(person->test(Vars::male)) {
+                        // if male
[…]

### 2022-07-12 FEM_CPP/Vars.cpp

@@ -735,7 +735,21 @@ VarsInfo::VarsInfo() {
     info[Vars::osteoe]              = VarInfo("osteoe", "Osteoporosis ever", VarTypes::Boolean);
     info[Vars::l2osteoe]            = VarInfo("l2osteoe", "Lag of Osteoporosis ever", VarTypes::Boolean);
     info[Vars::posteoe]             = VarInfo("posteoe", "Probability of Osteoporosis ever", VarTypes::Boolean);
-    info[Vars::alc_ldown_treated]   = VarInfo("alc_ldown_treated", "Recieved the alcohol consumption change intervention", VarTypes::Boolean);
+    info[Vars::alc_ldown_treated]   = VarInfo("alc_ldown_treated", "Received the alcohol consumption change intervention", VarTypes::Boolean);
+    info[Vars::mup_treated]         = VarInfo("mup_treated", "Received the minimum unit price of alcohol intervention", VarTypes::Boolean);
+    info[Vars::abstainerTime]       = VarInfo("abstainerTime", "Number waves spent in abstainer alcohol consumption group", VarTypes::Short);
+    info[Vars::moderateTime]        = VarInfo("moderateTime", "Number waves spent in moderate alcohol consumption group", VarTypes::Short);
+    info[Vars::increasingTime]      = VarInfo("increasingTime", "Number waves spent in increasingRisk alcohol consumption group", VarTypes::Short);
+    info[Vars::highTime]            = VarInfo("highTime", "Number waves spent in highRisk alcohol consumption group", VarTypes::Short);
+    info[Vars::abstainerProp]       = VarInfo("abstainerProp", "Proportion of time spent in abstainer consumption group", VarTypes::Double);
+    info[Vars::moderateProp]        = VarInfo("moderateProp", "Proportion of time spent in moderate consumption group", VarTypes::Double);
+    info[Vars::increasingProp]      = VarInfo("increasingProp", "Proportion of time spent in increasingRisk consumption group", VarTypes::Double);
+    info[Vars::highProp]            = VarInfo("highProp", "Proportion of time spent in highRisk consumption group", VarTypes::Double);
+    info[Vars::abstainerMC]         = VarInfo("abstainerMC", "Most common alcohol consumption group: Abstainer", VarTypes::Boolean);
+    info[Vars::moderateMC]          = VarInfo("moderateMC", "Most common alcohol consumption group: Moderate", VarTypes::Boolean);
[…]

### 2022-07-12 FEM_CPP/Vars.h

@@ -672,8 +672,22 @@ namespace Vars
         osteoe,                     ///< Osteoporosis ever
         l2osteoe,                   ///< Lag of Osteoporosis ever
         posteoe,                    ///< Probability of Osteoporosis ever
-        alc_ldown_treated,          ///< Recieved the alcohol lockdown consumption change intervention
-		NVars,                    ///< A counter for the number of valid variables
+        alc_ldown_treated,          ///< Received the alcohol lockdown consumption change intervention
+        mup_treated,                ///< Received the minimum unit pricing intervention treatment
+        abstainerTime,              ///< Number waves spent in abstainer alcohol consumption group
+        moderateTime,              ///< Number waves spent in moderate alcohol consumption group
+        increasingTime,            ///< Number waves spent in increasingRisk alcohol consumption group
+        highTime,                  ///< Number waves spent in highRisk alcohol consumption group
+        abstainerProp,             ///< Proportion of time spent in abstainer alcohol consumption group
+        moderateProp,              ///< Proportion of time spent in moderate alcohol consumption group
+        increasingProp,            ///< Proportion of time spent in increasingRisk alcohol consumption group
+        highProp,                  ///< Proportion of time spent in highRisk alcohol consumption group
+        abstainerMC,                ///< Most common alcohol consumption group: Abstainer
+        moderateMC,                ///< Most common alcohol consumption group: Moderate
+        increasingMC,                ///< Most common alcohol consumption group: IncreasingRisk
+        highMC,                ///< Most common alcohol consumption group: HighRisk
+        elsa_waves,             ///< Number of waves present in ELSA data
+		NVars,                     ///< A counter for the number of valid variables
 		_NONE                     ///< A special value for no variable, used in mappings where there is no valid second
 		};
 

### 2022-12-05 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added to the utilities script, and created new R Notebook for the loneliness and social isolation work

---
title: "Social_Isolation_and_Loneliness"
output: html_notebook
---

In this notebook, we prepare all the figures and tables for the paper 'Projecting the impact of loneliness and social isolation on the health outcomes of England's future elderly population'. We want to investigate how loneliness and social isolation will affect health outcomes, both together and separately, when we project an elderly population into the future using the English Future Elderly model.

# Setup

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(reactable)
require(haven)
require(ggplot2)
require(jtools)

```

## Data

```{r}
# Directories
base_output_dir <- "output/COMPLETE/"
input_dir <- "input_data/"

source('FEM_R/utilities.R')

# Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))

# Outputs
base_full <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort_full <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
```

# Sample Statistics

In this section, we will calculate a lot of sample statistics on the stock, replenishing, and transition populations, and generate a table that can be easily copied into latex. We will also investigate the distribution of both the loneliness and social isolation measures we have added to the FEM, and particularly look to see how both loneliness and social isolation is distributed among different demographic groups. I am also interested in the individual elements of the index of social isolation, because I want to investigate some of the reasons why social isolation may be distributed differently to loneliness.

## Sample Statistic Calculations

We want sample statistics on a number of characteristics of our input populations:
- Demographics
    - Age
    - BMI
    - Gender
    - Education
- Married/cohabiting %
- Loneliness score
- Social Isolation Index
- Risk Behaviours
    - Smoking
    - Alcohol frequency
    - Physical Activity
- Mental conditions
    - Depression
    - More?
- Chronic Diseases
    - Any
    - Hypertension
    - CVD
    - Diabetes
    - Cancer
    - Alzheimers/Dementia

Samples must be weighted by survey cross-sectional weight - `cwtresp`

```{r, echo=FALSE}
### Going to update the previous function from utilities.R here
# copy the data so we don't overwrite anything
s <- stock

# try it first to see what we get
#weighted.survey.means(s)

weighted.survey.means <- function(init.pop, transition=FALSE) {

  require(survey)

  if(!transition) {
    design <- svydesign(id = ~idauniq,
                        weights = ~weight,
                        data = init.pop)
  } else if(transition) {
    # Error from missing cwtresp var - drop all that are missing to solve
    init.pop <- init.pop[complete.cases(init.pop$cwtresp),]
    design <- svydesign(id = ~idauniq,
                        weights = ~cwtresp,
                        data = init.pop)
  }
  print(svymean(~age, design, na.rm=TRUE))
  print(svyvar(~age, design, na.rm=TRUE))
  print(svymean(~male, design, na.rm=TRUE))
  print(svymean(~bmi, design, na.rm=TRUE))
  print(svyvar(~bmi, design, na.rm=TRUE))
  print(svymean(~smoken, design, na.rm=TRUE))
  print(svymean(~smokev, design, na.rm=TRUE))
  print(svymean(~hsless, design, na.rm=TRUE))
  print(svymean(~college, design, na.rm=TRUE))
  print('---------------------')
  print(svymean(~cancre, design, na.rm=TRUE))
  print(svymean(~diabe, design, na.rm=TRUE))
  print(svymean(~hearte, design, na.rm=TRUE))
  print(svymean(~stroke, design, na.rm=TRUE))
  print(svymean(~lunge, design, na.rm=TRUE))
  print(svymean(~hibpe, design, na.rm=TRUE))
  print(svymean(~alzhe, design, na.rm=TRUE))
  print(svymean(~demene, design, na.rm=TRUE))
  print(svymean(~hchole, design, na.rm=TRUE))
}

s <- s %>% filter(!is.na(s$strat))

weighted.survey.means(s)

```

```{r}
## Lets try to make a table
require(survey)

# First create survey design object for weighted stats
design <- svydesign(id = ~idauniq,
                    weights = ~weight,
                    data = s)

# Now calculate means (and SE)
# Age + Sex + BMI + Smoking Status + Education MEANS
means <- svymean(~age+male+bmi+smoken+smokev+hsless+college, design, na.rm=TRUE)
mean.df <- as.data.frame(means, optional = TRUE)
mean.df <- tibble::rownames_to_column(mean.df, 'Variable')

# Need counts of non missing data per variable to calculate SD from SE
counts <- s %>% summarise_each(funs(sum(!is.na(.)))) %>%
  select(age, male, bmi, smoken, smokev, hsless, college) %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column('Variable') %>%
  rename(n = 'V1')

# Lovely, now merge
mean.df <- merge(counts, mean.df, by = 'Variable')

# Now we can calculate SD
mean.df$SD <- mean.df$SE * sqrt(mean.df$n)

print(mean.df)
```

## Distribution of Lnly
```{r}

```

## Distribution of SocIso

## Latex outputs

# Results

## Functions

## Baseline

### Life Years

### Disability-free Life Years

### Disease-free Life Years

## Interventions

### 2022-12-06 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added some functions for calculating Life Years and DFLYs from summary output. Ran these comparisons on baseline and intervention cases, not the outcome we wanted...

@@
 ```{r}
-# Directories
-base_output_dir <- "output/COMPLETE/"
+## Directories
+base_output_dir <- "output/LNLY_SOCISO/"
 input_dir <- "input_data/"
@@
 
-# Inputs
+## Inputs
 long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
@@
 
-# Outputs
+## Outputs
+### Baseline
 base_full <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
 cohort_full <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
+### Interventions
+no.lonely <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_cohort/ELSA_nolnly_cohort_summary.dta'))
+no.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nosociso_cohort/ELSA_nosociso_cohort_summary.dta'))
+no.lnly.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_nosociso_cohort/ELSA_nolnly_nosociso_cohort_summary.dta'))
 ```
@@
 
+Looking at three outcomes in particular, and maybe some other supporting stuff if its interesting enough.
+
+The three outcomes are:
+- Life Years
+- Disability-free Life Years
+- Disease-free Life Years
+
 ## Functions
+
+### Cohort
+
+Can work out Life Years and DFLYs from summary output for cohort populations.
+
+We do this by multiplying the number of people alive by 2 each year to get the total number of people living (life years) per wave (2 years each wave).
+We then get the total number of life years for all people at each wave, and can therefore work out the average life years per person by dividing the total by number of people alive at each wave.
+
+Using these calculation we can make comparisons between baseline and intervention scenarios in terms of average life years.
+
+```{r "LifeYears"}
+# Calculate average life years for an entire cohort from 50
+cohortLifeyear <- function(c) {
+  c.tot.lfyrs <- c$m_endpop_all * 2
+  c.sum.tot.lfyrs <- sum(c.tot.lfyrs)
+  c.avg.lfyr <- c.sum.tot.lfyrs / c$m_endpop_all[1]
+  return(c.avg.lfyr)
+}
+
+# Compare cohort LifeYears from 50 between two summary outputs (baseline vs int)
+lifeyearComparison <- function(c, i, int.name) {
+
+  c.avg.lfyr <- cohortLifeyear(c)
+
+  i.avg.lfyr <- cohortLifeyear(i)
+
+  print(int.name)
+  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.lfyr, digits=4)))
+  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.lfyr, digits=4)))
+  print(paste('Increase:                        ', signif(x = (i.avg.lfyr - c.avg.lfyr), digits=4)))
+}
+
+# This function calculates the average number of life years in the model when diagnosed with a chronic disease
+diseaseVarLifeyear <- function(c, var) {
+
+  prevVar <- paste0('p_', var, '_all')
+
+  c.tot.var.lfyrs <- c[,prevVar] * c$m_endpop_all * 2
+  c.sum.var.lfyrs <- sum(c.tot.var.lfyrs)
+  c.avg.var.lfyrs <- c.sum.var.lfyrs / c$m_endpop_all[1]
+
+  print(paste0(var, ':    ', signif(c.avg.var.lfyrs, digits=4)))
+}
+```
+
+```{r "Disability"}
+# Calculate disability life years
+disabilityLifeYearsCohort <- function(c) {
+  total_dis <- c$a_disabled_all * c$m_endpop_all
+  sum_tot_dis <- sum(total_dis)
+  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
+  return(avg_dis)
+}
+
+# Compare disability life years between scenarios
+disabilityLifeYearComp <- function(c, i, int.name) {
+  c.avg.dis <- disabilityLifeYearsCohort(c)
+  i.avg.dis <- disabilityLifeYearsCohort(i)
+  print(int.name)
+  print(paste('Cohort Average Disabled Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Disabled Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+
+# Calculate disability free life years
+disabilityFreeLifeYearsCohort <- function(c) {
+  total_disfree <- c$a_not_disabled_all * c$m_endpop_all
+  sum_tot_disfree <- sum(total_disfree)
+  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
+  return(avg_disfree)
+}
+
+# Compare disability free life years between scenarios
+disabilityFreeLifeYearsComp <- function(c, i, int.name) {
+  c.avg.dis <- disabilityFreeLifeYearsCohort(c)
+  i.avg.dis <- disabilityFreeLifeYearsCohort(i)
+  print(int.name)
+  print(paste('Cohort Average Disability Free Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Disability Free Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+```
+
+```{r "Disease"}
+# Calculate disease life years
+diseaseLifeYearsCohort <- function(c) {
+  total_dis <- c$a_anydisease_all * c$m_endpop_all
+  sum_tot_dis <- sum(total_dis)
+  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
+  return(avg_dis)
+}
+
+# Compare disease life years between scenarios
+diseaseLifeYearComp <- function(c, i, int.name) {
+  c.avg.dis <- diseaseLifeYearsCohort(c)
+  i.avg.dis <- diseaseLifeYearsCohort(i)
+  print(int.name)
+  print(paste('Cohort Average Diseased Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Diseased Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+
+# Calculate disease free life years
+diseaseFreeLifeYearsCohort <- function(c) {
+  total_disfree <- c$a_nodisease_all * c$m_endpop_all
+  sum_tot_disfree <- sum(total_disfree)
+  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
+  return(avg_disfree)
+}
+
+# Compare disease free life years between scenarios
+diseaseFreeLifeYearsComp <- function(c, i, int.name) {
+  c.avg.dis <- diseaseFreeLifeYearsCohort(c)
+  i.avg.dis <- diseaseFreeLifeYearsCohort(i)
+  print(int.name)
+  print(paste('Cohort Average Disease Free Lifeyears:        ', signif(c.avg.dis, digits=4)))
+  print(paste('Intervention Average Disease Free Lifeyears:  ', signif(i.avg.dis, digits=4)))
+  print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+}
+```
+
+### Whole Population
 
@@
 ### Life Years
+```{r}
+coh.ly <- cohortLifeyear(cohort_full)
+
+# Total lifeyears
+print(paste('Cohort Average Lifeyears:        ', signif(coh.ly, digits=4)))
+
+# Cohort disease life years
+diseaseVarLifeyear(cohort_full, 'cancre')
+diseaseVarLifeyear(cohort_full, 'diabe')
+diseaseVarLifeyear(cohort_full, 'hearte')
+diseaseVarLifeyear(cohort_full, 'lunge')
+diseaseVarLifeyear(cohort_full, 'hibpe')
+diseaseVarLifeyear(cohort_full, 'stroke')
+diseaseVarLifeyear(cohort_full, 'alzhe')
+diseaseVarLifeyear(cohort_full, 'demene')
+```
 
@@
 ## Interventions
+
+### Burden of Loneliness
+```{r}
+coh.ly <- cohortLifeyear(cohort_full)
+nolnly.ly <- cohortLifeyear(no.lonely)
+
+lifeyearComparison(cohort_full, no.lonely, 'No Loneliness')
+lifeyearComparison(cohort_full, no.sociso, 'No Social Isolation')
+lifeyearComparison(cohort_full, no.lnly.sociso, 'No Loneliness OR Social Isolation')
+```

### 2022-12-06 FEM_R/calculate_dflys_from_summary.Rmd
Added some functions for calculating Life Years and DFLYs from summary output. Ran these comparisons on baseline and intervention cases, not the outcome we wanted...

@@
 # Need to set up the environment
-workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
+workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/COMPLETE/"
 knitr::opts_knit$set(root.dir = workingDir)
@@
 # Hearte (cohort and pop)
-heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
+#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
 #heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')
@@
 # Smoking (cohort)
-smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
+#smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')
 
@@
 #obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
-obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')
+#obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')
 
 # Drinking (pop)
-nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
-nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
+#nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
+#nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
 #noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
@@
 # Cohort life years
-c <- cohort
+c <- cohort_full
 c.ly <- cohortLifeyear(c)
@@
 ```
+
+```{r}
+
+```

### 2022-12-07 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Additions to the R scripts for analysing loneliness and social isolation outputs

@@
 We want to investigate how loneliness and social isolation will affect health outcomes, both together and separately, when we project an elderly population into the future using the English Future Elderly model.
+
+# README!!!!!!!!!
+
+Need some investigation as the outcomes are not what we expect:
+- Check deaths by subgroup per wave
+    - Make sure the underlying data shows what we expect
+    - Should also extend this to see incidence of chronic disease by subgroup
+- Check the indices have been created successfully
+    - Loneliness should be fine, but we can test by recreating the summary measure from the three vars
+    - Social isolation could be wrong. Check each member of the index for any of these weird relationships
+- Look at the sample we are providing to the transition models
+    - Is the sample correct?
+Check the paper the social isolation index is based on and see what sample they used.
+Maybe we can recreate this and try to recreate the results they got for their cox PH model
 
@@
 require(jtools)
+require(ggfortify)
+require(survival)
+require(survminer)
+require(lazyeval)
 
+```
+
+```{r}
+source('../HAZARDS/source/hazard_utils.R')
+source('FEM_R/utilities.R')
 ```
@@
 input_dir <- "input_data/"
-
-source('FEM_R/utilities.R')
 
@@
 
+# Survival Analysis (For Testing)
+
+```{r}
+prepped.long <- prepare_ELSA(elsa_long = long, 
+                         wave_start = 1, 
+                         wave_end = 9, 
+                         print_counts = TRUE)
+```
+
+## Deaths
+### Subgroup
+```{r}
+deaths_by_subgroup <- function(elsa, subgroup) {
+
+  dead.subgroup <- elsa %>% 
+    filter(died == 1) %>%
+    group_by(sociso) %>%
+    count()
+
+  return(dead.subgroup)
+}
+
+test.out <- deaths_by_subgroup(prepped.long, 'sociso')
+test.out
+
+dead.subgroup <- prepped.long %>%
+  filter(died == 1) %>%
+  group_by(lnly) %>%
+  summarise(n = n())
+
+dead.subgroup
+
+prepped.long %>% count(lnly)
+```
+
+```{r}
+# Only extract rows when someone died
+dead <- prepped.long %>% filter(died == 1)
+
+# Report total number of deaths and number per wave
+print(paste0('Number of people who died during survey: ', nrow(dead), '/', length(unique(prepped.long$idauniq))))
+table(dead$wave)
+
+dead.alcstat <- dead %>%
+  group_by(wave, lnly) %>%
+  count() %>%
+  rename(Deaths = 'n')
+# 
+# # replace NA with 'unknown'
+# dead.alcstat$alcstat6 <- as.character(dead.alcstat$alcstat6)
+# dead.alcstat[is.na(dead.alcstat)] <- 'Unknown'
+# dead.alcstat$alcstat6 <- as.factor(dead.alcstat$alcstat6)
+# 
+# dead.alcstat
+# 
+# # Now plot deaths by alcstat6 at each wave
+# ggplot(data = dead.alcstat, mapping = aes(x = wave, y = Deaths, fill = alcstat6)) +
+#   geom_bar(stat = 'identity', position = 'dodge') +
+#   labs(title = 'Deaths stratified by alcstat6 categories',
+#        x = 'Wave',
+#        y = 'Number of Deaths')
+# 
+# ## GOOD (not really big problems but the code works fine)
+# # Now calculate the proportion of each group that die each wave
+# 
+# # First calculate n by group at each wave (including the unknowns)
+# counts.subgroup <- test %>% 
+#   group_by(wave, alcstat6) %>%
+#   count() %>%
+#   rename(Total_sample = 'n')
+# 
+# # Replace NA with 'unknown'
+# counts.subgroup$alcstat6 <- as.character(counts.subgroup$alcstat6)
+# counts.subgroup[is.na(counts.subgroup)] <- 'Unknown'
+# counts.subgroup$alcstat6 <- as.factor(counts.subgroup$alcstat6)
+# 
+# # now merge total and deaths together
+# counts <- merge(counts.subgroup, dead.alcstat, by = c('wave', 'alcstat6'), all = TRUE)
+# 
+# # finally, calculate proportion of deaths per subgroup
+# counts$death_prop <- (counts$Deaths / counts$Total_sample) * 100
+# 
+# # Now plot death_prop by alcstat6 at each wave
+# ggplot(data = counts, mapping = aes(x = wave, y = death_prop, fill = alcstat6)) +
+#   geom_bar(stat = 'identity', position = 'dodge') +
+#   labs(title = 'Proportion of deaths per wave within alcstat6 categories',
+#        x = 'Wave',
+#        y = 'Percentage of Sample')
+# # one more without unknown
+# ggplot(data = filter(counts, alcstat6 != 'Unknown'), mapping = aes(x = wave, y = death_prop, fill = alcstat6)) +
+#   geom_bar(stat = 'identity', position = 'dodge') +
+#   labs(title = 'Proportion of deaths per wave within alcstat6 categories',
+#        x = 'Wave',
+#        y = 'Percentage of Sample')
+```
+
 # Results
@@
 
-## Functions
+## Life Years Functions
 
@@
 
-### Burden of Loneliness
+### Burden of Chronic Diseases
+
+#### Life Years
 ```{r}
@@
 ```
+
+#### Disability Free Life Years
+```{r}
+disabilityFreeLifeYearsComp(cohort_full, no.lonely, 'No Loneliness')
+disabilityFreeLifeYearsComp(cohort_full, no.sociso, 'No Social Isolation')
+disabilityFreeLifeYearsComp(cohort_full, no.lnly.sociso, 'No Loneliness OR Social Isolation')
+```
+
+```{r}
+diseaseFreeLifeYearsComp(cohort_full, no.lonely, 'No Loneliness')
+diseaseFreeLifeYearsComp(cohort_full, no.sociso, 'No Social Isolation')
+diseaseFreeLifeYearsComp(cohort_full, no.lnly.sociso, 'No Loneliness OR Social Isolation')
+```

### 2023-02-23 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Changed path for new PC

@@
 # Need to set up the working directory
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
+workingDir <- "/home/luke/Documents/WORK/E_FEM/E_FEM/"
 knitr::opts_knit$set(root.dir = workingDir)

### 2023-02-24 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
Added function to utilities.R and some more variable testing of physically active vars in the rough_working_script

---
title: "R Notebook"
output: html_notebook
---

ROUGH WORK CHECKING SCRIPT FOR LONELINESS AND SOCIAL ISOLATION WORK.

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/WORK/E_FEM/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(reactable)
require(haven)
require(ggplot2)
require(jtools)
require(ggfortify)
require(survival)
require(survminer)
require(lazyeval)

```

```{r}
source('../HAZARDS/source/hazard_utils.R')
source('FEM_R/utilities.R')
```

```{r}
## Directories
base_output_dir <- "output/COMPLETE/"
input_dir <- "input_data/"
## Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))
## Outputs
### Baseline
base <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
### Cross-Validation
cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
#cv1.full <- read_dta(paste0(base_output_dir, 'ELSA_CV1/detailed_output/ELSA_CV1_summary.dta'))
cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
### Interventions
#no.lonely <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_cohort/ELSA_nolnly_cohort_summary.dta'))
#no.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nosociso_cohort/ELSA_nosociso_cohort_summary.dta'))
#no.lnly.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_nosociso_cohort/ELSA_nolnly_nosociso_cohort_summary.dta'))
```

## NEW PHYSICALLY_ACTIVE

```{r}
# First long file
physact.long <- long %>% 
  select(idauniq, wave, age, male, educ, workstat, employed, inactive, retired,
         physically_active) %>%
  replace_na(list(physically_active = -9)) %>%
  group_by(wave, physically_active) %>%
  summarise(n = n(), 
            age.mean = mean(age, na.rm=TRUE),
            age.median = median(age, na.rm=TRUE),
            age.5064 = sum(age >= 50 & age < 65, na.rm=TRUE),
            age.6579 = sum(age >= 65 & age < 80, na.rm=TRUE),
            age.80p = sum(age > 79, na.rm=TRUE),
            male.prop = (sum(male, na.rm=TRUE) / n),
            female = (n - sum(male, na.rm=TRUE)),
            educ.mean = mean(educ, na.rm=TRUE),
            educ1 = sum(educ == 1, na.rm=TRUE),
            educ2 = sum(educ == 2, na.rm=TRUE),
            educ3 = sum(educ == 3, na.rm=TRUE),
            workstat.mean = mean(workstat, na.rm=TRUE),
            employed = sum(employed, na.rm=TRUE),
            inactive = sum(inactive, na.rm=TRUE),
            retired = sum(retired, na.rm=TRUE))

physact.long$wave <- as.integer(physact.long$wave)
physact.long$physically_active <- as.factor(physact.long$physically_active)
```

#### By Age

```{r}
ggplot(data = physact.long, aes(x = wave, y = n, group = physically_active, colour = physically_active)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Number') +
  xlab('Wave') +
  ylab('Count')

ggplot(data = physact.long, aes(x = wave, y = age.mean, group = physically_active, colour = physically_active)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Mean Age') +
  xlab('Wave') +
  ylab('Count')

age.groups <- physact.long %>% 
  select(wave, physically_active, age.5064, 
                                      age.6579, age.80p) %>%
  pivot_longer(cols = age.5064:age.80p,
               names_to = 'age_group',
               values_to = 'count')

ggplot(data = age.groups, aes(x = wave, y = count, group = physically_active, colour = physically_active)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  facet_grid(age_group ~ .) +
  labs(title = 'Count by Age Group') +
  xlab('Wave') +
  ylab('Count')

# ggplot(data = physact.long, aes(x = wave, y = age.mean, group = physically_active, colour = physically_active)) +
#   geom_line() +
#   scale_x_continuous(breaks=c(1:9)) +
#   labs(title = 'Mean Age') +
#   xlab('Wave') +
#   ylab('Count')
```

#### By Sex

```{r}
ggplot(data = physact.long, aes(x = wave, y = male.prop, group = physically_active, colour = physically_active)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Number') +
  xlab('Wave') +
  ylab('Count')
```

#### By Educ

### Job Physically Active

New variable being used to determine physically active var above. Just want to 
plot and see.

```{r}
# First long file
jphysl.long <- long %>% 
  select(idauniq, wave, age, male, educ, workstat, employed, inactive, retired,
         jphysl) %>%
  group_by(wave, jphysl) %>%
  summarise(n = n(), 
            age.mean = mean(age),
            age.median = median(age),
            age.5064 = sum(age >= 50 & age < 65),
            age.6579 = sum(age >= 65 & age < 80),
            age.80p = sum(age > 79),
            male = sum(male),
            female = (n - sum(male)),
            educ.mean = mean(educ),
            educ1 = sum(educ == 1),
            educ2 = sum(educ == 2),
            educ3 = sum(educ == 3),
            workstat.mean = mean(workstat),
            employed = sum(employed),
            inactive = sum(inactive),
            retired = sum(retired))

jphysl.long$wave <- as.integer(physact.long$wave)
```

```{r}
ggplot(data = jphysl.long, aes(x = wave, y = count)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Count')
```

## NEW WORKSTAT

Newly recoded var is 3 levels:

- Employed
- Inactive
- Unemployed

Model won't run, saying l2inactive is not a valid variable. First thing to check
is that the variable has been properly created and check balance.

```{r}
# First long file
workstat.long <- long %>% 
  select(idauniq, wave, workstat, l2workstat, employed, 
         inactive, retired, l2employed, l2inactive, l2retired) %>%
  group_by(wave) %>%
  summarise(employed = sum(employed),
            inactive = sum(inactive),
            retired = sum(retired)) %>%
  pivot_longer(cols = employed:retired,
               names_to = 'workstat',
               values_to = 'count')

test.long$wave <- as.integer(test.long$wave)
```

```{r}
ggplot(data = workstat.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Count')
```

#### Outputs

```{r}
# First long file
test.CV1 <- cv1 %>% 
  select(year, a_workstat_all, p_employed_all, 
         p_inactive_all, p_retired_all) %>%
  group_by(year) %>%
  summarise(employed = sum(p_employed_all),
            inactive = sum(p_inactive_all),
            retired = sum(p_retired_all)) %>%
  pivot_longer(cols = employed:retired,
               names_to = 'workstat',
               values_to = 'prevalence')

test.long$wave <- as.integer(test.long$wave)
```

```{r}
ggplot(data = test.CV1, aes(x = year, y = count, group = workstat, color=workstat)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Prevalence')
```

```{r}
# First long file
test.CV1 <- long %>% 
  select(idauniq, wave, workstat, l2workstat, employed, 
         inactive, retired, l2employed, l2inactive, l2retired) %>%
  group_by(wave) %>%
  summarise(employed = sum(employed),
            inactive = sum(inactive),
            retired = sum(retired)) %>%
  pivot_longer(cols = employed:retired,
               names_to = 'workstat',
               values_to = 'count')

test.long$wave <- as.integer(test.long$wave)
```

```{r}
ggplot(data = test.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Count')
```

### 2023-02-24 FEM_R/variable_testing/workstat.Rmd
Started a directory for variable testing in R. Added script for workstat and physically_active

--- 2023-02-24 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
@@
 ---
-title: "R Notebook"
+title: "Workstat"
 output: html_notebook
 ---
-
-ROUGH WORK CHECKING SCRIPT FOR LONELINESS AND SOCIAL ISOLATION WORK.
 
@@
 ```
-
-## NEW PHYSICALLY_ACTIVE
-
-```{r}
-# First long file
-physact.long <- long %>% 
-  select(idauniq, wave, age, male, educ, workstat, employed, inactive, retired,
-         physically_active) %>%
-  replace_na(list(physically_active = -9)) %>%
-  group_by(wave, physically_active) %>%
-  summarise(n = n(), 
-            age.mean = mean(age, na.rm=TRUE),
-            age.median = median(age, na.rm=TRUE),
-            age.5064 = sum(age >= 50 & age < 65, na.rm=TRUE),
-            age.6579 = sum(age >= 65 & age < 80, na.rm=TRUE),
-            age.80p = sum(age > 79, na.rm=TRUE),
-            male.prop = (sum(male, na.rm=TRUE) / n),
-            female = (n - sum(male, na.rm=TRUE)),
-            educ.mean = mean(educ, na.rm=TRUE),
-            educ1 = sum(educ == 1, na.rm=TRUE),
-            educ2 = sum(educ == 2, na.rm=TRUE),
-            educ3 = sum(educ == 3, na.rm=TRUE),
-            workstat.mean = mean(workstat, na.rm=TRUE),
-            employed = sum(employed, na.rm=TRUE),
-            inactive = sum(inactive, na.rm=TRUE),
-            retired = sum(retired, na.rm=TRUE))
-
-physact.long$wave <- as.integer(physact.long$wave)
-physact.long$physically_active <- as.factor(physact.long$physically_active)
-```
-
-#### By Age
-
-```{r}
-ggplot(data = physact.long, aes(x = wave, y = n, group = physically_active, colour = physically_active)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Number') +
-  xlab('Wave') +
-  ylab('Count')
-
-ggplot(data = physact.long, aes(x = wave, y = age.mean, group = physically_active, colour = physically_active)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Mean Age') +
-  xlab('Wave') +
-  ylab('Count')
-
-age.groups <- physact.long %>% 
-  select(wave, physically_active, age.5064, 
-                                      age.6579, age.80p) %>%
-  pivot_longer(cols = age.5064:age.80p,
-               names_to = 'age_group',
-               values_to = 'count')
-
-ggplot(data = age.groups, aes(x = wave, y = count, group = physically_active, colour = physically_active)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  facet_grid(age_group ~ .) +
-  labs(title = 'Count by Age Group') +
-  xlab('Wave') +
-  ylab('Count')
-
-# ggplot(data = physact.long, aes(x = wave, y = age.mean, group = physically_active, colour = physically_active)) +
-#   geom_line() +
-#   scale_x_continuous(breaks=c(1:9)) +
-#   labs(title = 'Mean Age') +
-#   xlab('Wave') +
-#   ylab('Count')
-```
-
-#### By Sex
-
-```{r}
-ggplot(data = physact.long, aes(x = wave, y = male.prop, group = physically_active, colour = physically_active)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Number') +
-  xlab('Wave') +
-  ylab('Count')
-```
-
-#### By Educ
-
-### Job Physically Active
-
-New variable being used to determine physically active var above. Just want to 
-plot and see.
-
-```{r}
-# First long file
-jphysl.long <- long %>% 
-  select(idauniq, wave, age, male, educ, workstat, employed, inactive, retired,
-         jphysl) %>%
-  group_by(wave, jphysl) %>%
-  summarise(n = n(), 
-            age.mean = mean(age),
-            age.median = median(age),
-            age.5064 = sum(age >= 50 & age < 65),
-            age.6579 = sum(age >= 65 & age < 80),
-            age.80p = sum(age > 79),
-            male = sum(male),
-            female = (n - sum(male)),
-            educ.mean = mean(educ),
-            educ1 = sum(educ == 1),
-            educ2 = sum(educ == 2),
-            educ3 = sum(educ == 3),
-            workstat.mean = mean(workstat),
-            employed = sum(employed),
-            inactive = sum(inactive),
-            retired = sum(retired))
-
-jphysl.long$wave <- as.integer(physact.long$wave)
-```
-
-```{r}
-ggplot(data = jphysl.long, aes(x = wave, y = count)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Workstat in ELSA') +
-  xlab('Wave') +
-  ylab('Count')
-```
-
-## NEW WORKSTAT
 
@@
 ```
-
-```{r}
-# First long file
-test.CV1 <- long %>% 
-  select(idauniq, wave, workstat, l2workstat, employed, 
-         inactive, retired, l2employed, l2inactive, l2retired) %>%
-  group_by(wave) %>%
-  summarise(employed = sum(employed),
-            inactive = sum(inactive),
-            retired = sum(retired)) %>%
-  pivot_longer(cols = employed:retired,
-               names_to = 'workstat',
-               values_to = 'count')
-
-test.long$wave <- as.integer(test.long$wave)
-```
-
-```{r}
-ggplot(data = test.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Workstat in ELSA') +
-  xlab('Wave') +
-  ylab('Count')
-```

### 2023-02-24 FEM_R/model_analysis_core.Rmd

@@
 # Need to set up the working directory
-workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/COMPLETE/"
+workingDir <- "/home/luke/Documents/WORK/E_FEM/E_FEM/output/COMPLETE/"
 knitr::opts_knit$set(root.dir = workingDir)

### 2023-02-24 FEM_R/variable_testing/physically_active.Rmd

--- 2023-02-24 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
@@
 ```
-
-## NEW PHYSICALLY_ACTIVE
 
@@
 ```
-
-## NEW WORKSTAT
-
-Newly recoded var is 3 levels:
-
-- Employed
-- Inactive
-- Unemployed
-
-Model won't run, saying l2inactive is not a valid variable. First thing to check
-is that the variable has been properly created and check balance.
-
-```{r}
-# First long file
-workstat.long <- long %>% 
-  select(idauniq, wave, workstat, l2workstat, employed, 
-         inactive, retired, l2employed, l2inactive, l2retired) %>%
-  group_by(wave) %>%
-  summarise(employed = sum(employed),
-            inactive = sum(inactive),
-            retired = sum(retired)) %>%
-  pivot_longer(cols = employed:retired,
-               names_to = 'workstat',
-               values_to = 'count')
-
-test.long$wave <- as.integer(test.long$wave)
-```
-
-```{r}
-ggplot(data = workstat.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Workstat in ELSA') +
-  xlab('Wave') +
-  ylab('Count')
-```
-
-#### Outputs
-
-```{r}
-# First long file
-test.CV1 <- cv1 %>% 
-  select(year, a_workstat_all, p_employed_all, 
-         p_inactive_all, p_retired_all) %>%
-  group_by(year) %>%
-  summarise(employed = sum(p_employed_all),
-            inactive = sum(p_inactive_all),
-            retired = sum(p_retired_all)) %>%
-  pivot_longer(cols = employed:retired,
-               names_to = 'workstat',
-               values_to = 'prevalence')
-
-test.long$wave <- as.integer(test.long$wave)
-```
-
-```{r}
-ggplot(data = test.CV1, aes(x = year, y = count, group = workstat, color=workstat)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Workstat in ELSA') +
-  xlab('Wave') +
-  ylab('Prevalence')
-```
-
-```{r}
-# First long file
-test.CV1 <- long %>% 
-  select(idauniq, wave, workstat, l2workstat, employed, 
-         inactive, retired, l2employed, l2inactive, l2retired) %>%
-  group_by(wave) %>%
-  summarise(employed = sum(employed),
-            inactive = sum(inactive),
-            retired = sum(retired)) %>%
-  pivot_longer(cols = employed:retired,
-               names_to = 'workstat',
-               values_to = 'count')
-
-test.long$wave <- as.integer(test.long$wave)
-```
-
-```{r}
-ggplot(data = test.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
-  geom_line() +
-  scale_x_continuous(breaks=c(1:9)) +
-  labs(title = 'Workstat in ELSA') +
-  xlab('Wave') +
-  ylab('Count')
-```

### 2023-02-24 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
Updated the settings.csv and .txt files with the new vars for the loneliness and social isolation scenario runs. Also wrote a utilities function in R for collating detailed output files, and some more variable checking in the R working scripts. Also imputed some things in kludge.do, and need to investigate why lnly values are not being imputed properly

@@
 ## Directories
-base_output_dir <- "output/COMPLETE/"
+base_output_dir <- "output/LNLY_SOCISO_TEST/"
 input_dir <- "input_data/"
@@
 base <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
+base.detail <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
 cohort <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
 ### Cross-Validation
-cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
+#cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
 #cv1.full <- read_dta(paste0(base_output_dir, 'ELSA_CV1/detailed_output/ELSA_CV1_summary.dta'))
-cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
+#cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
 ### Interventions
@@
 #no.lnly.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_nosociso_cohort/ELSA_nolnly_nosociso_cohort_summary.dta'))
+```
+
+```{r}
+detailed_output_append <- function(base.out.dir, scenario) {
+  ## function to read in all files in a detailed output and collect into 1 object
+
+  path <- paste0('output/', base.out.dir, scenario, 'detailed_output/')
+
+  filelist <- list.files(path,
+                         full.names = TRUE)
+  return(do.call(rbind, lapply(filelist, read_dta)))
+}
+
+path <- 'output/LNLY_SOCISO_TEST/ELSA_core_base/detailed_output/'
+
+base.detail <- detailed_output_append(base.out.dir = base_output_dir, scenario = 'ELSA_core_base')
+
 ```

### 2023-02-24 FEM_R/utilities.R

@@ -37,6 +37,16 @@ lookfor <- function(data,
   else message("Nothing found. Sorry.")
 }
 
+########## DETAILED OUTPUT APPEND ##########
+
+detailed_output_append <- function(base.out.dir, scenario) {
+  ## function to read in all files in a detailed output and collect into 1 object
+  path <- paste0(base.out.dir, scenario, '/detailed_output/')
+  filelist <- list.files(path,
+                         full.names = TRUE)
+  return(do.call(rbind, lapply(filelist, read_dta)))
+}
+
 ########## SAMPLE STATISTICS ##########
 
 

### 2023-03-02 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
Fixed hotdecking not working properly for lnly, now no more missing values being treated as 0 by the model

--- 2023-02-24 FEM_R/Loneliness_Social_Isolation/rough_working_script.Rmd
@@
 base <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
+base.detail <- detailed_output_append(base.out.dir = base_output_dir, scenario = 'ELSA_core_base')
 cohort <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
 ### Cross-Validation
-cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
+#cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
 #cv1.full <- read_dta(paste0(base_output_dir, 'ELSA_CV1/detailed_output/ELSA_CV1_summary.dta'))
-cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
+#cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
 ### Interventions
@@
 ```
+
+## LONELINESS
+
+```{r}
+lnly.test <- base.detail %>%
+  select(hhidpn, year, lnly) %>%
+  group_by(year, lnly) %>%
+  count()
+
+lnly.test
+```
+
+It appears that values of 0 are being assigned to lnly at some point which should not happen! First check the stock dataset and go from there.
+
+```{r}
+lnly.stock.test <- stock %>%
+  select(idauniq, year, lnly) %>%
+  group_by(year, lnly) %>%
+  count()
+
+lnly.stock.test
+
+## Try the same in the long files
+lnly.long.test <- long %>%
+  select(idauniq, year, lnly) %>%
+  filter(year == 2012) %>%
+  group_by(year, lnly) %>%
+  count()
+
+lnly.long.test
+```
+
+Appears the lnly values of zero must come from these missing values, first check in stata as R loses the value labels when reading .dta files.
+
+Stata output looking at the long file:
+
+. keep if year == 2012
+(88,070 observations deleted)
+
+. tab lnly, m
+
+ Loneliness |
+ Score, Low |
+to High [1, |
+         3] |      Freq.     Percent        Cum.
+------------+-----------------------------------
+          1 |      5,156       44.96       44.96
+          2 |      2,309       20.14       65.10
+          3 |        366        3.19       68.29
+          . |        866        7.55       75.84
+         .c |      2,698       23.53       99.37
+         .m |         72        0.63      100.00
+------------+-----------------------------------
+      Total |     11,467      100.00
+
+Majority of missing cases here are due to not being in the self-completion sample. Makes sense. However what does not make sense is that kludge is not properly imputing the missing values.
 

### 2023-03-02 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Added a notebook with some variable and transition model testing

---
title: "FEM Transition Models"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/WORK/E_FEM/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)

```

```{r}
source('../HAZARDS/source/hazard_utils.R')
source('FEM_R/utilities.R')
```

Load in data needed to fit these models (transition pop).

```{r}
input_dir <- "input_data/"
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
```

# Mortality

```{r}
# Currently in the FEM (2/3/23) the mortality model only depends on demographics, loneliness, and social isolation
# This was to try and get the largest effect in mortality from interventions in these variables
# Lets try the minimal model first and go from there

# Mortality is a probit model with binary outcome

# make sure mortality if factor
trans$died <- factor(trans$died)

# check cross tabs with lnly and sociso
xtabs(formula = ~died + l2lnly, data = trans, addNA = TRUE)
xtabs(formula = ~died + l2sociso, data = trans, addNA = TRUE)
```

So we at least have some deaths from each group to work with. Now check the proportion of each group and see if there is a relationship between loneliness / social isolation and mortality.

```{r}
xtabs(formula = ~died + l2lnly, data = trans, addNA = TRUE)
xtabs(formula = ~died + l2sociso, data = trans, addNA = TRUE)
```

So we do have an association between the lag of loneliness and mortality.

Lets check the percentages.

```{r}
lnly <- as.data.frame(xtabs(formula = ~died + l2lnly, data = trans, addNA = TRUE))
sociso <- as.data.frame(xtabs(formula = ~died + l2sociso, data = trans, addNA = TRUE))

lnly.check <- lnly %>% 
  group_by(l2lnly) %>%
  mutate(total = sum(Freq)) %>%
  mutate(prop = (Freq / total) * 100)

sociso.check <- sociso %>% 
  group_by(l2sociso) %>%
  mutate(total = sum(Freq)) %>%
  mutate(prop = (Freq / total) * 100)
```

```{r}
lnly.check
```

Loneliness shows a pretty good relationship with mortality in the way that we expect. The lowest rate of mortality is in the loneliness == 1 group, which is low loneliness. Then we get higher for loneliness == 2, and higher still for loneliness == 3. Unfortunately the highest rate is in the missing group, where most people will be imputed or removed. Still happy that the relationship holds up.

```{r}
sociso.check
```

The previous output of this chunk wasn't very good, and prompted a closer look at the creation of the sociso index and the component variables. I found a problem in this definition (was a chain of replace commands that meant that the last replace was most important). Now the variable looks excellent in terms of mortality! We see the lowest rates of mortality in the sociso == 1 group (lowest / no social isolation), and higher rates of mortality as time goes on. Magnificent!

Time to fit some models. Start with mortality, and do the minimal (just demographics), lnly/sociso, and full version.

```{r}
form.mort.min <- 'died ~ male + white + hsless + college + l2age65l + l2age6574 + l2age75p'

probit.mort.min <- glm(formula = form.mort.min, family = binomial(link = 'probit'),
                    data = trans)

summary(probit.mort.min)
```

Now try an iteration with lnly and sociso.

```{r}
form.mort.min <- 'died ~ male + white + hsless + college + 
                  l2age65l + l2age6574 + l2age75p + 
                  l2lnly2 + l2lnly3 +
                  l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 + l2sociso6'

probit.mort.min <- glm(formula = form.mort.min, family = binomial(link = 'probit'),
                    data = trans)

summary(probit.mort.min)
```

### 2023-03-04 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Updated the settings.csv and .txt files with the new vars for the loneliness and social isolation scenario runs. Also wrote a utilities function in R for collating detailed output files, and some more variable checking in the R working scripts. Also imputed some things in kludge.do, and need to investigate why lnly values are not being imputed properly
Added a title to a section in Rmd

@@
 ## Directories
-base_output_dir <- "output/LNLY_SOCISO/"
+base_output_dir <- "output/LNLY_SOCISO_TEST/"
 input_dir <- "input_data/"
@@
 
+#### Disease Free Life Years
 ```{r}

### 2023-03-11 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Massive improvements to the loneliness model including lots of new variables, as well as changes to the reference factors for lnly and sociso which are particularly important in the mortality model. IT WORKS

@@
 require(haven)
+require(MASS)
 
@@
 
-# make sure mortality if factor
+# make sure responses are factor where necessary
 trans$died <- factor(trans$died)
+trans$lnly <- factor(trans$lnly)
+trans$sociso <- factor(trans$sociso)
+trans$male <- factor(trans$male)
+trans$hsless <- factor(trans$hsless)
+trans$college <- factor(trans$college)
+trans$l2employed <- factor(trans$l2employed)
+trans$l2inactive <- factor(trans$l2inactive)
+trans$l2physact <- factor(trans$l2physact)
 
@@
   mutate(prop = (Freq / total) * 100)
+
+#write.csv(lnly.check, )
 ```
@@
 
+Lets have one more go with these limited to waves 1-6, as no more deaths are recorded after wave 6.
+
+```{r}
+trans2 <- trans %>% filter(wave < 7)
+
+lnly2 <- as.data.frame(xtabs(formula = ~died + l2lnly, data = trans2, addNA = TRUE))
+sociso2 <- as.data.frame(xtabs(formula = ~died + l2sociso, data = trans2, addNA = TRUE))
+
+lnly.check2 <- lnly2 %>%
+  group_by(l2lnly) %>%
+  mutate(total = sum(Freq)) %>%
+  mutate(prop = (Freq / total) * 100)
+
+sociso.check2 <- sociso2 %>%
+  group_by(l2sociso) %>%
+  mutate(total = sum(Freq)) %>%
+  mutate(prop = (Freq / total) * 100)
+```
+
 Time to fit some models. Start with mortality, and do the minimal (just demographics), lnly/sociso, and full version.
@@
 ```{r}
-form.mort.min <- 'died ~ male + white + hsless + college + 
+form.mort.ls <- 'died ~ male + white + hsless + college + 
                   l2age65l + l2age6574 + l2age75p + 
-                  l2lnly2 + l2lnly3 +
-                  l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 + l2sociso6'
+                  l2lnly1 + l2lnly2 + l2lnly3 +
+                  l2sociso1 + l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 + l2sociso6'
 
-probit.mort.min <- glm(formula = form.mort.min, family = binomial(link = 'probit'),
+probit.mort.ls <- glm(formula = form.mort.ls, family = binomial(link = 'probit'),
                     data = trans)
 
-summary(probit.mort.min)
+summary(probit.mort.ls)
 ```
+
+Interesting.
+Mirrors the FEM transition model, although not exactly.
+Same sign for each var and similar relative magnitude (i.e.
+in the sociso levels).
+The sociso relationship is great now but the loneliness is not so clear cut.
+This actually mirrors the literature so we could be on to something here!
+Try a version where the reference factors are different, might be good to pick the worst of each var and then we can see where the relationship falls down.
+
+```{r}
+form.mort.ls2 <- 'died ~ male + white + hsless + college + 
+                  l2age65l + l2age6574 + l2age75p + 
+                  l2lnly1 + l2lnly2 +
+                  l2sociso1 + l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5'
+
+probit.mort.ls2 <- glm(formula = form.mort.ls2, family = binomial(link = 'probit'),
+                    data = trans)
+
+summary(probit.mort.ls2)
+```
+
+*Chefs kiss*.
+Exaclty what we want to see.
+Probability of mortality increases with higher rates of loneliness and social isolation.
+Appears that loneliness here is more highly associated with mortality than social isolation, but I have an idea about that.
+Perhaps, with loneliness being the subjective feeling and social isolation the objective situation, social isolation is more likely in people in palliative care or close to end of life already.
+These people are harder to follow in ELSA, and institutionalised people are removed from the survey.
+Maybe a contributing factor.
+
+Try one more version, with chronic diseases and stuff.
+
+```{r}
+form.mort.full <- 'died ~ male + white + hsless + college + 
+                  l2age65l + l2age6574 + l2age75p + 
+                  l2lnly1 + l2lnly2 +
+                  l2sociso1 + l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 +
+                  l2cancre + l2diabe + l2hearte + l2alzhe + l2demene + l2lunge +
+                  l2stroke'
+
+probit.mort.full <- glm(formula = form.mort.full, family = binomial(link = 'probit'),
+                    data = trans)
+
+summary(probit.mort.full)
+```
+
+# Loneliness and Social Isolation
+
+Now lets have a go at fitting models to lnly and sociso.
+
+## Loneliness
+
+```{r}
+# Get a sample comparable to the sample used in FEM
+trans.lnly <- trans %>%
+  dplyr::select(idauniq, wave, died, lnly, male, white, hsless, college, l2age65l, l2age6574, l2age75p, atotb, l2employed, l2inactive, l2physact) %>%
+  filter(wave > 1) %>%
+  filter(died == 0)
+trans.lnly.complete <- trans.lnly[complete.cases(trans.lnly),]
+
+## MINIMAL
+form.lnly.min <- 'lnly ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.lnly.min <- polr(formula = form.lnly.min,
+                    data = trans.lnly,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.lnly.min)
+
+## FULL
+form.lnly <- 'lnly ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + scale(atotb) + l2employed + l2inactive + l2physact'
+probit.lnly <- polr(formula = form.lnly,
+                    data = trans.lnly,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.lnly)
+```
+
+Now use these models as a testbed using ideas from literature.
+
+First up [Cohen-Mansfield et al (2015)](https://www.cambridge.org/core/journals/international-psychogeriatrics/article/abs/correlates-and-predictors-of-loneliness-in-olderadults-a-review-of-quantitative-results-informed-by-qualitative-insights/EF57448627F1EF5CFD89F83C3F188175).
+
+Shortlist of variables I already have or can easily test:
+- [x] gender
+- [x] age
+- [x] unmarried status
+- [x] income (wealth?)
+- [x] education
+- nursing home resident (may be issues with 0 analysis weight) (nhmliv)
+- [x] participates in social activities (socyr)
+- [x] care for someone in household (R*GCAREINHH1W)
+- [x] Childless (child)
+- [x] living alone / few individuals in household (H*HHRES)
+- [x] self reported health == poor (SRH == 1)
+- [x] functional status == poor (ADL == 3+, IADL == 2+)
+- [x] depression (psychiatric problems - emotional, nervous, or psychiatric)
+
+```{r}
+## TEST
+form.lnly.test <- 'lnly ~ male + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2cohab + l2widowed + l2single + l2employed + l2inactive + l2physact + l2anyadl + l2anyiadl + l2srh5 + l2psyche + l2hhres + l2socyr + l2gcareinhh1w + childless + scale(l2logatotb)'
+probit.lnly.test <- polr(formula = form.lnly.test,
+                    data = trans,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Test model:')
+summary(probit.lnly.test)
+```
+
+```{r}
+
+```

### 2023-03-11 FEM_CPP/Vars.h

@@ -30,8 +30,8 @@ namespace Vars
 		anyrx_meps,								///< Any prescriptions from MEPS model
 		arthre,					///< Arthritis ever, stored as Boolean
 		asthmae,				///< Asthma ever, stored as Boolean
-		atotb,                  ///< Total Family Wealth
-		itot,                    ///< Total Family Income
+		//atotb,                  ///< Total Family Wealth
+		//itot,                    ///< Total Family Income
 		births,                   ///< Count of childbirth events (0,1,2), stored as Short
 		black,                    ///< Non-Hispanic black [0, 1], stored as Boolean
 		bornus,                   ///< Born in the U.S.? [0, 1], stored as Boolean
@@ -211,8 +211,8 @@ namespace Vars
 		l2anyhi,                   ///< Lag of HI cov -gov/emp/other, [0, 1], stored as Boolean
 		l2arthre,					///< Lag of arthritis ever, stored as Boolean
 		l2asthmae,					///< Lag of asthma ever, stored as Boolean
-		l2atotb,                    ///< Lag of Total Family Wealth
-		l2itot,                      ///< Lag of Total Family Income
+		//l2atotb,                    ///< Lag of Total Family Wealth
+		//l2itot,                      ///< Lag of Total Family Income
 		l2bpcontrol,               ///< Lag of bpcontrol
 		l2bs_treated,              ///< Lag of treated with bariatric surgery, stored as Boolean
 		l2cancre,                  ///< Lag of Cancer, [0, 1], stored as Boolean
@@ -681,6 +681,17 @@ namespace Vars
         physact,                    ///< Physiaclly Active
         l2physact,                  ///< Lag of physically active
         pphysact,                   ///< Probability of Physically active
+        logatotb,                   ///< Log of atotb (total family wealth)
+        l2logatotb,                 ///< Lag of Log of atotb (total family wealth)
+        hhres,                      ///< Number of residents in household
+        l2hhres,                    ///< Lag of Number of residents in household
+        socyr,                      ///< Whether participates in social activities
+        l2socyr,                    ///< Lag of Whether participates in social activities
+        gcareinhh1w,                ///< Whether cared for someone in household in past week
+        l2gcareinhh1w,              ///< Lag of Whether cared for someone in household in past week
+        childless,                  ///< Does not have children
+        logitot,                    ///< Log(itot)
+        l2logitot,                  ///< Lag of Log(itot)
 		NVars,                     ///< A counter for the number of valid variables
[…]

### 2023-03-13 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Did some testing and improving the social isolation model, and fixed an error in the WORKING_SCRIPT where a dplyr function had been masked

@@
 counts <- s %>% summarise_each(funs(sum(!is.na(.)))) %>%
-  select(age, male, bmi, smoken, smokev, hsless, college) %>%
+  dplyr::select(age, male, bmi, smoken, smokev, hsless, college) %>%
   t() %>%

### 2023-03-13 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Did some testing and improving the social isolation model, and fixed an error in the WORKING_SCRIPT where a dplyr function had been masked

@@
 ## TEST
-form.lnly.test <- 'lnly ~ male + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2cohab + l2widowed + l2single + l2employed + l2inactive + l2physact + l2anyadl + l2anyiadl + l2srh5 + l2psyche + l2hhres + l2socyr + l2gcareinhh1w + childless + scale(l2logatotb)'
+form.lnly.test <- 'lnly ~ male + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2cohab + l2widowed + l2single + l2employed + l2inactive + l2physact + l2anyadl + l2anyiadl + l2srh5 + l2hhres + l2socyr + l2gcareinhh1w + childless + scale(l2logatotb) + l2cesd'
 probit.lnly.test <- polr(formula = form.lnly.test,
@@
 
+## Social Isolation
+
+First recreate the minimal and current models before trying anything new.
+
+```{r}
+# Get a sample comparable to the sample used in FEM
+trans.sociso <- trans %>%
+  dplyr::select(idauniq, wave, died, sociso, male, white, hsless, college, l2age65l, l2age6574, l2age75p, atotb, l2employed, l2inactive, l2physact) %>%
+  filter(wave > 1) %>%
+  filter(died == 0)
+trans.sociso.complete <- trans.sociso[complete.cases(trans.sociso),]
+
+## MINIMAL
+form.sociso.min <- 'sociso ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.sociso.min <- polr(formula = form.sociso.min,
+                    data = trans.sociso,
+                    Hess = TRUE,
+                    na.action = na.omit)
+# print('Minimal model:')
+summary(probit.sociso.min)
+
+## FULL
+form.sociso <- 'sociso ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + scale(atotb) + l2employed + l2inactive + l2physact'
+probit.sociso <- polr(formula = form.sociso,
+                    data = trans.sociso,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.sociso)
+```
+
+Now trying the predictors from the loneliness model minus some variables that would be too strongly correlated with the outcome (specifically socyr).
+
+```{r}
+## TEST
+form.sociso.test <- 'sociso ~ male + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2cohab + l2widowed + l2single + l2physact + l2srh5 + l2hhres + l2gcareinhh1w + childless + scale(l2logatotb) + l2cesd + l2sight + l2hearing + l2ahown'
+probit.sociso.test <- polr(formula = form.sociso.test,
+                    data = trans,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Test model:')
+summary(probit.sociso.test)
+```
+
 ```{r}

### 2023-03-16 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Expanded the mortality model with chronic diseases and risk behaviours

@@
 
+I like the previous model, lets also do a few test versions.
+
+```{r}
+form.mort.test <- 'died ~ male + white + hsless + college + 
+                  l2age65l + l2age6574 + l2age75p + 
+                  l2lnly1 + l2lnly2 +
+                  l2sociso1 + l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 +
+                  l2cancre + l2diabe + l2hearte + l2alzhe + l2demene + l2lunge +
+                  l2stroke + l2smokev + l2smoken + l2smokef'
+
+probit.mort.test <- glm(formula = form.mort.test, family = binomial(link = 'probit'),
+                    data = trans)
+
+summary(probit.mort.test)
+```
+
 # Loneliness and Social Isolation

### 2023-03-17 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added cognitive function variables and updated lots of transition models. Also added lots to the R notebook and utilities

@@
 require(tidyverse)
-require(reactable)
 require(haven)
@@
 ```{r}
-source('../HAZARDS/source/hazard_utils.R')
+#source('../HAZARDS/source/hazard_utils.R')
 source('FEM_R/utilities.R')
@@
 
+  df <- data.frame(outcome = c('LifeYears', 'LifeYears'),
+                   intervention = c(int.name, int.name),
+                   scenario = c('cohort', 'intervention'),
+                   lifeyear = c(c.avg.lfyr, i.avg.lfyr))
+
   print(int.name)
@@
   print(paste('Increase:                        ', signif(x = (i.avg.lfyr - c.avg.lfyr), digits=4)))
+
+  return(df)
 }
@@
   print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+  return(df)
 }
@@
   i.avg.dis <- disabilityFreeLifeYearsCohort(i)
+  df <- data.frame(outcome = c('DisabilityFree', 'DisabilityFree'),
+                   intervention = c(int.name, int.name),
+                   scenario = c('cohort', 'intervention'),
+                   lifeyear = c(c.avg.dis, i.avg.dis))
   print(int.name)
@@
   print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
+  return(df)
 }
@@
   i.avg.dis <- diseaseFreeLifeYearsCohort(i)
+  df <- data.frame(outcome = c('DiseaseFree', 'DiseaseFree'),
+                   intervention = c(int.name, int.name),
+                   scenario = c('cohort', 'intervention'),
+                   lifeyear = c(c.avg.dis, i.avg.dis))
   print(int.name)
@@
 
+[1] "No Loneliness"
+[1] "Cohort Average Disability Free Lifeyears:         12.55"
+[1] "Intervention Average Disability Free Lifeyears:   13.67"
+[1] "Increase:                                         1.121"
+[1] "No Social Isolation"
+[1] "Cohort Average Disability Free Lifeyears:         12.55"
+[1] "Intervention Average Disability Free Lifeyears:   12.63"
+[1] "Increase:                                         0.08628"
+[1] "No Loneliness OR Social Isolation"
+[1] "Cohort Average Disability Free Lifeyears:         12.55"
+[1] "Intervention Average Disability Free Lifeyears:   13.76"
+[1] "Increase:                                         1.208"
+
 #### Disease Free Life Years
@@
 ```
+[1] "No Loneliness"
+[1] "Cohort Average Disease Free Lifeyears:         5.48"
+[1] "Intervention Average Disease Free Lifeyears:   5.668"
+[1] "Increase:                                         0.1873"
+[1] "No Social Isolation"
+[1] "Cohort Average Disease Free Lifeyears:         5.48"
+[1] "Intervention Average Disease Free Lifeyears:   5.623"
+[1] "Increase:                                         0.1427"
+[1] "No Loneliness OR Social Isolation"
+[1] "Cohort Average Disease Free Lifeyears:         5.48"
+[1] "Intervention Average Disease Free Lifeyears:   5.852"
+[1] "Increase:                                         0.3715"
+#### Visualise
+
+Lets write some functions to visualise this information, so will need to collect them into dataframes first.
+
+```{r}
+
+ly.lnly <- lifeyearComparison(cohort_full, no.lonely, 'L')
+ly.sociso <- lifeyearComparison(cohort_full, no.sociso, 'S')
+ly.lnly.sociso <- lifeyearComparison(cohort_full, no.lnly.sociso, 'LS')
+
+ly <- rbind(ly.lnly, ly.sociso)
+ly.all <- rbind(ly.lnly, ly.sociso, ly.lnly.sociso)
+ly.wide <- ly %>% 
+  pivot_wider(names_from = scenario,
+                              values_from = lifeyear,
+                              names_repair = 'unique')
+colnames(ly.wide) <- c('outcome', 'scenario', 'cohort', 'intervention')
+
+disab.lnly <- disabilityFreeLifeYearsComp(cohort_full, no.lonely, 'L')
+disab.sociso <- disabilityFreeLifeYearsComp(cohort_full, no.sociso, 'S')
+disab.lnly.sociso <- disabilityFreeLifeYearsComp(cohort_full, no.lnly.sociso, 'LS')
+
+disab <- rbind(disab.lnly, disab.sociso)
+disab.all <- rbind(disab.lnly, disab.sociso, disab.lnly.sociso)
+disab.wide <- disab %>% 
+  pivot_wider(names_from = scenario,
+                              values_from = lifeyear,
+                              names_repair = 'unique')
+colnames(disab.wide) <- c('outcome', 'scenario', 'cohort', 'intervention')
+
+disease.lnly <- diseaseFreeLifeYearsComp(cohort_full, no.lonely, 'L')
+disease.sociso <- diseaseFreeLifeYearsComp(cohort_full, no.sociso, 'S')
+disease.lnly.sociso <- diseaseFreeLifeYearsComp(cohort_full, no.lnly.sociso, 'LS')
+
+disease <- rbind(disease.lnly, disease.sociso)
+disease.all <- rbind(disease.lnly, disease.sociso, disease.lnly.sociso)
+disease.wide <- disease %>% 
+  pivot_wider(names_from = scenario,
+                              values_from = lifeyear,
+                              names_repair = 'unique')
+colnames(disease.wide) <- c('outcome', 'scenario', 'cohort', 'intervention')
+
+lifeyears <- rbind(ly.wide, disab.wide, disease.wide)
+lifeyears['change'] <- lifeyears['intervention'] - lifeyears['cohort']
+```
+
+Now plot.
+
+```{r}
+ggplot(data = filter(lifeyears, outcome == 'LifeYears'),
+       aes(x = scenario, y = change, color = scenario, fill = scenario)) +
+  geom_col()
+
+ggplot(data = filter(lifeyears, outcome == 'DisabilityFree'),
+       aes(x = scenario, y = change, color = scenario, fill = scenario)) +
+  geom_col()
+
+ggplot(data = filter(lifeyears, outcome == 'DiseaseFree'),
+       aes(x = scenario, y = change, color = scenario, fill = scenario)) +
+  geom_col()
+
+combined <- ggplot(data = lifeyears, aes(x = factor(scenario, levels = c('L', 'S', 'LS')), y = change, fill = scenario)) +
+  geom_col() +
+  facet_wrap(~factor(outcome, levels = c('LifeYears', 'DisabilityFree', 'DiseaseFree'))) +
+  labs(x = 'Scenario', y = 'Increase') +
+  scale_fill_discrete(labels = c('No Loneliness', 'No Social Isolation'))
+  #scale_color_manual(labels = c('No Loneliness', 'No Loneliness or Social Isolation', 'No Social Isolation', values = c('red', 'green', 'blue')))
+
+combined
+
+# Save plot into FEM_R/Loneliness_Social_Isolation/plots/
+path <- 'FEM_R/Loneliness_Social_Isolation/plots/'
+
+ggsave(filename = 'ly_dfly_facet_L_S.png',
+      plot = combined,
+      path = path,
+      width = 9,
+      height = 6)
+```
+
+### Sub-populations
+
+Need to get the detailed output for this, but this will be a really nice thing to include.
+We can show the improvements in similar plots by different subgroups.
+Hopefully we'll see bigger improvements in specific subgroups, perhaps a certain age group or gender.
+The literature I was reading about this suggests that each intervention should affect both genders differently, so should be interesting.
+
+Lets start by appending detailed output and working on a function to get it into a plottable format.
+
+```{r}
+cohort <- detailed_output_append(base.out.dir = 'output/LNLY_SOCISO_TEST/',
+                              scenario = 'ELSA_core_cohort')
+```
+
+```{r}
+coh <- cohort
+
+#base2 <- base %>% group_by(hhidpn, mcrep) %>%
+#                  mutate(smoken_init = ())
+
+# generate lifeyear for people who have not died
+coh$lifeyear <- 1 - coh$died
+# generate the number of repetitions for each person
+coh$nreps <- max(coh$mcrep) + 1
+# now generate the lifeyear intermediate (lifeyears per person)
+coh2 <- coh %>% 
+  group_by(hhidpn, year) %>%
+  mutate(nreps = n())# %>%
+  #ungroup() %>%
+  #group_by(hhidpn) %>%
+  #summarise(n_LY = sum(lifeyear),
+  #       mean_LY = n_LY / nreps,
+  #       n_DFLY = sum(nodisease),
+  #       mean_DFLY = n_DFLY / nreps,
+  #       two_yr_LY = mean_LY * 2,
+  #       two_yr_DFLY = mean_DFLY * 2,
+  #       weight_intermed = sum(weight),
+  #       weight2 = weight_intermed / nreps)
+
+#mean <- coh %>% summarise(LY = mean(mean_LY),
+#                          DFLY = mean(mean_DFLY))
+```
+
+#### Gender
+```{r}
+
+print('Female')
+
+```
 

### 2023-03-17 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Added cognitive function variables and updated lots of transition models. Also added lots to the R notebook and utilities

--- 2023-03-13 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
@@
 long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
-```
-
-# Mortality
-
-```{r}
-# Currently in the FEM (2/3/23) the mortality model only depends on demographics, loneliness, and social isolation
-# This was to try and get the largest effect in mortality from interventions in these variables
-# Lets try the minimal model first and go from there
-
-# Mortality is a probit model with binary outcome
 
@@
 trans$l2physact <- factor(trans$l2physact)
+
+# trans needs to be anything after wave 1
+trans <- trans %>% 
+  filter(wave > 1)
+
+trans.notDead <- trans %>% 
+  filter(died == 0)
+```
+
+# Mortality
+
+```{r}
+# Currently in the FEM (2/3/23) the mortality model only depends on demographics, loneliness, and social isolation
+# This was to try and get the largest effect in mortality from interventions in these variables
+# Lets try the minimal model first and go from there
+
+# Mortality is a probit model with binary outcome
 
@@
 probit.mort.min <- glm(formula = form.mort.min, family = binomial(link = 'probit'),
-                    data = trans)
+                    data = trans2)
 
@@
 probit.mort.ls <- glm(formula = form.mort.ls, family = binomial(link = 'probit'),
-                    data = trans)
+                    data = trans2)
 
@@
 probit.mort.ls2 <- glm(formula = form.mort.ls2, family = binomial(link = 'probit'),
-                    data = trans)
+                    data = trans2)
 
@@
 probit.mort.full <- glm(formula = form.mort.full, family = binomial(link = 'probit'),
-                    data = trans)
+                    data = trans2)
 
 summary(probit.mort.full)
+```
+
+### Test
+
+I like the previous model, lets also do a few test versions.
+
+```{r}
+form.mort.test <- 'died ~ male + white + hsless + college + 
+                  l2age65l + l2age6574 + l2age75p + 
+                  l2lnly1 + l2lnly2 +
+                  l2sociso1 + l2sociso2 + l2sociso3 + l2sociso4 + l2sociso5 +
+                  l2cancre + l2diabe + l2hearte + l2demene + l2lunge +
+                  l2stroke + l2smokev + l2smoken + l2smokef +
+                  l2alcfreq1 + l2alcfreq2 + l2alcfreq3 + l2alcfreq5 + 
+                  l2alcfreq6 + l2alcfreq7 + l2alcfreq8'
+
+probit.mort.test <- glm(formula = form.mort.test, family = binomial(link = 'probit'),
+                    data = trans2)
+
+summary(probit.mort.test)
 ```
@@
 
+# Functional Limitations
+
+Lets have a go at adlstat and iadlstat.
+
+Data:
+```{r}
+trans.adlstat <- trans %>%
+  dplyr::select(idauniq, wave, died, adlstat, iadlstat, male, white, hsless, 
+                college, l2age65l, l2age6574, l2age75p, l2stroke, l2demene, 
+                l2alzhe, l2osteoe, l2catracte, l2lnly, l2lnly1, l2lnly2, 
+                l2sociso, l2cesd, l2cohab, l2widowed, l2single, l2married,
+                l2alcfreq, l2smoken, l2smokef, l2smokev, l2tr20, l2verbf,
+                l2orient) %>%
+  filter(wave > 1) %>%
+  filter(died == 0)
+trans.adlstat.complete <- trans.adlstat[complete.cases(trans.adlstat),]
+
+trans.adlstat.complete$adlstat <- as.factor(trans.adlstat.complete$adlstat)
+trans.adlstat.complete$iadlstat <- as.factor(trans.adlstat.complete$iadlstat)
+```
+
+## ADLSTAT
+
+```{r}
+## MINIMAL
+form.adlstat.min <- 'adlstat ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.adlstat.min <- polr(formula = form.adlstat.min,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.adlstat.min)
+
+## FULL
+form.adlstat <- 'adlstat ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2stroke + l2demene + l2alzhe + l2osteoe + l2catracte + l2lnly + l2sociso'
+probit.adlstat <- polr(formula = form.adlstat,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.adlstat)
+```
+
+Now testing.
+
+```{r}
+## TEST
+form.adlstat.test <- 'adlstat ~ male + white + hsless + college + 
+scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2stroke + l2demene + l2osteoe + l2catracte + l2lnly + l2sociso + l2cesd + l2married + 
+l2widowed + l2alcfreq + l2smokev + l2tr20 + l2verbf + l2orient'
+probit.adlstat.test <- polr(formula = form.adlstat.test,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Test model:')
+summary(probit.adlstat.test)
+```
+
+AIC: 64881.53
+AIC: 64857.35 : l2tr20
+AIC: 57179.64 : l2verbf
+AIC: 57008.17 : l2orient
+AIC: 57006.91 : removed l2single
+
+Stay there for now?
+
+## IADLSTAT
+
+```{r}
+## MINIMAL
+form.iadlstat.min <- 'iadlstat ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.iadlstat.min <- polr(formula = form.iadlstat.min,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.iadlstat.min)
+
+## FULL
+form.iadlstat <- 'iadlstat ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2stroke + l2demene + l2alzhe + l2osteoe + l2catracte + l2lnly + l2sociso'
+probit.iadlstat <- polr(formula = form.iadlstat,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.iadlstat)
+```
+
+Now test.
+
+```{r}
+## TEST
+form.iadlstat.test <- 'iadlstat ~ male + white + hsless + college + 
+scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2stroke + l2demene + l2osteoe + l2catracte + l2lnly + l2sociso + l2cesd + l2married + 
+l2widowed + l2alcfreq + l2smokev + l2tr20 + l2verbf + l2orient'
+probit.iadlstat.test <- polr(formula = form.iadlstat.test,
+                    data = trans.adlstat.complete,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Test model:')
+summary(probit.iadlstat.test)
+```
+
+# Cognitive Function Vars
+
+## TR20
+
+```{r}
+## MINIMAL
+form.tr20.min <- 'tr20 ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.tr20.min <- lm(formula = form.tr20.min,
+                    data = trans.notDead,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.tr20.min)
+
+## TEST
+form.tr20 <- 'tr20 ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.tr20 <- lm(formula = form.tr20,
+                    data = trans.notDead,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.tr20)
+```
+
+```{r}
+## TEST
+form.tr20.test <- 'tr20 ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2hearing + l2sight + l2adlstat + l2iadlstat + l2cesd + l2smoken + l2employed + l2retired'
+probit.tr20.test <- lm(formula = form.tr20.test,
+                    data = trans.notDead,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.tr20.test)
+
+# l2lnly
+##  + l2physact + l2hearing + l2adlstat + l2iadlstat + l2socyr + l2cesd + l2smoken + l2smoke
+```
+
 ```{r}

### 2023-03-17 FEM_CPP/Vars.cpp

@@ -779,6 +779,12 @@ VarsInfo::VarsInfo() {
     info[Vars::l2hearing]                  = VarInfo("l2hearing", "Lag of Self-rated hearing", VarTypes::Boolean);
     info[Vars::ahown]                    = VarInfo("ahown", "Whether owns home", VarTypes::Short);
     info[Vars::l2ahown]                  = VarInfo("l2ahown", "Lag of Whether owns home", VarTypes::Short);
+    info[Vars::tr20]                    = VarInfo("tr20", "Total word recall", VarTypes::Short);
+    info[Vars::l2tr20]                    = VarInfo("l2tr20", "Lag of Total word recall", VarTypes::Short);
+    info[Vars::verbf]                    = VarInfo("verbf", "Verbal Fluency", VarTypes::Short);
+    info[Vars::l2verbf]                    = VarInfo("l2verbf", "Lag of Verbal Fluency", VarTypes::Short);
+    info[Vars::orient]                    = VarInfo("orient", "Date Naming (orientation)", VarTypes::Short);
+    info[Vars::l2orient]                    = VarInfo("l2orient", "Lag of Date Naming (orientation)", VarTypes::Short);
 
 
 
@@ -968,6 +974,9 @@ VarsInfo::VarsInfo() {
     lag_map[Vars::sight] = Vars::l2sight;
     lag_map[Vars::hearing] = Vars::l2hearing;
     lag_map[Vars::ahown] = Vars::l2ahown;
+    lag_map[Vars::tr20] = Vars::l2tr20;
+    lag_map[Vars::verbf] = Vars::l2verbf;
+    lag_map[Vars::orient] = Vars::l2orient;
 
 
 

### 2023-03-17 FEM_R/utilities.R

@@ -37,6 +37,72 @@ lookfor <- function(data,
   else message("Nothing found. Sorry.")
 }
 
+########## PREPARING SUMMARY OUTPUT FILE ##########
+
+prepare_ELSA <- function(elsa_long, 
+                         wave_start = 1, 
+                         wave_end = 6, 
+                         print_counts = TRUE) {
+  
+  # Get important variables
+  # Select waves 1-6
+  # Check if present in both first and last
+  subset <- elsa_long %>% dplyr::select(idauniq, wave, died,
+                                        age, male, white,                    # demographics
+                                        educ, hsless, college,               # demographics
+                                        smokev, smoken, smokef,              # smoking status
+                                        hibpe, bmi, stroke, hearte, diabe,   # health measures
+                                        atotb, itot,                         # wealth/income
+                                        lnly, sociso,                        # loneliness/social isolation
+                                        cwtresp, inw, insc) %>%                    # survey stuff
+    filter(wave >= wave_start, wave <= wave_end) %>%
+    group_by(idauniq) %>%
+    mutate(n = n(),
+           in_first = ((dplyr::first(inw) == 1) & (dplyr::first(wave) == 1)),
+           first_age = dplyr::first(age),
+           died_wave_2 = ((wave == 2) & (died == 1)),
+           age_group = cut(x = first_age, 
+                           breaks = c(50, 60, 70, 80, 90, 100),
+                           labels = c('50-60', '60-70', '70-80', '80-90', '90-100'), 
+                           include.lowest = TRUE)) %>%
+    filter((in_first == TRUE), 
+           (first_age >= 50),
+           ((n > 1)  | (died_wave_2 == 1))) %>% # Make sure they were in for more than 1 wave (or died in second)
+    dplyr::select(-in_first, -n)
+  
+  # check and change var types
+  subset$idauniq <- as.factor(subset$idauniq)
+  subset$wave <- as.numeric(subset$wave)
+  subset$died <- as.numeric(subset$died)
+  subset$age <- as.numeric(subset$age)
+  subset$age_group <- as.factor(subset$age_group)
+  subset$male <- as.factor(subset$male)
+  subset$white <- as.factor(subset$white)
+  subset$educ <- as.factor(subset$educ)
[…]

### 2023-03-19 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Duplicated the original repl population 10 times to have a bigger start point for cohort sims. Also added lots to the FEM_R scripts, including LY outcomes from detailed outputs, confidence intervals on plots, and visualisation of additional "realistic" interventions. Realistic interventions also added to the code, but still WIP to fine tune and get an interesting result. Added transition models for sight, hearing, and cognitive scores. Also transitioning these variables now

@@
 weighted.survey.means(s)
-
 ```
@@
 print(mean.df)
+
+rm(design, means, mean.df, counts, s)
 ```
@@
 prepped.long %>% count(lnly)
+
+rm(test.out, dead.subgroup)
 ```
@@
 #        y = 'Percentage of Sample')
+
+rm(dead, dead.alcstat)
 ```
@@
 diseaseVarLifeyear(cohort_full, 'stroke')
-diseaseVarLifeyear(cohort_full, 'alzhe')
+#diseaseVarLifeyear(cohort_full, 'alzhe')
 diseaseVarLifeyear(cohort_full, 'demene')
 ```
-
-### Disability-free Life Years
-
-### Disease-free Life Years
 
@@
 ```{r}
-coh.ly <- cohortLifeyear(cohort_full)
-nolnly.ly <- cohortLifeyear(no.lonely)
+#coh.ly <- cohortLifeyear(cohort_full)
+#nolnly.ly <- cohortLifeyear(no.lonely)
 
@@
 lifeyears['change'] <- lifeyears['intervention'] - lifeyears['cohort']
+
+rm(ly, ly.all, ly.wide, ly.lnly, ly.sociso, ly.lnly.sociso)
+rm(disab.lnly, disab.sociso, disab.lnly.sociso, disab, disab.all, disab.wide)
+rm(disease.lnly, disease.sociso, disease.lnly.sociso, disease, disease.all, disease.wide)
 ```
@@
       height = 6)
+
+rm(combined, path)
+```
+
+### Using Detailed Output
+
+Lets start by appending detailed output and working on a function to get it into a plottable format.
+
+```{r}
+cohort <- detailed_output_append(base.out.dir = 'output/LNLY_SOCISO_TEST/',
+                              scenario = 'ELSA_core_cohort')
+```
+
+```{r}
+## Think it through - Steps:
+# 1. Add a variable for lifeyears
+  # Var needs to == 1 if not died and 0 if died
+  # With this var we can calculate the total number of lifeyears for each person
+# 2. Calculate lifeyears per person per repetition
+  # Sum lifeyears per person
+  # Divide sum lifeyears by number of reps
+coh.test <- ly_outcomes_from_detailed_output(cohort)
+
+ggplot(data = coh.test, aes(x = outcome, y = mean, fill = outcome)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2)
+
+rm(coh.test)
+```
+
+#### Comparisons
+
+First combine detailed outputs for each of the interventions we want to investigate.
+
+```{r}
+out.dir <- 'output/LNLY_SOCISO_TEST/'
+
+cohort <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_core_cohort')
+nolnly <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_nolnly_cohort')
+nosociso <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_nosociso_cohort')
+both <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_nolnly_nosociso_cohort')
+```
+
+```{r}
+# Calculate LY outcomes for all scenarios
+ly.coh <- ly_outcomes_from_detailed_output(cohort)
+ly.lnly <- ly_outcomes_from_detailed_output(nolnly)
+ly.sociso <- ly_outcomes_from_detailed_output(nosociso)
+ly.lnly.sociso <- ly_outcomes_from_detailed_output(both)
+# Add scenario column for combining
+ly.coh$scenario <- 'Baseline'
+ly.lnly$scenario <- 'No_Loneliness'
+ly.sociso$scenario <- 'No_Social_Isolation'
+ly.lnly.sociso$scenario <- 'Neither'
+
+# Group into list and rbind all together
+ly.list <- list(ly.coh, ly.lnly, ly.sociso, ly.lnly.sociso)
+ly.all <- do.call(rbind, ly.list)
+
+# arrange columns properly
+ly.all <- ly.all %>% select(scenario, everything())
+
+#rm(ly.coh, ly.lnly, ly.sociso, ly.lnly.sociso, ly.list)
+```
+
+Now we have a dataframe with the LY outcomes (and confidence intervals) for all scenarios. We need to reformat the dataset and plot these values with the confidence intervals clearly shown.
+
+First do LifeYears.
+
+```{r}
+ly.all$scenario <- factor(ly.all$scenario, levels = c('Baseline', 'No_Loneliness', 'No_Social_Isolation', 'Neither'))
+ly.all$outcome <- factor(ly.all$outcome, levels = c('LY', 'DiseaseFLY', 'DisabilityFLY'))
+
+ly.ly <- ly.all %>% 
+  filter(outcome == 'LY') %>%
+  select(-outcome)
+
+ggplot(data = ly.ly, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Life Years")
+
+ly.disease <- ly.all %>%
+  filter(outcome == 'DiseaseFLY') %>%
+  select(-outcome)
+
+ggplot(data = ly.disease, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Disease Free Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Disease Free Life Years")
+
+ly.disability <- ly.all %>%
+  filter(outcome == 'DisabilityFLY') %>%
+  select(-outcome)
+
+ggplot(data = ly.disability, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Disability Free Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Disability Free Life Years")
+```
+
+#### Change
+
+```{r}
+# Now do change plots starting with the ly.all object
+ly.change <- ly.all %>%
+  pivot_wider(names_from = 'scenario',
+              values_from = c('mean', 'min', 'max', 'margin')) %>%
+  mutate(L_change = mean_No_Loneliness - mean_Baseline,
+         S_change = mean_No_Social_Isolation - mean_Baseline,
+         LS_change = mean_Neither - mean_Baseline,
+         L_min = L_change - margin_No_Loneliness,
+         L_max = L_change + margin_No_Loneliness,
+         S_min = S_change - margin_No_Social_Isolation,
+         S_max = S_change + margin_No_Social_Isolation,
+         LS_min = LS_change - margin_Neither,
+         LS_max = LS_change + margin_Neither) %>%
+  pivot_longer(cols = L_change:LS_max,
+               names_to = 'scenario',
+               values_to = 'LifeYears')
+
+ # %>%
+ #  mutate(L_min = L - margin_No_Loneliness,
+ #         L_max = L + margin_No_Loneliness,
+ #         S_min = S - margin_No_Social_Isolation,
+ #         S_max = S + margin_No_Social_Isolation,
+ #         LS_min = LS - margin_Neither,
+ #         LS_max = LS + margin_Neither)
+
+#ggplot(data = ly.change, aes(x = outcome, y = LifeYears, fill = scenario)) +
+  #geom_col(position = 'dodge', stat = 'identity') +
+  #geom_errorbar(ymin = (LifeYears - margin), ymax = (LifeYears + margin))
 ```
@@
 
-Lets start by appending detailed output and working on a function to get it into a plottable format.
+#### Gender
 
-```{r}
-cohort <- detailed_output_append(base.out.dir = 'output/LNLY_SOCISO_TEST/',
-                              scenario = 'ELSA_core_cohort')
-```
-
-```{r}
-coh <- cohort
-
-#base2 <- base %>% group_by(hhidpn, mcrep) %>%
-#                  mutate(smoken_init = ())
-
-# generate lifeyear for people who have not died
-coh$lifeyear <- 1 - coh$died
-# generate the number of repetitions for each person
-coh$nreps <- max(coh$mcrep) + 1
-# now generate the lifeyear intermediate (lifeyears per person)
-coh2 <- coh %>% 
-  group_by(hhidpn, year) %>%
-  mutate(nreps = n())# %>%
-  #ungroup() %>%
-  #group_by(hhidpn) %>%
-  #summarise(n_LY = sum(lifeyear),
-  #       mean_LY = n_LY / nreps,
-  #       n_DFLY = sum(nodisease),
-  #       mean_DFLY = n_DFLY / nreps,
-  #       two_yr_LY = mean_LY * 2,
-  #       two_yr_DFLY = mean_DFLY * 2,
-  #       weight_intermed = sum(weight),
-  #       weight2 = weight_intermed / nreps)
-
-#mean <- coh %>% summarise(LY = mean(mean_LY),
-#                          DFLY = mean(mean_DFLY))
-```
-
-#### Gender
 ```{r}
@@
 ```{r}
-
+test <- detailed_output_append(base.out.dir = 'output/COMPLETE/',
+                              scenario = 'ELSA_core_cohort')
 ```

### 2023-03-19 FEM_R/Loneliness_Social_Isolation/transition_models_testing.Rmd
Duplicated the original repl population 10 times to have a bigger start point for cohort sims. Also added lots to the FEM_R scripts, including LY outcomes from detailed outputs, confidence intervals on plots, and visualisation of additional "realistic" interventions. Realistic interventions also added to the code, but still WIP to fine tune and get an interesting result. Added transition models for sight, hearing, and cognitive scores. Also transitioning these variables now

@@
 
+# Sight
+
+NOTE: Higher values of sight (self-rated eyesight) are worse. 1 == Excellent, 5 == Poor, 6 == Blind.
+
+```{r}
+trans.notDead$sight <- as.factor(trans.notDead$sight)
+
+## MINIMAL
+form.sight.min <- 'sight ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.sight.min <- polr(formula = form.sight.min,
+                    data = trans.notDead,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.sight.min)
+
+## TEST
+form.sight <- 'sight ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2catracte + l2logbmi_l30 + l2logbmi_30p + l2hchole + l2hibpe + l2alcfreq + l2smoken + l2smokev + l2stroke'
+probit.sight <- polr(formula = form.sight,
+                    data = trans.notDead,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.sight)
+```
+
+AIC: 143800.44 : Initial
+AIC: 128596.87 : Added alcfreq
+AIC: 128133.34 : Added smoken smokev
+AIC: 128039.81 : Added stroke
+
+# Hearing
+
+Using this paper - [Kiely et al.](https://academic.oup.com/biomedgerontology/article/67/9/997/513794?login=false).
+
+```{r}
+trans.notDead$hearing <- as.factor(trans.notDead$hearing)
+
+## MINIMAL
+form.hearing.min <- 'hearing ~ male + scale(l2age65l) + scale(l2age6574) + scale(l2age75p)'
+probit.hearing.min <- polr(formula = form.hearing.min,
+                    data = trans.notDead,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Minimal model:')
+summary(probit.hearing.min)
+
+## TEST
+form.hearing <- 'hearing ~ male + white + hsless + college + scale(l2age65l) + scale(l2age6574) + scale(l2age75p) + l2logbmi_l30 + l2logbmi_30p + l2smoken + l2smokev + l2hibpe + l2diabe + l2stroke + l2sight'
+probit.hearing <- polr(formula = form.hearing,
+                    data = trans.notDead,
+                    Hess = TRUE,
+                    na.action = na.omit)
+print('Full model:')
+summary(probit.hearing)
+```
+
+AIC: 152711.79 : initial
+
 ```{r}

### 2023-03-19 FEM_R/utilities.R

@@ -199,7 +199,7 @@ ly_outcomes_from_detailed_output <- function(coh) {
     pivot_longer(cols = everything(),
                  names_sep = '_',
                  names_to = c('outcome', 'statistic')) %>%
-    filter(statistic %in% c('mean', 'min', 'max')) %>%
+    filter(statistic %in% c('mean', 'min', 'max', 'margin')) %>%
     pivot_wider(values_from = 'value',
                 names_from = 'statistic')
   

### 2023-03-19 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added lots of analysis for the detailed output intervention stuff, as well as generated new populations (and fixed some old ones) in reweight_ELSA_repl.do

@@
 ```{r}
+# save these large files in case I overwrite the output stuff
+write.csv(cohort, file = 'input_data/detailed_output/cohort_append_large.csv')
+write.csv(nolnly, file = 'input_data/detailed_output/nolnly_append_large.csv')
+write.csv(nosociso, file = 'input_data/detailed_output/nosociso_append_large.csv')
+write.csv(both, file = 'input_data/detailed_output/both_append_large.csv')
+```
+
+```{r}
 # Calculate LY outcomes for all scenarios
@@
          LS_max = LS_change + margin_Neither) %>%
+  select(outcome, contains('L_'), contains('S_'), contains('LS_'), -contains('No_Social_Isolation')) %>%
   pivot_longer(cols = L_change:LS_max,
-               names_to = 'scenario',
-               values_to = 'LifeYears')
+               names_sep = '_',
+               names_to = c('scenario', 'statistic'),
+               values_to = 'LifeYears') %>%
+  pivot_wider(names_from = 'statistic',
+              values_from = 'LifeYears')
 
- # %>%
- #  mutate(L_min = L - margin_No_Loneliness,
- #         L_max = L + margin_No_Loneliness,
- #         S_min = S - margin_No_Social_Isolation,
- #         S_max = S + margin_No_Social_Isolation,
- #         LS_min = LS - margin_Neither,
- #         LS_max = LS + margin_Neither)
+ly.change$scenario <- factor(ly.change$scenario, levels = c('L', 'S', 'LS'))
 
-#ggplot(data = ly.change, aes(x = outcome, y = LifeYears, fill = scenario)) +
-  #geom_col(position = 'dodge', stat = 'identity') +
-  #geom_errorbar(ymin = (LifeYears - margin), ymax = (LifeYears + margin))
+combined.plot <- ggplot(data = ly.change, aes(x = scenario, y = change, fill=scenario)) +
+  geom_col() +
+  facet_wrap(~outcome) +
+  geom_errorbar(aes(ymin = min, ymax = max)) +
+  labs(title = 'Life Years, Disease Free Life Years, and Disability Free Life Years') +
+  xlab('Scenario') +
+  ylab('Change (years)')
+
+combined.plot
+
+path <- 'FEM_R/Loneliness_Social_Isolation/plots/'
+
+ggsave(filename = 'ly_dfly_facet_all_confints.png',
+      plot = combined.plot,
+      path = path,
+      width = 9,
+      height = 6)
+```
+
+### Realistic Intervention
+
+```{r}
+lnly.real <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_real_lnly')
+lnly.sociso <- detailed_output_append(base.out.dir = out.dir,
+                              scenario = 'ELSA_real_sociso')
+```
+
+```{r}
+ly.lnly.real <- ly_outcomes_from_detailed_output(lnly.real)
+ly.sociso.real <- ly_outcomes_from_detailed_output(lnly.sociso)
+
+ly.lnly.real$scenario <- 'Loneliness'
+ly.sociso.real$scenario <- 'Social_Isolation'
+
+ly.comb.list <- list(ly.coh, ly.lnly.real, ly.sociso.real)
+ly.combined <- do.call(rbind, ly.comb.list)
+
+# arrange columns properly
+ly.combined <- ly.combined %>% select(scenario, everything())
+
+ly.combined$scenario <- factor(ly.combined$scenario, levels = c('Baseline', 'Loneliness', 'Social_Isolation'))
+ly.combined$outcome <- factor(ly.combined$outcome, levels = c('LY', 'DiseaseFLY', 'DisabilityFLY'))
+```
+
+#### LY & DFLYs
+```{r}
+ly.ly <- ly.combined %>% 
+  filter(outcome == 'LY') %>%
+  select(-outcome)
+
+ggplot(data = ly.ly, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Life Years")
+
+ly.disease <- ly.combined %>%
+  filter(outcome == 'DiseaseFLY') %>%
+  select(-outcome)
+
+ggplot(data = ly.disease, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Disease Free Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Disease Free Life Years")
+
+ly.disability <- ly.combined %>%
+  filter(outcome == 'DisabilityFLY') %>%
+  select(-outcome)
+
+ggplot(data = ly.disability, aes(x = scenario, y = mean, fill = scenario)) +
+  geom_bar(position = 'dodge', stat = 'identity') +
+  geom_errorbar(aes(ymin = min, ymax = max),
+                width = 0.2) +
+  labs(title = 'Disability Free Life Year Comparison between Intervention Scenarios') +
+  xlab("Scenario") +
+  ylab("Mean Disability Free Life Years")
+```
+
+Now change plots.
+
+```{r}
+# Now do change plots starting with the ly.all object
+ly.change.real <- ly.combined %>%
+  pivot_wider(names_from = 'scenario',
+              values_from = c('mean', 'min', 'max', 'margin')) %>%
+  mutate(L_change = mean_Loneliness - mean_Baseline,
+         S_change = mean_Social_Isolation - mean_Baseline,
+         L_min = L_change - margin_Loneliness,
+         L_max = L_change + margin_Loneliness,
+         S_min = S_change - margin_Social_Isolation,
+         S_max = S_change + margin_Social_Isolation) %>%
+  select(outcome, contains('L_'), contains('S_')) %>%
+  pivot_longer(cols = L_change:S_max,
+               names_sep = '_',
+               names_to = c('scenario', 'statistic'),
+               values_to = 'LifeYears') %>%
+  pivot_wider(names_from = 'statistic',
+              values_from = 'LifeYears')
+
+ly.change.real$scenario <- factor(ly.change.real$scenario, levels = c('L', 'S', 'LS'))
+
+combined.plot <- ggplot(data = ly.change.real, aes(x = scenario, y = change, fill=scenario)) +
+  geom_col() +
+  facet_wrap(~outcome) +
+  geom_errorbar(aes(ymin = min, ymax = max)) +
+  labs(title = 'Life Years, Disease Free Life Years, and Disability Free Life Years') +
+  xlab('Scenario') +
+  ylab('Change (years)')
+
+combined.plot
 ```
@@
 The literature I was reading about this suggests that each intervention should affect both genders differently, so should be interesting.
+
+As a first step, lets get the breakdown by gender and age group of loneliness and social isolation (for planning interventions).
+
+```{r}
+lnly.check <- repl %>%
+  select(hhidpn, year, age, male, lnly) %>%
+  filter(year == 2012) %>%
+  group_by(lnly, male) %>%
+  count()
+
+lnly.check2 <- repl %>%
+  filter(year == 2012) %>%
+  select(male, lnly) %>%
+  group_by(lnly) %>%
+  summarise(n = n(),
+            male = sum(male) / n,
+            female = 1 - male)
+
+lnly.check3 <- long %>%
+  select(male, lnly) %>%
+  group_by(lnly) %>%
+  summarise(n = n(),
+            male = sum(male) / n,
+            female = 1 - male)
+
+lnly.check4 <- stock %>%
+  select(male, lnly) %>%
+  group_by(lnly) %>%
+  summarise(n = n(),
+            male = sum(male) / n,
+            female = 1 - male)
+
+lnly.check5 <- repl %>%
+  filter(year == 2012) %>%
+  select(hhidpn, year, age, male, childless, lnly) %>%
+  group_by(lnly, male, childless) %>%
+  count()
+```
 
